### 2016-05-11 nips_2016.tex
first commit, adding nips template plus ipython notebook stuff

\documentclass{article}

% if you need to pass options to natbib, use, e.g.: \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
% to avoid loading the natbib package, add option nonatbib: \usepackage[nonatbib]{nips_2016}

\PassOptionsToPackage{numbers, compress}{natbib}
\usepackage{nips_2016}

% to compile a camera-ready version, add the [final] option, e.g.:
%\usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings}
\graphicspath{{figures/}}

\lstset{
numbers=left, 
numberstyle=\small, 
numbersep=8pt, 
frame = single, 
language=matlab, 
framexleftmargin=20pt}

\newtheorem{lemma}{Lemma}

\title{Fast Mini-Batch MCMC for Big Data Bayesian Posterior Inference}

% The \author macro works with any number of authors. There are two commands used to separate the
% names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the lines. Using \AND
% forces a line break at that point. So, if LaTeX puts 3 of 4 authors names on the first line, and
% the last on the second line, try using \AND instead of \And before the third author name.

\author{
  David S.~Hippocampus\thanks{Use footnote for providing further
    information about author (webpage, alternative
    address)---\emph{not} for acknowledging funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
Markov chain Monte Carlo (MCMC) methods have become one of the most popular techniques in machine
learning, but they can be intractably slow when doing Bayesian posterior estimation due to the need
to use information from all training data points during each sampling iteration. We propose a
mini-batch approach that uses information from a small subset of samples per iteration. This, along
with a new Metropolis-Hastings test, will allow us to sample faster to obtain higher-quality
parameter estimates. We implement our sampler in BIDMach and benchmark it with the state of the art.
{\color{blue} Daniel: I hope we can do this.}
{\color{blue} Xinlei: Compare with "cutting MH test budget" paper?}
\end{abstract}

\section{Introduction}

MCMC has become one of the most popular techniques in machine learning...

In this paper, we resolve these challenges\footnote{{\color{blue} Daniel: we can write these
``challenges'' later once the project becomes clearer}.} by proposing an MCMC algorithm for
Bayesian parameter estimation in which proposals are fast and mini-batch based. Our M-H test
acceptance probability is about 50 percent, reasonably high, which means our samples mix well.We
experimentally show that our approach is better than the state of the art.

{\color{blue} Daniel: we can fill out this section later.}

\section{Related Work}

In this section, we review different results on improving MCMC methods with a focus on those dealing
with mini-batches of large datasets.

Cutting the Metropolis-Hastings Budget paper \cite{ cutting_mh_2014} proposed that there is a bias-variance trade-off in standard MCMC methods. Bias occurs when sampling from the wrong distribution, however, bias usually decreases fast. Variance occurs when the number of sampling is not enough since we can collect only a limited number of samples in a given amount of time. The recently proposed Stochastic Gradient Langevin Dynamics (SGLD) \cite{langevin_2011} and Stochastic Gradient Fisher Scoring methods are biased because they omit the required Metropolis-Hastings tests.

Hamiltonian Monte Carlo (HMC) methods attempt to simulate the ``physics'' of the
system\footnote{Intuitively, we can think of a probability distribution as being a distribution over
particles in the system, where particles are the values of random variables.} in order to generate
distant, high-quality sample proposals (those are the best kind of samples). Since they simulate the
physical dynamics of the system, total energy is preserved and their proposals are always
accepted and there is no need for a Metropolis-Hastings correction/test\footnote{I'm not sure why
this follows but I haven't read the details.}. Unfortunately, HMC methods require the use of a
gradient computation based on all of the data per iteration. (This means the advantage of HMC
methods is that they generate high-quality proposals that mix well; their downside is that they are
unsuitable for big data, which might involve Bayesian posterior inference over that data.)

The Stochastic Gradient Hamiltonian Monte Carlo (SGHMC) method~
\cite{sghmc_2014} tries to compromise
by using a \emph{mini-batch version} of HMC, so each iteration, only the mini-batch is used to
compute the gradient.  SGHMC also includes an extra friction term to counteract the effect of noise
on the system (or gradient) that is added due to using only a small subset of data. (The ``naive''
way of replacing the batch with the mini-batch \emph{without} friction means the system deviates
from the target/desired distribution.) ~\cite{sghmc_2014} assumes the injected noise is Gaussian (we
make a similar assumption) and appeals to the Central Limit Theorem and second-order Langevin
dynamics\footnote{There is a class of MCMC method techniques called \emph{Langevin dynamics}~\cite{mcmc_hamiltonian_2010}. Those algorithms involve computing a gradient update for the parameter of interest $\theta$ and then using the updated $\theta$ as the next sample (i.e., this is like the proposal test). Moreover, by simulating the Hamiltonian dynamics of the system, there is no need for a rejection test. To converge to a \emph{full (posterior) distribution} of $\theta$ and not a single point (i.e., the MAP estimate of $\theta$, or the mode of the distribution) langevin dynamics adds Gaussian noise each iteration to cause the $\theta$ samples to ``jump around.'' The work of~\cite{langevin_2011} is closer to ours because it applies the normal Langevin dynamics algorithm with one difference: the gradient updates are based on a \emph{mini-batch} of the original data.} to show that SGHMC maintains the desired target distribution as the stationary
distribution. Our methods are similar in that we use mini-batches of data, but the difference is
that they focus on improving a particular flavor of MCMC, \emph{Hamiltonian} Monte Carlo (which uses Langevin dynamics to add noise), and they
also try avoiding MH tests, while we are focused on improving \emph{general} MCMC by means of a \emph{more efficient} MH proposal. These are different ways of solving the same problem (i.e., the problem of using mini-batches of
large data for MCMC).

The work of~\cite{cutting_mh_2014} also concerns speeding up MCMC 
by using mini-batches of data. Their argument is that the standard MCMC algorithm for Bayesian
posterior inference (i.e., to compute $p(\theta \mid x_1, \ldots, x_N)$ for i.i.d. data points) is
unsuitable for large datasets because the likelihood computation requires the use of all points,
$p(x_1, \ldots, x_n \mid \theta) = \prod_{i=1}^Np(x_i\mid \theta)$. Using all of this data just to
decide whether the new sample $\theta$ should be accepted or rejected is computationally
inefficient. In theory, given infinite time, using all the data points is fine because it results in
an unbiased estimate of the target distribution and the variance of the resulting sampled values is
low given enough time.\footnote{Note that~\cite{cutting_mh_2014} primarily use MCMC to estimate just
the expectation of the target distribution, but their analysis should extend to other values of the
target distribution for arbitrary input $\theta$. In other words, they use MCMC to estimate
$\mathbb{E}[p(\theta \mid x_1, \ldots, x_N)]$, but we are mostly interested in $p(\theta \mid x_1,
\ldots, x_N)$.} In reality, we have limited time. The key argument in~\cite{cutting_mh_2014} is that
given a fixed budget, it may be better to use $n$ data points per iteration (where $n \ll N$), which
will sacrifice some bias but will perhaps improve variance by allowing us to sample for $\theta$
more often, since we will get $\theta_1, \ldots, \theta_T$ as opposed to the standard case using all
$N$ samples but with only $\theta_1, \ldots, \theta_t$ samples ($t \ll T$).

More formally, the algorithm they develop is a sequential hypothesis testing algorithm. During each
iteration, their algorithm generates a new sample $\theta'$ from the proposal distribution. They
start with a small mini-batch of the data, and test the hypothesis (based on this small mini-batch)
that $\theta'$ should be accepted or rejected based the likelihood ratio. (Intuitively, we can
easily tell if $\theta'$ will be accepted/rejected if the mini-batch of data results in likelihood
drastically different from the likelihood of the current sample $\theta$.) If the hypothesis test is
too uncertain, then we increase the mini-batch size and apply the test again, and we repeat this
increment until we accept or reject $\theta'$ with high confidence (based on a Student-t
distribution). The downside of this algorithm is the need to keep incrementing the mini-batch size in one iteration; in the worst case, they may not be able to tell with confidence to accept or
reject until we get nearly all $N$ of the data points, which defeats the purpose of the algorithm.
We propose our own algorithm that avoids this problem. {\color{blue} Daniel: does this make sense as
a ``weakness'' of their approach?  Ideally I would like to have a comparison between our method and
theirs.}

\section{Our Algorithm}\label{sec:our_algorithm}

\subsection{The General MCMC Method}\label{sec:general_mcmc}

The general MCMC method proceeds as follows. For a desired random vector $\theta$, we have a
\emph{target distribution} we wish to compute, $p(\theta \mid x_1, \ldots, x_N)$, based on $N$
i.i.d. data points. Since exact evaluation is intractable, we generate a chain of correlated samples
$\theta_1, \ldots, \theta_T$ for some large $T$, and approximate $p$ by using counts of samples. For
each iteration $t$, we start with our current $\theta_t$. We use a \emph{proposal distribution}
$q(\theta' \mid \theta_t)$ to determine a new candidate $\theta'$. With probability $P_a$, we accept
it and set $\theta_{t+1} = \theta'$; otherwise, we repeat the previous value $\theta_{t+1} =
\theta_t$. Traditionally, $P_a$ is computed as follows:
\begin{equation}\label{eq:traditional}
P_a = \min\left\{ 1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid \theta_t)}
\right\} = \min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i ; \theta')q(\theta_t \mid \theta')}{p(\theta_t)\prod_{i=1}^N p(x_i ; \theta_t)q(\theta' \mid
\theta_t)} \right\},
\end{equation}
where $f(\theta_t)=p(\theta_t|x_1,\ldots,x_N) \propto p(\theta_t)\prod_{i=1}^N p(x_i
\mid \theta_t)$ which roughly tells us how ``likely'' we are to have a $\theta_t$ sample. The
normalizing constants cancel out, so the requirement for $P_a$ is that $f$ be proportional to the
true target distribution.

One then draws a uniform random variable $u$ from ${\rm Unif}[0,1]$ and accepts $\theta_{t+1} =
\theta'$ if $u < P_a$, and uses $\theta_{t+1} = \theta_t$ otherwise. The $P_a$ from
Equation~\ref{eq:traditional} satisfies ``detailed balance'' which, roughly speaking, means that if
we sample long enough using the heuristic above, we will arrive at a stationary distribution.

The $P_a$ is guaranteed to converge to the true $p$ distribution given sufficiently many examples
(though a burn-in period and/or taking every $n$th sample may be necessary). Unfortunately,
computing $f$ requires the use of all $N$ training data points. In this paper, we derive a new fast,
approximate Metropolis-Hastings test which does not use all $N$ points during each test.

\subsection{A Different Metropolis-Hastings Test}\label{sec:faster}

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{john_bair_fig01}
  \caption{
  To the left is the logistic distribution, representing a possible acceptance test. As $\Delta U$
  approaches infinity, the acceptance rate approaches 1. To the right, we have a new variable $X$
  and its density, along with an acceptance test $\Delta U + X$, which we describe in detail in
  Section~\ref{sec:our_algorithm} (note that we use $\Delta$ instead of $\Delta U$ for simplicity).
  {\color{blue}
  Daniel: I think that the diagram to the left should use the logistic function, i.e., replace
  ``$\exp(\Delta U)$'' with ``$(1+\exp(\Delta U))^{-1}$.''
  }
  }
  \label{fig:part1}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{john_bair_fig02}
  \caption{
  To the left, we have $\Delta U$ (which is simplified as $\Delta$ in
  Section~\ref{sec:our_algorithm}), along with two noise terms that come from the new variable we
  define $X$. The details are in Section~\ref{sec:our_algorithm}.
  }
  \label{fig:part2}
\end{figure}

Our starting point is the following Lemma.

\begin{lemma}\label{lem:detailed_balance}
Let $\Delta = \log \left(\frac{f(\theta') q(\theta_t \mid
\theta')}{f(\theta_t) q(\theta'\mid \theta_t)} \right)$, where $f$ is proportional to the desired target distribution and $q$ is our chosen
proposal distribution. Any acceptance function $g$ such that $g(\Delta) = \exp(\Delta) g(-\Delta )$
satisfies detailed balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) = f(\theta')p(\theta_t
\mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the probability of jumping from $\theta_x$ to $\theta_y$ in our chain.
\end{lemma}

\begin{proof}
We begin by deriving $p(\theta' \mid \theta_t)$. This is equivalent to the probability of proposing
$\theta'$ and then accepting it, so
\begin{equation}\label{eq:one_way}
p(\theta' \mid \theta_t) = q(\theta' \mid \theta_t)g(\Delta).
\end{equation}
Similarly, we have
\begin{equation}\label{eq:other_way}
p(\theta_t \mid \theta') = q(\theta_t \mid \theta')g(-\Delta).
\end{equation}
Notice that the probability of accepting a transition from $\theta'$ to $\theta_t$ is $g(-\Delta)$
because this inverts the fraction inside the logarithm term of $\Delta$.  By assumption, we can
expand $g(\Delta) = \exp(\Delta)g(-\Delta)$ in Equation~\ref{eq:one_way}.  Doing this, and combining
the result of Equation~\ref{eq:other_way}, we get
\begin{equation}\label{eq:combined}
g(-\Delta) = \frac{p(\theta' \mid \theta_t)}{q(\theta' \mid \theta_t)\exp(\Delta)} = \frac{p(\theta_t \mid \theta')}{q(\theta_t \mid \theta')}.
\end{equation}
Rearranging terms and expanding $\exp(\Delta)$, we have
\begin{equation}\label{eq:rearrange}
\frac{p(\theta' \mid \theta_t) f(\theta_t) q(\theta' \mid \theta_t)}{q(\theta' \mid \theta_t) f(\theta') q(\theta_t \mid \theta')} = \frac{p(\theta_t \mid \theta')}{ q(\theta_t \mid \theta')}.
\end{equation}
Cancellations result in $f(\theta') p(\theta_t \mid \theta') = f(\theta_t) p(\theta' \mid \theta_t)$. Thus, detailed balance is satisfied.\\
\end{proof}

Some straightforward arithmetic shows that the standard Metropolis-Hastings acceptance function
$g(\Delta) = \min\{1, e^\Delta \} = \min\left\{1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid
\theta_t)}\right\}$ satisfies the condition $g(\Delta) =
\exp(\Delta)g(-\Delta)$ and consequently, results in detailed balance.

For reasons that will be clear later, we choose an alternative $g$, the logistic function:
$g(\Delta) = (1+\exp(-\Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
condition in Lemma~\ref{lem:detailed_balance}.

We can sample from the logistic function using the following procedure. Let $u$ be drawn from ${\rm
Unif}[0,1]$. At any given iteration, we can compute $\Delta$, and we accept the new candidate
$\theta'$ if $g(\Delta) > u$, and reject otherwise. Let us define the random variable $X =
g^{-1}(u)$ where again, $g$ is the logistic function. We claim that $X$ has CDF function $F_X(x) =
g(x)$, i.e., its CDF is precisely the logistic function. This is because for some $X = x$, we have
\[
F_X(x) = {\rm Pr}(X \le x) = {\rm Pr}(g^{-1}(u) \le x) = {\rm Pr}(u \le g(x)) = \int_{0}^{g(x)} 1 dx = g(x),
\]
as the density of the uniform random variable here is simply one. Thus, the criteria to accept the
candidate $\theta'$ is equivalent to whether $\Delta > X$, where $X$ has CDF of logistic function.
(If this isn't immediately obvious, note that we can pre-multiply $g^{-1}$ to $g(\Delta) \ge u$, and
our test is $\Delta > X$.) Since $X$ is symmetric about zero, the acceptance criteria can also be
expressed as $\Delta + X>0$, as shown in Figure~\ref{fig:part1}.

{\color{blue}
Daniel: The above makes sense, but I still don't understand why we had to show that the CDF of $X$
is the logistic function, because I never use that when I claim that our acceptance test is now
$\Delta > X$.  Do we need that assumption so that, for instance, we can decompose $X$ later the way
we do it, with $X_{\rm noise}$ having the same variance as $\epsilon$? That might explain it.
}

Unfortunately, there's a problem with the $\Delta + X > 0$ test: it requires us to compute $\Delta$,
so our test will be slow! We can get around this by defining a new scalar-valued term, $\Delta'$, which is
computed the same way as $\Delta$, but uses far fewer samples. In other words, $\Delta'$ is computed
based on a \emph{mini-batch} of data, and it is an approximation of $\Delta$, so we can express the
relationship as $\Delta' = \Delta + \epsilon$ for a noise term $\epsilon$. \emph{The key is that
$\epsilon$ follows a normal distribution}, $\epsilon \sim \mathcal{N}(0, \sigma^2)$, because for i.i.d. data $x_1, \ldots, x_N$, the $\Delta$
is expressed as the sum of log probabilities (look at Equation~\ref{eq:traditional} and apply the
logarithm to the product over the $N$ data points, to get a sum of log terms). All we do with
$\Delta$ is take far fewer of the summations, and normally-distributed noise is there as a product
of the Central Limit Theorem.

To sample accurately from $\Delta'$, we need to slightly change our acceptance criteria. We can
decompose $X$ as $X = X_{\rm norm} +X_{\rm correction}$, where the first term is a ``noise term''
which has the same variance as $\epsilon$, and the second term is a ``residual'' term. The criteria
to accept, previously expressed as $\Delta + X > 0$, can be rewritten as
\begin{equation}\label{eq:criteria}
\Delta + X = \Delta + X_{\rm norm} + X_{\rm correction} \approx \Delta' + X_{\rm correction} >0.
\end{equation}
This explains Figure~\ref{fig:part2} (though it really should be an approximation as we can't guarantee that our $X_{\rm norm}$ term is exactly the correct amount of noise), where the left hand side has distributions representing normal
noise ($X_{\rm norm}$, which is labeled as ${\rm Noise}_U$, but they are the same thing) and a
``correction'' term labeled $X_{\rm correction}$. Be aware that all of these ``additions'' are to be
interpreted as convolutions when computing the sum of densities.

Notice that we have several options we can control. For instance, we can adjust the mini-batch size;
increasing the mini-batch size will cause the $X_{\rm noise}$ distribution to shrink and become
narrower.

{\color{blue} Daniel: I think it's becoming clear but I'm still trying to connect all of the steps.
During each iteration of MCMC, we compute $\Delta'$ and \emph{sample} (right?) an $X$ (which we can
do by sampling $u$ and then doing $X = g^{-1}(u)$. This gives us $\Delta' + X$ but our MH test
    actually requires $X_{\rm correction}$, so first we must estimate $\sigma^2$, the variance of
    $X_{\rm noise} \sim \mathcal{N}(0, \sigma^2)$ (perhaps from the mini-batch data, somehow?) and
    then somehow ``remove'' that component from $X$ to get $X_{\rm correction}$? Then we have our
    acceptance test. Is this right? Or, of course, we don't have to sample $X$ at all, but just
    figure out a way to sample $X_{\rm correction}$ directly.

I think we talked about estimating $\sigma^2$. The variance $\sigma^2$ that we're concerned about is
a scalar value based on the sum of log probabilities of different samples, so one way we could
estimate the variance is by taking all samples of a mini-batch and seeing how much those individual
$\log p(x_i \mid \theta)$ vary. Or we could look at $k$ mini-batches of data which will give us
$\Delta_1, \ldots, \Delta_k$ and compute the empirical $\sigma^2$ (of course that defeats the
purpose of using mini-batches).  }

{\color{blue} Xinlei: I agree with you that out MH test requires $X_{correction}$. My understanding
is that we can actually estimate $X$ from the whole data set, and the probability distribution of
$X_{norm}$ is what we proposed? So the determination of $X_{correction}$ is a deterministic
deconvolution problem.  }

\section{Implementation}\label{sec:implementation}

We implement our system within the open-source BIDMach project~\cite{canny2013bidmach}. Note to
ourselves: do NOT modify the \texttt{Learner.scala} code -- just make it a new ``Model''.

\section{Experiments}\label{sec:experiments}

\subsection{Experiment 1}

\subsection{Experiment 2}

\section{Conclusion}\label{sec:conclusion}

\small
\bibliography{nips_2016}
\bibliographystyle{ieeetr}

\clearpage
\appendix

\section{Section for Appendix Information}

{\color{blue}
Daniel: In case we want to use this.
}

% Daniel: here are example LaTeX codes for figures and tables if we want to use them.

%\begin{figure}[h]
%  \centering
%  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
%  \caption{Sample figure caption.}
%\end{figure}

%\begin{table}[t]
%  \caption{Sample table title}
%  \label{sample-table}
%  \centering
%  \begin{tabular}{lll}
%    \toprule
%    \multicolumn{2}{c}{Part}                   \\
%    \cmidrule{1-2}
%    Name     & Description     & Size ($\mu$m) \\
%    \midrule
%    Dendrite & Input terminal  & $\sim$100     \\
%    Axon     & Output terminal & $\sim$10      \\
%    Soma     & Cell body       & up to $10^6$  \\
%    \bottomrule
%  \end{tabular}
%\end{table}

%\usepackage[pdftex]{graphicx} ...
%\includegraphics[width=0.8\linewidth]{myfile.pdf}

\end{document}

### 2016-05-11 MATLAB_stuff/deconvolve.m

@@ -1,0 +1,56 @@
+% Some parameters ...
+n = 1000;
+dist = 10;
+ascale = 2;
+
+% Creates x = [-10.00, -9.99, -9.98, ... , 9.98, 9.99, 10.00] with n=1000, dist=10.
+x = ((-n):n)/n*dist;
+
+% Creates ldist = [0, ..., 0.25 , ..., 0] w/a distribution that seems Gaussian.
+ldist = 1./(2+exp(ascale*x)+exp(-ascale*x));
+
+% Almost the same as ldist but the middle values seem to have a higher peak (0.6-ish).
+ndist = normpdf(x, 0, 1.4/ascale);
+
+% The same as ldist except with about 500 zeros padded on each side.
+ldistfull = [zeros(1, n/2) , ldist, zeros(1, n/2-1)];
+
+% For each column, we set a different 2000 block slice (out of 3000) to be ldist.
+convmat = zeros(3*n, n);
+for i = 1:n
+  convmat(i:(i+2*n), i) = ndist';
+end
+
+% Making everything nonzero ...
+tail = ldistfull' + 3e-3;
+
+% A (6000 x 1001) matrix but IDK why we need this.
+A = [[convmat ; -convmat] , -[tail ; tail]];
+
+% A (6000 x 1) vector, w/normal-like distributions on top/bottom (bottom is negated).
+b = [ldistfull'; -ldistfull'];
+
+% OK ... why on earth are we using a linear programming solver? I'm missing the context.
+f = [zeros(n,1); 1];    % A (1001 x 1) vector, IDK why.
+lb = zeros(n+1,1);      % A (1001 x 1) vector, perhaps A*(lb) = b?
+
+% Returns y that minimizes f^Ty s.t. Ay <= b; 'lb' is a component-wise lower_bound on y.
+y=linprog(f,A,b,[],[],lb);
+
+% This is interesting, we convolve it ... I guess this is a nice way of doing a convolution?
+z = convmat * y(1:n);
+zs = z((n/2):(5*n/2))';
+
+% The "semilogy" plots data with log scale for y-axis, but IDK what this means.
+figure(1)
+semilogy(x, ldist, x, zs);
+
+% Easiest figure to understand. Looks like y is sharp in a few places, zero in most of them.
+figure(2)
+plot(y);
+
+% relerr reported as 0.0101
+relerr = y(n+1)
+
+% Third figure does ... something, IDK what.
+figure(3);plot(x,zs./ldist,x,10*ldist)

### 2016-05-11 MATLAB_stuff/deconvolve2.m

@@ -1,0 +1,44 @@
+% OK a bunch of the usual parameters ...
+n = 1000;
+dist = 10;
+ascale = 2;
+iclip = 20;
+
+% Again, produce some x-coordinates, but it's odd, we go 0, ..., n, but then -n, ..., 0. ???
+x = [(0:n), ((-n):-1)]/n*dist;
+
+% So ... [0.5, ..., 0, ..., 0.5], goes down to zero, stays at zero, then goes back up. ???
+ldist = ascale./(2+exp(ascale*x)+exp(-ascale*x));
+
+% As expected, is [0.8, ..., 0, ..., 0.8] due to how values closer to zero are at the ends of "x".
+ndist = normpdf(x, 0, 1/ascale);
+
+% r0 = 1001 random nubers, in a row-vector, then rr has a symmetric version of r0 in second half.
+% But we never use these?
+r0 = rand(1,n+1);
+rr = [r0, r0(end:-1:2)];
+
+% Plots ldist and ndist, very straightforward. Note: greener/lighter curve is ndist, the second item.
+figure(1);plot(1:length(x), ldist, 1:length(x), ndist, 'LineWidth', 3);legend('ldist','ndist');
+
+% Now we're doing a Fast Fourier Transform of ldist and ndist, why?
+ff = real(fft(ldist));
+nff = real(fft(ndist));
+
+% At least we can see the (absolute value) of the FFT, but the plot isn't that informative. The LOG
+% plot is MUCH better and we see some weird stuff at the bottom.
+figure(2);plot(1:length(x),abs(ff),1:length(x),abs(nff), 'LineWidth', 2); legend('abs(ff)','abs(nff)');
+figure(3);semilogy(1:length(x),abs(ff),1:length(x),abs(nff), 'LineWidth', 2); legend('abs(ff)','abs(nff)');
+
+% Then we form u (component-wise division, then zero out elements, then do _inverse_ fft ...). ???
+fdiv = ff./nff;
+fdiv(iclip:(length(fdiv)-iclip+2)) = 0;
+u = ifft(fdiv);
+figure(4);plot(u, 'Linewidth', 2); legend('u');
+
+% Then we do something to form v and plot it. We plot it with ldist, which is interesting. ???
+v = ifft(fdiv.*nff);
+figure(5);semilogy(1:length(x), ldist, 1:length(x), v, 'Linewidth', 2);
+legend('v');
+
+figure(6); plot(x, u, 'Linewidth', 2); legend('plot(x,u)');

### 2016-05-11 MATLAB_stuff/deconvolve3.m

@@ -1,0 +1,23 @@
+% OK a bunch of the usual parameters ...
+n = 1000;
+dist = 10;
+ascale = 2;
+iclip = 20;
+x = [(0:n), ((-n):-1)]/n*dist;
+%x = [(-n:n)]/n * dist; % So now it goes from, e.g., -10 to 10 by 0.01.
+ldist = ascale./(2+exp(ascale*x)+exp(-ascale*x));
+ndist = normpdf(x, 0, 1/ascale);
+figure(1);plot(1:length(x), ldist, 1:length(x), ndist, 'LineWidth', 3);legend('ldist','ndist');xlim([0,2000]);
+
+ff = real(fft(ldist));
+nff = real(fft(ndist));
+figure(2);plot(1:length(x),abs(ff),1:length(x),abs(nff), 'LineWidth', 2); legend('abs(ff)','abs(nff)');xlim([0,2000]);
+figure(3);semilogy(1:length(x),abs(ff),1:length(x),abs(nff), 'LineWidth', 2); legend('abs(ff)','abs(nff)');xlim([0,2000]);
+
+fdiv = ff./nff;
+fdiv(iclip:(length(fdiv)-iclip+2)) = 0;
+u = ifft(fdiv);
+figure(4);plot(u, 'Linewidth', 2); legend('u');xlim([0,2000]);
+v = ifft(fdiv.*nff);
+figure(5);semilogy(1:length(x), ldist, 1:length(x), v, 'Linewidth', 2);legend('ldist', 'v');xlim([0,2000]);
+figure(6);plot(1:length(x), ldist, 1:length(x), v, 'Linewidth', 2);legend('ldist', 'v');xlim([0,2000]);

### 2016-05-12 nips_2016.tex
Updated writeup to include new list of references

@@
 theirs.}
+
+Some additional references to use:
+
+\begin{itemize}
+    \item \cite{icml2014c1_bardenet14} (ICML 2014) I think we definitely want to consider this. It
+    seems to be ``an alternative to the Cutting the MH budget approach'' and so it might also be
+    considered as a direct competitor.
+    \item \cite{conf/icml/AhnBW12} (ICML 2012) is the ``successor'' to Stochastic Gradient Langevin
+    Dynamics.
+    \item \cite{conf/uai/MaclaurinA14} (UAI 2014) seems like it would be interesting to state its
+    limitations.
+    \item \cite{mcmc_hamiltonian_2010} is a long paper and is written like a chapter in a textbook,
+    so it's for pedagogical reasons. It describes MCMC with Hamiltonian dynamics. It is useful to
+    gain some background on a lot of the HMC-related papers that have come since its publication
+    (2010).
+    \item \cite{stochastic_thermostats_2014} (NIPS 2014) might be considered as a ``competitor''
+    though only for HMC cases.
+    \item \cite{sgmcmc_2015} (NIPS 2015) might want to give this a mention, where it would fit in
+    their recipe.
+\end{itemize}
 

### 2016-05-13 nips_2016.tex
Updated abstract of paper

@@
 
-\title{Fast Mini-Batch MCMC for Big Data Bayesian Posterior Inference}
+\title{Fast Mini-Batch MCMC for Big Data Parameter Optimization}
 
@@
 \begin{document}
-% \nipsfinalcopy is no longer used
+
+\begin{comment}
+\end{comment}
 
@@
 \begin{abstract}
-Markov chain Monte Carlo (MCMC) methods have become one of the most popular techniques in machine
-learning, but they can be intractably slow when doing Bayesian posterior estimation due to the need
-to use information from all training data points during each sampling iteration. We propose a
-mini-batch approach that uses information from a small subset of samples per iteration. This, along
-with a new Metropolis-Hastings test, will allow us to sample faster to obtain higher-quality
-parameter estimates. We implement our sampler in BIDMach and benchmark it with the state of the art.
-{\color{blue} Daniel: I hope we can do this.}
-{\color{blue} Xinlei: Compare with "cutting MH test budget" paper?}
+Markov chain Monte Carlo (MCMC) methods are one of the most popular techniques in machine learning
+for Bayesian inference. With large datasets, however, MCMC methods can be intractably slow due to
+the need to compute the likelihood of all the data items. Borrowing from the optimization
+literature, one intuitive way to make MCMC faster is to compute necessary gradients or likelihood
+ratios based only on a random subset of the data each iteration. Such stochastic MCMC has received
+increased research attention in recent years, but it is relatively difficult to develop formal
+guarantees for those algorithms. Some stochastic MCMC methods also involve adaptively increasing the
+minibatch size, which can nullify the benefits of the method if too much data is used. In addition,
+with large datasets, it is unclear if obtaining a full posterior distribution is necessary or better
+than deriving just a single MAP or point estimate of the parameters. In this paper, we present an
+alternative perspective on stochastic MCMC based on deriving high-quality point estimates of
+parameters in the presence of big data. To this end, we present a different Metropolis-Hastings test
+designed to be more efficient in the state space. We also augment the target posterior distribution
+with a temperature parameter, alleviating the need to adjust minibatch sizes.  We theoretically
+show the benefits of our approach, and verify with empirical benchmarks on synthetic data and neural
+network optimization.
 \end{abstract}
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 
-MCMC has become one of the most popular techniques in machine learning...
+Markov chain Monte Carlo (MCMC) sampling is one of the most popular techniques for Bayesian
+computation. In general, an MCMC algorithm will propose samples from a proposal distribution $q$,
+and decide whether to accept or reject the new sample based on a Metropolis-Hastings
+test~\cite{Metropolis1953,hastings70}. With traditional Bayesian posterior inference, the posterior
+probability $p(\theta \mid x_1, \ldots x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ based
+on i.i.d. data $\{x_i\}_{i=1}^n$ is the target distribution which one aims to approximate by
+sampling. This formula is problematic when $n$ is large (e.g., on the order of billions) because the
+likelihood must be evaluated based on each individual point.
 
-In this paper, we resolve these challenges\footnote{{\color{blue} Daniel: we can write these
-``challenges'' later once the project becomes clearer}.} by proposing an MCMC algorithm for
-Bayesian parameter estimation in which proposals are fast and mini-batch based. Our M-H test
-acceptance probability is about 50 percent, reasonably high, which means our samples mix well.We
-experimentally show that our approach is better than the state of the art.
+and~\cite{RobbinsMonro1951}.
 
-{\color{blue} Daniel: we can fill out this section later.}
+In this paper, we resolve these challenges by proposing an MCMC algorithm for Bayesian parameter
+estimation in which proposals are fast and mini-batch based. Our M-H test acceptance probability is
+about 50 percent, reasonably high, which means our samples mix well.We experimentally show that our
+approach is better than the state of the art.
 
-\section{Related Work}
+Our goal is using MH testing is in the context of MCMC model optimization, especially for neural
+networks. We are also interested in using it for other big data problems. We are not interested in
+exact inference of the posterior model parameters for a very large dataset.
+
+The example you have been using, the mixture of two gaussians, illustrates this point well. The
+"full data" posterior  based on 300 points has two sharp modes. The "cutting the MH budget" paper
+takes on average about half the data points to make steps on this distribution. That's intuitive.
+But it is also not very useful. The more data you have, the sharper the modes based on all the data.
+On a non-toy problem (e.g. > 10,000 points), the posterior would collapse to two points from which
+the MCMC simulation could likely never escape. And the "cutting the budget" method would still
+require almost all the data to make steps.
+
+For any real big-data problem (billions of points) the story would be the same.
+
+We should argue instead that for model parameter inference, the goal is either MAP or ML estimation.
+In both cases the goal is to find a good posterior mode. But the exact form of the distribution is
+arbitrary, and e.g. non-unit powers of the distribution are fine. That corresponds to changing the
+temperature of the distribution, or changing the scale factor in log delta space (which the factor
+N/n is doing).
+
+So rather than fixing the posterior temperature and trying to use adjustable mini-batch size to
+achieve it (which ends up reducing to nearly-full-batch updates) we are proposing to run at the
+natural temperature induced by a given minibatch size. We can always increase the mini-batch size to
+reduce the temperature. But there is no need to do this beyond the point where the posterior mode is
+sharp and well-defined. That will occur at some fixed number of points (e.g. a few thousand,
+depending on the problem), and far less than the full dataset.
+
+\section{Related Work}\label{sec:related_work}
 
@@
 
+\subsection{Experiment 3}
+
 \section{Conclusion}\label{sec:conclusion}

### 2016-05-13 nips_2016.tex
Added appendix with some details on gaussians, will email to john tomorrow around noontime

@@
 \usepackage{nips_2016}
-
 % to compile a camera-ready version, add the [final] option, e.g.:
@@
 
-\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings}
+\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem}
 \graphicspath{{figures/}}
@@
 % names and addresses of multiple authors: \And and \AND.
-%
 % Using \And between authors leaves it to LaTeX to determine where to break the lines. Using \AND
@@
 
+%%  David S.~Hippocampus\thanks{Use footnote for providing further
+%%    information about author (webpage, alternative
+%%    address)---\emph{not} for acknowledging funding agencies.} \\
+%%  Department of Computer Science\\
+%%  Cranberry-Lemon University\\
+%%  Pittsburgh, PA 15213 \\
+%%  \texttt{hippo@cs.cranberry-lemon.edu} \\
+
 \author{
-  David S.~Hippocampus\thanks{Use footnote for providing further
-    information about author (webpage, alternative
-    address)---\emph{not} for acknowledging funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  %% examples of more authors
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
+  John Canny \\
+  Department of Computer Science \\
+  University of California, Berkeley \\
+  \texttt{canny@berkeley.edu}
+  \And
+  Haoyu Chen \\
+  Department of Computer Science \\
+  University of California, Berkeley \\
+  \texttt{haoyuchen@berkeley.edu}
+  \And
+  Daniel Seita \\
+  Department of Computer Science \\
+  University of California, Berkeley \\
+  \texttt{seita@berkeley.edu}
+  \And
+  Xinlei Pan \\
+  Department of Bioengineering \\
+  University of California, Berkeley \\
+  \texttt{xinleipan@berkeley.edu}
+  \And 
+  Biye Jiang \\
+  Department of Computer Science \\
+  University of California, Berkeley \\
+  \texttt{bjiang@berkeley.edu}
 }
@@
 
-and~\cite{RobbinsMonro1951}.
+We can use stochasticity~\cite{RobbinsMonro1951}.
 
@@
 
-Cutting the Metropolis-Hastings Budget paper \cite{ cutting_mh_2014} proposed that there is a bias-variance trade-off in standard MCMC methods.
-Bias occurs when sampling from the wrong distribution, however, bias usually decreases fast.
-Variance occurs when the number of sampling is not enough since we can collect only a limited number of samples in a given amount of time.
-The recently proposed Stochastic Gradient Langevin Dynamics (SGLD) \cite{langevin_2011} and Stochastic Gradient Fisher Scoring methods are biased because they omit the required Metropolis-Hastings tests.
-
 Hamiltonian Monte Carlo (HMC) methods attempt to simulate the ``physics'' of the
@@
 
-The Stochastic Gradient Hamiltonian Monte Carlo (SGHMC) method~
-\cite{sghmc_2014} tries to compromise
-by using a \emph{mini-batch version} of HMC, so each iteration, only the mini-batch is used to
-compute the gradient.  SGHMC also includes an extra friction term to counteract the effect of noise
-on the system (or gradient) that is added due to using only a small subset of data. (The ``naive''
-way of replacing the batch with the mini-batch \emph{without} friction means the system deviates
-from the target/desired distribution.) ~\cite{sghmc_2014} assumes the injected noise is Gaussian (we
-make a similar assumption) and appeals to the Central Limit Theorem and second-order Langevin
-dynamics\footnote{There is a class of MCMC method techniques called \emph{Langevin dynamics}~\cite{mcmc_hamiltonian_2010}.
-Those algorithms involve computing a gradient update for the parameter of interest $\theta$ and then using the updated $\theta$ as the next sample (i.e., this is like the proposal test).
-Moreover, by simulating the Hamiltonian dynamics of the system, there is no need for a rejection test.
-To converge to a \emph{full (posterior) distribution} of $\theta$ and not a single point (i.e., the MAP estimate of $\theta$, or the mode of the distribution) langevin dynamics adds Gaussian noise each iteration to cause the $\theta$ samples to ``jump around.'' The work of~\cite{langevin_2011} is closer to ours because it applies the normal Langevin dynamics algorithm with one difference: the gradient updates are based on a \emph{mini-batch} of the original data.} to show that SGHMC maintains the desired target distribution as the stationary
-distribution. Our methods are similar in that we use mini-batches of data, but the difference is
-that they focus on improving a particular flavor of MCMC, \emph{Hamiltonian} Monte Carlo (which uses Langevin dynamics to add noise), and they
-also try avoiding MH tests, while we are focused on improving \emph{general} MCMC by means of a \emph{more efficient} MH proposal.
-These are different ways of solving the same problem (i.e., the problem of using mini-batches of
-large data for MCMC).
+The Stochastic Gradient Hamiltonian Monte Carlo (SGHMC) method~ \cite{sghmc_2014} tries to
+compromise by using a \emph{mini-batch version} of HMC, so each iteration, only the mini-batch is
+used to compute the gradient.  SGHMC also includes an extra friction term to counteract the effect
+of noise on the system (or gradient) that is added due to using only a small subset of data. (The
+``naive'' way of replacing the batch with the mini-batch \emph{without} friction means the system
+deviates from the target/desired distribution.) ~\cite{sghmc_2014} assumes the injected noise is
+Gaussian (we make a similar assumption) and appeals to the Central Limit Theorem and second-order
+Langevin dynamics\footnote{There is a class of MCMC method techniques called \emph{Langevin
+dynamics}~\cite{mcmc_hamiltonian_2010}. Those algorithms involve computing a gradient update for the
+parameter of interest $\theta$ and then using the updated $\theta$ as the next sample (i.e., this is
+like the proposal test). Moreover, by simulating the Hamiltonian dynamics of the system, there is no
+need for a rejection test. To converge to a \emph{full (posterior) distribution} of $\theta$ and not
+a single point (i.e., the MAP estimate of $\theta$, or the mode of the distribution) langevin
+dynamics adds Gaussian noise each iteration to cause the $\theta$ samples to ``jump around.'' The
+work of~\cite{langevin_2011} is closer to ours because it applies the normal Langevin dynamics
+algorithm with one difference: the gradient updates are based on a \emph{mini-batch} of the original
+data.} to show that SGHMC maintains the desired target distribution as the stationary distribution.
+Our methods are similar in that we use mini-batches of data, but the difference is that they focus
+on improving a particular flavor of MCMC, \emph{Hamiltonian} Monte Carlo (which uses Langevin
+dynamics to add noise), and they also try avoiding MH tests, while we are focused on improving
+\emph{general} MCMC by means of a \emph{more efficient} MH proposal. These are different ways of
+solving the same problem (i.e., the problem of using mini-batches of large data for MCMC).
 
-The work of~\cite{cutting_mh_2014} also concerns speeding up MCMC 
-by using mini-batches of data. Their argument is that the standard MCMC algorithm for Bayesian
-posterior inference (i.e., to compute $p(\theta \mid x_1, \ldots, x_N)$ for i.i.d. data points) is
-unsuitable for large datasets because the likelihood computation requires the use of all points,
-$p(x_1, \ldots, x_n \mid \theta) = \prod_{i=1}^Np(x_i\mid \theta)$. Using all of this data just to
-decide whether the new sample $\theta$ should be accepted or rejected is computationally
-inefficient. In theory, given infinite time, using all the data points is fine because it results in
-an unbiased estimate of the target distribution and the variance of the resulting sampled values is
-low given enough time.\footnote{Note that~\cite{cutting_mh_2014} primarily use MCMC to estimate just
-the expectation of the target distribution, but their analysis should extend to other values of the
-target distribution for arbitrary input $\theta$. In other words, they use MCMC to estimate
+The work of~\cite{cutting_mh_2014} also concerns speeding up MCMC by using mini-batches of data.
+Their argument is that the standard MCMC algorithm for Bayesian posterior inference (i.e., to
+compute $p(\theta \mid x_1, \ldots, x_N)$ for i.i.d. data points) is unsuitable for large datasets
+because the likelihood computation requires the use of all points, $p(x_1, \ldots, x_n \mid \theta)
+= \prod_{i=1}^Np(x_i\mid \theta)$. Using all of this data just to decide whether the new sample
+$\theta$ should be accepted or rejected is computationally inefficient. In theory, given infinite
+time, using all the data points is fine because it results in an unbiased estimate of the target
+distribution and the variance of the resulting sampled values is low given enough
+time.\footnote{Note that~\cite{cutting_mh_2014} primarily use MCMC to estimate just the expectation
+of the target distribution, but their analysis should extend to other values of the target
+distribution for arbitrary input $\theta$. In other words, they use MCMC to estimate
 $\mathbb{E}[p(\theta \mid x_1, \ldots, x_N)]$, but we are mostly interested in $p(\theta \mid x_1,
@@
 increment until we accept or reject $\theta'$ with high confidence (based on a Student-t
-distribution). The downside of this algorithm is the need to keep incrementing the mini-batch size in one iteration; in the worst case, they may not be able to tell with confidence to accept or
+distribution). The downside of this algorithm is the need to keep incrementing the mini-batch size
+in one iteration; in the worst case, they may not be able to tell with confidence to accept or
 reject until we get nearly all $N$ of the data points, which defeats the purpose of the algorithm.
@@
 \begin{lemma}\label{lem:detailed_balance}
-Let $\Delta = \log \left(\frac{f(\theta') q(\theta_t \mid
-\theta')}{f(\theta_t) q(\theta'\mid \theta_t)} \right)$, where $f$ is proportional to the desired target distribution and $q$ is our chosen
-proposal distribution. Any acceptance function $g$ such that $g(\Delta) = \exp(\Delta) g(-\Delta )$
-satisfies detailed balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) = f(\theta')p(\theta_t
-\mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the probability of jumping from $\theta_x$ to $\theta_y$ in our chain.
+Let $\Delta = \log \left(\frac{f(\theta') q(\theta_t \mid \theta')}{f(\theta_t) q(\theta'\mid
+\theta_t)} \right)$, where $f$ is proportional to the desired target distribution and $q$ is our
+chosen proposal distribution. Any acceptance function $g$ such that $g(\Delta) = \exp(\Delta)
+g(-\Delta )$ satisfies detailed balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) =
+f(\theta')p(\theta_t \mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the probability of jumping
+from $\theta_x$ to $\theta_y$ in our chain.
 \end{lemma}
@@
 
-\subsection{Experiment 1}
+\subsection{Mixture of Gaussians}\label{ssec:gaussians}
 
-\subsection{Experiment 2}
+We start with a simple mixture of Gaussians.
 
-\subsection{Experiment 3}
+\begin{figure}[ht]
+  \centering
+  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  %\includegraphics[width=0.5\linewidth]{empty}
+  \caption{This will be a sequence of images for the Gaussian mixture.}
+\end{figure}
 
-\section{Conclusion}\label{sec:conclusion}
+{\color{blue}
+Daniel: TODO
+}
+
+\subsection{Logistic Regression}\label{ssec:logistic}
+
+\begin{figure}[ht]
+  \centering
+  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  %\includegraphics[width=0.5\linewidth]{empty}
+  \caption{This will be a sequence of images for logistic regression.}
+\end{figure}
+
+{\color{blue}
+Daniel: TODO
+}
+
+\subsection{Neural Network Optimization}\label{ssec:nets}
+
+\begin{figure}[ht]
+  \centering
+  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  %\includegraphics[width=0.5\linewidth]{empty}
+  \caption{This will hopefully show results for neural networks on BIDMach.}
+\end{figure}
+
+{\color{blue}
+Daniel: TODO
+}
+
+\section{Conclusions}\label{sec:conclusion}
+
+{\color{blue}
+Daniel: TODO
+}
 
@@
 
-\section{Section for Appendix Information}
+\begin{center}
+{\Large Outline of Appendix}
+\end{center}
+
+In this appendix, we describe the following topics, with an emphasis on clarity and understanding:
+
+\begin{itemize}[noitemsep]
+    \item Detailed discussion of the experiment from Section~\ref{ssec:gaussians}.
+    \item Detailed discussion of the experiment from Section~\ref{ssec:logistic}.
+    \item Detailed discussion of the experiment from Section~\ref{ssec:nets}.
+\end{itemize}
+
+\section{Gaussian Mixture Experiment Details}
+
+\subsection{Mathematical Assumptions}
+
+We borrow this example from~\cite{langevin_2011}. Our parameter is a two dimensional vector, denoted
+as $\theta = (\theta_1, \theta_2)$, where we have
+\begin{equation}
+\theta_1 \sim \mathcal{N}(0, \sigma_1^2) \quad \mbox{ and } \quad \theta_2 \sim \mathcal{N}(0, \sigma_2^2)
+\end{equation}
+where $\mathcal{N}$ indicates the normal distribution (more generally, the multivariate normal). We
+consider the above as our prior. Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and
+$\sigma_2^2=1$, so the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
+log prior probability we endow on $\theta$ is
+\begin{equation}
+\log p(\theta) = \log \left(\frac{1}{2\pi\sqrt{10}}\right) - \frac{1}{2}\theta^T\Sigma^{-1}\theta.
+\end{equation}
+To generate the data, we use the following Gaussian mixture with tied means:
+\begin{equation}\label{eq:x_points}
+x_i \sim \frac{1}{2}\mathcal{N}(\theta_1, \sigma_x^2) + \frac{1}{2}\mathcal{N}(\theta_1+\theta_2, \sigma_x^2)
+\end{equation}
+where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$. This means the log likelihood
+of a single data instance is
+\begin{equation}
+\log p(x_i \mid \theta) = \log\left(\frac{1}{4\sqrt{\pi}}\right) +
+\log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
+(\theta_1+\theta_2))^2\right)\right)
+\end{equation}
+Here is the problem statement. Given some number of i.i.d. data points $x_1, x_2, \ldots, x_N$
+generated according to~(\ref{eq:x_points}), determine the posterior distribution\footnote{Note that,
+as is typical with Bayesian analysis, we do not concern ourselves with the denominator of the
+original (non-log) posterior.} of $\theta$:
+\begin{equation}\label{eq:log_post}
+\log p(\theta \mid x_1,\ldots,x_n) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
+\end{equation}
+Alternatively, if there are too many data points, we may opt to instead pick a point estimate of
+$\theta$, generally the MAP estimate. (If $N$ is extremely large, it will cause the posterior to
+peak sharply at its modes, reducing distribution estimates to point estimates.) Note that in many
+cases, we will need to take a \emph{minibatch estimate} of~(\ref{eq:log_post}). In that case, the
+literature generally uses
+\begin{equation}\label{eq:scaling_factor}
+\log p(\theta \mid x_1,\ldots,x_n) \approx \log p(\theta) + \frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
+\end{equation}
+where we only use $n \ll N$ samples, but we must scale up the likelihood contribution by $N/n$. If we
+didn't add this scaling factor, then the contribution of the likelihood terms would be weaker. We
+we will discuss this scaling factor in more detail later.
+
+Finally, recall earlier that we needed to define $\Delta$. In our context, $\Delta$ is:
+\begin{align}
+\Delta &= \log \left(\frac{f(\theta') q(\theta_t \mid \theta')}{f(\theta_t) q(\theta'\mid \theta_t)} \right) \\
+&= \log p(\theta') - \log p(\theta_t) + \sum_{i=1}^N(\log p(x_i \mid \theta') - \log p(x_i \mid \theta_t)) + \log\left(\frac{q(\theta_t \mid \theta')}{q(\theta' \mid \theta_t)}\right).
+\end{align}
+
+\subsection{Experiment Observations}
+
+Figure~\ref{fig:contour} describes a contour plot of the posterior based on the points we have
+sampled. The goal is to get distributions that either cover this posterior well (i.e., identifying
+that there are two modes), or if the data is too large and we want MAP estimates, we wish to see the
+methods stay at one of the modes.
+
+In the following, we only consider random walk proposals, so the term in $\Delta$ containing the
+$q$s disappears.  Random walk proposals are bad, but since the proposal is poor, good performance
+can only be obtained with a strong MH test, hence why we can use this as a reasonable starting
+benchmark.
+
+We use the settings:
+
+\begin{itemize}[noitemsep]
+    \item We use $N=10000$ samples of $x_i$.
+    \item The covariance of the Gaussian random walk is ${\rm diag}(0.15, 0.15)$.
+    \item The mini-batch size is ...
+    \item (more settings) ...
+    \item For the adaptive sampling method, the tolerance for deciding on a test is ....
+    \item The number of iterations is .... One iteration involves one $\theta'$ proposal and a test
+    to accept/reject.
+\end{itemize}
+
+Figure~\ref{fig:scatter} compares scatter plots of the ``Cutting the MH'' test (hereafter called
+``adaptive minibatch'') versus ``Our Method''.
+
+Figure~\ref{fig:cutting_mh} describes the histograms of the adaptive minibatch's batch sizes.
+
+Figure~\ref{fig:diagnostics} describes the histograms of various statistics for our method.
+
+\begin{figure}[ht]
+  \centering
+  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  %\includegraphics[width=0.5\linewidth]{empty}
+  \caption{TODO contour plots}
+  \label{fig:contour}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  %\includegraphics[width=0.5\linewidth]{empty}
+  \caption{Scatter plots}
+  \label{fig:scatter}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  %\includegraphics[width=0.5\linewidth]{empty}
+  \caption{Diagnostic Plots 1 (e.g., histigrams of cutting MH)}
+  \label{fig:cutting_mh}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  %\includegraphics[width=0.5\linewidth]{empty}
+  \caption{Diagnostic Plots 2 (sd, delta, etc)}
+  \label{fig:diagnostics}
+\end{figure}
 
 {\color{blue}
-Daniel: In case we want to use this.
+Daniel: TODO FINISH
+}
+
+\section{Logistic Regression Experiment Details}
+
+{\color{blue}
+Daniel: TODO
+}
+
+\section{Neural Network Experiment Details}
+
+{\color{blue}
+Daniel: TODO
 }
@@
 % Daniel: here are example LaTeX codes for figures and tables if we want to use them.
-
 %\begin{figure}[h]
@@
 %\end{figure}
-
 %\begin{table}[t]
@@
 %\end{table}
-
 %\usepackage[pdftex]{graphicx} ...

### 2016-05-13 nips_2016.tex
Appendix A on Gaussian Mixtures

@@
 \bibliographystyle{ieeetr}
+\normalsize
 
@@
 
-We borrow this example from~\cite{langevin_2011}. Our parameter is a two dimensional vector, denoted
-as $\theta = (\theta_1, \theta_2)$, where we have
+We borrow this example from~\cite{langevin_2011}. Our parameter is a 2-D vector $\theta =
+(\theta_1, \theta_2)$, where
 \begin{equation}
@@
 \begin{equation}\label{eq:log_post}
-\log p(\theta \mid x_1,\ldots,x_n) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
+\log p(\theta \mid x_1,\ldots,x_N) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
 \end{equation}
@@
 \begin{equation}\label{eq:scaling_factor}
-\log p(\theta \mid x_1,\ldots,x_n) \approx \log p(\theta) + \frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
+\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
 \end{equation}
 where we only use $n \ll N$ samples, but we must scale up the likelihood contribution by $N/n$. If we
-didn't add this scaling factor, then the contribution of the likelihood terms would be weaker. We
-we will discuss this scaling factor in more detail later.
+didn't add this scaling factor, then the contribution of the likelihood terms would be weaker. There
+is another thing we need to consider about that, which is the \emph{temperature} of our
+distribution. In general, we will want to add $T > 1$ so that our posterior is
+$p(\theta)((\prod_{i=1}^n p(x_i\mid \theta))^{N/n})^{1/T}$, resulting in the log
+posterior of 
+\begin{equation}\label{eq:log_prior_temp}
+\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
+\end{equation}
+which has the extra $1/T$ to decrease the scale factor.  Equation~(\ref{eq:log_prior_temp}) is what
+we will be using for our experiments, because we need to consider warmer distributions.
+{\color{blue} Daniel: not totally confident on this but I'll try.}
 
@@
 \end{align}
+But again, with \emph{minibatches} of data, we usually do not sum over all $N$ terms and instead sum
+over $n$ terms, but with the extra $(1/T)(N/n)$ scaling factor, from
+Equation~(\ref{eq:log_prior_temp}). We generally denote that approximation as $\Delta'$ in the paper.
 
@@
 
-Figure~\ref{fig:contour} describes a contour plot of the posterior based on the points we have
-sampled. The goal is to get distributions that either cover this posterior well (i.e., identifying
-that there are two modes), or if the data is too large and we want MAP estimates, we wish to see the
-methods stay at one of the modes.
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{contour_v1}
+  \caption{The posterior distribution, from 50 to 100k samples, with temperature set at 1.}
+  \label{fig:contour1}
+\end{figure}
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{contour_v2}
+  \caption{The posterior distribution, with $N=10000$ but with temperature $T$ varying.}
+  \label{fig:contour2}
+\end{figure}
 
-In the following, we only consider random walk proposals, so the term in $\Delta$ containing the
-$q$s disappears.  Random walk proposals are bad, but since the proposal is poor, good performance
-can only be obtained with a strong MH test, hence why we can use this as a reasonable starting
-benchmark.
+In this experiment, we compare our method with the one from~\cite{cutting_mh_2014}, hereafter called
+the ``adaptive sampling approach''. We apply them to the Gaussian Mixture model previously
+discussed.
 
-We use the settings:
+Figure~\ref{fig:contour1} shows simulated contour plots of the posterior based on the $N$ data
+points, for varying $N$, with the temperature set at $T=1$. Note that because we're using all $N$
+points here, the scale factor $N/n=1$. As $N$ increases, the posterior converges to a multimodal
+distribution with modes at $(0,1)$ and $(1,-1)$.
+
+Figure~\ref{fig:contour2} is similar, except this time we fix the number of samples at $N=10000$,
+but show how changing the temperature $T$ affects the distribution. A larger $T$ implies a flatter
+posterior, one that (weakly) peaks in between the two true modes.
+
+Our ultimate goal is to either (a) get a posterior distribution that covers the posterior well, or
+(b) to get samples that stay trapped at a mode, for a point estimate of $\theta$.  Whether our goal
+is (a) or (b) depends on our objectives and the context of the problem (e.g., big data usually
+implies (b) is OK even if it would not satisfy the average Bayesian).
+
+For both our method and the adaptive sampling method, we use random walk proposals. (Thus, the term
+in $\Delta$ containing the $q$s disappears.)  Random walk proposals are bad, but since the proposal
+is poor, good performance can only be obtained with a strong MH test, hence why we can use this as a
+reasonable starting benchmark.
+
+We test use the following settings:
 
@@
     \item We use $N=10000$ samples of $x_i$.
-    \item The covariance of the Gaussian random walk is ${\rm diag}(0.15, 0.15)$.
-    \item The mini-batch size is ...
-    \item (more settings) ...
-    \item For the adaptive sampling method, the tolerance for deciding on a test is ....
-    \item The number of iterations is .... One iteration involves one $\theta'$ proposal and a test
-    to accept/reject.
+    \item The covariance of the Gaussian random walk is ${\rm diag}(0.1, 0.1)$.
+    \item We use 10000 iterations. One iteration involves one $\theta'$ proposal and a test to
+    accept/reject.
+    \item The mini-batch size is set at 100, so it is one percent of the total data. For the
+    adaptive sampling approach, this is the amount that the minibatch size would keep increasing if
+    we needed more precision. Our approach, of courses, keeps the size fixed.
+    \item For adaptive sampling, the tolerance for deciding on a test is 0.05. The temperature for
+    that is also set at $T=1$ because that algorithm was not designed to deal with temperatures.
+    \item For our approach, to get an estimate of the standard deviation of $\Delta'$ (remember,
+    $\Delta'$ is a minibatch approximation of $\Delta$) we take five other minibatches (randomly
+    selected from the data) and take the standard deviation of those values.
 \end{itemize}
 
+For our approach, we run using three different temperature values, $T=1$, $T=10$, and $T=100$.
+
 Figure~\ref{fig:scatter} compares scatter plots of the ``Cutting the MH'' test (hereafter called
-``adaptive minibatch'') versus ``Our Method''.
-
-Figure~\ref{fig:cutting_mh} describes the histograms of the adaptive minibatch's batch sizes.
-
-Figure~\ref{fig:diagnostics} describes the histograms of various statistics for our method.
+``adaptive minibatch'') versus ``Our Method'', where we use three different temperature settings.
+There doesn't seem to be too much noticeable difference in our method for the three different
+temperature setting values, but our approach is a lot different from the adaptive sampling method
+(this is because of rejection rates, more on that later).
 
@@
   \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  %\includegraphics[width=0.5\linewidth]{empty}
-  \caption{TODO contour plots}
-  \label{fig:contour}
+  \includegraphics[width=1\linewidth]{scatter_v01}
+  \caption{Scatter plots of accepted $\theta$ for (in order) the adaptive sampling approach, and our
+  method (temperatures 1, 10, and 100).}
+  \label{fig:scatter}
+\end{figure}
+
+Before taking a look at our approach, we first examine the adaptive sampling approach. \textbf{That
+approach rejected a lot of samples; out of 10000 iterations, it only accepted 773 times}.
+Figure~\ref{fig:mb_sizes} describes the histograms of the adaptive minibatch's batch sizes. (Note
+that the histogram y-axis values are different.) In many cases, using the initial batch size of 100
+sufficed to make an accept/reject decision, but sometimes, this method needed to use the entire 10k
+data set. Decreasing the tolerance (from its current 0.05 value) would have required larger
+minibatches. Incidentally, it makes sense that the method rejects a lot of samples, because the
+random walk proposal usually provides us with poor points.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.75\linewidth]{adaptive_sampling_sizes_v01.png}
+  \caption{Minibatch sizes for adaptive sampling when it was accepting or rejecting.}
+  \label{fig:mb_sizes}
+\end{figure}
+
+Now we analyze our approach in more detail. Our method accepts far more samples than the adaptive
+sampling method, \textbf{accepting 3285, 3318, and 3332 out of 10k}, for temperature values 1, 10,
+and 100, respectively.
+
+For diagnostics purposes, Figures~\ref{fig:diagnostics1},~\ref{fig:diagnostics2},
+and~\ref{fig:diagnostics3} express histograms (for all three temperature runs) of the $\Delta'$,
+${\rm std}(\Delta')$, and $X_{\rm corr}$ values, respectively.
+
+There are a few immediate observations. First, increasing the temperature size by a factor of 10
+will decrease $\Delta'$ by a factor of 10, which makes sense. That will also directly affect the
+standard deviation estimate of $\Delta'$. Our $X_{\rm corr}$ distribution is symmetric about zero,
+which is also expected (and it also exhibits a decrease in a factor of 10 for an increase
+in temperature).
+
+{\color{blue}
+One concern I have is that our $\Delta'$ values do not seem to be ``sufficiently negative''. If our
+$\Delta'$ values are negative (but large in absolute value) that means our MH test will reject more
+often. But then this raises another concern: didn't we say at one point that we \emph{should} be
+getting high acceptance rates, around 50 percent (in part because of the shape of the logistic
+function)? Doesn't that depend on the quality of the proposal distribution? If our method is
+designed to accept a lot regardless of the proposal distribution, I don't see how its performance
+can be considered equal to the adaptive sampling approach, or even a standard MH test. In fact, the
+best observation I can derive from this is that our approach and the adaptive sampling (or any
+standard MH test) will accept the high likelihood cases, but our case will accept the borderline
+ones more often.
+}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1\linewidth]{our_deltas_v01.png}
+  \caption{Our $\Delta'$ values for three different temperature values.}
+  \label{fig:diagnostics1}
 \end{figure}
@@
   \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  %\includegraphics[width=0.5\linewidth]{empty}
-  \caption{Scatter plots}
-  \label{fig:scatter}
+  \includegraphics[width=1\linewidth]{our_sds_v01.png}
+  \caption{Our ${\rm std}(\Delta')$ values for three different temperature values.}
+  \label{fig:diagnostics2}
 \end{figure}
@@
   \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  %\includegraphics[width=0.5\linewidth]{empty}
-  \caption{Diagnostic Plots 1 (e.g., histigrams of cutting MH)}
-  \label{fig:cutting_mh}
+  \includegraphics[width=1\linewidth]{our_xcorrs_v01.png}
+  \caption{Our $X_{\rm corr}$ values for three different temperature values.}
+  \label{fig:diagnostics3}
 \end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  %\includegraphics[width=0.5\linewidth]{empty}
-  \caption{Diagnostic Plots 2 (sd, delta, etc)}
-  \label{fig:diagnostics}
-\end{figure}
-
-{\color{blue}
-Daniel: TODO FINISH
-}
 

### 2016-05-15 nips_2016.tex
Update to paper draft. More organized, new title, sections 1 and 2 in draft stage, more references.

@@
 
-\title{Fast Mini-Batch MCMC for Big Data Parameter Optimization}
+\title{A New Metropolis-Hastings Test for Fast Minibatch MCMC Inference and Optimization}
 
@@
 for Bayesian inference. With large datasets, however, MCMC methods can be intractably slow due to
-the need to compute the likelihood of all the data items. Borrowing from the optimization
-literature, one intuitive way to make MCMC faster is to compute necessary gradients or likelihood
-ratios based only on a random subset of the data each iteration. Such stochastic MCMC has received
-increased research attention in recent years, but it is relatively difficult to develop formal
-guarantees for those algorithms. Some stochastic MCMC methods also involve adaptively increasing the
-minibatch size, which can nullify the benefits of the method if too much data is used. In addition,
-with large datasets, it is unclear if obtaining a full posterior distribution is necessary or better
-than deriving just a single MAP or point estimate of the parameters. In this paper, we present an
-alternative perspective on stochastic MCMC based on deriving high-quality point estimates of
-parameters in the presence of big data. To this end, we present a different Metropolis-Hastings test
-designed to be more efficient in the state space. We also augment the target posterior distribution
-with a temperature parameter, alleviating the need to adjust minibatch sizes.  We theoretically
-show the benefits of our approach, and verify with empirical benchmarks on synthetic data and neural
-network optimization.
+the need to compute the likelihood of all data points each iteration. Borrowing from the
+optimization field, one way to make MCMC faster is to compute the posterior based on a random subset
+of the data each iteration. Such stochastic MCMC has received increased research attention in recent
+years, but these methods generally require adaptively increasing the minibatch size each iteration,
+which can nullify the benefits if too much data is needed. In addition, with large datasets and
+model optimization, it is unclear if obtaining a full posterior distribution is desirable over
+getting a high-quality point MAP or ML estimate of the parameters.  In this paper, we present a new
+Metropolis-Hastings test to fit in our alternative perspective on stochastic MCMC methods for big
+data model optimization. We demonstrate the benefits of our approach with theoretical and empirical
+results.
 \end{abstract}
@@
 Markov chain Monte Carlo (MCMC) sampling is one of the most popular techniques for Bayesian
-computation. In general, an MCMC algorithm will propose samples from a proposal distribution $q$,
-and decide whether to accept or reject the new sample based on a Metropolis-Hastings
-test~\cite{Metropolis1953,hastings70}. With traditional Bayesian posterior inference, the posterior
-probability $p(\theta \mid x_1, \ldots x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ based
-on i.i.d. data $\{x_i\}_{i=1}^n$ is the target distribution which one aims to approximate by
-sampling. This formula is problematic when $n$ is large (e.g., on the order of billions) because the
-likelihood must be evaluated based on each individual point.
+computation. MCMC algorithms propose samples from a proposal distribution $q$, and decide whether to
+accept or reject them based on a Metropolis-Hastings test~\cite{Metropolis1953,hastings70}. With
+traditional Bayesian posterior inference, the posterior distribution $p(\theta \mid x_1, \ldots,
+x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ of parameter $\theta$ based on conditionally
+independent data $\{x_i\}_{i=1}^n$ is the target distribution which one aims to approximate by
+sampling $\theta$ values. When $n$ is large (e.g., in the billions), it is computationally difficult
+to compute the exact posterior because the likelihood must be evaluated on each $x_i$.
 
-We can use stochasticity~\cite{RobbinsMonro1951}.
+Recent success in machine learning optimization in the presence of large datasets have been in part
+due to the ubiquitous use of stochastic gradient methods, where a subset of the data is used each
+iteration to compute a noisy estimate of the gradient. These Robbins-Munro
+algorithms~\cite{RobbinsMonro1951} have historically not been used with MCMC methods. A recent
+result from~\cite{cutting_mh_2014} attempted to bridge the gap between minibatch optimization and
+Bayesian inference by proposing an \emph{adaptive minibatch} method. Their method uses a subset of
+the $N$ data points for the MH test each iteration, but requires incrementally increasing the
+minibatch size so that it is more confident in its prediction on accepting or rejecting.
 
-In this paper, we resolve these challenges by proposing an MCMC algorithm for Bayesian parameter
-estimation in which proposals are fast and mini-batch based. Our M-H test acceptance probability is
-about 50 percent, reasonably high, which means our samples mix well.We experimentally show that our
-approach is better than the state of the art.
+We argue for an alternative approach to using MCMC methods in the context of big data problems. In
+general, model optimization involves finding one point estimate of optimal parameters based on a
+user's goals, which typically refer to MAP or ML estimates. In addition, the more data one has, the
+sharper the posterior distribution's modes. In the extreme case, the posterior would collapse to the
+points corresponding to the modes, and it would be difficult for MCMC methods to escape.
 
-Our goal is using MH testing is in the context of MCMC model optimization, especially for neural
-networks. We are also interested in using it for other big data problems. We are not interested in
-exact inference of the posterior model parameters for a very large dataset.
+Therefore, we are not interested in exact inference of the posterior model parameters for a very
+large dataset, but rather, we want to perform either MAP or ML estimation. In both cases the goal is
+to find a good posterior mode, but the exact form of the distribution is arbitrary. In particular,
+non-unit powers of the distribution are fine. That corresponds to changing the temperature of the
+distribution to make it ``flatter''.
 
-The example you have been using, the mixture of two gaussians, illustrates this point well. The
-"full data" posterior  based on 300 points has two sharp modes. The "cutting the MH budget" paper
-takes on average about half the data points to make steps on this distribution. That's intuitive.
-But it is also not very useful. The more data you have, the sharper the modes based on all the data.
-On a non-toy problem (e.g. > 10,000 points), the posterior would collapse to two points from which
-the MCMC simulation could likely never escape. And the "cutting the budget" method would still
-require almost all the data to make steps.
+Rather than fixing the posterior temperature and trying to use adjustable mini-batch size to achieve
+it (which ends up reducing to nearly-full-batch updates) we propose to run at the natural
+temperature induced by a given minibatch size. We can always increase the minibatch size to reduce
+the temperature, but there is no need to do this beyond the point where the posterior mode is sharp
+and well-defined. That will occur at some fixed number of points and far less than the full dataset.
 
-For any real big-data problem (billions of points) the story would be the same.
+To this end, we present a new Metropolis-Hastings test that can fit in this framework of minibatch
+MCMC methods. We present theoretical and experimental results on our method and show that it can be
+the method of choice for minibatch MCMC .
 
-We should argue instead that for model parameter inference, the goal is either MAP or ML estimation.
-In both cases the goal is to find a good posterior mode. But the exact form of the distribution is
-arbitrary, and e.g. non-unit powers of the distribution are fine. That corresponds to changing the
-temperature of the distribution, or changing the scale factor in log delta space (which the factor
-N/n is doing).
+\section{Preliminaries and Related Work}\label{sec:related_work}
 
-So rather than fixing the posterior temperature and trying to use adjustable mini-batch size to
-achieve it (which ends up reducing to nearly-full-batch updates) we are proposing to run at the
-natural temperature induced by a given minibatch size. We can always increase the mini-batch size to
-reduce the temperature. But there is no need to do this beyond the point where the posterior mode is
-sharp and well-defined. That will occur at some fixed number of points (e.g. a few thousand,
-depending on the problem), and far less than the full dataset.
-
-\section{Related Work}\label{sec:related_work}
-
-In this section, we review different results on improving MCMC methods with a focus on those dealing
-with mini-batches of large datasets.
-
-Hamiltonian Monte Carlo (HMC) methods attempt to simulate the ``physics'' of the
-system\footnote{Intuitively, we can think of a probability distribution as being a distribution over
-particles in the system, where particles are the values of random variables.} in order to generate
-distant, high-quality sample proposals (those are the best kind of samples). Since they simulate the
-physical dynamics of the system, total energy is preserved and their proposals are always
-accepted and there is no need for a Metropolis-Hastings correction/test\footnote{I'm not sure why
-this follows but I haven't read the details.}. Unfortunately, HMC methods require the use of a
-gradient computation based on all of the data per iteration. (This means the advantage of HMC
-methods is that they generate high-quality proposals that mix well; their downside is that they are
-unsuitable for big data, which might involve Bayesian posterior inference over that data.)
-
-The Stochastic Gradient Hamiltonian Monte Carlo (SGHMC) method~ \cite{sghmc_2014} tries to
-compromise by using a \emph{mini-batch version} of HMC, so each iteration, only the mini-batch is
-used to compute the gradient.  SGHMC also includes an extra friction term to counteract the effect
-of noise on the system (or gradient) that is added due to using only a small subset of data. (The
-``naive'' way of replacing the batch with the mini-batch \emph{without} friction means the system
-deviates from the target/desired distribution.) ~\cite{sghmc_2014} assumes the injected noise is
-Gaussian (we make a similar assumption) and appeals to the Central Limit Theorem and second-order
-Langevin dynamics\footnote{There is a class of MCMC method techniques called \emph{Langevin
-dynamics}~\cite{mcmc_hamiltonian_2010}. Those algorithms involve computing a gradient update for the
-parameter of interest $\theta$ and then using the updated $\theta$ as the next sample (i.e., this is
-like the proposal test). Moreover, by simulating the Hamiltonian dynamics of the system, there is no
-need for a rejection test. To converge to a \emph{full (posterior) distribution} of $\theta$ and not
-a single point (i.e., the MAP estimate of $\theta$, or the mode of the distribution) langevin
-dynamics adds Gaussian noise each iteration to cause the $\theta$ samples to ``jump around.'' The
-work of~\cite{langevin_2011} is closer to ours because it applies the normal Langevin dynamics
-algorithm with one difference: the gradient updates are based on a \emph{mini-batch} of the original
-data.} to show that SGHMC maintains the desired target distribution as the stationary distribution.
-Our methods are similar in that we use mini-batches of data, but the difference is that they focus
-on improving a particular flavor of MCMC, \emph{Hamiltonian} Monte Carlo (which uses Langevin
-dynamics to add noise), and they also try avoiding MH tests, while we are focused on improving
-\emph{general} MCMC by means of a \emph{more efficient} MH proposal. These are different ways of
-solving the same problem (i.e., the problem of using mini-batches of large data for MCMC).
-
-The work of~\cite{cutting_mh_2014} also concerns speeding up MCMC by using mini-batches of data.
-Their argument is that the standard MCMC algorithm for Bayesian posterior inference (i.e., to
-compute $p(\theta \mid x_1, \ldots, x_N)$ for i.i.d. data points) is unsuitable for large datasets
-because the likelihood computation requires the use of all points, $p(x_1, \ldots, x_n \mid \theta)
-= \prod_{i=1}^Np(x_i\mid \theta)$. Using all of this data just to decide whether the new sample
-$\theta$ should be accepted or rejected is computationally inefficient. In theory, given infinite
-time, using all the data points is fine because it results in an unbiased estimate of the target
-distribution and the variance of the resulting sampled values is low given enough
-time.\footnote{Note that~\cite{cutting_mh_2014} primarily use MCMC to estimate just the expectation
-of the target distribution, but their analysis should extend to other values of the target
-distribution for arbitrary input $\theta$. In other words, they use MCMC to estimate
-$\mathbb{E}[p(\theta \mid x_1, \ldots, x_N)]$, but we are mostly interested in $p(\theta \mid x_1,
-\ldots, x_N)$.} In reality, we have limited time. The key argument in~\cite{cutting_mh_2014} is that
-given a fixed budget, it may be better to use $n$ data points per iteration (where $n \ll N$), which
-will sacrifice some bias but will perhaps improve variance by allowing us to sample for $\theta$
-more often, since we will get $\theta_1, \ldots, \theta_T$ as opposed to the standard case using all
-$N$ samples but with only $\theta_1, \ldots, \theta_t$ samples ($t \ll T$).
-
-More formally, the algorithm they develop is a sequential hypothesis testing algorithm. During each
-iteration, their algorithm generates a new sample $\theta'$ from the proposal distribution. They
-start with a small mini-batch of the data, and test the hypothesis (based on this small mini-batch)
-that $\theta'$ should be accepted or rejected based the likelihood ratio. (Intuitively, we can
-easily tell if $\theta'$ will be accepted/rejected if the mini-batch of data results in likelihood
-drastically different from the likelihood of the current sample $\theta$.) If the hypothesis test is
-too uncertain, then we increase the mini-batch size and apply the test again, and we repeat this
-increment until we accept or reject $\theta'$ with high confidence (based on a Student-t
-distribution). The downside of this algorithm is the need to keep incrementing the mini-batch size
-in one iteration; in the worst case, they may not be able to tell with confidence to accept or
-reject until we get nearly all $N$ of the data points, which defeats the purpose of the algorithm.
-We propose our own algorithm that avoids this problem. {\color{blue} Daniel: does this make sense as
-a ``weakness'' of their approach?  Ideally I would like to have a comparison between our method and
-theirs.}
-
-Some additional references to use:
-
-\begin{itemize}
-    \item \cite{icml2014c1_bardenet14} (ICML 2014) I think we definitely want to consider this. It
-    seems to be ``an alternative to the Cutting the MH budget approach'' and so it might also be
-    considered as a direct competitor.
-    \item \cite{conf/icml/AhnBW12} (ICML 2012) is the ``successor'' to Stochastic Gradient Langevin
-    Dynamics.
-    \item \cite{conf/uai/MaclaurinA14} (UAI 2014) seems like it would be interesting to state its
-    limitations.
-    \item \cite{mcmc_hamiltonian_2010} is a long paper and is written like a chapter in a textbook,
-    so it's for pedagogical reasons. It describes MCMC with Hamiltonian dynamics. It is useful to
-    gain some background on a lot of the HMC-related papers that have come since its publication
-    (2010).
-    \item \cite{stochastic_thermostats_2014} (NIPS 2014) might be considered as a ``competitor''
-    though only for HMC cases.
-    \item \cite{sgmcmc_2015} (NIPS 2015) might want to give this a mention, where it would fit in
-    their recipe.
-\end{itemize}
-
-\section{Our Algorithm}\label{sec:our_algorithm}
-
-\subsection{The General MCMC Method}\label{sec:general_mcmc}
-
-The general MCMC method proceeds as follows. For a desired random vector $\theta$, we have a
-\emph{target distribution} we wish to compute, $p(\theta \mid x_1, \ldots, x_N)$, based on $N$
-i.i.d. data points. Since exact evaluation is intractable, we generate a chain of correlated samples
-$\theta_1, \ldots, \theta_T$ for some large $T$, and approximate $p$ by using counts of samples. For
-each iteration $t$, we start with our current $\theta_t$. We use a \emph{proposal distribution}
-$q(\theta' \mid \theta_t)$ to determine a new candidate $\theta'$. With probability $P_a$, we accept
-it and set $\theta_{t+1} = \theta'$; otherwise, we repeat the previous value $\theta_{t+1} =
-\theta_t$. Traditionally, $P_a$ is computed as follows:
+The standard MCMC method proceeds as follows~\cite{gilks1996markov,brooks2011handbook}. For a
+parameter $\theta$ and conditionally independent data, we wish to compute the target distribution
+$p(\theta \mid x_1, \ldots, x_N)$. Since exact evaluation is intractable, we generate a chain of
+correlated samples $\theta_1, \ldots, \theta_T$ for some large $T$, and approximate $p$ by using
+counts of samples. For each iteration $t$, we start with our current $\theta_t$. We use a
+\emph{proposal distribution} $q(\theta' \mid \theta_t)$ to determine a new candidate $\theta'$. With
+probability $P_a$, we accept it and set $\theta_{t+1} = \theta'$; otherwise, we repeat the previous
+value $\theta_{t+1} = \theta_t$. Traditionally, $P_a$ is computed as follows:
 \begin{equation}\label{eq:traditional}
@@
 \end{equation}
-where $f(\theta_t)=p(\theta_t|x_1,\ldots,x_N) \propto p(\theta_t)\prod_{i=1}^N p(x_i
-\mid \theta_t)$ which roughly tells us how ``likely'' we are to have a $\theta_t$ sample. The
-normalizing constants cancel out, so the requirement for $P_a$ is that $f$ be proportional to the
-true target distribution.
+where $f(\theta_t)=p(\theta_t|x_1,\ldots,x_N) \propto p(\theta_t)\prod_{i=1}^N p(x_i \mid
+\theta_t)$. The technical requirement for $P_a$ is that $f$ be proportional to the true target
+distribution.
 
@@
 \theta'$ if $u < P_a$, and uses $\theta_{t+1} = \theta_t$ otherwise. The $P_a$ from
-Equation~\ref{eq:traditional} satisfies ``detailed balance'' which, roughly speaking, means that if
-we sample long enough using the heuristic above, we will arrive at a stationary distribution.
+Equation~\ref{eq:traditional} satisfies detailed balance, so if one samples long enough using the
+heuristic above, one will arrive at a stationary distribution.  The $P_a$ is guaranteed to converge
+to the true $p$ distribution given sufficiently many examples (though a burn-in period and/or taking
+every $n$th sample may be necessary).
 
-The $P_a$ is guaranteed to converge to the true $p$ distribution given sufficiently many examples
-(though a burn-in period and/or taking every $n$th sample may be necessary). Unfortunately,
-computing $f$ requires the use of all $N$ training data points. In this paper, we derive a new fast,
-approximate Metropolis-Hastings test which does not use all $N$ points during each test.
+Unfortunately, computing $f$ requires the use of all $N$ training data points. Moreover, it is
+difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
+balance. There has been research on ways to conduct exact accept/reject tests using unbiased
+estimators of the likelihood~\cite{Andrieu09thepseudo-marginal}. However, unbiased estimators of the
+likelihood that can be computed from minibatches of data, such as the Kennedy-Bhanot
+estimator~\cite{PhysRevD} have very high variance for large datasets. Because of this, once we get a
+very high estimate of the likelihood, almost all proposed moves are rejected and the algorithm gets
+stuck.
 
-\subsection{A Different Metropolis-Hastings Test}\label{sec:faster}
+{\color{blue}
+Daniel: I am worried. We need to motivate why it is not easy to use minibatch MCMC, and what other
+researchers have tried to do about that.  The last paragraph is from~\cite{cutting_mh_2014}, and it
+seems to argue that people have not used minibatch MCMC before ``the algorithm gets stuck''. But
+isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes? If we wanted
+to get stuck at modes with large datasets, why didn't we just do what previous work did?
+}
+
+The work of~\cite{cutting_mh_2014} proposes an adaptive minibatch approach by using a sequential
+hypothesis testing algorithm. During each iteration, they start with a small minibatch of data and
+test the hypothesis that the sample $\theta'$ should be accepted or rejected based the likelihood
+ratio. If the test cannot make a decision over a certain confidence threshold, then they increase
+the minibatch size and test again. This process repeats until a decision.  The downside of this
+algorithm is the need to keep incrementing the mini-batch size in one iteration; in the worst case,
+all $N$ data points may be needed in a single iteration. A related approach
+is~\cite{icml2014c1_bardenet14}, which has slightly more robust theoretical properties, but comes at
+the cost of more computational time. The work of~\cite{conf/uai/MaclaurinA14} presents an
+\emph{exact} approach, but has the sometimes unreasonable requirement of a lower bound on the
+per-datum likelihood factors.
+
+There is a separate line of work related to optimizing MCMC about simulating the physics of a
+probability distribution. By viewing random variables as ``particles'' in a system, it is possible
+to apply Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which generate both high
+quality and distant proposals, getting the best of both worlds. These require a full gradient
+computation of the posterior, which means they also suffer from the problem of needing to know the
+likelihood at each point. Stochastic versions of HMC~\cite{sghmc_2014,stochastic_thermostats_2014}
+use a minibatch of data each iteration but require controlling the noise from sampling so the sample
+sequence does not diverge. A similar approach is to use Langevin
+Dynamics~\cite{langevin_2011,conf/icml/AhnBW12}, which does not use the momentum terms in HMC.
+
+Ultimately, minibatch SGHMC and Langevin Dynamics is an orthogonal contribution to ours, two ways of
+attempting to solve the same problem of MCMC methods with big datasets. We could augment their
+methods with our MH test, to get the benefit of their high quality proposals. Note that while their
+methods in theory do not need an MH test, in practice, they need one due to discretization of the
+system. Furthermore, tests may also be needed for when the algorithms are run with a constant step
+size.
+
+%% Daniel: I don't know how to integrate this easily.
+%% \cite{sgmcmc_2015} (NIPS 2015) might want to give this a mention, where it would fit in their
+%% recipe.
+
+\section{A New Metropolis-Hastings Test}\label{sec:our_algorithm}
 
@@
   define $X$. The details are in Section~\ref{sec:our_algorithm}.
+  {\color{blue}
+  Daniel: I think this needs to be merged into Figure~\ref{fig:part1}.
+  }
   }
@@
 
-\section{Implementation}\label{sec:implementation}
+\section{Theoretical Results}\label{sec:theory}
+
+TODO insert Haoyu's proof.
 

### 2016-05-15 nips_2016.tex
Copied Haoyus proof in the paper draft. Now I will revise Sections 3 and 4.

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
 
@@
 
-Rather than fixing the posterior temperature and trying to use adjustable mini-batch size to achieve
+Rather than fixing the posterior temperature and trying to use adjustable minibatch size to achieve
 it (which ends up reducing to nearly-full-batch updates) we propose to run at the natural
@@
 
+{\color{blue}
+Daniel: This section needs work in describing our argument carefully.
+}
+
 \section{Preliminaries and Related Work}\label{sec:related_work}
 
-The standard MCMC method proceeds as follows~\cite{gilks1996markov,brooks2011handbook}. For a
-parameter $\theta$ and conditionally independent data, we wish to compute the target distribution
-$p(\theta \mid x_1, \ldots, x_N)$. Since exact evaluation is intractable, we generate a chain of
-correlated samples $\theta_1, \ldots, \theta_T$ for some large $T$, and approximate $p$ by using
-counts of samples. For each iteration $t$, we start with our current $\theta_t$. We use a
-\emph{proposal distribution} $q(\theta' \mid \theta_t)$ to determine a new candidate $\theta'$. With
-probability $P_a$, we accept it and set $\theta_{t+1} = \theta'$; otherwise, we repeat the previous
-value $\theta_{t+1} = \theta_t$. Traditionally, $P_a$ is computed as follows:
+In the standard MCMC method~\cite{gilks1996markov,brooks2011handbook} for Bayesian inference with
+parameter $\theta$ and conditionally independent data, the goal is to compute the target
+distribution $p(\theta \mid x_1, \ldots, x_N)$.  To do so, one generates a chain of (correlated)
+samples $\theta_1, \ldots, \theta_T$ for large $T$ and approximates $p$ using the sample counts.
+Each iteration $t$ has a current $\theta_t$, and a \emph{proposal distribution} $q(\theta' \mid
+\theta_t)$ determines a new candidate $\theta'$. With probability $P_a$, the sample is accepted, so
+$\theta_{t+1} = \theta'$. Otherwise, $\theta_{t+1} = \theta_t$. This can be implemented by drawing
+a uniform random variable $u \sim {\rm Unif}[0,1]$ and accepting if $u < P_a$. Traditionally, $P_a$ is
 \begin{equation}\label{eq:traditional}
@@
 \end{equation}
-where $f(\theta_t)=p(\theta_t|x_1,\ldots,x_N) \propto p(\theta_t)\prod_{i=1}^N p(x_i \mid
-\theta_t)$. The technical requirement for $P_a$ is that $f$ be proportional to the true target
-distribution.
-
-One then draws a uniform random variable $u$ from ${\rm Unif}[0,1]$ and accepts $\theta_{t+1} =
-\theta'$ if $u < P_a$, and uses $\theta_{t+1} = \theta_t$ otherwise. The $P_a$ from
-Equation~\ref{eq:traditional} satisfies detailed balance, so if one samples long enough using the
-heuristic above, one will arrive at a stationary distribution.  The $P_a$ is guaranteed to converge
-to the true $p$ distribution given sufficiently many examples (though a burn-in period and/or taking
-every $n$th sample may be necessary).
+where $f(\theta_t)=p(\theta_t \mid x_1,\ldots,x_N)$. The $P_a$ from Equation~\ref{eq:traditional}
+satisfies detailed balance, so if one samples long enough, one will arrive at a stationary
+distribution matching the posterior, though a burn-in period and/or taking every $n$th sample may be
+helpful.
 
@@
 researchers have tried to do about that.  The last paragraph is from~\cite{cutting_mh_2014}, and it
-seems to argue that people have not used minibatch MCMC before ``the algorithm gets stuck''. But
-isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes? If we wanted
-to get stuck at modes with large datasets, why didn't we just do what previous work did?
+seems to argue that people have not used minibatch MCMC before because ``the algorithm gets stuck''.
+But isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes? If we
+wanted to get stuck at modes with large datasets, why didn't we just do what previous work did?
 }
@@
 algorithm is the need to keep incrementing the mini-batch size in one iteration; in the worst case,
-all $N$ data points may be needed in a single iteration. A related approach
-is~\cite{icml2014c1_bardenet14}, which has slightly more robust theoretical properties, but comes at
-the cost of more computational time. The work of~\cite{conf/uai/MaclaurinA14} presents an
-\emph{exact} approach, but has the sometimes unreasonable requirement of a lower bound on the
-per-datum likelihood factors.
+all $N$ data points may be needed in a single iteration. A similar approach
+from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties, but comes at the
+cost of more computational time. The work of~\cite{conf/uai/MaclaurinA14} presents another minibatch
+MCMC method, but which gets the \emph{exact} posterior, with the only requirement being that there
+needs to be a lower bound on the log likelihood for each data point. They augment the likelihood
+with auxiliary random variables, but this can hurt mixing.
 
@@
 %% recipe.
+
+%% Daniel: for a further discussion on the 'Firefly' paper, see this blog post:
+%% https://xianblog.wordpress.com/2014/04/02/firefly-monte-carlo/
+%% Note: this guy also blogged about our earlier ICLR 2015 paper.
 
@@
 
-TODO insert Haoyu's proof.
+Since our proposed accept probability function $g(\Delta) = (1+ \exp(-\Delta))^{-1}$ satisfies
+detail balance, if $\Delta$ is computed from the complete data, the distribution of the sampled data
+will be the same as our target distribution $f(.)$. Define the accept probability computed by the
+complete data as $P_a = g(\Delta) = \Pr(\Delta + X > 0)$. However, if our proposed mini-batch
+testing method, we use $\Delta'$, which comes from mini-batch data, to estimate the accept
+probability. Define the accept probability we get by $\Delta'$ as $P_a' = \Pr(\Delta' +
+X_{correction} > 0)$. In this section, we first prove that if the our estimation of $X_{norm}$ is
+correct, our accept probability $P_a'$ is the same as the $P_a$. Secondly, we define the transition
+kernel of complete data at sampling step $i$ is $T_i$, the transition kernel of mini-batch data at
+sampling step $i$ is $T_i'$. We show the distance of $T_i$ and $T_i'$ is proportional to our
+estimation of $X_{norm}$. Thirdly, we define the stationary distribution obtained by $T_i$ is $\pi$.
+We show the by decreasing the step size, the distribution of our sampled data converge to the $\pi$.
 
-We implement our system within the open-source BIDMach project~\cite{canny2013bidmach}. Note to
-ourselves: do NOT modify the \texttt{Learner.scala} code -- just make it a new ``Model''.
+\begin{theorem}
+If our estimation of $X_{norm}$ is ideally correct, the accept probability between our mini-batch
+algorithm and the complete data are the same, i.e. $P_a = P_a'$.
+\end{theorem}
+
+\begin{proof}
+\[
+P_a = \Pr (\Delta + X > 0) = \Pr (\Delta + X_{norm} + X_{correction} > 0) = \Pr (\Delta' +
+X_{correction} > 0) = P_a'.
+\]
+\end{proof}
+
+\begin{lemma}
+The distribution of $\Delta'$ is approximately Gaussian, if we choose the mini-batch data randomly. 
+\end{lemma}
+
+\begin{proof}
+By definition, $\Delta' = \log \left( \frac{f(\theta')P(\theta_t | \theta')}{f(\theta_t) P(\theta' |
+\theta_t)} \right)$, where $f(\theta)$ is the posterior distribution, i.e. $f(\theta) =
+p_{\theta}(\theta) \prod_i p(x_i | \theta)$. By plugging the $f(\theta)$, we have $\Delta' = \log
+p_{\theta}(\theta') - \log p_{\theta}(\theta_t) + \log P(\theta_t|\theta') - \log
+P(\theta'|\theta_t) + \sum_i (\log p(x_i|\theta') - \log p(x_i|\theta_t))$. If we fix the $\theta_t$
+and $\theta'$, the randomness only comes from the choice of mini-batch data, i.e. $\sum_i (\log
+p(x_i|\theta') - \log p(x_i|\theta_t))$. Since we choose the mini-batch data randomly, the
+distribution of $\sum_i (\log p(x_i|\theta') - \log p(x_i|\theta_t))$ converge to Gaussian
+distribution by the Central Limit Theorem.
+\end{proof}
+
+Let's define the our estimation of $X_{norm}$ is $X_{norm}'$, and $X+{\xi} = X_{norm} - X_{norm}'$.
+We also define the cumulative distribution function of $X$ is $F_X(.)$.
+
+\begin{lemma}
+If $|X_\xi| < \zeta$ and $\nabla F_X < k$, the total variance distance between $T_i$ and $T_i'$ is
+bounded by $2\zeta k$.
+\end{lemma}
+
+\begin{proof}
+We firsts prove the difference of accept rate is bounded: $|P_a - P_a'| \leq 2\zeta k$ where, $P_a =
+\Pr (\Delta + X > 0)$, $P_a' = \Pr(\Delta' + X_{corr} > 0)$, $X= X_{corr} + X_{norm}$, and $\Delta'
+= \Delta + X_{norm} + X_\xi$. Given $X_\xi = \xi$ the $|P_a - P_a'|$ is:
+\begin{align*}
+|P_a - P_a'| &= |\Pr(\Delta + X >0) - \Pr(\Delta' + X_{corr} > 0)|\\
+& = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{norm} + \xi + X_{corr} >0) |    \\
+&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| = |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
+&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2k \xi.
+\end{align*}
+Since $|X_\xi| < \zeta$, $|P_a - P_a'| < 2k\zeta$.
+
+Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
+$||T_i - T'_i||\leq 2k\zeta$. We can write the $T_i(\theta_t|\theta') = P_a(\theta_t, \theta')
+q(\theta'|\theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$ is
+the Dirac delta function. We can write the distance between two proposal kernel is:
+
+\begin{align*}
+||T_i - T'_i || &= \int_{\theta'} d\theta' |\int_{\theta_t} d\theta_t ((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) | \\
+& = \int_{\theta'} d\theta' |\int _{\theta_t} d\theta_t (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) | \\ 
+& \leq 2 \zeta k \int_{\theta'}d\theta' | \int_{\theta_t} d\theta_t (q(\theta'|\theta_t) + q(\theta_t|\theta'))| = 2 \zeta k.
+\end{align*}
+\end{proof}
+
+Since the accuracy of our transition kernel is related to how accurate our estimation of $X_{\rm
+norm}$, we show the error of our estimation is roughly proportional to the step size of our
+proposer.
+
+\begin{lemma}
+For one step sampling from $\theta_t$ to $\theta'$, if the jumping step  $|\theta_t - \theta'|$ is
+smaller than $\epsilon$, and the gradient of the log-likelihood is bounded by a constant factor,
+i.e. $|\nabla (\log \Pr(x_i| \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2
+k^2 m$, where $m is the mini-batch size$.
+\end{lemma}
+
+\begin{proof}
+Similarly as the proof in the Lemma 2, We write the $\Delta'$ as: $\Delta' = \sum_i (\log
+p(x_i|\theta') - \log p(x_i|\theta_t)) + \log P(\theta_t | \theta') - P(\theta'|\theta_t) + \log
+p_{\theta}(\theta') - \log p_{\theta}(\theta_t)$. Since only the $\sum_i (\log p(x_i|\theta') - \log
+p(x_i|\theta_t))$ brings randomness, we define function $g_i(\theta)$ as $\log p(x_i|\theta)$. Then
+we have:
+\begin{align*} 
+|\log \Pr(x_i | \theta') - \log \Pr (x_i |\theta_t)| & = |g(\theta') - g(\theta_t)| \\
+&= |\nabla g(\theta_t) (\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
+\end{align*}
+Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
+\begin{align*}
+var(\Delta') &= var(\sum_i (\log \Pr(x_i | \theta') - \log \Pr (x_i |\theta_t))) = var(\sum_i g_i(\theta') - g_i(\theta_t)) \\
+& \leq \sum_i var(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
+\end{align*}
+where $m$ is the minibatch size.
+\end{proof}
+
+Since $var(\Delta') < \epsilon^2 k^2 m$, we can assume the maximum value of $X_{\xi}$ is
+proportional to $sd(\Delta)$, i.e. $\epsilon$. Thus, we can assume $\zeta=\epsilon C$, where $C$ is
+a constant factor.
+
+\begin{theorem}
+Assume the stationary distribution $\pi$ satisfies $||T \circ P_0 - \pi|| \leq \eta ||P_0 - \pi||$,
+where $\eta \in [0, 1)$ for all distribution $P_0$. And assume there exists $\rho_t < 1$ such that
+$||T'_i(x, .) - \pi_i'|| \leq 2\rho_i$. We can get $|| T_t' \circ T_{t-1}' \circ \cdots \circ T_1' P_0
+- \pi_0 || \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $,
+where $\alpha_i = \frac{\epsilon_i k }{1-\eta}$.
+\end{theorem}
+
+\begin{proof}
+Firstly, "Austerity in MCMC Land" shows that with $||T \circ P_0 - \pi|| \leq \eta ||P_0 - \pi||$
+and $||T-T'_i|| \leq \epsilon_i k$, the norm between the stationary distribution is $||\pi -
+\pi_i'||\leq \frac{\epsilon_i k}{1-\eta}$.
+
+Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| T_t' \circ T_{t-1}' \circ \cdots
+\circ T_1' P_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
+\alpha_s$, where $\alpha = \frac{\epsilon_i k}{1-\eta}$. Thus, we get:
+\[
+ \| T_t' \circ T_{t-1}' \circ \cdots \circ T_1' P_0 - \pi_0 \| \leq \|T_t' \circ T_{t-1}' \circ
+ \cdots \circ T_1' P_0 - \pi_t\| + \|\pi_t - \pi_0\| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u
+ (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t.
+\]
+\end{proof}
 
@@
 
+We implement our system within the open-source BIDMach project~\cite{canny2013bidmach}. Note to
+ourselves: do NOT modify the \texttt{Learner.scala} code -- just make it a new ``Model''.
+
 \section{Conclusions}\label{sec:conclusion}
@@
 }
+
+Straightforward directions for future work include more extensive experiments. More elaborate
+extensions could be combining our results with Hamiltonian Monte Carlo methods, or integrating
+parallel MCMC~\cite{conf/uai/AngelinoKWSA14} concepts into our work.
 

### 2016-05-15 nips_2016.tex
major update to sections 1 through 4. Obviously not done but lots of progress in a draft form.

--- 2016-05-15 nips_2016.tex
@@
 \usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem}
+\usepackage[]{algorithm2e}
 \graphicspath{{figures/}}
@@
 non-unit powers of the distribution are fine. That corresponds to changing the temperature of the
-distribution to make it ``flatter''.
-
-Rather than fixing the posterior temperature and trying to use adjustable mini-batch size to achieve
-it (which ends up reducing to nearly-full-batch updates) we propose to run at the natural
-temperature induced by a given minibatch size. We can always increase the minibatch size to reduce
-the temperature, but there is no need to do this beyond the point where the posterior mode is sharp
-and well-defined. That will occur at some fixed number of points and far less than the full dataset.
+distribution to make it ``flatter''.  Rather than fixing the posterior temperature and trying to use
+adjustable minibatch size to achieve it (which ends up reducing to nearly-full-batch updates) we
+propose to run at the natural temperature induced by a given minibatch size. We can always increase
+the minibatch size to reduce the temperature, but there is no need to do this beyond the point where
+the posterior mode is sharp and well-defined. That will occur at some fixed number of points and far
+less than the full dataset.
 
@@
 MCMC methods. We present theoretical and experimental results on our method and show that it can be
-the method of choice for minibatch MCMC .
+the method of choice for minibatch MCMC. The contributions of this paper are as follows:
+
+\begin{enumerate}[noitemsep]
+    \item We develop a new minibatch MH test which satisfies detailed balance.
+    \item Our MH test is more efficient than the adaptive sampling method.
+    \item The test can incorporate temperature considerations to control mixing rates.
+\end{enumerate}
+
+{\color{blue}
+Daniel: This section needs work in describing our argument carefully. I will decrease the focus on
+temperature and increase focus on the new MH test once the results are clearer.
+}
 
@@
 
-The standard MCMC method proceeds as follows~\cite{gilks1996markov,brooks2011handbook}. For a
-parameter $\theta$ and conditionally independent data, we wish to compute the target distribution
-$p(\theta \mid x_1, \ldots, x_N)$. Since exact evaluation is intractable, we generate a chain of
-correlated samples $\theta_1, \ldots, \theta_T$ for some large $T$, and approximate $p$ by using
-counts of samples. For each iteration $t$, we start with our current $\theta_t$. We use a
-\emph{proposal distribution} $q(\theta' \mid \theta_t)$ to determine a new candidate $\theta'$. With
-probability $P_a$, we accept it and set $\theta_{t+1} = \theta'$; otherwise, we repeat the previous
-value $\theta_{t+1} = \theta_t$. Traditionally, $P_a$ is computed as follows:
+In the standard MCMC method~\cite{gilks1996markov,brooks2011handbook} for Bayesian inference with
+parameter $\theta$ and conditionally independent data, the goal is to compute the target
+distribution $p(\theta \mid x_1, \ldots, x_N)$.  To do so, one generates a chain of (correlated)
+samples $\theta_1, \ldots, \theta_T$ for large $T$ and approximates $p$ using the sample counts.
+Each iteration $t$ has a current $\theta_t$, and a \emph{proposal distribution} $q(\theta' \mid
+\theta_t)$ determines a new candidate $\theta'$. With probability $P_a$, the sample is accepted, so
+$\theta_{t+1} = \theta'$. Otherwise, $\theta_{t+1} = \theta_t$. This can be implemented by drawing
+a uniform random variable $u \sim {\rm Unif}[0,1]$ and accepting if $u < P_a$. Traditionally, $P_a$ is
 \begin{equation}\label{eq:traditional}
 P_a = \min\left\{ 1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid \theta_t)}
-\right\} = \min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i ; \theta')q(\theta_t \mid \theta')}{p(\theta_t)\prod_{i=1}^N p(x_i ; \theta_t)q(\theta' \mid
+\right\} = \min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
+\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid
 \theta_t)} \right\},
 \end{equation}
-where $f(\theta_t)=p(\theta_t|x_1,\ldots,x_N) \propto p(\theta_t)\prod_{i=1}^N p(x_i \mid
-\theta_t)$. The technical requirement for $P_a$ is that $f$ be proportional to the true target
-distribution.
-
-One then draws a uniform random variable $u$ from ${\rm Unif}[0,1]$ and accepts $\theta_{t+1} =
-\theta'$ if $u < P_a$, and uses $\theta_{t+1} = \theta_t$ otherwise. The $P_a$ from
-Equation~\ref{eq:traditional} satisfies detailed balance, so if one samples long enough using the
-heuristic above, one will arrive at a stationary distribution.  The $P_a$ is guaranteed to converge
-to the true $p$ distribution given sufficiently many examples (though a burn-in period and/or taking
-every $n$th sample may be necessary).
+where $f(\theta_t)=p(\theta_t \mid x_1,\ldots,x_N)$. The $P_a$ from Equation~\ref{eq:traditional}
+satisfies detailed balance, so if one samples long enough, one will arrive at a stationary
+distribution matching the posterior, though a burn-in period and/or taking every $n$th sample may be
+helpful.
 
@@
 researchers have tried to do about that.  The last paragraph is from~\cite{cutting_mh_2014}, and it
-seems to argue that people have not used minibatch MCMC before ``the algorithm gets stuck''. But
-isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes? If we wanted
-to get stuck at modes with large datasets, why didn't we just do what previous work did?
+seems to argue that people have not used minibatch MCMC before because ``the algorithm gets stuck''.
+But \emph{isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes}? If
+we wanted to get stuck at modes with large datasets, why didn't we just do what previous work did?
+The only way around this is if our algorithm is getting stuck at ``better'' modes, but then we need
+to design our experiments to test this.
 }
@@
 
-Our starting point is the following Lemma.
+In this section, we introduce our new Metropolis-Hastings test and show how to use it to achieve
+sufficient mixing and/or to reach better posterior modes. To start, we introduce two key values we
+will use frequently: $\Delta$ and $\Delta'$:
+\begin{equation}\label{eq:deltas}
+\Delta = \log \left( \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
+\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid
+\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta')\prod_{i=1}^n p(x_i \mid \theta')q(\theta_t \mid
+\theta')}{p(\theta_t)\prod_{i=1}^n p(x_i \mid \theta_t)q(\theta' \mid
+\theta_t)} \right)
+\end{equation}
+where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}, so the terms here
+consider the likelihood ratios in log space. The only difference between $\Delta$ and $\Delta'$ is
+that they use $N$ and $n$ terms in the likelihood ratio, respectively. We assume $n \ll N$ (so
+$\Delta'$ is substantially faster to compute) and that the $n$ data points $x_i$ represent a random
+minibatch of data.
+
+To motivate the use of $\Delta$, we turn to Lemma~\ref{lem:detailed_balance}.
 
 \begin{lemma}\label{lem:detailed_balance}
-Let $\Delta = \log \left(\frac{f(\theta') q(\theta_t \mid \theta')}{f(\theta_t) q(\theta'\mid
-\theta_t)} \right)$, where $f$ is proportional to the desired target distribution and $q$ is our
-chosen proposal distribution. Any acceptance function $g$ such that $g(\Delta) = \exp(\Delta)
-g(-\Delta )$ satisfies detailed balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) =
-f(\theta')p(\theta_t \mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the probability of jumping
-from $\theta_x$ to $\theta_y$ in our chain.
+Let $\Delta$ be defined as in Equation~\ref{eq:deltas}. Any acceptance function $g$ such that
+$g(\Delta) = \exp(\Delta) g(-\Delta )$ satisfies detailed balance. That is, $f(\theta_t)p(\theta'
+\mid \theta_t) = f(\theta')p(\theta_t \mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the
+probability of jumping from $\theta_x$ to $\theta_y$ in our chain.
 \end{lemma}
@@
 We begin by deriving $p(\theta' \mid \theta_t)$. This is equivalent to the probability of proposing
-$\theta'$ and then accepting it, so
-\begin{equation}\label{eq:one_way}
-p(\theta' \mid \theta_t) = q(\theta' \mid \theta_t)g(\Delta).
-\end{equation}
-Similarly, we have
-\begin{equation}\label{eq:other_way}
-p(\theta_t \mid \theta') = q(\theta_t \mid \theta')g(-\Delta).
-\end{equation}
-Notice that the probability of accepting a transition from $\theta'$ to $\theta_t$ is $g(-\Delta)$
-because this inverts the fraction inside the logarithm term of $\Delta$.  By assumption, we can
-expand $g(\Delta) = \exp(\Delta)g(-\Delta)$ in Equation~\ref{eq:one_way}.  Doing this, and combining
-the result of Equation~\ref{eq:other_way}, we get
+$\theta'$ and then accepting it, so $p(\theta' \mid \theta_t) = q(\theta' \mid \theta_t)g(\Delta).$
+Similarly, $p(\theta_t \mid \theta') = q(\theta_t \mid \theta')g(-\Delta).$ Notice that the
+probability of accepting a transition from $\theta'$ to $\theta_t$ is $g(-\Delta)$ because this
+inverts the fraction inside the logarithm term of $\Delta$.  By assumption, we can expand $g(\Delta)
+= \exp(\Delta)g(-\Delta)$ in $p(\theta' \mid \theta_t)$. Doing this, and combining that with the
+definition of $p(\theta_t \mid \theta')$, we get
 \begin{equation}\label{eq:combined}
@@
 \end{equation}
-Cancellations result in $f(\theta') p(\theta_t \mid \theta') = f(\theta_t) p(\theta' \mid \theta_t)$. Thus, detailed balance is satisfied.\\
+Cancellations result in $f(\theta') p(\theta_t \mid \theta') = f(\theta_t) p(\theta' \mid
+\theta_t)$. Thus, detailed balance is satisfied.
 \end{proof}
@@
 Some straightforward arithmetic shows that the standard Metropolis-Hastings acceptance function
-$g(\Delta) = \min\{1, e^\Delta \} = \min\left\{1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid
-\theta_t)}\right\}$ satisfies the condition $g(\Delta) =
+$g(\Delta) = \min\{1, e^\Delta \} = \min\left\{1, \frac{f(\theta')q(\theta_t \mid
+\theta')}{f(\theta_t)q(\theta' \mid \theta_t)}\right\}$ satisfies the condition $g(\Delta) =
 \exp(\Delta)g(-\Delta)$ and consequently, results in detailed balance.
 
-For reasons that will be clear later, we choose an alternative $g$, the logistic function:
-$g(\Delta) = (1+\exp(-\Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
-condition in Lemma~\ref{lem:detailed_balance}.
-
-We can sample from the logistic function using the following procedure. Let $u$ be drawn from ${\rm
-Unif}[0,1]$. At any given iteration, we can compute $\Delta$, and we accept the new candidate
-$\theta'$ if $g(\Delta) > u$, and reject otherwise. Let us define the random variable $X =
-g^{-1}(u)$ where again, $g$ is the logistic function. We claim that $X$ has CDF function $F_X(x) =
-g(x)$, i.e., its CDF is precisely the logistic function. This is because for some $X = x$, we have
+For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
+(1+\exp(-\Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the condition in
+Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily sample from
+it using the following procedure. Let $u$ be a uniform random variable $u \sim {\rm Unif}[0,1]$. At
+any iteration with a current parameter $\theta_t$ and a proposed sample $\theta'$, we can compute
+$\Delta$. We accept the candidate if $g(\Delta) > u$, and reject otherwise. To see why this works,
+define a new random variable $X = g^{-1}(u)$ where again, $g$ is the logistic function. Notice that
+$X$ has cumulative distribution function $F_X(x) = g(x)$. In other words, the CDF is precisely the
+logistic function. This is because for some $X = x$, we have
 \[
@@
 as the density of the uniform random variable here is simply one. Thus, the criteria to accept the
-candidate $\theta'$ is equivalent to whether $\Delta > X$, where $X$ has CDF of logistic function.
-(If this isn't immediately obvious, note that we can pre-multiply $g^{-1}$ to $g(\Delta) \ge u$, and
-our test is $\Delta > X$.) Since $X$ is symmetric about zero, the acceptance criteria can also be
-expressed as $\Delta + X>0$, as shown in Figure~\ref{fig:part1}.
+candidate $\theta'$ is equivalent to whether $\Delta > X$. Since $X$ is symmetric about zero, the
+acceptance criteria is equivalently expressed as $\Delta + X>0$, as shown in Figure~\ref{fig:part1}.
 
 {\color{blue}
-Daniel: The above makes sense, but I still don't understand why we had to show that the CDF of $X$
-is the logistic function, because I never use that when I claim that our acceptance test is now
-$\Delta > X$.  Do we need that assumption so that, for instance, we can decompose $X$ later the way
-we do it, with $X_{\rm noise}$ having the same variance as $\epsilon$? That might explain it.
+Daniel: The above makes sense, but I \emph{still} don't understand why we had to show that the CDF
+of $X$ is the logistic function, because I never use that when I claim that our acceptance test is
+now $\Delta > X$. All I need to say is this: we can pre-multiply $g^{-1}$ to $g(\Delta) \ge u$, and
+our test is $\Delta > X$. Right???
 }
 
-Unfortunately, there's a problem with the $\Delta + X > 0$ test: it requires us to compute $\Delta$,
-so our test will be slow! We can get around this by defining a new scalar-valued term, $\Delta'$, which is
-computed the same way as $\Delta$, but uses far fewer samples. In other words, $\Delta'$ is computed
-based on a \emph{mini-batch} of data, and it is an approximation of $\Delta$, so we can express the
-relationship as $\Delta' = \Delta + \epsilon$ for a noise term $\epsilon$. \emph{The key is that
-$\epsilon$ follows a normal distribution}, $\epsilon \sim \mathcal{N}(0, \sigma^2)$, because for i.i.d. data $x_1, \ldots, x_N$, the $\Delta$
-is expressed as the sum of log probabilities (look at Equation~\ref{eq:traditional} and apply the
-logarithm to the product over the $N$ data points, to get a sum of log terms). All we do with
-$\Delta$ is take far fewer of the summations, and normally-distributed noise is there as a product
-of the Central Limit Theorem.
+With large datasets, $\Delta$ is intractable to compute. An intuitive alternative is to use
+$\Delta'$ instead. The following lemma characterizes the distribution of $\Delta'$.
 
-To sample accurately from $\Delta'$, we need to slightly change our acceptance criteria. We can
-decompose $X$ as $X = X_{\rm norm} +X_{\rm correction}$, where the first term is a ``noise term''
-which has the same variance as $\epsilon$, and the second term is a ``residual'' term. The criteria
-to accept, previously expressed as $\Delta + X > 0$, can be rewritten as
+\begin{lemma}\label{lem:gaussian}
+If the minibatch data is chosen randomly, then the distribution of $\Delta'$ is Gaussian.
+\end{lemma}
+
+\begin{proof}
+By expanding the definition of $\Delta'$ from Equation~\ref{eq:deltas}, we get
+\[
+\Delta' = \log p(\theta') - \log p(\theta_t) + \log q(\theta_t \mid \theta') - \log q(\theta' \mid \theta_t) +
+\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t)).
+\]
+If we fix $\theta_t$ and $\theta'$, the randomness in our $\Delta'$ samples only comes from the
+minibatch. Since the minibatch is chosen randomly, the distribution of $\sum_{i=1}^n (\log p(x_i\mid
+\theta') - \log p(x_i\mid \theta_t))$ converge to a Gaussian by the Central Limit Theorem.
+\end{proof}
+
+From Lemma~\ref{lem:gaussian}, since $\Delta'$ is a noisy approximation of $\Delta$, the
+relationship can be expressed as
+\begin{equation}\label{eq:relationship}
+\Delta' = \Delta + \epsilon, \quad \epsilon \sim \mathcal{N}(0, \sigma^2),
+\end{equation}
+for a Gaussian noise term $\epsilon$.  The extra noise in $\Delta'$ means we can no longer perform
+the $\Delta + X$ test, so we need to slightly change our acceptance criteria. We can decompose
+$X$ as $X = X_{\rm norm}+X_{\rm corr}$, where the first term is a zero-mean ``noise term'' which
+has the same variance as $\epsilon$, and the second term is a zero-mean ``correction'' term.
+These two convolve to form $X$. The criteria to accept can be rewritten as
 \begin{equation}\label{eq:criteria}
-\Delta + X = \Delta + X_{\rm norm} + X_{\rm correction} \approx \Delta' + X_{\rm correction} >0.
+\Delta + X = \Delta + X_{\rm norm} + X_{\rm corr} \approx \Delta' + X_{\rm corr} >0.
 \end{equation}
-This explains Figure~\ref{fig:part2} (though it really should be an approximation as we can't guarantee that our $X_{\rm norm}$ term is exactly the correct amount of noise), where the left hand side has distributions representing normal
+This deconvolution process is visualized in Figure~\ref{fig:part2}. {\color{blue} (Daniel: note that
+we will be fixing this figure, so the visualization should hopefully be a little clearer.)} We
+decompose $X$ this way so that we can estimate the two components separately.  Notice that if we
+increase the mini-batch size, this will cause the $X_{\rm noise}$ distribution to shrink and become
+narrower.
+
+Furthermore, since our proposed acceptance function $g(\Delta) = (1+ \exp(-\Delta))^{-1}$ satisfies
+detailed balance (see Lemma~\ref{lem:detailed_balance}), if $\Delta$ is computed from the complete
+data, the distribution of the sampled data will match our target distribution $f$, the posterior,
+under standard MCMC assumptions (e.g., that the chain is run sufficiently long enough and mixes
+well). Now consider the case of using minibatches of data. Let our true and estimated acceptance
+probabilities be $P_a = g(\Delta) = \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} >
+0)$.  If our estimate of $X_{\rm norm}$ is exact, then $P_a' = P_a$, which follows by making the
+inequality in Equation~\ref{eq:criteria} an equality.
+
+Algorithm~\ref{alg:our_algorithm} concisely describes our MH test.
+
+{\color{blue}
+Daniel: I really need to make this clear and do the algorithm description.
+}
+
+\begin{comment}
+though it really should be an approximation as we can't guarantee that our $X_{\rm norm}$ term is
+exactly the correct amount of noise), where the left hand side has distributions representing normal
 noise ($X_{\rm norm}$, which is labeled as ${\rm Noise}_U$, but they are the same thing) and a
@@
 
-Notice that we have several options we can control. For instance, we can adjust the mini-batch size;
-increasing the mini-batch size will cause the $X_{\rm noise}$ distribution to shrink and become
-narrower.
-
-{\color{blue} Daniel: I think it's becoming clear but I'm still trying to connect all of the steps.
-During each iteration of MCMC, we compute $\Delta'$ and \emph{sample} (right?) an $X$ (which we can
-do by sampling $u$ and then doing $X = g^{-1}(u)$. This gives us $\Delta' + X$ but our MH test
-    actually requires $X_{\rm correction}$, so first we must estimate $\sigma^2$, the variance of
-    $X_{\rm noise} \sim \mathcal{N}(0, \sigma^2)$ (perhaps from the mini-batch data, somehow?) and
-    then somehow ``remove'' that component from $X$ to get $X_{\rm correction}$? Then we have our
-    acceptance test. Is this right? Or, of course, we don't have to sample $X$ at all, but just
-    figure out a way to sample $X_{\rm correction}$ directly.
+I think it's becoming clear but I'm still trying to connect all of the steps.  During each iteration
+of MCMC, we compute $\Delta'$ and \emph{sample} (right?) an $X$ (which we can do by sampling $u$ and
+then doing $X = g^{-1}(u)$. This gives us $\Delta' + X$ but our MH test actually requires $X_{\rm
+correction}$, so first we must estimate $\sigma^2$, the variance of $X_{\rm noise} \sim
+\mathcal{N}(0, \sigma^2)$ (perhaps from the mini-batch data, somehow?) and then somehow ``remove''
+that component from $X$ to get $X_{\rm correction}$? Then we have our acceptance test. Is this
+right? Or, of course, we don't have to sample $X$ at all, but just figure out a way to sample
+$X_{\rm correction}$ directly.
 
@@
 $\Delta_1, \ldots, \Delta_k$ and compute the empirical $\sigma^2$ (of course that defeats the
-purpose of using mini-batches).  }
+purpose of using mini-batches).
 
-{\color{blue} Xinlei: I agree with you that out MH test requires $X_{correction}$. My understanding
+I agree with you that out MH test requires $X_{correction}$. My understanding
 is that we can actually estimate $X$ from the whole data set, and the probability distribution of
 $X_{norm}$ is what we proposed? So the determination of $X_{correction}$ is a deterministic
-deconvolution problem.  }
+deconvolution problem.
+\end{comment}
+
+\begin{algorithm}[t]
+\KwData{this text}
+\KwResult{how to write algorithm with \LaTeX2e } initialization\;
+\While{not at end of this document}{
+    read current\;
+    \eIf{understand}{
+        go to next section\;
+        current section becomes this one\;
+    }{
+        go back to the beginning of current section\;
+    }
+}
+\caption{A description of our MH test.}
+\label{alg:our_algorithm}
+\end{algorithm}
 
@@
 
-TODO insert Haoyu's proof.
+In this section, we explore how well our algorithm will converge to the posterior distribution.
+Define our estimate of $X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm norm} - X_{\rm
+norm}'$, so that $\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as the CDF of $X$.
+Let the transition kernel of complete data at sampling step $i$ be $T_i$, and let the transition
+kernel of minibatch data at sampling step $i$ is $T_i'$.  We show that the distance of $T_i$ and
+$T_i'$ is proportional to our estimate of $X_{\rm norm}$. We also define the stationary distribution
+obtained by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution of our
+sampled data converges to $\pi$.
 
-We implement our system within the open-source BIDMach project~\cite{canny2013bidmach}. Note to
-ourselves: do NOT modify the \texttt{Learner.scala} code -- just make it a new ``Model''.
+\begin{lemma}
+If $|X_\xi| < \zeta$ and $\nabla F_X < k$, the total variance distance between $T_i$ and $T_i'$ is
+bounded by $2\zeta k$.
+\end{lemma}
+
+\begin{proof}
+We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta k$
+for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
+\begin{align*}
+|P_a - P_a'| &= |\Pr(\Delta + X >0) - \Pr(\Delta' + X_{\rm corr} > 0)|\\
+& = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
+&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| = |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
+&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2k \xi,
+\end{align*}
+where the last line uses Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have
+smaller absolute value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| <
+2k\zeta$.
+
+Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
+$\|T_i - T'_i\|_1 \leq 2k\zeta$ with $\| \cdot \|_1$ the standard $L_1$ norm. We can write the
+$T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta') q(\theta'\mid \theta_t) +
+(1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$ is the Dirac delta
+function. The distance between two proposal kernels is
+\begin{align*}
+\|T_i - T'_i \|_1 &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
+& = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
+& \leq 2 \zeta k \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta'))  d\theta_t \right| d\theta'= 2 \zeta k,
+\end{align*}
+as desired.
+\end{proof}
+
+Since the accuracy of our transition kernel is related to the accuracy of our $X_{\rm norm}$
+estimate, we can show that error of our estimation is roughly proportional to the step size of our
+proposer.
+
+\begin{lemma}
+For one step sampling from $\theta_t$ to $\theta'$, if the jumping step  $|\theta_t - \theta'| <
+\epsilon$, and the gradient of the log-likelihood is bounded by a constant factor, i.e. $|\nabla
+(\log \Pr(x_i| \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 m$, where
+$m$ is the minibatch size.
+\end{lemma}
+
+\begin{proof}
+From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
+\theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
+$g_i(\theta) = \log p(x_i\mid \theta)$. Then by Taylor's Theorem:
+\[
+|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t) (\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
+\]
+Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
+\begin{align*}
+{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^n (\log \Pr(x_i | \theta') - \log \Pr (x_i |\theta_t))\right] = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \\
+& \leq \sum_{i=1}^n {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
+\end{align*}
+as desired.
+\end{proof}
+
+Since ${\rm Var}(\Delta') < \epsilon^2 k^2 m$, we can assume the maximum value of $X_{\xi}$ is
+proportional to the standard deviation of $\Delta$, i.e. $\epsilon$. Thus, we can assume
+$\zeta=\epsilon C$, where $C$ is a constant factor.
+
+\begin{theorem}
+Assume the stationary distribution $\pi$ satisfies $\|T \circ P_0 - \pi\| \leq \eta \|P_0 - \pi\|$,
+where $\eta \in [0, 1)$ for all distribution $P_0$. And assume there exists $\rho_t < 1$ such that
+$\|T'_i(x, \cdot) - \pi_i'\| \leq 2\rho_i$. We can get $\| T_t' \circ T_{t-1}' \circ \cdots \circ T_1' P_0
+- \pi_0 \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $,
+where $\alpha_i = \frac{\epsilon_i k }{1-\eta}$.
+\end{theorem}
+
+\begin{proof}
+First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|T \circ P_0 - \pi\| \leq \eta \|P_0
+- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i k$, the norm between the stationary distribution is $\|\pi
+- \pi_i'\|\leq \frac{\epsilon_i k}{1-\eta}$.
+
+Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| T_t' \circ T_{t-1}' \circ \cdots
+\circ T_1' P_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
+\alpha_s$, where $\alpha = \frac{\epsilon_i k}{1-\eta}$. Thus, we get:
+\begin{align*}
+ \| T_t' \circ \cdots \circ T_1' P_0 - \pi_0 \| &\leq \|T_t' \circ \cdots \circ T_1' P_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
+ &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
+\end{align*}
+as desired.
+\end{proof}
 
@@
 {\color{blue}
-One concern I have is that our $\Delta'$ values do not seem to be ``sufficiently negative''. If our
+Daniel: One concern I have is that our $\Delta'$ values do not seem to be ``sufficiently negative''. If our
 $\Delta'$ values are negative (but large in absolute value) that means our MH test will reject more

### 2016-05-16 nips_2016.tex
slight updates to the writeup, added in some comments for 5.1

@@
 
-In this section, we explore how well our algorithm will converge to the posterior distribution.
-Define our estimate of $X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm norm} - X_{\rm
-norm}'$, so that $\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as the CDF of $X$.
-Let the transition kernel of complete data at sampling step $i$ be $T_i$, and let the transition
-kernel of minibatch data at sampling step $i$ is $T_i'$.  We show that the distance of $T_i$ and
-$T_i'$ is proportional to our estimate of $X_{\rm norm}$. We also define the stationary distribution
-obtained by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution of our
-sampled data converges to $\pi$.
+In this section, we explore the convergence properties of our algorithm.  Define our estimate of
+$X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm norm} - X_{\rm norm}'$, so that
+$\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as the CDF of $X$.  Let the
+transition kernel of complete data at sampling step $i$ be $T_i$, and let the transition kernel of
+minibatch data at sampling step $i$ is $T_i'$.  We show that the distance of $T_i$ and $T_i'$ is
+proportional to our estimate of $X_{\rm norm}$. We also define the stationary distribution obtained
+by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution of our sampled data
+converges to $\pi$. We use something called the \emph{total variation distance} between two kernels,
+denoted as $\| \cdot \|$.
 
 \begin{lemma}
-If $|X_\xi| < \zeta$ and $\nabla F_X < k$, the total variance distance between $T_i$ and $T_i'$ is
-bounded by $2\zeta k$.
+If $|X_\xi| < \zeta$ and $\nabla F_X < k$, then $\|T_i-T_i'\| \le 2\zeta k$.
 \end{lemma}
@@
 Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
-$\|T_i - T'_i\|_1 \leq 2k\zeta$ with $\| \cdot \|_1$ the standard $L_1$ norm. We can write the
-$T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta') q(\theta'\mid \theta_t) +
-(1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$ is the Dirac delta
-function. The distance between two proposal kernels is
+$\|T_i - T'_i\| \leq 2k\zeta$. We write $T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
+q(\theta'\mid \theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$
+is the Dirac delta function. The total variation distance between two proposal kernels is
 \begin{align*}
-\|T_i - T'_i \|_1 &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
+\|T_i - T'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
 & = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
@@
 Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
-\begin{align*}
-{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^n (\log \Pr(x_i | \theta') - \log \Pr (x_i |\theta_t))\right] = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \\
-& \leq \sum_{i=1}^n {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
-\end{align*}
+\[
+{\rm Var}(\Delta') = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \leq
+\sum_{i=1}^n {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
+\]
 as desired.
@@
 {\color{blue}
-Daniel: TODO
+Daniel: I can borrow some of the description from the Appendix into here, but I need to know what we
+are testing. What is the purpose of this experiment? What do we want to show? Once I know that,
+writing this part will be easier, and for the Gaussian case in particular, the experiment is pretty
+simple and can be run on short notice.
+
+For additional details, see Appendix~\ref{app:gaussian}.
 }
@@
 Daniel: TODO
+
+For additional details, see Appendix~\ref{app:logistic}.
 }
@@
 Daniel: TODO
+
+For additional details, see Appendix~\ref{app:nnet}.
 }
@@
 
-\section{Gaussian Mixture Experiment Details}
+\section{Gaussian Mixture Experiment Details}\label{app:gaussian}
 
@@
 
-\section{Logistic Regression Experiment Details}
+\section{Logistic Regression Experiment Details}\label{app:logistic}
 
@@
 
-\section{Neural Network Experiment Details}
+\section{Neural Network Experiment Details}\label{app:nnet}
 

### 2016-05-16 nips_2016.tex
add temptuerare

@@
 \begin{equation}\label{eq:deltas}
-\Delta = \log \left( \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
-\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid
-\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta')\prod_{i=1}^n p(x_i \mid \theta')q(\theta_t \mid
-\theta')}{p(\theta_t)\prod_{i=1}^n p(x_i \mid \theta_t)q(\theta' \mid
+\Delta = \log \left( \frac{p(\theta') \left ( \prod_{i=1}^N p(x_i \mid \theta') \right)^t q(\theta_t \mid
+\theta')}{p(\theta_t)\left (\prod_{i=1}^N p(x_i \mid \theta_t) \right) ^t q(\theta' \mid
+\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta') \left ( \prod_{i=1}^n p(x_i \mid \theta') \right ) ^t q(\theta_t \mid
+\theta')}{p(\theta_t) \left ( \prod_{i=1}^n p(x_i \mid \theta_t) \right ) ^t q(\theta' \mid
 \theta_t)} \right)
@@
 \Delta' = \log p(\theta') - \log p(\theta_t) + \log q(\theta_t \mid \theta') - \log q(\theta' \mid \theta_t) +
-\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t)).
+\frac{1}{t}\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t)).
 \]
 If we fix $\theta_t$ and $\theta'$, the randomness in our $\Delta'$ samples only comes from the
-minibatch. Since the minibatch is chosen randomly, the distribution of $\sum_{i=1}^n (\log p(x_i\mid
+minibatch. Since the minibatch is chosen randomly, the distribution of $\frac{1}{t}\sum_{i=1}^n (\log p(x_i\mid
 \theta') - \log p(x_i\mid \theta_t))$ converge to a Gaussian by the Central Limit Theorem.
@@
 \epsilon$, and the gradient of the log-likelihood is bounded by a constant factor, i.e. $|\nabla
-(\log \Pr(x_i| \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 m$, where
-$m$ is the minibatch size.
+(\log \Pr(x_i| \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 m /t^2$, where
+$m$ is the minibatch size, and $t$ is the temperature.
 \end{lemma}
@@
 \begin{proof}
-From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
+From Lemma~\ref{lem:gaussian}, only the $\frac{1}{t}\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
 \theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
@@
 \[
-{\rm Var}(\Delta') = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \leq
-\sum_{i=1}^n {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
+{\rm Var}(\Delta') = {\rm Var}\left[\frac{1}{t}\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \leq
+\sum_{i=1}^n \frac{1}{t^2} {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2 /t^2,
 \]
@@
 
-Since ${\rm Var}(\Delta') < \epsilon^2 k^2 m$, we can assume the maximum value of $X_{\xi}$ is
-proportional to the standard deviation of $\Delta$, i.e. $\epsilon$. Thus, we can assume
-$\zeta=\epsilon C$, where $C$ is a constant factor.
+Since ${\rm Var}(\Delta') < \epsilon^2 k^2 m / t^2$, we can assume the maximum value of $X_{\xi}$ is
+proportional to the standard deviation of $\Delta$, i.e. $\epsilon / t$. Thus, we can assume
+$\zeta=\epsilon C/t$, where $C$ is a constant factor.
 
@@
 - \pi_0 \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $,
-where $\alpha_i = \frac{\epsilon_i k }{1-\eta}$.
+where $\alpha_i = \frac{\epsilon_i C }{t_i(1-\eta)}$.
 \end{theorem}
@@
 First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|T \circ P_0 - \pi\| \leq \eta \|P_0
-- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i k$, the norm between the stationary distribution is $\|\pi
-- \pi_i'\|\leq \frac{\epsilon_i k}{1-\eta}$.
+- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i C/t_i$, the norm between the stationary distribution is $\|\pi
+- \pi_i'\|\leq \frac{\epsilon_i C}{t_i(1-\eta)}$.
 
@@
 \circ T_1' P_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
-\alpha_s$, where $\alpha = \frac{\epsilon_i k}{1-\eta}$. Thus, we get:
+\alpha_s$, where $\alpha = \frac{\epsilon_i C}{t_i(1-\eta)}$. Thus, we get:
 \begin{align*}
@@
 \end{proof}
+
+The Theorem above indicates that the difference of the distburition obtained by applying the
 

### 2016-05-16 nips_2016.tex
update nips_2016.tex

@@
 
-{\color{blue}
-Daniel: TODO
-}
-
+We applied our mini-batch Metropolis Hastings algorithm to a Bayesian logistic regression model,
+338 comparing with the adaptive sampling method. We tested both our method and the adaptive sampling
+339 method using a random walk proposer q(θ|θt) = N (θt, σ2 ). Since the random walk proposer does RW
+340 not contain any information about the target distribution, it only relies on the Metropolis Hastings
+341 test to converge to the correct distribution. Therefore, random walk proposer is the ideal choice for
+342 illustrating our algorithm and comparing with the adaptive sampling method.
 \section{Neural Network Experiment Details}\label{app:nnet}

### 2016-05-16 nips_2016.tex
clean the parts before the experriments

--- 2016-05-16 nips_2016.tex
@@
 \begin{equation}\label{eq:deltas}
-\Delta = \log \left( \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
-\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid
-\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta')\prod_{i=1}^n p(x_i \mid \theta')q(\theta_t \mid
-\theta')}{p(\theta_t)\prod_{i=1}^n p(x_i \mid \theta_t)q(\theta' \mid
+\Delta = \log \left( \frac{p(\theta')  \prod_{i=1}^N p(x_i \mid \theta')  q(\theta_t \mid
+\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)  q(\theta' \mid
+\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta') \prod_{i=1}^n p(x_i \mid \theta') q(\theta_t \mid
+\theta')}{p(\theta_t) \prod_{i=1}^n p(x_i \mid \theta_t) q(\theta' \mid
 \theta_t)} \right)
@@
 For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
-(1+\exp(-\Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the condition in
+(1+\exp(-\beta \Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the condition in
 Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily sample from
-it using the following procedure. Let $u$ be a uniform random variable $u \sim {\rm Unif}[0,1]$. At
+it using the following procedure. At
 any iteration with a current parameter $\theta_t$ and a proposed sample $\theta'$, we can compute
-$\Delta$. We accept the candidate if $g(\Delta) > u$, and reject otherwise. To see why this works,
-define a new random variable $X = g^{-1}(u)$ where again, $g$ is the logistic function. Notice that
+$\Delta$.
+Let $u$ be a value sampled from uniform distribution $U \sim {\rm Unif}[0,1]$.
+We accept the candidate if $g(\Delta) > u$, and reject otherwise.
+This testing process is equivalent to sample a value $x$ from distribution whose CDF is $g(x)$, and accept if $\Delta + x > 0$, reject otherwise.
+To see why this works,
+define a new random variable $X = g^{-1}(U)$ where again, $g$ is the logistic function. Notice that
 $X$ has cumulative distribution function $F_X(x) = g(x)$. In other words, the CDF is precisely the
-logistic function. This is because for some $X = x$, we have
+logistic function. This is because for arbitrary $X = x$, we have
 \[
@@
 }
+{\color{red}
+Haoyu: If you don't show the CDF of $X$ is $g()$, how do you define the $X$.
+In order to use rule: $\Delta +X >0$, you need to tell people how to get $X$, thus, we need the distribution of $X$.
+If you just say $Delta > g^{-1}(u)$, then what's X?
+Then how can you decomposite the $X$ to get the $X_{corr}$?
+So we need to at least annouce the distribution of $X$ is $g()$.
+}
 
@@
 \begin{lemma}\label{lem:gaussian}
-If the minibatch data is chosen randomly, then the distribution of $\Delta'$ is Gaussian.
+If the minibatch data is chosen randomly, then the distribution of $\Delta'$ is approximately Gaussian.
 \end{lemma}
@@
 \begin{equation}\label{eq:relationship}
-\Delta' = \Delta + \epsilon, \quad \epsilon \sim \mathcal{N}(0, \sigma^2),
+\Delta' = \Delta + \varepsilon, \quad \varepsilon \sim \mathcal{N}(0, \sigma^2),
 \end{equation}
-for a Gaussian noise term $\epsilon$.  The extra noise in $\Delta'$ means we can no longer perform
+for a Gaussian noise term $\varepsilon$.  The extra noise in $\Delta'$ means we can no longer perform
 the $\Delta + X$ test, so we need to slightly change our acceptance criteria. We can decompose
 $X$ as $X = X_{\rm norm}+X_{\rm corr}$, where the first term is a zero-mean ``noise term'' which
-has the same variance as $\epsilon$, and the second term is a zero-mean ``correction'' term.
+has the same variance as $\varepsilon$, and the second term is a zero-mean ``correction'' term.
 These two convolve to form $X$. The criteria to accept can be rewritten as
@@
 
-Furthermore, since our proposed acceptance function $g(\Delta) = (1+ \exp(-\Delta))^{-1}$ satisfies
+Furthermore, since our proposed acceptance function $g(\Delta) = (1+ \exp(-\beta \Delta))^{-1}$ satisfies
 detailed balance (see Lemma~\ref{lem:detailed_balance}), if $\Delta$ is computed from the complete
@@
 Algorithm~\ref{alg:our_algorithm} concisely describes our MH test.
+
+Our mini-batch MH test can not only be used to perform MCMC, but also can be used to obtain parameter samples around MAP/ML by adding temperature $t$ to the likelihood function.
+Thus, the target distribution become $f = \left ( \prod_{i=1}^N p(x_i \mid \theta') \right )^{t_i}$.
+We can decrease the temperature at each iteration to decline the mixing rate, so that our samples will converge to the MAP/ML point.
 
@@
 minibatch data at sampling step $i$ is $T_i'$.  We show that the distance of $T_i$ and $T_i'$ is
-proportional to our estimate of $X_{\rm norm}$. We also define the stationary distribution obtained
+proportional to our estimate error $X_{\xi}$. We also define the stationary distribution obtained
 by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution of our sampled data
@@
 
-Since ${\rm Var}(\Delta') < \epsilon^2 k^2 m$, we can assume the maximum value of $X_{\xi}$ is
+Since ${\rm Var}(\Delta') < \epsilon^2 k^2 m $, we can assume the maximum value of $X_{\xi}$ is
 proportional to the standard deviation of $\Delta$, i.e. $\epsilon$. Thus, we can assume
@@
 - \pi_0 \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $,
-where $\alpha_i = \frac{\epsilon_i k }{1-\eta}$.
+where $\alpha_i = \frac{\epsilon_i C }{1-\eta}$.
 \end{theorem}
@@
 First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|T \circ P_0 - \pi\| \leq \eta \|P_0
-- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i k$, the norm between the stationary distribution is $\|\pi
-- \pi_i'\|\leq \frac{\epsilon_i k}{1-\eta}$.
+- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i C$, the norm between the stationary distribution is $\|\pi
+- \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
 
@@
 \circ T_1' P_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
-\alpha_s$, where $\alpha = \frac{\epsilon_i k}{1-\eta}$. Thus, we get:
+\alpha_s$, where $\alpha = \frac{\epsilon_i C}{1-\eta}$. Thus, we get:
 \begin{align*}
@@
 
+The Theorem above indicates that the difference of the target distribution and the sampled distribution have two components: $\rho_i$ which are determined by how efficient of $T'_i$, and $\alpha_i$ which are determined by error of our transition kernel.
+Both of these two terms are related to the step size.
+If step size is large, the $T'_i$ is very efficient, but the error of $T'_i$ increases.
+Otherwise, if step size is too small, the $T_i'$ is low efficient, but the error of $T'_i$ is low.
+
 \section{Experiments}\label{sec:experiments}
 
+In this section, we use three experiments to reveal the benefits of our mini-batch MH testing. 
 \subsection{Mixture of Gaussians}\label{ssec:gaussians}

### 2016-05-16 nips_2016.tex
clean the parts before the experriments

--- 2016-05-16 nips_2016.tex
@@
 
-In this section, we use three experiments to reveal the benefits of our mini-batch MH testing. 
+In this section, we use three experiments to reveal the benefits of our mini-batch MH testing. For the first experiment, we illustrate that our mini-batch MH testing can 
 \subsection{Mixture of Gaussians}\label{ssec:gaussians}
@@
 
-{\color{blue}
-Daniel: TODO
-}
-
+We applied our mini-batch Metropolis Hastings algorithm to a Bayesian logistic regression model,
+338 comparing with the adaptive sampling method. We tested both our method and the adaptive sampling
+339 method using a random walk proposer $q(\theta \mid \theta_t) = N (\theta_t, \sigma^2)$. Since the random walk proposer does random walk
+340 not contain any information about the target distribution, it only relies on the Metropolis Hastings
+341 test to converge to the correct distribution. Therefore, random walk proposer is the ideal choice for
+342 illustrating our algorithm and comparing with the adaptive sampling method.
 \section{Neural Network Experiment Details}\label{app:nnet}

### 2016-05-16 nips_2016.tex
start experiment

@@
 
-In this section, we use three experiments to reveal the benefits of our mini-batch MH testing. For the first experiment, we illustrate that our mini-batch MH testing can 
+In this section, we use three experiments to reveal the benefits of our mini-batch MH testing.
+For the first experiment, we illustrate that our mini-batch MH testing can obtain the posterior distribution which is very close to the ideal one.
+For the second experiment, we use Logistic regression to reveal that our mini-batch sampler is much more efficient than adaptive minibatch approach.
+The third experiment shows that with the help of temperature, our mini-batch MH testing can be used to obtain the parameters which are close to MAP.
 \subsection{Mixture of Gaussians}\label{ssec:gaussians}

### 2016-05-16 nips_2016.tex
1st exp

@@
 
-We start with a simple mixture of Gaussians.
+We start with a simple mixture of Gaussians, which has the tied means:
+
+\[
+\theta_1 \sim N(0, \sigma_1^2); \theta_2 \sim N(0, \sigma_2^2); x_i \sim \frac{1}{2}N(\theta_1, \sigma_x^2) + \frac{1}{2} N(\theta_2, \sigma_x^2)
+\]
+where $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.
+200 data points are generated from the model with $\theta_1 = 0$ and $\theta_2 = 1$.
+The theoretical posterior distribution can be seen in Fig.~3a.
+We use random walk as the proposer.
+We set the step size 0.1 as a constant.
+We compare three scenarios: with our mini-batch MH testing (see Fig.~3b), without our mini-batch MH testing (see Fig.~3c), and with the standard MH testing (see Fig.~3d).
 
@@
 
-{\color{blue}
-Daniel: I can borrow some of the description from the Appendix into here, but I need to know what we
-are testing. What is the purpose of this experiment? What do we want to show? Once I know that,
-writing this part will be easier, and for the Gaussian case in particular, the experiment is pretty
-simple and can be run on short notice.
-
-For additional details, see Appendix~\ref{app:gaussian}.
-}
-
+From the figures above, we find that with the help of our mini-batch MH testing we can obtain the sampled distribution which are very close to target one.
+We feature draw our estimation of standard derivation of $X_{\rm norm}$ over iterations, see Fig.~3e.
+We find with even constant step size, the standard derivation decreases, because the gradient of posterior distribution declines and make the scale of $\Delta'$ deminished when the sampled point is apporching MAP.
+Thus, the error our transition kernel will declines with even constant step size as discribed in Sect.~ 4.
 \subsection{Logistic Regression}\label{ssec:logistic}
-
+In this experiment, we use logistic regression to classify the MNIST dataset.
+For similification, we only train the binary classifer for digits 1 vs 7.
+The dataset consisted of 12007 training data and 1000 testing data.
+Similarly, we use random walk as the proposer.
+We compare our mini-batch MH testing with the adaptive mini-batch MH testing.
+Figure 4 shows the loglikelihood value and perdiction accuracy over training process.
+Since the adaptive MH testing method needs to increase the mini-batch size adaptively, we use the cumulative number of data instance each algorithm consumes as the horizontal axis.
 \begin{figure}[ht]

### 2016-05-16 nips_2016.tex
add the finalized figs for exp2

@@
 Since the adaptive MH testing method needs to increase the mini-batch size adaptively, we use the cumulative number of data instance each algorithm consumes as the horizontal axis.
+We find our mini-batch MH testing can achieve the similar (or even better loglikelihood/prediction accuracy) with consuming much fewer data points.
+Thus, we can assert that our mini-batch MH testing is much more efficient than the adaptive one.
+
+In this experiment, we also illustrate that by appling the temperature, we can achieve slight faster convergence than the constant temperature one.
 \begin{figure}[ht]
@@
 
+We use this experiment to illustarte another property of our mini-batch MH testing:
+
 {\color{blue}
 Daniel: TODO
+Add the roofline description.
 
@@
 Daniel: TODO
-
+1) describe the data background
+2) describe the SGHMC proposer, and our other hyperparams
+3) show the results and analysis
 For additional details, see Appendix~\ref{app:nnet}.

### 2016-05-16 nips_2016.tex
add figs
updates to abstract

@@
 \usepackage{microtype}      % microtypography
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 
-\title{A New Metropolis-Hastings Test for Fast Minibatch MCMC Inference and Optimization}
+\title{A Simple Minibatch Acceptance Test for MCMC}
 
@@
 \begin{abstract}
-Markov chain Monte Carlo (MCMC) methods are one of the most popular techniques in machine learning
-for Bayesian inference. With large datasets, however, MCMC methods can be intractably slow due to
-the need to compute the likelihood of all data points each iteration. Borrowing from the
-optimization field, one way to make MCMC faster is to compute the posterior based on a random subset
-of the data each iteration. Such stochastic MCMC has received increased research attention in recent
-years, but these methods generally require adaptively increasing the minibatch size each iteration,
-which can nullify the benefits if too much data is needed. In addition, with large datasets and
-model optimization, it is unclear if obtaining a full posterior distribution is desirable over
-getting a high-quality point MAP or ML estimate of the parameters.  In this paper, we present a new
-Metropolis-Hastings test to fit in our alternative perspective on stochastic MCMC methods for big
-data model optimization. We demonstrate the benefits of our approach with theoretical and empirical
-results.
+Markov chain Monte Carlo (MCMC) methods have many applications in
+machine learning. We are particularly interested in their application
+to modeling very large datasets, where it is impractical to
+perform Metropolis-Hastings tests on an entire dataset. Instead we
+describe a novel acceptance test which is applied to minibatches of
+data. Previous work on minibatch Metropolis-Hastings use adaptive
+batch sizes, which may use far more than one minibatch of data and/or
+yield few samples.  Here we describe a simpler test on fixed-sized
+minibatches. Instead of adaptive batch size, this test requires a
+precondition on the variance of the log proposal probability
+ratio. Given a measurement of this variance, minibatch size can be
+chosen to support the test in one step, and yielding a number of
+samples proportional to the batch/minibatch ratio. Alternatively the
+variance condition can be used to adjust the temperature of the
+sampling distribution to a ``natural'' value to provide regular
+samples at a given minibatch size. The resulting proposal/test is only
+slightly more complex than a simple SGD update. In this paper we
+derive the test, discuss its implementation, and present several
+experiments.
+
 \end{abstract}
@@
 
-Markov chain Monte Carlo (MCMC) sampling is one of the most popular techniques for Bayesian
-computation. MCMC algorithms propose samples from a proposal distribution $q$, and decide whether to
-accept or reject them based on a Metropolis-Hastings test~\cite{Metropolis1953,hastings70}. With
-traditional Bayesian posterior inference, the posterior distribution $p(\theta \mid x_1, \ldots,
-x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ of parameter $\theta$ based on conditionally
-independent data $\{x_i\}_{i=1}^n$ is the target distribution which one aims to approximate by
-sampling $\theta$ values. When $n$ is large (e.g., in the billions), it is computationally difficult
-to compute the exact posterior because the likelihood must be evaluated on each $x_i$.
+Markov chain Monte Carlo (MCMC) sampling is a powerful method for
+computation on intractable distributions. MCMC algorithms propose
+samples from a proposal distribution $q$ which is in general different
+from the target distribution $p$, and decide whether to accept or
+reject them based on an acceptance test. The acceptance test is
+usually a Metropolis-Hastings
+test~\cite{Metropolis1953,hastings70}. For example, for Bayesian
+posterior inference, the posterior distribution $p(\theta \mid x_1,
+\ldots, x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ of
+parameter $\theta$ based on conditionally independent data
+$\{x_i\}_{i=1}^n$ is the target distribution which one aims to
+approximate by sampling $\theta$ values. When $n$ is large (e.g., in
+the billions), it is computationally difficult to compute the exact
+posterior because the likelihood must be evaluated on each $x_i$.
 
@@
 In this experiment, we also illustrate that by appling the temperature, we can achieve slight faster convergence than the constant temperature one.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.32\textwidth}
+        \includegraphics[width=\textwidth]{./figures/acc_hyc.png}
+        \caption{prediction accuracy}
+        \label{fig:logistic_pa}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.32\textwidth}
+        \includegraphics[width=\textwidth]{./figures/ll_hyc.png}
+        \caption{loglikelihood}
+        \label{fig:logistic_ll}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+    %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.32\textwidth}
+        \includegraphics[width=\textwidth]{./figures/roofline_hyc.png}
+        \caption{roofline property}
+        \label{fig:logistic_roof}
+    \end{subfigure}
+    \caption{Experiments of logistic regression}\label{fig:logistic}
+\end{figure}
+
 \begin{figure}[ht]

### 2016-05-16 nips_2016.tex
updates to abstract

@@
 $\{x_i\}_{i=1}^n$ is the target distribution which one aims to
-approximate by sampling $\theta$ values. When $n$ is large (e.g., in
-the billions), it is computationally difficult to compute the exact
-posterior because the likelihood must be evaluated on each $x_i$.
+approximate by sampling $\theta$ values. When $n$ is large, it is
+very expensive to compute the exact posterior because the
+likelihood must be evaluated on every $x_i$.
 
-Recent success in machine learning optimization in the presence of large datasets have been in part
-due to the ubiquitous use of stochastic gradient methods, where a subset of the data is used each
-iteration to compute a noisy estimate of the gradient. These Robbins-Munro
-algorithms~\cite{RobbinsMonro1951} have historically not been used with MCMC methods. A recent
-result from~\cite{cutting_mh_2014} attempted to bridge the gap between minibatch optimization and
-Bayesian inference by proposing an \emph{adaptive minibatch} method. Their method uses a subset of
-the $N$ data points for the MH test each iteration, but requires incrementally increasing the
-minibatch size so that it is more confident in its prediction on accepting or rejecting.
+Many state-of-the-art machine learning methods, and deep learning in
+particular, are based on minibatch updates (such as SGD) to a model.may
+Minibatch updates produce many improvements in the model for each pass
+over the dataset, and have high sample efficiency. They also map very
+well onto hardware such as GPUs. MCMC methods in contrast, typically
+require calculations over the full dataset to produce a new sample. A
+recent result from~\cite{cutting_mh_2014} attempted to bridge the gap
+between minibatch optimization and MCMC by proposing an \emph{adaptive
+  minibatch} method. Their method uses a subset of the $N$ data points
+for the MH test each iteration. However the number of points is not
+known a priori, and is adjusted dynamically during every test. This
+undermines many of the practical advantages of minibatch calculations
+- GPU kernels work best for certain minibatch sizes, memory blocks
+can be reused when minibatch size is fixed etc.
 
-We argue for an alternative approach to using MCMC methods in the context of big data problems. In
-general, model optimization involves finding one point estimate of optimal parameters based on a
-user's goals, which typically refer to MAP or ML estimates. In addition, the more data one has, the
-sharper the posterior distribution's modes. In the extreme case, the posterior would collapse to the
-points corresponding to the modes, and it would be difficult for MCMC methods to escape.
+Here we develop an alternative minibatch acceptance test for MCMC
+methods which uses only one fixed-size minibatch per test. Instead of
+dynamically adjusting the batch size, we add a prequisite for the test
+based on the variance of the log acceptance probability. If this
+variance is small enough, the test requires only a single step. This
+variance changes slowly on typical problems, and we can estimate it
+before performaning the accpeptance test. The minibatch variance
+descreases with the minibatch size, so the variance measurement can be
+used to set the minimum minibatch size.
 
-Therefore, we are not interested in exact inference of the posterior model parameters for a very
-large dataset, but rather, we want to perform either MAP or ML estimation. In both cases the goal is
-to find a good posterior mode, but the exact form of the distribution is arbitrary. In particular,
-non-unit powers of the distribution are fine. That corresponds to changing the temperature of the
-distribution to make it ``flatter''.  Rather than fixing the posterior temperature and trying to use
-adjustable minibatch size to achieve it (which ends up reducing to nearly-full-batch updates) we
-propose to run at the natural temperature induced by a given minibatch size. We can always increase
-the minibatch size to reduce the temperature, but there is no need to do this beyond the point where
-the posterior mode is sharp and well-defined. That will occur at some fixed number of points and far
-less than the full dataset.
+There is an additional advantage when using the test for Bayesian
+posterior inference. As dataset size grows, the posterior distribution
+sharpens and will often concentrate to a few sharp peaks. Samples from
+this distribution will come from those peaks, and will easily become
+stuck in one of them (see section 5.1).  The samples are not exploring
+the posterior parameter space, and are unlikely to concentrate at a
+strong local optimum. They therefore fail to accurately represent a
+sample of the posterior because of inadequate search (inadequate
+mixing). We argue that it is much more natural to sample the
+distribution at a different (higher) temperature determined by the
+minibatch size rather than the dataset size. This temperature is
+determined in a natural way by the variance prerequisite for our
+test. Starting at a high temperature (small minibatch size) which
+mixes easily, one can reduce the temperature (increasing minibatch
+size) until one reaches the desired target distribution temperature.
+Although this approach also involves changing minibath size, the
+changes increase minibatch size monotonically, and only a few changes
+are performed in the course of a sampling session.
+Annealing can also be used for MAP and ML estimation because changing temperature preserves
+posterior modes.
 
-To this end, we present a new Metropolis-Hastings test that can fit in this framework of minibatch
-MCMC methods. We present theoretical and experimental results on our method and show that it can be
-the method of choice for minibatch MCMC. The contributions of this paper are as follows:
+The contributions of this paper are as follows:
 
 \begin{enumerate}[noitemsep]
-    \item We develop a new minibatch MH test which satisfies detailed balance.
-    \item Our MH test is more efficient than the adaptive sampling method.
-    \item The test can incorporate temperature considerations to control mixing rates.
+    \item We develop a new minibatch acceptance test which satisfies detailed balance.
+    \item We compare performance of our new test and the adaptive sampling method.
+    \item Experiments using the approach for posterior estimation.
 \end{enumerate}
 
-{\color{blue}
-Daniel: This section needs work in describing our argument carefully. I will decrease the focus on
-temperature and increase focus on the new MH test once the results are clearer.
-}
+%{\color{blue}
+%Daniel: This section needs work in describing our argument carefully. I will decrease the focus on
+%temperature and increase focus on the new MH test once the results are clearer.
+%}
 

### 2016-05-16 nips_2016.tex
Slight updates to Section 1 and abstract following John's suggestions, also removed my discussion and haoyus discussion on the whole X = g inverse

@@
 \begin{document}
-
-\begin{comment}
-\end{comment}
-
 \maketitle
@@
 to modeling very large datasets, where it is impractical to
-perform Metropolis-Hastings tests on an entire dataset. Instead we
+perform Metropolis-Hastings tests on the full data. To do this, we
 describe a novel acceptance test which is applied to minibatches of
 data. Previous work on minibatch Metropolis-Hastings use adaptive
-batch sizes, which may use far more than one minibatch of data and/or
+batch sizes, which may use far more than one minibatch of data and therefore
 yield few samples.  Here we describe a simpler test on fixed-sized
-minibatches. Instead of adaptive batch size, this test requires a
+minibatches. Instead of adaptive batch sizes, our test requires a
 precondition on the variance of the log proposal probability
-ratio. Given a measurement of this variance, minibatch size can be
-chosen to support the test in one step, and yielding a number of
-samples proportional to the batch/minibatch ratio. Alternatively the
+ratio. Given a measurement of this variance, a minibatch size can be
+chosen to support the test in one step, which yields a number of
+samples proportional to the batch/minibatch ratio. Alternatively, the
 variance condition can be used to adjust the temperature of the
@@
 experiments.
-
 \end{abstract}
@@
 Many state-of-the-art machine learning methods, and deep learning in
-particular, are based on minibatch updates (such as SGD) to a model.may
-Minibatch updates produce many improvements in the model for each pass
+particular, are based on minibatch updates (such as SGD) to a model.
+Minibatch updates produce many improvements to the model for each pass
 over the dataset, and have high sample efficiency. They also map very
-well onto hardware such as GPUs. MCMC methods in contrast, typically
+well onto hardware such as GPUs. In contrast, MCMC methods typically
 require calculations over the full dataset to produce a new sample. A
@@
 known a priori, and is adjusted dynamically during every test. This
-undermines many of the practical advantages of minibatch calculations
-- GPU kernels work best for certain minibatch sizes, memory blocks
-can be reused when minibatch size is fixed etc.
+undermines many of the practical advantages of minibatch calculations: 
+GPU kernels work best for certain minibatch sizes, memory blocks
+can be reused when minibatch size is fixed, etc.
 
-Here we develop an alternative minibatch acceptance test for MCMC
+In this paper, we develop an alternative minibatch acceptance test for MCMC
 methods which uses only one fixed-size minibatch per test. Instead of
-dynamically adjusting the batch size, we add a prequisite for the test
+dynamically adjusting the batch size, we add a prerequisite for the test
 based on the variance of the log acceptance probability. If this
@@
 variance changes slowly on typical problems, and we can estimate it
-before performaning the accpeptance test. The minibatch variance
-descreases with the minibatch size, so the variance measurement can be
+before performing the test. The minibatch variance
+decreases as the minibatch size grows, so the variance measurement can be
 used to set the minimum minibatch size.
 
-There is an additional advantage when using the test for Bayesian
-posterior inference. As dataset size grows, the posterior distribution
-sharpens and will often concentrate to a few sharp peaks. Samples from
+There is another advantage when using our test for Bayesian
+posterior inference. As the dataset size grows, the posterior distribution
+sharpens and often concentrates to a few sharp peaks. Samples from
 this distribution will come from those peaks, and will easily become
-stuck in one of them (see section 5.1).  The samples are not exploring
+stuck in one of them (see Section~\ref{ssec:gaussians}).  The samples are not exploring
 the posterior parameter space, and are unlikely to concentrate at a
 strong local optimum. They therefore fail to accurately represent a
-sample of the posterior because of inadequate search (inadequate
+sample of the posterior because of inadequate search (i.e.,
 mixing). We argue that it is much more natural to sample the
-distribution at a different (higher) temperature determined by the
+distribution at a higher temperature determined by the
 minibatch size rather than the dataset size. This temperature is
@@
 size) until one reaches the desired target distribution temperature.
-Although this approach also involves changing minibath size, the
+Although this approach also involves changing minibatch size, the
 changes increase minibatch size monotonically, and only a few changes
@@
 
-The contributions of this paper are as follows:
+To be precise, the contributions of this paper are as follows:
 
@@
     \item We compare performance of our new test and the adaptive sampling method.
-    \item Experiments using the approach for posterior estimation.
+    \item We experiment using the approach for posterior estimation.
 \end{enumerate}
-
-%{\color{blue}
-%Daniel: This section needs work in describing our argument carefully. I will decrease the focus on
-%temperature and increase focus on the new MH test once the results are clearer.
-%}
 
@@
 
-Ultimately, minibatch SGHMC and Langevin Dynamics is an orthogonal contribution to ours, two ways of
-attempting to solve the same problem of MCMC methods with big datasets. We could augment their
-methods with our MH test, to get the benefit of their high quality proposals. Note that while their
-methods in theory do not need an MH test, in practice, they need one due to discretization of the
-system. Furthermore, tests may also be needed for when the algorithms are run with a constant step
-size.
+Minibatch SGHMC and Langevin Dynamics are orthogonal to our work; two ways of attempting
+to solve the same problem of MCMC methods with big datasets. We could augment their methods with our
+MH test, to get the benefit of their high quality proposals. Note that while their methods in theory
+do not need an MH test, in practice, they need one due to discretization of the system. Furthermore,
+    tests may also be needed for when the algorithms are run with a constant step size.
 
@@
 sufficient mixing and/or to reach better posterior modes. To start, we introduce two key values we
-will use frequently: $\Delta$ and $\Delta'$:
+will use frequently, $\Delta$ and $\Delta'$:
 \begin{equation}\label{eq:deltas}
@@
 For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
-(1+\exp(-\beta \Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the condition in
-Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily sample from
-it using the following procedure. At
-any iteration with a current parameter $\theta_t$ and a proposed sample $\theta'$, we can compute
-$\Delta$.
-Let $u$ be a value sampled from uniform distribution $U \sim {\rm Unif}[0,1]$.
-We accept the candidate if $g(\Delta) > u$, and reject otherwise.
-This testing process is equivalent to sample a value $x$ from distribution whose CDF is $g(x)$, and accept if $\Delta + x > 0$, reject otherwise.
-To see why this works,
-define a new random variable $X = g^{-1}(U)$ where again, $g$ is the logistic function. Notice that
-$X$ has cumulative distribution function $F_X(x) = g(x)$. In other words, the CDF is precisely the
-logistic function. This is because for arbitrary $X = x$, we have
+(1+\exp(-\beta \Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
+condition in Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily
+sample from it using the following procedure. At any iteration with a current parameter $\theta_t$
+and a candidate sample $\theta'$, we can compute $\Delta$. Let $u$ be a uniform random variable $u
+\sim {\rm Unif}[0,1]$. We accept $\theta'$ if $g(\Delta) > u$, and reject otherwise. This process is
+equivalent to sampling a random variable $X$ with cumulative distribution function $F_X(x) = g(x)$,
+and accepting if $\Delta > X$ and rejecting otherwise. To see why, define $X = g^{-1}(u)$ where
+again, $g$ is the logistic function. The CDF of $X$ is precisely the logistic function because for
+arbitrary $X = x$, we have
 \[
@@
 acceptance criteria is equivalently expressed as $\Delta + X>0$, as shown in Figure~\ref{fig:part1}.
-
-{\color{blue}
-Daniel: The above makes sense, but I \emph{still} don't understand why we had to show that the CDF
-of $X$ is the logistic function, because I never use that when I claim that our acceptance test is
-now $\Delta > X$. All I need to say is this: we can pre-multiply $g^{-1}$ to $g(\Delta) \ge u$, and
-our test is $\Delta > X$. Right???
-}
-{\color{red}
-Haoyu: If you don't show the CDF of $X$ is $g()$, how do you define the $X$.
-In order to use rule: $\Delta +X >0$, you need to tell people how to get $X$, thus, we need the distribution of $X$.
-If you just say $Delta > g^{-1}(u)$, then what's X?
-Then how can you decomposite the $X$ to get the $X_{corr}$?
-So we need to at least annouce the distribution of $X$ is $g()$.
-}
 
@@
 
-In this section, we use three experiments to reveal the benefits of our mini-batch MH testing.
-For the first experiment, we illustrate that our mini-batch MH testing can obtain the posterior distribution which is very close to the ideal one.
-For the second experiment, we use Logistic regression to reveal that our mini-batch sampler is much more efficient than adaptive minibatch approach.
-The third experiment shows that with the help of temperature, our mini-batch MH testing can be used to obtain the parameters which are close to MAP.
+In this section, we use three experiments to reveal the benefits of our mini-batch MH testing. For
+the first experiment, we illustrate that our mini-batch MH testing can obtain the posterior
+distribution which is very close to the ideal one. For the second experiment, we use Logistic
+regression to reveal that our mini-batch sampler is much more efficient than adaptive minibatch
+approach. The third experiment shows that with the help of temperature, our mini-batch MH testing
+can be used to obtain the parameters which are close to MAP.
+
 \subsection{Mixture of Gaussians}\label{ssec:gaussians}
 
-We start with a simple mixture of Gaussians, which has the tied means:
+{\color{blue}
+Daniel: TODO Have to show that with more data, the posterior peaks (because it is referenced by the
+introduction).
+}
 
+We start with a simple mixture of Gaussians, which have tied means:
 \[
@@
 \]
-where $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.
-200 data points are generated from the model with $\theta_1 = 0$ and $\theta_2 = 1$.
-The theoretical posterior distribution can be seen in Fig.~3a.
-We use random walk as the proposer.
-We set the step size 0.1 as a constant.
-We compare three scenarios: with our mini-batch MH testing (see Fig.~3b), without our mini-batch MH testing (see Fig.~3c), and with the standard MH testing (see Fig.~3d).
+where $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. 200 data points are generated from the
+model with $\theta_1 = 0$ and $\theta_2 = 1$. The theoretical posterior distribution can be seen in
+Fig.~3a. We use random walk as the proposer. We set the step size 0.1 as a constant. We compare
+three scenarios: with our mini-batch MH testing (see Fig.~3b), without our mini-batch MH testing
+(see Fig.~3c), and with the standard MH testing (see Fig.~3d).
 
@@
 
-From the figures above, we find that with the help of our mini-batch MH testing we can obtain the sampled distribution which are very close to target one.
-We feature draw our estimation of standard derivation of $X_{\rm norm}$ over iterations, see Fig.~3e.
-We find with even constant step size, the standard derivation decreases, because the gradient of posterior distribution declines and make the scale of $\Delta'$ deminished when the sampled point is apporching MAP.
-Thus, the error our transition kernel will declines with even constant step size as discribed in Sect.~ 4.
+From the figures above, we find that with the help of our mini-batch MH testing we can obtain the
+sampled distribution which are very close to target one. We feature draw our estimation of standard
+derivation of $X_{\rm norm}$ over iterations, see Fig.~3e. We find with even constant step size, the
+standard derivation decreases, because the gradient of posterior distribution declines and make the
+scale of $\Delta'$ deminished when the sampled point is apporching MAP. Thus, the error our
+transition kernel will declines with even constant step size as discribed in Sect.~ 4.
+
 \subsection{Logistic Regression}\label{ssec:logistic}
-In this experiment, we use logistic regression to classify the MNIST dataset.
-For similification, we only train the binary classifer for digits 1 vs 7.
-The dataset consisted of 12007 training data and 1000 testing data.
-Similarly, we use random walk as the proposer.
-We compare our mini-batch MH testing with the adaptive mini-batch MH testing.
-Figure 4 shows the loglikelihood value and perdiction accuracy over training process.
-Since the adaptive MH testing method needs to increase the mini-batch size adaptively, we use the cumulative number of data instance each algorithm consumes as the horizontal axis.
-We find our mini-batch MH testing can achieve the similar (or even better loglikelihood/prediction accuracy) with consuming much fewer data points.
-Thus, we can assert that our mini-batch MH testing is much more efficient than the adaptive one.
+
+In this experiment, we use logistic regression to classify the MNIST dataset. For similification, we
+only train the binary classifer for digits 1 vs 7. The dataset consisted of 12007 training data and
+1000 testing data. Similarly, we use random walk as the proposer. We compare our mini-batch MH
+testing with the adaptive mini-batch MH testing. Figure 4 shows the loglikelihood value and
+perdiction accuracy over training process. Since the adaptive MH testing method needs to increase
+the mini-batch size adaptively, we use the cumulative number of data instance each algorithm
+consumes as the horizontal axis. We find our mini-batch MH testing can achieve the similar (or even
+better loglikelihood/prediction accuracy) with consuming much fewer data points. Thus, we can assert
+that our mini-batch MH testing is much more efficient than the adaptive one.
 

### 2016-05-17 nips_2016.tex
Updated the Algorithm 1 portion of the NIPS draft

@@
 \]
-If we fix $\theta_t$ and $\theta'$, the randomness in our $\Delta'$ samples only comes from the
-minibatch. Since the minibatch is chosen randomly, the distribution of $\sum_{i=1}^n (\log p(x_i\mid
-\theta') - \log p(x_i\mid \theta_t))$ converge to a Gaussian by the Central Limit Theorem.
+During a given iteration, $\theta_t$ and $\theta'$ are fixed. Therefore, the randomness in $\Delta'$
+comes only from the minibatch. Since the minibatch is chosen randomly, the distribution of
+$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t))$ converges to a Gaussian by the
+Central Limit Theorem.
 \end{proof}
@@
 \end{equation}
-This deconvolution process is visualized in Figure~\ref{fig:part2}. {\color{blue} (Daniel: note that
-we will be fixing this figure, so the visualization should hopefully be a little clearer.)} We
-decompose $X$ this way so that we can estimate the two components separately.  Notice that if we
-increase the mini-batch size, this will cause the $X_{\rm noise}$ distribution to shrink and become
-narrower.
+This deconvolution process is visualized in Figure~\ref{fig:part2}. We decompose $X$ this way so
+that we can estimate $X_{\rm norm}$ from the data, and then add the correction $X_{\rm corr}$ later.
+Notice that we have some control over the true value of $X_{\rm norm}$, even though we cannot
+generally guarantee computing it exactly. In particular, increasing the minibatch size will cause
+the $X_{\rm norm}$ distribution to become narrower. {\color{blue} (Daniel: note that we will be
+fixing this figure, so the visualization should hopefully be a little clearer, along with the
+corresponding paragraph description. In particular, this paragraph needs an informal description of
+Algorithm~\ref{alg:our_algorithm}.)} Algorithm~\ref{alg:our_algorithm} concisely describes our MH
+test.
 
-Furthermore, since our proposed acceptance function $g(\Delta) = (1+ \exp(-\beta \Delta))^{-1}$ satisfies
+Since our proposed acceptance function $g(\Delta) = (1+ \exp(-\beta \Delta))^{-1}$ satisfies
 detailed balance (see Lemma~\ref{lem:detailed_balance}), if $\Delta$ is computed from the complete
@@
 
-Algorithm~\ref{alg:our_algorithm} concisely describes our MH test.
-
-Our mini-batch MH test can not only be used to perform MCMC, but also can be used to obtain parameter samples around MAP/ML by adding temperature $t$ to the likelihood function.
-Thus, the target distribution become $f = \left ( \prod_{i=1}^N p(x_i \mid \theta') \right )^{t_i}$.
-We can decrease the temperature at each iteration to decline the mixing rate, so that our samples will converge to the MAP/ML point.
+It is also important to note that our minibatch MH test can also be used to obtain parameter
+estimates close to the MAP and/or ML estimaes by augmenting the likelihood function with a
+temperature term $T > 1$. Thus, the target distribution would become
+\begin{equation}\label{eq:temperature}
+f(\theta') = p(\theta')\left ( \prod_{i=1}^N p(x_i \mid \theta') \right)^{1/T}.
+\end{equation}
+We can decrease the temperature as the MCMC sampling proceeds, thus slowing the mixing rate and
+sharpening the peaks of the distribution, so that the samples coverge to the desired point estimate.
 
 {\color{blue}
-Daniel: I really need to make this clear and do the algorithm description.
+Daniel: Haoyu and I are going to make this section a lot more clearer, along with the algorithm
+description.
 }
 
-\begin{comment}
-though it really should be an approximation as we can't guarantee that our $X_{\rm norm}$ term is
-exactly the correct amount of noise), where the left hand side has distributions representing normal
-noise ($X_{\rm norm}$, which is labeled as ${\rm Noise}_U$, but they are the same thing) and a
-``correction'' term labeled $X_{\rm correction}$. Be aware that all of these ``additions'' are to be
-interpreted as convolutions when computing the sum of densities.
-
-I think it's becoming clear but I'm still trying to connect all of the steps.  During each iteration
-of MCMC, we compute $\Delta'$ and \emph{sample} (right?) an $X$ (which we can do by sampling $u$ and
-then doing $X = g^{-1}(u)$. This gives us $\Delta' + X$ but our MH test actually requires $X_{\rm
-correction}$, so first we must estimate $\sigma^2$, the variance of $X_{\rm noise} \sim
-\mathcal{N}(0, \sigma^2)$ (perhaps from the mini-batch data, somehow?) and then somehow ``remove''
-that component from $X$ to get $X_{\rm correction}$? Then we have our acceptance test. Is this
-right? Or, of course, we don't have to sample $X$ at all, but just figure out a way to sample
-$X_{\rm correction}$ directly.
-
-I think we talked about estimating $\sigma^2$. The variance $\sigma^2$ that we're concerned about is
-a scalar value based on the sum of log probabilities of different samples, so one way we could
-estimate the variance is by taking all samples of a mini-batch and seeing how much those individual
-$\log p(x_i \mid \theta)$ vary. Or we could look at $k$ mini-batches of data which will give us
-$\Delta_1, \ldots, \Delta_k$ and compute the empirical $\sigma^2$ (of course that defeats the
-purpose of using mini-batches).
-
-I agree with you that out MH test requires $X_{correction}$. My understanding
-is that we can actually estimate $X$ from the whole data set, and the probability distribution of
-$X_{norm}$ is what we proposed? So the determination of $X_{correction}$ is a deterministic
-deconvolution problem.
-\end{comment}
+\SetKwInOut{Input}{Input}
+\SetKwInOut{Output}{Output}
 
 \begin{algorithm}[t]
-\KwData{this text}
-\KwResult{how to write algorithm with \LaTeX2e } initialization\;
-\While{not at end of this document}{
-    read current\;
-    \eIf{understand}{
-        go to next section\;
-        current section becomes this one\;
+\Input{Samples $T$, number of $\Delta'$ estimates $K$, minibatch size $m$, pre-computed $X_{\rm corr}$ values.}
+\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ for either posterior inference or point estimation.}
+Initialize $\theta_1$\;
+\For{$t=\{2, \ldots, T\}$}{
+    Create a new empty list for $\Delta'$ estimates, $d = []$\;
+    \For{$k = \{1, \ldots, K\}$}{
+        Sample a random minibatch of size $m$ from the complete data\;
+        Compute $\Delta'$ from this minibatch (see Equation~\ref{eq:deltas}), and add $\Delta'$ to $d$\;
+    }
+    Compute ${\rm std}(d)$, the standard deviation of the $\Delta'$ elements from list $d$\;
+    Using a random minibatch of $m$ data points, compute a \emph{new} $\Delta'$; call it $\Delta_{\rm real}'$\;
+    Sample $u \sim {\rm Unif}[0,1]$ and extract the corresponding $X_{\rm corr}$ from pre-computed data\;
+    Re-scale the correction error: $X_{\rm corr} = X_{\rm corr} \cdot {\rm std}(d)$\;
+    {\color{blue} Daniel: TODO, check if we need some variance correction or something here.}\;
+    \eIf{$X_{\rm corr} + \Delta_{\rm real}' > 0$}{
+        Accept the candidate, $\theta_{t+1} = \theta'$\;
     }{
-        go back to the beginning of current section\;
+        Reject and re-use the old one, $\theta_{t+1} = \theta_t$\;
     }
 }
-\caption{A description of our MH test.}
+\caption{A description of our MH test. {\color{blue} Daniel: IN PROGRESS!}}
 \label{alg:our_algorithm}

### 2016-05-17 nips_2016.tex
keep writing experiments

--- 2016-05-16 nips_2016.tex
@@
 
-We use this experiment to illustarte another property of our mini-batch MH testing:
+We use this experiment to illustarte another property of our mini-batch MH testing: when the size of mini-batch increase to a threshold, the accuracy / loglikelihood of sampled data at a fix iteration number will not improve with larger mini-batch size, see Fig.~4c.
+This is because the variance of $\Delta'$ decrease with large mini-batch, and the error of our mini-batch MH testing is proportional to our estimation error of variance of $\Delta'$.
+When the $var(\Delta')$ decrease to a threshold, the error of our mini-batch MH testing can be ignored.
 
 {\color{blue}
-Daniel: TODO
-Add the roofline description.
 
@@
 \subsection{Neural Network Optimization}\label{ssec:nets}
-
+In this experiment, we apply about mini-batch MH testing with SGHMC proposer to sample the parameter instances from 
 \begin{figure}[ht]

### 2016-05-17 nips_2016.tex
adding exp

--- 2016-05-17 nips_2016.tex
@@
 \PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage{nips_2016}
+%\usepackage{nips_2016}
 % to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{nips_2016}
+\usepackage[final]{nips_2016}
 
@@
 \author{
-  John Canny \\
-  Department of Computer Science \\
-  University of California, Berkeley \\
-  \texttt{canny@berkeley.edu}
-  \And
   Haoyu Chen \\
@@
   \texttt{bjiang@berkeley.edu}
+  \And
+  John Canny \\
+  Department of Computer Science \\
+  University of California, Berkeley \\
+  \texttt{canny@berkeley.edu}
 }
@@
 \subsection{Neural Network Optimization}\label{ssec:nets}
-In this experiment, we apply about mini-batch MH testing with SGHMC proposer to sample the parameter instances from 
+In this experiment, we apply about mini-batch MH testing with SGHMC proposer to sample the parameter instances from a fully connected neural network.
 \begin{figure}[ht]

### 2016-05-17 nips_2016.tex
still writing exp

@@
 \begin{algorithm}[t]
-\KwData{this text}
-\KwResult{how to write algorithm with \LaTeX2e } initialization\;
-\While{not at end of this document}{
-    read current\;
-    \eIf{understand}{
-        go to next section\;
-        current section becomes this one\;
+\KwData{$\{x_i\}$ under a target distribution}
+\KwResult{a serise of sampled $\{\theta_i\}$ under the posterior distribution} 
+set the $var(\Delta') = 1.0$ and deconvolve $X$ to get the disribution of $X_{corr}$\;
+\While{not enough sampled data}{
+	estimate the $sd(\Delta')$\;
+    propose a candidate $\theta'$\;
+    compute the noisy $\Delta'$\;
+    sample a $x$ from $X_{corr}$;
+    \eIf{$x*sd(\Delta') + \Delta' > 0$}{
+        accpet the $\theta'$;
     }{
-        go back to the beginning of current section\;
+        reject the $\theta'$;
     }
@@
 
-\begin{figure}[ht]
-  \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  %\includegraphics[width=0.5\linewidth]{empty}
-  \caption{This will be a sequence of images for logistic regression.}
-\end{figure}
-
 We use this experiment to illustarte another property of our mini-batch MH testing: when the size of mini-batch increase to a threshold, the accuracy / loglikelihood of sampled data at a fix iteration number will not improve with larger mini-batch size, see Fig.~4c.
@@
 In this experiment, we apply about mini-batch MH testing with SGHMC proposer to sample the parameter instances from a fully connected neural network.
+We use Higgs data set from University of California Irvine.
+It is a classification problem to distinguish between a signal process which produces Higgs bosons and a background process which does not.
+We build a 4 layers neural network to tackle this problem.
+The first layer is a input layer, the last layer is a softmax layer, the intermediate layers are linear followed by Sigmoid, norm, dropout, starting and ending with linear operation.
+
+We choose the 
 \begin{figure}[ht]

### 2016-05-17 nips_2016.tex
fix some error in appendix logistic regression part

@@
 
-We applied our mini-batch Metropolis Hastings algorithm to a Bayesian logistic regression model,
-338 comparing with the adaptive sampling method. We tested both our method and the adaptive sampling
-339 method using a random walk proposer $q(\theta \mid \theta_t) = N (\theta_t, \sigma^2)$. Since the random walk proposer does random walk
-340 not contain any information about the target distribution, it only relies on the Metropolis Hastings
-341 test to converge to the correct distribution. Therefore, random walk proposer is the ideal choice for
-342 illustrating our algorithm and comparing with the adaptive sampling method.
+We applied our mini-batch Metropolis Hastings algorithm to a Bayesian logistic regression model, comparing with the adaptive sampling method.
+We tested both our method and the adaptive sampling method using a random walk proposer $q(\theta \mid \theta_t) = N (\theta_t, \sigma^2)$.
+Since the random walk proposer does not contain any information about the target distribution, it only relies on the Metropolis-Hastings test to converge to the correct distribution.
+Therefore, random walk proposer is the ideal choice for illustrating our algorithm and comparing with the adaptive sampling method.
+
 \section{Neural Network Experiment Details}\label{app:nnet}

### 2016-05-17 nips_2016.tex
Adding in some of my earlier changes for the writeup

@@
 \PassOptionsToPackage{numbers, compress}{natbib}
-%\usepackage{nips_2016}
+\usepackage{nips_2016}
 % to compile a camera-ready version, add the [final] option, e.g.:
-\usepackage[final]{nips_2016}
+%\usepackage[final]{nips_2016}
 
@@
 
-Algorithm~\ref{alg:our_algorithm} concisely describes our MH test.
-
-Our mini-batch MH test can not only be used to perform MCMC, but also can be used to obtain parameter samples around MAP/ML by adding temperature $t$ to the likelihood function.
-Thus, the target distribution become $f = \left ( \prod_{i=1}^N p(x_i \mid \theta') \right )^{t_i}$.
-We can decrease the temperature at each iteration to decline the mixing rate, so that our samples will converge to the MAP/ML point.
+It is also important to note that our minibatch MH test can also be used to obtain parameter
+estimates close to the MAP and/or ML estimaes by augmenting the likelihood function with a
+temperature term $T > 1$. Thus, the target distribution would become
+\begin{equation}
+\label{eq:temperature} f(\theta') = p(\theta')\left ( \prod_{i=1}^N p(x_i \mid \theta')\right)^{\frac{1}{T}}.
+\end{equation}
+We can decrease the temperature as the MCMC sampling proceeds, thus slowing the mixing rate and
+sharpening the peaks of the distribution, so that the samples coverge to the desired point estimate.
 
 {\color{blue}
-Daniel: I really need to make this clear and do the algorithm description.
+Daniel: we really need to make this clear and do the algorithm description.
 }
-
-\begin{comment}
-though it really should be an approximation as we can't guarantee that our $X_{\rm norm}$ term is
-exactly the correct amount of noise), where the left hand side has distributions representing normal
-noise ($X_{\rm norm}$, which is labeled as ${\rm Noise}_U$, but they are the same thing) and a
-``correction'' term labeled $X_{\rm correction}$. Be aware that all of these ``additions'' are to be
-interpreted as convolutions when computing the sum of densities.
-
-I think it's becoming clear but I'm still trying to connect all of the steps.  During each iteration
-of MCMC, we compute $\Delta'$ and \emph{sample} (right?) an $X$ (which we can do by sampling $u$ and
-then doing $X = g^{-1}(u)$. This gives us $\Delta' + X$ but our MH test actually requires $X_{\rm
-correction}$, so first we must estimate $\sigma^2$, the variance of $X_{\rm noise} \sim
-\mathcal{N}(0, \sigma^2)$ (perhaps from the mini-batch data, somehow?) and then somehow ``remove''
-that component from $X$ to get $X_{\rm correction}$? Then we have our acceptance test. Is this
-right? Or, of course, we don't have to sample $X$ at all, but just figure out a way to sample
-$X_{\rm correction}$ directly.
-
-I think we talked about estimating $\sigma^2$. The variance $\sigma^2$ that we're concerned about is
-a scalar value based on the sum of log probabilities of different samples, so one way we could
-estimate the variance is by taking all samples of a mini-batch and seeing how much those individual
-$\log p(x_i \mid \theta)$ vary. Or we could look at $k$ mini-batches of data which will give us
-$\Delta_1, \ldots, \Delta_k$ and compute the empirical $\sigma^2$ (of course that defeats the
-purpose of using mini-batches).
-
-I agree with you that out MH test requires $X_{correction}$. My understanding
-is that we can actually estimate $X$ from the whole data set, and the probability distribution of
-$X_{norm}$ is what we proposed? So the determination of $X_{correction}$ is a deterministic
-deconvolution problem.
-\end{comment}
 
@@
     compute the noisy $\Delta'$\;
-    sample a $x$ from $X_{corr}$;
+    sample a $x$ from $X_{corr}$\;
     \eIf{$x*sd(\Delta') + \Delta' > 0$}{
-        accpet the $\theta'$;
+        Accept the $\theta'$;
     }{
-        reject the $\theta'$;
+        Reject the $\theta'$;
     }

### 2016-05-17 generateXcorr/ecdf_decompo.m
add new ecdf data

@@ -1,0 +1,53 @@
+function [x, ecdf] = ecdf_decompo(sd, isdraw)
+
+if ~exist('isdraw'), isdraw = 1; end
+if ~exist('ratio'), ratio = 0.995; end
+
+n = 2000;
+iclip = 10;
+alpha = sd;
+beta = 1.0;
+dist = 10;
+
+g = @(x, beta) beta*exp(-beta*x)./(1+exp(-beta*x)).^2;
+
+x = [(0:n), ((-n):-1)]/n*dist;
+
+ldist = g(x, beta);
+ldist = ldist * abs(x(2)-x(1));
+
+ndist = normpdf(x, 0, alpha);
+ndist = ndist * abs(x(2)-x(1));
+if (isdraw)
+    figure(2); plot(x, ldist, x , ndist,'r');
+    legend('logistic', 'noisy')
+end
+ff = (fft(ldist));
+
+nff = (fft(ndist));
+
+fdiv = ff./nff;
+fdiv(iclip:(length(fdiv)-iclip+2)) = 0;
+
+u = ifft(fdiv);
+u(u<0) = 0;
+% show the x_corr
+u= u/ trapz(x,u);
+n_de = 2000;
+x_new = ((1:(2*n_de))/n_de - 1)* dist;
+u_new = interp1(x, u, x_new, 'linear');
+% linear iterplot
+Ecdf = zeros(length(u_new)-1,1);
+for i = 1:1:(length(u_new)-1)
+    Ecdf(i) = trapz(x_new(1:(i+1)), u_new(1:(i+1)));
+end
+Ecdf = Ecdf' + (1:length(Ecdf))/length(Ecdf)*0.000001*max(Ecdf);
+Ecdf = Ecdf / max(Ecdf);
+
+if (isdraw)
+    figure(1); plot(x, u, 'b'); 
+end
+x = x_new(2:length(x_new));
+ecdf = Ecdf;
+ecdf = ecdf / max(ecdf);
+end

### 2016-05-17 generateXcorr/generateECDF.m

@@ -1,0 +1,16 @@
+% generate the ecdf matrix
+clear
+clc
+close all
+N = 100;    % 100 hundred curves
+[xtmp, cdftmp] = ecdf_decompo(1.0,0);
+n = length(xtmp);
+res = zeros(N+1, n);
+res(1, :) = xtmp;
+maxSD = 1.2;
+minSD = 0.1;
+sdval = zeros(N);
+for i = 1:1:N
+    sdval(i) =  minSD + (maxSD - minSD) / N *(i-1);
+    [x, cdf] = ecdf_decompo(sdval(i), 0);
+end

### 2016-05-17 nips_2016.tex
Replaced two figures with one nicer deconvolution visualization

@@
   \centering
-  \includegraphics[width=\textwidth]{john_bair_fig01}
+  \includegraphics[width=\textwidth]{mh_convolution_diagram}
   \caption{
-  To the left is the logistic distribution, representing a possible acceptance test. As $\Delta U$
-  approaches infinity, the acceptance rate approaches 1. To the right, we have a new variable $X$
-  and its density, along with an acceptance test $\Delta U + X$, which we describe in detail in
-  Section~\ref{sec:our_algorithm} (note that we use $\Delta$ instead of $\Delta U$ for simplicity).
   {\color{blue}
-  Daniel: I think that the diagram to the left should use the logistic function, i.e., replace
-  ``$\exp(\Delta U)$'' with ``$(1+\exp(\Delta U))^{-1}$.''
+  Daniel: this image is finalized, but change this caption.
   }
   }
-  \label{fig:part1}
-\end{figure}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\textwidth]{john_bair_fig02}
-  \caption{
-  To the left, we have $\Delta U$ (which is simplified as $\Delta$ in
-  Section~\ref{sec:our_algorithm}), along with two noise terms that come from the new variable we
-  define $X$. The details are in Section~\ref{sec:our_algorithm}.
-  {\color{blue}
-  Daniel: I think this needs to be merged into Figure~\ref{fig:part1}.
-  }
-  }
-  \label{fig:part2}
+  \label{fig:deconvolution}
 \end{figure}
@@
 candidate $\theta'$ is equivalent to whether $\Delta > X$. Since $X$ is symmetric about zero, the
-acceptance criteria is equivalently expressed as $\Delta + X>0$, as shown in Figure~\ref{fig:part1}.
+acceptance criteria is equivalently expressed as $\Delta + X>0$.
 
@@
 \end{equation}
-This deconvolution process is visualized in Figure~\ref{fig:part2}. We decompose $X$ this way so
-that we can estimate $X_{\rm norm}$ from the data, and then add the correction $X_{\rm corr}$ later.
-Notice that we have some control over the true value of $X_{\rm norm}$, even though we cannot
-generally guarantee computing it exactly. In particular, increasing the minibatch size will cause
-the $X_{\rm norm}$ distribution to become narrower. {\color{blue} (Daniel: note that we will be
-fixing this figure, so the visualization should hopefully be a little clearer, along with the
+This deconvolution process is visualized in Figure~\ref{fig:deconvolution}. We decompose $X$ this
+way so that we can estimate $X_{\rm norm}$ from the data, and then add the correction $X_{\rm corr}$
+later.  Notice that we have some control over the true value of $X_{\rm norm}$, even though we
+cannot generally guarantee computing it exactly. In particular, increasing the minibatch size will
+cause the $X_{\rm norm}$ distribution to become narrower. {\color{blue} (Daniel: note that we will
+be fixing this figure, so the visualization should hopefully be a little clearer, along with the
 corresponding paragraph description. In particular, this paragraph needs an informal description of

### 2016-05-17 nips_2016.tex
Adding a few references, modified Section 2. Now I will do Section 3, then work on experiments

@@
 difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
-balance. There has been research on ways to conduct exact accept/reject tests using unbiased
-estimators of the likelihood~\cite{Andrieu09thepseudo-marginal}. However, unbiased estimators of the
-likelihood that can be computed from minibatches of data, such as the Kennedy-Bhanot
-estimator~\cite{PhysRevD} have very high variance for large datasets. Because of this, once we get a
-very high estimate of the likelihood, almost all proposed moves are rejected and the algorithm gets
-stuck.
+balance. To reconcile these competing objectives,~\cite{cutting_mh_2014} propose an adaptive
+minibatch MCMC algorithm which uses a sequential hypothesis test. During each iteration, the
+algorithm starts with a small minibatch of data and tests the hypothesis that the sample $\theta'$
+should be accepted or rejected based on the likelihood ratio. If the test cannot make a decision
+over a certain confidence threshold, then they increase the minibatch size and test again. This
+process repeats until a decision.  The downside of this algorithm is the need to keep incrementing
+the mini-batch size in one iteration; in the worst case, all $N$ data points may be needed in a
+single iteration.
 
-{\color{blue}
-Daniel: I am worried. We need to motivate why it is not easy to use minibatch MCMC, and what other
-researchers have tried to do about that.  The last paragraph is from~\cite{cutting_mh_2014}, and it
-seems to argue that people have not used minibatch MCMC before because ``the algorithm gets stuck''.
-But \emph{isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes}? If
-we wanted to get stuck at modes with large datasets, why didn't we just do what previous work did?
-The only way around this is if our algorithm is getting stuck at ``better'' modes, but then we need
-to design our experiments to test this.
-}
-
-The work of~\cite{cutting_mh_2014} proposes an adaptive minibatch approach by using a sequential
-hypothesis testing algorithm. During each iteration, they start with a small minibatch of data and
-test the hypothesis that the sample $\theta'$ should be accepted or rejected based the likelihood
-ratio. If the test cannot make a decision over a certain confidence threshold, then they increase
-the minibatch size and test again. This process repeats until a decision.  The downside of this
-algorithm is the need to keep incrementing the mini-batch size in one iteration; in the worst case,
-all $N$ data points may be needed in a single iteration. A similar approach
-from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties, but comes at the
-cost of more computational time. The work of~\cite{conf/uai/MaclaurinA14} presents another minibatch
-MCMC method, but which gets the \emph{exact} posterior, with the only requirement being that there
-needs to be a lower bound on the log likelihood for each data point. They augment the likelihood
-with auxiliary random variables, but this can hurt mixing.
+A similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical
+properties, but comes at the cost of more computational time. The work
+of~\cite{conf/uai/MaclaurinA14} presents another minibatch MCMC method, but which gets the
+\emph{exact} posterior, with the only requirement being that there needs to be a lower bound on the
+log likelihood for each data point. They augment the likelihood with auxiliary random variables, but
+this can hurt mixing.
 
@@
 do not need an MH test, in practice, they need one due to discretization of the system. Furthermore,
-    tests may also be needed for when the algorithms are run with a constant step size.
-
-%% Daniel: I don't know how to integrate this easily.
-%% \cite{sgmcmc_2015} (NIPS 2015) might want to give this a mention, where it would fit in their
-%% recipe.
+tests may also be needed for when the algorithms are run with a constant step size. {\color{blue}
+Daniel: not totally confident on this paragraph, how much SGHMC/LD do we discuss?}
 
@@
 %% https://xianblog.wordpress.com/2014/04/02/firefly-monte-carlo/
-%% Note: this guy also blogged about our earlier ICLR 2015 paper.
+%% Note: this guy also blogged about our earlier ICLR 2015 paper. He's the third author of the SAME
+%% paper, which explains why he knows this stuff.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \centering
@@
 
-In this experiment, we use logistic regression to classify the MNIST dataset. For similification, we
-only train the binary classifer for digits 1 vs 7. The dataset consisted of 12007 training data and
-1000 testing data. Similarly, we use random walk as the proposer. We compare our mini-batch MH
-testing with the adaptive mini-batch MH testing. Figure 4 shows the loglikelihood value and
-perdiction accuracy over training process. Since the adaptive MH testing method needs to increase
-the mini-batch size adaptively, we use the cumulative number of data instance each algorithm
-consumes as the horizontal axis. We find our mini-batch MH testing can achieve the similar (or even
-better loglikelihood/prediction accuracy) with consuming much fewer data points. Thus, we can assert
-that our mini-batch MH testing is much more efficient than the adaptive one.
+In this experiment, we use logistic regression to classify the MNIST
+dataset~\cite{lecun-mnisthandwrittendigit-2010}. For simplification, we only train the binary
+classifier for digits 1 vs 7. The dataset consisted of 12007 training data and 1000 testing data.
+Similarly, we use random walk as the proposer. We compare our mini-batch MH testing with the
+adaptive mini-batch MH testing. Figure 4 shows the log likelihood value and prediction accuracy over
+training process. Since the adaptive MH testing method needs to increase the mini-batch size
+adaptively, we use the cumulative number of data instance each algorithm consumes as the horizontal
+axis. We find our mini-batch MH testing can achieve the similar (or even better log
+likelihood/prediction accuracy) with consuming much fewer data points. Thus, we can assert that our
+mini-batch MH testing is much more efficient than the adaptive one.
 
-In this experiment, we also illustrate that by appling the temperature, we can achieve slight faster convergence than the constant temperature one.
+In this experiment, we also illustrate that by applying the temperature, we can achieve slight
+faster convergence than the constant temperature one.
 
-\begin{figure}
+\begin{figure}[t]
     \centering
@@
 
-We use this experiment to illustarte another property of our mini-batch MH testing: when the size of mini-batch increase to a threshold, the accuracy / loglikelihood of sampled data at a fix iteration number will not improve with larger mini-batch size, see Fig.~4c.
-This is because the variance of $\Delta'$ decrease with large mini-batch, and the error of our mini-batch MH testing is proportional to our estimation error of variance of $\Delta'$.
-When the $var(\Delta')$ decrease to a threshold, the error of our mini-batch MH testing can be ignored.
+We use this experiment to illustrate another property of our mini-batch MH testing: when the size of
+mini-batch increase to a threshold, the accuracy / log likelihood of sampled data at a fix iteration
+number will not improve with larger mini-batch size, see Fig.~4c. This is because the variance of
+$\Delta'$ decrease with large mini-batch, and the error of our mini-batch MH testing is proportional
+to our estimation error of variance of $\Delta'$. When the $var(\Delta')$ decrease to a threshold,
+the error of our mini-batch MH testing can be ignored.
 
 {\color{blue}
-
-For additional details, see Appendix~\ref{app:logistic}.
+Daniel/Haoyu: For additional details, see Appendix~\ref{app:logistic}.
 }
@@
 \subsection{Neural Network Optimization}\label{ssec:nets}
+
 In this experiment, we apply about mini-batch MH testing with SGHMC proposer to sample the parameter instances from a fully connected neural network.
@@
 We choose the 
-\begin{figure}[ht]
+\begin{figure}[t]
   \centering
@@
 
+In this paper, we have derived a new MH test for minibatch MCMC methods. We demonstrated how a
+simple deconvolution process allows us to use a minibatch approximation to the full data tests. We
+experimentally show the benefits our MH test.
+
+Straightforward directions for future work include running additional experiments, with a particular
+focus on investigation of the variance precondition.  More elaborate extensions include combining
+our results with Hamiltonian Monte Carlo methods, providing a recipe for how to use our algorithm
+(following the general framework of~\cite{sgmcmc_2015}), or integrating parallel
+MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts into our work.
+
 {\color{blue}
-Daniel: TODO
+Daniel: TODO, clean this up.
 }
-
-Straightforward directions for future work include more extensive experiments. More elaborate
-extensions could be combining our results with Hamiltonian Monte Carlo methods, or integrating
-parallel MCMC~\cite{conf/uai/AngelinoKWSA14} concepts into our work.
 

### 2016-05-18 nips_2016.tex
Made major changes to Section 3 to clarify things but still in a draft stage. Also added my python notebook to generate Figure 1.

@@
 difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
-balance. To reconcile these competing objectives,~\cite{cutting_mh_2014} propose an adaptive
+balance. To reconcile these competing objectives,~\cite{cutting_mh_2014} proposes an adaptive
 minibatch MCMC algorithm which uses a sequential hypothesis test. During each iteration, the
@@
 \section{A New Metropolis-Hastings Test}\label{sec:our_algorithm}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\textwidth]{mh_convolution_diagram}
-  \caption{
-  {\color{blue}
-  Daniel: this image is finalized, but change this caption.
-  }
-  }
-  \label{fig:deconvolution}
-\end{figure}
 
@@
 For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
-(1+\exp(-\beta \Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
+(1+\exp(-\Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
 condition in Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily
@@
 equivalent to sampling a random variable $X$ with cumulative distribution function $F_X(x) = g(x)$,
-and accepting if $\Delta > X$ and rejecting otherwise. To see why, define $X = g^{-1}(u)$ where
-again, $g$ is the logistic function. The CDF of $X$ is precisely the logistic function because for
-arbitrary $X = x$, we have
+and accepting if $\Delta > X$ and rejecting otherwise. We can define $X = g^{-1}(u)$ so that its CDF
+is the logistic function, which follows because for arbitrary $X = x$, we have
 \[
@@
 \]
-as the density of the uniform random variable here is simply one. Thus, the criteria to accept the
-candidate $\theta'$ is equivalent to whether $\Delta > X$. Since $X$ is symmetric about zero, the
-acceptance criteria is equivalently expressed as $\Delta + X>0$.
+as the density of $u$ is one. Thus, the criteria to accept the candidate $\theta'$ is equivalent to
+whether $\Delta > X$. Moreover, $X$ is symmetric about zero, so the acceptance criteria can also be
+expressed as $\Delta + X>0$.
 
@@
 \begin{lemma}\label{lem:gaussian}
-If the minibatch data is chosen randomly, then the distribution of $\Delta'$ is approximately Gaussian.
+If the minibatch data is chosen randomly and sufficiently large, then the distribution of $\Delta'$ is approximately Gaussian.
 \end{lemma}
@@
 comes only from the minibatch. Since the minibatch is chosen randomly, the distribution of
-$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t))$ converges to a Gaussian by the
-Central Limit Theorem.
+$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t))$ converges to a Gaussian as $n
+\to \infty$ by the Central Limit Theorem.
 \end{proof}
 
+% Daniel: I decided to put the figure here, so it's closer to the text that references it.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\textwidth]{mh_convolution_diagram}
+    \caption{
+    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
+    standard logistic distribution ($\mu=0$ and $s=1$). The three distributions correspond to three
+    different standard deviation (SD) values of the $X_{\rm norm}$ distribution (the colors carry
+    over to the $X_{\rm corr}$ plot, whose SD values correspond to the $X_{\rm norm}$ SD values).
+    This figure must be viewed in color.
+    }
+    \label{fig:deconvolution}
+    \vspace{-10pt}
+\end{figure}
+
 From Lemma~\ref{lem:gaussian}, since $\Delta'$ is a noisy approximation of $\Delta$, the
-relationship can be expressed as
+relationship is expressed as
 \begin{equation}\label{eq:relationship}
@@
 \end{equation}
-for a Gaussian noise term $\varepsilon$.  The extra noise in $\Delta'$ means we can no longer perform
-the $\Delta + X$ test, so we need to slightly change our acceptance criteria. We can decompose
-$X$ as $X = X_{\rm norm}+X_{\rm corr}$, where the first term is a zero-mean ``noise term'' which
-has the same variance as $\varepsilon$, and the second term is a zero-mean ``correction'' term.
-These two convolve to form $X$. The criteria to accept can be rewritten as
+for a Gaussian noise term $\varepsilon$.  The extra noise means we can no longer directly perform
+the $\Delta + X$ test, so we need to slightly change our acceptance criteria. We can decompose $X$
+as
+\begin{equation}\label{eq:deconvolution}
+X = X_{\rm norm}+X_{\rm corr},
+\end{equation}
+where the first term has the same distribution as $\varepsilon$ and the second term is a zero-mean
+``correction'' term.  These two add (i.e., convolve) to form $X$. The criteria to accept can be rewritten as
 \begin{equation}\label{eq:criteria}
@@
 \end{equation}
-This deconvolution process is visualized in Figure~\ref{fig:deconvolution}. We decompose $X$ this
-way so that we can estimate $X_{\rm norm}$ from the data, and then add the correction $X_{\rm corr}$
-later.  Notice that we have some control over the true value of $X_{\rm norm}$, even though we
-cannot generally guarantee computing it exactly. In particular, increasing the minibatch size will
-cause the $X_{\rm norm}$ distribution to become narrower. {\color{blue} (Daniel: note that we will
-be fixing this figure, so the visualization should hopefully be a little clearer, along with the
-corresponding paragraph description. In particular, this paragraph needs an informal description of
-Algorithm~\ref{alg:our_algorithm}.)} Algorithm~\ref{alg:our_algorithm} concisely describes our MH
-test.
+The approximation exists in Equation~\ref{eq:criteria} because, while we know the complete
+distribution of $X$, the $X_{\rm norm}$ distribution must be estimated from minibatches of data.
+Since $X_{\rm norm}$ is known a priori to be a zero-mean Gaussian, this task reduces to estimating
+its standard deviation.
+
+It is important to understand why we use Equations~\ref{eq:deconvolution} and~\ref{eq:criteria}. The
+deconvolution process allows us to use $\Delta' + X_{\rm corr}$ as a proxy for $\Delta + X$. It
+works because if we know the distribution of $\varepsilon$ (from Equation~\ref{eq:relationship},
+which again reduces to knowing its standard deviation), then we can pretend that our $\Delta'$ is in
+fact $\Delta$. To ``insert'' $\varepsilon$ into our test, we deconvolve $X$ so that one of its
+components, $X_{\rm norm}$, plays the part of $\varepsilon$. This raises two immediate tasks: during
+a given iteration, with current sample $\theta_t$ and proposed sample $\theta'$, how should we get
+the densities of $X_{\rm norm}$ and $X_{\rm corr}$?
+
+\begin{enumerate}[noitemsep]
+    \item To estimate the variance of $X_{\rm norm}$, we generate $K$ values of $\Delta'$, each
+    using a different random minibatch of $n$ data points (but each using the same $\theta_t$ and
+    $\theta'$ parameters). In our experiments (see Section~\ref{sec:experiments}), $K$ is usually
+    small, around five, and $n$ is usually a few hundred for the Central Limit Theorem to apply.
+    Using these $K$ values, we estimate the empirical standard deviation ${\rm std}(\Delta')$.
+
+    \item Once we have estimated ${\rm std}(\Delta')$, we can determine the distribution of
+    $X_{\rm corr}$ from standard deconvolution techniques. These rely on the well-known fact that
+    the Fourier transform of a convolution is the product of the Fourier transforms. For our
+    experiments, we pre-compute $X_{\rm corr}$ values and store them in matrices so that during MCMC
+    runs, we sample a standard uniform random variable and use that to directly obtain $X_{\rm
+    corr}$.
+\end{enumerate}
+
+Figure~\ref{fig:deconvolution} demonstrates this deconvolution process and provides three examples
+of densities for $X_{\rm norm}$ and $X_{\rm corr}$. We know that the densities of $X_{\rm norm}$ and
+$X_{\rm corr}$ convolve to form the density of a standard logistic random variable with mean $\mu=0$
+and scale $s=1$. If $X_{\rm norm}$ has small standard deviation, then the deconvolution to form
+$X_{\rm corr}$ will result in a wider distribution. Likewise, with a larger standard deviation for
+$X_{\rm norm}$, the $X_{\rm corr}$ distribution is narrower, but note that it becomes increasingly
+unstable with bumps in the distribution, because the logistic curve has fatter tails than the
+Gaussian curve. For this reason, we cap the standard deviation possibilities of $X_{\rm norm}$ at
+$\approx 1.2$. We also set the minimum standard deviation of $X_{\rm norm}$ to be 0.1, because below
+that, the $X_{\rm corr}$ distribution essentially matches the distribution of the logistic.
+
+Algorithm~\ref{alg:our_algorithm} concisely describes our MH test.
 
@@
 0)$.  If our estimate of $X_{\rm norm}$ is exact, then $P_a' = P_a$, which follows by making the
-inequality in Equation~\ref{eq:criteria} an equality.
+approximation in Equation~\ref{eq:criteria} an equality.
 
 It is also important to note that our minibatch MH test can also be used to obtain parameter
-estimates close to the MAP and/or ML estimaes by augmenting the likelihood function with a
+estimates close to the MAP and/or ML estimates by augmenting the likelihood function with a
 temperature term $T > 1$. Thus, the target distribution would become
@@
 We can decrease the temperature as the MCMC sampling proceeds, thus slowing the mixing rate and
-sharpening the peaks of the distribution, so that the samples coverge to the desired point estimate.
-
-{\color{blue}
-Daniel: we really need to make this clear and do the algorithm description.
-}
+sharpening the peaks of the distribution, so that the samples converge to the desired point estimate.
 
@@
 \KwData{$\{x_i\}$ under a target distribution}
-\KwResult{a serise of sampled $\{\theta_i\}$ under the posterior distribution} 
-set the $var(\Delta') = 1.0$ and deconvolve $X$ to get the disribution of $X_{corr}$\;
+\KwResult{A series of sampled $\{\theta_i\}$ under the posterior distribution} 
+set the $var(\Delta') = 1.0$ and deconvolve $X$ to get the distribution of $X_{corr}$\;
 \While{not enough sampled data}{
@@
 \end{algorithm}
+
+{\color{blue}
+Daniel: lots of things to do for this section, such as revising the Algorithm 1 description,
+breaking this up into subsections (it's a bit long), and also just making sure it is clear for
+readers to understand.
+}
 

### 2016-05-18 nips_2016.tex
writing exp3

@@
 classifier for digits 1 vs 7. The dataset consisted of 12007 training data and 1000 testing data.
-Similarly, we use random walk as the proposer. We compare our mini-batch MH testing with the
+Similarly, we use random walk as the proposer.
+We initilize the temperature as 3000, and decrease it by $t_i = t_0 / (i + 1)^{0.5}$, where $i$ represent the $t^{th}$ iteration.
+Thus, the standard derivation of $\Delta'$ is smaller than the threshold $\pi/\sqrt{3}$.
+We compare our mini-batch MH testing with the
 adaptive mini-batch MH testing. Figure 4 shows the log likelihood value and prediction accuracy over
@@
 mini-batch MH testing is much more efficient than the adaptive one.
-
-In this experiment, we also illustrate that by applying the temperature, we can achieve slight
-faster convergence than the constant temperature one.
 
@@
 
-In this experiment, we apply about mini-batch MH testing with SGHMC proposer to sample the parameter instances from a fully connected neural network.
+In this experiment, we apply about mini-batch MH testing with SGHMC (Stochastic Gradient Hamiltonian Monte Carlo, \cite{sghmc_2014}) proposer to sample the parameter instances from a fully connected neural network.
 We use Higgs data set from University of California Irvine.
@@
 
-We choose the 
+We choose the Adaptive Gradient Descent method ~\cite{adapGrad} as the comparison baseline.
+Since the frequencies of the features are quite different, we use the same strategy in the Adaptive Gradient to rescale the gradient before apply SGHMC.
+We implement our system within the open-source BIDMach project~\cite{canny2013bidmach}.
+
+Figure~5 illustrates our experiment results.  
 \begin{figure}[t]
@@
 \end{figure}
-
-{\color{blue}
-Daniel: TODO
-1) describe the data background
-2) describe the SGHMC proposer, and our other hyperparams
-3) show the results and analysis
-For additional details, see Appendix~\ref{app:nnet}.
-}
-
-We implement our system within the open-source BIDMach project~\cite{canny2013bidmach}. Note to
-ourselves: do NOT modify the \texttt{Learner.scala} code -- just make it a new ``Model''.
 

### 2016-05-18 nips_2016.tex
revise the algo

@@
 \KwResult{A series of sampled $\{\theta_i\}$ under the posterior distribution} 
-set the $var(\Delta') = 1.0$ and deconvolve $X$ to get the distribution of $X_{corr}$\;
+pre-compute the distribution of $X_{corr}$ under the different value of $var(\Delta')$\;
 \While{not enough sampled data}{
 	estimate the $sd(\Delta')$\;
+	choose the distribution of $X_corr$ from the pre-computed distribution family\;
     propose a candidate $\theta'$\;
@@
     sample a $x$ from $X_{corr}$\;
-    \eIf{$x*sd(\Delta') + \Delta' > 0$}{
+    \eIf{$x + \Delta' > 0$}{
         Accept the $\theta'$;

### 2016-05-18 nips_2016.tex
Added figure about Gaussians to the paper, also fixed the convolution figure and its caption

@@
     \centering
-    \includegraphics[width=\textwidth]{mh_convolution_diagram}
+    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
     \caption{
     Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
-    standard logistic distribution ($\mu=0$ and $s=1$). The three distributions correspond to three
-    different standard deviation (SD) values of the $X_{\rm norm}$ distribution (the colors carry
-    over to the $X_{\rm corr}$ plot, whose SD values correspond to the $X_{\rm norm}$ SD values).
-    This figure must be viewed in color.
+    standard logistic distribution ($\mu=0$, $s=1$); we use three standard deviation (SD) values of
+    $X_{\rm norm}$. The red $X_{\rm norm}$ and $X_{\rm corr}$ curves convolve to form the logistic,
+    etc. The $y$-axis is capped at 0.5 for readability. This figure must be viewed in color.
     }
@@
     \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  %\includegraphics[width=0.5\linewidth]{empty}
-  \caption{This will be a sequence of images for the Gaussian mixture.}
+    \includegraphics[width=1\linewidth]{cloud_v01.png}
+    \caption{
+    {\color{blue}
+    Daniel: I'll fix the caption later. Basically, true log posterior, standard/adaptive procedures,
+    and our method, using MB=200 out of the 10k samples here. All three used the same proposal
+    distribution, a RW with a Gaussian of 0.03.
+    }
+    }
 \end{figure}

### 2016-05-18 nips_2016.tex
Adding another figure for Gaussian Mixtures. Going to sleep tomorrow, will try and do as much writing as possible tomorrow

@@
 stuck in one of them (see Section~\ref{ssec:gaussians}).  The samples are not exploring
-the posterior parameter space, and are unlikely to concentrate at a
+the posterior parameter %vspace, and are unlikely to concentrate at a
 strong local optimum. They therefore fail to accurately represent a
@@
 where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}, so the terms here
-consider the likelihood ratios in log space. The only difference between $\Delta$ and $\Delta'$ is
+consider the likelihood ratios in log %vspace. The only difference between $\Delta$ and $\Delta'$ is
 that they use $N$ and $n$ terms in the likelihood ratio, respectively. We assume $n \ll N$ (so
@@
     \label{fig:deconvolution}
-    \vspace{-10pt}
+    %\vspace{-10pt}
 \end{figure}
@@
 }
-\caption{A description of our MH test.}
+\caption{A description of our MH test. {\color{blue} TODO}}
 \label{alg:our_algorithm}
+%\vspace{-10pt}
 \end{algorithm}
@@
     Daniel: I'll fix the caption later. Basically, true log posterior, standard/adaptive procedures,
-    and our method, using MB=200 out of the 10k samples here. All three used the same proposal
-    distribution, a RW with a Gaussian of 0.03.
+    and our method, using MB=200 out of the 10k samples here (also 10k passes over data). All three
+    used the same proposal distribution, a RW with a Gaussian of 0.03. Temperature is at 110.
     }
     }
+    %\vspace{-10pt}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=1\linewidth]{adaptive_and_ours_information_v01.png}
+    \caption{
+    {\color{blue}
+    Daniel: I'll fix the caption later. First plot is capped at 1k but the first bin goes up to
+    8000. For the SDs, I skipped the ones that didn't satisfy the requirement (a.k.a. repeating the
+    previous sample). The third is when I re-ran the experiments but with different minibatches, to
+    see how much I would have to repeat. These values were all with one run, though, I might
+    consider just taking several (about five?) and doing error bar stuff.
+    }
+    }
+    %\vspace{-10pt}
 \end{figure}
@@
     \caption{Experiments of logistic regression}\label{fig:logistic}
+    %\vspace{-10pt}
 \end{figure}
@@
     \caption{This will hopefully show results for neural networks on BIDMach.}
+    %\vspace{-10pt}
 \end{figure}

### 2016-05-18 nips_2016.tex
Moved Section 4 proofs into appendix. When I get to the office I am doing a lot of rewriting of Sections 2, 3, and 5

@@
 
-\begin{lemma}
+\begin{lemma}\label{lem:theory1}
 If $|X_\xi| < \zeta$ and $\nabla F_X < k$, then $\|T_i-T_i'\| \le 2\zeta k$.
@@
 \begin{proof}
-We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta k$
-for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
-\begin{align*}
-|P_a - P_a'| &= |\Pr(\Delta + X >0) - \Pr(\Delta' + X_{\rm corr} > 0)|\\
-& = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
-&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| = |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
-&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2k \xi,
-\end{align*}
-where the last line uses Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have
-smaller absolute value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| <
-2k\zeta$.
-
-Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
-$\|T_i - T'_i\| \leq 2k\zeta$. We write $T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
-q(\theta'\mid \theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$
-is the Dirac delta function. The total variation distance between two proposal kernels is
-\begin{align*}
-\|T_i - T'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
-& = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
-& \leq 2 \zeta k \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta'))  d\theta_t \right| d\theta'= 2 \zeta k,
-\end{align*}
-as desired.
+See Appendix~\ref{app:theory1}.
 \end{proof}
@@
 
-\begin{lemma}
+\begin{lemma}\label{lem:theory2}
 For one step sampling from $\theta_t$ to $\theta'$, if the jumping step  $|\theta_t - \theta'| <
@@
 \begin{proof}
-From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
-\theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
-$g_i(\theta) = \log p(x_i\mid \theta)$. Then by Taylor's Theorem:
-\[
-|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t) (\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
-\]
-Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
-\[
-{\rm Var}(\Delta') = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \leq
-\sum_{i=1}^n  {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
-\]
-as desired.
+See Appendix~\ref{app:theory2}.
 \end{proof}
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:theory3}
 Assume the stationary distribution $\pi$ satisfies $\|T \circ P_0 - \pi\| \leq \eta \|P_0 - \pi\|$,
@@
 \begin{proof}
-First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|T \circ P_0 - \pi\| \leq \eta \|P_0
-- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i C$, the norm between the stationary distribution is $\|\pi
-- \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
-
-Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| T_t' \circ T_{t-1}' \circ \cdots
-\circ T_1' P_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
-\alpha_s$, where $\alpha = \frac{\epsilon_i C}{1-\eta}$. Thus, we get:
-\begin{align*}
- \| T_t' \circ \cdots \circ T_1' P_0 - \pi_0 \| &\leq \|T_t' \circ \cdots \circ T_1' P_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
- &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
-\end{align*}
-as desired.
+See Appendix~\ref{app:theory3}.
 \end{proof}
 
-The Theorem above indicates that the difference of the target distribution and the sampled distribution have two components: $\rho_i$ which are determined by how efficient of $T'_i$, and $\alpha_i$ which are determined by error of our transition kernel.
-Both of these two terms are related to the step size.
-If step size is large, the $T'_i$ is very efficient, but the error of $T'_i$ increases.
+Theorem~\ref{thm:theory3} indicates that the difference between the target and sampled distributions
+consists of two components: $\rho_i$, which is determined by how efficient of $T'_i$, and the
+$\alpha_i$s, which are determined by error of our transition kernel. Both are related to the step
+size. If step size is large, the $T'_i$ is very efficient, but the error of $T'_i$ increases.
 Otherwise, if step size is too small, the $T_i'$ is low efficient, but the error of $T'_i$ is low.
@@
 \begin{itemize}[noitemsep]
+    \item Some proof details on results in Section~\ref{sec:theory}.
     \item Detailed discussion of the experiment from Section~\ref{ssec:gaussians}.
@@
 \end{itemize}
+
+\section{Proofs}\label{app:proofs}
+
+\subsection{Proof of Lemma~\ref{lem:theory1}}\label{app:theory1}
+
+We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta k$
+for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
+\begin{align*}
+|P_a - P_a'| &= |\Pr(\Delta + X >0) - \Pr(\Delta' + X_{\rm corr} > 0)|\\
+& = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
+&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| = |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
+&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2k \xi,
+\end{align*}
+where the last line uses Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have
+smaller absolute value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| <
+2k\zeta$.
+
+Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
+$\|T_i - T'_i\| \leq 2k\zeta$. We write $T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
+q(\theta'\mid \theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$
+is the Dirac delta function. The total variation distance between two proposal kernels is
+\begin{align*}
+\|T_i - T'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
+& = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
+& \leq 2 \zeta k \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta'))  d\theta_t \right| d\theta'= 2 \zeta k,
+\end{align*}
+as desired.
+
+\subsection{Proof of Lemma~\ref{lem:theory2}}\label{app:theory2}
+
+From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
+\theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
+$g_i(\theta) = \log p(x_i\mid \theta)$. Then by Taylor's Theorem:
+\[
+|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t) (\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
+\]
+Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
+\[
+{\rm Var}(\Delta') = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \leq
+\sum_{i=1}^n  {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
+\]
+as desired.
+
+\subsection{Proof of Theorem~\ref{thm:theory3}}\label{app:theory3}
+
+First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|T \circ P_0 - \pi\| \leq \eta \|P_0
+- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i C$, the norm between the stationary distribution is $\|\pi
+- \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
+
+Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| T_t' \circ T_{t-1}' \circ \cdots
+\circ T_1' P_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
+\alpha_s$, where $\alpha = \frac{\epsilon_i C}{1-\eta}$. Thus, we get:
+\begin{align*}
+ \| T_t' \circ \cdots \circ T_1' P_0 - \pi_0 \| &\leq \|T_t' \circ \cdots \circ T_1' P_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
+ &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
+\end{align*}
+as desired.
 

### 2016-05-18 nips_2016.tex
Updates to Section 2

@@
 
-In the standard MCMC method~\cite{gilks1996markov,brooks2011handbook} for Bayesian inference with
-parameter $\theta$ and conditionally independent data, the goal is to compute the target
-distribution $p(\theta \mid x_1, \ldots, x_N)$.  To do so, one generates a chain of (correlated)
-samples $\theta_1, \ldots, \theta_T$ for large $T$ and approximates $p$ using the sample counts.
-Each iteration $t$ has a current $\theta_t$, and a \emph{proposal distribution} $q(\theta' \mid
+In standard MCMC methods~\cite{gilks1996markov,brooks2011handbook} for Bayesian inference with
+parameter $\theta$ and conditionally independent data, the goal is to compute the distribution
+$p(\theta \mid x_1, \ldots, x_N)$.  To do so, one generates a chain of (correlated) samples
+$\theta_1, \ldots, \theta_T$ for large $T$ and approximates $p$ using the sample counts.  Each
+iteration $t$ has a current $\theta_t$, and a \emph{proposal distribution} $q(\theta' \mid
 \theta_t)$ determines a new candidate $\theta'$. With probability $P_a$, the sample is accepted, so
-$\theta_{t+1} = \theta'$. Otherwise, $\theta_{t+1} = \theta_t$. This can be implemented by drawing
-a uniform random variable $u \sim {\rm Unif}[0,1]$ and accepting if $u < P_a$. Traditionally, $P_a$ is
+$\theta_{t+1} = \theta'$. Otherwise, $\theta_{t+1} = \theta_t$. This is done by drawing a uniform
+random variable $u \sim {\rm Unif}[0,1]$ and accepting if $u < P_a$. Traditionally, $P_a$ is
 \begin{equation}\label{eq:traditional}
@@
 process repeats until a decision.  The downside of this algorithm is the need to keep incrementing
-the mini-batch size in one iteration; in the worst case, all $N$ data points may be needed in a
-single iteration.
+the mini-batch size; in the worst case, all $N$ data points may be needed in a single iteration.  A
+similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties,
+but comes at the cost of more computational time.
 
-A similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical
-properties, but comes at the cost of more computational time. The work
-of~\cite{conf/uai/MaclaurinA14} presents another minibatch MCMC method, but which gets the
-\emph{exact} posterior, with the only requirement being that there needs to be a lower bound on the
-log likelihood for each data point. They augment the likelihood with auxiliary random variables, but
-this can hurt mixing.
+The work of~\cite{conf/uai/MaclaurinA14} presents another minibatch MCMC method which gets the
+\emph{exact} posterior, so long as there exists a (cheap) lower bound on the log likelihood for each
+data point. Their approach is only a starting point, however, and it can be difficult to derive
+these lower bounds in general cases. The extra auxiliary variables they introduce also hurt mixing.
 
-There is a separate line of work related to optimizing MCMC about simulating the physics of a
-probability distribution. By viewing random variables as ``particles'' in a system, it is possible
-to apply Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which generate both high
-quality and distant proposals, getting the best of both worlds. These require a full gradient
-computation of the posterior, which means they also suffer from the problem of needing to know the
-likelihood at each point. Stochastic versions of HMC~\cite{sghmc_2014,stochastic_thermostats_2014}
-use a minibatch of data each iteration but require controlling the noise from sampling so the sample
-sequence does not diverge. A similar approach is to use Langevin
-Dynamics~\cite{langevin_2011,conf/icml/AhnBW12}, which does not use the momentum terms in HMC.
-
-Minibatch SGHMC and Langevin Dynamics are orthogonal to our work; two ways of attempting
-to solve the same problem of MCMC methods with big datasets. We could augment their methods with our
-MH test, to get the benefit of their high quality proposals. Note that while their methods in theory
-do not need an MH test, in practice, they need one due to discretization of the system. Furthermore,
-tests may also be needed for when the algorithms are run with a constant step size. {\color{blue}
-Daniel: not totally confident on this paragraph, how much SGHMC/LD do we discuss?}
+There is a separate line of MCMC work on simulating the physics of a probability distribution. By
+viewing random variables as particles in a system, one can apply Hamiltonian Monte Carlo
+(HMC)~\cite{mcmc_hamiltonian_2010} methods which generate the best case scenario for proposals: high
+quality \emph{and} distant.  HMC methods require a full gradient computation of the posterior, which
+means they also face the problem of computing the likelihood at each point. Stochastic versions of
+HMC~\cite{sghmc_2014,stochastic_thermostats_2014} use a minibatch of data each iteration but require
+controlling the noise from sampling to avoid divergence. Langevin
+Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} is a similar approach which does not use the
+momentum terms in HMC.  Minibatch SGHMC and Langevin Dynamics are orthogonal to our work and are
+other ways of attempting to solve the same problem of MCMC methods on big data. We can combine our
+methods with these results, as we demonstrate in Section~\ref{ssec:nets}.
 

### 2016-05-18 nips_2016.tex
Updated half of section 3, the first four pages of the writeup look ok, but I need to fix the last part of section 3

@@
 
-In this section, we introduce our new Metropolis-Hastings test and show how to use it to achieve
-sufficient mixing and/or to reach better posterior modes. To start, we introduce two key values we
-will use frequently, $\Delta$ and $\Delta'$:
+For our new MH test, we use two key values, $\Delta$ and $\Delta'$:
 \begin{equation}\label{eq:deltas}
@@
 \end{equation}
-where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}, so the terms here
-consider the likelihood ratios in log %vspace. The only difference between $\Delta$ and $\Delta'$ is
-that they use $N$ and $n$ terms in the likelihood ratio, respectively. We assume $n \ll N$ (so
+where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}. Thus, $\Delta$ and
+$\Delta'$ are the likelihood ratios in log space. Note that the only difference between $\Delta$ and
+$\Delta'$ is that they use $N$ and $n$ data terms $x_i$, respectively. We assume $n \ll N$ (so
 $\Delta'$ is substantially faster to compute) and that the $n$ data points $x_i$ represent a random
 minibatch of data.
+
+\subsection{The Full Data Test}\label{ssec:deltas_full}
 
@@
 inverts the fraction inside the logarithm term of $\Delta$.  By assumption, we can expand $g(\Delta)
-= \exp(\Delta)g(-\Delta)$ in $p(\theta' \mid \theta_t)$. Doing this, and combining that with the
-definition of $p(\theta_t \mid \theta')$, we get
+= \exp(\Delta)g(-\Delta)$ in $p(\theta' \mid \theta_t)$. Doing this, and combining the result with
+the definition of $p(\theta_t \mid \theta')$, we get
 \begin{equation}\label{eq:combined}
@@
 
-Some straightforward arithmetic shows that the standard Metropolis-Hastings acceptance function
-$g(\Delta) = \min\{1, e^\Delta \} = \min\left\{1, \frac{f(\theta')q(\theta_t \mid
-\theta')}{f(\theta_t)q(\theta' \mid \theta_t)}\right\}$ satisfies the condition $g(\Delta) =
-\exp(\Delta)g(-\Delta)$ and consequently, results in detailed balance.
+As a sanity check, the standard Metropolis-Hastings acceptance function $g(\Delta) = \min\{1,
+e^\Delta \} = \min\left\{1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid
+\theta_t)}\right\}$ satisfies the condition $g(\Delta) = \exp(\Delta)g(-\Delta)$.
 
 For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
-(1+\exp(-\Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
-condition in Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily
-sample from it using the following procedure. At any iteration with a current parameter $\theta_t$
-and a candidate sample $\theta'$, we can compute $\Delta$. Let $u$ be a uniform random variable $u
-\sim {\rm Unif}[0,1]$. We accept $\theta'$ if $g(\Delta) > u$, and reject otherwise. This process is
+(1+\exp(-\Delta))^{-1}$. Straightforward arithmetic shows that it satisfies the condition in
+Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily sample from
+it using the following procedure. At any iteration with a current parameter $\theta_t$ and a
+candidate sample $\theta'$, we can compute $\Delta$. Let $u$ be a uniform random variable $u \sim
+{\rm Unif}[0,1]$. We accept $\theta'$ if $g(\Delta) > u$, and reject otherwise. This process is
 equivalent to sampling a random variable $X$ with cumulative distribution function $F_X(x) = g(x)$,
@@
 
-With large datasets, $\Delta$ is intractable to compute. An intuitive alternative is to use
-$\Delta'$ instead. The following lemma characterizes the distribution of $\Delta'$.
+\subsection{The Minibatch Version}\label{ssec:deltas_minibatch}
+
+With large datasets, $\Delta$ is intractable to compute. An intuitive fix is to instead use the
+minibatch version $\Delta'$. The following lemma characterizes the distribution of $\Delta'$.
 
@@
 for a Gaussian noise term $\varepsilon$.  The extra noise means we can no longer directly perform
-the $\Delta + X$ test, so we need to slightly change our acceptance criteria. We can decompose $X$
-as
+the $\Delta + X$ test, so we need to slightly change our acceptance criteria. Our insight is to
+decompose $X$ as
 \begin{equation}\label{eq:deconvolution}
@@
 \end{equation}
-where the first term has the same distribution as $\varepsilon$ and the second term is a zero-mean
-``correction'' term.  These two add (i.e., convolve) to form $X$. The criteria to accept can be rewritten as
+where we assume $X_{\rm norm}$ is a zero-mean Gaussian and $X_{\rm corr}$ is a zero-mean
+``correction'' term.  These two add (i.e., their distributions convolve) to form $X$. The criteria
+to accept is now
 \begin{equation}\label{eq:criteria}
@@
 \end{equation}
-The approximation exists in Equation~\ref{eq:criteria} because, while we know the complete
-distribution of $X$, the $X_{\rm norm}$ distribution must be estimated from minibatches of data.
-Since $X_{\rm norm}$ is known a priori to be a zero-mean Gaussian, this task reduces to estimating
-its standard deviation.
+The approximation exists because $X_{\rm norm}$ is an estimate of $\varepsilon$.
 
@@
 deconvolution process allows us to use $\Delta' + X_{\rm corr}$ as a proxy for $\Delta + X$. It
-works because if we know the distribution of $\varepsilon$ (from Equation~\ref{eq:relationship},
-which again reduces to knowing its standard deviation), then we can pretend that our $\Delta'$ is in
-fact $\Delta$. To ``insert'' $\varepsilon$ into our test, we deconvolve $X$ so that one of its
-components, $X_{\rm norm}$, plays the part of $\varepsilon$. This raises two immediate tasks: during
-a given iteration, with current sample $\theta_t$ and proposed sample $\theta'$, how should we get
-the densities of $X_{\rm norm}$ and $X_{\rm corr}$?
+works because if we know the distribution\footnote{Computing the distributions of $\varepsilon$ and
+$X_{\rm norm}$ reduces to finding their standard deviations.} of $\varepsilon$ (from
+Equation~\ref{eq:relationship}), then we can pretend that our $\Delta'$ is in fact $\Delta$. To
+``insert'' $\varepsilon$ into our test, we deconvolve $X$ so that one of its components, $X_{\rm
+norm}$, plays the part of $\varepsilon$. In light of our revised MH test, the previous discussion
+raises the question: in an iteration of MCMC with current $\theta_t$ and proposed $\theta'$, how do
+we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample $X_{\rm
+corr}$ appropriately. This procedure consists of two major steps.
 
 \begin{enumerate}[noitemsep]
-    \item To estimate the variance of $X_{\rm norm}$, we generate $K$ values of $\Delta'$, each
-    using a different random minibatch of $n$ data points (but each using the same $\theta_t$ and
-    $\theta'$ parameters). In our experiments (see Section~\ref{sec:experiments}), $K$ is usually
-    small, around five, and $n$ is usually a few hundred for the Central Limit Theorem to apply.
-    Using these $K$ values, we estimate the empirical standard deviation ${\rm std}(\Delta')$.
+    \item We first need to estimate the standard deviation of $X_{\rm norm}$. We generate $K$ values
+    of $\Delta'$, each using a different random minibatch of $n$ data points, but each using the
+    same $\theta_t$ and $\theta'$. In our experiments (see Section~\ref{sec:experiments}), $K$ is
+    usually small, around 5-10, and $n$ is usually a few hundred for the Central Limit Theorem to
+    apply.  Using these $K$ values, we use the empirical standard deviation ${\rm std}(\Delta')$.
 
-    \item Once we have estimated ${\rm std}(\Delta')$, we can determine the distribution of
+    \item Once we have ${\rm std}(\Delta')$, we can determine the distribution of
     $X_{\rm corr}$ from standard deconvolution techniques. These rely on the well-known fact that
-    the Fourier transform of a convolution is the product of the Fourier transforms. For our
-    experiments, we pre-compute $X_{\rm corr}$ values and store them in matrices so that during MCMC
-    runs, we sample a standard uniform random variable and use that to directly obtain $X_{\rm
-    corr}$.
+    the Fourier transform of a convolution is the product of the Fourier transforms. In practice,
+    we pre-compute distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to
+    facilitate the sampling process.
 \end{enumerate}

### 2016-05-18 nips_2016.tex
Substantially updated section 3. Now sections 1-3, maybe 4, look good, first 5-6 pages look passable

@@
 \theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)  q(\theta' \mid
-\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta') \prod_{i=1}^n p(x_i \mid \theta') q(\theta_t \mid
-\theta')}{p(\theta_t) \prod_{i=1}^n p(x_i \mid \theta_t) q(\theta' \mid
-\theta_t)} \right)
+\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta') (\prod_{i=1}^n p(x_i \mid
+\theta'))^{\frac{N}{n}} q(\theta_t \mid \theta')}{p(\theta_t) (\prod_{i=1}^n p(x_i \mid
+\theta_t))^{\frac{N}{n}} q(\theta' \mid \theta_t)} \right)
 \end{equation}
@@
 $\Delta'$ are the likelihood ratios in log space. Note that the only difference between $\Delta$ and
-$\Delta'$ is that they use $N$ and $n$ data terms $x_i$, respectively. We assume $n \ll N$ (so
-$\Delta'$ is substantially faster to compute) and that the $n$ data points $x_i$ represent a random
-minibatch of data.
+$\Delta'$ is that they use $N$ and $n$ data terms $x_i$, respectively, and that the latter uses an
+$N/n$ scaling term. We assume $n \ll N$ (so $\Delta'$ is substantially faster to compute) and that
+the $n$ data points $x_i$ represent a random minibatch of data.
 
@@
 \Delta' = \log p(\theta') - \log p(\theta_t) + \log q(\theta_t \mid \theta') - \log q(\theta' \mid \theta_t) +
-\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t)).
+\frac{N}{n}\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t)).
 \]
@@
     Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
-    standard logistic distribution ($\mu=0$, $s=1$); we use three standard deviation (SD) values of
-    $X_{\rm norm}$. The red $X_{\rm norm}$ and $X_{\rm corr}$ curves convolve to form the logistic,
-    etc. The $y$-axis is capped at 0.5 for readability. This figure must be viewed in color.
+    standard logistic distribution. We use three standard deviation values of $X_{\rm norm}$. The
+    two red curves convolve to form the logistic, etc. The $y$-axis is capped at 0.5 for
+    readability. This figure must be viewed in color.
     }
@@
 \end{equation}
-The approximation exists because $X_{\rm norm}$ is an estimate of $\varepsilon$.
+The approximation exists because $X_{\rm norm}$ is an estimate of $\varepsilon$. If $X_{\rm norm}$
+is exact, then we get the same MH test with $\Delta'$ as we do with the full data version $\Delta$.
 
@@
 raises the question: in an iteration of MCMC with current $\theta_t$ and proposed $\theta'$, how do
-we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample $X_{\rm
-corr}$ appropriately. This procedure consists of two major steps.
+we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample the
+\emph{values} of $X_{\rm corr}$ appropriately. This procedure consists of two major steps.
 
@@
     same $\theta_t$ and $\theta'$. In our experiments (see Section~\ref{sec:experiments}), $K$ is
-    usually small, around 5-10, and $n$ is usually a few hundred for the Central Limit Theorem to
-    apply.  Using these $K$ values, we use the empirical standard deviation ${\rm std}(\Delta')$.
+    small, around 5-10, and $n$ is usually a few hundred for the Central Limit Theorem to apply.
+    Using these $K$ values, we use the empirical standard deviation ${\rm std}(\Delta')$.
 
-    \item Once we have ${\rm std}(\Delta')$, we can determine the distribution of
-    $X_{\rm corr}$ from standard deconvolution techniques. These rely on the well-known fact that
-    the Fourier transform of a convolution is the product of the Fourier transforms. In practice,
-    we pre-compute distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to
-    facilitate the sampling process.
+    \item Once we have ${\rm std}(\Delta')$, we can determine the distribution of $X_{\rm corr}$
+    from standard deconvolution techniques. These rely on the well-known fact that the Fourier
+    transform of a convolution is the product of the Fourier transforms. In practice, we pre-compute
+    distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to facilitate the
+    sampling process.
 \end{enumerate}
 
-Figure~\ref{fig:deconvolution} demonstrates this deconvolution process and provides three examples
-of densities for $X_{\rm norm}$ and $X_{\rm corr}$. We know that the densities of $X_{\rm norm}$ and
-$X_{\rm corr}$ convolve to form the density of a standard logistic random variable with mean $\mu=0$
-and scale $s=1$. If $X_{\rm norm}$ has small standard deviation, then the deconvolution to form
-$X_{\rm corr}$ will result in a wider distribution. Likewise, with a larger standard deviation for
-$X_{\rm norm}$, the $X_{\rm corr}$ distribution is narrower, but note that it becomes increasingly
-unstable with bumps in the distribution, because the logistic curve has fatter tails than the
-Gaussian curve. For this reason, we cap the standard deviation possibilities of $X_{\rm norm}$ at
-$\approx 1.2$. We also set the minimum standard deviation of $X_{\rm norm}$ to be 0.1, because below
-that, the $X_{\rm corr}$ distribution essentially matches the distribution of the logistic.
+Algorithm~\ref{alg:our_algorithm} describes our MH test within the MCMC algorithm.
 
-Algorithm~\ref{alg:our_algorithm} concisely describes our MH test.
+\subsection{Discussions and Practical Considerations of the New MH Test}\label{ssec:discussion}
 
-Since our proposed acceptance function $g(\Delta) = (1+ \exp(-\beta \Delta))^{-1}$ satisfies
-detailed balance (see Lemma~\ref{lem:detailed_balance}), if $\Delta$ is computed from the complete
-data, the distribution of the sampled data will match our target distribution $f$, the posterior,
-under standard MCMC assumptions (e.g., that the chain is run sufficiently long enough and mixes
-well). Now consider the case of using minibatches of data. Let our true and estimated acceptance
-probabilities be $P_a = g(\Delta) = \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} >
-0)$.  If our estimate of $X_{\rm norm}$ is exact, then $P_a' = P_a$, which follows by making the
-approximation in Equation~\ref{eq:criteria} an equality.
-
-It is also important to note that our minibatch MH test can also be used to obtain parameter
-estimates close to the MAP and/or ML estimates by augmenting the likelihood function with a
-temperature term $T > 1$. Thus, the target distribution would become
-\begin{equation}
-\label{eq:temperature} f(\theta') = p(\theta')\left ( \prod_{i=1}^N p(x_i \mid \theta')\right)^{\frac{1}{T}}.
-\end{equation}
-We can decrease the temperature as the MCMC sampling proceeds, thus slowing the mixing rate and
-sharpening the peaks of the distribution, so that the samples converge to the desired point estimate.
-
+\SetKwInOut{Input}{Input}
+\SetKwInOut{Output}{Output}
 \begin{algorithm}[t]
-\KwData{$\{x_i\}$ under a target distribution}
-\KwResult{A series of sampled $\{\theta_i\}$ under the posterior distribution} 
-pre-compute the distribution of $X_{corr}$ under the different value of $var(\Delta')$\;
-\While{not enough sampled data}{
-	estimate the $sd(\Delta')$\;
-	choose the distribution of $X_corr$ from the pre-computed distribution family\;
-    propose a candidate $\theta'$\;
-    compute the noisy $\Delta'$\;
-    sample a $x$ from $X_{corr}$\;
-    \eIf{$x + \Delta' > 0$}{
-        Accept the $\theta'$;
+\Input{Number of samples $T$, number of $\Delta'$ estimates $K$, minibatch size $m$, pre-computed $X_{\rm
+corr}$ distributions, $\{x_i\}_{i=1}^n$ values from a target distribution, and initial sample $\theta_0$.}
+\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ for either posterior inference or
+point estimation.}
+\For{$t=\{1, \ldots, T\}$}{
+    Propose a candidate $\theta'$ and create an empty list for $\Delta'$ estimates, $d = []$\;
+    \For{$k = \{1, \ldots, K\}$}{
+        Sample a random minibatch of size $m$ from the complete data\;
+        Compute $\Delta'$ from this minibatch (see Equation~\ref{eq:deltas}), and add $\Delta'$ to $d$\;
+    }
+    Compute ${\rm std}(d)$, the standard deviation estimate of $\Delta'$ from list $d$\;
+    \eIf{${\rm std}(d) \ge 1.2$}{
+        Skip loop (set $\theta_{t+1}=\theta_t$) and apply variance preconditioning fix (if desired)\;
     }{
-        Reject the $\theta'$;
+        Choose the closest $X_{\rm corr}$ distribution, and sample a value $X_c$ from it\;
+        Using another random minibatch of $m$ data points, compute a \emph{new} $\Delta'$ (call it $\Delta_{\rm real}'$)\;
+        \eIf{$X_c + \Delta_{\rm real}' > 0$}{
+            Accept the candidate, $\theta_{t+1} = \theta'$\;
+        }{
+            Reject and re-use the old one, $\theta_{t+1} = \theta_t$\;
         }
     }
-\caption{A description of our MH test. {\color{blue} TODO}}
+}
+\caption{A description of our MH test within the MCMC algorithm.}
 \label{alg:our_algorithm}
@@
 
-{\color{blue}
-Daniel: lots of things to do for this section, such as revising the Algorithm 1 description,
-breaking this up into subsections (it's a bit long), and also just making sure it is clear for
-readers to understand.
-}
+\textbf{The Convolution}. Figure~\ref{fig:deconvolution} demonstrates
+Equation~\ref{eq:deconvolution} and provides three (color-coded) examples of densities for $X_{\rm
+norm}$ and $X_{\rm corr}$. The density of $X$, a logistic random variable with mean $\mu = 0$ and
+scale $s=1$, is known, as are the $X_{\rm norm}$ densities for different standard deviations. The
+deconvolution provides us with the $X_{\rm corr}$ distributions. Note, however, that as the standard
+deviation of $X_{\rm norm}$ increases, the $X_{\rm corr}$ distribution becomes increasingly unstable
+and ``bumpier'', because the logistic curve has fatter tails than Gaussians. For this reason, we cap
+the standard deviation possibilities of $X_{\rm norm}$ at $\approx 1.2$ in our experiments. In our
+experiments, we discretize the possible standard deviations of $X_{\rm norm}$, and we also limit the
+densities to consider the range $[-10,10]$ (as shown in Figure~\ref{eq:deconvolution}).
+
+\textbf{Variance Preconditioning}. Returning to the discussion from Section~\ref{sec:introduction},
+our test requires a variance check. Equation~\ref{eq:deconvolution} means that the variance of $X$,
+which is $\pi^2/3\approx 3.29$ is an upper bound on the variances of $X_{\rm norm}$. Therefore, the
+test cannot work if the variance of $X_{\rm norm}$ is too large. Since the standard deviation of $X$
+is about $\sqrt{3.29}\approx 1.81$, this bounds ${\rm std}(X_{\rm norm})$ (as well as ${\rm
+std}(\Delta')$). From the discussion above, we bound the standard deviation at 1.2 so that $X_{\rm
+corr}$ can handle the remaining noise. Recall from Section~\ref{ssec:deltas_minibatch} that we
+estimate the ${\rm std}(X_{\rm norm})$ each iteration. If the variance/standard deviation test
+fails, then we can (a) skip the iteration, (b) increase the minibatch size, (c) change the proposal
+to decrease step sizes, or (d) increase temperature, as we now discuss.
+
+\textbf{Temperature}. One way to satisfy the variance precondition is to increase the temperature of
+the target distribution. For a temperature $T>1$, the augmented target distribution would become
+\begin{equation}\label{eq:log_temperature}
+\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
+\end{equation}
+with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
+smaller, thus reducing variance. The flatter posterior results in easier mixing of the samples, and
+as the samples approach a mode, one can decrease the temperature to sharpen the posterior peaks.
 
@@
 
-In this section, we explore the convergence properties of our algorithm.  Define our estimate of
-$X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm norm} - X_{\rm norm}'$, so that
-$\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as the CDF of $X$.  Let the
-transition kernel of complete data at sampling step $i$ be $T_i$, and let the transition kernel of
-minibatch data at sampling step $i$ is $T_i'$.  We show that the distance of $T_i$ and $T_i'$ is
-proportional to our estimate error $X_{\xi}$. We also define the stationary distribution obtained
-by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution of our sampled data
-converges to $\pi$. We use something called the \emph{total variation distance} between two kernels,
-denoted as $\| \cdot \|$.
+In this section, we explore the convergence properties of our algorithm. Let our true and estimated
+acceptance probabilities be $P_a = g(\Delta) = \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm
+corr} > 0)$. Define our estimate of $X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm
+norm} - X_{\rm norm}'$, so that $\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as
+the CDF of $X$.  Let the transition kernel of complete data at sampling step $i$ be $T_i$, and let
+the transition kernel of minibatch data at sampling step $i$ is $T_i'$.  We show that the distance
+of $T_i$ and $T_i'$ is proportional to our estimate error $X_{\xi}$. We also define the stationary
+distribution obtained by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution
+of our sampled data converges to $\pi$. We use something called the \emph{total variation distance}
+between two kernels, denoted as $\| \cdot \|$.
 

### 2016-05-18 nips_2016.tex
revise the xp

--- 2016-05-18 nips_2016.tex
@@
 $X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm norm} - X_{\rm norm}'$, so that
-$\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as the CDF of $X$.  Let the
+$\Delta' = \Delta + X_{\rm norm} + X_\xi$.
+We assume the standard derivation of $X_{\rm norm}$ is less than $\pi/\sqrt{3}$, so that we can always apply deconvolution process in Eq~\cite{eq:deconvolution} without using a temperature.
+We also denote $F_X$ as the CDF of $X$.
+Let the
 transition kernel of complete data at sampling step $i$ be $T_i$, and let the transition kernel of
@@
 
-Since the accuracy of our transition kernel is related to the accuracy of our $X_{\rm norm}$
-estimate, we can show that error of our estimation is roughly proportional to the step size of our
-proposer.
+Thus, the accuracy of our transition kernel is related to the accuracy of our $X_{\rm norm}$
+estimate. We now show that error of our estimation is roughly proportional to the step size of our proposer.
 
@@
 Since ${\rm Var}(\Delta') < \epsilon^2 k^2 m $, we can assume the maximum value of $X_{\xi}$ is
-proportional to the standard deviation of $\Delta$, i.e. $\epsilon$. Thus, we can assume
+proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we can assume
 $\zeta=\epsilon C$, where $C$ is a constant factor.

### 2016-05-18 nips_2016.tex

--- 2016-05-18 nips_2016.tex
@@
 corr} > 0)$. Define our estimate of $X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm
-norm} - X_{\rm norm}'$, so that $\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as
+norm} - X_{\rm norm}'$, so that $\Delta' = \Delta + X_{\rm norm} + X_\xi$.
+We assume the standard derivation of $X_{\rm norm}$ is less than $\pi/\sqrt{3}$, so that we can always apply deconvolution process in Eq~\cite{eq:deconvolution} without using a temperature.
+We also denote $F_X$ as
 the CDF of $X$.  Let the transition kernel of complete data at sampling step $i$ be $T_i$, and let

### 2016-05-18 nips_2016.tex
clean the theorm

@@
 We also denote $F_X$ as
-the CDF of $X$.  Let the transition kernel of complete data at sampling step $i$ be $T_i$, and let
-the transition kernel of minibatch data at sampling step $i$ is $T_i'$.  We show that the distance
-of $T_i$ and $T_i'$ is proportional to our estimate error $X_{\xi}$. We also define the stationary
-distribution obtained by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution
+the CDF of $X$.  Let the transition kernel of complete data at sampling step $i$ be $P_i$, and let
+the transition kernel of minibatch data at sampling step $i$ is $P_i'$.
+When $P_i$ is apply on one distribution $D$, it generates a new distribution $D'$, we define the transition operation between $P_i$ and $D$ as ``$\circ$''.
+We show that the distance
+of $P_i$ and $P_i'$ is proportional to our estimate error $X_{\xi}$. We also define the stationary
+distribution obtained by $P_i$ is $\pi$.  We show the by decreasing the step size, the distribution
 of our sampled data converges to $\pi$. We use something called the \emph{total variation distance}
@@
 \begin{lemma}\label{lem:theory1}
-If $|X_\xi| < \zeta$ and $\nabla F_X < k$, then $\|T_i-T_i'\| \le 2\zeta k$.
+If $|X_\xi| < \zeta$ and $\nabla F_X < k$, then $\|P_i-P_i'\| \le 2\zeta k$.
 \end{lemma}
@@
 \begin{theorem}\label{thm:theory3}
-Assume the stationary distribution $\pi$ satisfies $\|T \circ P_0 - \pi\| \leq \eta \|P_0 - \pi\|$,
-where $\eta \in [0, 1)$ for all distribution $P_0$. And assume there exists $\rho_t < 1$ such that
-$\|T'_i(x, \cdot) - \pi_i'\| \leq 2\rho_i$. We can get $\| T_t' \circ T_{t-1}' \circ \cdots \circ T_1' P_0
+Assume the stationary distribution $\pi$ satisfies $\|T \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$,
+where $\eta \in [0, 1)$ for all distribution $D_0$. And assume there exists $\rho_t < 1$ such that
+$\|P'_i(x, \cdot) - \pi_i'\| \leq 2\rho_i$. We can get $\| P_t' \circ P_{t-1}' \circ \cdots \circ P_1' D_0
 - \pi_0 \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $,
@@
 Theorem~\ref{thm:theory3} indicates that the difference between the target and sampled distributions
-consists of two components: $\rho_i$, which is determined by how efficient of $T'_i$, and the
+consists of two components: $\rho_i$, which is determined by how efficient of $P'_i$, and the
 $\alpha_i$s, which are determined by error of our transition kernel. Both are related to the step
-size. If step size is large, the $T'_i$ is very efficient, but the error of $T'_i$ increases.
-Otherwise, if step size is too small, the $T_i'$ is low efficient, but the error of $T'_i$ is low.
+size. If step size is large, the $P'_i$ is very efficient, but the error of $P'_i$ increases.
+Otherwise, if step size is too small, the $P_i'$ is low efficient, but the error of $P'_i$ is low.
 
@@
 Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
-$\|T_i - T'_i\| \leq 2k\zeta$. We write $T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
+$\|P_i - P'_i\| \leq 2k\zeta$. We write $P_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
 q(\theta'\mid \theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$
@@
 \begin{align*}
-\|T_i - T'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
+\|P_i - P'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
 & = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
@@
 
-First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|T \circ P_0 - \pi\| \leq \eta \|P_0
-- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i C$, the norm between the stationary distribution is $\|\pi
+First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0
+- \pi\|$ and $\|P_i-P'_i\| \leq \epsilon_i C$, the norm between the stationary distribution is $\|\pi
 - \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
 
-Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| T_t' \circ T_{t-1}' \circ \cdots
-\circ T_1' P_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
+Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| P_t' \circ P_{t-1}' \circ \cdots
+\circ P_1' D_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
 \alpha_s$, where $\alpha = \frac{\epsilon_i C}{1-\eta}$. Thus, we get:
 \begin{align*}
- \| T_t' \circ \cdots \circ T_1' P_0 - \pi_0 \| &\leq \|T_t' \circ \cdots \circ T_1' P_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
+ \| P_t' \circ \cdots \circ P_1' D_0 - \pi_0 \| &\leq \|P_t' \circ \cdots \circ P_1' D_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
  &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,

### 2016-05-18 nips_2016.tex
update the fig at experiment2
Did some of Section 4 but need to fix typos

@@
 with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
-smaller, thus reducing variance. The flatter posterior results in easier mixing of the samples, and
-as the samples approach a mode, one can decrease the temperature to sharpen the posterior peaks.
+smaller, thus reducing variance. The flatter posterior results in easier mixing.
 
@@
 
-In this section, we explore the convergence properties of our algorithm. Let our true and estimated
-acceptance probabilities be $P_a = g(\Delta) = \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm
-corr} > 0)$. Define our estimate of $X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm
-norm} - X_{\rm norm}'$, so that $\Delta' = \Delta + X_{\rm norm} + X_\xi$.
-We assume the standard derivation of $X_{\rm norm}$ is less than $\pi/\sqrt{3}$, so that we can always apply deconvolution process in Eq~\cite{eq:deconvolution} without using a temperature.
-We also denote $F_X$ as
-the CDF of $X$.  Let the transition kernel of complete data at sampling step $i$ be $P_i$, and let
-the transition kernel of minibatch data at sampling step $i$ is $P_i'$.
-When $P_i$ is apply on one distribution $D$, it generates a new distribution $D'$, we define the transition operation between $P_i$ and $D$ as ``$\circ$''.
-We show that the distance
-of $P_i$ and $P_i'$ is proportional to our estimate error $X_{\xi}$. We also define the stationary
-distribution obtained by $P_i$ is $\pi$.  We show the by decreasing the step size, the distribution
-of our sampled data converges to $\pi$. We use something called the \emph{total variation distance}
-between two kernels, denoted as $\| \cdot \|$.
+In this section, we explore the convergence properties of our MH test. We first introduce some
+notation. Let our true and approximated acceptance probabilities be $P_a = g(\Delta) = \Pr(\Delta +
+X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} > 0)$.  Define $X_{\xi} = \varepsilon - X_{\rm norm}$
+which means, invoking Equations~\ref{eq:relationship} and~\ref{eq:deconvolution}, that $X_\xi$ is
+the random variable characterizing the accuracy of our $X_{\rm norm}$.  Thus, $\Delta' = \Delta +
+X_{\rm norm} + X_\xi$. We assume the standard deviation of $X_{\rm norm}$ is less than
+$\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}. We denote $F_X(x)$ as the CDF of
+$X$. (Note that $X$ is a \emph{scalar} random variable.)
+
+We use some additional terms from standard Markov chain theory~\cite{Meyn2009}. Let the
+\emph{transition kernels} of MCMC on complete and minibatch data at step $i$ be $P_i$ and $P_i'$,
+respectively. When these kernels (e.g., $P_i$) are applied on a probability distribution $D_1$, they
+generate a new distribution $D_2$, which we write as $P_i \circ D_1 = D_2$. Define $\pi$ to be the
+stationary distribution obtained by $P_i$.  Finally, we indicate the \emph{total variation distance}
+between two kernels using $\| \cdot \|$, e.g., $\|P_i-P_i'\|$.
+
+Our first result shows that the distance between $P_i$ and $P_i'$ is proportional to our estimation
+error $X_{\xi}$.
 
 \begin{lemma}\label{lem:theory1}
-If $|X_\xi| < \zeta$ and $\nabla F_X < k$, then $\|P_i-P_i'\| \le 2\zeta k$.
+If $|X_\xi| < \zeta$ and $|\nabla F_X(x)| < \ell$ for all $x$, then $\|P_i-P_i'\| \le 2\zeta \ell$.
 \end{lemma}
@@
 
-Thus, the accuracy of our transition kernel is related to the accuracy of our $X_{\rm norm}$
-estimate. We now show that error of our estimation is roughly proportional to the step size of our proposer.
+Thus, the accuracy of our transition kernel $P_i'$ is related to the quality of our estimated
+$X_{\rm norm}$. We next show that ${\rm Var}(\Delta')$ is roughly proportional to the step size of
+our proposer.
 
@@
 \epsilon$, and the gradient of the log-likelihood is bounded by a constant factor, i.e. $|\nabla
-(\log \Pr(x_i| \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 m$, where
+(\log \Pr(x_i\mid \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 m$, where
 $m$ is the minibatch size.
@@
 
-Since ${\rm Var}(\Delta') < \epsilon^2 k^2 m $, we can assume the maximum value of $X_{\xi}$ is
+Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 m $, we can assume the maximum value of $X_{\xi}$ is
 proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we can assume
@@
 
+We finally introduce Theorem~\ref{thm:theory3} to characterize the convergence of our samples.
+
 \begin{theorem}\label{thm:theory3}
-Assume the stationary distribution $\pi$ satisfies $\|T \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$,
-where $\eta \in [0, 1)$ for all distribution $D_0$. And assume there exists $\rho_t < 1$ such that
-$\|P'_i(x, \cdot) - \pi_i'\| \leq 2\rho_i$. We can get $\| P_t' \circ P_{t-1}' \circ \cdots \circ P_1' D_0
-- \pi_0 \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $,
-where $\alpha_i = \frac{\epsilon_i C }{1-\eta}$.
+Assume $\pi$ satisfies $\|T \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$, where $\eta \in [0, 1)$ for
+all distributions $D_0$. Also assume there exists $0 < \rho_t < 1$ such that $\|P'_i(x, \cdot) -
+\pi_i'\| \leq 2\rho_i$. We can get $\| P_t' \circ P_{t-1}' \circ \cdots \circ P_1' \circ D_0 - \pi
+\| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $, where
+$\alpha_i = \frac{\epsilon_i C }{1-\eta}$.
 \end{theorem}
@@
 \begin{proof}
-See Appendix~\ref{app:theory3}.
+See Appendix~\ref{app:theory3}. {\color{blue} Daniel: Haoyu, please fix the Typos in Theorem 1
+or introduce them (e.g., what is $\pi_i$?}
 \end{proof}
@@
 Theorem~\ref{thm:theory3} indicates that the difference between the target and sampled distributions
-consists of two components: $\rho_i$, which is determined by how efficient of $P'_i$, and the
-$\alpha_i$s, which are determined by error of our transition kernel. Both are related to the step
-size. If step size is large, the $P'_i$ is very efficient, but the error of $P'_i$ increases.
-Otherwise, if step size is too small, the $P_i'$ is low efficient, but the error of $P'_i$ is low.
+consists of two components: $\rho_i$, which is determined by the efficiency of $P'_i$, and the
+$\alpha_i$s, which are determined by the error of our transition kernels. These involve competing
+objectives: the larger the step size size, the more efficient the $P_i'$ but its error increases.
+The reverse happens with a step size too small.
 
@@
     \centering
-    \begin{subfigure}[b]{0.32\textwidth}
-        \includegraphics[width=\textwidth]{./figures/acc_hyc.png}
-        \caption{prediction accuracy}
-        \label{fig:logistic_pa}
-    \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.32\textwidth}
-        \includegraphics[width=\textwidth]{./figures/ll_hyc.png}
-        \caption{loglikelihood}
-        \label{fig:logistic_ll}
-    \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-    %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.32\textwidth}
-        \includegraphics[width=\textwidth]{./figures/roofline_hyc.png}
-        \caption{roofline property}
-        \label{fig:logistic_roof}
-    \end{subfigure}
-    \caption{Experiments of logistic regression}\label{fig:logistic}
+    \includegraphics[width=1\linewidth]{exp2.png}
+    \caption{Experiment results of Logistic regression}
     %\vspace{-10pt}
@@
 
-We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta k$
-for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
+We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta
+\ell$ for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
 \begin{align*}
@@
 &= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| = |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
-&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2k \xi,
+&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2\ell \xi,
 \end{align*}
@@
 smaller absolute value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| <
-2k\zeta$.
+2 \zeta \ell$.
 
 Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
-$\|P_i - P'_i\| \leq 2k\zeta$. We write $P_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
+$\|P_i - P'_i\| \leq 2\zeta \ell$. We write $P_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
 q(\theta'\mid \theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$
@@
 & = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
-& \leq 2 \zeta k \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta'))  d\theta_t \right| d\theta'= 2 \zeta k,
+& \leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta')) d\theta_t \right| d\theta'= 2 \zeta \ell,
 \end{align*}
@@
 \[
-|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t) (\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
+|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t)^T(\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
 \]
@@
 {\rm Var}(\Delta') = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \leq
-\sum_{i=1}^n  {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
+\sum_{i=1}^n  {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 k^2m,
 \]

### 2016-05-18 nips_2016.tex
fix the typo in theoretical part

@@
 generate a new distribution $D_2$, which we write as $P_i \circ D_1 = D_2$. Define $\pi$ to be the
-stationary distribution obtained by $P_i$.  Finally, we indicate the \emph{total variation distance}
+stationary distribution obtained by $P_i$, and $\pi_i'$ to be the stationary distribution obtained by $P_i'$.  Finally, we indicate the \emph{total variation distance}
 between two kernels using $\| \cdot \|$, e.g., $\|P_i-P_i'\|$.
@@
 \epsilon$, and the gradient of the log-likelihood is bounded by a constant factor, i.e. $|\nabla
-(\log \Pr(x_i\mid \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 m$, where
-$m$ is the minibatch size.
+(\log \Pr(x_i\mid \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 (m-\frac{m(m-1)}{N-1})$, where
+$m$ is the minibatch size, and $N$ is the total data size.
 \end{lemma}
@@
 \begin{theorem}\label{thm:theory3}
-Assume $\pi$ satisfies $\|T \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$, where $\eta \in [0, 1)$ for
+Assume $\pi$ satisfies $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$, where $\eta \in [0, 1)$ for
 all distributions $D_0$. Also assume there exists $0 < \rho_t < 1$ such that $\|P'_i(x, \cdot) -
@@
 \theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
-$g_i(\theta) = \log p(x_i\mid \theta)$. Then by Taylor's Theorem:
+$g_i(\theta) = \log p(x_i\mid \theta)$, and $q_i = g_i(\theta') - g_i(\theta_t)$. Then by Taylor's Theorem:
 \[
-|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t)^T(\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
+|q| = |g(\theta') - g(\theta_t)| = |\nabla g(\theta_t)^T(\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
 \]
 Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
-\[
-{\rm Var}(\Delta') = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \leq
-\sum_{i=1}^n  {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 k^2m,
-\]
+\begin{align*}
+{\rm Var}(\Delta') & = {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta_t)\right] = {\rm Var}(\sum_i^m q_i) = m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
+&= m{\rm Var}(q_i) - \frac{m(m-1)}{N-1} {\rm Var}(q_i) \leq 4(m - \frac{m(m-1)}{N-1})k^2 \epsilon^2, \\
+\end{align*}
 as desired.

### 2016-05-18 nips_2016.tex
Shrunk two figures for Section 5.1 (Gaussians)

@@
 \begin{proof}
-See Appendix~\ref{app:theory3}. {\color{blue} Daniel: Haoyu, please fix the Typos in Theorem 1
-or introduce them (e.g., what is $\pi_i$?}
+See Appendix~\ref{app:theory3}.
 \end{proof}
@@
 
-In this section, we use three experiments to reveal the benefits of our mini-batch MH testing. For
-the first experiment, we illustrate that our mini-batch MH testing can obtain the posterior
-distribution which is very close to the ideal one. For the second experiment, we use Logistic
-regression to reveal that our mini-batch sampler is much more efficient than adaptive minibatch
-approach. The third experiment shows that with the help of temperature, our mini-batch MH testing
-can be used to obtain the parameters which are close to MAP.
+We conduct three sets of experiments to explore the benefits of our minibatch MH test and to
+benchmark it with previous work. In Section~\ref{ssec:gaussians}, we show that our test enables
+samples to converge to the posterior distribution of a Gaussian mixture model. In
+Section~\ref{ssec:logistic}, we analyze the efficiency of our sampler on logistic regression. In
+Section~\ref{ssec:nets}, we show that our minibatch MH can obtain good MAP estimates for deep
+learning.
 
 \subsection{Mixture of Gaussians}\label{ssec:gaussians}
-
-{\color{blue}
-Daniel: TODO Have to show that with more data, the posterior peaks (because it is referenced by the
-introduction).
-}
 

### 2016-05-19 nips_2016.tex
More revisions to experiment. Revised one figure, fixed some stuff in appendix on proofs, and slight adjustments to labeling figures

@@
 samples to converge to the posterior distribution of a Gaussian mixture model. In
-Section~\ref{ssec:logistic}, we analyze the efficiency of our sampler on logistic regression. In
-Section~\ref{ssec:nets}, we show that our minibatch MH can obtain good MAP estimates for deep
-learning.
+Section~\ref{ssec:logistic}, we analyze its efficiency on logistic regression. In
+Section~\ref{ssec:nets}, we apply the test for deep learning.
+Appendices~\ref{app:gaussian},~\ref{app:logistic}, and~\ref{app:nnet} contain more detailed
+information on these experiments.
 
@@
 scale of $\Delta'$ deminished when the sampled point is apporching MAP. Thus, the error our
-transition kernel will declines with even constant step size as discribed in Sect.~ 4.
+transition kernel will declines with even constant step size as discribed in Section~\ref{sec:theory}.
 
@@
 classifier for digits 1 vs 7. The dataset consisted of 12007 training data and 1000 testing data.
-Similarly, we use random walk as the proposer.
-We initilize the temperature as 3000, and decrease it by $t_i = t_0 / (i + 1)^{0.5}$, where $i$ represent the $t^{th}$ iteration.
-Thus, the standard derivation of $\Delta'$ is smaller than the threshold $\pi/\sqrt{3}$.
-We compare our mini-batch MH testing with the
-adaptive mini-batch MH testing. Figure 4 shows the log likelihood value and prediction accuracy over
-training process. Since the adaptive MH testing method needs to increase the mini-batch size
-adaptively, we use the cumulative number of data instance each algorithm consumes as the horizontal
-axis. We find our mini-batch MH testing can achieve the similar (or even better log
-likelihood/prediction accuracy) with consuming much fewer data points. Thus, we can assert that our
-mini-batch MH testing is much more efficient than the adaptive one.
+Similarly, we use random walk as the proposer. We initilize the temperature as 3000, and decrease it
+by $t_i = t_0 / (i + 1)^{0.5}$, where $i$ represent the $t^{th}$ iteration. Thus, the standard
+derivation of $\Delta'$ is smaller than the threshold $\pi/\sqrt{3}$.  We compare our mini-batch MH
+testing with the adaptive mini-batch MH testing. Figure 4 shows the log likelihood value and
+prediction accuracy over training process. Since the adaptive MH testing method needs to increase
+the mini-batch size adaptively, we use the cumulative number of data instance each algorithm
+consumes as the horizontal axis. We find our mini-batch MH testing can achieve the similar (or even
+better log likelihood/prediction accuracy) with consuming much fewer data points. Thus, we can
+assert that our mini-batch MH testing is much more efficient than the adaptive one.
 
@@
     \caption{Experiment results of Logistic regression}
+    \label{fig:logistic_fig}
     %\vspace{-10pt}
@@
 $\Delta'$ decrease with large mini-batch, and the error of our mini-batch MH testing is proportional
-to our estimation error of variance of $\Delta'$. When the $var(\Delta')$ decrease to a threshold,
-the error of our mini-batch MH testing can be ignored.
+to our estimation error of variance of $\Delta'$. When the ${\rm Var}(\Delta')$ decrease to a
+threshold, the error of our mini-batch MH testing can be ignored.
 
-{\color{blue}
-Daniel/Haoyu: For additional details, see Appendix~\ref{app:logistic}.
-}
+Figure~\ref{fig:logistic_fig} shows our results ... {\color{blue} Daniel: TODO finish description}
 
@@
 
-In this experiment, we apply about mini-batch MH testing with SGHMC (Stochastic Gradient Hamiltonian Monte Carlo, \cite{sghmc_2014}) proposer to sample the parameter instances from a fully connected neural network.
-We use Higgs data set from University of California Irvine.
-It is a classification problem to distinguish between a signal process which produces Higgs bosons and a background process which does not.
-We build a 4 layers neural network to tackle this problem.
-The first layer is a input layer, the last layer is a softmax layer, the intermediate layers are linear followed by Sigmoid, norm, dropout, starting and ending with linear operation.
+In this experiment, we apply about mini-batch MH testing with SGHMC (Stochastic Gradient Hamiltonian
+Monte Carlo~\cite{sghmc_2014}) proposer to sample the parameter instances from a fully connected
+neural network. We use Higgs data set from University of California Irvine. It is a classification
+problem to distinguish between a signal process which produces Higgs bosons and a background process
+which does not. We build a 4 layers neural network to tackle this problem. The first layer is a
+input layer, the last layer is a softmax layer, the intermediate layers are linear followed by
+Sigmoid, norm, dropout, starting and ending with linear operation.
 
-We choose the Adaptive Gradient Descent method ~\cite{adapGrad} as the comparison baseline.
-Since the frequencies of the features are quite different, we use the same strategy in the Adaptive Gradient to rescale the gradient before apply SGHMC.
-We implement our system within the open-source BIDMach project~\cite{canny2013bidmach}.
+We choose the Adaptive Gradient Descent method~\cite{adapGrad} as the comparison baseline. Since
+the frequencies of the features are quite different, we use the same strategy in the Adaptive
+Gradient to rescale the gradient before apply SGHMC. We implement our system within the open-source
+BIDMach project~\cite{canny2013bidmach}.
 
-Figure~5 illustrates our experiment results.  
+Figure~\ref{fig:nnet_fig} illustrates our experiment results.
+
 \begin{figure}[t]
@@
     \caption{This will hopefully show results for neural networks on BIDMach.}
+    \label{fig:nnet_fig}
     %\vspace{-10pt}
@@
 {\color{blue}
-Daniel: TODO, clean this up.
+Daniel: TODO, clean this up!
 }
@@
 \]
-Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
+Thus, given $\theta$ and $\theta'$, the variance of $\Delta'$ can be written as:
 \begin{align*}
-{\rm Var}(\Delta') & = {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta_t)\right] = {\rm Var}(\sum_i^m q_i) = m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
-&= m{\rm Var}(q_i) - \frac{m(m-1)}{N-1} {\rm Var}(q_i) \leq 4(m - \frac{m(m-1)}{N-1})k^2 \epsilon^2, \\
+{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta_t)\right] = {\rm Var}\left(\sum_i^m q_i\right) \\
+&= m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
+&= m{\rm Var}(q_i) - \frac{m(m-1)}{N-1} {\rm Var}(q_i) \\
+&\leq 4\left(m - \frac{m(m-1)}{N-1}\right)k^2 \epsilon^2,
 \end{align*}

### 2016-05-19 nips_2016.tex
Did lots of work on Section 5.1

@@
 benchmark it with previous work. In Section~\ref{ssec:gaussians}, we show that our test enables
-samples to converge to the posterior distribution of a Gaussian mixture model. In
+samples to converge to the posterior distribution of a heated Gaussian mixture model. In
 Section~\ref{ssec:logistic}, we analyze its efficiency on logistic regression. In
@@
 Appendices~\ref{app:gaussian},~\ref{app:logistic}, and~\ref{app:nnet} contain more detailed
-information on these experiments.
+information on these respective experiments.
 
@@
 
-We start with a simple mixture of Gaussians, which have tied means:
-\[
-\theta_1 \sim N(0, \sigma_1^2); \theta_2 \sim N(0, \sigma_2^2); x_i \sim \frac{1}{2}N(\theta_1, \sigma_x^2) + \frac{1}{2} N(\theta_2, \sigma_x^2)
-\]
-where $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. 200 data points are generated from the
-model with $\theta_1 = 0$ and $\theta_2 = 1$. The theoretical posterior distribution can be seen in
-Fig.~3a. We use random walk as the proposer. We set the step size 0.1 as a constant. We compare
-three scenarios: with our mini-batch MH testing (see Fig.~3b), without our mini-batch MH testing
-(see Fig.~3c), and with the standard MH testing (see Fig.~3d).
+We start with a simple Gaussian mixture model, borrowing an experiment from~\cite{langevin_2011}.
+The parameter is 2-D, $\theta = (\theta_1,\theta_2)$, and the parameter/data generation process is
+\begin{equation}\label{eq:data_generation}
+(\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
+0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
+\end{equation}
+We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta = (0,1)$, we draw 10000
+data points so that the target distribution is $p(\theta)\prod_{i=1}^{10000}p(x_i\mid \theta)$, with
+the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This turns
+out to be a rather sharp posterior, so to make our test applicable, we apply a temperature $T=110$
+to reduce ${\rm std}(\Delta')$. Taking logs, we get the new target as shown in the far left of
+Figure~\ref{fig:gauss_mix_1}.
+
+We run MCMC with our MH test with $m=200$. We also run this using standard full-batch MCMC
+(``standard sampling'') with the standard MH test, and the method from~\cite{cutting_mh_2014}
+(``adaptive sampling''). For the latter, we also use $m=200$ and increment minibatches by that
+amount within an iteration if necessary. The tolerance for making a decision is $\epsilon=0.1$. To
+make comparisons easier, all three use the same random walk proposer with covariance $\Sigma={\rm
+diag}(0.03, 0.03)$. This is a poor proposer, but it is sufficient for our purposes as the quality of
+the samples will be due to the MH test. All methods are run 10000 times to collect 10000 samples.
 
@@
     \caption{
-    {\color{blue}
-    Daniel: I'll fix the caption later. Basically, true log posterior, standard/adaptive procedures,
-    and our method, using MB=200 out of the 10k samples here (also 10k passes over data). All three
-    used the same proposal distribution, a RW with a Gaussian of 0.03. Temperature is at 110.
+    The log posterior contours (temperature 110) and three scatter plots of sampled $\theta$ values.
     }
-    }
+    \label{fig:gauss_mix_1}
     %\vspace{-10pt}
@@
     \caption{
-    {\color{blue}
-    Daniel: I'll fix the caption later. First plot is capped at 1k but the first bin goes up to
-    8000. For the SDs, I skipped the ones that didn't satisfy the requirement (a.k.a. repeating the
-    previous sample). The third is when I re-ran the experiments but with different minibatches, to
-    see how much I would have to repeat. These values were all with one run, though, I might
-    consider just taking several (about five?) and doing error bar stuff.
+    Histograms of minibatch sizes and ${\rm std}(\Delta')$ values, along with ``${\rm std}(\Delta')$
+    violations'' (right).
     }
-    }
+    \label{fig:gauss_mix_2}
     %\vspace{-10pt}
@@
 
-From the figures above, we find that with the help of our mini-batch MH testing we can obtain the
-sampled distribution which are very close to target one. We feature draw our estimation of standard
-derivation of $X_{\rm norm}$ over iterations, see Fig.~3e. We find with even constant step size, the
-standard derivation decreases, because the gradient of posterior distribution declines and make the
-scale of $\Delta'$ deminished when the sampled point is apporching MAP. Thus, the error our
-transition kernel will declines with even constant step size as discribed in Section~\ref{sec:theory}.
+Figure~\ref{fig:gauss_mix_1} shows three scatter plots of the resulting $\theta$ samples for the
+three methods, with darker regions indicating a greater density of points. All three methods obtain
+the same rough form of the posterior, so our MH test can indeed obtain the posterior just as the
+other methods do.
+
+Figure~\ref{fig:gauss_mix_2} (to the left) shows a histogram of the final minibatch sizes used by
+the adaptive subsampling method each iteration (capped at 600 for readability).  Usually, the method
+can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
+in roughly 50 times more computation. For this particular experiment, that happened eight times.  In
+contrast, our method keeps the minibatch size fixed at 200, but requires that ${\rm std}(\Delta') <
+1.2$. Figure~\ref{fig:gauss_mix_2} (in the middle) shows a histogram of the estimated ${\rm
+std}(\Delta')$ values each iteration. Only 75 iterations resulted in ${\rm std}(\Delta') \ge 1.2$
+and for those, we skipped the rest of the iteration (setting $\theta_{t+1}=\theta_t$). To smooth the
+${\rm std}(\Delta')$ values, we use moving average updates. The third plot in
+Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta') \ge 1.2$. For six
+minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with our MH test five times and the
+number of violations averaged. As minibatch sizes increases, the number of violations decreases,
+showing that our test is more applicable.
 
@@
 simple deconvolution process allows us to use a minibatch approximation to the full data tests. We
-experimentally show the benefits our MH test.
-
-Straightforward directions for future work include running additional experiments, with a particular
-focus on investigation of the variance precondition.  More elaborate extensions include combining
-our results with Hamiltonian Monte Carlo methods, providing a recipe for how to use our algorithm
-(following the general framework of~\cite{sgmcmc_2015}), or integrating parallel
-MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts into our work.
-
-{\color{blue}
-Daniel: TODO, clean this up!
-}
+experimentally show the benefits of our test on Gaussian mixtures, logistic regression, and deep
+learning.  Straightforward directions for future work include running more experiments with a
+particular focus on investigation of the variance precondition.  More elaborate extensions include
+combining our results with Hamiltonian Monte Carlo methods, providing a recipe for how to use our
+algorithm (following the framework of~\cite{sgmcmc_2015}), or integrating parallel
+MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
 

### 2016-05-19 nips_2016.tex
Minor change, log likelihood --> likelihood
update author information

@@
   Xinlei Pan \\
-  Department of Bioengineering \\
-  University of California, Berkeley \\
+  Graduate Program in Bioengineering \\
+  UC Berkeley-UC San Francisco \\
   \texttt{xinleipan@berkeley.edu}
@@
 The work of~\cite{conf/uai/MaclaurinA14} presents another minibatch MCMC method which gets the
-\emph{exact} posterior, so long as there exists a (cheap) lower bound on the log likelihood for each
+\emph{exact} posterior, so long as there exists a (cheap) lower bound on the likelihood for each
 data point. Their approach is only a starting point, however, and it can be difficult to derive
@@
 $\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}. We denote $F_X(x)$ as the CDF of
-$X$. (Note that $X$ is a \emph{scalar} random variable.)
+$X$.
 

### 2016-05-19 nips_2016.tex
Modified section 5.2 a bit

@@
   Xinlei Pan \\
-  Graduate Program in Bioengineering \\
-  UC Berkeley-UC San Francisco \\
+  Department of Bioengineering \\
+  University of California, Berkeley \\
   \texttt{xinleipan@berkeley.edu}
@@
 process repeats until a decision.  The downside of this algorithm is the need to keep incrementing
-the mini-batch size; in the worst case, all $N$ data points may be needed in a single iteration.  A
+the minibatch size; in the worst case, all $N$ data points may be needed in a single iteration.  A
 similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties,
@@
 
-In this section, we explore the convergence properties of our MH test. We first introduce some
-notation. Let our true and approximated acceptance probabilities be $P_a = g(\Delta) = \Pr(\Delta +
-X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} > 0)$.  Define $X_{\xi} = \varepsilon - X_{\rm norm}$
-which means, invoking Equations~\ref{eq:relationship} and~\ref{eq:deconvolution}, that $X_\xi$ is
-the random variable characterizing the accuracy of our $X_{\rm norm}$.  Thus, $\Delta' = \Delta +
-X_{\rm norm} + X_\xi$. We assume the standard deviation of $X_{\rm norm}$ is less than
-$\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}. We denote $F_X(x)$ as the CDF of
-$X$.
+In this section, we explore the convergence properties of our MH test. Due to space constraints, all
+our proofs are in Appendix~\ref{app:proofs}.
+
+We first introduce some notation. Let our true and approximated acceptance probabilities be $P_a =
+g(\Delta) = \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} > 0)$.  Define $X_{\xi} =
+\varepsilon - X_{\rm norm}$ which means, invoking Equations~\ref{eq:relationship}
+and~\ref{eq:deconvolution}, that $X_\xi$ is the random variable characterizing the accuracy of our
+$X_{\rm norm}$.  Thus, $\Delta' = \Delta + X_{\rm norm} + X_\xi$. We assume the standard deviation
+of $X_{\rm norm}$ is less than $\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}.
+We denote $F_X(x)$ as the CDF of $X$.
 
@@
 generate a new distribution $D_2$, which we write as $P_i \circ D_1 = D_2$. Define $\pi$ to be the
-stationary distribution obtained by $P_i$, and $\pi_i'$ to be the stationary distribution obtained by $P_i'$.  Finally, we indicate the \emph{total variation distance}
-between two kernels using $\| \cdot \|$, e.g., $\|P_i-P_i'\|$.
+stationary distribution obtained by $P_i$, and $\pi_i'$ to be the stationary distribution obtained
+by $P_i'$.  Finally, we indicate the \emph{total variation distance} between two kernels using $\|
+\cdot \|$, e.g., $\|P_i-P_i'\|$.
 
@@
 
-\begin{proof}
-See Appendix~\ref{app:theory1}.
-\end{proof}
+%% \begin{proof}
+%% See Appendix~\ref{app:theory1}.
+%% \end{proof}
 
@@
 
-\begin{proof}
-See Appendix~\ref{app:theory2}.
-\end{proof}
+%% \begin{proof}
+%% See Appendix~\ref{app:theory2}.
+%% \end{proof}
 
@@
 
-\begin{proof}
-See Appendix~\ref{app:theory3}.
-\end{proof}
+%% \begin{proof}
+%% See Appendix~\ref{app:theory3}.
+%% \end{proof}
 
@@
 can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
-in roughly 50 times more computation. For this particular experiment, that happened eight times.  In
+in more computation (some of it wasted). For this particular experiment, that happened eight times.  In
 contrast, our method keeps the minibatch size fixed at 200, but requires that ${\rm std}(\Delta') <
@@
 Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta') \ge 1.2$. For six
-minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with our MH test five times and the
-number of violations averaged. As minibatch sizes increases, the number of violations decreases,
+minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with our MH test five times and
+averaged the number of violations. As minibatch sizes increases, the number of violations decreases,
 showing that our test is more applicable.
@@
 
-In this experiment, we use logistic regression to classify the MNIST
-dataset~\cite{lecun-mnisthandwrittendigit-2010}. For simplification, we only train the binary
-classifier for digits 1 vs 7. The dataset consisted of 12007 training data and 1000 testing data.
-Similarly, we use random walk as the proposer. We initilize the temperature as 3000, and decrease it
-by $t_i = t_0 / (i + 1)^{0.5}$, where $i$ represent the $t^{th}$ iteration. Thus, the standard
-derivation of $\Delta'$ is smaller than the threshold $\pi/\sqrt{3}$.  We compare our mini-batch MH
-testing with the adaptive mini-batch MH testing. Figure 4 shows the log likelihood value and
-prediction accuracy over training process. Since the adaptive MH testing method needs to increase
-the mini-batch size adaptively, we use the cumulative number of data instance each algorithm
-consumes as the horizontal axis. We find our mini-batch MH testing can achieve the similar (or even
-better log likelihood/prediction accuracy) with consuming much fewer data points. Thus, we can
-assert that our mini-batch MH testing is much more efficient than the adaptive one.
+We next use logistic regression for the binary classification of 1s versus 7s in the MNIST
+dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007 and 1000 training and testing
+points, respectively (we used a random subset of the test data). The propose is again a random walk
+with covariance matrix $0.1I$ for the $784\times 784$ identity matrix $I$. We initialize the
+posterior temperature at $T_0=3000$, and decrease it according to $T_i = T_0 /\sqrt{i + 1}$ for
+iteration $i \ge 1$. We set the minibatch size $m=50$ and compare our minibatch MH test with the
+adaptive minibatch.
+
+Figure~\ref{fig:logistic_fig} shows the prediction accuracy and log likelihood on the test set as a
+function of the cumulative training data points processed. Our test increments the cumulative data
+by a fixed amount per iteration, but the adaptive method may require more data per iteration.  We
+see that our minibatch MH test is more efficient; it has similar or better performance while
+consuming fewer data points.
 
@@
 
-We use this experiment to illustrate another property of our mini-batch MH testing: when the size of
-mini-batch increase to a threshold, the accuracy / log likelihood of sampled data at a fix iteration
-number will not improve with larger mini-batch size, see Fig.~4c. This is because the variance of
-$\Delta'$ decrease with large mini-batch, and the error of our mini-batch MH testing is proportional
+We use this experiment to illustrate another property of our minibatch MH testing: when the size of
+minibatch increase to a threshold, the accuracy / log likelihood of sampled data at a fix iteration
+number will not improve with larger minibatch size, see Fig.~4c. This is because the variance of
+$\Delta'$ decrease with large minibatch, and the error of our minibatch MH testing is proportional
 to our estimation error of variance of $\Delta'$. When the ${\rm Var}(\Delta')$ decrease to a
-threshold, the error of our mini-batch MH testing can be ignored.
-
-Figure~\ref{fig:logistic_fig} shows our results ... {\color{blue} Daniel: TODO finish description}
+threshold, the error of our minibatch MH testing can be ignored.
 
@@
 
-In this experiment, we apply about mini-batch MH testing with SGHMC (Stochastic Gradient Hamiltonian
+In this experiment, we apply about minibatch MH testing with SGHMC (Stochastic Gradient Hamiltonian
 Monte Carlo~\cite{sghmc_2014}) proposer to sample the parameter instances from a fully connected
@@
     accept/reject.
-    \item The mini-batch size is set at 100, so it is one percent of the total data. For the
+    \item The minibatch size is set at 100, so it is one percent of the total data. For the
     adaptive sampling approach, this is the amount that the minibatch size would keep increasing if
@@
 
-We applied our mini-batch Metropolis Hastings algorithm to a Bayesian logistic regression model, comparing with the adaptive sampling method.
-We tested both our method and the adaptive sampling method using a random walk proposer $q(\theta \mid \theta_t) = N (\theta_t, \sigma^2)$.
-Since the random walk proposer does not contain any information about the target distribution, it only relies on the Metropolis-Hastings test to converge to the correct distribution.
-Therefore, random walk proposer is the ideal choice for illustrating our algorithm and comparing with the adaptive sampling method.
+We applied our minibatch Metropolis Hastings algorithm to a Bayesian logistic regression model,
+comparing with the adaptive sampling method. We tested both our method and the adaptive sampling
+method using a random walk proposer $q(\theta \mid \theta_t) = N (\theta_t, \sigma^2)$. Since the
+random walk proposer does not contain any information about the target distribution, it only relies
+on the Metropolis-Hastings test to converge to the correct distribution. Therefore, random walk
+proposer is the ideal choice for illustrating our algorithm and comparing with the adaptive sampling
+method.
 

### 2016-05-19 nips_2016.tex
Finished section 5.2. Now Haoyu and I will finish Section 5.3

@@
     \includegraphics[width=1\linewidth]{exp2.png}
-    \caption{Experiment results of Logistic regression}
+    \caption{Logistic regression performance (accuracy/log likelihood) and minibatch size analysis.}
     \label{fig:logistic_fig}
@@
 
-We use this experiment to illustrate another property of our minibatch MH testing: when the size of
-minibatch increase to a threshold, the accuracy / log likelihood of sampled data at a fix iteration
-number will not improve with larger minibatch size, see Fig.~4c. This is because the variance of
-$\Delta'$ decrease with large minibatch, and the error of our minibatch MH testing is proportional
-to our estimation error of variance of $\Delta'$. When the ${\rm Var}(\Delta')$ decrease to a
-threshold, the error of our minibatch MH testing can be ignored.
+The third plot of Figure~\ref{fig:logistic_fig} shows performance based on minibatch size, when the
+performance is evaluated on the test set after training iterations 700, 1000, and 1500. For all
+three iteration counts, when the minibatch size increases beyond about 10, the performance of the
+sampled data will not improve with a large minibatch size. This makes sense because the error of our
+MH test is proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch
+increases, ${\rm std}(\Delta')$ decreases. Beyond a certain standard deviation threshold, the error
+in our minibatch MH test is negligible.
 
@@
     \centering
-    \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-    %\includegraphics[width=0.5\linewidth]{empty}
-    \caption{This will hopefully show results for neural networks on BIDMach.}
+    \includegraphics[width=1\linewidth]{exp3}
+    \caption{TODO This will hopefully show results for neural networks on BIDMach.}
     \label{fig:nnet_fig}

### 2016-05-19 nips_2016.tex
exp3

@@
 Monte Carlo~\cite{sghmc_2014}) proposer to sample the parameter instances from a fully connected
-neural network. We use Higgs data set from University of California Irvine. It is a classification
-problem to distinguish between a signal process which produces Higgs bosons and a background process
-which does not. We build a 4 layers neural network to tackle this problem. The first layer is a
+neural network. We use Higgs data set from University of California Irvine. We build a 4 layers neural network to tackle this problem. The first layer is a
 input layer, the last layer is a softmax layer, the intermediate layers are linear followed by
@@
 Figure~\ref{fig:nnet_fig} illustrates our experiment results.
+Fig~\ref{fig:nnet_fig}a reveals that SGHMC with our mini batch MH testing achieves high log likelihood than the baseline.
+But by comparing the SGHMC with and without our mini batch testing, the benefits are limited.
+Because we decrease the stepsize quickly, so accept rate of SGHMC is very high that the benefits of testing are not obvious.
 
@@
     \includegraphics[width=1\linewidth]{exp3}
-    \caption{TODO This will hopefully show results for neural networks on BIDMach.}
+    \caption{Log likelihood vs number of passes for optimization of neural network}
     \label{fig:nnet_fig}
-    %\vspace{-10pt}
+    \vspace{-10pt}
 \end{figure}

### 2016-05-19 nips_2016.tex
OK Haoyu and I are calling it a night for now.

--- 2016-05-19 nips_2016.tex
@@
 
-In this experiment, we apply about minibatch MH testing with SGHMC (Stochastic Gradient Hamiltonian
-Monte Carlo~\cite{sghmc_2014}) proposer to sample the parameter instances from a fully connected
-neural network. We use Higgs data set from University of California Irvine. It is a classification
-problem to distinguish between a signal process which produces Higgs bosons and a background process
-which does not. We build a 4 layers neural network to tackle this problem. The first layer is a
-input layer, the last layer is a softmax layer, the intermediate layers are linear followed by
-Sigmoid, norm, dropout, starting and ending with linear operation.
+In this experiment, we apply our minibatch MH test with a Stochastic Gradient Hamiltonian
+Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the parameter instances from a fully connected
+neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}. The network we
+use for this binary classification problem has four layers and is fully connected. The first layer
+is the input, the last layer is the softmax, and the intermediate layers apply the sigmoid
+activation unit.  We also employ dropout and batch normalization~\cite{icml2015_ioffe15}.
 
-We choose the Adaptive Gradient Descent method~\cite{adapGrad} as the comparison baseline. Since
-the frequencies of the features are quite different, we use the same strategy in the Adaptive
-Gradient to rescale the gradient before apply SGHMC. We implement our system within the open-source
-BIDMach project~\cite{canny2013bidmach}.
+For our comparison baseline, we train the same network architecture using the Adaptive Gradient
+Descent method~\cite{adapGrad}. Since the frequencies of the features are quite different, we use
+the same strategy in the Adaptive Gradient to rescale the gradient before apply SGHMC. We implement
+our MH test and Adaptive Gradient in the open-source BIDMach project~\cite{canny2013bidmach}.
 
-Figure~\ref{fig:nnet_fig} illustrates our experiment results.
+Figure~\ref{fig:nnet_fig} illustrates our experiment results and reveals that SGHMC with our
+minibatch MH test achieves higher log likelihood than the baseline. By comparing SGHMC with and
+without our minibatch MH test, we see that our MH test has limited benefit. Because we decrease the
+step size quickly, the acceptance rate of SGHMC is already high, so there is little need to test.
+% Daniel's note: Haoyu said he cannot increase the step size without it violating the variance
+% condition.
 
@@
     \includegraphics[width=1\linewidth]{exp3}
-    \caption{TODO This will hopefully show results for neural networks on BIDMach.}
+    \caption{Log likelihood versus number of passes for optimization of neural network.}
     \label{fig:nnet_fig}
-    %\vspace{-10pt}
+    \vspace{-10pt}
 \end{figure}

### 2016-05-19 nips_2016.tex
Fixed and improved Appendices A and B

@@
     \item Some proof details on results in Section~\ref{sec:theory}.
-    \item Detailed discussion of the experiment from Section~\ref{ssec:gaussians}.
-    \item Detailed discussion of the experiment from Section~\ref{ssec:logistic}.
-    \item Detailed discussion of the experiment from Section~\ref{ssec:nets}.
+    \item More information on Section~\ref{ssec:gaussians}.
+    \item More information on Section~\ref{ssec:logistic}.
+    \item More information on Section~\ref{ssec:nets}.
 \end{itemize}
@@
 & = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
-&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| = |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
-&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2\ell \xi,
+&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| \\
+&= |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
+&= |\nabla F(-\Delta) \xi + o(\xi^2)|\\
+&\leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2\ell \xi,
 \end{align*}
-where the last line uses Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have
-smaller absolute value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| <
-2 \zeta \ell$.
+where we use Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have smaller absolute
+value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| < 2 \zeta \ell$.
 
@@
 \|P_i - P'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
-& = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
-& \leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta')) d\theta_t \right| d\theta'= 2 \zeta \ell,
+&= \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
+&\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta')) d\theta_t \right| d\theta'\\
+&= 2 \zeta \ell,
 \end{align*}
@@
 \begin{align*}
-{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta_t)\right] = {\rm Var}\left(\sum_i^m q_i\right) \\
+{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta_t)\right] = {\rm Var}\left(\sum_{i=1}^m q_i\right) \\
 &= m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
@@
 
-Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| P_t' \circ P_{t-1}' \circ \cdots
-\circ P_1' D_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
-\alpha_s$, where $\alpha = \frac{\epsilon_i C}{1-\eta}$. Thus, we get:
+Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have 
+\[
+\| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t \| \leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t
+\rho_u (1-\alpha_u)\right\} \rho_s \alpha_s,
+\]
+where $\alpha = \frac{\epsilon_i C}{1-\eta}$.  Thus, we get:
 \begin{align*}
- \| P_t' \circ \cdots \circ P_1' D_0 - \pi_0 \| &\leq \|P_t' \circ \cdots \circ P_1' D_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
+ \| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_0 \| &\leq \|P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
  &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
@@
 
-\subsection{Mathematical Assumptions}
-
-We borrow this example from~\cite{langevin_2011}. Our parameter is a 2-D vector $\theta =
+In this section, we go over the math details on the Gaussian mixture model problem borrowed
+from~\cite{langevin_2011}. Our parameter is a 2-D vector $\theta =
 (\theta_1, \theta_2)$, where
@@
 \end{equation}
-Here is the problem statement. Given some number of i.i.d. data points $x_1, x_2, \ldots, x_N$
-generated according to~(\ref{eq:x_points}), determine the posterior distribution\footnote{Note that,
-as is typical with Bayesian analysis, we do not concern ourselves with the denominator of the
-original (non-log) posterior.} of $\theta$:
+Here is the problem statement: given some number of conditionally independent data points $x_1, x_2,
+\ldots, x_N$ generated according to~(\ref{eq:x_points}), determine the posterior
+distribution of $\theta$:
 \begin{equation}\label{eq:log_post}
@@
 where we only use $n \ll N$ samples, but we must scale up the likelihood contribution by $N/n$. If we
-didn't add this scaling factor, then the contribution of the likelihood terms would be weaker. There
-is another thing we need to consider about that, which is the \emph{temperature} of our
-distribution. In general, we will want to add $T > 1$ so that our posterior is
-$p(\theta)((\prod_{i=1}^n p(x_i\mid \theta))^{N/n})^{1/T}$, resulting in the log
-posterior of 
+didn't add this scaling factor, then the contribution of the likelihood terms would be weaker.
+
+One technique we use is adding a \emph{temperature} to our distribution. In general, we will want to
+add $T > 1$ so that our posterior is $p(\theta)((\prod_{i=1}^n p(x_i\mid \theta))^{N/n})^{1/T}$,
+resulting in the log posterior of 
 \begin{equation}\label{eq:log_prior_temp}
@@
 which has the extra $1/T$ to decrease the scale factor. Equation~(\ref{eq:log_prior_temp}) is what
-we will be using for our experiments, because we need to consider warmer distributions.
-{\color{blue} Daniel: not totally confident on this but I'll try.}
-
-Finally, recall earlier that we needed to define $\Delta$. In our context, $\Delta$ is:
-\begin{align}
-\Delta &= \log \left(\frac{f(\theta') q(\theta_t \mid \theta')}{f(\theta_t) q(\theta'\mid \theta_t)} \right) \\
-&= \log p(\theta') - \log p(\theta_t) + \sum_{i=1}^N(\log p(x_i \mid \theta') - \log p(x_i \mid \theta_t)) + \log\left(\frac{q(\theta_t \mid \theta')}{q(\theta' \mid \theta_t)}\right).
-\end{align}
-But again, with \emph{minibatches} of data, we usually do not sum over all $N$ terms and instead sum
-over $n$ terms, but with the extra $(1/T)(N/n)$ scaling factor, from
-Equation~(\ref{eq:log_prior_temp}). We generally denote that approximation as $\Delta'$ in the paper.
-
-\subsection{Experiment Observations}
+we use for our experiments, because warmer distributions help us satisfy our ${\rm std}(\Delta') <
+1.2$ requirement.
 
@@
 
-In this experiment, we compare our method with the one from~\cite{cutting_mh_2014}, hereafter called
-the ``adaptive sampling approach''. We apply them to the Gaussian Mixture model previously
-discussed.
-
-Figure~\ref{fig:contour1} shows simulated contour plots of the posterior based on the $N$ data
-points, for varying $N$, with the temperature set at $T=1$. Note that because we're using all $N$
-points here, the scale factor $N/n=1$. As $N$ increases, the posterior converges to a multimodal
-distribution with modes at $(0,1)$ and $(1,-1)$.
-
+To gain some intuition on what the posterior looks like, Figure~\ref{fig:contour1} shows simulated
+contour plots of the posterior based on varying numbers of data points $N$, with the temperature set
+at $T=1$. Note that because we are using all $N$ points here, the scale factor $N/n=1$. As $N$
+increases, the posterior converges to a multimodal distribution with modes at $(0,1)$ and $(1,-1)$.
 Figure~\ref{fig:contour2} is similar, except this time we fix the number of samples at $N=10000$,
@@
 posterior, one that (weakly) peaks in between the two true modes.
-
-Our ultimate goal is to either (a) get a posterior distribution that covers the posterior well, or
-(b) to get samples that stay trapped at a mode, for a point estimate of $\theta$.  Whether our goal
-is (a) or (b) depends on our objectives and the context of the problem (e.g., big data usually
-implies (b) is OK even if it would not satisfy the average Bayesian).
-
-For both our method and the adaptive sampling method, we use random walk proposals. (Thus, the term
-in $\Delta$ containing the $q$s disappears.)  Random walk proposals are bad, but since the proposal
-is poor, good performance can only be obtained with a strong MH test, hence why we can use this as a
-reasonable starting benchmark.
-
-We test use the following settings:
-
-\begin{itemize}[noitemsep]
-    \item We use $N=10000$ samples of $x_i$.
-    \item The covariance of the Gaussian random walk is ${\rm diag}(0.1, 0.1)$.
-    \item We use 10000 iterations. One iteration involves one $\theta'$ proposal and a test to
-    accept/reject.
-    \item The minibatch size is set at 100, so it is one percent of the total data. For the
-    adaptive sampling approach, this is the amount that the minibatch size would keep increasing if
-    we needed more precision. Our approach, of courses, keeps the size fixed.
-    \item For adaptive sampling, the tolerance for deciding on a test is 0.05. The temperature for
-    that is also set at $T=1$ because that algorithm was not designed to deal with temperatures.
-    \item For our approach, to get an estimate of the standard deviation of $\Delta'$ (remember,
-    $\Delta'$ is a minibatch approximation of $\Delta$) we take five other minibatches (randomly
-    selected from the data) and take the standard deviation of those values.
-\end{itemize}
-
-For our approach, we run using three different temperature values, $T=1$, $T=10$, and $T=100$.
-
-Figure~\ref{fig:scatter} compares scatter plots of the ``Cutting the MH'' test (hereafter called
-``adaptive minibatch'') versus ``Our Method'', where we use three different temperature settings.
-There doesn't seem to be too much noticeable difference in our method for the three different
-temperature setting values, but our approach is a lot different from the adaptive sampling method
-(this is because of rejection rates, more on that later).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1\linewidth]{scatter_v01}
-  \caption{Scatter plots of accepted $\theta$ for (in order) the adaptive sampling approach, and our
-  method (temperatures 1, 10, and 100).}
-  \label{fig:scatter}
-\end{figure}
-
-Before taking a look at our approach, we first examine the adaptive sampling approach. \textbf{That
-approach rejected a lot of samples; out of 10000 iterations, it only accepted 773 times}.
-Figure~\ref{fig:mb_sizes} describes the histograms of the adaptive minibatch's batch sizes. (Note
-that the histogram y-axis values are different.) In many cases, using the initial batch size of 100
-sufficed to make an accept/reject decision, but sometimes, this method needed to use the entire 10k
-data set. Decreasing the tolerance (from its current 0.05 value) would have required larger
-minibatches. Incidentally, it makes sense that the method rejects a lot of samples, because the
-random walk proposal usually provides us with poor points.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.75\linewidth]{adaptive_sampling_sizes_v01.png}
-  \caption{Minibatch sizes for adaptive sampling when it was accepting or rejecting.}
-  \label{fig:mb_sizes}
-\end{figure}
-
-Now we analyze our approach in more detail. Our method accepts far more samples than the adaptive
-sampling method, \textbf{accepting 3285, 3318, and 3332 out of 10k}, for temperature values 1, 10,
-and 100, respectively.
-
-For diagnostics purposes, Figures~\ref{fig:diagnostics1},~\ref{fig:diagnostics2},
-and~\ref{fig:diagnostics3} express histograms (for all three temperature runs) of the $\Delta'$,
-${\rm std}(\Delta')$, and $X_{\rm corr}$ values, respectively.
-
-There are a few immediate observations. First, increasing the temperature size by a factor of 10
-will decrease $\Delta'$ by a factor of 10, which makes sense. That will also directly affect the
-standard deviation estimate of $\Delta'$. Our $X_{\rm corr}$ distribution is symmetric about zero,
-which is also expected (and it also exhibits a decrease in a factor of 10 for an increase
-in temperature).
-
-{\color{blue}
-Daniel: One concern I have is that our $\Delta'$ values do not seem to be ``sufficiently negative''. If our
-$\Delta'$ values are negative (but large in absolute value) that means our MH test will reject more
-often. But then this raises another concern: didn't we say at one point that we \emph{should} be
-getting high acceptance rates, around 50 percent (in part because of the shape of the logistic
-function)? Doesn't that depend on the quality of the proposal distribution? If our method is
-designed to accept a lot regardless of the proposal distribution, I don't see how its performance
-can be considered equal to the adaptive sampling approach, or even a standard MH test. In fact, the
-best observation I can derive from this is that our approach and the adaptive sampling (or any
-standard MH test) will accept the high likelihood cases, but our case will accept the borderline
-ones more often.
-}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1\linewidth]{our_deltas_v01.png}
-  \caption{Our $\Delta'$ values for three different temperature values.}
-  \label{fig:diagnostics1}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1\linewidth]{our_sds_v01.png}
-  \caption{Our ${\rm std}(\Delta')$ values for three different temperature values.}
-  \label{fig:diagnostics2}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1\linewidth]{our_xcorrs_v01.png}
-  \caption{Our $X_{\rm corr}$ values for three different temperature values.}
-  \label{fig:diagnostics3}
-\end{figure}
 

### 2016-05-19 nips_2016.tex
Deleted out of date appendix. Will finish Appendix C and D later today

@@
 
-We applied our minibatch Metropolis Hastings algorithm to a Bayesian logistic regression model,
-comparing with the adaptive sampling method. We tested both our method and the adaptive sampling
-method using a random walk proposer $q(\theta \mid \theta_t) = N (\theta_t, \sigma^2)$. Since the
-random walk proposer does not contain any information about the target distribution, it only relies
-on the Metropolis-Hastings test to converge to the correct distribution. Therefore, random walk
-proposer is the ideal choice for illustrating our algorithm and comparing with the adaptive sampling
-method.
+{\color{blue}
+Daniel: TODO
+}
 

### 2016-05-19 nips_2016.tex
finish append C and D

@@
 \section{Logistic Regression Experiment Details}\label{app:logistic}
+In this section, we go over the details of our logistic regression experiment.
+The feature vector is the pix value of one image, which are normalized between 0 and 1.0.
+For simplification, we only have digits one (i.e.
+output $y = 1$) and seven (i.e.
+output $y = -1$).
+The probability of the $i^{th}$ output $y_i \in \{-1,+1\}$ with the input vector $x_i$ is $p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)}$, where $\theta$ are the parameters.
+We use the uniform prior distribution to represent that we do not have more information about $\theta$.
 
-{\color{blue}
-Daniel: TODO
-}
+The proposer is random walk proposer, which can be modelled as: $\theta' = \theta_i + \mathcal{N}(0, \sigma^2I)$, where $\theta_i$ is the previous parameter vector, $\theta'$ is the proposed parameter vector, and we choose the variance $\sigma^2 = 0.01$.
+We initialize the $\theta$ to be all ones, so that we need sample more to access to the MAP.
+
+For our minibatch MH testing, in order to control the ${\rm Sd} (\Delta') \leq 1.2$, we initialize the temperature $T_0 = 3000$, and decrease it by rule: $T_i =  T_0 /(i + 1)^{3} \mbox{ for } i \leq N/2 \mbox{ and } = T_0 /(i + 1)^{0.2} \mbox{ for } i >N/2 $, where $N=4000$ is total number of iteration.
+If the estimation of ${\rm Sd} (\Delta') > 1.2$, we ignore the current iterartion.
+The our estimation of ${\rm Sd} (\Delta')$ versus iteration number can be found in Fig.~\cite{fig:appendixExp2}.
+We choose our minibatch size equal to 50.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{./figures/appendixExp2}
+  \caption{Addition results for logistic regression experiment}
+  \label{fig:appendixExp2}
+\end{figure}
+
+For adaptive MH testing, all the parameters are the same except they do not have the temperature term.
+The minibatch size of adaptive MH testing is initialized as 50, but it may increase.
+The Fig.~\cite{fig:appendixExp2}.
+shows the histogram of the actual minibatch size in adaptive MH testing.
 
@@
 
-{\color{blue}
-Daniel: TODO
-}
+In this section, we show the details of our neural network experiment.
+We use four-layer fully connected neural network to solve this classification problem.
+The first layer is an input layer, the last layer is softmax layer, each intermediate layer is linear, sigmoid, and batch normailization.
+We use the SGHMC as the proposer for our minibatch MH testing.
+For the baseline, we use a tuned adaptive gradient descent optimizer, whose stepsize changes by $0.01 /(i+1)^{0.4}$.
+All of methods have minibatch size equal to 200.
+The total number of data instance is 1,000,000.
 
+For SGHMC, we use the simplified update equation ~\cite{sghmc_2014}:
+\[\Delta \theta = v \mbox{ , } \Delta v = -\eta \nabla U(\theta) - \alpha v + \mathcal{N}(0, 2(\alpha -\hat{\beta}) \eta), \]
+where $v$ is auxiliary momentum variables, $\theta$ is our parameters, $\nabla U(\theta)$ is the gradient.
+We set the hyperparameter $\eta_i = 0.01 /(i+1)^{0.4}$, and $\alpha=0.1$.
+We use empirical Fisher information $V(\theta)$ to estimate value of $\hat{\beta}$, so that $\hat{\beta}=\frac{1}{2}\eta V(\theta)$ ~\cite{conf/icml/AhnBW12}.
+
+In order to control the ${\rm Sd}(\Delta')$, we initialize temperature as 1000, and set temperature at iteration $i$ is $T_i = 1000/(i+1)^{0.5}$.
+But if temperature is smaller than one, we will just use one as its value.
 % Daniel: here are example LaTeX codes for figures and tables if we want to use them.

### 2016-05-19 nips_2016.tex
Revised Appendics C and D. Only main thing I have in mind is to fix the conflicting temperature T definitions

--- 2016-05-19 nips_2016.tex
@@
 the minibatch size; in the worst case, all $N$ data points may be needed in a single iteration.  A
-similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties,
-but comes at the cost of more computational time.
+similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties
+based on concentration inequalities, but comes at the cost of more computational time.
 
@@
 \begin{itemize}[noitemsep]
-    \item Some proof details on results in Section~\ref{sec:theory}.
+    \item Proofs for material in Section~\ref{sec:theory}.
     \item More information on Section~\ref{ssec:gaussians}.
@@
 
-{\color{blue}
-Daniel: TODO
-}
+In this section, we go over some details of our logistic regression experiment. The feature vector
+for an image consists of its pixel values, normalized between 0 and 1. For simplicity, we only
+consider the binary classification case, so we only use digits one (denoted as output $y = 1$) and
+seven (denoted as $y = -1$). The probability of the $i^{\rm th}$ output $y_i \in \{-1,+1\}$ with the
+input vector $x_i$ is
+\begin{equation}
+p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)},
+\end{equation}
+where $\theta$ is the 784-length parameter vector.
+
+For our experiment, we impose a uniform prior to represent our lack of knowledge about $\theta$.  We
+use a random walk proposer, which can be modeled as $\theta' = \theta_i + \mathcal{N}(0,
+\sigma^2I)$, where $\theta_i$ is the current sample, $\theta'$ is the proposed sample, and we choose
+the variance to be a constant $\sigma^2 = 0.01$ for all components. We initialize $\theta_0$ to be a
+vector of all ones, and set our minibatch size as $m=50$.
+
+For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we
+initialize the temperature $T_0 = 3000$, and decrease it by the rule $T_i =  T_0 /(i + 1)^{3} \mbox{
+for } i \leq N/2 \mbox{ and } = T_0 /(i + 1)^{0.2} \mbox{ for } i >N/2 $, where $N=4000$ is
+the total number of iterations. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
+current iteration. Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
+versus iteration count.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{./figures/appendixExp2}
+  \caption{Additional results for the logistic regression experiment.}
+  \label{fig:appendixExp2}
+\end{figure}
+
+For adaptive MH testing, our experimental settings are the same as with our MH test, except we do
+not impose a temperature. The minibatch size of adaptive MH testing is also initialized as 50, but
+it may increase by that amount each iteration. Figure~\ref{fig:appendixExp2} shows the histogram of
+the actual minibatch size at the end of each iteration in adaptive MH testing.
 
@@
 
-{\color{blue}
-Daniel: TODO
-}
+For our neural network experiment, we use the architecture discussed in Section~\ref{ssec:nets}. We
+use the SGHMC as the proposer for our minibatch MH testing. For the baseline, we use a tuned
+adaptive gradient descent optimizer, whose step size changes by $0.01 /(i+1)^{0.4}$. Both methods
+have minibatch sizes set at $m=200$. There are one million total data instances $x_i$.
+
+For SGHMC, we use the simplified update equations~\cite{sghmc_2014}:
+\begin{equation}
+\Delta \theta = v, \quad \quad\Delta v = -\eta \nabla U(\theta) - \alpha v + \mathcal{N}(0, 2(\alpha
+-\hat{\beta}) \eta),
+\end{equation}
+where $v$ represents auxiliary momentum variables, and $\nabla U(\theta)$ is the gradient of the
+system. We set the hyperparameters to be $\eta_i = 0.01 /(i+1)^{0.4}$, and $\alpha=0.1$. We use the
+empirical Fisher~\cite{conf/icml/AhnBW12} information $V(\theta)$ to estimate the value of
+$\hat{\beta}$, so that $\hat{\beta}=\frac{1}{2}\eta V(\theta)$.
+
+In order to control ${\rm std}(\Delta')$, we initialize the temperature at 1000, and adjust it
+at iteration $i$ according to $T_i = \max\{1, 1000/(i+1)^{0.5}\}$.
 

### 2016-05-19 nips_2016.tex
Standardized the temperature usage. I'll submit this one as the first draft to NIPS

@@
 with covariance matrix $0.1I$ for the $784\times 784$ identity matrix $I$. We initialize the
-posterior temperature at $T_0=3000$, and decrease it according to $T_i = T_0 /\sqrt{i + 1}$ for
-iteration $i \ge 1$. We set the minibatch size $m=50$ and compare our minibatch MH test with the
-adaptive minibatch.
+posterior temperature at $T_0=3000$, and decrease it according to $T_i = T_0 /(i + 1)^{3}$ for $i
+\leq N/2$ and $T_i = T_0 /(i + 1)^{0.2}$ for $i >N/2$, where $N=4000$ is the total number of
+iterations. We set the minibatch size $m=50$ and compare with adaptive sampling.
 
@@
     \label{fig:logistic_fig}
-    %\vspace{-10pt}
+    \vspace{-10pt}
 \end{figure}
@@
 For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we
-initialize the temperature $T_0 = 3000$, and decrease it by the rule $T_i =  T_0 /(i + 1)^{3} \mbox{
-for } i \leq N/2 \mbox{ and } = T_0 /(i + 1)^{0.2} \mbox{ for } i >N/2 $, where $N=4000$ is
-the total number of iterations. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
-current iteration. Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
-versus iteration count.
+initialize the temperature $T_0 = 3000$, and decrease it by the rule $T_i = T_0 /(i + 1)^{3}$ for $i
+\leq N/2$ and $T_i = T_0 /(i + 1)^{0.2}$ for $i >N/2$, where $N=4000$ is the total number of
+iterations. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the current iteration.
+Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values versus iteration
+count.
 

### 2016-05-19 nips_2016.tex
OK Haoyu was using a constant temperature after all. Finally clarified the confusion. Lets use this as the first NIPS submission

@@
 points, respectively (we used a random subset of the test data). The propose is again a random walk
-with covariance matrix $0.1I$ for the $784\times 784$ identity matrix $I$. We initialize the
-posterior temperature at $T_0=3000$, and decrease it according to $T_i = T_0 /(i + 1)^{3}$ for $i
-\leq N/2$ and $T_i = T_0 /(i + 1)^{0.2}$ for $i >N/2$, where $N=4000$ is the total number of
-iterations. We set the minibatch size $m=50$ and compare with adaptive sampling.
+with covariance matrix $0.1I$ for the $784\times 784$ identity matrix $I$. We set the posterior
+temperature at $T=3000$. We set the minibatch size $m=50$ and compare with adaptive sampling.
 
@@
 
-For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we
-initialize the temperature $T_0 = 3000$, and decrease it by the rule $T_i = T_0 /(i + 1)^{3}$ for $i
-\leq N/2$ and $T_i = T_0 /(i + 1)^{0.2}$ for $i >N/2$, where $N=4000$ is the total number of
-iterations. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the current iteration.
-Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values versus iteration
-count.
+For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we use a
+constant temperature $T = 3000$. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
+current iteration.  Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
+versus iteration count.
 

### 2016-05-19 nips_2016.tex
revise a typo

@@
 
-Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 m $, we can assume the maximum value of $X_{\xi}$ is
+Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 (m - \frac{m(m-1)}{N-1}) $, we can assume the maximum value of $X_{\xi}$ is
 proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we can assume

### 2016-05-20 nips_2016.tex
several minor changes. still a few more to go

@@
 posterior inference, the posterior distribution $p(\theta \mid x_1,
-\ldots, x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ of
+\ldots, x_N) \propto p(\theta)\prod_{i=1}^N p(x_i \mid \theta)$ of
 parameter $\theta$ based on conditionally independent data
-$\{x_i\}_{i=1}^n$ is the target distribution which one aims to
-approximate by sampling $\theta$ values. When $n$ is large, it is
+$\{x_i\}_{i=1}^N$ is the target distribution which one aims to
+approximate by sampling $\theta$ values. When $N$ is large, it is
 very expensive to compute the exact posterior because the
@@
 Minibatch updates produce many improvements to the model for each pass
-over the dataset, and have high sample efficiency. They also map very
+over the dataset, and therefore have high sample efficiency. They also map very
 well onto hardware such as GPUs. In contrast, MCMC methods typically
@@
 
-There is another advantage when using our test for Bayesian
-posterior inference. As the dataset size grows, the posterior distribution
-sharpens and often concentrates to a few sharp peaks. Samples from
-this distribution will come from those peaks, and will easily become
-stuck in one of them (see Section~\ref{ssec:gaussians}).  The samples are not exploring
-the posterior parameter %vspace, and are unlikely to concentrate at a
-strong local optimum. They therefore fail to accurately represent a
-sample of the posterior because of inadequate search (i.e.,
-mixing). We argue that it is much more natural to sample the
-distribution at a higher temperature determined by the
-minibatch size rather than the dataset size. This temperature is
-determined in a natural way by the variance prerequisite for our
-test. Starting at a high temperature (small minibatch size) which
-mixes easily, one can reduce the temperature (increasing minibatch
-size) until one reaches the desired target distribution temperature.
-Although this approach also involves changing minibatch size, the
-changes increase minibatch size monotonically, and only a few changes
-are performed in the course of a sampling session.
-Annealing can also be used for MAP and ML estimation because changing temperature preserves
-posterior modes.
+There is another advantage when using our test for Bayesian posterior inference. As the dataset size
+grows, the posterior distribution sharpens and often concentrates to a few sharp peaks. Samples from
+this distribution will come from those peaks, and will easily become stuck in one of them. The
+samples are not exploring the posterior parameter space, and are unlikely to concentrate at a strong
+local optimum. They therefore fail to accurately represent a sample of the posterior because of
+inadequate search (i.e., mixing). We argue that it is much more natural to sample the distribution
+at a higher temperature determined by the minibatch size rather than the dataset size. Starting at a
+high temperature (small minibatch size) which mixes easily, one can reduce the temperature
+(increasing minibatch size) until one reaches the desired target distribution temperature.  Although
+this approach also involves changing minibatch size, the changes increase minibatch size
+monotonically, and only a few changes are performed in the course of a sampling session.  Annealing
+can also be used for MAP and ML estimation because changing temperature preserves posterior modes.
 
@@
 \end{equation}
-where $f(\theta_t)=p(\theta_t \mid x_1,\ldots,x_N)$. The $P_a$ from Equation~\ref{eq:traditional}
-satisfies detailed balance, so if one samples long enough, one will arrive at a stationary
-distribution matching the posterior, though a burn-in period and/or taking every $n$th sample may be
-helpful.
+where $f(\theta_t)=p(\theta_t \mid x_1,\ldots,x_N)$. This $P_a$ satisfies detailed balance, so if
+one samples long enough, one will arrive at a stationary distribution matching the posterior, though
+a burn-in period and/or only using samples at regular intervals is often done in practice.
 
@@
 algorithm starts with a small minibatch of data and tests the hypothesis that the sample $\theta'$
-should be accepted or rejected based on the likelihood ratio. If the test cannot make a decision
-over a certain confidence threshold, then they increase the minibatch size and test again. This
-process repeats until a decision.  The downside of this algorithm is the need to keep incrementing
-the minibatch size; in the worst case, all $N$ data points may be needed in a single iteration.  A
+should be accepted or rejected based on a probability ratio. If the test cannot make a decision
+over a certain confidence threshold, then the minibatch size is increased and the test repeats. This
+process continues until a decision. The downside of this algorithm is the need to keep incrementing
+the minibatch size; in the worst case, all $N$ data points may be needed to decide on $\theta'$. A
 similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties
@@
 HMC~\cite{sghmc_2014,stochastic_thermostats_2014} use a minibatch of data each iteration but require
-controlling the noise from sampling to avoid divergence. Langevin
+controlling the minibatch noise to avoid divergence. Langevin
 Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} is a similar approach which does not use the
@@
 where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}. Thus, $\Delta$ and
-$\Delta'$ are the likelihood ratios in log space. Note that the only difference between $\Delta$ and
+$\Delta'$ are probability ratios in log space. Note that the only difference between $\Delta$ and
 $\Delta'$ is that they use $N$ and $n$ data terms $x_i$, respectively, and that the latter uses an
 $N/n$ scaling term. We assume $n \ll N$ (so $\Delta'$ is substantially faster to compute) and that
-the $n$ data points $x_i$ represent a random minibatch of data.
+the $n$ data points $x_i$ represent a random minibatch without replacement.
 
@@
 \begin{lemma}\label{lem:detailed_balance}
-Let $\Delta$ be defined as in Equation~\ref{eq:deltas}. Any acceptance function $g$ such that
-$g(\Delta) = \exp(\Delta) g(-\Delta )$ satisfies detailed balance. That is, $f(\theta_t)p(\theta'
-\mid \theta_t) = f(\theta')p(\theta_t \mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the
-probability of jumping from $\theta_x$ to $\theta_y$ in our chain.
+An acceptance function $g$ such that $g(\Delta) = \exp(\Delta) g(-\Delta )$ satisfies detailed
+balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) = f(\theta')p(\theta_t \mid \theta')$, where
+$p(\theta_y \mid \theta_x)$ is the probability of jumping from $\theta_x$ to $\theta_y$ in our
+chain.
 \end{lemma}
@@
 as the density of $u$ is one. Thus, the criteria to accept the candidate $\theta'$ is equivalent to
-whether $\Delta > X$. Moreover, $X$ is symmetric about zero, so the acceptance criteria can also be
-expressed as $\Delta + X>0$.
+whether $\Delta > X$. Moreover, the density of $X$ (a standard logistic random variable) is
+symmetric about zero, so the acceptance criteria can also be expressed as $\Delta + X>0$.
 
@@
 
-With large datasets, $\Delta$ is intractable to compute. An intuitive fix is to instead use the
-minibatch version $\Delta'$. The following lemma characterizes the distribution of $\Delta'$.
+With large datasets, $\Delta$ is intractable to compute. An intuitive fix is to replace $\Delta$
+with the minibatch version, $\Delta'$, for our MH test. The following lemma characterizes the distribution of $\Delta'$.
 
 \begin{lemma}\label{lem:gaussian}
-If the minibatch data is chosen randomly and sufficiently large, then the distribution of $\Delta'$ is approximately Gaussian.
+If the minibatch data is chosen randomly without replacement and has sufficiently many elements,
+then the distribution of $\Delta'$ is approximately Gaussian.
 \end{lemma}
@@
 \end{equation}
-for a Gaussian noise term $\varepsilon$.  The extra noise means we can no longer directly perform
-the $\Delta + X$ test, so we need to slightly change our acceptance criteria. Our insight is to
+for a Gaussian noise term $\varepsilon$.  The noise means we cannot just replace $\Delta+X$ with
+$\Delta'+X$ for our test, so we need to slightly change our acceptance criteria. Our insight is to
 decompose $X$ as
@@
     of $\Delta'$, each using a different random minibatch of $n$ data points, but each using the
-    same $\theta_t$ and $\theta'$. In our experiments (see Section~\ref{sec:experiments}), $K$ is
-    small, around 5-10, and $n$ is usually a few hundred for the Central Limit Theorem to apply.
-    Using these $K$ values, we use the empirical standard deviation ${\rm std}(\Delta')$.
+    same $\theta_t$ and $\theta'$. In our experiments (see Section~\ref{sec:experiments}) we make $K$
+    small, around 5-10, for speed purposes, and $n$ is usually a few hundred for the Central Limit
+    Theorem to apply.  Using these $K$ values, we compute the empirical standard deviation ${\rm
+    std}(\Delta')$.
 
@@
     from standard deconvolution techniques. These rely on the well-known fact that the Fourier
-    transform of a convolution is the product of the Fourier transforms. In practice, we pre-compute
-    distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to facilitate the
-    sampling process.
+    transform of a convolution is the product of the individual Fourier transforms. In practice, we
+    pre-compute distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to
+    facilitate the sampling process.
 \end{enumerate}
@@
 \Input{Number of samples $T$, number of $\Delta'$ estimates $K$, minibatch size $m$, pre-computed $X_{\rm
-corr}$ distributions, $\{x_i\}_{i=1}^n$ values from a target distribution, and initial sample $\theta_0$.}
+corr}$ distributions, $\{x_i\}_{i=1}^N$ values from a target distribution, and initial sample $\theta_0$.}
 \Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ for either posterior inference or
-point estimation.}
+MAP/ML estimation.}
 \For{$t=\{1, \ldots, T\}$}{
@@
 deviation of $X_{\rm norm}$ increases, the $X_{\rm corr}$ distribution becomes increasingly unstable
-and ``bumpier'', because the logistic curve has fatter tails than Gaussians. For this reason, we cap
+and ``bumpier'', because the logistic curve has fatter tails than Gaussians. In addition, we cap
 the standard deviation possibilities of $X_{\rm norm}$ at $\approx 1.2$ in our experiments. In our
 experiments, we discretize the possible standard deviations of $X_{\rm norm}$, and we also limit the
-densities to consider the range $[-10,10]$ (as shown in Figure~\ref{eq:deconvolution}).
+densities to consider the range $[-10,10]$ as shown in Figure~\ref{fig:deconvolution}.
 
@@
 with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
-smaller, thus reducing variance. The flatter posterior results in easier mixing.
+smaller in absolute value, thus reducing variance. The flatter posterior results in easier mixing.
 
@@
 \begin{lemma}\label{lem:theory2}
-For one step sampling from $\theta_t$ to $\theta'$, if the jumping step $|\theta_t - \theta'| <
-\epsilon$, and the gradient of the log-likelihood is bounded by a constant factor, i.e. $|\nabla
-(\log \Pr(x_i\mid \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 (m-\frac{m(m-1)}{N-1})$, where
-$m$ is the minibatch size, and $N$ is the total data size.
+For one step sampling from $\theta_t$ to $\theta'$, if the jumping step $\|\theta_t - \theta'\|_2 <
+\epsilon$, and the gradient of the log likelihood is bounded by a constant factor, i.e. $\|\nabla
+(\log \Pr(x_i\mid \theta))\|_2 < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2
+(m-\frac{m(m-1)}{N-1})$, where $m$ is the minibatch size, and $N$ is the total data size.
 \end{lemma}
@@
 
-Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 (m - \frac{m(m-1)}{N-1}) $, we can assume the maximum value of $X_{\xi}$ is
-proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we can assume
-$\zeta=\epsilon C$, where $C$ is a constant factor.
+Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 (m - \frac{m(m-1)}{N-1}) $, we can assume the maximum
+value of $X_{\xi}$ is proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we
+can assume $\zeta=\epsilon C$, where $C$ is a constant factor.
 
@@
 $\alpha_i$s, which are determined by the error of our transition kernels. These involve competing
-objectives: the larger the step size size, the more efficient the $P_i'$ but its error increases.
-The reverse happens with a step size too small.
+objectives: the larger the step size, the more efficient the $P_i'$ but its error increases.  The
+reverse happens with a step size too small.
 
@@
 data points so that the target distribution is $p(\theta)\prod_{i=1}^{10000}p(x_i\mid \theta)$, with
-the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This turns
-out to be a rather sharp posterior, so to make our test applicable, we apply a temperature $T=110$
-to reduce ${\rm std}(\Delta')$. Taking logs, we get the new target as shown in the far left of
-Figure~\ref{fig:gauss_mix_1}.
+the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This
+results in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we apply a
+temperature $T=110$ to reduce ${\rm std}(\Delta')$. Taking logs, we get the target as shown in the
+far left of Figure~\ref{fig:gauss_mix_1}.
 
@@
 diag}(0.03, 0.03)$. This is a poor proposer, but it is sufficient for our purposes as the quality of
-the samples will be due to the MH test. All methods are run 10000 times to collect 10000 samples.
+the samples will be primarily due to the MH test. All methods are run 10000 times to collect 10000
+samples.
 
@@
 
-Figure~\ref{fig:gauss_mix_1} shows three scatter plots of the resulting $\theta$ samples for the
-three methods, with darker regions indicating a greater density of points. All three methods obtain
-the same rough form of the posterior, so our MH test can indeed obtain the posterior just as the
-other methods do.
+Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$ samples for the three
+methods, with darker regions indicating a greater density of points. All three methods obtain the
+same rough form of the posterior, so our MH test can indeed result in the same posterior as the
+other two methods.
 
@@
 can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
-in more computation (some of it wasted). For this particular experiment, that happened eight times.  In
+in lots of wasted computation. For this particular experiment, that happened eight times.  In
 contrast, our method keeps the minibatch size fixed at 200, but requires that ${\rm std}(\Delta') <
 1.2$. Figure~\ref{fig:gauss_mix_2} (in the middle) shows a histogram of the estimated ${\rm
-std}(\Delta')$ values each iteration. Only 75 iterations resulted in ${\rm std}(\Delta') \ge 1.2$
-and for those, we skipped the rest of the iteration (setting $\theta_{t+1}=\theta_t$). To smooth the
-${\rm std}(\Delta')$ values, we use moving average updates. The third plot in
-Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta') \ge 1.2$. For six
-minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with our MH test five times and
-averaged the number of violations. As minibatch sizes increases, the number of violations decreases,
-showing that our test is more applicable.
+std}(\Delta')$ values each iteration. Only 75 iterations (out of 10000) resulted in ${\rm
+std}(\Delta') \ge 1.2$ and for these, we skipped the rest of the iteration (setting
+$\theta_{t+1}=\theta_t$). To smooth the ${\rm std}(\Delta')$ values, we use moving average updates.
+The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
+\ge 1.2$ based on minibatch sizes. For six sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
+our MH test five times and averaged the number of ``violations.'' As the size increases, the
+number of violations decreases, as expected.
 
@@
 dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007 and 1000 training and testing
-points, respectively (we used a random subset of the test data). The propose is again a random walk
+points, respectively (we used a random subset of the test data). The proposer is again a random walk
 with covariance matrix $0.1I$ for the $784\times 784$ identity matrix $I$. We set the posterior
-temperature at $T=3000$. We set the minibatch size $m=50$ and compare with adaptive sampling.
+temperature at $T=3000$. We set the minibatch size $m=50$ and compare with adaptive sampling with
+tolerance 0.02.
 
@@
 
-The third plot of Figure~\ref{fig:logistic_fig} shows performance based on minibatch size, when the
-performance is evaluated on the test set after training iterations 700, 1000, and 1500. For all
-three iteration counts, when the minibatch size increases beyond about 10, the performance of the
-sampled data will not improve with a large minibatch size. This makes sense because the error of our
-MH test is proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch
-increases, ${\rm std}(\Delta')$ decreases. Beyond a certain standard deviation threshold, the error
-in our minibatch MH test is negligible.
+The third plot of Figure~\ref{fig:logistic_fig} shows performance based on minibatch size. The
+parameter is evaluated on the test set after training iterations 700, 1000, and 1500. For those
+three cases, when the minibatch size increases beyond about 10, the performance of the sampled data
+will not improve with a large minibatch size. This makes sense because the error of our MH test is
+proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch increases, ${\rm
+std}(\Delta')$ decreases. Beyond a certain standard deviation (or minibatch) threshold, the error in
+our minibatch MH test is negligible.
 
@@
 In this experiment, we apply our minibatch MH test with a Stochastic Gradient Hamiltonian
-Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the parameter instances from a fully connected
-neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}. The network we
-use for this binary classification problem has four layers and is fully connected. The first layer
-is the input, the last layer is the softmax, and the intermediate layers apply the sigmoid
-activation unit.  We also employ dropout and batch normalization~\cite{icml2015_ioffe15}.
+Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the complete parameters for a fully
+connected neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}, which
+is a binary classification task with one million instances, each of which are 28-dimensional. The
+network we use has four layers: the first is the input, the last is the softmax, and the
+intermediate ones apply the sigmoid activation unit.  We also employ dropout and batch
+normalization~\cite{icml2015_ioffe15}.
 
@@
 Descent method~\cite{adapGrad}. Since the frequencies of the features are quite different, we use
-the same strategy in the Adaptive Gradient to rescale the gradient before apply SGHMC. We implement
-our MH test and Adaptive Gradient in the open-source BIDMach project~\cite{canny2013bidmach}.
+the same strategy as in Adaptive Gradient to rescale the gradient before applying SGHMC. We
+implement our MH test and Adaptive Gradient in the open-source BIDMach
+project~\cite{canny2013bidmach}.
 

### 2016-05-20 nips_2016.tex
Changed text to say it is on a log scale

@@
 Figure~\ref{fig:gauss_mix_2} (to the left) shows a histogram of the final minibatch sizes used by
-the adaptive subsampling method each iteration (capped at 600 for readability).  Usually, the method
+the adaptive subsampling method each iteration (on a log scale for readability).  Usually, the method
 can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting

### 2016-05-20 nips_2016.tex
Improved the theoretical explanation a tiny bit.

@@
 
-We use some additional terms from standard Markov chain theory~\cite{Meyn2009}. Let the
-\emph{transition kernels} of MCMC on complete and minibatch data at step $i$ be $P_i$ and $P_i'$,
-respectively. When these kernels (e.g., $P_i$) are applied on a probability distribution $D_1$, they
-generate a new distribution $D_2$, which we write as $P_i \circ D_1 = D_2$. Define $\pi$ to be the
-stationary distribution obtained by $P_i$, and $\pi_i'$ to be the stationary distribution obtained
-by $P_i'$.  Finally, we indicate the \emph{total variation distance} between two kernels using $\|
-\cdot \|$, e.g., $\|P_i-P_i'\|$.
+We use some terms from standard Markov chain theory~\cite{Meyn2009}. Let the \emph{transition
+kernels} of MCMC on complete and minibatch data at step $i$ be $P_i$ and $P_i'$, respectively. When
+these kernels (e.g., $P_i$) are applied on a probability distribution $D_1$, they generate a new
+distribution $D_2$, which we write as $P_i \circ D_1 = D_2$. Define $\pi$ to be the stationary
+distribution obtained by $P_i$, and $\pi_i'$ to be the stationary distribution obtained by $P_i'$
+($\pi'$ needs a subscript in this case).  Finally, we indicate the \emph{total variation distance}
+between two kernels using $\| \cdot \|$, e.g., $\|P_i-P_i'\|$.
 
@@
 value of $X_{\xi}$ is proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we
-can assume $\zeta=\epsilon C$, where $C$ is a constant factor.
-
-We finally introduce Theorem~\ref{thm:theory3} to characterize the convergence of our samples.
+can assume $\zeta=\epsilon C$, where $C$ is a constant factor. We finally introduce
+Theorem~\ref{thm:theory3} to characterize the convergence of our samples.
 
@@
 \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $, where
-$\alpha_i = \frac{\epsilon_i C }{1-\eta}$.
+$\alpha_i = \frac{\epsilon_i C }{1-\eta}$ and $\epsilon_i$ is the step size at iteration $i$.
 \end{theorem}

### 2016-05-20 nips_2016.tex
updated lots of the theory stuff, typos etc

@@
 With large datasets, $\Delta$ is intractable to compute. An intuitive fix is to replace $\Delta$
-with the minibatch version, $\Delta'$, for our MH test. The following lemma characterizes the distribution of $\Delta'$.
+with the minibatch version, $\Delta'$, for our MH test. The following lemma characterizes the
+distribution of $\Delta'$.
 
@@
 comes only from the minibatch. Since the minibatch is chosen randomly, the distribution of
-$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t))$ converges to a Gaussian as $n
-\to \infty$ by the Central Limit Theorem.
+$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t))$ converges to a Gaussian for
+sufficiently large $N$ by the Central Limit Theorem.
 \end{proof}
@@
     standard logistic distribution. We use three standard deviation values of $X_{\rm norm}$. The
-    two red curves convolve to form the logistic, etc. The $y$-axis is capped at 0.5 for
-    readability. This figure must be viewed in color.
+    two red (respectively, blue, green) curves form the logistic. The $y$-axis is capped at 0.5.
+    This figure must be viewed in color.
     }
@@
     same $\theta_t$ and $\theta'$. In our experiments (see Section~\ref{sec:experiments}) we make $K$
-    small, around 5-10, for speed purposes, and $n$ is usually a few hundred for the Central Limit
-    Theorem to apply.  Using these $K$ values, we compute the empirical standard deviation ${\rm
-    std}(\Delta')$.
+    small, around 5-10, for speed purposes, and $n$ is usually 50-200.  Using these $K$ values, we
+    compute the empirical standard deviation ${\rm std}(\Delta')$.
 
@@
 which is $\pi^2/3\approx 3.29$ is an upper bound on the variances of $X_{\rm norm}$. Therefore, the
-test cannot work if the variance of $X_{\rm norm}$ is too large. Since the standard deviation of $X$
-is about $\sqrt{3.29}\approx 1.81$, this bounds ${\rm std}(X_{\rm norm})$ (as well as ${\rm
+test cannot work if the variance of $X_{\rm norm}$ is too large. The standard deviation of $X$
+is about $\sqrt{3.29}\approx 1.81$, so this bounds ${\rm std}(X_{\rm norm})$ (as well as ${\rm
 std}(\Delta')$). From the discussion above, we bound the standard deviation at 1.2 so that $X_{\rm
@@
 ($\pi'$ needs a subscript in this case).  Finally, we indicate the \emph{total variation distance}
-between two kernels using $\| \cdot \|$, e.g., $\|P_i-P_i'\|$.
+between two kernels using $\| \cdot \|$.
 
-Our first result shows that the distance between $P_i$ and $P_i'$ is proportional to our estimation
-error $X_{\xi}$.
+Our first result shows that the distance $\|P_i\circ D - P_i'\circ D\| \le 4\zeta \ell$ for all
+distributions $D$.
+
+% Our first result shows that the distance between $P_i$ and $P_i'$ is proportional to our estimation
+% error $X_{\xi}$.
 
 \begin{lemma}\label{lem:theory1}
-If $|X_\xi| < \zeta$ and $|\nabla F_X(x)| < \ell$ for all $x$, then $\|P_i-P_i'\| \le 2\zeta \ell$.
+If $|X_\xi| < \zeta$ and $|\nabla F_X(x)| < \ell$ for all $x$, then $\|P_i\circ D-P_i'\circ D\| \le
+4\zeta \ell \;\;\forall D$.
 \end{lemma}
@@
 Assume $\pi$ satisfies $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$, where $\eta \in [0, 1)$ for
-all distributions $D_0$. Also assume there exists $0 < \rho_t < 1$ such that $\|P'_i(x, \cdot) -
-\pi_i'\| \leq 2\rho_i$. We can get $\| P_t' \circ P_{t-1}' \circ \cdots \circ P_1' \circ D_0 - \pi
+all distributions $D_0$. Also assume there exists $0 < \rho_t < 1$ such that $\|P'_i \circ D -
+\pi_i'\| \leq 2\rho_i \;\; \forall D$. We can get $\| P_t' \circ P_{t-1}' \circ \cdots \circ P_1' \circ D_0 - \pi
 \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $, where
@@
 
-We run MCMC with our MH test with $m=200$. We also run this using standard full-batch MCMC
-(``standard sampling'') with the standard MH test, and the method from~\cite{cutting_mh_2014}
-(``adaptive sampling''). For the latter, we also use $m=200$ and increment minibatches by that
-amount within an iteration if necessary. The tolerance for making a decision is $\epsilon=0.1$. To
-make comparisons easier, all three use the same random walk proposer with covariance $\Sigma={\rm
-diag}(0.03, 0.03)$. This is a poor proposer, but it is sufficient for our purposes as the quality of
-the samples will be primarily due to the MH test. All methods are run 10000 times to collect 10000
-samples.
+We run MCMC with our MH test using minibatch size $m=200$. We also run this using standard
+full-batch MCMC (``standard sampling'') with the standard test from Equation~\ref{eq:traditional}, and the method
+from~\cite{cutting_mh_2014} (``adaptive sampling''). For the latter, we also use $m=200$ and
+increment minibatches by that amount within an iteration if necessary. The tolerance for making a
+decision is $\epsilon=0.1$. To make comparisons easier, all three use the same random walk proposer
+with covariance $\Sigma={\rm diag}(0.03, 0.03)$. This is a poor proposer, but it is sufficient for
+our purposes as the quality of the samples will be primarily due to the MH test. All methods are run
+10000 times to collect 10000 samples.
 
@@
 intermediate ones apply the sigmoid activation unit.  We also employ dropout and batch
-normalization~\cite{icml2015_ioffe15}.
+normalization~\cite{icml2015_ioffe15}.  In order to control ${\rm std}(\Delta')$, we initialize the
+temperature at 1000, and adjust it at iteration $i$ according to $T_i = \max\{1,
+1000/(i+1)^{0.5}\}$.
 
@@
 the same strategy as in Adaptive Gradient to rescale the gradient before applying SGHMC. We
-implement our MH test and Adaptive Gradient in the open-source BIDMach
-project~\cite{canny2013bidmach}.
+implement our MH test and SGHMCin the open-source BIDMach project~\cite{canny2013bidmach}.
 
@@
 
-Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
-$\|P_i - P'_i\| \leq 2\zeta \ell$. We write $P_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
+Next, we show the distance of new distributions obtained by applying $P_i$ and $P_i'$ on an
+arbitrary distribution $D$ once is bounded if we use the same proposer, i.e.  $\|P_i\circ D -
+P'_i\circ D\| \leq 4\zeta \ell \;\; \forall D$. We write $P_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
 q(\theta'\mid \theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$
-is the Dirac delta function. The total variation distance between two proposal kernels is
+is the Dirac delta function. The total variation distance between two generated distributions is
 \begin{align*}
-\|P_i - P'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
-&= \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
-&\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta')) d\theta_t \right| d\theta'\\
-&= 2 \zeta \ell,
+\|P_i\circ D - P'_i\circ D\| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) dD(\theta_t) \right| d\theta' \\
+&= \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) dD(\theta_t) \right|  d\theta'\\ 
+&\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + \delta_D(\theta_t - \theta')) dD(\theta_t) \right| d\theta'\\
+&\leq 2 \zeta \ell \int_{\theta'} ( D'(\theta') + D(\theta') ) d\theta'\\
+&= 4 \zeta \ell,
 \end{align*}
-as desired.
+where $D'$ is the distribution generated if we always accept the proposed point. (In the last line,
+we integrate two probability density functions, resulting in a value of two.) This matches our
+desired bound.
 
@@
 First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0
-- \pi\|$ and $\|P_i-P'_i\| \leq \epsilon_i C$, the norm between the stationary distribution is $\|\pi
-- \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
+- \pi\|$ and $\|P_i\circ D-P'_i\circ D\| \leq \epsilon_i C$ for all $D$. The norm between the
+stationary distributions is $\|\pi - \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
 
@@
 empirical Fisher~\cite{conf/icml/AhnBW12} information $V(\theta)$ to estimate the value of
-$\hat{\beta}$, so that $\hat{\beta}=\frac{1}{2}\eta V(\theta)$.
-
-In order to control ${\rm std}(\Delta')$, we initialize the temperature at 1000, and adjust it
-at iteration $i$ according to $T_i = \max\{1, 1000/(i+1)^{0.5}\}$.
+$\hat{\beta}$, so that $\hat{\beta}=\frac{1}{2}\eta V(\theta)$. In order to control ${\rm
+std}(\Delta')$, we initialize the temperature at 1000, and adjust it at iteration $i$ according to
+$T_i = \max\{1, 1000/(i+1)^{0.5}\}$.
 

### 2016-05-23 nips_2016_jfc.tex
added changes from JFC

--- 2016-05-20 nips_2016.tex
@@
 machine learning. We are particularly interested in their application
-to modeling very large datasets, where it is impractical to
-perform Metropolis-Hastings tests on the full data. To do this, we
-describe a novel acceptance test which is applied to minibatches of
-data. Previous work on minibatch Metropolis-Hastings use adaptive
-batch sizes, which may use far more than one minibatch of data and therefore
-yield few samples.  Here we describe a simpler test on fixed-sized
-minibatches. Instead of adaptive batch sizes, our test requires a
-precondition on the variance of the log proposal probability
-ratio. Given a measurement of this variance, a minibatch size can be
-chosen to support the test in one step, which yields a number of
-samples proportional to the batch/minibatch ratio. Alternatively, the
-variance condition can be used to adjust the temperature of the
-sampling distribution to a ``natural'' value to provide regular
-samples at a given minibatch size. The resulting proposal/test is only
-slightly more complex than a simple SGD update. In this paper we
-derive the test, discuss its implementation, and present several
-experiments.
+to modeling very large datasets where it is impractical to perform
+Metropolis-Hastings tests on the full data. We describe a novel
+acceptance test which can be applied to fixed-size minibatches of
+data. Previous works on reducing sample count for the
+Metropolis-Hastings test require a data-dependent number of samples
+which is not fixed a-priori.  They do not work on bounded-size samples
+and therefore do not provide the typical practical benefits of
+minibatch methods.  Our novel test requires an upper bound on the
+variance of the log probability, but given this condition it can be
+applied to fixed-size minibatches of data. We show via experiments
+that if this condition is not met, that the earlier methods consume an
+large number of samples.  Our test is extremely simple and fast,
+bringing the benefits of minibatch methods to MCMC inference.
+Alternatively, the variance condition can be used to adjust the
+temperature of the sampling distribution to a ``natural'' value to
+provide regular samples at a given minibatch size. The resulting
+proposal/test is only slightly more complex than a simple SGD
+update. In this paper we derive the test, discuss its implementation,
+and present several experiments.
 \end{abstract}
@@
 test~\cite{Metropolis1953,hastings70}. For example, for Bayesian
-posterior inference, the posterior distribution $p(\theta \mid x_1,
-\ldots, x_N) \propto p(\theta)\prod_{i=1}^N p(x_i \mid \theta)$ of
-parameter $\theta$ based on conditionally independent data
-$\{x_i\}_{i=1}^N$ is the target distribution which one aims to
-approximate by sampling $\theta$ values. When $N$ is large, it is
-very expensive to compute the exact posterior because the
+posterior inference, the sampling target distribution over the
+parameters $\theta$ is the posterior distribution $p(\theta \mid x_1,
+\ldots, x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ of
+based on conditionally independent data $\{x_i\}_{i=1}^n$. When $n$ is
+large, it is very expensive to compute the exact posterior because the
 likelihood must be evaluated on every $x_i$.
@@
 Minibatch updates produce many improvements to the model for each pass
-over the dataset, and therefore have high sample efficiency. They also map very
+over the dataset, and have high sample efficiency. They also map very
 well onto hardware such as GPUs. In contrast, MCMC methods typically
-require calculations over the full dataset to produce a new sample. A
-recent result from~\cite{cutting_mh_2014} attempted to bridge the gap
-between minibatch optimization and MCMC by proposing an \emph{adaptive
-  minibatch} method. Their method uses a subset of the $N$ data points
-for the MH test each iteration. However the number of points is not
-known a priori, and is adjusted dynamically during every test. This
-undermines many of the practical advantages of minibatch calculations: 
-GPU kernels work best for certain minibatch sizes, memory blocks
-can be reused when minibatch size is fixed, etc.
+require calculations over the full dataset to produce a new sample.
+Recent results
+\cite{cutting_mh_2014,conf/uai/MaclaurinA14,icml2014c1_bardenet14}
+attempted to bridge the gap between minibatch optimization and MCMC by
+data-adaptive subsampling. These methods uses a
+subset of the $N$ data points for the MH test each iteration. However
+the number of points is not known a priori, and is adjusted
+dynamically during every test. This undermines many of the practical
+advantages of minibatch calculations: GPU kernels work best for
+certain minibatch sizes, memory blocks can be reused when minibatch
+size is fixed, etc.
 
@@
 
-There is another advantage when using our test for Bayesian posterior inference. As the dataset size
-grows, the posterior distribution sharpens and often concentrates to a few sharp peaks. Samples from
-this distribution will come from those peaks, and will easily become stuck in one of them. The
-samples are not exploring the posterior parameter space, and are unlikely to concentrate at a strong
-local optimum. They therefore fail to accurately represent a sample of the posterior because of
-inadequate search (i.e., mixing). We argue that it is much more natural to sample the distribution
-at a higher temperature determined by the minibatch size rather than the dataset size. Starting at a
-high temperature (small minibatch size) which mixes easily, one can reduce the temperature
-(increasing minibatch size) until one reaches the desired target distribution temperature.  Although
-this approach also involves changing minibatch size, the changes increase minibatch size
-monotonically, and only a few changes are performed in the course of a sampling session.  Annealing
-can also be used for MAP and ML estimation because changing temperature preserves posterior modes.
+There is another advantage when using our test for Bayesian
+posterior inference. As the dataset size grows, the posterior distribution
+sharpens and often concentrates to a few sharp peaks. Samples from
+this distribution will come from those peaks, and will easily become
+stuck in one of them (see Section~\ref{ssec:gaussians}).  The samples are not exploring
+the posterior parameter space, and are unlikely to concentrate at a
+strong local optimum. They therefore fail to accurately represent a
+sample of the posterior because of inadequate search (i.e.,
+mixing). We argue that it is much more natural to sample the
+distribution at a higher temperature determined by the
+minibatch size rather than the dataset size. This temperature is
+determined in a natural way by the variance prerequisite for our
+test. Starting at a high temperature (small minibatch size) which
+mixes easily, one can reduce the temperature (increasing minibatch
+size) until one reaches the desired target distribution temperature.
+Although this approach also involves changing minibatch size, the
+changes increase minibatch size monotonically, and only a few changes
+are performed in the course of a sampling session.
+Annealing can also be used for MAP and ML estimation because changing temperature preserves
+posterior modes.
 
@@
 \begin{enumerate}[noitemsep]
-    \item We develop a new minibatch acceptance test which satisfies detailed balance.
-    \item We compare performance of our new test and the adaptive sampling method.
+\item We develop a new acceptance test which works on fixed-size data samples, given
+  an auxiliary condition. 
+    \item We compare performance of our new test and adaptive sampling methods.
     \item We experiment using the approach for posterior estimation.
@@
 \end{equation}
-where $f(\theta_t)=p(\theta_t \mid x_1,\ldots,x_N)$. This $P_a$ satisfies detailed balance, so if
-one samples long enough, one will arrive at a stationary distribution matching the posterior, though
-a burn-in period and/or only using samples at regular intervals is often done in practice.
+where $f(\theta_t)=p(\theta_t \mid x_1,\ldots,x_N)$. The $P_a$ from Equation~\ref{eq:traditional}
+satisfies detailed balance, so if one samples long enough, one will arrive at a stationary
+distribution matching the posterior, though a burn-in period and/or taking every $n$th sample may be
+helpful.
 
-Unfortunately, computing $f$ requires the use of all $N$ training data points. Moreover, it is
-difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
-balance. To reconcile these competing objectives,~\cite{cutting_mh_2014} proposes an adaptive
-minibatch MCMC algorithm which uses a sequential hypothesis test. During each iteration, the
-algorithm starts with a small minibatch of data and tests the hypothesis that the sample $\theta'$
-should be accepted or rejected based on a probability ratio. If the test cannot make a decision
-over a certain confidence threshold, then the minibatch size is increased and the test repeats. This
-process continues until a decision. The downside of this algorithm is the need to keep incrementing
-the minibatch size; in the worst case, all $N$ data points may be needed to decide on $\theta'$. A
-similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties
-based on concentration inequalities, but comes at the cost of more computational time.
+Unfortunately, computing $f$ requires the use of all $N$ training data
+points. Moreover, it is difficult to design tests using substantially
+fewer than $N$ points that also satisfy detailed balance. To reconcile
+these competing objectives,~\cite{cutting_mh_2014} proposes an
+adaptive minibatch MCMC algorithm which uses a sequential hypothesis
+test. During each iteration, the algorithm starts with a small
+minibatch of data and adds additional samples until the test either
+succeeds or fails. in the worst case, all $N$ data points may be
+needed in a single iteration.  A similar approach is explored in
+~\cite{icml2014c1_bardenet14} but as the authors there note, the
+method is ``of limited usefulness'' when the log probability variance
+is high, and the method may consume all or most of the samples. The
+work of~\cite{conf/uai/MaclaurinA14} presents another approach to
+reducing sample size for MH testing. In their case, auxiliary
+variables are introduced which allow exact sampling from the
+posterior. Their method assumes a lower bound on the posterior
+probability. Unfortunately, in the most challenging situations (high
+variance in the log probability), this lower bound, if it exists at
+all, will be much smaller than the mean probability. As a consequence,
+this method will consume most of the data samples.
 
-The work of~\cite{conf/uai/MaclaurinA14} presents another minibatch MCMC method which gets the
-\emph{exact} posterior, so long as there exists a (cheap) lower bound on the likelihood for each
-data point. Their approach is only a starting point, however, and it can be difficult to derive
-these lower bounds in general cases. The extra auxiliary variables they introduce also hurt mixing.
-
-There is a separate line of MCMC work on simulating the physics of a probability distribution. By
-viewing random variables as particles in a system, one can apply Hamiltonian Monte Carlo
-(HMC)~\cite{mcmc_hamiltonian_2010} methods which generate the best case scenario for proposals: high
-quality \emph{and} distant.  HMC methods require a full gradient computation of the posterior, which
-means they also face the problem of computing the likelihood at each point. Stochastic versions of
-HMC~\cite{sghmc_2014,stochastic_thermostats_2014} use a minibatch of data each iteration but require
-controlling the minibatch noise to avoid divergence. Langevin
-Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} is a similar approach which does not use the
-momentum terms in HMC.  Minibatch SGHMC and Langevin Dynamics are orthogonal to our work and are
-other ways of attempting to solve the same problem of MCMC methods on big data. We can combine our
-methods with these results, as we demonstrate in Section~\ref{ssec:nets}.
+MCMC methods can use physics-inspired state-space
+extensions. By viewing random variables as particles in a physical
+system and adding virtual momentum variables, one realizes the
+Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} approach.
+These methods generate high-efficiency proposals with large steps and
+high acceptance rates. Traditional HMC methods require a full gradient
+computation for proposals, but recent extensions
+HMC~\cite{sghmc_2014,stochastic_thermostats_2014} use minibatches of
+data. Langevin Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} is a
+similar approach which does not use the momentum terms in HMC, but
+typically produces smaller steps and random-walk behavior. In both
+cases, these methods avoided the expense of MH acceptance tests which
+are much larger than the minibatch proposals. But because of this they
+are forced to limit step size so that their approximate proposals
+generate the true distribution in the limit. With our fast test,
+these methods can be made more efficient by using larger
+step sizes without compromising accuracy.
+We demonstrate this in Section~\ref{ssec:nets}.
 
@@
 
-For our new MH test, we use two key values, $\Delta$ and $\Delta'$:
+For our new MH test, we define $\Delta$ and $\Delta'$ which are logs of transition probability
+ratios:
 \begin{equation}\label{eq:deltas}
@@
 \end{equation} 
-where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}. Thus, $\Delta$ and
-$\Delta'$ are probability ratios in log space. Note that the only difference between $\Delta$ and
-$\Delta'$ is that they use $N$ and $n$ data terms $x_i$, respectively, and that the latter uses an
-$N/n$ scaling term. We assume $n \ll N$ (so $\Delta'$ is substantially faster to compute) and that
-the $n$ data points $x_i$ represent a random minibatch without replacement.
+where $p$ and $q$ are target and proposal distributions respectively
+as in Equation~\ref{eq:traditional}. Note that the difference
+between $\Delta$ and $\Delta'$ is that they use $N$ and $n$ data terms
+$x_i$, respectively, and that the latter uses an $N/n$ scaling
+term. The scaling term ensures that $\Delta'$ is an unbiased estimator of $\Delta$.
+ We assume $n \ll N$ (so $\Delta'$ is substantially faster to
+compute) and that the $n$ data points $x_i$ represent a random
+minibatch of data.
 
@@
 \begin{lemma}\label{lem:detailed_balance}
-An acceptance function $g$ such that $g(\Delta) = \exp(\Delta) g(-\Delta )$ satisfies detailed
-balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) = f(\theta')p(\theta_t \mid \theta')$, where
-$p(\theta_y \mid \theta_x)$ is the probability of jumping from $\theta_x$ to $\theta_y$ in our
-chain.
+Let $\Delta$ be defined as in Equation~\ref{eq:deltas}. Any acceptance function $g$ such that
+$g(\Delta) = \exp(\Delta) g(-\Delta )$ satisfies detailed balance. That is, $f(\theta_t)p(\theta'
+\mid \theta_t) = f(\theta')p(\theta_t \mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the
+probability of jumping from $\theta_x$ to $\theta_y$ in our chain.
 \end{lemma}
@@
 
-For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
-(1+\exp(-\Delta))^{-1}$. Straightforward arithmetic shows that it satisfies the condition in
-Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily sample from
-it using the following procedure. At any iteration with a current parameter $\theta_t$ and a
-candidate sample $\theta'$, we can compute $\Delta$. Let $u$ be a uniform random variable $u \sim
-{\rm Unif}[0,1]$. We accept $\theta'$ if $g(\Delta) > u$, and reject otherwise. This process is
-equivalent to sampling a random variable $X$ with cumulative distribution function $F_X(x) = g(x)$,
-and accepting if $\Delta > X$ and rejecting otherwise. We can define $X = g^{-1}(u)$ so that its CDF
-is the logistic function, which follows because for arbitrary $X = x$, we have
+For our MH test, the key is that we use a different $g$, the logistic
+function: $g(\Delta) = (1+\exp(-\Delta))^{-1}$. Straightforward
+arithmetic shows that it satisfies the condition in
+Lemma~\ref{lem:detailed_balance}.  The logistic function is ideal for
+this task because we can easily sample from it using the following
+procedure: At any iteration with a current parameter $\theta_t$ and a
+candidate sample $\theta'$, we can compute $\Delta$. Let $u$ be a
+uniform random variable $u \sim {\rm Unif}[0,1]$. We accept $\theta'$
+if $g(\Delta) > u$, and reject otherwise. This process is equivalent
+to sampling a random variable $X$ with cumulative distribution
+function $F_X(x) = g(x)$, and accepting if $\Delta > X$ and rejecting
+otherwise. We can define $X = g^{-1}(u)$ so that its CDF is the
+logistic function, which follows because for arbitrary $X = x$, we
+have
 \[
@@
 as the density of $u$ is one. Thus, the criteria to accept the candidate $\theta'$ is equivalent to
-whether $\Delta > X$. Moreover, the density of $X$ (a standard logistic random variable) is
-symmetric about zero, so the acceptance criteria can also be expressed as $\Delta + X>0$.
+whether $\Delta > X$. Moreover, $X$ is symmetric about zero, so the acceptance criteria can also be
+expressed as $\Delta + X>0$.
 
@@
 
-With large datasets, $\Delta$ is intractable to compute. An intuitive fix is to replace $\Delta$
-with the minibatch version, $\Delta'$, for our MH test. The following lemma characterizes the distribution of $\Delta'$.
+With large datasets, $\Delta$ is intractable to compute. An intuitive fix is to instead use the
+minibatch approximation $\Delta'$. The following lemma characterizes the distribution of $\Delta'$.
 
 \begin{lemma}\label{lem:gaussian}
-If the minibatch data is chosen randomly without replacement and has sufficiently many elements,
-then the distribution of $\Delta'$ is approximately Gaussian.
+If the minibatch data is chosen randomly and sufficiently large, then the distribution of $\Delta'$ is approximately Normal.
 \end{lemma}
@@
 \]
-During a given iteration, $\theta_t$ and $\theta'$ are fixed. Therefore, the randomness in $\Delta'$
-comes only from the minibatch. Since the minibatch is chosen randomly, the distribution of
-$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t))$ converges to a Gaussian as $n
+During a given iteration, $\theta_t$ and $\theta'$ are fixed. Therefore, the variation in $\Delta'$
+comes only from the samples in the minibatch. Since each term in the sum is IID, the distribution of
+$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t))$ converges to a Normal distribtuion as $n
 \to \infty$ by the Central Limit Theorem.
@@
 \begin{equation}\label{eq:relationship}
-\Delta' = \Delta + \varepsilon, \quad \varepsilon \sim \mathcal{N}(0, \sigma^2),
+\Delta' \approx \Delta + \varepsilon, \quad \varepsilon \sim \mathcal{N}(0, \sigma^2),
 \end{equation}
-for a Gaussian noise term $\varepsilon$.  The noise means we cannot just replace $\Delta+X$ with
-$\Delta'+X$ for our test, so we need to slightly change our acceptance criteria. Our insight is to
+for a Normal noise term $\varepsilon$.  The extra noise means we can no longer directly perform
+the $\Delta + X$ test, so we need to slightly change our acceptance criteria. Our insight is to
 decompose $X$ as
@@
 \end{equation}
-where we assume $X_{\rm norm}$ is a zero-mean Gaussian and $X_{\rm corr}$ is a zero-mean
-``correction'' term.  These two add (i.e., their distributions convolve) to form $X$. The criteria
+where we assume $X_{\rm norm}$ is a zero-mean Normally-distributed and $X_{\rm corr}$ is a zero-mean
+``correction'' term.  These two add to form $X$. The criteria
 to accept is now
@@
 
-It is important to understand why we use Equations~\ref{eq:deconvolution} and~\ref{eq:criteria}. The
-deconvolution process allows us to use $\Delta' + X_{\rm corr}$ as a proxy for $\Delta + X$. It
-works because if we know the distribution\footnote{Computing the distributions of $\varepsilon$ and
-$X_{\rm norm}$ reduces to finding their standard deviations.} of $\varepsilon$ (from
-Equation~\ref{eq:relationship}), then we can pretend that our $\Delta'$ is in fact $\Delta$. To
-``insert'' $\varepsilon$ into our test, we deconvolve $X$ so that one of its components, $X_{\rm
-norm}$, plays the part of $\varepsilon$. In light of our revised MH test, the previous discussion
-raises the question: in an iteration of MCMC with current $\theta_t$ and proposed $\theta'$, how do
-we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample the
-\emph{values} of $X_{\rm corr}$ appropriately. This procedure consists of two major steps.
+Equations~\ref{eq:deconvolution} and~\ref{eq:criteria} yield an extremely simple test: Namely,
+compute the approximate log transition probability ratio $\Delta'$, add the correction random
+variable, and compare with zero. It remains to be determined how, and under what conditions
+we can do the decomposition~\ref{eq:deconvolution}. First we recall that the distribution
+of a sum of indendent random variables is the convolution of their distributions. From equation
+\cite{eq:deconvolution}, we know the distributions of both $X_{\rm norm}$ and $X$. It follows
+that $X_{\rm corr}$ is the {\em deconvolution} of $X$ by $X_{\rm norm}$. Such a decomposition
+need not exist. For instance, \ref{eq:deconvolution} implies that
+\begin{equation}
+  {\rm var}(X) = {\rm var}(X_{\rm norm}) + {\rm var}(X_{\rm corr})
+\end{equation}
+so we must have ${\rm var}(X) > {\rm var}(X_{\rm norm})$ to have a chance of success. In fact
+since these distributions are known, the variance information is enough to determine whether
+the test is applicable. The correction sample can be generated as follows:
 
@@
     of $\Delta'$, each using a different random minibatch of $n$ data points, but each using the
-    same $\theta_t$ and $\theta'$. In our experiments (see Section~\ref{sec:experiments}) we make $K$
-    small, around 5-10, for speed purposes, and $n$ is usually a few hundred for the Central Limit
-    Theorem to apply.  Using these $K$ values, we compute the empirical standard deviation ${\rm
-    std}(\Delta')$.
+    same $\theta_t$ and $\theta'$. In our experiments (see Section~\ref{sec:experiments}), $K$ is
+    small, around 5-10, and $n$ is usually a few hundred for the Central Limit Theorem to apply.
+    Using these $K$ values, we use the empirical standard deviation ${\rm std}(\Delta')$.
 
@@
     from standard deconvolution techniques. These rely on the well-known fact that the Fourier
-    transform of a convolution is the product of the individual Fourier transforms. In practice, we
-    pre-compute distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to
-    facilitate the sampling process.
+    transform of a convolution is the product of the Fourier transforms. In practice, we pre-compute
+    distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to facilitate the
+    sampling process.
 \end{enumerate}
@@
 \begin{algorithm}[t]
-\Input{Number of samples $T$, number of $\Delta'$ estimates $K$, minibatch size $m$, pre-computed $X_{\rm
-corr}$ distributions, $\{x_i\}_{i=1}^N$ values from a target distribution, and initial sample $\theta_0$.}
-\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ for either posterior inference or
-MAP/ML estimation.}
+\Input{Number of samples $T$, variance estimation window $W$, minibatch size $m$, pre-computed $X_{\rm
+corr}$ distributions, $\{x_i\}_{i=1}^n$ values from a target distribution, and initial sample $\theta_0$.}
+\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$}
 \For{$t=\{1, \ldots, T\}$}{
     Propose a candidate $\theta'$ and create an empty list for $\Delta'$ estimates, $d = []$\;
-    \For{$k = \{1, \ldots, K\}$}{
+    \For{$k = \{1, \ldots, W\}$}{
         Sample a random minibatch of size $m$ from the complete data\;
@@
         }{
-            Reject and re-use the old one, $\theta_{t+1} = \theta_t$\;
+            Reject and re-use the sample, $\theta_{t+1} = \theta_t$\;
         }
@@
 deviation of $X_{\rm norm}$ increases, the $X_{\rm corr}$ distribution becomes increasingly unstable
-and ``bumpier'', because the logistic curve has fatter tails than Gaussians. In addition, we cap
+and ``bumpier'', because the logistic curve has fatter tails than the normal distribution. For this reason, we cap
 the standard deviation possibilities of $X_{\rm norm}$ at $\approx 1.2$ in our experiments. In our
 experiments, we discretize the possible standard deviations of $X_{\rm norm}$, and we also limit the
-densities to consider the range $[-10,10]$ as shown in Figure~\ref{fig:deconvolution}.
+densities to consider the range $[-10,10]$ (as shown in Figure~\ref{eq:deconvolution}).
 
@@
 with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
-smaller in absolute value, thus reducing variance. The flatter posterior results in easier mixing.
+smaller, thus reducing variance. The flatter posterior results in easier mixing.
+
+\section{Accelerating the Test}
+The test as stated is dominated by the time to estimate the variance of $\Delta'$ using $W$
+samples per step. An alternative is to use a moving-average estimate of variance, using only
+a single sample per iteration. However, the variance of $\Delta$ may be quite complex since
+it involves interactions through the proposal distribution, between new and old states $\theta$.
+We can simplify it by decomposing it:
+
+\begin{equation}
+   \Delta' = \log \left(p(\theta') (\prod_{i=1}^n p(x_i \mid \theta'))^{\frac{N}{n}}\right) -
+   \log \left(p(\theta_t) (\prod_{i=1}^n p(x_i \mid \theta_t))^{\frac{N}{n}}\right) +
+   \log \left(\frac{
+   q(\theta' \mid \theta_t)}{q(\theta_t \mid \theta')}\right)
+\end{equation} 
+which can be written as
+\begin{equation}
+  \Delta' = X(\theta') + X(\theta_t) + X(q)
+\end{equation}
+where $X(\theta')$, $X(\theta_t)$ and $X(q)$ correspond to the three
+terms above. We can make independent estimates of $X(\theta')$ and
+$X(\theta_t)$ by using different minibatches for each.  This will
+likely increase the variance of $\Delta'$ but also render it more
+stable, since it is then depends only on the variance of the log
+likihood in the parameter region of interest.
 
@@
 \begin{lemma}\label{lem:theory2}
-For one step sampling from $\theta_t$ to $\theta'$, if the jumping step $\|\theta_t - \theta'\|_2 <
-\epsilon$, and the gradient of the log likelihood is bounded by a constant factor, i.e. $\|\nabla
-(\log \Pr(x_i\mid \theta))\|_2 < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2
-(m-\frac{m(m-1)}{N-1})$, where $m$ is the minibatch size, and $N$ is the total data size.
+For one step sampling from $\theta_t$ to $\theta'$, if the jumping step $|\theta_t - \theta'| <
+\epsilon$, and the gradient of the log-likelihood is bounded by a constant factor, i.e. $|\nabla
+(\log \Pr(x_i\mid \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 (m-\frac{m(m-1)}{N-1})$, where
+$m$ is the minibatch size, and $N$ is the total data size.
 \end{lemma}
@@
 
-Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 (m - \frac{m(m-1)}{N-1}) $, we can assume the maximum
-value of $X_{\xi}$ is proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we
-can assume $\zeta=\epsilon C$, where $C$ is a constant factor.
+Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 m $, we can assume the maximum value of $X_{\xi}$ is
+proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we can assume
+$\zeta=\epsilon C$, where $C$ is a constant factor.
 
@@
 $\alpha_i$s, which are determined by the error of our transition kernels. These involve competing
-objectives: the larger the step size, the more efficient the $P_i'$ but its error increases.  The
-reverse happens with a step size too small.
+objectives: the larger the step size size, the more efficient the $P_i'$ but its error increases.
+The reverse happens with a step size too small.
 
@@
 data points so that the target distribution is $p(\theta)\prod_{i=1}^{10000}p(x_i\mid \theta)$, with
-the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This
-results in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we apply a
-temperature $T=110$ to reduce ${\rm std}(\Delta')$. Taking logs, we get the target as shown in the
-far left of Figure~\ref{fig:gauss_mix_1}.
+the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This turns
+out to be a rather sharp posterior, so to make our test applicable, we apply a temperature $T=110$
+to reduce ${\rm std}(\Delta')$. Taking logs, we get the new target as shown in the far left of
+Figure~\ref{fig:gauss_mix_1}.
 
@@
 diag}(0.03, 0.03)$. This is a poor proposer, but it is sufficient for our purposes as the quality of
-the samples will be primarily due to the MH test. All methods are run 10000 times to collect 10000
-samples.
+the samples will be due to the MH test. All methods are run 10000 times to collect 10000 samples.
 
@@
 
-Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$ samples for the three
-methods, with darker regions indicating a greater density of points. All three methods obtain the
-same rough form of the posterior, so our MH test can indeed result in the same posterior as the
-other two methods.
+Figure~\ref{fig:gauss_mix_1} shows three scatter plots of the resulting $\theta$ samples for the
+three methods, with darker regions indicating a greater density of points. All three methods obtain
+the same rough form of the posterior, so our MH test can indeed obtain the posterior just as the
+other methods do.
 
 Figure~\ref{fig:gauss_mix_2} (to the left) shows a histogram of the final minibatch sizes used by
-the adaptive subsampling method each iteration (on a log scale for readability).  Usually, the method
+the adaptive subsampling method each iteration (capped at 600 for readability).  Usually, the method
 can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
-in lots of wasted computation. For this particular experiment, that happened eight times.  In
+in more computation (some of it wasted). For this particular experiment, that happened eight times.  In
 contrast, our method keeps the minibatch size fixed at 200, but requires that ${\rm std}(\Delta') <
 1.2$. Figure~\ref{fig:gauss_mix_2} (in the middle) shows a histogram of the estimated ${\rm
-std}(\Delta')$ values each iteration. Only 75 iterations (out of 10000) resulted in ${\rm
-std}(\Delta') \ge 1.2$ and for these, we skipped the rest of the iteration (setting
-$\theta_{t+1}=\theta_t$). To smooth the ${\rm std}(\Delta')$ values, we use moving average updates.
-The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
-\ge 1.2$ based on minibatch sizes. For six sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
-our MH test five times and averaged the number of ``violations.'' As the size increases, the
-number of violations decreases, as expected.
+std}(\Delta')$ values each iteration. Only 75 iterations resulted in ${\rm std}(\Delta') \ge 1.2$
+and for those, we skipped the rest of the iteration (setting $\theta_{t+1}=\theta_t$). To smooth the
+${\rm std}(\Delta')$ values, we use moving average updates. The third plot in
+Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta') \ge 1.2$. For six
+minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with our MH test five times and
+averaged the number of violations. As minibatch sizes increases, the number of violations decreases,
+showing that our test is more applicable.
 
@@
 dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007 and 1000 training and testing
-points, respectively (we used a random subset of the test data). The proposer is again a random walk
-with covariance matrix $0.1I$ for the $784\times 784$ identity matrix $I$. We set the posterior
-temperature at $T=3000$. We set the minibatch size $m=50$ and compare with adaptive sampling with
-tolerance 0.02.
+points, respectively (we used a random subset of the test data). The propose is again a random walk
+with covariance matrix $0.1I$ for the $784\times 784$ identity matrix $I$. We initialize the
+posterior temperature at $T_0=3000$, and decrease it according to $T_i = T_0 /\sqrt{i + 1}$ for
+iteration $i \ge 1$. We set the minibatch size $m=50$ and compare our minibatch MH test with the
+adaptive minibatch.
 
@@
     \label{fig:logistic_fig}
-    \vspace{-10pt}
+    %\vspace{-10pt}
 \end{figure}
 
-The third plot of Figure~\ref{fig:logistic_fig} shows performance based on minibatch size. The
-parameter is evaluated on the test set after training iterations 700, 1000, and 1500. For those
-three cases, when the minibatch size increases beyond about 10, the performance of the sampled data
-will not improve with a large minibatch size. This makes sense because the error of our MH test is
-proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch increases, ${\rm
-std}(\Delta')$ decreases. Beyond a certain standard deviation (or minibatch) threshold, the error in
-our minibatch MH test is negligible.
+The third plot of Figure~\ref{fig:logistic_fig} shows performance based on minibatch size, when the
+performance is evaluated on the test set after training iterations 700, 1000, and 1500. For all
+three iteration counts, when the minibatch size increases beyond about 10, the performance of the
+sampled data will not improve with a large minibatch size. This makes sense because the error of our
+MH test is proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch
+increases, ${\rm std}(\Delta')$ decreases. Beyond a certain standard deviation threshold, the error
+in our minibatch MH test is negligible.
 
@@
 In this experiment, we apply our minibatch MH test with a Stochastic Gradient Hamiltonian
-Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the complete parameters for a fully
-connected neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}, which
-is a binary classification task with one million instances, each of which are 28-dimensional. The
-network we use has four layers: the first is the input, the last is the softmax, and the
-intermediate ones apply the sigmoid activation unit.  We also employ dropout and batch
-normalization~\cite{icml2015_ioffe15}.
+Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the parameter instances from a fully connected
+neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}. The network we
+use for this binary classification problem has four layers and is fully connected. The first layer
+is the input, the last layer is the softmax, and the intermediate layers apply the sigmoid
+activation unit.  We also employ dropout and batch normalization~\cite{icml2015_ioffe15}.
 
@@
 Descent method~\cite{adapGrad}. Since the frequencies of the features are quite different, we use
-the same strategy as in Adaptive Gradient to rescale the gradient before applying SGHMC. We
-implement our MH test and Adaptive Gradient in the open-source BIDMach
-project~\cite{canny2013bidmach}.
+the same strategy in the Adaptive Gradient to rescale the gradient before apply SGHMC. We implement
+our MH test and Adaptive Gradient in the open-source BIDMach project~\cite{canny2013bidmach}.
 
@@
 \begin{itemize}[noitemsep]
-    \item Proofs for material in Section~\ref{sec:theory}.
-    \item More information on Section~\ref{ssec:gaussians}.
-    \item More information on Section~\ref{ssec:logistic}.
-    \item More information on Section~\ref{ssec:nets}.
+    \item Some proof details on results in Section~\ref{sec:theory}.
+    \item Detailed discussion of the experiment from Section~\ref{ssec:gaussians}.
+    \item Detailed discussion of the experiment from Section~\ref{ssec:logistic}.
+    \item Detailed discussion of the experiment from Section~\ref{ssec:nets}.
 \end{itemize}
@@
 & = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
-&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| \\
-&= |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
-&= |\nabla F(-\Delta) \xi + o(\xi^2)|\\
-&\leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2\ell \xi,
+&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| = |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
+&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2\ell \xi,
 \end{align*}
-where we use Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have smaller absolute
-value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| < 2 \zeta \ell$.
+where the last line uses Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have
+smaller absolute value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| <
+2 \zeta \ell$.
 
@@
 \|P_i - P'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
-&= \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
-&\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta')) d\theta_t \right| d\theta'\\
-&= 2 \zeta \ell,
+& = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
+& \leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta')) d\theta_t \right| d\theta'= 2 \zeta \ell,
 \end{align*}
@@
 \begin{align*}
-{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta_t)\right] = {\rm Var}\left(\sum_{i=1}^m q_i\right) \\
+{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta_t)\right] = {\rm Var}\left(\sum_i^m q_i\right) \\
 &= m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
@@
 
-Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have 
-\[
-\| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t \| \leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t
-\rho_u (1-\alpha_u)\right\} \rho_s \alpha_s,
-\]
-where $\alpha = \frac{\epsilon_i C}{1-\eta}$.  Thus, we get:
+Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| P_t' \circ P_{t-1}' \circ \cdots
+\circ P_1' D_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
+\alpha_s$, where $\alpha = \frac{\epsilon_i C}{1-\eta}$. Thus, we get:
 \begin{align*}
- \| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_0 \| &\leq \|P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
+ \| P_t' \circ \cdots \circ P_1' D_0 - \pi_0 \| &\leq \|P_t' \circ \cdots \circ P_1' D_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
  &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
@@
 
-In this section, we go over the math details on the Gaussian mixture model problem borrowed
-from~\cite{langevin_2011}. Our parameter is a 2-D vector $\theta =
+\subsection{Mathematical Assumptions}
+
+We borrow this example from~\cite{langevin_2011}. Our parameter is a 2-D vector $\theta =
 (\theta_1, \theta_2)$, where
@@
 \end{equation}
-Here is the problem statement: given some number of conditionally independent data points $x_1, x_2,
-\ldots, x_N$ generated according to~(\ref{eq:x_points}), determine the posterior
-distribution of $\theta$:
+Here is the problem statement. Given some number of i.i.d. data points $x_1, x_2, \ldots, x_N$
+generated according to~(\ref{eq:x_points}), determine the posterior distribution\footnote{Note that,
+as is typical with Bayesian analysis, we do not concern ourselves with the denominator of the
+original (non-log) posterior.} of $\theta$:
 \begin{equation}\label{eq:log_post}
@@
 where we only use $n \ll N$ samples, but we must scale up the likelihood contribution by $N/n$. If we
-didn't add this scaling factor, then the contribution of the likelihood terms would be weaker.
-
-One technique we use is adding a \emph{temperature} to our distribution. In general, we will want to
-add $T > 1$ so that our posterior is $p(\theta)((\prod_{i=1}^n p(x_i\mid \theta))^{N/n})^{1/T}$,
-resulting in the log posterior of 
+didn't add this scaling factor, then the contribution of the likelihood terms would be weaker. There
+is another thing we need to consider about that, which is the \emph{temperature} of our
+distribution. In general, we will want to add $T > 1$ so that our posterior is
+$p(\theta)((\prod_{i=1}^n p(x_i\mid \theta))^{N/n})^{1/T}$, resulting in the log
+posterior of 
 \begin{equation}\label{eq:log_prior_temp}
@@
 which has the extra $1/T$ to decrease the scale factor.  Equation~(\ref{eq:log_prior_temp}) is what
-we use for our experiments, because warmer distributions help us satisfy our ${\rm std}(\Delta') <
-1.2$ requirement.
+we will be using for our experiments, because we need to consider warmer distributions.
+{\color{blue} Daniel: not totally confident on this but I'll try.}
+
+Finally, recall earlier that we needed to define $\Delta$. In our context, $\Delta$ is:
+\begin{align}
+\Delta &= \log \left(\frac{f(\theta') q(\theta_t \mid \theta')}{f(\theta_t) q(\theta'\mid \theta_t)} \right) \\
+&= \log p(\theta') - \log p(\theta_t) + \sum_{i=1}^N(\log p(x_i \mid \theta') - \log p(x_i \mid \theta_t)) + \log\left(\frac{q(\theta_t \mid \theta')}{q(\theta' \mid \theta_t)}\right).
+\end{align}
+But again, with \emph{minibatches} of data, we usually do not sum over all $N$ terms and instead sum
+over $n$ terms, but with the extra $(1/T)(N/n)$ scaling factor, from
+Equation~(\ref{eq:log_prior_temp}). We generally denote that approximation as $\Delta'$ in the paper.
+
+\subsection{Experiment Observations}
 
@@
 
-To gain some intuition on what the posterior looks like, Figure~\ref{fig:contour1} shows simulated
-contour plots of the posterior based on varying numbers of data points $N$, with the temperature set
-at $T=1$. Note that because we are using all $N$ points here, the scale factor $N/n=1$. As $N$
-increases, the posterior converges to a multimodal distribution with modes at $(0,1)$ and $(1,-1)$.
+In this experiment, we compare our method with the one from~\cite{cutting_mh_2014}, hereafter called
+the ``adaptive sampling approach''. We apply them to the Gaussian Mixture model previously
+discussed.
+
+Figure~\ref{fig:contour1} shows simulated contour plots of the posterior based on the $N$ data
+points, for varying $N$, with the temperature set at $T=1$. Note that because we're using all $N$
+points here, the scale factor $N/n=1$. As $N$ increases, the posterior converges to a multimodal
+distribution with modes at $(0,1)$ and $(1,-1)$.
+
 Figure~\ref{fig:contour2} is similar, except this time we fix the number of samples at $N=10000$,
@@
 
+Our ultimate goal is to either (a) get a posterior distribution that covers the posterior well, or
+(b) to get samples that stay trapped at a mode, for a point estimate of $\theta$.  Whether our goal
+is (a) or (b) depends on our objectives and the context of the problem (e.g., big data usually
+implies (b) is OK even if it would not satisfy the average Bayesian).
+
+For both our method and the adaptive sampling method, we use random walk proposals. (Thus, the term
+in $\Delta$ containing the $q$s disappears.)  Random walk proposals are bad, but since the proposal
+is poor, good performance can only be obtained with a strong MH test, hence why we can use this as a
+reasonable starting benchmark.
+
+We test use the following settings:
+
+\begin{itemize}[noitemsep]
+    \item We use $N=10000$ samples of $x_i$.
+    \item The covariance of the Gaussian random walk is ${\rm diag}(0.1, 0.1)$.
+    \item We use 10000 iterations. One iteration involves one $\theta'$ proposal and a test to
+    accept/reject.
+    \item The minibatch size is set at 100, so it is one percent of the total data. For the
+    adaptive sampling approach, this is the amount that the minibatch size would keep increasing if
+    we needed more precision. Our approach, of courses, keeps the size fixed.
+    \item For adaptive sampling, the tolerance for deciding on a test is 0.05. The temperature for
+    that is also set at $T=1$ because that algorithm was not designed to deal with temperatures.
+    \item For our approach, to get an estimate of the standard deviation of $\Delta'$ (remember,
+    $\Delta'$ is a minibatch approximation of $\Delta$) we take five other minibatches (randomly
+    selected from the data) and take the standard deviation of those values.
+\end{itemize}
+
+For our approach, we run using three different temperature values, $T=1$, $T=10$, and $T=100$.
+
+Figure~\ref{fig:scatter} compares scatter plots of the ``Cutting the MH'' test (hereafter called
+``adaptive minibatch'') versus ``Our Method'', where we use three different temperature settings.
+There doesn't seem to be too much noticeable difference in our method for the three different
+temperature setting values, but our approach is a lot different from the adaptive sampling method
+(this is because of rejection rates, more on that later).
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1\linewidth]{scatter_v01}
+  \caption{Scatter plots of accepted $\theta$ for (in order) the adaptive sampling approach, and our
+  method (temperatures 1, 10, and 100).}
+  \label{fig:scatter}
+\end{figure}
+
+Before taking a look at our approach, we first examine the adaptive sampling approach. \textbf{That
+approach rejected a lot of samples; out of 10000 iterations, it only accepted 773 times}.
+Figure~\ref{fig:mb_sizes} describes the histograms of the adaptive minibatch's batch sizes. (Note
+that the histogram y-axis values are different.) In many cases, using the initial batch size of 100
+sufficed to make an accept/reject decision, but sometimes, this method needed to use the entire 10k
+data set. Decreasing the tolerance (from its current 0.05 value) would have required larger
+minibatches. Incidentally, it makes sense that the method rejects a lot of samples, because the
+random walk proposal usually provides us with poor points.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.75\linewidth]{adaptive_sampling_sizes_v01.png}
+  \caption{Minibatch sizes for adaptive sampling when it was accepting or rejecting.}
+  \label{fig:mb_sizes}
+\end{figure}
+
+Now we analyze our approach in more detail. Our method accepts far more samples than the adaptive
+sampling method, \textbf{accepting 3285, 3318, and 3332 out of 10k}, for temperature values 1, 10,
+and 100, respectively.
+
+For diagnostics purposes, Figures~\ref{fig:diagnostics1},~\ref{fig:diagnostics2},
+and~\ref{fig:diagnostics3} express histograms (for all three temperature runs) of the $\Delta'$,
+${\rm std}(\Delta')$, and $X_{\rm corr}$ values, respectively.
+
+There are a few immediate observations. First, increasing the temperature size by a factor of 10
+will decrease $\Delta'$ by a factor of 10, which makes sense. That will also directly affect the
+standard deviation estimate of $\Delta'$. Our $X_{\rm corr}$ distribution is symmetric about zero,
+which is also expected (and it also exhibits a decrease in a factor of 10 for an increase
+in temperature).
+
+{\color{blue}
+Daniel: One concern I have is that our $\Delta'$ values do not seem to be ``sufficiently negative''. If our
+$\Delta'$ values are negative (but large in absolute value) that means our MH test will reject more
+often. But then this raises another concern: didn't we say at one point that we \emph{should} be
+getting high acceptance rates, around 50 percent (in part because of the shape of the logistic
+function)? Doesn't that depend on the quality of the proposal distribution? If our method is
+designed to accept a lot regardless of the proposal distribution, I don't see how its performance
+can be considered equal to the adaptive sampling approach, or even a standard MH test. In fact, the
+best observation I can derive from this is that our approach and the adaptive sampling (or any
+standard MH test) will accept the high likelihood cases, but our case will accept the borderline
+ones more often.
+}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1\linewidth]{our_deltas_v01.png}
+  \caption{Our $\Delta'$ values for three different temperature values.}
+  \label{fig:diagnostics1}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1\linewidth]{our_sds_v01.png}
+  \caption{Our ${\rm std}(\Delta')$ values for three different temperature values.}
+  \label{fig:diagnostics2}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1\linewidth]{our_xcorrs_v01.png}
+  \caption{Our $X_{\rm corr}$ values for three different temperature values.}
+  \label{fig:diagnostics3}
+\end{figure}
+
 \section{Logistic Regression Experiment Details}\label{app:logistic}
 
-In this section, we go over some details of our logistic regression experiment. The feature vector
-for an image consists of its pixel values, normalized between 0 and 1. For simplicity, we only
-consider the binary classification case, so we only use digits one (denoted as output $y = 1$) and
-seven (denoted as $y = -1$). The probability of the $i^{\rm th}$ output $y_i \in \{-1,+1\}$ with the
-input vector $x_i$ is
-\begin{equation}
-p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)},
-\end{equation}
-where $\theta$ is the 784-length parameter vector.
-
-For our experiment, we impose a uniform prior to represent our lack of knowledge about $\theta$.  We
-use a random walk proposer, which can be modeled as $\theta' = \theta_i + \mathcal{N}(0,
-\sigma^2I)$, where $\theta_i$ is the current sample, $\theta'$ is the proposed sample, and we choose
-the variance to be a constant $\sigma^2 = 0.01$ for all components. We initialize $\theta_0$ to be a
-vector of all ones, and set our minibatch size as $m=50$.
-
-For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we use a
-constant temperature $T = 3000$. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
-current iteration.  Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
-versus iteration count.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=1\linewidth]{./figures/appendixExp2}
-  \caption{Additional results for the logistic regression experiment.}
-  \label{fig:appendixExp2}
-\end{figure}
-
-For adaptive MH testing, our experimental settings are the same as with our MH test, except we do
-not impose a temperature. The minibatch size of adaptive MH testing is also initialized as 50, but
-it may increase by that amount each iteration. Figure~\ref{fig:appendixExp2} shows the histogram of
-the actual minibatch size at the end of each iteration in adaptive MH testing.
+We applied our minibatch Metropolis Hastings algorithm to a Bayesian logistic regression model,
+comparing with the adaptive sampling method. We tested both our method and the adaptive sampling
+method using a random walk proposer $q(\theta \mid \theta_t) = N (\theta_t, \sigma^2)$. Since the
+random walk proposer does not contain any information about the target distribution, it only relies
+on the Metropolis-Hastings test to converge to the correct distribution. Therefore, random walk
+proposer is the ideal choice for illustrating our algorithm and comparing with the adaptive sampling
+method.
 
@@
 
-For our neural network experiment, we use the architecture discussed in Section~\ref{ssec:nets}. We
-use the SGHMC as the proposer for our minibatch MH testing. For the baseline, we use a tuned
-adaptive gradient descent optimizer, whose step size changes by $0.01 /(i+1)^{0.4}$. Both methods
-have minibatch sizes set at $m=200$. There are one million total data instances $x_i$.
-
-For SGHMC, we use the simplified update equations~\cite{sghmc_2014}:
-\begin{equation}
-\Delta \theta = v, \quad \quad\Delta v = -\eta \nabla U(\theta) - \alpha v + \mathcal{N}(0, 2(\alpha
--\hat{\beta}) \eta),
-\end{equation}
-where $v$ represents auxiliary momentum variables, and $\nabla U(\theta)$ is the gradient of the
-system. We set the hyperparameters to be $\eta_i = 0.01 /(i+1)^{0.4}$, and $\alpha=0.1$. We use the
-empirical Fisher~\cite{conf/icml/AhnBW12} information $V(\theta)$ to estimate the value of
-$\hat{\beta}$, so that $\hat{\beta}=\frac{1}{2}\eta V(\theta)$.
-
-In order to control ${\rm std}(\Delta')$, we initialize the temperature at 1000, and adjust it
-at iteration $i$ according to $T_i = \max\{1, 1000/(i+1)^{0.5}\}$.
+{\color{blue}
+Daniel: TODO
+}
 

### 2016-06-02 generateXcorr/generateXcorrCDF.m
added new generator

@@ -1,0 +1,64 @@
+% Compute the pdf of the X_Corr variable such that X_norm + X_corr ~ X_logistic
+% This version minimizes the CDF error.
+
+n = 100;                                             % resolution of discrete X_norm and X_logistic (actually 2n)
+dist = 20;                                           % real x range encoded over these 2n values (-dist, dist)
+nn = 1;                                              % multiplier for additional values outside the central range
+sigma = 1.1;                                         % std deviation of the normal distribution
+
+x = ((-n):n)/n*dist;                                            % x values
+
+lcdfm =  1./(1+exp(-x));                                        % logistic cdf at the central values
+lcdf = [zeros(1, nn*n), lcdfm, ones(1, nn*n-1)];                % logistic cdf over the full range
+
+ncdfm =  normcdf(x, 0, sigma);                                  % normal cdf at the central values
+ncdf = [zeros(1, (nn+1)*n), ncdfm, ones(1, (nn+1)*n-1)];        % normal cdf over the full range
+
+ndr = 2*nn+2;                                                   % convenient constant
+convmat = zeros(ndr*n, 2*n);                                    % convolution matrix
+for i = 1:(2*n)
+  convmat(:, i) = (ncdf(i:(i+ndr*n-1)))';
+end
+
+% variables [xcorr, eps];
+
+tail = lcdf'+ 2.0e-3;                                           % epsilon multiplier for relative error minimization
+tail = ones(ndr*n,1);                                           % epsilon multiplier absolute error minimization
+
+A = [[convmat ; -convmat] , [-tail ; -tail]];                   % A and b inequality constraints, Ay <= b. Equivalent to
+b = [lcdf'; -lcdf'];                                            % convmat * y - lcdf <= eps and
+                                                                % convmat * y - lcdf >= -eps
+
+Aeq = [ones(1, 2*n), 0];                                        % equality constraints specify sum(y) = 1
+beq = 1;                                                        % but they dont seem to help
+
+f = [zeros(2*n,1); 1];                                          % objective to minimize = eps
+lb = zeros(2*n+1,1);                                            % lower bound all zeros;
+
+% y=linprog(f,A,b,Aeq,beq,lb);
+
+options = optimoptions('linprog');
+% options = optimoptions(@linprog,'Algorithm','active-set')
+% options.Algorithm='interior-point';
+% options.OptimalityTolerance=1.0e-20;
+% options.MaxIterations= 200;
+
[…]

### 2016-06-15 nips_2016.tex
added main tex file back
midway to revise the paperr

@@
 process continues until a decision. The downside of this algorithm is the need to keep incrementing
-the minibatch size; in the worst case, all $N$ data points may be needed to decide on $\theta'$. A
+the minibatch size; in the worst case, all $N$ data points may be needed to decide on $\theta'$.
+We prove that the lower bound of average number of data instance consumed for one iteration is $O(\sqrt{N})$, where $N$ is the number of all data points, see Lemma~\ref{lem:efficienty_cutting}.
+A
 similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties
 based on concentration inequalities, but comes at the cost of more computational time.
+
+\begin{lemma}\label{lem:efficienty_cutting}
+If the sample mean $\bar{l}$ has uniform distribution from $[-L,L]$, the lower bound of expected data consumed of adaptive sampling method for one iteration is $\frac{\eta s_{l}Q}{2L}\sqrt{N}$, where $\bar{l} = \frac{1}{n}\sum_{i=1}^{n} \left (\log p(x_i; \theta') - \log p(x_i| \theta_t)\right )$, $\eta = \phi^{-1}_{n-1}(1-\varepsilon)$, $Q=\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')}$, $s_l$ is the population standard deviation of $l_i$, $\phi_{n-1}()$ is the CDF of t distribution with degree of freedom $n-1$, and $n$ is the minibatch size.
+\end{lemma}
 
@@
 we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample the
-\emph{values} of $X_{\rm corr}$ appropriately. This procedure consists of two major steps.
+\emph{values} of $X_{\rm corr}$ appropriately. This procedure consists of three major steps.
 
 \begin{enumerate}[noitemsep]
-    \item We first need to estimate the standard deviation of $X_{\rm norm}$. We generate $K$ values
-    of $\Delta'$, each using a different random minibatch of $n$ data points, but each using the
-    same $\theta_t$ and $\theta'$. In our experiments (see Section~\ref{sec:experiments}) we make $K$
-    small, around 5-10, for speed purposes, and $n$ is usually 50-200.  Using these $K$ values, we
-    compute the empirical standard deviation ${\rm std}(\Delta')$.
+    \item We first need to estimate the standard deviation of $X_{\rm norm}$ (i.e.
+the standard deviation of $\Delta'$).
+We use one minibatch data to complete this task.
+From the proof of Lemma~\ref{lem:gaussian}, we know the randomness comes from $\frac{N}{n}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta_t) \right)$.
+Thus, ${\rm var}(\Delta') = \frac{N^2}{n^2}\sum_{i=1}^n{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right ) = \frac{N^2}{n}{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right )$.
+We use data instances in one minibatch to estimate the ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right )$.
+Since there are $n$ data instances in one minibatch, the estimation of ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right )$ has high accuracy.
+In our experiments (see Section~\ref{sec:experiments}) we also use the complete data to obtain the population variance of $\left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right )$, and use the nonparametric Bootstrap to estimate the ${\rm std(std}(\Delta'))$, so that we can construct the confidence interval of ${\rm std}(\Delta')$ to reveal the accuracy of our estimation.
+
+	\item If the ${\rm std}(\Delta') > 1.2$, the decomposition of $X$ (i.e.
+Equation~\ref{eq:deconvolution}) is unstable.
+We consume more minibatches to decrease the ${\rm std}(\Delta')$.
+We update the value of $\Delta'$ by $\Delta' = (k\Delta'_{\rm old} + \Delta'_{\rm new}) / (k+1)$, where $\Delta'_{\rm old}$ is the existing $\Delta'$ which uses $k$ minibatches to computed (usually $k=1$), and $\Delta'_{\rm new}$ is the $\Delta'$ computed from the new minibatch.
+We also update the value of ${\rm std}(\Delta')$ by ${\rm std}(\Delta') = \sqrt{\left ( k^2 {\rm var}(\Delta'_{\rm old}) + {\rm var}(\Delta'_{\rm new}) \right ) / (k+1)^2}$.
+Thus, we can decrease the ${\rm std}(\Delta')$ by consuming more data until the ${\rm std}(\Delta') > 1.2$.
+With help of temperature (see Section~\ref{ssec:discussion}), we will show that usually the mean of minibatch consumed per iteration is below 2 in Section~\ref{sec:experiments}.
+Later our efficiency analysis in Section~\ref{sec:theory} will reveal that the number of data instance consumed by our algorithm is much fewer than that of exisitng algorithms.
 
@@
 \For{$t=\{1, \ldots, T\}$}{
-    Propose a candidate $\theta'$ and create an empty list for $\Delta'$ estimates, $d = []$\;
-    \For{$k = \{1, \ldots, K\}$}{
-        Sample a random minibatch of size $m$ from the complete data\;
-        Compute $\Delta'$ from this minibatch (see Equation~\ref{eq:deltas}), and add $\Delta'$ to $d$\;
+    Propose a candidate $\theta'$\;
+    Compute the $\Delta'$ from one minibatch data\;
+    Compute the ${\rm std}(\Delta')$ from another minibatch data\;
+    \While {${\rm std}(d) \ge 1.2$} {
+    	Update the value of $\Delta'$ and ${\rm std}(\Delta')$ from one new minibatch\;
     }
-    Compute ${\rm std}(d)$, the standard deviation estimate of $\Delta'$ from list $d$\;
-    \eIf{${\rm std}(d) \ge 1.2$}{
-        Skip loop (set $\theta_{t+1}=\theta_t$) and apply variance preconditioning fix (if desired)\;
-    }{
+
         Choose the closest $X_{\rm corr}$ distribution, and sample a value $X_c$ from it\;
-        Using another random minibatch of $m$ data points, compute a \emph{new} $\Delta'$ (call it $\Delta_{\rm real}'$)\;
-        \eIf{$X_c + \Delta_{\rm real}' > 0$}{
+
+        \eIf{$X_c + \Delta' > 0$}{
             Accept the candidate, $\theta_{t+1} = \theta'$\;
@@
         }
-    }
+
 }
@@
 
-In this section, we explore the convergence properties of our MH test. Due to space constraints, all
+In this section, we firstly explore the convergence properties of our MH test. Secondly, we compare the efficiency between our MH test and adaptive MH test. Due to space constraints, all
 our proofs are in Appendix~\ref{app:proofs}.
@@
 
+We explore the efficiency of our minibatch MH test.
+Theorem~\ref{thm:efficiency} indicates that our minibatch MH test only requires a constant number of instances for each step, however, recall the Lemma~\ref{lem:efficienty_cutting}, the adaptive minibatch MH test requires $O(\sqrt{N})$ number of data instance for each step.
+Thus, We show that our proposed minibatch MH test is much higher efficient than the known minibatch algorithms.
+
+\begin{theorem}\label{thm:efficiency}
+We require $\frac{ C^2 \rho^2 }{\sigma^6 \varepsilon^2}$ minibatch size to ensure $|P_a -P_a'| < \varepsilon$, where $C$ is a positive constant which is smaller than 0.4748, $\sigma^2 $ and $\rho$ is the second and third moment of $(\nabla\log \Pr(x_i|\theta_t) - \nabla\log\Pr(x_i|\theta')) - E\left[ \nabla\log \Pr(x_i|\theta_t) - \nabla\log\Pr(x_i|\theta')\right]$, $P_a$ is the accpet rate from the complete data, $P_a'$ is the accpet rate from minibatch data by our algorithm, $\theta_t$ is the current parameter, and $\theta'$ is the proposed parameter.
+
+\end{theorem}
+
 \section{Experiments}\label{sec:experiments}
@@
 
-Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$ samples for the three
-methods, with darker regions indicating a greater density of points. All three methods obtain the
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=1\linewidth]{CI_v01.png}
+    \caption{
+    Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
+    }
+    \label{fig:gauss_mix_3}
+    %\vspace{-10pt}
+\end{figure}
+
+Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$ samples for the four
+methods, with darker regions indicating a greater density of points. The first three methods obtain the
 same rough form of the posterior, so our MH test can indeed result in the same posterior as the
 other two methods.
+However, the fourth method (i.e.
+just treat the $\Delta'$ which comes from minibatch as $\Delta$ which comes from the complete data) reveals wider spread distribution than the posterior.
+From Figure~\ref{fig:gauss_mix_1} we illustrate our proposed method can obtain the samples which have almost the same distirbution as the standard MH test from the complete data, adaptive minibatch test, and the true posterior.
+Meanwhile, we show the benefits of our minibatch method compared with the navie minibatch method.
 
@@
 can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
-in lots of wasted computation. For this particular experiment, that happened eight times.  In
-contrast, our method keeps the minibatch size fixed at 200, but requires that ${\rm std}(\Delta') <
-1.2$. Figure~\ref{fig:gauss_mix_2} (in the middle) shows a histogram of the estimated ${\rm
-std}(\Delta')$ values each iteration. Only 75 iterations (out of 10000) resulted in ${\rm
-std}(\Delta') \ge 1.2$ and for these, we skipped the rest of the iteration (setting
-$\theta_{t+1}=\theta_t$). To smooth the ${\rm std}(\Delta')$ values, we use moving average updates.
+in lots of wasted computation.
+For this particular experiment, that happened eight times.
+The mean of adaptive minibatch size is 520 which is roughly close to $\sqrt{N}$ (i.e.
+100).
+This number validates the claims of Theorem~\ref{thm:efficiency}.
+In
+contrast, our method keeps the minibatch size around 200. Figure~\ref{fig:gauss_mix_2} (in the middle) shows a histogram of the minibatch size, and the mean of our minibatch size is 247. 
 The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
-\ge 1.2$ based on minibatch sizes. For six sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
+\ge 1.2$ for only consuming one mimnibatch data. For six minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
 our MH test five times and averaged the number of ``violations.'' As the size increases, the
 number of violations decreases, as expected.
+
+Figure~\ref{fig:gauss_mix_3} (to the left) shows a historgram of ${\rm std}(\Delta')$.
+Figure~\ref{fig:gauss_mix_3} (to the middle) shows the confidence interval of our estimated ${\rm std}(\Delta')$ at each iteration \footnote{In order to show the confidence interval clearly, we use moving smooth to smooth the confidence interval, and plot each value by interval of 50 iterations}.
+The blue line represents the standard deviation comes from the complete data, which can be treated as the ``true'' ${\rm std}(\Delta')$, the green line represents our estimation of ${\rm std}(\Delta')$ from the minibatch, and the two red dash lines represent the confidence interval of ${\rm std}(\Delta')$ by adding / subtracking one ${\rm std(std(}\Delta'))$.
+This figure indicates that our estimation of ${\rm std}(\Delta')$ from one minibatch data is accurate.
+Figure~\ref{fig:gauss_mix_3} (to the right) shows the value of $\rho/\sigma^3$ at each iteration.
+Recall Theorem~\ref{thm:efficiency}, the number of data instance required is $\frac{4 \ell^2 C^2 \rho^2 }{\sigma^6 \varepsilon^2}$.
+This figure reveals that $\rho/\sigma^3$ is rougly a constant, which means the data required for each iteration is a constant value has nothing to do with the $N$.
 
@@
 our minibatch MH test is negligible.
+
+Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{} to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability of $\rho/\sigma^3$.
 
@@
 
+\subsection{Proof of Lemma~\ref{lem:efficienty_cutting}}\label{app:lem_efficiency}
+
+If the minibatch size is $n$, we fisrt compute the probability that more than $n$ data instances are required for adaptive sampling method for one iteration. We denote this probaility $P_{rej}$.
+\[P_{rej} = \Pr \left ( 1-\phi_{n-1} \left( \frac{|\bar{l} - \frac{1}{N} \log (u\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')})|}{s} \right )  \geq \varepsilon \right ) \]
+
+where $s = \frac{s_l}{\sqrt{n}}\sqrt{1-\frac{n-1}{N-1}}$. \\
+We define random variable $Z = \frac{1}{N}\log (uQ)$, where $Q=\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')}$.
+Then $Z$ has probability density function $f_Z(z) = N \exp(zN), z \in (-\infty, \frac{1}{N} \log Q]$.
+
+\begin{align*} 
+P_{rej}&=  \Pr \left( 1-\varepsilon \geq \phi_{n-1} \left( \frac{|\bar{l} - Z|}{s_l \sqrt{\frac{N-n}{N-1}}} \sqrt{n} \right) \right) \\ 
+&=  \Pr \left(\phi^{-1}_{n-1} \left (1-\varepsilon \right) \frac{s_l }{\sqrt{n}} \sqrt{\frac{N-n}{N-1}} \geq |\bar{l} -Z| \right) = \Pr \left( |\bar{l} - Z| \leq \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \right) \\
+&= \Pr \left(Z - \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \leq \bar{l} \leq Z + \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \right) \\
+&= \int_{-\infty} ^{\frac{\log Q}{N}} N \exp (zN)dz \int_{z - \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}}} ^{z+ \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}}} \frac{1}{2L} d\bar{l} = \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \int_{-\infty} ^{\frac{\log Q}{N}} N \exp(zN) dz \\
+&= \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} Q
+\end{align*}
+ If we define the random variable $X$ the number of instance for one minibatch, we have $\Pr (X>n) = P_{rej} = \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} Q, \mbox{ for } n \in [1,N], n \in Z^+$.
+Since $X$ is nonnegative, we have $E(X) \geq \Pr (X >x)x, \forall x$.
+If we choose $x=N/2$, we have $E(X) \geq \frac{\eta s_l}{L\sqrt{N/2}}\sqrt{\frac{N-N/2}{N-1}} Q \frac{N}{2} \geq \frac{\eta s_{l}Q}{2L}\sqrt{N}$.
+
 \subsection{Proof of Lemma~\ref{lem:theory1}}\label{app:theory1}
@@
 as desired.
+
+\subsection{Proof of Theorem~\ref{thm:efficiency}}\label{app:theory4}
+From Berry-Esseen theorem, we know $|F_n(x) - \Phi(x)| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$, where $\Phi()$ is the CDF of standard normal distribution.
 

### 2016-06-15 nips_2016.tex
finish my revise Haoyu

@@
 
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=1\linewidth]{CI_v02.png}
+    \caption{
+    Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
+    }
+    \label{fig:exp_3}
+    %\vspace{-10pt}
+\end{figure}
+
 The third plot of Figure~\ref{fig:logistic_fig} shows performance based on minibatch size. The
@@
 
-Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{} to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability of $\rho/\sigma^3$.
+Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{fig:exp_3} to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability of $\rho/\sigma^3$.
 
@@
 \subsection{Proof of Theorem~\ref{thm:efficiency}}\label{app:theory4}
-From Berry-Esseen theorem, we know $|F_n(x) - \Phi(x)| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$, where $\Phi()$ is the CDF of standard normal distribution.
+Since only the $\frac{N}{nT}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta_t) \right)$ term brings randomness into $\Delta'$.
+Define $X_i = \frac{N}{T}\left(\log p(x_i|\theta') - \log p (x_i|theta_t) \right)$, $\tilde{X_i} = X_i - E(X_i)$, and $Y_n = \frac{\sum_{i=1}^n \tilde{X_i}}{n}$ (i.e.
+$Y_n$ is the $\Delta' - E(\Delta')$).
+From Berry-Esseen theorem, $|F_n(x) - \Phi(x)| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$, where $\Phi()$ is the CDF of standard normal distribution, $F_n()$ is the CDF of $Y_n$, $C$ is the constant parameter which is below 0.4748, $\rho = E(|\tilde{X_i}|^3)$, and $\sigma^2 = E(\tilde{X_i}^2)$.
+Since $\Delta'$ is an unbiased estimator of $\Delta$, the CDFs between $\Delta'$ and $\Delta + X_{\rm norm}$ has the same max gap (i.e.
+$\frac{C\rho}{\sigma^3\sqrt{n}}$).
+
+From Equation~\ref{eq:criteria}, we show we want to use $\Delta' + X_{\rm corr}$ to approximate $\Delta + X_{\rm norm} + X_{\rm corr}$.
+Since $|F_{\Delta'} - F_{\Delta + X_{\rm norm}}| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$, for the worst case, assume $\exists \xi, ~{\rm s.t.} |f_{\Delta'}(\xi) - f_{\Delta + X_{\rm norm}} (\xi)| = \max_x |f_{\Delta'}(x) - f_{\Delta + X_{\rm norm}}(x)| \overset{\Delta}{=} \eta $, where $f()$ represents the PDF function.
+Because PDF of $X_{\rm corr}$ has the positive value and unit area, we have:
+\begin{align*} 
+& |(f_{\Delta'}*f_{X_{\rm corr}})(\xi) - (f_{\Delta+X_{\rm norm}}*f_{X_{\rm corr}})(\xi)| \\
+& = |\int_{-\infty}^{\infty}f_{\Delta'}(\xi-\tau)f_{X_{\rm corr}}(\tau)d\tau - \int_{-\infty}^{\infty}f_{\Delta + X_{\rm norm}} (\xi -\tau) f_{X_{\rm corr}} (\tau) d\tau| \\
+& = |\int_{-\infty}^{\infty} (f_{\Delta'}(\xi-\tau) - f_{\Delta + X_{\rm norm}} (\xi -\tau)) f_{X_{\rm corr}} (\tau) d\tau| \\
+& \leq \int_{-\infty}^{\infty} |f_{\Delta'}(\xi-\tau) - f_{\Delta + X_{\rm norm}} (\xi -\tau) |f_{X_{\rm corr}} (\tau) d\tau  \\
+& \leq \int_{-\infty}^{\infty} \eta f_{X_{\rm corr}} (\tau) d\tau \leq \eta
+\end{align*}
+
+where ``*'' represents the convolution operation.
+
+Thus, after convolution with $X_{\rm corr}$, the maximum gap between PDF of the $\Delta'$ and $\Delta + X_{\rm norm}$ decreases.
+Thus, $|F_{\Delta' + X_{\rm corr}} - F_{\Delta + X_{\rm norm} + X_{\rm corr}}| \leq |F_{\Delta'} - F_{\Delta + X_{\rm norm}}| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$.
+Thus, $|P_a - P_a'| = |\Pr (\Delta + X_{\rm norm} + X_{\rm corr} > 0) - \Pr (\Delta' + X_{\rm corr} > 0)|=|F_{\Delta + X_{\rm norm} + X_{\rm corr}}(0) - F_{\Delta' + X_{\rm corr}}(0)| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$.
+We can choose $n$ so that $\frac{C\rho}{\sigma^3\sqrt{n}} < \varepsilon$ to ensure $|P_a - P_a'| < \varepsilon$.
+Thus, the requirement for $n$ is $n \geq \frac{C^2\rho^2}{\sigma^6 \varepsilon^2}$.
 

### 2016-06-15 nips_2016.tex
updated a typo

@@
 Recall Theorem~\ref{thm:efficiency}, the number of data instance required is $\frac{4 \ell^2 C^2 \rho^2 }{\sigma^6 \varepsilon^2}$.
-This figure reveals that $\rho/\sigma^3$ is rougly a constant, which means the data required for each iteration is a constant value has nothing to do with the $N$.
+This figure reveals that $\rho/\sigma^3$ is rougly a constant, which means the data required for each iteration is a constant value which has nothing to do with the $N$.
 

### 2016-06-21 nips_2016.tex
update by JFC

@@
 
-\title{A Simple Minibatch Acceptance Test for MCMC}
+\title{An Efficient Minibatch Acceptance Test for MCMC}
 
@@
 to modeling very large datasets, where it is impractical to
-perform Metropolis-Hastings tests on the full data. To do this, we
-describe a novel acceptance test which is applied to minibatches of
-data. Previous work on minibatch Metropolis-Hastings use adaptive
-batch sizes, which may use far more than one minibatch of data and therefore
-yield few samples.  Here we describe a simpler test on fixed-sized
-minibatches. Instead of adaptive batch sizes, our test requires a
-precondition on the variance of the log proposal probability
-ratio. Given a measurement of this variance, a minibatch size can be
-chosen to support the test in one step, which yields a number of
-samples proportional to the batch/minibatch ratio. Alternatively, the
-variance condition can be used to adjust the temperature of the
-sampling distribution to a ``natural'' value to provide regular
-samples at a given minibatch size. The resulting proposal/test is only
-slightly more complex than a simple SGD update. In this paper we
+perform Metropolis-Hastings tests on the full data. Previous work on reducing
+the cost of Metropolis-Hastings tests yield variable data reductions
+and in the worst case require the entire dataset. 
+Here we present a simple test which requires constant size minibatches
+if an auxiliary condition is met. We describe several cases under which
+this condition can be satisfied. Our new test is naturally tolerant
+of ``noisy'' estimates of the transition probability ratio. To achieve
+this, we replace the classical Metropolis-Hastings test with a smooth
+acceptance function which is a sigmoid (a logistic function). 
+The resulting test can be combined with minibatch proposals to yield
+updates with the same complexity as a simple SGD update. In this paper we
 derive the test, discuss its implementation, and present several
@@
 Minibatch updates produce many improvements to the model for each pass
-over the dataset, and therefore have high sample efficiency. They also map very
-well onto hardware such as GPUs. In contrast, MCMC methods typically
-require calculations over the full dataset to produce a new sample. A
-recent result from~\cite{cutting_mh_2014} attempted to bridge the gap
-between minibatch optimization and MCMC by proposing an \emph{adaptive
-  minibatch} method. Their method uses a subset of the $N$ data points
-for the MH test each iteration. However the number of points is not
-known a priori, and is adjusted dynamically during every test. This
-undermines many of the practical advantages of minibatch calculations: 
-GPU kernels work best for certain minibatch sizes, memory blocks
-can be reused when minibatch size is fixed, etc.
-
-In this paper, we develop an alternative minibatch acceptance test for MCMC
-methods which uses only one fixed-size minibatch per test. Instead of
-dynamically adjusting the batch size, we add a prerequisite for the test
-based on the variance of the log acceptance probability. If this
-variance is small enough, the test requires only a single step. This
-variance changes slowly on typical problems, and we can estimate it
-before performing the test. The minibatch variance
-decreases as the minibatch size grows, so the variance measurement can be
-used to set the minimum minibatch size.
-
-There is another advantage when using our test for Bayesian posterior inference. As the dataset size
-grows, the posterior distribution sharpens and often concentrates to a few sharp peaks. Samples from
-this distribution will come from those peaks, and will easily become stuck in one of them. The
-samples are not exploring the posterior parameter space, and are unlikely to concentrate at a strong
-local optimum. They therefore fail to accurately represent a sample of the posterior because of
-inadequate search (i.e., mixing). We argue that it is much more natural to sample the distribution
-at a higher temperature determined by the minibatch size rather than the dataset size. Starting at a
-high temperature (small minibatch size) which mixes easily, one can reduce the temperature
-(increasing minibatch size) until one reaches the desired target distribution temperature.  Although
-this approach also involves changing minibatch size, the changes increase minibatch size
-monotonically, and only a few changes are performed in the course of a sampling session.  Annealing
-can also be used for MAP and ML estimation because changing temperature preserves posterior modes.
-
-To be precise, the contributions of this paper are as follows:
-
-\begin{enumerate}[noitemsep]
+over the dataset, and therefore have high sample efficiency. They also
+map very well onto hardware such as GPUs. In contrast, MCMC methods
+typically require calculations over the full dataset to produce a new
+sample.  Recent results
+from~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
+(bounded error) acceptance tests based on the classical M-H test. The
+tests depend on minibatch statistics, and on the value of an
+additional random variable. The amount of data consumed for each test
+varies significantly from one minibatch to the next, and
+non-repeatably on the random variable. By contrast,
+\cite{conf/uai/MaclaurinA14} performs exact tests but requires a lower
+bound on parameter distribution across its domain. The amount of data
+reduction depends on the accuracy of this bound, and such bounds are
+only available for relatively simple distributions.
+
+Here we derive a new test which incorporates the variability in
+minibatch statistics as {\em a natural part of the test}. Because of
+this, the amount of data required for each test is fixed with high
+probability and in expected value. Our test requires a bound on the
+(relative) variance of the log acceptance probability ratio. This variance
+changes slowly on typical problems, and we can estimate it before
+performing the test.
+
+If the variance test fails, there are three
+ways to mitigate the situation:
+\begin{itemize}
+
+\item Increase the temperature of the target distribution. Log likelihoods
+  scale as $1/T$, and so the variance of the likelihood ratio will
+  vary as $1/T^2$.
+
+\item Increase the minibatch size. Relative variance scales as $1/k$ vs the
+  minibatch size $k$.
+
+\item For continuous probability distributions, reduce the proposal step
+  size. Since the condition is on variance in the {\em difference} in
+  log probabilities of prior and proposed likelihoods, smaller proposal steps
+  will lead to smaller changes in probability and smaller changes in the
+  variance of those changes.
+\end{itemize}
+
+Note that changes to the distribution temperature may only need to be
+temporary.  The distribution can be annealed from high-to-low
+temperatures to maintain the variance condition. High temperature is
+natural with small minibatches and large step sizes, since the
+precision of the acceptance test is limited by the amount of data
+available. As the distribution reaches a stationary state and is cooled,
+particle motion can be slowed with larger minibatches or smaller proposal
+steps. The contributions of this paper are as follows:
+
+\begin{itemize}
 \item We develop a new minibatch acceptance test which satisfies detailed balance.
-    \item We compare performance of our new test and the adaptive sampling method.
+\item we developed effectively measurable bounds for both absolute and relative
+  test error. 
+\item We compare performance of our new test and prior approaches. 
 \item We experiment using the approach for posterior estimation.
-\end{enumerate}
+\end{itemize}
 
@@
 $p(\theta \mid x_1, \ldots, x_N)$.  To do so, one generates a chain of (correlated) samples
-$\theta_1, \ldots, \theta_T$ for large $T$ and approximates $p$ using the sample counts.  Each
+$\theta_1, \ldots, \theta_T$ for large $T$ and approximates $p$ using the samples.  Each
 iteration $t$ has a current $\theta_t$, and a \emph{proposal distribution} $q(\theta' \mid
@@
 
-Unfortunately, computing $f$ requires the use of all $N$ training data points. Moreover, it is
-difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
-balance. To reconcile these competing objectives,~\cite{cutting_mh_2014} proposes an adaptive
-minibatch MCMC algorithm which uses a sequential hypothesis test. During each iteration, the
-algorithm starts with a small minibatch of data and tests the hypothesis that the sample $\theta'$
-should be accepted or rejected based on a probability ratio. If the test cannot make a decision
-over a certain confidence threshold, then the minibatch size is increased and the test repeats. This
-process continues until a decision. The downside of this algorithm is the need to keep incrementing
-the minibatch size; in the worst case, all $N$ data points may be needed to decide on $\theta'$.
-We prove that the lower bound of average number of data instance consumed for one iteration is $O(\sqrt{N})$, where $N$ is the number of all data points, see Lemma~\ref{lem:efficienty_cutting}.
-A
-similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties
-based on concentration inequalities, but comes at the cost of more computational time.
+Unfortunately, computing $f$ requires the use of all $N$ training data
+points. Moreover, it is difficult to design tests using substantially
+fewer than $N$ points that also satisfy detailed balance. To reconcile
+these competing
+objectives,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
+approximate Metropolis-Hasting tests using sequential hypothesis
+testing. During each iteration, they start with a small minibatch of
+data and tests the hypothesis that the sample $\theta'$ should be
+accepted or rejected based on a probability ratio involving the
+additional random variable $u\in [0,1]$. If the test cannot make a
+decision with sufficient confidence, then the minibatch size is
+increased and the test repeats. This process continues until a
+decision. The bounds depend on either a central limit theorem
+\cite{cutting_mh_2014} or a concentration bound
+\cite{icml2014c1_bardenet14} requiring direct bounds on the log
+likelihood ratio.  The problem with both methods is the expense of
+resolving small differences.  In the worst case, all $N$ data points
+may be needed, and as we show next this worst case can occur
+about $\sqrt{N}$ times during the performance of $N$ tests.
+
+We next prove that for some simple distributions, the lower
+bound of average number of data instance consumed for one iteration of
+~\cite{cutting_mh_2014,icml2014c1_bardenet14} is
+$\Omega(\sqrt{N})$, where $N$ is the number of data points, see
+Lemma~\ref{lem:efficienty_cutting}.
 
 \begin{lemma}\label{lem:efficienty_cutting}
-If the sample mean $\bar{l}$ has uniform distribution from $[-L,L]$, the lower bound of expected data consumed of adaptive sampling method for one iteration is $\frac{\eta s_{l}Q}{2L}\sqrt{N}$, where $\bar{l} = \frac{1}{n}\sum_{i=1}^{n} \left (\log p(x_i; \theta') - \log p(x_i| \theta_t)\right )$, $\eta = \phi^{-1}_{n-1}(1-\varepsilon)$, $Q=\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')}$, $s_l$ is the population standard deviation of $l_i$, $\phi_{n-1}()$ is the CDF of t distribution with degree of freedom $n-1$, and $n$ is the minibatch size.
+If the sample mean $\bar{l}$ has uniform distribution on $[-L,L]$,
+the lower bound of expected data consumed of adaptive sampling method
+for one iteration is $\frac{\eta s_{l}Q}{2L}\sqrt{N}$, where $\bar{l}
+= \frac{1}{n}\sum_{i=1}^{n} \left (\log p(x_i; \theta') - \log p(x_i|
+\theta_t)\right )$, $\eta = \phi^{-1}_{n-1}(1-\varepsilon)$,
+$Q=\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')}$,
+$s_l$ is the population standard deviation of $l_i$, $\phi_{n-1}()$ is
+the CDF of t distribution with degree of freedom $n-1$, and $n$ is the
+minibatch size.
 \end{lemma}
 
-The work of~\cite{conf/uai/MaclaurinA14} presents another minibatch MCMC method which gets the
-\emph{exact} posterior, so long as there exists a (cheap) lower bound on the likelihood for each
-data point. Their approach is only a starting point, however, and it can be difficult to derive
-these lower bounds in general cases. The extra auxiliary variables they introduce also hurt mixing.
-
-There is a separate line of MCMC work on simulating the physics of a probability distribution. By
-viewing random variables as particles in a system, one can apply Hamiltonian Monte Carlo
-(HMC)~\cite{mcmc_hamiltonian_2010} methods which generate the best case scenario for proposals: high
-quality \emph{and} distant.  HMC methods require a full gradient computation of the posterior, which
-means they also face the problem of computing the likelihood at each point. Stochastic versions of
-HMC~\cite{sghmc_2014,stochastic_thermostats_2014} use a minibatch of data each iteration but require
-controlling the minibatch noise to avoid divergence. Langevin
-Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} is a similar approach which does not use the
-momentum terms in HMC.  Minibatch SGHMC and Langevin Dynamics are orthogonal to our work and are
-other ways of attempting to solve the same problem of MCMC methods on big data. We can combine our
-methods with these results, as we demonstrate in Section~\ref{ssec:nets}.
+\begin{proof} see appendix
+  \end{proof}
+There is a separate line of MCMC work drawing principles from
+statistical physics. By viewing random variables as particles in a
+system, one can apply Hamiltonian Monte Carlo
+(HMC)~\cite{mcmc_hamiltonian_2010} methods which generate high
+acceptance \emph{and} distant proposals when run on full batches of
+data. Recently Langevin
+Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has been applied to
+Bayesian estimation on minibatches of data. This simplified dynamics
+uses local proposals and avoids MH tests by using small proposal steps
+whose acceptance approaches 1 in the limit. However, the constraint on
+proposal step size is severe, and the state space exploration reduces
+to a random walk. Full minibatch HMC for minibatches was recently
+described in ~\cite{sghmc_2014} which allows momentum-augmented
+proposals with larger step sizes. However, step sizes are still
+limited by the need to run accurately without MH tests.
+% \cite{stochastic_thermostats_2014}
+Our work opens the door to applying those methods with much
+more agressive step sizes without loss of accuracy. We demonstrate
+this in Section~\ref{ssec:nets}.
 
@@
 \begin{algorithm}[t]
-\Input{Number of samples $T$, number of $\Delta'$ estimates $K$, minibatch size $m$, pre-computed $X_{\rm
-corr}$ distributions, $\{x_i\}_{i=1}^N$ values from a target distribution, and initial sample $\theta_0$.}
+\Input{Number of samples $T$, minibatch size $m$, pre-computed $X_{\rm
+corr}$ distribution, and initial sample $\theta_0$.}
 \Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ for either posterior inference or
@@
 \For{$t=\{1, \ldots, T\}$}{
-    Propose a candidate $\theta'$\;
-    Compute the $\Delta'$ from one minibatch data\;
-    Compute the ${\rm std}(\Delta')$ from another minibatch data\;
-    \While {${\rm std}(d) \ge 1.2$} {
-    	Update the value of $\Delta'$ and ${\rm std}(\Delta')$ from one new minibatch\;
+    Propose a candidate $\theta'$\ from proposal distribution\;
+    Draw a minibatch of $m$ samples, compute $\Delta'(\theta',\theta_t)$ and ${\rm std}(\Delta')$ for this sample\;
+    \While {${\rm std}(\Delta') \ge 1.2$} {
+    	Draw $m$ more samples to augment the current minibatch, and update $\Delta'$ and ${\rm std}(\Delta')$ \;
     }
-
-        Choose the closest $X_{\rm corr}$ distribution, and sample a value $X_c$ from it\;
-
-        \eIf{$X_c + \Delta' > 0$}{
+    Let $\sigma = \sqrt{1.2 - {\rm std}(\Delta')^2}$, draw a normal variable $X_n \in Norm(0,\sigma)$ and
+    $X_c$ from the correction distribution\;
+
+        \eIf{$X_c + X_n + \Delta' > 0$}{
             Accept the candidate, $\theta_{t+1} = \theta'$\;
         }{
-            Reject and re-use the old one, $\theta_{t+1} = \theta_t$\;
+            Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
         }

### 2016-06-21 nips_2016.tex
correct typos

@@
 \begin{itemize}
-\item We develop a new minibatch acceptance test which satisfies detailed balance.
+\item We developed a new minibatch acceptance test which satisfies detailed balance.
 \item we developed effectively measurable bounds for both absolute and relative
   test error. 
-\item We compare performance of our new test and prior approaches. 
-\item We experiment using the approach for posterior estimation.
+\item We compared performance of our new test and prior approaches. 
+\item We did experiments using the approach for posterior estimation.
 \end{itemize}
@@
 Our work opens the door to applying those methods with much
-more agressive step sizes without loss of accuracy. We demonstrate
+more aggressive step sizes without loss of accuracy. We demonstrate
 this in Section~\ref{ssec:nets}.
@@
 With help of temperature (see Section~\ref{ssec:discussion}), we will show that usually the mean of minibatch consumed per iteration is below 2 in Section~\ref{sec:experiments}.
-Later our efficiency analysis in Section~\ref{sec:theory} will reveal that the number of data instance consumed by our algorithm is much fewer than that of exisitng algorithms.
+Later our efficiency analysis in Section~\ref{sec:theory} will reveal that the number of data instance consumed by our algorithm is much fewer than that of existing algorithms.
 
@@
 \begin{theorem}\label{thm:efficiency}
-We require $\frac{ C^2 \rho^2 }{\sigma^6 \varepsilon^2}$ minibatch size to ensure $|P_a -P_a'| < \varepsilon$, where $C$ is a positive constant which is smaller than 0.4748, $\sigma^2 $ and $\rho$ is the second and third moment of $(\nabla\log \Pr(x_i|\theta_t) - \nabla\log\Pr(x_i|\theta')) - E\left[ \nabla\log \Pr(x_i|\theta_t) - \nabla\log\Pr(x_i|\theta')\right]$, $P_a$ is the accpet rate from the complete data, $P_a'$ is the accpet rate from minibatch data by our algorithm, $\theta_t$ is the current parameter, and $\theta'$ is the proposed parameter.
+We require $\frac{ C^2 \rho^2 }{\sigma^6 \varepsilon^2}$ minibatch size to ensure $|P_a -P_a'| < \varepsilon$, where $C$ is a positive constant which is smaller than 0.4748, $\sigma^2 $ and $\rho$ is the second and third moment of $(\nabla\log \Pr(x_i|\theta_t) - \nabla\log\Pr(x_i|\theta')) - E\left[ \nabla\log \Pr(x_i|\theta_t) - \nabla\log\Pr(x_i|\theta')\right]$, $P_a$ is the accept rate from the complete data, $P_a'$ is the accept rate from minibatch data by our algorithm, $\theta_t$ is the current parameter, and $\theta'$ is the proposed parameter.
 
@@
 just treat the $\Delta'$ which comes from minibatch as $\Delta$ which comes from the complete data) reveals wider spread distribution than the posterior.
-From Figure~\ref{fig:gauss_mix_1} we illustrate our proposed method can obtain the samples which have almost the same distirbution as the standard MH test from the complete data, adaptive minibatch test, and the true posterior.
-Meanwhile, we show the benefits of our minibatch method compared with the navie minibatch method.
+From Figure~\ref{fig:gauss_mix_1} we illustrate our proposed method can obtain the samples which have almost the same distribution as the standard MH test from the complete data, adaptive minibatch test, and the true posterior.
+Meanwhile, we show the benefits of our minibatch method compared with the naive minibatch method.
 
@@
 The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
-\ge 1.2$ for only consuming one mimnibatch data. For six minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
+\ge 1.2$ for only consuming one minibatch data. For six minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
 our MH test five times and averaged the number of ``violations.'' As the size increases, the
@@
 
-Figure~\ref{fig:gauss_mix_3} (to the left) shows a historgram of ${\rm std}(\Delta')$.
+Figure~\ref{fig:gauss_mix_3} (to the left) shows a histogram of ${\rm std}(\Delta')$.
 Figure~\ref{fig:gauss_mix_3} (to the middle) shows the confidence interval of our estimated ${\rm std}(\Delta')$ at each iteration \footnote{In order to show the confidence interval clearly, we use moving smooth to smooth the confidence interval, and plot each value by interval of 50 iterations}.
-The blue line represents the standard deviation comes from the complete data, which can be treated as the ``true'' ${\rm std}(\Delta')$, the green line represents our estimation of ${\rm std}(\Delta')$ from the minibatch, and the two red dash lines represent the confidence interval of ${\rm std}(\Delta')$ by adding / subtracking one ${\rm std(std(}\Delta'))$.
+The blue line represents the standard deviation comes from the complete data, which can be treated as the ``true'' ${\rm std}(\Delta')$, the green line represents our estimation of ${\rm std}(\Delta')$ from the minibatch, and the two red dash lines represent the confidence interval of ${\rm std}(\Delta')$ by adding / subtracting one ${\rm std(std(}\Delta'))$.
 This figure indicates that our estimation of ${\rm std}(\Delta')$ from one minibatch data is accurate.
@@
 Recall Theorem~\ref{thm:efficiency}, the number of data instance required is $\frac{4 \ell^2 C^2 \rho^2 }{\sigma^6 \varepsilon^2}$.
-This figure reveals that $\rho/\sigma^3$ is rougly a constant, which means the data required for each iteration is a constant value which has nothing to do with the $N$.
+This figure reveals that $\rho/\sigma^3$ is roughly a constant, which means the data required for each iteration is a constant value which has nothing to do with the $N$.
 
@@
 
-If the minibatch size is $n$, we fisrt compute the probability that more than $n$ data instances are required for adaptive sampling method for one iteration. We denote this probaility $P_{rej}$.
+If the minibatch size is $n$, we first compute the probability that more than $n$ data instances are required for adaptive sampling method for one iteration. We denote this probability $P_{rej}$.
 \[P_{rej} = \Pr \left ( 1-\phi_{n-1} \left( \frac{|\bar{l} - \frac{1}{N} \log (u\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')})|}{s} \right )  \geq \varepsilon \right ) \]

### 2016-06-21 nips_2016.tex
updated MH_test

--- 2016-06-21 nips_2016.tex
@@
 corr}$ distribution, and initial sample $\theta_0$.}
-\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ for either posterior inference or
+\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ from $p(\theta)$\;
 MAP/ML estimation.}
@@
     Draw a minibatch of $m$ samples, compute $\Delta'(\theta',\theta_t)$ and ${\rm std}(\Delta')$ for this sample\;
-    \While {${\rm std}(\Delta') \ge 1.2$} {
+    \While {${\rm std}(\Delta') \ge 1$} {
     	Draw $m$ more samples to augment the current minibatch, and update $\Delta'$ and ${\rm std}(\Delta')$ \;
     }
-    Let $\sigma = \sqrt{1.2 - {\rm std}(\Delta')^2}$, draw a normal variable $X_n \in Norm(0,\sigma)$ and
+    Let $\sigma = \sqrt{1 - {\rm std}(\Delta')^2}$, draw a normal variable $X_n \in Norm(0,\sigma)$ and
     $X_c$ from the correction distribution\;

### 2016-07-05 nips_2016.tex
fixed lower bound

@@
 reject them based on an acceptance test. The acceptance test is
-usually a Metropolis-Hastings
+usually a Metropolis
 test~\cite{Metropolis1953,hastings70}. For example, for Bayesian
@@
 
-In standard MCMC methods~\cite{gilks1996markov,brooks2011handbook} for Bayesian inference with
-parameter $\theta$ and conditionally independent data, the goal is to compute the distribution
-$p(\theta \mid x_1, \ldots, x_N)$.  To do so, one generates a chain of (correlated) samples
-$\theta_1, \ldots, \theta_T$ for large $T$ and approximates $p$ using the samples.  Each
-iteration $t$ has a current $\theta_t$, and a \emph{proposal distribution} $q(\theta' \mid
-\theta_t)$ determines a new candidate $\theta'$. With probability $P_a$, the sample is accepted, so
-$\theta_{t+1} = \theta'$. Otherwise, $\theta_{t+1} = \theta_t$. This is done by drawing a uniform
-random variable $u \sim {\rm Unif}[0,1]$ and accepting if $u < P_a$. Traditionally, $P_a$ is
+In standard MCMC methods~\cite{gilks1996markov,brooks2011handbook}, a difficult-to-compute
+probability distribution $p(\theta)$ is sampled using a Markov chain $\theta_1,\ldots,\theta_n$. The 
+sample $\theta_{t+1}$ at time $t+1$ is generated using a candidate $\theta'$ from a (simpler) proposal
+distribution $q(\theta'\mid \theta_t)$, filtered by an acceptance test. The acceptance test is usually
+a Metropolis test. The Metropolis test has acceptance probability:
 \begin{equation}\label{eq:traditional}
-P_a = \min\left\{ 1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid \theta_t)}
-\right\} = \min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
-\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid
-\theta_t)} \right\},
-\end{equation}
-where $f(\theta_t)=p(\theta_t \mid x_1,\ldots,x_N)$. This $P_a$ satisfies detailed balance, so if
-one samples long enough, one will arrive at a stationary distribution matching the posterior, though
-a burn-in period and/or only using samples at regular intervals is often done in practice.
-
-Unfortunately, computing $f$ requires the use of all $N$ training data
-points. Moreover, it is difficult to design tests using substantially
-fewer than $N$ points that also satisfy detailed balance. To reconcile
-these competing
+\alpha(\theta',\theta_t) = \min\left\{ 1, \frac{p(\theta')q(\theta_t \mid \theta')}{p(\theta_t)q(\theta' \mid \theta_t)}
+\right\}
+\end{equation}
+and with probability $\alpha(\theta',\theta_t)$, we accept $\theta'$
+and set $\theta_{t+1} = \theta'$, otherwise set
+$\theta_{t+1}=\theta_t$.  The test is often implemented with an
+auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison $u < \alpha(\theta',\theta_t)$
+
+The acceptance test guarantees detailed balance, so that
+\begin{equation}
+  p(\theta)q(\theta',\theta)\alpha(\theta',\theta) = p(\theta')q(\theta,\theta')\alpha(\theta,\theta')
+\end{equation}
+The detailed balance condition, together with ergodicity, guarantee that the
+Markov chain has a unique stationary distribution $\pi(\theta) = p(\theta)$.
+
+For Bayesian inference, we would like to sample from a parameter distribution for $\theta$ based on some
+observed data $x_1,\ldots,x_N$. i.e. we want to sample from 
+$p(\theta \mid x_1, \ldots, x_N)$. The acceptance probability is now:
+\begin{equation}
+\alpha(\theta',\theta_t) = 
+\min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
+\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid\theta_t)} \right\},
+\end{equation}
+where $p(\theta)$ is a prior, and $p(x_i \mid \theta)$ are the
+probabilities of the observations. Computing samples this way requires
+the use of all $N$ training data points. Moreover, it is difficult to
+design tests using substantially fewer than $N$ points that also
+satisfy detailed balance. To reconcile these competing
 objectives,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
@@
 testing. During each iteration, they start with a small minibatch of
-data and tests the hypothesis that the sample $\theta'$ should be
-accepted or rejected based on a probability ratio involving the
-additional random variable $u\in [0,1]$. If the test cannot make a
-decision with sufficient confidence, then the minibatch size is
-increased and the test repeats. This process continues until a
-decision. The bounds depend on either a central limit theorem
-\cite{cutting_mh_2014} or a concentration bound
-\cite{icml2014c1_bardenet14} requiring direct bounds on the log
-likelihood ratio.  The problem with both methods is the expense of
-resolving small differences.  In the worst case, all $N$ data points
-may be needed, and as we show next this worst case can occur
-about $\sqrt{N}$ times during the performance of $N$ tests.
-
-We next prove that for some simple distributions, the lower
+data and test the hypothesis that the sample $\theta'$ based on an
+approximate version of the test $u < \alpha(\theta',\theta_t)$. If the
+approximate test cannot make a decision with sufficient confidence,
+then the minibatch size is increased and the test repeats. This
+process continues until a decision. The bounds depend on either a
+asymptotic central limit theorem \cite{cutting_mh_2014} or a
+concentration bound \cite{icml2014c1_bardenet14}, the latter requiring
+direct bounds on the log likelihood ratio.  The problem with both
+methods is the expense of resolving small differences.  In the worst
+case, all $N$ data points may be needed, and as we show next this
+worst case can occur about $\sqrt{N}$ times during the performance of
+$N$ tests.
+
+Following ~\cite{icml2014c1_bardenet14}, we write the test $u < \alpha(\theta',\theta_t)$ in
+the equivalent form $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$ where
+\begin{equation}
+\Lambda(\theta,\theta') = \frac{1}{N}\sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+\hspace{0.2in} {\rm and} \hspace{0.2in}
+\psi(u,\theta,\theta') = \frac{1}{N}\log\left(u\frac{q(\theta'|\theta)p(\theta)}{q(\theta|\theta')p(\theta')}\right)
+\end{equation}
+
+By the central limit theorem, we expect $\Lambda(\theta_t,\theta')$ to be approximately
+Normal. To reduce computational effort, an unbiased estimate of $\Lambda(\theta,\theta')$
+based on a minibatch can be used:
+
+\begin{equation}
+\Lambda^*(\theta,\theta') = \frac{1}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+\end{equation}
+
+The acceptance test then becomes a statistical test of the hypothesis
+that $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$ by
+establishing that $\Lambda^*(\theta,\theta')$ is substantially larger
+than $\psi(u,\theta,\theta')$.  In ~\cite{cutting_mh_2014} an
+asymptotic central limit argument was used to derive this gap, while
+in \cite{icml2014c1_bardenet14} a concentration bound was used. In
+both cases, the resulting tests were shown to give useful reductions
+in number of samples required over using the full dataset. But no
+worst-case bounds were given on average number of samples needed.
+
+We next show that for some simple distributions, the lower
 bound of average number of data instance consumed for one iteration of
-~\cite{cutting_mh_2014,icml2014c1_bardenet14} is
-$\Omega(\sqrt{N})$, where $N$ is the number of data points, see
-Lemma~\ref{lem:efficienty_cutting}.
+~\cite{cutting_mh_2014} and ~\cite{icml2014c1_bardenet14} is
+$\Omega(\sqrt{N})$, where $N$ is the number of data points.
+
+Consider a simple model where samples $x_i$ for $i=1,\ldots,N$ are
+drawn from a Normal distribution $\mathcal{N}(\theta,1)$ where
+$\theta=0.5$, and with a uniform prior on $\theta$. The log likelihood ratio is
+\begin{equation}
+\Lambda^*(\theta,\theta') =  \frac{1}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta_t)}=
+  %%\frac{1}{2}\left( (x_i-\theta')^2 - (x_i - \theta_t)^2\right) =
+  (\theta'-\theta_t)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta_t-\frac{\theta'-\theta_t}{2}\right)
+\end{equation}
+which is normally distributed over selection of the Normal samples $x_i$.
+Since the $x_i$ have unit variance, their mean has variance $1/b$, and the variance $\sigma^2(\Lambda^*)$
+of $\Lambda^*(\theta,\theta')$ is $(\theta'-\theta_t)^2/b$. In order to pass
+a hypothesis test that $\Lambda > \psi$, we will need to establish a
+large enough gap (several $\sigma(\Lambda^*)$) between $\Lambda^*(\theta',\theta_t)$
+and $\psi(u,\theta',\theta_t)$.
+
+We would like to sample from the posterior distribution of $\theta$
+which is a normal distribution centered on the sample mean, and with
+variance $1/N$.  The target distribution is therefore
+$\mathcal{N}(\mu,1/N)$ where $\mu$ is the sample mean.  An efficient
+proposal distribution has the same variance as the target distribution
+$\mathcal{N}(\mu,1/N)$, so we take $q(\theta'\mid\theta_t) =
+\phi((\theta'-\theta_t)N)$, where $\phi(x)$ is a Normal density
+function with zero mean and unit variance. This proposal is symmetric
+$q(\theta',\theta_t)=q(\theta_t,\theta')$, and we assumed uniform
+prior for $\theta$, so $\psi(u,\theta',\theta_t)$ reduces to $1/N\log
+u$.
 
 \begin{lemma}\label{lem:efficienty_cutting}
-If the sample mean $\bar{l}$ has uniform distribution on $[-L,L]$,
-the lower bound of expected data consumed of adaptive sampling method
-for one iteration is $\frac{\eta s_{l}Q}{2L}\sqrt{N}$, where $\bar{l}
-= \frac{1}{n}\sum_{i=1}^{n} \left (\log p(x_i; \theta') - \log p(x_i|
-\theta_t)\right )$, $\eta = \phi^{-1}_{n-1}(1-\varepsilon)$,
-$Q=\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')}$,
-$s_l$ is the population standard deviation of $l_i$, $\phi_{n-1}()$ is
-the CDF of t distribution with degree of freedom $n-1$, and $n$ is the
-minibatch size.
+  For the model above, there
+  exists a fixed constant $c$ such that with probability $ \geq c$ over the joint
+  distribution of $(\theta_t, \theta', u)$, the
+  test from \cite{icml2014c1_bardenet14} consumes all $N$ samples. 
 \end{lemma}
 
-\begin{proof} see appendix
+\begin{proof} Assume $(\theta' - \theta_t) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and
+  $\theta_t -0.5 \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events
+  occur with probability $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and $p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively. 
+  Choose $u_0$
+  to satisfy
+  \begin{equation}
+    \log u_0 = -N(\theta'-\theta_t)\left(\theta_t-0.5+\frac{\theta'-\theta_t}{2}\right) 
+  \end{equation}
+  and rewrite the test now as
+  \begin{equation}
+\frac{1}{b}\sum_{i=1}^b x_i-\left(\theta_t+\frac{\theta'-\theta_t}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) > 0
+  \end{equation}
+  The above choice of $u_0$ ensures that the terms in parenthesis above sum to 0.5. Since the $x_i$ have mean 0.5, the
+  test will never correctly succeed.
+
+  Furthermore, if we sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be sufficiently different
+  from 0.5 to allow the test to succeed.
+
+ The choices above for $\theta_t$ and $\theta'$ guarantee that
+  \begin{equation}
+    \log u_0 \in -[0.5,1][0.75,1.5] = -[0.375,1.5]
+        \end{equation}
+and consider the range of $u$ values
+  \begin{equation}
+    \log u \in \log u_0 + [-1,0.375]
+  \end{equation}
+  the size of this range is at least $\exp([-2.5,-1.125]) = [0082,0.286]$ and occurs
+  with probability at least $p_2=0.2044$. The samples $\theta_t$, $(\theta'-\theta_t)$
+  and $u$ are drawn independently and so the total probability of these events is
+  $c = p_0 p_1 p_2 = 0.009$.
 \end{proof}
@@
 described in ~\cite{sghmc_2014} which allows momentum-augmented
+
 proposals with larger step sizes. However, step sizes are still
@@
 
-\section{Theoretical Results}\label{sec:theory}
-
-In this section, we firstly explore the convergence properties of our MH test. Secondly, we compare the efficiency between our MH test and adaptive MH test. Due to space constraints, all
-our proofs are in Appendix~\ref{app:proofs}.
-
-We first introduce some notation. Let our true and approximated acceptance probabilities be $P_a =
-g(\Delta) = \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} > 0)$.  Define $X_{\xi} =
-\varepsilon - X_{\rm norm}$ which means, invoking Equations~\ref{eq:relationship}
-and~\ref{eq:deconvolution}, that $X_\xi$ is the random variable characterizing the accuracy of our
-$X_{\rm norm}$.  Thus, $\Delta' = \Delta + X_{\rm norm} + X_\xi$. We assume the standard deviation
-of $X_{\rm norm}$ is less than $\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}.
+\section{Analysis}\label{sec:theory}
+
+In this section, we first perform an error analysis for our new
+acceptance test. Second, we follow the implications of this analysis for accuracy of
+the estimated target distribution. Third, we compare the speed and accuracy of
+our test with prior work. Due to space constraints, all our proofs
+are in Appendix~\ref{app:proofs}.
+
+We first introduce some notation. Let our true (full dataset) and
+approximate (minibatch) acceptance probabilities be $P_a = g(\Delta) =
+\Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} > 0)$.
+Define $X_{\xi} = \varepsilon - X_{\rm norm}$ which means, invoking
+Equations~\ref{eq:relationship} and~\ref{eq:deconvolution}, that
+$X_\xi$ is the random variable characterizing the accuracy of our
+$X_{\rm norm}$.  Thus, $\Delta' = \Delta + X_{\rm norm} + X_\xi$. We
+assume the standard deviation of $X_{\rm norm}$ is less than
+$\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}.
 We denote $F_X(x)$ as the CDF of $X$.

### 2016-07-05 nips_2016.tex
updates to first section

@@
 Markov chain Monte Carlo (MCMC) sampling is a powerful method for
-computation on intractable distributions. MCMC algorithms propose
-samples from a proposal distribution $q$ which is in general different
-from the target distribution $p$, and decide whether to accept or
-reject them based on an acceptance test. The acceptance test is
-usually a Metropolis
-test~\cite{Metropolis1953,hastings70}. For example, for Bayesian
-posterior inference, the posterior distribution $p(\theta \mid x_1,
-\ldots, x_N) \propto p(\theta)\prod_{i=1}^N p(x_i \mid \theta)$ of
-parameter $\theta$ based on conditionally independent data
-$\{x_i\}_{i=1}^N$ is the target distribution which one aims to
-approximate by sampling $\theta$ values. When $N$ is large, it is
-very expensive to compute the exact posterior because the
-likelihood must be evaluated on every $x_i$.
+computation on intractable distributions. We are interested primarily
+in large dataset applications, where the goal is to sample a posterior
+distribution $p(\theta \mid x_1, \ldots, x_N)$ of parameter $\theta$,
+and where the number of data instances $N$ is large.  The
+Metroplis-Hastings method (M-H) generates sample candidates from a
+proposal distribution $q$ which is in general different from the
+target distribution $p$, and decides whether to accept or reject them
+based on an acceptance test. The acceptance test is usually a
+Metropolis test~\cite{Metropolis1953,hastings70}. Conventional
+Metropolis-Hastings requires all $N$ data instances to generate one
+posterior sample.
 
@@
 Minibatch updates produce many improvements to the model for each pass
-over the dataset, and therefore have high sample efficiency. They also
-map very well onto hardware such as GPUs. In contrast, MCMC methods
-typically require calculations over the full dataset to produce a new
-sample.  Recent results
-from~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
-(bounded error) acceptance tests based on the classical M-H test. The
-tests depend on minibatch statistics, and on the value of an
-additional random variable. The amount of data consumed for each test
-varies significantly from one minibatch to the next, and
-non-repeatably on the random variable. By contrast,
+over the dataset, and have high sample efficiency. They also
+map very well onto hardware such as GPUs. In contrast, M-H requires
+calculations over the full dataset to produce a new sample.  Recent
+results from~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
+approximate (bounded error) acceptance tests using subsets
+(minibatches) of the full dataset. The tests depend on minibatch
+statistics, and on the value of an additional random variable $u$. The
+amount of data consumed for each test varies significantly from one
+minibatch to the next, and depends on the current sample, the proposed
+sample, and on the random variable $u$. By contrast,
 \cite{conf/uai/MaclaurinA14} performs exact tests but requires a lower
@@
 probability and in expected value. Our test requires a bound on the
-(relative) variance of the log acceptance probability ratio. This variance
-changes slowly on typical problems, and we can estimate it before
-performing the test.
-
-If the variance test fails, there are three
-ways to mitigate the situation:
+(relative) variance of the log acceptance probability. This variance
+is small when the proposal distribution is efficient, that is when samples
+diffuse at an optimal rate.
+
+The variance of the log acceptance probability can be controlled in
+three ways, and these allow the test to be applied to modest-sized
+minibatches. The first two allow the test to be applied when the
+minibatch size is fixed a-priori.
+
 \begin{itemize}
@@
 
+\item For continuous probability distributions, reduce the proposal
+  step size and variance (for stochastic proposals).
+
 \item Increase the minibatch size. Relative variance scales as $1/k$ vs the
-  minibatch size $k$.
-
-\item For continuous probability distributions, reduce the proposal step
-  size. Since the condition is on variance in the {\em difference} in
-  log probabilities of prior and proposed likelihoods, smaller proposal steps
-  will lead to smaller changes in probability and smaller changes in the
-  variance of those changes.
+  minibatch size $k$. Increased minibatch size also reduces the error of
+  the test.
+
 \end{itemize}
 
-Note that changes to the distribution temperature may only need to be
-temporary.  The distribution can be annealed from high-to-low
-temperatures to maintain the variance condition. High temperature is
-natural with small minibatches and large step sizes, since the
-precision of the acceptance test is limited by the amount of data
-available. As the distribution reaches a stationary state and is cooled,
-particle motion can be slowed with larger minibatches or smaller proposal
-steps. The contributions of this paper are as follows:
+The contributions of this paper are as follows:
 
 \begin{itemize}
-\item We developed a new minibatch acceptance test which satisfies detailed balance.
-\item we developed effectively measurable bounds for both absolute and relative
-  test error. 
-\item We compared performance of our new test and prior approaches. 
-\item We did experiments using the approach for posterior estimation.
+\item We develop a new, more efficient (in samples per test) minibatch acceptance test which incorporates
+online  estimates of its error probability.
+\item We analyze the test for accuracy and speed.
+\item We compared performance of our new test and prior approaches on several datasets. We demonstrate the test is several orders
+  of magnitude more efficient than prior work measured as data consumed per test, and that it does not suffer
+  from long-tailed data requirements (up to the dataset size). 
 \end{itemize}
@@
 
-In standard MCMC methods~\cite{gilks1996markov,brooks2011handbook}, a difficult-to-compute
+In the Metropolis-Hastings method~\cite{gilks1996markov,brooks2011handbook}, a difficult-to-compute
 probability distribution $p(\theta)$ is sampled using a Markov chain $\theta_1,\ldots,\theta_n$. The 
@@
 \begin{equation}\label{eq:traditional}
-\alpha(\theta',\theta_t) = \min\left\{ 1, \frac{p(\theta')q(\theta_t \mid \theta')}{p(\theta_t)q(\theta' \mid \theta_t)}
+\alpha(\theta_t,\theta') = \min\left\{ 1, \frac{p(\theta')q(\theta_t \mid \theta')}{p(\theta_t)q(\theta' \mid \theta_t)}
 \right\}
 \end{equation}
-and with probability $\alpha(\theta',\theta_t)$, we accept $\theta'$
+and with probability $\alpha(\theta_t,\theta')$, we accept $\theta'$
 and set $\theta_{t+1} = \theta'$, otherwise set
 $\theta_{t+1}=\theta_t$.  The test is often implemented with an
-auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison $u < \alpha(\theta',\theta_t)$
+auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison $u < \alpha(\theta_t,\theta')$
 
@@
 \begin{equation}
-  p(\theta)q(\theta',\theta)\alpha(\theta',\theta) = p(\theta')q(\theta,\theta')\alpha(\theta,\theta')
+  p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
 \end{equation}
@@
 \begin{equation}
-\alpha(\theta',\theta_t) = 
+\alpha(\theta_t,\theta') = 
 \min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
@@
 data and test the hypothesis that the sample $\theta'$ based on an
-approximate version of the test $u < \alpha(\theta',\theta_t)$. If the
+approximate version of the test $u < \alpha(\theta_t,\theta')$. If the
 approximate test cannot make a decision with sufficient confidence,
@@
 case, all $N$ data points may be needed, and as we show next this
-worst case can occur about $\sqrt{N}$ times during the performance of
+worst case can occur about $\Omega(N)$ times during the performance of
 $N$ tests.
 
-Following ~\cite{icml2014c1_bardenet14}, we write the test $u < \alpha(\theta',\theta_t)$ in
-the equivalent form $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$ where
-\begin{equation}
-\Lambda(\theta,\theta') = \frac{1}{N}\sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+Following ~\cite{icml2014c1_bardenet14}, we write the test $u < \alpha(\theta_t,\theta')$ in
+the equivalent form $\Lambda(\theta_t,\theta') > \psi(u,\theta_t,\theta')$ where
+\begin{equation}
+\Lambda(\theta_t,\theta') = \frac{1}{N}\sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta_t)}\right)  
 \hspace{0.2in} {\rm and} \hspace{0.2in}
-\psi(u,\theta,\theta') = \frac{1}{N}\log\left(u\frac{q(\theta'|\theta)p(\theta)}{q(\theta|\theta')p(\theta')}\right)
+\psi(u,\theta_t,\theta') = \frac{1}{N}\log\left(u\frac{q(\theta'|\theta_t)p(\theta_t)}{q(\theta_t|\theta')p(\theta')}\right)
 \end{equation}
@@
 By the central limit theorem, we expect $\Lambda(\theta_t,\theta')$ to be approximately
-Normal. To reduce computational effort, an unbiased estimate of $\Lambda(\theta,\theta')$
+Normal. To reduce computational effort, an unbiased estimate of $\Lambda(\theta_t,\theta')$
 based on a minibatch can be used:
@@
 \begin{equation}
-\Lambda^*(\theta,\theta') = \frac{1}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+\Lambda^*(\theta_t,\theta') = \frac{1}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta_t)}\right)  
 \end{equation}
@@
 The acceptance test then becomes a statistical test of the hypothesis
-that $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$ by
-establishing that $\Lambda^*(\theta,\theta')$ is substantially larger
-than $\psi(u,\theta,\theta')$.  In ~\cite{cutting_mh_2014} an
+that $\Lambda(\theta_t,\theta') > \psi(u,\theta_t,\theta')$ by
+establishing that $\Lambda^*(\theta_t,\theta')$ is substantially larger
+than $\psi(u,\theta_t,\theta')$.  In ~\cite{cutting_mh_2014} an
 asymptotic central limit argument was used to derive this gap, while
@@
 ~\cite{cutting_mh_2014} and ~\cite{icml2014c1_bardenet14} is
-$\Omega(\sqrt{N})$, where $N$ is the number of data points.
+$\Omega(N)$, where $N$ is the number of data points.
 
@@
 \begin{equation}
-\Lambda^*(\theta,\theta') =  \frac{1}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta_t)}=
+\Lambda^*(\theta_t,\theta') =  \frac{1}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta_t)}=
   %%\frac{1}{2}\left( (x_i-\theta')^2 - (x_i - \theta_t)^2\right) =
@@
 proposal distribution has the same variance as the target distribution
-$\mathcal{N}(\mu,1/N)$, so we take $q(\theta'\mid\theta_t) =
-\phi((\theta'-\theta_t)N)$, where $\phi(x)$ is a Normal density
+$\mathcal{N}(\theta'-\theta_t,1/N)$, so we take $q(\theta'\mid\theta_t) =
+\phi((\theta'-\theta_t)N)$, where $\phi(x)$ is the Normal density
 function with zero mean and unit variance. This proposal is symmetric
@@
   For the model above, there
-  exists a fixed constant $c$ such that with probability $ \geq c$ over the joint
+  exists a fixed (independent of $N$) constant $c$ such that with probability $ \geq c$ over the joint
   distribution of $(\theta_t, \theta', u)$, the
@@
   $\theta_t -0.5 \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events
-  occur with probability $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and $p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively. 
-  Choose $u_0$
+  occur with probability $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and $p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively,
+  and guarantee that $\Lambda^*(\theta',\theta_t)$ is negative, which guarantees that we can find
+  a $u$ so that $\psi(u,\theta',\theta_t)$ equals the expected value of $\Lambda^*(\theta',\theta_t)$.
+  Specifically, choose $u_0$
   to satisfy
   \begin{equation}
+    \log u_0 = N(\Lambda^*(\theta',\theta_t)-\psi(1,\theta',\theta_t))
+  \end{equation}
+  which evaluates to 
+  \begin{equation}
     \log u_0 = -N(\theta'-\theta_t)\left(\theta_t-0.5+\frac{\theta'-\theta_t}{2}\right) 
@@
   \begin{equation}
-\frac{1}{b}\sum_{i=1}^b x_i-\left(\theta_t+\frac{\theta'-\theta_t}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) > 0
-  \end{equation}
+    \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta_t+\frac{\theta'-\theta_t}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right)
+    \hspace{0.1in} \not\approx \hspace{0.1in} 0
+  \end{equation}
+  where $\not\approx$ means ``significantly different from'' under the distribution over samples of $x_i$. 
   The above choice of $u_0$ ensures that the terms in parenthesis above sum to 0.5. Since the $x_i$ have mean 0.5, the
@@
   \begin{equation}
-    \log u \in \log u_0 + [-1,0.375]
-  \end{equation}
-  the size of this range is at least $\exp([-2.5,-1.125]) = [0082,0.286]$ and occurs
-  with probability at least $p_2=0.2044$. The samples $\theta_t$, $(\theta'-\theta_t)$
-  and $u$ are drawn independently and so the total probability of these events is
+    \log u \in \log u_0 + [-0.5,0.375]
+  \end{equation}
+  the size of the range in $u$ is at least $\exp([-2,-1.125]) = [0082,0.286]$ and occurs
+  with probability at least $p_2=0.2044$. With $u$ in this range, we rewrite the test as:
+  \begin{equation}
+    \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
+  \end{equation}
+  so that the LHS has expected value zero.  Given our choice of intervals for the variables, the RHS is in the range $1/\sqrt{N}[-1,0.75]$.
+  The standard deviation of the LHS given the interval constraints is at least $0.5/\sqrt{b}$. And
+  so the gap between LHS and RHS is at most $2\sqrt{b/N}$ standard deviations.
+
+  The samples $\theta_t$, $(\theta'-\theta_t)$
+  and $u$ are drawn independently and so the probability of the conjunction of these events is
   $c = p_0 p_1 p_2 = 0.009$.
@@
 described in ~\cite{sghmc_2014} which allows momentum-augmented
-
 proposals with larger step sizes. However, step sizes are still
@@
 
-\section{A New Metropolis-Hastings Test}\label{sec:our_algorithm}
-
-For our new MH test, we use two key values, $\Delta$ and $\Delta'$:
-\begin{equation}\label{eq:deltas}
-\Delta = \log \left( \frac{p(\theta')  \prod_{i=1}^N p(x_i \mid \theta')  q(\theta_t \mid
-\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)  q(\theta' \mid
-\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta') (\prod_{i=1}^n p(x_i \mid
-\theta'))^{\frac{N}{n}} q(\theta_t \mid \theta')}{p(\theta_t) (\prod_{i=1}^n p(x_i \mid
-\theta_t))^{\frac{N}{n}} q(\theta' \mid \theta_t)} \right)
-\end{equation}
-where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}. Thus, $\Delta$ and
-$\Delta'$ are probability ratios in log space. Note that the only difference between $\Delta$ and
-$\Delta'$ is that they use $N$ and $n$ data terms $x_i$, respectively, and that the latter uses an
-$N/n$ scaling term. We assume $n \ll N$ (so $\Delta'$ is substantially faster to compute) and that
-the $n$ data points $x_i$ represent a random minibatch without replacement.
-
-\subsection{The Full Data Test}\label{ssec:deltas_full}
-
-To motivate the use of $\Delta$, we turn to Lemma~\ref{lem:detailed_balance}.
+\section{A New Metropolis-Hastings Acceptance Test}\label{sec:our_algorithm}
+
+For our new M-H test, we denote the exact and approximate log likelihood ratios as $\Delta$ and $\Delta^*$.
+First, $\Delta$ is defined as:
+\begin{equation}\label{eq:delta0}
+    \Delta(x,\theta_t,\theta') =
+\log \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
+  \theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid\theta_t)}
+  \end{equation}
+and we can separate out terms dependent and independent of the data $x$ as:
+\begin{equation}\label{eq:delta1}
+\Delta(x,\theta_t,\theta') = N\psi(1,\theta_t,\theta') + \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta_t)}
+\end{equation}
+\begin{equation}\label{eq:delta2}
+    \Delta^*(x,\theta_t,\theta') =
+N\psi(1,\theta_t,\theta') + \frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta_t)}
+\end{equation}
+Note that $\Delta$ and $\Delta^*$ are evaluated on the full dataset and a minibatch of size $b$
+respectively. The scaling term $N$ produces true log likelihood ratios whereas $\Lambda-\psi$
+is a scaled log ratio. $\Delta^*$ is an unbiased estimator of $\Delta$.
+
+\subsection{A Full Data Acceptance Test}\label{ssec:deltas_full}
+
+The key to our test is to smooth the acceptance function.  We consider tests other than the Metropolis
+test that satisfy the conditions needed for accurate posterior estimation. This condition is
+the detailed balance condition defined earlier. Expressed in terms of $\Delta$, the
+classical Metropolis test is $f(\Delta) = {\rm min}(\exp(\Delta),1)$. More generally we have:
 
@@
 balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) = f(\theta')p(\theta_t \mid \theta')$, where
-$p(\theta_y \mid \theta_x)$ is the probability of jumping from $\theta_x$ to $\theta_y$ in our
+$p(\theta_y \mid \theta_x)$ is the probability of jumping from $\theta_x$ to $\theta_y$ in the
 chain.
@@
 \begin{proof}
-We begin by deriving $p(\theta' \mid \theta_t)$. This is equivalent to the probability of proposing
+First note that the transition probability $p(\theta' \mid \theta_t)$ is the probability of proposing
 $\theta'$ and then accepting it, so $p(\theta' \mid \theta_t) = q(\theta' \mid \theta_t)g(\Delta).$
@@
 e^\Delta \} = \min\left\{1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid
-\theta_t)}\right\}$ satisfies the condition $g(\Delta) = \exp(\Delta)g(-\Delta)$.
-
-For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
+  \theta_t)}\right\}$ satisfies the condition $g(\Delta) = \exp(\Delta)g(-\Delta)$. In fact it is
+the optimal acceptance function in terms of acceptance rate, since it accepts with probability
+1 for $\Delta > 0$.
+
+\subsubsection{Logistic Acceptance Function}\label{ssec:deltas_full}
+For our new MH test we use the logistic function: $g(\Delta) =
 (1+\exp(-\Delta))^{-1}$. Straightforward arithmetic shows that it satisfies the condition in
@@
 
-\subsection{The Minibatch Version}\label{ssec:deltas_minibatch}
+\subsection{A Minibatch Acceptance Test}\label{ssec:deltas_minibatch}
 

### 2016-07-05 nips_2016.tex
updates to first section

@@
 \usepackage{amsfonts}       % blackboard math symbols
+\usepackage{amssymb}       % blackboard math symbols
 \usepackage{nicefrac}       % compact symbols for 1/2, etc.
@@
 $\theta_{t+1}=\theta_t$.  The test is often implemented with an
-auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison $u < \alpha(\theta_t,\theta')$
-
-The acceptance test guarantees detailed balance, so that
-\begin{equation}
+auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison $u < \alpha(\theta_t,\theta')$.
+For notational simplicity, from now on we will drop the subscript $t$ for the current sample $\theta_t$ and
+denote is simply as $\theta$.
+
+The acceptance test guarantees detailed balance, which means
+\begin{equation}\label{detailed_balance1}
+  p(\theta)p(\theta,\theta') = p(\theta')p(\theta',\theta)
+\end{equation}
+where $p(\theta,\theta')$ is the probability of a transition from state $\theta$ to $\theta'$. Here
+$p(\theta,\theta') = q(\theta'\mid\theta)\alpha(\theta,\theta')$ so the detailed balance equation becomes
+\begin{equation}\label{detailed_balance2}
   p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
@@
 \begin{equation}
-\alpha(\theta_t,\theta') = 
-\min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
-\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid\theta_t)} \right\},
-\end{equation}
-where $p(\theta)$ is a prior, and $p(x_i \mid \theta)$ are the
+\alpha(\theta,\theta') = 
+\min\left\{ 1, \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
+\theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)} \right\},
+\end{equation}
+where $p_0(\theta)$ is a prior, and $p(x_i \mid \theta)$ are the
 probabilities of the observations. Computing samples this way requires
-the use of all $N$ training data points. Moreover, it is difficult to
-design tests using substantially fewer than $N$ points that also
-satisfy detailed balance. To reconcile these competing
-objectives,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
+the use of all $N$ training data points, but this is very expensive
+for large datasets. To address this challenge,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
 approximate Metropolis-Hasting tests using sequential hypothesis
@@
 data and test the hypothesis that the sample $\theta'$ based on an
-approximate version of the test $u < \alpha(\theta_t,\theta')$. If the
+approximate version of the test $u < \alpha(\theta,\theta')$. If the
 approximate test cannot make a decision with sufficient confidence,
@@
 
-Following ~\cite{icml2014c1_bardenet14}, we write the test $u < \alpha(\theta_t,\theta')$ in
-the equivalent form $\Lambda(\theta_t,\theta') > \psi(u,\theta_t,\theta')$ where
-\begin{equation}
-\Lambda(\theta_t,\theta') = \frac{1}{N}\sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta_t)}\right)  
+Following ~\cite{icml2014c1_bardenet14}, we write the test $u < \alpha(\theta,\theta')$ in
+the equivalent form $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$ where
+\begin{equation}
+\Lambda(\theta,\theta') = \frac{1}{N}\sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
 \hspace{0.2in} {\rm and} \hspace{0.2in}
-\psi(u,\theta_t,\theta') = \frac{1}{N}\log\left(u\frac{q(\theta'|\theta_t)p(\theta_t)}{q(\theta_t|\theta')p(\theta')}\right)
-\end{equation}
-
-By the central limit theorem, we expect $\Lambda(\theta_t,\theta')$ to be approximately
-Normal. To reduce computational effort, an unbiased estimate of $\Lambda(\theta_t,\theta')$
+\psi(u,\theta,\theta') = \frac{1}{N}\log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right)
+\end{equation}
+
+To reduce computational effort, an unbiased estimate of $\Lambda(\theta,\theta')$
 based on a minibatch can be used:
@@
 \begin{equation}
-\Lambda^*(\theta_t,\theta') = \frac{1}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta_t)}\right)  
-\end{equation}
-
+\Lambda^*(\theta,\theta') = \frac{1}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+\end{equation}
+
+The minibatch should contain randomly selected samples, and
+by the central limit theorem, we expect $\Lambda^*(\theta,\theta')$ to be approximately
+Normal. 
 The acceptance test then becomes a statistical test of the hypothesis
-that $\Lambda(\theta_t,\theta') > \psi(u,\theta_t,\theta')$ by
-establishing that $\Lambda^*(\theta_t,\theta')$ is substantially larger
-than $\psi(u,\theta_t,\theta')$.  In ~\cite{cutting_mh_2014} an
+that $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$ by
+establishing that $\Lambda^*(\theta,\theta')$ is substantially larger
+than $\psi(u,\theta,\theta')$.  In ~\cite{cutting_mh_2014} an
 asymptotic central limit argument was used to derive this gap, while
@@
 We next show that for some simple distributions, the lower
-bound of average number of data instance consumed for one iteration of
+bound of average number of data instances consumed for one iteration of
 ~\cite{cutting_mh_2014} and ~\cite{icml2014c1_bardenet14} is
@@
 \begin{equation}
-\Lambda^*(\theta_t,\theta') =  \frac{1}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta_t)}=
-  %%\frac{1}{2}\left( (x_i-\theta')^2 - (x_i - \theta_t)^2\right) =
-  (\theta'-\theta_t)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta_t-\frac{\theta'-\theta_t}{2}\right)
+\Lambda^*(\theta,\theta') =  \frac{1}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)}=
+  %%\frac{1}{2}\left( (x_i-\theta')^2 - (x_i - \theta)^2\right) =
+  (\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
 \end{equation}
@@
 Since the $x_i$ have unit variance, their mean has variance $1/b$, and the variance $\sigma^2(\Lambda^*)$
-of $\Lambda^*(\theta,\theta')$ is $(\theta'-\theta_t)^2/b$. In order to pass
+of $\Lambda^*(\theta,\theta')$ is $(\theta'-\theta)^2/b$. In order to pass
 a hypothesis test that $\Lambda > \psi$, we will need to establish a
-large enough gap (several $\sigma(\Lambda^*)$) between $\Lambda^*(\theta',\theta_t)$
-and $\psi(u,\theta',\theta_t)$.
+large enough gap (several $\sigma(\Lambda^*)$) between $\Lambda^*(\theta',\theta)$
+and $\psi(u,\theta',\theta)$.
 
@@
 proposal distribution has the same variance as the target distribution
-$\mathcal{N}(\theta'-\theta_t,1/N)$, so we take $q(\theta'\mid\theta_t) =
-\phi((\theta'-\theta_t)N)$, where $\phi(x)$ is the Normal density
+$\mathcal{N}(\theta'-\theta,1/N)$, so we take $q(\theta'\mid\theta) =
+\phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal density
 function with zero mean and unit variance. This proposal is symmetric
-$q(\theta',\theta_t)=q(\theta_t,\theta')$, and we assumed uniform
-prior for $\theta$, so $\psi(u,\theta',\theta_t)$ reduces to $1/N\log
+$q(\theta',\theta)=q(\theta,\theta')$, and we assumed uniform
+prior for $\theta$, so $\psi(u,\theta',\theta)$ reduces to $1/N\log
 u$.
@@
   exists a fixed (independent of $N$) constant $c$ such that with probability $ \geq c$ over the joint
-  distribution of $(\theta_t, \theta', u)$, the
+  distribution of $(\theta, \theta', u)$, the
   test from \cite{icml2014c1_bardenet14} consumes all $N$ samples. 
@@
 
-\begin{proof} Assume $(\theta' - \theta_t) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and
-  $\theta_t -0.5 \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events
+\begin{proof} Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and
+  $\theta -0.5 \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events
   occur with probability $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and $p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively,
-  and guarantee that $\Lambda^*(\theta',\theta_t)$ is negative, which guarantees that we can find
-  a $u$ so that $\psi(u,\theta',\theta_t)$ equals the expected value of $\Lambda^*(\theta',\theta_t)$.
+  and guarantee that $\Lambda^*(\theta',\theta)$ is negative, which guarantees that we can find
+  a $u$ so that $\psi(u,\theta',\theta)$ equals the expected value of $\Lambda^*(\theta',\theta)$.
   Specifically, choose $u_0$
@@
   \begin{equation}
-    \log u_0 = N(\Lambda^*(\theta',\theta_t)-\psi(1,\theta',\theta_t))
+    \log u_0 = N(\Lambda^*(\theta',\theta)-\psi(1,\theta',\theta))
   \end{equation}
@@
   \begin{equation}
-    \log u_0 = -N(\theta'-\theta_t)\left(\theta_t-0.5+\frac{\theta'-\theta_t}{2}\right) 
+    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right) 
   \end{equation}
@@
   \begin{equation}
-    \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta_t+\frac{\theta'-\theta_t}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right)
+    \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right)
     \hspace{0.1in} \not\approx \hspace{0.1in} 0
@@
 
- The choices above for $\theta_t$ and $\theta'$ guarantee that
+ The choices above for $\theta$ and $\theta'$ guarantee that
   \begin{equation}
@@
 
-  The samples $\theta_t$, $(\theta'-\theta_t)$
+  The samples $\theta$, $(\theta'-\theta)$
   and $u$ are drawn independently and so the probability of the conjunction of these events is
@@
 \begin{equation}\label{eq:delta0}
-    \Delta(x,\theta_t,\theta') =
-\log \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
-  \theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid\theta_t)}
+  \Delta(\theta,\theta') = \log\frac{p(\theta')q(\theta\mid\theta')}{p(\theta)q(\theta'\mid\theta)}
+\end{equation}
+and in the Bayesian case
+\begin{equation}\label{eq:delta1}
+\Delta(x,\theta,\theta')  =
+\log \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
+  \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)}
   \end{equation}
 and we can separate out terms dependent and independent of the data $x$ as:
-\begin{equation}\label{eq:delta1}
-\Delta(x,\theta_t,\theta') = N\psi(1,\theta_t,\theta') + \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta_t)}
-\end{equation}
 \begin{equation}\label{eq:delta2}
-    \Delta^*(x,\theta_t,\theta') =
-N\psi(1,\theta_t,\theta') + \frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta_t)}
+\Delta(x,\theta,\theta') = N\psi(1,\theta,\theta') + \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)}
+\end{equation}
+A minibatch estimator of $\Delta$ can be defined as:
+\begin{equation}\label{eq:delta3}
+    \Delta^*(x,\theta,\theta') =
+N\psi(1,\theta,\theta') + \frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)}
 \end{equation}
 Note that $\Delta$ and $\Delta^*$ are evaluated on the full dataset and a minibatch of size $b$
-respectively. The scaling term $N$ produces true log likelihood ratios whereas $\Lambda-\psi$
-is a scaled log ratio. $\Delta^*$ is an unbiased estimator of $\Delta$.
-
-\subsection{A Full Data Acceptance Test}\label{ssec:deltas_full}
-
-The key to our test is to smooth the acceptance function.  We consider tests other than the Metropolis
+respectively. The scaling term $\frac{N}{b}$ ensures that $\Delta^*(x,\theta,\theta')$ is an
+unbiased estimator of $\Delta(x,\theta,\theta')$.
+
+\subsection{A Full Dataset Acceptance Test}\label{ssec:deltas_full}
+
+The key to our test is a smooth acceptance function.  We consider tests other than the Metropolis
 test that satisfy the conditions needed for accurate posterior estimation. This condition is
@@
 \begin{lemma}\label{lem:detailed_balance}
-An acceptance function $g$ such that $g(\Delta) = \exp(\Delta) g(-\Delta )$ satisfies detailed
-balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) = f(\theta')p(\theta_t \mid \theta')$, where
-$p(\theta_y \mid \theta_x)$ is the probability of jumping from $\theta_x$ to $\theta_y$ in the
-chain.
+  If $g(s)$ is any function such that $g(s) = \exp(s) g(-s)$, then the acceptance function $\alpha(\theta,\theta') \triangleq  g(\Delta(\theta,\theta'))$ satisfies detailed balance. That is,
+  \begin{equation}\label{detbal}
+    p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
+    \end{equation}
+
 \end{lemma}
@@
 \begin{proof}
-First note that the transition probability $p(\theta' \mid \theta_t)$ is the probability of proposing
-$\theta'$ and then accepting it, so $p(\theta' \mid \theta_t) = q(\theta' \mid \theta_t)g(\Delta).$
-Similarly, $p(\theta_t \mid \theta') = q(\theta_t \mid \theta')g(-\Delta).$ Notice that the
-probability of accepting a transition from $\theta'$ to $\theta_t$ is $g(-\Delta)$ because this
-inverts the fraction inside the logarithm term of $\Delta$.  By assumption, we can expand $g(\Delta)
-= \exp(\Delta)g(-\Delta)$ in $p(\theta' \mid \theta_t)$. Doing this, and combining the result with
-the definition of $p(\theta_t \mid \theta')$, we get
-\begin{equation}\label{eq:combined}
-g(-\Delta) = \frac{p(\theta' \mid \theta_t)}{q(\theta' \mid \theta_t)\exp(\Delta)} = \frac{p(\theta_t \mid \theta')}{q(\theta_t \mid \theta')}.
-\end{equation}
-Rearranging terms and expanding $\exp(\Delta)$, we have
-\begin{equation}\label{eq:rearrange}
-\frac{p(\theta' \mid \theta_t) f(\theta_t) q(\theta' \mid \theta_t)}{q(\theta' \mid \theta_t) f(\theta') q(\theta_t \mid \theta')} = \frac{p(\theta_t \mid \theta')}{ q(\theta_t \mid \theta')}.
-\end{equation}
-Cancellations result in $f(\theta') p(\theta_t \mid \theta') = f(\theta_t) p(\theta' \mid
-\theta_t)$. Thus, detailed balance is satisfied.
+  Start with the LHS of (\ref{detbal}), substitute for $\alpha(\theta,\theta')$, and then apply the
+  identity $g(s) = \exp(s)g(-s)$:
+  \begin{equation}\label{detbal1}
+    \begin{split}
+      p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') =~ & p(\theta)q(\theta'\mid\theta)g(\Delta(\theta,\theta'))\\
+      =~ & p(\theta)q(\theta'\mid\theta)\exp(\Delta(\theta,\theta'))g(-\Delta(\theta,\theta'))
+    \end{split}
+  \end{equation}
+  Note then that $-\Delta(\theta,\theta') = \Delta(\theta',\theta)$ because $\Delta$ is the log of ratios of
+  terms involving $\theta$ and $\theta'$ and exchanging the two inverts all these ratios, negating the
+  logarithm. Continuing:
+  \begin{equation}\label{detbal2}
+    \begin{split}
+    p(\theta)q(\theta'\mid\theta)\exp(\Delta(\theta,\theta'))g(-\Delta(\theta,\theta')) = ~ & p(\theta)q(\theta'\mid\theta)\exp(\Delta(\theta,\theta'))g(\Delta(\theta',\theta))\\
+    = ~ & p(\theta)q(\theta'\mid\theta)\exp(\Delta(\theta,\theta'))\alpha(\theta',\theta)
+    \end{split}
+  \end{equation}
+  Substituting now for $\Delta(\theta,\theta')$ from equation (\ref{eq:delta0}) gives
+  \begin{equation*}\label{detbal3}
+    p(\theta)q(\theta'\mid\theta)\frac{p(\theta')q(\theta\mid\theta')}{p(\theta)q(\theta'\mid\theta)}\alpha(\theta',\theta)
+      = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta) = RHS
+  \end{equation*}
 \end{proof}
 
-As a sanity check, the standard Metropolis-Hastings acceptance function $g(\Delta) = \min\{1,
-e^\Delta \} = \min\left\{1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid
-  \theta_t)}\right\}$ satisfies the condition $g(\Delta) = \exp(\Delta)g(-\Delta)$. In fact it is
-the optimal acceptance function in terms of acceptance rate, since it accepts with probability
-1 for $\Delta > 0$.
+As a sanity check, choosing $g(s) = \min\{1, \exp(s) \}$ produces the
+standard Metropolis acceptance test $\alpha(\theta,\theta') = g(\Delta(\theta,\theta')) = 
+\min\left\{1, \frac{p(\theta')q(\theta \mid
+  \theta')}{p(\theta)q(\theta' \mid \theta)}\right\}$. $g(s)$ satisfies
+the condition $g(s) = \exp(s)g(-s)$, and therefore detailed
+balance. In fact it is the optimal acceptance function in terms of
+acceptance rate, since it accepts with probability 1 for $\Delta > 0$.
 
 \subsubsection{Logistic Acceptance Function}\label{ssec:deltas_full}
-For our new MH test we use the logistic function: $g(\Delta) =
-(1+\exp(-\Delta))^{-1}$. Straightforward arithmetic shows that it satisfies the condition in
-Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily sample from
-it using the following procedure. At any iteration with a current parameter $\theta_t$ and a
-candidate sample $\theta'$, we can compute $\Delta$. Let $u$ be a uniform random variable $u \sim
-{\rm Unif}[0,1]$. We accept $\theta'$ if $g(\Delta) > u$, and reject otherwise. This process is
-equivalent to sampling a random variable $X$ with cumulative distribution function $F_X(x) = g(x)$,
-and accepting if $\Delta > X$ and rejecting otherwise. We can define $X = g^{-1}(u)$ so that its CDF
-is the logistic function, which follows because for arbitrary $X = x$, we have
-\[
-F_X(x) = {\rm Pr}(X \le x) = {\rm Pr}(g^{-1}(u) \le x) = {\rm Pr}(u \le g(x)) = \int_{0}^{g(x)} 1 dx = g(x),
-\]
-as the density of $u$ is one. Thus, the criteria to accept the candidate $\theta'$ is equivalent to
-whether $\Delta > X$. Moreover, the density of $X$ (a standard logistic random variable) is
-symmetric about zero, so the acceptance criteria can also be expressed as $\Delta + X>0$.
+For our new MH test we use the logistic function: $g(s) =
+(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the condition in
+Lemma~\ref{lem:detailed_balance}.  While it is slightly less efficient than the Metropolis
+test when used on the full dataset, we will see soon that its smoothness allows it to naturally
+tolerate substantial variance in its input argument. This in turn will lead to a much more efficient
+test on subsets of data.
+
+Assume we begin with the current sample $\theta$  and a
+candidate sample $\theta'$. Let $u$ be a uniform random variable $u \sim
+\mathcal{U}[0,1]$. We accept $\theta'$ if $g(\Delta(\theta,\theta')) > u$, and reject otherwise.
+Since $g(s)$ is monotonically increasing, its inverse $g^{-1}(s)$ is well-defined and
+unique. So an equivalent test is to accept $\theta'$ iff
+\begin{equation}
+  \Delta(\theta,\theta') > v = g^{-1}(u)
+\end{equation}
+where $v$ is a random variable with the logistic distribution (its CDF is the logistic function). To
+see this notice that $\frac{du}{dv} = g'$, that $g'$ is the density corresponding to a logistic
+CDF, and finally that $\frac{du}{dv}$ is the density of $v$. The density of $v$ is symmetric,
+so we can equivalently test whether
+\begin{equation}
+  \Delta(\theta,\theta') + v > 0
+\end{equation}
+for a logistic random $v$.
 
@@
 
-With large datasets, $\Delta$ is intractable to compute. An intuitive fix is to replace $\Delta$
-with the minibatch version, $\Delta'$, for our MH test. The following lemma characterizes the
-distribution of $\Delta'$.
+We now move to acceptance testing using the minibatch estimator
+$\Delta^*(x,\theta,\theta')$. From equation (\ref{eq:delta3}),
+$\Delta^*(x,\theta,\theta')$ can be represented as a constant term
+plus a sum of $b$ iid terms of the form
+$\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $N$ increases, $\Delta^*(\theta,\theta')$
+therefore has a distribution which approaches a normal distribution by the central limit
+theorem. Later we will provide specific bounds between the CDF of $\Delta^*(\theta,\theta')$
+and a Normal CDF, but for now we use an aymptotic argument.
 
@@
 \[
-\Delta' = \log p(\theta') - \log p(\theta_t) + \log q(\theta_t \mid \theta') - \log q(\theta' \mid \theta_t) +
-\frac{N}{n}\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t)).
+\Delta' = \log p(\theta') - \log p(\theta) + \log q(\theta \mid \theta') - \log q(\theta' \mid \theta) +
+\frac{N}{n}\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta)).
 \]
-During a given iteration, $\theta_t$ and $\theta'$ are fixed. Therefore, the randomness in $\Delta'$
+During a given iteration, $\theta$ and $\theta'$ are fixed. Therefore, the randomness in $\Delta'$
 comes only from the minibatch. Since the minibatch is chosen randomly, the distribution of
-$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t))$ converges to a Gaussian for
+$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta))$ converges to a Gaussian for
 sufficiently large $N$ by the Central Limit Theorem.
@@
 norm}$, plays the part of $\varepsilon$. In light of our revised MH test, the previous discussion
-raises the question: in an iteration of MCMC with current $\theta_t$ and proposed $\theta'$, how do
+raises the question: in an iteration of MCMC with current $\theta$ and proposed $\theta'$, how do
 we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample the
@@
 We use one minibatch data to complete this task.
-From the proof of Lemma~\ref{lem:gaussian}, we know the randomness comes from $\frac{N}{n}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta_t) \right)$.
-Thus, ${\rm var}(\Delta') = \frac{N^2}{n^2}\sum_{i=1}^n{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right ) = \frac{N^2}{n}{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right )$.
-We use data instances in one minibatch to estimate the ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right )$.
-Since there are $n$ data instances in one minibatch, the estimation of ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right )$ has high accuracy.
-In our experiments (see Section~\ref{sec:experiments}) we also use the complete data to obtain the population variance of $\left (  \log p(x_i|\theta') - \log p(x_i|\theta_t) \right )$, and use the nonparametric Bootstrap to estimate the ${\rm std(std}(\Delta'))$, so that we can construct the confidence interval of ${\rm std}(\Delta')$ to reveal the accuracy of our estimation.
+From the proof of Lemma~\ref{lem:gaussian}, we know the randomness comes from $\frac{N}{n}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta) \right)$.
+Thus, ${\rm var}(\Delta') = \frac{N^2}{n^2}\sum_{i=1}^n{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right ) = \frac{N^2}{n}{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$.
+We use data instances in one minibatch to estimate the ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$.
+Since there are $n$ data instances in one minibatch, the estimation of ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$ has high accuracy.
+In our experiments (see Section~\ref{sec:experiments}) we also use the complete data to obtain the population variance of $\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$, and use the nonparametric Bootstrap to estimate the ${\rm std(std}(\Delta'))$, so that we can construct the confidence interval of ${\rm std}(\Delta')$ to reveal the accuracy of our estimation.
 
@@
 corr}$ distribution, and initial sample $\theta_0$.}
-\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ from $p(\theta)$\;
+\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta\}$ from $p(\theta)$\;
 MAP/ML estimation.}
@@
     Propose a candidate $\theta'$\ from proposal distribution\;
-    Draw a minibatch of $m$ samples, compute $\Delta'(\theta',\theta_t)$ and ${\rm std}(\Delta')$ for this sample\;
+    Draw a minibatch of $m$ samples, compute $\Delta'(\theta',\theta)$ and ${\rm std}(\Delta')$ for this sample\;
     \While {${\rm std}(\Delta') \ge 1$} {
@@
         }{
-            Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
+            Reject and re-use the old sample, $\theta_{t+1} = \theta$\;
         }
@@
 \begin{lemma}\label{lem:theory2}
-For one step sampling from $\theta_t$ to $\theta'$, if the jumping step $\|\theta_t - \theta'\|_2 <
+For one step sampling from $\theta$ to $\theta'$, if the jumping step $\|\theta - \theta'\|_2 <
 \epsilon$, and the gradient of the log likelihood is bounded by a constant factor, i.e. $\|\nabla
@@
 \begin{theorem}\label{thm:efficiency}
-We require $\frac{ C^2 \rho^2 }{\sigma^6 \varepsilon^2}$ minibatch size to ensure $|P_a -P_a'| < \varepsilon$, where $C$ is a positive constant which is smaller than 0.4748, $\sigma^2 $ and $\rho$ is the second and third moment of $(\nabla\log \Pr(x_i|\theta_t) - \nabla\log\Pr(x_i|\theta')) - E\left[ \nabla\log \Pr(x_i|\theta_t) - \nabla\log\Pr(x_i|\theta')\right]$, $P_a$ is the accept rate from the complete data, $P_a'$ is the accept rate from minibatch data by our algorithm, $\theta_t$ is the current parameter, and $\theta'$ is the proposed parameter.
+We require $\frac{ C^2 \rho^2 }{\sigma^6 \varepsilon^2}$ minibatch size to ensure $|P_a -P_a'| < \varepsilon$, where $C$ is a positive constant which is smaller than 0.4748, $\sigma^2 $ and $\rho$ is the second and third moment of $(\nabla\log \Pr(x_i|\theta) - \nabla\log\Pr(x_i|\theta')) - E\left[ \nabla\log \Pr(x_i|\theta) - \nabla\log\Pr(x_i|\theta')\right]$, $P_a$ is the accept rate from the complete data, $P_a'$ is the accept rate from minibatch data by our algorithm, $\theta$ is the current parameter, and $\theta'$ is the proposed parameter.
 
@@
 If the minibatch size is $n$, we first compute the probability that more than $n$ data instances are required for adaptive sampling method for one iteration. We denote this probability $P_{rej}$.
-\[P_{rej} = \Pr \left ( 1-\phi_{n-1} \left( \frac{|\bar{l} - \frac{1}{N} \log (u\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')})|}{s} \right )  \geq \varepsilon \right ) \]
+\[P_{rej} = \Pr \left ( 1-\phi_{n-1} \left( \frac{|\bar{l} - \frac{1}{N} \log (u\frac{p(\theta)q(\theta'|\theta)}{p(\theta')q(\theta|\theta')})|}{s} \right )  \geq \varepsilon \right ) \]
 
 where $s = \frac{s_l}{\sqrt{n}}\sqrt{1-\frac{n-1}{N-1}}$. \\
-We define random variable $Z = \frac{1}{N}\log (uQ)$, where $Q=\frac{p(\theta_t)q(\theta'|\theta_t)}{p(\theta')q(\theta_t|\theta')}$.
+We define random variable $Z = \frac{1}{N}\log (uQ)$, where $Q=\frac{p(\theta)q(\theta'|\theta)}{p(\theta')q(\theta|\theta')}$.
 Then $Z$ has probability density function $f_Z(z) = N \exp(zN), z \in (-\infty, \frac{1}{N} \log Q]$.
@@
 arbitrary distribution $D$ once is bounded if we use the same proposer, i.e.  $\|P_i\circ D -
-P'_i\circ D\| \leq 4\zeta \ell \;\; \forall D$. We write $P_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
-q(\theta'\mid \theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$
+P'_i\circ D\| \leq 4\zeta \ell \;\; \forall D$. We write $P_i(\theta \mid \theta') = P_a(\theta, \theta')
+q(\theta'\mid \theta) + (1-P_a(\theta,\theta')) \delta_D(\theta' - \theta)$, where $\delta_D$
 is the Dirac delta function. The total variation distance between two generated distributions is
 \begin{align*}
-\|P_i\circ D - P'_i\circ D\| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) dD(\theta_t) \right| d\theta' \\
-&= \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) dD(\theta_t) \right|  d\theta'\\ 
-&\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + \delta_D(\theta_t - \theta')) dD(\theta_t) \right| d\theta'\\
+\|P_i\circ D - P'_i\circ D\| &= \int_{\theta'} \left| \int_{\theta}((P_a-P_a') q(\theta'|\theta) + (1-P_a - 1+P_a') \delta_D(\theta -\theta')) dD(\theta) \right| d\theta' \\
+&= \int_{\theta'} \left| \int_{\theta} (P_a- P_a')(q(\theta'|\theta) - \delta_D(\theta - \theta')) dD(\theta) \right|  d\theta'\\ 
+&\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta}(q(\theta'|\theta) + \delta_D(\theta - \theta')) dD(\theta) \right| d\theta'\\
 &\leq 2 \zeta \ell \int_{\theta'} ( D'(\theta') + D(\theta') ) d\theta'\\
@@
 From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
-\theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
-$g_i(\theta) = \log p(x_i\mid \theta)$, and $q_i = g_i(\theta') - g_i(\theta_t)$. Then by Taylor's Theorem:
+\theta))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
+$g_i(\theta) = \log p(x_i\mid \theta)$, and $q_i = g_i(\theta') - g_i(\theta)$. Then by Taylor's Theorem:
 \[
-|q| = |g(\theta') - g(\theta_t)| = |\nabla g(\theta_t)^T(\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
+|q| = |g(\theta') - g(\theta)| = |\nabla g(\theta)^T(\theta' - \theta) + o(\theta' - \theta)^2| \leq 2\epsilon k.
 \]
@@
 \begin{align*}
-{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta_t)\right] = {\rm Var}\left(\sum_{i=1}^m q_i\right) \\
+{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta)\right] = {\rm Var}\left(\sum_{i=1}^m q_i\right) \\
 &= m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
@@
 \subsection{Proof of Theorem~\ref{thm:efficiency}}\label{app:theory4}
-Since only the $\frac{N}{nT}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta_t) \right)$ term brings randomness into $\Delta'$.
+Since only the $\frac{N}{nT}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta) \right)$ term brings randomness into $\Delta'$.
 Define $X_i = \frac{N}{T}\left(\log p(x_i|\theta') - \log p (x_i|theta_t) \right)$, $\tilde{X_i} = X_i - E(X_i)$, and $Y_n = \frac{\sum_{i=1}^n \tilde{X_i}}{n}$ (i.e.

### 2016-07-06 nips_2016.tex
updates to first section

@@
 Assume we begin with the current sample $\theta$  and a
-candidate sample $\theta'$. Let $u$ be a uniform random variable $u \sim
-\mathcal{U}[0,1]$. We accept $\theta'$ if $g(\Delta(\theta,\theta')) > u$, and reject otherwise.
+candidate sample $\theta'$. Let $V$ be a uniform random variable $V \sim
+\mathcal{U}(0,1)$. We accept $\theta'$ if $g(\Delta(\theta,\theta')) > V$, and reject otherwise.
 Since $g(s)$ is monotonically increasing, its inverse $g^{-1}(s)$ is well-defined and
@@
 \begin{equation}
-  \Delta(\theta,\theta') > v = g^{-1}(u)
-\end{equation}
-where $v$ is a random variable with the logistic distribution (its CDF is the logistic function). To
-see this notice that $\frac{du}{dv} = g'$, that $g'$ is the density corresponding to a logistic
-CDF, and finally that $\frac{du}{dv}$ is the density of $v$. The density of $v$ is symmetric,
+  \Delta(\theta,\theta') > X = g^{-1}(V)
+\end{equation}
+where $X$ is a random variable with the logistic distribution (its CDF is the logistic function). To
+see this notice that $\frac{dV}{dX} = g'$, that $g'$ is the density corresponding to a logistic
+CDF, and finally that $\frac{dV}{dX}$ is the density of $X$. The density of $X$ is symmetric,
 so we can equivalently test whether
 \begin{equation}
-  \Delta(\theta,\theta') + v > 0
-\end{equation}
-for a logistic random $v$.
+  \label{theexacttest}
+  \Delta(\theta,\theta') + X > 0
+\end{equation}
+for a logistic random $X$.
 
@@
 theorem. Later we will provide specific bounds between the CDF of $\Delta^*(\theta,\theta')$
-and a Normal CDF, but for now we use an aymptotic argument.
-
-\begin{lemma}\label{lem:gaussian}
-If the minibatch data is chosen randomly without replacement and has sufficiently many elements,
-then the distribution of $\Delta'$ is approximately Gaussian.
-\end{lemma}
-
-\begin{proof}
-By expanding the definition of $\Delta'$ from Equation~\ref{eq:deltas}, we get
-\[
-\Delta' = \log p(\theta') - \log p(\theta) + \log q(\theta \mid \theta') - \log q(\theta' \mid \theta) +
-\frac{N}{n}\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta)).
-\]
-During a given iteration, $\theta$ and $\theta'$ are fixed. Therefore, the randomness in $\Delta'$
-comes only from the minibatch. Since the minibatch is chosen randomly, the distribution of
-$\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta))$ converges to a Gaussian for
-sufficiently large $N$ by the Central Limit Theorem.
-\end{proof}
-
-% Daniel: I decided to put the figure here, so it's closer to the text that references it.
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
-    \caption{
-    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
-    standard logistic distribution. We use three standard deviation values of $X_{\rm norm}$. The
-    two red (respectively, blue, green) curves form the logistic. The $y$-axis is capped at 0.5.
-    This figure must be viewed in color.
-    }
-    \label{fig:deconvolution}
-    %\vspace{-10pt}
-\end{figure}
-
-From Lemma~\ref{lem:gaussian}, since $\Delta'$ is a noisy approximation of $\Delta$, the
-relationship is expressed as
+and a Normal CDF, but for now we use an asymptotic argument.
+
+In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$, we
+can write
 \begin{equation}\label{eq:relationship}
-\Delta' = \Delta + \varepsilon, \quad \varepsilon \sim \mathcal{N}(0, \sigma^2),
-\end{equation}
-for a Gaussian noise term $\varepsilon$.  The noise means we cannot just replace $\Delta+X$ with
-$\Delta'+X$ for our test, so we need to slightly change our acceptance criteria. Our insight is to
-decompose $X$ as
+\Delta^* = \Delta + X_{\rm norm}, \quad X_{\rm norm} \sim \mathcal{\bar{N}}(0, \sigma^2(\Delta^*)),
+\end{equation}
+where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is approximately
+normal with variance $\sigma^2(\Delta^*)$. 
+But to perform the test in equation (\ref{theexacttest}) we want $\Delta + X$ for
+a logistic random $X$ (call it $X_{\rm logistic}$ from now on). 
+Our approach is to decompose $X_{\rm logistic}$ as
 \begin{equation}\label{eq:deconvolution}
-X = X_{\rm norm}+X_{\rm corr},
+X_{\rm logistic} = X_{\rm norm}+X_{\rm corr},
 \end{equation}
 where we assume $X_{\rm norm}$ is a zero-mean Gaussian and $X_{\rm corr}$ is a zero-mean
-``correction'' term.  These two add (i.e., their distributions convolve) to form $X$. The criteria
-to accept is now
+``correction'' term.  These two add (i.e., their distributions convolve) to form $X_{\rm logistic}$. The acceptance
+test is now:
 \begin{equation}\label{eq:criteria}
-\Delta + X = \Delta + X_{\rm norm} + X_{\rm corr} \approx \Delta' + X_{\rm corr} >0.
-\end{equation}
-The approximation exists because $X_{\rm norm}$ is an estimate of $\varepsilon$. If $X_{\rm norm}$
-is exact, then we get the same MH test with $\Delta'$ as we do with the full data version $\Delta$.
+\Delta + X_{\rm logistic} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
+\end{equation}
+If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0, \sigma^2(\Delta^*))$,
+the above test is exact as well. And as we will see later,  if there is a maximum error $\epsilon$ between
+the CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0, \sigma^2(\Delta^*))$,
+then the acceptance test has an error of at most $\epsilon$.
 

### 2016-07-06 nips_2016.tex
updates to first section

@@
 \end{equation}
+i.e. given an approximate estimate $\Delta^*$ from the current data minibatch (assuming its
+variance is small enough) we test acceptance by adding a random variable $X_{\rm corr}$ and
+then accept $\theta'$ if the result is positive.
+
 If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0, \sigma^2(\Delta^*))$,
@@
 then the acceptance test has an error of at most $\epsilon$.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
+    \caption{
+    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
+    standard logistic distribution. We use three standard deviation values of $X_{\rm norm}$. The
+    two red curves convolve to form the logistic, etc. The $y$-axis is capped at 0.5 for
+    readability. This figure must be viewed in color.
+    }
+    \label{fig:deconvolution}
+    %\vspace{-10pt}
+\end{figure}
 

### 2016-07-06 nips_2016.tex
updates to first section

@@
 
-\title{An Efficient Minibatch Acceptance Test for MCMC}
+\title{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
 
@@
 machine learning. We are particularly interested in their application
-to modeling very large datasets, where it is impractical to
-perform Metropolis-Hastings tests on the full data. Previous work on reducing
+to modeling very large datasets, where it is impractical to perform
+Metropolis-Hastings tests on the full data. Previous work on reducing
 the cost of Metropolis-Hastings tests yield variable data reductions
-and in the worst case require the entire dataset. 
-Here we present a simple test which requires constant size minibatches
-if an auxiliary condition is met. We describe several cases under which
-this condition can be satisfied. Our new test is naturally tolerant
-of ``noisy'' estimates of the transition probability ratio. To achieve
-this, we replace the classical Metropolis-Hastings test with a smooth
-acceptance function which is a sigmoid (a logistic function). 
+and in the worst case require the entire dataset.  Here we present a
+simple test which requires constant size minibatches if an auxiliary
+condition is met. We describe several cases under which this condition
+can be satisfied. Our new test is naturally tolerant of ``noisy''
+estimates of the transition probability ratio. To achieve this, we
+replace the classical Metropolis-Hastings acceptance test with a
+smooth acceptance function which is a sigmoid (logistic)
+function of the log likelihood ratio.
 The resulting test can be combined with minibatch proposals to yield
-updates with the same complexity as a simple SGD update. In this paper we
-derive the test, discuss its implementation, and present several
+updates with the same complexity as a simple SGD update. In this paper
+we derive the test, discuss its implementation, and present several
 experiments.
@@
 
+\subsection{An Example}
 Consider a simple model where samples $x_i$ for $i=1,\ldots,N$ are
@@
   distribution of $(\theta, \theta', u)$, the
-  test from \cite{icml2014c1_bardenet14} consumes all $N$ samples. 
+  tests from \cite{icml2014c1_bardenet14} and \cite{cutting_mh_2014} consume all $N$ samples. 
 \end{lemma}
@@
 \end{proof}
+
+Similar results can be proved for other distributions and proposals, by identifying
+regions in product space $(\theta,\theta'-\theta,u)$ such that the hypothesis test
+needs to separate nearly-equal values.
+
+It follows that the accelerated M-H tests in \cite{icml2014c1_bardenet14} and
+\cite{cutting_mh_2014} require at least a constant fraction $\geq c$
+in the amount of data consumed per test, compared to full-dataset tests.
+i.e. their speed-up is at most $1/c$.
+
+\subsection{MCMC Posterior Inference}
 There is a separate line of MCMC work drawing principles from

### 2016-07-06 nips_2016.tex
most proofs done

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}
 
@@
 
-\subsubsection{Logistic Acceptance Function}\label{ssec:deltas_full}
-For our new MH test we use the logistic function: $g(s) =
-(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the condition in
-Lemma~\ref{lem:detailed_balance}.  While it is slightly less efficient than the Metropolis
-test when used on the full dataset, we will see soon that its smoothness allows it to naturally
-tolerate substantial variance in its input argument. This in turn will lead to a much more efficient
-test on subsets of data.
+\subsection{Logistic Acceptance Function}
+For our new MH test we use the logistic function:
+$g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it
+satisfies the condition in Lemma~\ref{lem:detailed_balance}.  While it
+is slightly less efficient than the Metropolis test when used on the
+full dataset, we will see soon that its smoothness allows it to
+naturally tolerate substantial variance in its input argument. This in
+turn will lead to a much more efficient test on subsets of data.
 
@@
 But to perform the test in equation (\ref{theexacttest}) we want $\Delta + X$ for
-a logistic random $X$ (call it $X_{\rm logistic}$ from now on). 
-Our approach is to decompose $X_{\rm logistic}$ as
+a logistic random $X$ (call it $X_{\rm log}$ from now on). 
+Our approach is to decompose $X_{\rm log}$ as
 \begin{equation}\label{eq:deconvolution}
-X_{\rm logistic} = X_{\rm norm}+X_{\rm corr},
-\end{equation}
-where we assume $X_{\rm norm}$ is a zero-mean Gaussian and $X_{\rm corr}$ is a zero-mean
-``correction'' term.  These two add (i.e., their distributions convolve) to form $X_{\rm logistic}$. The acceptance
-test is now:
+X_{\rm log} = X_{\rm norm}+X_{\rm corr},
+\end{equation}
+where we assume $X_{\rm norm}$ is a zero-mean Gaussian and $X_{\rm
+  corr}$ is a zero-mean ``correction'' term with distribution $C(X)$.
+The two variables are added (i.e., their distributions convolve) to
+form $X_{\rm log}$. It will be convenient to express this relationship
+in terms of the CDFs $N(X_{\rm norm}) = \Phi(X_{\rm norm}/\sigma)$ and
+$S(X_{\rm log})$, where $\Phi(.)$ is the Normal CDF with zero mean and
+unit variance and $S(.)$ is the standard logistic function.
+
+Equation (\ref{eq:deconvolution}) implies that $S = N * C$ where $*$ denotes convolution. We will later
+derive a numerical representation of $C$ by deconvolution, and using this we can generate
+samples of $X_{\rm corr}$.
+
+  The acceptance test is now:
 \begin{equation}\label{eq:criteria}
-\Delta + X_{\rm logistic} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
-\end{equation}
-i.e. given an approximate estimate $\Delta^*$ from the current data minibatch (assuming its
+\Delta + X_{\rm log} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
+\end{equation}
+i.e. given an estimate $\Delta^*$ from the current data minibatch (assuming its
 variance is small enough) we test acceptance by adding a random variable $X_{\rm corr}$ and
@@
 
-It is important to understand why we use Equations~\ref{eq:deconvolution} and~\ref{eq:criteria}. The
-deconvolution process allows us to use $\Delta' + X_{\rm corr}$ as a proxy for $\Delta + X$. It
-works because if we know the distribution\footnote{Computing the distributions of $\varepsilon$ and
-$X_{\rm norm}$ reduces to finding their standard deviations.} of $\varepsilon$ (from
-Equation~\ref{eq:relationship}), then we can pretend that our $\Delta'$ is in fact $\Delta$. To
-``insert'' $\varepsilon$ into our test, we deconvolve $X$ so that one of its components, $X_{\rm
-norm}$, plays the part of $\varepsilon$. In light of our revised MH test, the previous discussion
-raises the question: in an iteration of MCMC with current $\theta$ and proposed $\theta'$, how do
-we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample the
-\emph{values} of $X_{\rm corr}$ appropriately. This procedure consists of three major steps.
-
-\begin{enumerate}[noitemsep]
-    \item We first need to estimate the standard deviation of $X_{\rm norm}$ (i.e.
+%% It is important to understand why we use Equations~\ref{eq:deconvolution} and~\ref{eq:criteria}. The
+%% deconvolution process allows us to use $\Delta' + X_{\rm corr}$ as a proxy for $\Delta + X$. It
+%% works because if we know the distribution\footnote{Computing the distributions of $\varepsilon$ and
+%% $X_{\rm norm}$ reduces to finding their standard deviations.} of $\varepsilon$ (from
+%% Equation~\ref{eq:relationship}), then we can pretend that our $\Delta'$ is in fact $\Delta$. To
+%% ``insert'' $\varepsilon$ into our test, we deconvolve $X$ so that one of its components, $X_{\rm
+%% norm}$, plays the part of $\varepsilon$. In light of our revised MH test, the previous discussion
+%% raises the question: in an iteration of MCMC with current $\theta$ and proposed $\theta'$, how do
+%% we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample the
+%% \emph{values} of $X_{\rm corr}$ appropriately. This procedure consists of three major steps.
+
+%% \begin{enumerate}[noitemsep]
+%%     \item We first need to estimate the standard deviation of $X_{\rm norm}$ (i.e.
 the standard deviation of $\Delta'$).
@@
 
-	\item If the ${\rm std}(\Delta') > 1.2$, the decomposition of $X$ (i.e.
+%% 	\item If the ${\rm std}(\Delta') > 1.2$, the decomposition of $X$ (i.e.
 Equation~\ref{eq:deconvolution}) is unstable.
@@
 
-    \item Once we have ${\rm std}(\Delta')$, we can determine the distribution of $X_{\rm corr}$
-    from standard deconvolution techniques. These rely on the well-known fact that the Fourier
-    transform of a convolution is the product of the individual Fourier transforms. In practice, we
-    pre-compute distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to
-    facilitate the sampling process.
-\end{enumerate}
-
-Algorithm~\ref{alg:our_algorithm} describes our MH test within the MCMC algorithm.
-
-\subsection{Discussions and Practical Considerations of the New MH Test}\label{ssec:discussion}
+%%     \item Once we have ${\rm std}(\Delta')$, we can determine the distribution of $X_{\rm corr}$
+%%     from standard deconvolution techniques. These rely on the well-known fact that the Fourier
+%%     transform of a convolution is the product of the individual Fourier transforms. In practice, we
+%%     pre-compute distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to
+%%     facilitate the sampling process.
+%% \end{enumerate}
+
+Algorithm~\ref{alg:our_algorithm} describes the M-H iteration using our acceptance test.
 
@@
 \SetKwInOut{Output}{Output}
+
 \begin{algorithm}[t]
 \Input{Number of samples $T$, minibatch size $m$, pre-computed $X_{\rm
-corr}$ distribution, and initial sample $\theta_0$.}
-\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta\}$ from $p(\theta)$\;
-MAP/ML estimation.}
+corr}$ distribution, and initial sample $\theta_1$.}
+\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ from $p(\theta)$\;}
 \For{$t=\{1, \ldots, T\}$}{
-    Propose a candidate $\theta'$\ from proposal distribution\;
-    Draw a minibatch of $m$ samples, compute $\Delta'(\theta',\theta)$ and ${\rm std}(\Delta')$ for this sample\;
-    \While {${\rm std}(\Delta') \ge 1$} {
-    	Draw $m$ more samples to augment the current minibatch, and update $\Delta'$ and ${\rm std}(\Delta')$ \;
+    Propose a candidate $\theta'$\ from proposal distribution $q(\theta'\mid\theta_t)$\;
+    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(x,\theta_t,\theta')$ and sample variance $\sigma^2(\Delta^*)$\;
+    \While {$\sigma^2(\Delta^*) \ge 1$} {
+    	Draw $m$ more samples to augment the current minibatch, and update $\Delta^*$ and $\sigma^2(\Delta^*)$ \;
     }
-    Let $\sigma = \sqrt{1 - {\rm std}(\Delta')^2}$, draw a normal variable $X_n \in Norm(0,\sigma)$ and
+    Let $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$, draw a normal variable $X_n \in \mathcal{N}(0,\sigma_n^2)$ and
     $X_c$ from the correction distribution\;
 
-        \eIf{$X_c + X_n + \Delta' > 0$}{
+        \eIf{$\Delta^* + X_n + X_c > 0$}{
             Accept the candidate, $\theta_{t+1} = \theta'$\;
         }{
-            Reject and re-use the old sample, $\theta_{t+1} = \theta$\;
+            Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
         }
@@
 }
-\caption{A description of our MH test within the MCMC algorithm.}
+\caption{A description of M-H sampling with our acceptance test.}
 \label{alg:our_algorithm}
@@
 
-\textbf{The Convolution}. Figure~\ref{fig:deconvolution} demonstrates
-Equation~\ref{eq:deconvolution} and provides three (color-coded) examples of densities for $X_{\rm
-norm}$ and $X_{\rm corr}$. The density of $X$, a logistic random variable with mean $\mu = 0$ and
-scale $s=1$, is known, as are the $X_{\rm norm}$ densities for different standard deviations. The
-deconvolution provides us with the $X_{\rm corr}$ distributions. Note, however, that as the standard
-deviation of $X_{\rm norm}$ increases, the $X_{\rm corr}$ distribution becomes increasingly unstable
-and ``bumpier'', because the logistic curve has fatter tails than Gaussians. In addition, we cap
-the standard deviation possibilities of $X_{\rm norm}$ at $\approx 1.2$ in our experiments. In our
-experiments, we discretize the possible standard deviations of $X_{\rm norm}$, and we also limit the
-densities to consider the range $[-10,10]$ as shown in Figure~\ref{fig:deconvolution}.
-
-\textbf{Variance Preconditioning}. Returning to the discussion from Section~\ref{sec:introduction},
-our test requires a variance check. Equation~\ref{eq:deconvolution} means that the variance of $X$,
-which is $\pi^2/3\approx 3.29$ is an upper bound on the variances of $X_{\rm norm}$. Therefore, the
-test cannot work if the variance of $X_{\rm norm}$ is too large. The standard deviation of $X$
-is about $\sqrt{3.29}\approx 1.81$, so this bounds ${\rm std}(X_{\rm norm})$ (as well as ${\rm
-std}(\Delta')$). From the discussion above, we bound the standard deviation at 1.2 so that $X_{\rm
-corr}$ can handle the remaining noise. Recall from Section~\ref{ssec:deltas_minibatch} that we
-estimate the ${\rm std}(X_{\rm norm})$ each iteration. If the variance/standard deviation test
-fails, then we can (a) skip the iteration, (b) increase the minibatch size, (c) change the proposal
-to decrease step sizes, or (d) increase temperature, as we now discuss.
-
-\textbf{Temperature}. One way to satisfy the variance precondition is to increase the temperature of
-the target distribution. For a temperature $T>1$, the augmented target distribution would become
-\begin{equation}\label{eq:log_temperature}
-\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
-\end{equation}
-with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
-smaller in absolute value, thus reducing variance. The flatter posterior results in easier mixing.
-
-\section{Analysis}\label{sec:theory}
-
-In this section, we first perform an error analysis for our new
-acceptance test. Second, we follow the implications of this analysis for accuracy of
-the estimated target distribution. Third, we compare the speed and accuracy of
-our test with prior work. Due to space constraints, all our proofs
-are in Appendix~\ref{app:proofs}.
-
-We first introduce some notation. Let our true (full dataset) and
-approximate (minibatch) acceptance probabilities be $P_a = g(\Delta) =
-\Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} > 0)$.
-Define $X_{\xi} = \varepsilon - X_{\rm norm}$ which means, invoking
-Equations~\ref{eq:relationship} and~\ref{eq:deconvolution}, that
-$X_\xi$ is the random variable characterizing the accuracy of our
-$X_{\rm norm}$.  Thus, $\Delta' = \Delta + X_{\rm norm} + X_\xi$. We
-assume the standard deviation of $X_{\rm norm}$ is less than
-$\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}.
-We denote $F_X(x)$ as the CDF of $X$.
-
-We use some terms from standard Markov chain theory~\cite{Meyn2009}. Let the \emph{transition
-kernels} of MCMC on complete and minibatch data at step $i$ be $P_i$ and $P_i'$, respectively. When
-these kernels (e.g., $P_i$) are applied on a probability distribution $D_1$, they generate a new
-distribution $D_2$, which we write as $P_i \circ D_1 = D_2$. Define $\pi$ to be the stationary
-distribution obtained by $P_i$, and $\pi_i'$ to be the stationary distribution obtained by $P_i'$
-($\pi'$ needs a subscript in this case).  Finally, we indicate the \emph{total variation distance}
-between two kernels using $\| \cdot \|$.
-
-Our first result shows that the distance $\|P_i\circ D - P_i'\circ D\| \le 4\zeta \ell$ for all
-distributions $D$.
-
-% Our first result shows that the distance between $P_i$ and $P_i'$ is proportional to our estimation
-% error $X_{\xi}$.
-
-\begin{lemma}\label{lem:theory1}
-If $|X_\xi| < \zeta$ and $|\nabla F_X(x)| < \ell$ for all $x$, then $\|P_i\circ D-P_i'\circ D\| \le
-4\zeta \ell \;\;\forall D$.
+\section{Analysis}\label{sec:analysis}
+
+In this section we derive error bounds for our M-H test, and for the approximate target
+distribution that it generates. We discuss generation of the correction variable in the
+next section. To cut to the chase, it is possible to generate samples $X_{\rm corr}$ with
+a CDF error of at most single-precision floating point error, and possibly better. We
+therefore treat the $X_{\rm corr}$ variable as a sample from the exact correction distribution
+and we will not analyze its errors.
+
+In the most similar prior works, \cite{cutting_mh_2014} uses asymptotic arguments based
+on the central limit theorem to argue that its approximate acceptance test error tends to
+zero as batch size increases. But no quantitative bounds are given. In
+\cite{icml2014c1_bardenet14}, explicit bounds are given, but they depend on a bound on
+the quantity:
+\begin{equation}
+  \max_{1\leq i\leq N}|\log p(x_i|\theta') - \log p(x_i|\theta)|
+\end{equation}
+while such bounds can be derived easily for simple models, its unclear how to derive them
+for a complex model such as a neural network.
+
+In this paper, we rely on quantitave forms of the central limit theorem which rely on measurable
+statistics from the minibatch $\{x_i\}$. Thus a sampler using our approach does not need to ``see''
+data beyond the current minibatch. This supports use of the sampler on very large datasets, and
+in online mode where the dataset is presented as a stream.
+
+The outline of this section is as follows: we first present bounds on the absolute
+and relative error (in CDF) of the distribution of $\Delta^*$ vs a Normal distribution. We then show that
+these errors bounds are preserved after the addition of other random variables, in particular
+$X_n$ and $X_c$. From this it follows that the acceptance test has the same error bound.
+
+The quantitative central-limit result below comes from \cite{explicit-clt05}:
+
+\begin{lemma}\label{lem:quant_clt}
+  Let $X_1,\ldots,X_n$ be
+a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and
+variance $\sigma^2(\hat{X})$ where:
+\begin{equation}
+  \hat{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad   \sigma(\hat{X}) = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\hat{X})^2\right)^{\frac{1}{2}}
+\end{equation}
+then $t=\hat{X}/\sigma(\hat{X})$ has an approximate Student's distribution which approaches a normal
+distribution in the limit. Then from \cite{explicit-clt05}:
+\begin{equation}\label{eq:clt-bounds}
+ \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
+  \end{equation}
 \end{lemma}
 
-%% \begin{proof}
-%% See Appendix~\ref{app:theory1}.
-%% \end{proof}
-
-Thus, the accuracy of our transition kernel $P_i'$ is related to the quality of our estimated
-$X_{\rm norm}$. We next show that ${\rm Var}(\Delta')$ is roughly proportional to the step size of
-our proposer.
-
-\begin{lemma}\label{lem:theory2}
-For one step sampling from $\theta$ to $\theta'$, if the jumping step $\|\theta - \theta'\|_2 <
-\epsilon$, and the gradient of the log likelihood is bounded by a constant factor, i.e. $\|\nabla
-(\log \Pr(x_i\mid \theta))\|_2 < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2
-(m-\frac{m(m-1)}{N-1})$, where $m$ is the minibatch size, and $N$ is the total data size.
+So as long as we know the first and third absolute moments $E|X|$ and $E|X|^3$, we can bound the
+error of the normal approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of variables
+$y = x \sigma(\hat{X})$, equation (\ref{eq:clt-bounds}) becomes
+\begin{equation}\label{eq:clt-bounds2}
+   \sup_y\left|{\rm Pr}(\hat{X}< y) - \Phi\left(\frac{y}{\sigma(\hat{X})}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
+\end{equation}
+which shows that the distribution of $\hat{X}$ approaches the normal distribution $\mathcal{N}(0,\sigma^2(\hat{X}))$
+whose variance is $\sigma^2(\hat{X})$, measured from the sample.
+
+To apply this to our test, let $\mu = \frac{1}{N}\sum_{i=1}^N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$ which is
+the expected log data probability ratio across the entire dataset. Then
+Let $X_i = N(\log\frac{p(x_i|\theta')}{p(x_i|\theta)}-\mu)$, are zero-mean, IID variables.
+Take $n=b$ the minibatch size, and then:
+\begin{equation}
+  \Delta^*(x,\theta,\theta') = \hat{X} + \Delta(x,\theta,\theta')
+\end{equation}
+so that $\sigma(\hat{X}) = \sigma(\Delta^*)$.
+\begin{corollary}
+  We can now substitute into equation (\ref{eq:clt-bounds2}) and displace by the mean, giving:
+\begin{equation}\label{eq:clt-bounds3}
+   \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} = \epsilon(\theta,\theta',b)
+\end{equation}
+\end{corollary}
+
+We have shown that the distribution of $\Delta^*$ approximates a
+Normal distribution with mean $\Delta$ and variance
+$\sigma^2(\Delta^*)$, and bounded its error with estimable
+quantities. Both $E|X|$ and $E|X|^3$ can be estimated as means of
+$|X_i - \hat{X}|$ and $|X_i - \hat{X}|^3$ respectively on each minibatch. We expect this
+will often be accurate enough on minibatches with hundreds or
+thousands of points, but otherwise bootstrap CIs can be computed from
+those sequences. Since the bounds are monotone in $E|X|$ and $E|X|^3$,
+using upper bootstrap CI limits will provide high-confidence error bounds.
+
+\begin{lemma} Let $P(x)$ and $Q(x)$ be two cumulative distributions satisfying $\sup_x|P(x)-Q(x)|\leq \epsilon$ with
+  $x$ in some real range. Let $R(y)$ be the {\em density} of another
+  random variable $y$. Let $P'$ be the convolution $P*R$ and $Q'$ be
+  the convolution $Q*R$. Then $P'(z)$ (resp. $Q'(z)$) is the CDF of sum
+  $z=x+y$ of independent random variables $x$ with CDF $P(x)$ (resp. $Q(x)$) and y
+  with density $R(y)$.  Then
+  \begin{equation}
+    \sup_x|P'(x)-Q'(x)|\leq \epsilon
+  \end{equation}
   \end{lemma}
-
-%% \begin{proof}
-%% See Appendix~\ref{app:theory2}.
-%% \end{proof}
-
-Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 (m - \frac{m(m-1)}{N-1}) $, we can assume the maximum
-value of $X_{\xi}$ is proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we
-can assume $\zeta=\epsilon C$, where $C$ is a constant factor. We finally introduce
-Theorem~\ref{thm:theory3} to characterize the convergence of our samples.
-
-\begin{theorem}\label{thm:theory3}
-Assume $\pi$ satisfies $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$, where $\eta \in [0, 1)$ for
-all distributions $D_0$. Also assume there exists $0 < \rho_t < 1$ such that $\|P'_i \circ D -
-\pi_i'\| \leq 2\rho_i \;\; \forall D$. We can get $\| P_t' \circ P_{t-1}' \circ \cdots \circ P_1' \circ D_0 - \pi
-\| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $, where
-$\alpha_i = \frac{\epsilon_i C }{1-\eta}$ and $\epsilon_i$ is the step size at iteration $i$.
-\end{theorem}
-
-%% \begin{proof}
-%% See Appendix~\ref{app:theory3}.
-%% \end{proof}
-
-Theorem~\ref{thm:theory3} indicates that the difference between the target and sampled distributions
-consists of two components: $\rho_i$, which is determined by the efficiency of $P'_i$, and the
-$\alpha_i$s, which are determined by the error of our transition kernels. These involve competing
-objectives: the larger the step size, the more efficient the $P_i'$ but its error increases.  The
-reverse happens with a step size too small.
-
-We explore the efficiency of our minibatch MH test.
+\begin{proof}
+  \begin{equation}
+    \begin{split}
+      P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx
+    \end{split}
+  \end{equation}
+  and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
+  \begin{equation}
+- \epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon
+  \end{equation}
+\end{proof}
+
+\begin{corollary}
+  If $|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})| \leq \epsilon(\theta,\theta',b)$,
+  then
+  \begin{equation}
+    |{\rm Pr}(\Delta^*+X_n+X_c < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
+  \end{equation}
+  where $S(x)$ is the standard logistic function, and $X_n$ and $X_c$ are generated as per Algorithm 1. 
+  \end{corollary}
+%% \subsection{Discussions and Practical Considerations of the New MH Test}\label{ssec:discussion}
+\begin{proof}
+  Apply the previous lemma twice. First take:
+  \begin{equation}
+    P(y) = Pr(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
+  \end{equation}
+  and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$ where $\sigma_n^2
+  = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of $P$ and $Q$
+  \begin{equation}
+    P'(y) = Pr(\Delta^*+X_n < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
+  \end{equation}
+  Now we convolve with the distribution of $X_c$ which gives:
+  \begin{equation}
+    P''(y) = Pr(\Delta^*+X_n+X_c < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
+  \end{equation}
+  Both steps preserve the error bound $\epsilon(\theta,\theta',b)$
+\end{proof}
+
+%% \textbf{The Convolution}. Figure~\ref{fig:deconvolution} demonstrates
+%% Equation~\ref{eq:deconvolution} and provides three (color-coded) examples of densities for $X_{\rm
+%% norm}$ and $X_{\rm corr}$. The density of $X$, a logistic random variable with mean $\mu = 0$ and
+%% scale $s=1$, is known, as are the $X_{\rm norm}$ densities for different standard deviations. The
+%% deconvolution provides us with the $X_{\rm corr}$ distributions. Note, however, that as the standard
+%% deviation of $X_{\rm norm}$ increases, the $X_{\rm corr}$ distribution becomes increasingly unstable
+%% and ``bumpier'', because the logistic curve has fatter tails than Gaussians. In addition, we cap
+%% the standard deviation possibilities of $X_{\rm norm}$ at $\approx 1.2$ in our experiments. In our
+%% experiments, we discretize the possible standard deviations of $X_{\rm norm}$, and we also limit the
+%% densities to consider the range $[-10,10]$ as shown in Figure~\ref{fig:deconvolution}.
+
+%% \textbf{Variance Preconditioning}. Returning to the discussion from Section~\ref{sec:introduction},
+%% our test requires a variance check. Equation~\ref{eq:deconvolution} means that the variance of $X$,
+%% which is $\pi^2/3\approx 3.29$ is an upper bound on the variances of $X_{\rm norm}$. Therefore, the
+%% test cannot work if the variance of $X_{\rm norm}$ is too large. The standard deviation of $X$
+%% is about $\sqrt{3.29}\approx 1.81$, so this bounds ${\rm std}(X_{\rm norm})$ (as well as ${\rm
+%% std}(\Delta')$). From the discussion above, we bound the standard deviation at 1.2 so that $X_{\rm
+%% corr}$ can handle the remaining noise. Recall from Section~\ref{ssec:deltas_minibatch} that we
+%% estimate the ${\rm std}(X_{\rm norm})$ each iteration. If the variance/standard deviation test
+%% fails, then we can (a) skip the iteration, (b) increase the minibatch size, (c) change the proposal
+%% to decrease step sizes, or (d) increase temperature, as we now discuss.
+
+%% \textbf{Temperature}. One way to satisfy the variance precondition is to increase the temperature of
+%% the target distribution. For a temperature $T>1$, the augmented target distribution would become
+%% \begin{equation}\label{eq:log_temperature}
+%% \log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
+%% \end{equation}
+%% with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
+%% smaller in absolute value, thus reducing variance. The flatter posterior results in easier mixing.
+
+%% \section{Analysis}\label{sec:theory}
+
+%% In this section, we first perform an error analysis for our new
+%% acceptance test. Second, we follow the implications of this analysis for accuracy of
+%% the estimated target distribution. Third, we compare the speed and accuracy of
+%% our test with prior work. Due to space constraints, all our proofs
+%% are in Appendix~\ref{app:proofs}.
+
+%% We first introduce some notation. Let our true (full dataset) and
+%% approximate (minibatch) acceptance probabilities be $P_a = g(\Delta) =
+%% \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} > 0)$.
+%% Define $X_{\xi} = \varepsilon - X_{\rm norm}$ which means, invoking
+%% Equations~\ref{eq:relationship} and~\ref{eq:deconvolution}, that
+%% $X_\xi$ is the random variable characterizing the accuracy of our
+%% $X_{\rm norm}$.  Thus, $\Delta' = \Delta + X_{\rm norm} + X_\xi$. We
+%% assume the standard deviation of $X_{\rm norm}$ is less than
+%% $\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}.
+%% We denote $F_X(x)$ as the CDF of $X$.
+
+%% We use some terms from standard Markov chain theory~\cite{Meyn2009}. Let the \emph{transition
+%% kernels} of MCMC on complete and minibatch data at step $i$ be $P_i$ and $P_i'$, respectively. When
+%% these kernels (e.g., $P_i$) are applied on a probability distribution $D_1$, they generate a new
+%% distribution $D_2$, which we write as $P_i \circ D_1 = D_2$. Define $\pi$ to be the stationary
+%% distribution obtained by $P_i$, and $\pi_i'$ to be the stationary distribution obtained by $P_i'$
+%% ($\pi'$ needs a subscript in this case).  Finally, we indicate the \emph{total variation distance}
+%% between two kernels using $\| \cdot \|$.
+
+%% Our first result shows that the distance $\|P_i\circ D - P_i'\circ D\| \le 4\zeta \ell$ for all
+%% distributions $D$.
+
+%% % Our first result shows that the distance between $P_i$ and $P_i'$ is proportional to our estimation
+%% % error $X_{\xi}$.
+
+%% \begin{lemma}\label{lem:theory1}
+%% If $|X_\xi| < \zeta$ and $|\nabla F_X(x)| < \ell$ for all $x$, then $\|P_i\circ D-P_i'\circ D\| \le
+%% 4\zeta \ell \;\;\forall D$.
+%% \end{lemma}
+
+%% %% \begin{proof}
+%% %% See Appendix~\ref{app:theory1}.
+%% %% \end{proof}
+
+%% Thus, the accuracy of our transition kernel $P_i'$ is related to the quality of our estimated
+%% $X_{\rm norm}$. We next show that ${\rm Var}(\Delta')$ is roughly proportional to the step size of
+%% our proposer.
+
+%% \begin{lemma}\label{lem:theory2}
+%% For one step sampling from $\theta$ to $\theta'$, if the jumping step $\|\theta - \theta'\|_2 <
+%% \epsilon$, and the gradient of the log likelihood is bounded by a constant factor, i.e. $\|\nabla
+%% (\log \Pr(x_i\mid \theta))\|_2 < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2
+%% (m-\frac{m(m-1)}{N-1})$, where $m$ is the minibatch size, and $N$ is the total data size.
+%% \end{lemma}
+
+%% %% \begin{proof}
+%% %% See Appendix~\ref{app:theory2}.
+%% %% \end{proof}
+
+%% Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 (m - \frac{m(m-1)}{N-1}) $, we can assume the maximum
+%% value of $X_{\xi}$ is proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we
+%% can assume $\zeta=\epsilon C$, where $C$ is a constant factor. We finally introduce
+%% Theorem~\ref{thm:theory3} to characterize the convergence of our samples.
+
+%% \begin{theorem}\label{thm:theory3}
+%% Assume $\pi$ satisfies $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$, where $\eta \in [0, 1)$ for
+%% all distributions $D_0$. Also assume there exists $0 < \rho_t < 1$ such that $\|P'_i \circ D -
+%% \pi_i'\| \leq 2\rho_i \;\; \forall D$. We can get $\| P_t' \circ P_{t-1}' \circ \cdots \circ P_1' \circ D_0 - \pi
+%% \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $, where
+%% $\alpha_i = \frac{\epsilon_i C }{1-\eta}$ and $\epsilon_i$ is the step size at iteration $i$.
+%% \end{theorem}
+
+%% %% \begin{proof}
+%% %% See Appendix~\ref{app:theory3}.
+%% %% \end{proof}
+
+%% Theorem~\ref{thm:theory3} indicates that the difference between the target and sampled distributions
+%% consists of two components: $\rho_i$, which is determined by the efficiency of $P'_i$, and the
+%% $\alpha_i$s, which are determined by the error of our transition kernels. These involve competing
+%% objectives: the larger the step size, the more efficient the $P_i'$ but its error increases.  The
+%% reverse happens with a step size too small.
+
+%% We explore the efficiency of our minibatch MH test.
 Theorem~\ref{thm:efficiency} indicates that our minibatch MH test only requires a constant number of instances for each step, however, recall the Lemma~\ref{lem:efficienty_cutting}, the adaptive minibatch MH test requires $O(\sqrt{N})$ number of data instance for each step.
@@
 
-\begin{theorem}\label{thm:efficiency}
-We require $\frac{ C^2 \rho^2 }{\sigma^6 \varepsilon^2}$ minibatch size to ensure $|P_a -P_a'| < \varepsilon$, where $C$ is a positive constant which is smaller than 0.4748, $\sigma^2 $ and $\rho$ is the second and third moment of $(\nabla\log \Pr(x_i|\theta) - \nabla\log\Pr(x_i|\theta')) - E\left[ \nabla\log \Pr(x_i|\theta) - \nabla\log\Pr(x_i|\theta')\right]$, $P_a$ is the accept rate from the complete data, $P_a'$ is the accept rate from minibatch data by our algorithm, $\theta$ is the current parameter, and $\theta'$ is the proposed parameter.
-
-\end{theorem}
+%% \begin{theorem}\label{thm:efficiency}
+%% We require $\frac{ C^2 \rho^2 }{\sigma^6 \varepsilon^2}$ minibatch size to ensure $|P_a -P_a'| < \varepsilon$, where $C$ is a positive constant which is smaller than 0.4748, $\sigma^2 $ and $\rho$ is the second and third moment of $(\nabla\log \Pr(x_i|\theta) - \nabla\log\Pr(x_i|\theta')) - E\left[ \nabla\log \Pr(x_i|\theta) - \nabla\log\Pr(x_i|\theta')\right]$, $P_a$ is the accept rate from the complete data, $P_a'$ is the accept rate from minibatch data by our algorithm, $\theta$ is the current parameter, and $\theta'$ is the proposed parameter.
+
+%% \end{theorem}
 
@@
 \begin{itemize}[noitemsep]
-    \item Proofs for material in Section~\ref{sec:theory}.
     \item More information on Section~\ref{ssec:gaussians}.
@@
 
-\section{Proofs}\label{app:proofs}
-
-\subsection{Proof of Lemma~\ref{lem:efficienty_cutting}}\label{app:lem_efficiency}
-
-If the minibatch size is $n$, we first compute the probability that more than $n$ data instances are required for adaptive sampling method for one iteration. We denote this probability $P_{rej}$.
-\[P_{rej} = \Pr \left ( 1-\phi_{n-1} \left( \frac{|\bar{l} - \frac{1}{N} \log (u\frac{p(\theta)q(\theta'|\theta)}{p(\theta')q(\theta|\theta')})|}{s} \right )  \geq \varepsilon \right ) \]
-
-where $s = \frac{s_l}{\sqrt{n}}\sqrt{1-\frac{n-1}{N-1}}$. \\
-We define random variable $Z = \frac{1}{N}\log (uQ)$, where $Q=\frac{p(\theta)q(\theta'|\theta)}{p(\theta')q(\theta|\theta')}$.
+%% \section{Proofs}\label{app:proofs}
+
+%% \subsection{Proof of Lemma~\ref{lem:efficienty_cutting}}\label{app:lem_efficiency}
+
+%% If the minibatch size is $n$, we first compute the probability that more than $n$ data instances are required for adaptive sampling method for one iteration. We denote this probability $P_{rej}$.
+%% \[P_{rej} = \Pr \left ( 1-\phi_{n-1} \left( \frac{|\bar{l} - \frac{1}{N} \log (u\frac{p(\theta)q(\theta'|\theta)}{p(\theta')q(\theta|\theta')})|}{s} \right )  \geq \varepsilon \right ) \]
+
+%% where $s = \frac{s_l}{\sqrt{n}}\sqrt{1-\frac{n-1}{N-1}}$. \\
+%% We define random variable $Z = \frac{1}{N}\log (uQ)$, where $Q=\frac{p(\theta)q(\theta'|\theta)}{p(\theta')q(\theta|\theta')}$.
 Then $Z$ has probability density function $f_Z(z) = N \exp(zN), z \in (-\infty, \frac{1}{N} \log Q]$.
 
-\begin{align*} 
-P_{rej}&=  \Pr \left( 1-\varepsilon \geq \phi_{n-1} \left( \frac{|\bar{l} - Z|}{s_l \sqrt{\frac{N-n}{N-1}}} \sqrt{n} \right) \right) \\ 
-&=  \Pr \left(\phi^{-1}_{n-1} \left (1-\varepsilon \right) \frac{s_l }{\sqrt{n}} \sqrt{\frac{N-n}{N-1}} \geq |\bar{l} -Z| \right) = \Pr \left( |\bar{l} - Z| \leq \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \right) \\
-&= \Pr \left(Z - \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \leq \bar{l} \leq Z + \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \right) \\
-&= \int_{-\infty} ^{\frac{\log Q}{N}} N \exp (zN)dz \int_{z - \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}}} ^{z+ \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}}} \frac{1}{2L} d\bar{l} = \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \int_{-\infty} ^{\frac{\log Q}{N}} N \exp(zN) dz \\
-&= \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} Q
-\end{align*}
- If we define the random variable $X$ the number of instance for one minibatch, we have $\Pr (X>n) = P_{rej} = \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} Q, \mbox{ for } n \in [1,N], n \in Z^+$.
+%% \begin{align*} 
+%% P_{rej}&=  \Pr \left( 1-\varepsilon \geq \phi_{n-1} \left( \frac{|\bar{l} - Z|}{s_l \sqrt{\frac{N-n}{N-1}}} \sqrt{n} \right) \right) \\ 
+%% &=  \Pr \left(\phi^{-1}_{n-1} \left (1-\varepsilon \right) \frac{s_l }{\sqrt{n}} \sqrt{\frac{N-n}{N-1}} \geq |\bar{l} -Z| \right) = \Pr \left( |\bar{l} - Z| \leq \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \right) \\
+%% &= \Pr \left(Z - \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \leq \bar{l} \leq Z + \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \right) \\
+%% &= \int_{-\infty} ^{\frac{\log Q}{N}} N \exp (zN)dz \int_{z - \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}}} ^{z+ \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}}} \frac{1}{2L} d\bar{l} = \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \int_{-\infty} ^{\frac{\log Q}{N}} N \exp(zN) dz \\
+%% &= \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} Q
+%% \end{align*}
+%%  If we define the random variable $X$ the number of instance for one minibatch, we have $\Pr (X>n) = P_{rej} = \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} Q, \mbox{ for } n \in [1,N], n \in Z^+$.
 Since $X$ is nonnegative, we have $E(X) \geq \Pr (X >x)x, \forall x$.
@@
 
-\subsection{Proof of Lemma~\ref{lem:theory1}}\label{app:theory1}
-
-We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta
-\ell$ for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
-\begin{align*}
-|P_a - P_a'| &= |\Pr(\Delta + X >0) - \Pr(\Delta' + X_{\rm corr} > 0)|\\
-& = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
-&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| \\
-&= |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
-&= |\nabla F(-\Delta) \xi + o(\xi^2)|\\
-&\leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2\ell \xi,
-\end{align*}
-where we use Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have smaller absolute
-value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| < 2 \zeta \ell$.
-
-Next, we show the distance of new distributions obtained by applying $P_i$ and $P_i'$ on an
-arbitrary distribution $D$ once is bounded if we use the same proposer, i.e.  $\|P_i\circ D -
-P'_i\circ D\| \leq 4\zeta \ell \;\; \forall D$. We write $P_i(\theta \mid \theta') = P_a(\theta, \theta')
-q(\theta'\mid \theta) + (1-P_a(\theta,\theta')) \delta_D(\theta' - \theta)$, where $\delta_D$
-is the Dirac delta function. The total variation distance between two generated distributions is
-\begin{align*}
-\|P_i\circ D - P'_i\circ D\| &= \int_{\theta'} \left| \int_{\theta}((P_a-P_a') q(\theta'|\theta) + (1-P_a - 1+P_a') \delta_D(\theta -\theta')) dD(\theta) \right| d\theta' \\
-&= \int_{\theta'} \left| \int_{\theta} (P_a- P_a')(q(\theta'|\theta) - \delta_D(\theta - \theta')) dD(\theta) \right|  d\theta'\\ 
-&\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta}(q(\theta'|\theta) + \delta_D(\theta - \theta')) dD(\theta) \right| d\theta'\\
-&\leq 2 \zeta \ell \int_{\theta'} ( D'(\theta') + D(\theta') ) d\theta'\\
-&= 4 \zeta \ell,
-\end{align*}
-where $D'$ is the distribution generated if we always accept the proposed point. (In the last line,
-we integrate two probability density functions, resulting in a value of two.) This matches our
-desired bound.
-
-\subsection{Proof of Lemma~\ref{lem:theory2}}\label{app:theory2}
-
-From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
-\theta))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
-$g_i(\theta) = \log p(x_i\mid \theta)$, and $q_i = g_i(\theta') - g_i(\theta)$. Then by Taylor's Theorem:
-\[
-|q| = |g(\theta') - g(\theta)| = |\nabla g(\theta)^T(\theta' - \theta) + o(\theta' - \theta)^2| \leq 2\epsilon k.
-\]
-Thus, given $\theta$ and $\theta'$, the variance of $\Delta'$ can be written as:
-\begin{align*}
-{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta)\right] = {\rm Var}\left(\sum_{i=1}^m q_i\right) \\
-&= m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
-&= m{\rm Var}(q_i) - \frac{m(m-1)}{N-1} {\rm Var}(q_i) \\
-&\leq 4\left(m - \frac{m(m-1)}{N-1}\right)k^2 \epsilon^2,
-\end{align*}
-as desired.
-
-\subsection{Proof of Theorem~\ref{thm:theory3}}\label{app:theory3}
-
-First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0
-- \pi\|$ and $\|P_i\circ D-P'_i\circ D\| \leq \epsilon_i C$ for all $D$. The norm between the
-stationary distributions is $\|\pi - \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
-
-Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have 
-\[
-\| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t \| \leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t
-\rho_u (1-\alpha_u)\right\} \rho_s \alpha_s,
-\]
-where $\alpha = \frac{\epsilon_i C}{1-\eta}$.  Thus, we get:
-\begin{align*}
- \| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_0 \| &\leq \|P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
- &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
-\end{align*}
-as desired.
-
-\subsection{Proof of Theorem~\ref{thm:efficiency}}\label{app:theory4}
-Since only the $\frac{N}{nT}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta) \right)$ term brings randomness into $\Delta'$.
+%% \subsection{Proof of Lemma~\ref{lem:theory1}}\label{app:theory1}
+
+%% We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta
+%% \ell$ for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
+%% \begin{align*}
+%% |P_a - P_a'| &= |\Pr(\Delta + X >0) - \Pr(\Delta' + X_{\rm corr} > 0)|\\
+%% & = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
+%% &= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| \\
+%% &= |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
+%% &= |\nabla F(-\Delta) \xi + o(\xi^2)|\\
+%% &\leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2\ell \xi,
+%% \end{align*}
+%% where we use Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have smaller absolute
+%% value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| < 2 \zeta \ell$.
+
+%% Next, we show the distance of new distributions obtained by applying $P_i$ and $P_i'$ on an
+%% arbitrary distribution $D$ once is bounded if we use the same proposer, i.e.  $\|P_i\circ D -
+%% P'_i\circ D\| \leq 4\zeta \ell \;\; \forall D$. We write $P_i(\theta \mid \theta') = P_a(\theta, \theta')
+%% q(\theta'\mid \theta) + (1-P_a(\theta,\theta')) \delta_D(\theta' - \theta)$, where $\delta_D$
+%% is the Dirac delta function. The total variation distance between two generated distributions is
+%% \begin{align*}
+%% \|P_i\circ D - P'_i\circ D\| &= \int_{\theta'} \left| \int_{\theta}((P_a-P_a') q(\theta'|\theta) + (1-P_a - 1+P_a') \delta_D(\theta -\theta')) dD(\theta) \right| d\theta' \\
+%% &= \int_{\theta'} \left| \int_{\theta} (P_a- P_a')(q(\theta'|\theta) - \delta_D(\theta - \theta')) dD(\theta) \right|  d\theta'\\ 
+%% &\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta}(q(\theta'|\theta) + \delta_D(\theta - \theta')) dD(\theta) \right| d\theta'\\
+%% &\leq 2 \zeta \ell \int_{\theta'} ( D'(\theta') + D(\theta') ) d\theta'\\
+%% &= 4 \zeta \ell,
+%% \end{align*}
+%% where $D'$ is the distribution generated if we always accept the proposed point. (In the last line,
+%% we integrate two probability density functions, resulting in a value of two.) This matches our
+%% desired bound.
+
+%% \subsection{Proof of Lemma~\ref{lem:theory2}}\label{app:theory2}
+
+%% From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
+%% \theta))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
+%% $g_i(\theta) = \log p(x_i\mid \theta)$, and $q_i = g_i(\theta') - g_i(\theta)$. Then by Taylor's Theorem:
+%% \[
+%% |q| = |g(\theta') - g(\theta)| = |\nabla g(\theta)^T(\theta' - \theta) + o(\theta' - \theta)^2| \leq 2\epsilon k.
+%% \]
+%% Thus, given $\theta$ and $\theta'$, the variance of $\Delta'$ can be written as:
+%% \begin{align*}
+%% {\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta)\right] = {\rm Var}\left(\sum_{i=1}^m q_i\right) \\
+%% &= m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
+%% &= m{\rm Var}(q_i) - \frac{m(m-1)}{N-1} {\rm Var}(q_i) \\
+%% &\leq 4\left(m - \frac{m(m-1)}{N-1}\right)k^2 \epsilon^2,
+%% \end{align*}
+%% as desired.
+
+%% \subsection{Proof of Theorem~\ref{thm:theory3}}\label{app:theory3}
+
+%% First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0
+%% - \pi\|$ and $\|P_i\circ D-P'_i\circ D\| \leq \epsilon_i C$ for all $D$. The norm between the
+%% stationary distributions is $\|\pi - \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
+
+%% Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have 
+%% \[
+%% \| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t \| \leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t
+%% \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s,
+%% \]
+%% where $\alpha = \frac{\epsilon_i C}{1-\eta}$.  Thus, we get:
+%% \begin{align*}
+%%  \| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_0 \| &\leq \|P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
+%%  &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
+%% \end{align*}
+%% as desired.
+
+%% \subsection{Proof of Theorem~\ref{thm:efficiency}}\label{app:theory4}
+%% Since only the $\frac{N}{nT}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta) \right)$ term brings randomness into $\Delta'$.
 Define $X_i = \frac{N}{T}\left(\log p(x_i|\theta') - \log p (x_i|theta_t) \right)$, $\tilde{X_i} = X_i - E(X_i)$, and $Y_n = \frac{\sum_{i=1}^n \tilde{X_i}}{n}$ (i.e.
@@
 
-From Equation~\ref{eq:criteria}, we show we want to use $\Delta' + X_{\rm corr}$ to approximate $\Delta + X_{\rm norm} + X_{\rm corr}$.
+%% From Equation~\ref{eq:criteria}, we show we want to use $\Delta' + X_{\rm corr}$ to approximate $\Delta + X_{\rm norm} + X_{\rm corr}$.
 Since $|F_{\Delta'} - F_{\Delta + X_{\rm norm}}| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$, for the worst case, assume $\exists \xi, ~{\rm s.t.} |f_{\Delta'}(\xi) - f_{\Delta + X_{\rm norm}} (\xi)| = \max_x |f_{\Delta'}(x) - f_{\Delta + X_{\rm norm}}(x)| \overset{\Delta}{=} \eta $, where $f()$ represents the PDF function.
 Because PDF of $X_{\rm corr}$ has the positive value and unit area, we have:
-\begin{align*} 
-& |(f_{\Delta'}*f_{X_{\rm corr}})(\xi) - (f_{\Delta+X_{\rm norm}}*f_{X_{\rm corr}})(\xi)| \\
-& = |\int_{-\infty}^{\infty}f_{\Delta'}(\xi-\tau)f_{X_{\rm corr}}(\tau)d\tau - \int_{-\infty}^{\infty}f_{\Delta + X_{\rm norm}} (\xi -\tau) f_{X_{\rm corr}} (\tau) d\tau| \\
-& = |\int_{-\infty}^{\infty} (f_{\Delta'}(\xi-\tau) - f_{\Delta + X_{\rm norm}} (\xi -\tau)) f_{X_{\rm corr}} (\tau) d\tau| \\
-& \leq \int_{-\infty}^{\infty} |f_{\Delta'}(\xi-\tau) - f_{\Delta + X_{\rm norm}} (\xi -\tau) |f_{X_{\rm corr}} (\tau) d\tau  \\
-& \leq \int_{-\infty}^{\infty} \eta f_{X_{\rm corr}} (\tau) d\tau \leq \eta
-\end{align*}
-
-where ``*'' represents the convolution operation.
-
-Thus, after convolution with $X_{\rm corr}$, the maximum gap between PDF of the $\Delta'$ and $\Delta + X_{\rm norm}$ decreases.
+%% \begin{align*} 
+%% & |(f_{\Delta'}*f_{X_{\rm corr}})(\xi) - (f_{\Delta+X_{\rm norm}}*f_{X_{\rm corr}})(\xi)| \\
+%% & = |\int_{-\infty}^{\infty}f_{\Delta'}(\xi-\tau)f_{X_{\rm corr}}(\tau)d\tau - \int_{-\infty}^{\infty}f_{\Delta + X_{\rm norm}} (\xi -\tau) f_{X_{\rm corr}} (\tau) d\tau| \\
+%% & = |\int_{-\infty}^{\infty} (f_{\Delta'}(\xi-\tau) - f_{\Delta + X_{\rm norm}} (\xi -\tau)) f_{X_{\rm corr}} (\tau) d\tau| \\
+%% & \leq \int_{-\infty}^{\infty} |f_{\Delta'}(\xi-\tau) - f_{\Delta + X_{\rm norm}} (\xi -\tau) |f_{X_{\rm corr}} (\tau) d\tau  \\
+%% & \leq \int_{-\infty}^{\infty} \eta f_{X_{\rm corr}} (\tau) d\tau \leq \eta
+%% \end{align*}
+
+%% where ``*'' represents the convolution operation.
+
+%% Thus, after convolution with $X_{\rm corr}$, the maximum gap between PDF of the $\Delta'$ and $\Delta + X_{\rm norm}$ decreases.
 Thus, $|F_{\Delta' + X_{\rm corr}} - F_{\Delta + X_{\rm norm} + X_{\rm corr}}| \leq |F_{\Delta'} - F_{\Delta + X_{\rm norm}}| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$.

### 2016-07-11 nips_2016.tex
added more test scripts

@@
 \begin{equation}\label{eq:traditional}
-\alpha(\theta_t,\theta') = \min\left\{ 1, \frac{p(\theta')q(\theta_t \mid \theta')}{p(\theta_t)q(\theta' \mid \theta_t)}
-\right\}
-\end{equation}
-and with probability $\alpha(\theta_t,\theta')$, we accept $\theta'$
+\alpha(\theta_t,\theta') = \frac{p(\theta')q(\theta_t \mid \theta')}{p(\theta_t)q(\theta' \mid \theta_t)} \wedge 1
+\end{equation}
+where $a \wedge b$ denotes $\min(a,b)$.  With probability $\alpha(\theta_t,\theta')$, we accept $\theta'$
 and set $\theta_{t+1} = \theta'$, otherwise set
 $\theta_{t+1}=\theta_t$.  The test is often implemented with an
-auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison $u < \alpha(\theta_t,\theta')$.
+auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison $u < \alpha(\theta_t,\theta')$. $\mathcal{U}(a,b)$ denotes the uniform distribution on the interval $[a,b]$. 
 For notational simplicity, from now on we will drop the subscript $t$ for the current sample $\theta_t$ and
@@
 \begin{equation}\label{detailed_balance1}
-  p(\theta)p(\theta,\theta') = p(\theta')p(\theta',\theta)
-\end{equation}
-where $p(\theta,\theta')$ is the probability of a transition from state $\theta$ to $\theta'$. Here
-$p(\theta,\theta') = q(\theta'\mid\theta)\alpha(\theta,\theta')$ so the detailed balance equation becomes
+  p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')
+\end{equation}
+where $p(\theta'\mid\theta)$ is the probability of a transition from state $\theta$ to $\theta'$. Here
+$p(\theta'\mid\theta) = q(\theta'\mid\theta)\alpha(\theta,\theta')$ so the detailed balance equation becomes
 \begin{equation}\label{detailed_balance2}
@@
 \alpha(\theta,\theta') = 
-\min\left\{ 1, \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
-\theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)} \right\},
+ \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
+\theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)} \wedge 1
 \end{equation}
@@
 Following ~\cite{icml2014c1_bardenet14}, we write the test $u < \alpha(\theta,\theta')$ in
-the equivalent form $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$ where
-\begin{equation}
-\Lambda(\theta,\theta') = \frac{1}{N}\sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+the equivalent form $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$. Our definitions differ
+from those in \cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later on. 
+\begin{equation}
+\Lambda(\theta,\theta') = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
 \hspace{0.2in} {\rm and} \hspace{0.2in}
-\psi(u,\theta,\theta') = \frac{1}{N}\log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right)
+\psi(u,\theta,\theta') = \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right)
 \end{equation}
@@
 \begin{equation}
-\Lambda^*(\theta,\theta') = \frac{1}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
-\end{equation}
-
-The minibatch should contain randomly selected samples, and
-by the central limit theorem, we expect $\Lambda^*(\theta,\theta')$ to be approximately
-Normal. 
-The acceptance test then becomes a statistical test of the hypothesis
-that $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$ by
-establishing that $\Lambda^*(\theta,\theta')$ is substantially larger
-than $\psi(u,\theta,\theta')$.  In ~\cite{cutting_mh_2014} an
-asymptotic central limit argument was used to derive this gap, while
-in \cite{icml2014c1_bardenet14} a concentration bound was used. In
-both cases, the resulting tests were shown to give useful reductions
-in number of samples required over using the full dataset. But no
+\Lambda^*(\theta,\theta') = \frac{N}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+\end{equation}
+
+Finally, it will be convenient in what follows to define the individual components that
+contribute to the sums above:
+
+\begin{equation}\label{eq:individual_terms}
+\Lambda_i(\theta,\theta') = N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+\end{equation}
+
+So $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over the
+entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean over the minibatch
+of $\Lambda_i(\theta,\theta')$.
+
+Since the minibatch contains randomly selected samples $x_i$, the
+values $\Lambda_i$ are independent, IID random variables.  By the central
+limit theorem, we expect $\Lambda^*(\theta,\theta')$ to be
+approximately Normal.  The acceptance test then becomes a statistical
+test of the hypothesis that $\Lambda(\theta,\theta') >
+\psi(u,\theta,\theta')$ by establishing that
+$\Lambda^*(\theta,\theta')$ is substantially larger than
+$\psi(u,\theta,\theta')$.  In ~\cite{cutting_mh_2014} an asymptotic
+central limit argument was used to derive this gap, while in
+\cite{icml2014c1_bardenet14} a concentration bound was used. In both
+cases, the resulting tests were shown to give useful reductions in
+number of samples required over using the full dataset. But no
 worst-case bounds were given on average number of samples needed.
@@
 \begin{equation}
-\Lambda^*(\theta,\theta') =  \frac{1}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)}=
+\Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)}=
   %%\frac{1}{2}\left( (x_i-\theta')^2 - (x_i - \theta)^2\right) =
-  (\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
+  N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
 \end{equation}
@@
 Since the $x_i$ have unit variance, their mean has variance $1/b$, and the variance $\sigma^2(\Lambda^*)$
-of $\Lambda^*(\theta,\theta')$ is $(\theta'-\theta)^2/b$. In order to pass
+of $\Lambda^*(\theta,\theta')$ is $(\theta'-\theta)^2N/b$. In order to pass
 a hypothesis test that $\Lambda > \psi$, we will need to establish a
@@
 variance $1/N$.  The target distribution is therefore
-$\mathcal{N}(\mu,1/N)$ where $\mu$ is the sample mean.  An efficient
+$\mathcal{N}(\mu,1/N)$ where $\mu$ is the sample mean, and $\mathcal{N}(\mu,\sigma^2)$ is a
+normal distribution with mean $\mu$ and variance $\sigma^2$.  An efficient
 proposal distribution has the same variance as the target distribution
@@
 function with zero mean and unit variance. This proposal is symmetric
-$q(\theta',\theta)=q(\theta,\theta')$, and we assumed uniform
-prior for $\theta$, so $\psi(u,\theta',\theta)$ reduces to $1/N\log
+$q(\theta'\mid\theta)=q(\theta\mid\theta')$, and we assumed uniform
+prior for $\theta$, so $\psi(u,\theta',\theta)$ reduces to $\log
 u$.
@@
   \end{equation}
-  the size of the range in $u$ is at least $\exp([-2,-1.125]) = [0082,0.286]$ and occurs
-  with probability at least $p_2=0.2044$. With $u$ in this range, we rewrite the test as:
+  the size of the range in $u$ is at least $\exp([-2,-1.125]) = [0.13534,0.32465]$ and occurs
+  with probability at least $p_2=0.18932$. With $u$ in this range, we rewrite the test as:
   \begin{equation}
@@
   and $u$ are drawn independently and so the probability of the conjunction of these events is
-  $c = p_0 p_1 p_2 = 0.009$.
+  $c = p_0 p_1 p_2 = 0.0085$.
 \end{proof}
@@
 \begin{equation}\label{eq:delta1}
-\Delta(x,\theta,\theta')  =
+\Delta(\theta,\theta')  =
 \log \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
@@
 \begin{equation}\label{eq:delta2}
-\Delta(x,\theta,\theta') = N\psi(1,\theta,\theta') + \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)}
+  \Delta(\theta,\theta') = \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') =
+  \Lambda(\theta,\theta') -  \psi(1,\theta,\theta')
 \end{equation}
@@
 \begin{equation}\label{eq:delta3}
-    \Delta^*(x,\theta,\theta') =
-N\psi(1,\theta,\theta') + \frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)}
+    \Delta^*(\theta,\theta') =
+\frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') =
+\Lambda^*(\theta,\theta') -     \psi(1,\theta,\theta')
 \end{equation}
 Note that $\Delta$ and $\Delta^*$ are evaluated on the full dataset and a minibatch of size $b$
-respectively. The scaling term $\frac{N}{b}$ ensures that $\Delta^*(x,\theta,\theta')$ is an
-unbiased estimator of $\Delta(x,\theta,\theta')$.
+respectively. The scaling term $\frac{N}{b}$ ensures that $\Delta^*(\theta,\theta')$ is an
+unbiased estimator of $\Delta(\theta,\theta')$.
 
@@
 
-As a sanity check, choosing $g(s) = \min\{1, \exp(s) \}$ produces the
+As a sanity check, choosing $g(s) = \exp(s) \wedge 1$ produces the
 standard Metropolis acceptance test $\alpha(\theta,\theta') = g(\Delta(\theta,\theta')) = 
-\min\left\{1, \frac{p(\theta')q(\theta \mid
-  \theta')}{p(\theta)q(\theta' \mid \theta)}\right\}$. $g(s)$ satisfies
+\frac{p(\theta')q(\theta \mid
+  \theta')}{p(\theta)q(\theta' \mid \theta)}\wedge 1$. $g(s)$ satisfies
 the condition $g(s) = \exp(s)g(-s)$, and therefore detailed
@@
 We now move to acceptance testing using the minibatch estimator
-$\Delta^*(x,\theta,\theta')$. From equation (\ref{eq:delta3}),
-$\Delta^*(x,\theta,\theta')$ can be represented as a constant term
-plus a sum of $b$ iid terms of the form
-$\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $N$ increases, $\Delta^*(\theta,\theta')$
+$\Delta^*(\theta,\theta')$. From equation (\ref{eq:delta3}),
+$\Delta^*(\theta,\theta')$ can be represented as a constant term
+plus the mean of $b$ IID terms $\Lambda_i(\theta,\theta')$ of the form
+$N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $b$ increases, $\Delta^*(\theta,\theta')$
 therefore has a distribution which approaches a normal distribution by the central limit
@@
 \end{equation}
-where we assume $X_{\rm norm}$ is a zero-mean Gaussian and $X_{\rm
-  corr}$ is a zero-mean ``correction'' term with distribution $C(X)$.
-The two variables are added (i.e., their distributions convolve) to
-form $X_{\rm log}$. It will be convenient to express this relationship
-in terms of the CDFs $N(X_{\rm norm}) = \Phi(X_{\rm norm}/\sigma)$ and
-$S(X_{\rm log})$, where $\Phi(.)$ is the Normal CDF with zero mean and
-unit variance and $S(.)$ is the standard logistic function.
-
-Equation (\ref{eq:deconvolution}) implies that $S = N * C$ where $*$ denotes convolution. We will later
+where we assume $X_{\rm norm}$ has zero-mean, Normal distribution with
+variance $\sigma^2$, and $X_{\rm corr}$ is a zero-mean ``correction''
+term with density $C_{\sigma}(X)$.  The two variables are added
+(i.e., their distributions convolve) to form $X_{\rm log}$. It will be
+convenient to express this relationship in terms of the CDFs $N_{\sigma}(X_{\rm
+  norm}) = \Phi(X_{\rm norm}/\sigma)$ and $S(X_{\rm log})$, where
+$\Phi(.)$ is the Normal CDF with zero mean and variance $\sigma^2$ and
+$S(.)$ is the standard logistic function.
+
+Equation (\ref{eq:deconvolution}) implies that $S = N_{\sigma} * C_{\sigma}$ where $*$ denotes convolution. We will later
 derive a numerical representation of $C$ by deconvolution, and using this we can generate
@@
 \begin{algorithm}[t]
-\Input{Number of samples $T$, minibatch size $m$, pre-computed $X_{\rm
-corr}$ distribution, and initial sample $\theta_1$.}
+\Input{Number of samples $T$, minibatch size $m$, error bound $\delta$, pre-computed correction $C_1(X)$ distribution, initial sample $\theta_1$.}
 \Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ from $p(\theta)$\;}
@@
     Propose a candidate $\theta'$\ from proposal distribution $q(\theta'\mid\theta_t)$\;
-    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(x,\theta_t,\theta')$ and sample variance $\sigma^2(\Delta^*)$\;
-    \While {$\sigma^2(\Delta^*) \ge 1$} {
-    	Draw $m$ more samples to augment the current minibatch, and update $\Delta^*$ and $\sigma^2(\Delta^*)$ \;
+    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $\sigma^2(\Delta^*)$\;
+    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from equation (\ref{eq:clt-bounds3})\;
+    \While {$\sigma^2(\Delta^*) \ge 1$ {\bf or} $\epsilon > \delta$}{
+    	Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $\sigma^2(\Delta^*)$ and $\epsilon$ estimates\;
     }
-    Let $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$, draw a normal variable $X_n \in \mathcal{N}(0,\sigma_n^2)$ and
-    $X_c$ from the correction distribution\;
-
-        \eIf{$\Delta^* + X_n + X_c > 0$}{
+    Draw $X_{\rm norm} \in \mathcal{N}(0,1-\sigma^2(\Delta*))$ and
+    $X_{\rm corr}$ from the correction distribution $C_1(X)$\;
+
+        \eIf{$\Delta^* + X_{\rm norm} + X_{\rm corr} > 0$}{
             Accept the candidate, $\theta_{t+1} = \theta'$\;
@@
 these errors bounds are preserved after the addition of other random variables, in particular
-$X_n$ and $X_c$. From this it follows that the acceptance test has the same error bound.
+$X_{\rm norm}$ and $X_{\rm corr}$. From this it follows that the acceptance test has the same error bound.
 
@@
 
-To apply this to our test, let $\mu = \frac{1}{N}\sum_{i=1}^N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$ which is
-the expected log data probability ratio across the entire dataset. Then
-Let $X_i = N(\log\frac{p(x_i|\theta')}{p(x_i|\theta)}-\mu)$, are zero-mean, IID variables.
+To apply this to our test, let $X_i = \Lambda_i(\theta,\theta') - \Lambda(\theta,\theta')$,
+then the $X_i$ are zero-mean, IID variables.
 Take $n=b$ the minibatch size, and then:
 \begin{equation}
-  \Delta^*(x,\theta,\theta') = \hat{X} + \Delta(x,\theta,\theta')
+\hat{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta') 
 \end{equation}
@@
 quantities. Both $E|X|$ and $E|X|^3$ can be estimated as means of
-$|X_i - \hat{X}|$ and $|X_i - \hat{X}|^3$ respectively on each minibatch. We expect this
+$|\Lambda_i - \Lambda|$ and $|\Lambda_i - \Lambda|^3$ respectively on each minibatch. We expect this
 will often be accurate enough on minibatches with hundreds or
@@
 \begin{corollary}
-  If $|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})| \leq \epsilon(\theta,\theta',b)$,
+  If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})| \leq \epsilon(\theta,\theta',b)$,
   then
   \begin{equation}
-    |{\rm Pr}(\Delta^*+X_n+X_c < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
-  \end{equation}
-  where $S(x)$ is the standard logistic function, and $X_n$ and $X_c$ are generated as per Algorithm 1. 
+    \sup_y|{\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
+  \end{equation}
+  where $S(x)$ is the standard logistic function, and $X_{\rm norm}$ and $X_{\rm corr}$ are generated as per Algorithm 1. 
   \end{corollary}
@@
   \begin{equation}
-    P'(y) = Pr(\Delta^*+X_n < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
-  \end{equation}
-  Now we convolve with the distribution of $X_c$ which gives:
-  \begin{equation}
-    P''(y) = Pr(\Delta^*+X_n+X_c < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
-  \end{equation}
-  Both steps preserve the error bound $\epsilon(\theta,\theta',b)$
+    P'(y) = Pr(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
+  \end{equation}
+  Now we convolve with the distribution of $X_{\rm corr}$ which gives:
+  \begin{equation}
+    P''(y) = Pr(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
+  \end{equation}
+  Both steps preserve the error bound
+  $\epsilon(\theta,\theta',b)$. Finally $S(y-\Delta)$ is a logistic
+  CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm Pr}(\Delta +
+  X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We
+  conclude that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm
+    corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
+    log} > 0)$ by at most $\epsilon$.
+
 \end{proof}

### 2016-07-18 nips_2016.tex
mnist sim

@@
 normal distribution with mean $\mu$ and variance $\sigma^2$.  An efficient
-proposal distribution has the same variance as the target distribution
-$\mathcal{N}(\theta'-\theta,1/N)$, so we take $q(\theta'\mid\theta) =
+proposal distribution has the same variance as the target distribution in one dimension
+\cite{OptimalScaling01}, so take the proposal
+$\mathcal{N}(\theta'-\theta,1/N)$, which is implemented as $q(\theta'\mid\theta) =
 \phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal density

### 2016-07-19 nips_2016.tex
more edits

@@
 Metropolis-Hastings tests on the full data. Previous work on reducing
-the cost of Metropolis-Hastings tests yield variable data reductions
-and in the worst case require the entire dataset.  Here we present a
-simple test which requires constant size minibatches if an auxiliary
-condition is met. We describe several cases under which this condition
-can be satisfied. Our new test is naturally tolerant of ``noisy''
-estimates of the transition probability ratio. To achieve this, we
-replace the classical Metropolis-Hastings acceptance test with a
-smooth acceptance function which is a sigmoid (logistic)
-function of the log likelihood ratio.
-The resulting test can be combined with minibatch proposals to yield
+the cost of Metropolis-Hastings tests yield variable data consumed per
+sample, with only constant factor reductions vs. using the full
+dataset for each sample.  Here we present a method that can be tuned
+to provide arbitrarily small batch sizes, by adjusting either proposal
+step size or temperature. Our approach uses the natural noise present
+in minibatch likelihood esimates to furnish the randomness in a
+Metropolis-Hastings test. Our test uses the noise-tolerant Barker
+acceptance test with a novel additive correction variable.  The
+resulting test can be combined with minibatch proposals to yield
 updates with the same complexity as a simple SGD update. In this paper
-we derive the test, discuss its implementation, and present several
-experiments.
+we derive the test, analyze its performance, discuss its
+implementation, and present several experiments.
 \end{abstract}
@@
 this, the amount of data required for each test is fixed with high
-probability and in expected value. Our test requires a bound on the
-(relative) variance of the log acceptance probability. This variance
-is small when the proposal distribution is efficient, that is when samples
-diffuse at an optimal rate.
-
-The variance of the log acceptance probability can be controlled in
-three ways, and these allow the test to be applied to modest-sized
-minibatches. The first two allow the test to be applied when the
-minibatch size is fixed a-priori.
+probability and in expected value. We use a Barker
+test function \cite{Barker65} rather than a Metropolis test, which
+makes the test naturally error tolerant. Similar ideas were presented
+in \cite{Bardenet15} section 6.3. But the asymptotic test statistic
+CDF and the Barker function are different, which leads to fixed
+errors for the approach in \cite{Bardenet15}. Here we show that the
+difference between the distributions can be corrected with an
+additive random variable. This leads to a test which is fast, and
+whose error can be made arbitrarily small.
+
+Our test is applicable when the variance (over data samples) of the
+log acceptance probability is small enough (less than 1). Its not
+clear at first why this quantity should be so bounded. But we will see
+that it is ``natural'' for well-specified models running
+Metroplis-Hastings sampling with optimal proposals
+\cite{OptimalScaling01} on a full dataset. When we reduce the amount
+of data for the test, the variance goes up, and we have to reduce
+it in one of several ways. Either:
 
@@
   scale as $1/T$, and so the variance of the likelihood ratio will
-  vary as $1/T^2$.
+  vary as $1/T^2$. Our model is no longer well-specified (we are doing inference
+  at a temperature different from that assumed during data generation), but
+  higher temperature can be advantageous for parameter exploration.
 
 \item For continuous probability distributions, reduce the proposal
-  step size and variance (for stochastic proposals).
-
-\item Increase the minibatch size. Relative variance scales as $1/k$ vs the
-  minibatch size $k$. Increased minibatch size also reduces the error of
+  step size and variance (for stochastic proposals) compared to an optimal
+  proposal. The variance of the log acceptance probability scales as the
+  square of proposal step size.
+
+\item Increase the minibatch size. log acceptance variance scales as $1/k$ vs the
+  minibatch size $k$. Increased minibatch size also reduces the error rate for
   the test.
@@
 
+Its worth discussing at this point what are the goals (typically) of
+M-H sampling on very large datasets.  By the Bernstein-von Mises
+theorem, the posterior distribution of the parameter $\theta$ for a
+Bayesian inference task is asymptotically normal, and has variance
+that scales inversely with the number of data samples $N$. This mode
+is extremely sharp for large datasets, which may contain millions or
+billions of samples. Simply sampling from this distribution is one
+application, but an efficient proposal distribution
+\cite{OptimalScaling01} has similar variance to the target
+distribution and will diffuse to it extremely slowly from an random
+initialization value. If there are any other strong modes, it is very
+likely for the sampler to find one of them and become trapped in it
+when run at the normal distribution temperature (T=1). A common
+solution is to anneal the sampler, running first at high temperature
+(scaling log likelihoods by 1/T) which flattens the likelihood
+landscape.  This is turn reduces the variance of the log acceptance
+probability and allows our acceptance test to be applied.
+
+A second question concerns step size. Once we have fixed temperature,
+our variance constraint implies that we have to trade-off proposal
+step size $s$ and batch size $b$ ($b \propto p^2$). i.e. we can make
+many small steps, or one large step, with a given batch of data. One
+of the primary drivers of this work is our belief in the value of
+small steps. For applications to neural networks or other models where
+the posterior is multimodel, posterior inference is arguably a search
+process. Covering the search space densely with small steps is much
+more valuable than few sparse steps toward the nearest optimum.
+
 The contributions of this paper are as follows:
@@
 \begin{itemize}
-\item We develop a new, more efficient (in samples per test) minibatch acceptance test which incorporates
-online  estimates of its error probability.
+\item We develop a new, more efficient (in samples per test) minibatch acceptance test with quantifiable error bounds. The test
+  uses a novel additive correction variable to implement a Barker test based on minibatch mean and variance. 
 \item We analyze the test for accuracy and speed.
@@
   of magnitude more efficient than prior work measured as data consumed per test, and that it does not suffer
-  from long-tailed data requirements (up to the dataset size). 
+  from long-tailed batch sizes (up to the dataset size). 
 \end{itemize}
@@
   norm}) = \Phi(X_{\rm norm}/\sigma)$ and $S(X_{\rm log})$, where
-$\Phi(.)$ is the Normal CDF with zero mean and variance $\sigma^2$ and
+$\Phi(.)$ is the Normal CDF with zero mean and unit variance and
 $S(.)$ is the standard logistic function.
@@
 \end{algorithm}
+
+\section{Computing the Correction Distribution}\label{sec:correction}
+Our goal is to compute the distribution of the correction variable $X_{\rm corr}$ such that
+$X_{\rm norm} + X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \in \mathcal{N}(0,\sigma)$
+and $X_{\rm log}$ has a standard logistic CDF $(1+\exp(-X))^{-1}$.
+In the next section we show that the accuracy of the test depends on the absolute
+error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and the distribution of $X_{\rm log}$.
+So we minimize this in our construction.
+
+Our goal then is to minimize
+\begin{equation}
+  |\Phi_{\sigma} * C_{\sigma} - S|
+\end{equation}
+where $\Phi_{\sigma}(X) = \Phi(X/\sigma)$ and $\Phi$ is the standard
+normal CDF, $S(X)$ is the logistic function, $C_{\sigma}(X)$ is the
+{\em density} of the correction distribution, and $*$ denotes
+convolution. In continuous form, this value is
+\begin{equation}
+  \label{eq:conv1}
+  \sup_X\left|\int_{-V}^{V}\Phi_{\sigma}(X-Y) C_{\sigma}(Y) dY - S(X)\right|
+  \end{equation}
+Assuming $X_{\rm corr}$ lies in the interval $[-V,V]$. For computation of $C_{\sigma}(Y)$, we
+discretize $X$ and $Y$ into $4N+1$ and $2N+1$ values respectively, with $X \in [-2V,2V]$ and
+$Y \in [-V,V]$. Writing $X_i = i * V / N$ and $Y_j = j * V / N$, the condition in
+equation (\ref{eq:conv1}) can be written:
+\begin{equation}
+  \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|
+\end{equation}
+and defining a matrix $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, vectors $U_j = C_{\sigma}(Y_j)$ and $V_i = S(X_i)$, the
+above is equivalent to minimizing the 1-norm:
+\begin{equation}
+  \label{eq:error_norm}
+  ||MU-V||_1
+\end{equation}
+wrt $U$. We have an additional constraint that $U_j > 0$ for all $j$, since $U$ represents
+a density. We first explored optimizing this system with linear programming to find the $U$ such that:
+\begin{equation}
+  \begin{split}
+    & \min(\epsilon) \qquad \hbox{where} \\
+    & -\epsilon \leq MU - V \leq \epsilon \\
+    & U \geq 0
+  \end{split}
+\end{equation}
+which was tractable up to a few hundred dimensions. However, the discretization error is bounded by
+the curvature of the underlying functions which are here slightly less than one. i.e.
+the errors are of the order of $(V/N)^2$. Here we chose $V=20$ to provide adequate containment of
+the distributions (the CDFs are extremely close to either zero or one outside this range). So
+with 200 points, we have a discretization error of approximately 0.01. This is quite high. To yield
+higher resolution and lower error, we switched to a least squares solution. 
+Thus we seek to find $U$ that minimizes
+\begin{equation}
+  ||MU-V||_2
+\end{equation}
+Its not clear that this will yield a good 1-norm for equation (\ref{eq:error_norm}), but in practice it does.
+Because the solution of this system is not very stable, especially in high dimensions, we actually
+minimize a regularized version
+\begin{equation}
+  ||MU-V||_2 + \lambda||U||_2
+\end{equation}
+the solution is
+\begin{equation}
+  U = (M^T M + \lambda I)^{-1}V
+\end{equation}
+with this approach, there is no guarantee that $U \geq 0$. However, we have some
+flexibility in the choice of $\sigma$ in equation (\ref{eq:conv1}). As we decrease
+the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount
+and is in fact the result of convolution with a gaussian whose variance is the difference.
+Thus as $\sigma$ decreases, $C(X)$ grows and approaches the derivative of a
+logistic function at $\sigma = 0$. It retains some very weak negative values for $\sigma > 0$
+but removal of those values leads to very small error.
+\begin{table}[h]
+  \centering
+  \begin{tabular}{|l|l|l|l|}
+    \hline
+     N & $\sigma$ & $\lambda$ & L1 error  \\
+        \hline
+     4000 & 0.9 & 1 & 1.0e-4  \\
+     4000 & 0.8 & 0.03 & 5.0e-6 \\
+     \hline
+    \end{tabular}
+\end{table}
 

### 2016-07-20 nips_2016.tex
tweaks to norm2log table
added O(1/n) error bound

@@
 this, the amount of data required for each test is fixed with high
-probability and in expected value. We use a Barker
-test function \cite{Barker65} rather than a Metropolis test, which
-makes the test naturally error tolerant. Similar ideas were presented
-in \cite{Bardenet15} section 6.3. But the asymptotic test statistic
-CDF and the Barker function are different, which leads to fixed
-errors for the approach in \cite{Bardenet15}. Here we show that the
-difference between the distributions can be corrected with an
+probability and in expected value. We use a Barker test function
+\cite{Barker65} rather than a Metropolis test, which makes the test
+naturally error tolerant. The idea of using a noise-tolerant test
+using Barker's test function were suggested but not explored
+empirically in \cite{Bardenet15} section 6.3. But the asymptotic test
+statistic CDF and the Barker function are different, which leads to
+fixed errors for the approach in \cite{Bardenet15}. Here we show that
+the difference between the distributions can be corrected with an
 additive random variable. This leads to a test which is fast, and
@@
 \cite{OptimalScaling01} has similar variance to the target
-distribution and will diffuse to it extremely slowly from an random
-initialization value. If there are any other strong modes, it is very
-likely for the sampler to find one of them and become trapped in it
-when run at the normal distribution temperature (T=1). A common
-solution is to anneal the sampler, running first at high temperature
-(scaling log likelihoods by 1/T) which flattens the likelihood
-landscape.  This is turn reduces the variance of the log acceptance
-probability and allows our acceptance test to be applied.
+distribution and will diffuse to it extremely slowly from an
+initialization value which is (likely to be) many standard deviations
+away. If there are any other strong modes, it is very likely for the
+sampler to find one of them and become trapped in it when run at the
+normal distribution temperature (T=1). A common solution is to anneal
+the sampler, running first at high temperature (scaling log
+likelihoods by 1/T) which flattens the likelihood landscape.  This is
+turn reduces the variance of the log acceptance probability and allows
+our acceptance test to be applied.
 
@@
 process. Covering the search space densely with small steps is much
-more valuable than few sparse steps toward the nearest optimum.
+more valuable than few sparse steps toward the nearest optimum. In
+this mode, Metropolis-Hastings can be used in similar fashion to
+Stochastic Gradient Descent. The goal in SGD is to make gradual
+progress to a posterior mode with each step, taking small steps so
+that the cumulative displacement has progressively lower variance. A
+substantial part of the computational work of MCMC on massive datasets
+will be similarly in reaching a stationary distribution, which really
+means finding a deep posterior mode. Taking noisy small steps will
+nevertheless make steady progress to a posterior mode since their bias
+is in that direction. We demonstrate this behavior with our experiments.
 
@@
 \end{lemma}
-
-\begin{proof} Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and
-  $\theta -0.5 \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events
-  occur with probability $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and $p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively,
-  and guarantee that $\Lambda^*(\theta',\theta)$ is negative, which guarantees that we can find
-  a $u$ so that $\psi(u,\theta',\theta)$ equals the expected value of $\Lambda^*(\theta',\theta)$.
-  Specifically, choose $u_0$
-  to satisfy
-  \begin{equation}
-    \log u_0 = N(\Lambda^*(\theta',\theta)-\psi(1,\theta',\theta))
-  \end{equation}
-  which evaluates to 
-  \begin{equation}
-    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right) 
-  \end{equation}
-  and rewrite the test now as
-  \begin{equation}
-    \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right)
-    \hspace{0.1in} \not\approx \hspace{0.1in} 0
-  \end{equation}
-  where $\not\approx$ means ``significantly different from'' under the distribution over samples of $x_i$. 
-  The above choice of $u_0$ ensures that the terms in parenthesis above sum to 0.5. Since the $x_i$ have mean 0.5, the
-  test will never correctly succeed.
-
-  Furthermore, if we sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be sufficiently different
-  from 0.5 to allow the test to succeed.
-
- The choices above for $\theta$ and $\theta'$ guarantee that
-  \begin{equation}
-    \log u_0 \in -[0.5,1][0.75,1.5] = -[0.375,1.5]
-        \end{equation}
-and consider the range of $u$ values
-  \begin{equation}
-    \log u \in \log u_0 + [-0.5,0.375]
-  \end{equation}
-  the size of the range in $u$ is at least $\exp([-2,-1.125]) = [0.13534,0.32465]$ and occurs
-  with probability at least $p_2=0.18932$. With $u$ in this range, we rewrite the test as:
-  \begin{equation}
-    \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
-  \end{equation}
-  so that the LHS has expected value zero.  Given our choice of intervals for the variables, the RHS is in the range $1/\sqrt{N}[-1,0.75]$.
-  The standard deviation of the LHS given the interval constraints is at least $0.5/\sqrt{b}$. And
-  so the gap between LHS and RHS is at most $2\sqrt{b/N}$ standard deviations.
-
-  The samples $\theta$, $(\theta'-\theta)$
-  and $u$ are drawn independently and so the probability of the conjunction of these events is
-  $c = p_0 p_1 p_2 = 0.0085$.
-\end{proof}
+Proof is given in the appendix.
 
@@
 
-\begin{proof}
-  Start with the LHS of (\ref{detbal}), substitute for $\alpha(\theta,\theta')$, and then apply the
-  identity $g(s) = \exp(s)g(-s)$:
-  \begin{equation}\label{detbal1}
-    \begin{split}
-      p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') =~ & p(\theta)q(\theta'\mid\theta)g(\Delta(\theta,\theta'))\\
-      =~ & p(\theta)q(\theta'\mid\theta)\exp(\Delta(\theta,\theta'))g(-\Delta(\theta,\theta'))
-    \end{split}
-  \end{equation}
-  Note then that $-\Delta(\theta,\theta') = \Delta(\theta',\theta)$ because $\Delta$ is the log of ratios of
-  terms involving $\theta$ and $\theta'$ and exchanging the two inverts all these ratios, negating the
-  logarithm. Continuing:
-  \begin{equation}\label{detbal2}
-    \begin{split}
-    p(\theta)q(\theta'\mid\theta)\exp(\Delta(\theta,\theta'))g(-\Delta(\theta,\theta')) = ~ & p(\theta)q(\theta'\mid\theta)\exp(\Delta(\theta,\theta'))g(\Delta(\theta',\theta))\\
-    = ~ & p(\theta)q(\theta'\mid\theta)\exp(\Delta(\theta,\theta'))\alpha(\theta',\theta)
-    \end{split}
-  \end{equation}
-  Substituting now for $\Delta(\theta,\theta')$ from equation (\ref{eq:delta0}) gives
-  \begin{equation*}\label{detbal3}
-    p(\theta)q(\theta'\mid\theta)\frac{p(\theta')q(\theta\mid\theta')}{p(\theta)q(\theta'\mid\theta)}\alpha(\theta',\theta)
-      = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta) = RHS
-  \end{equation*}
-\end{proof}
-
+This result is used in \cite{Barker65} to define the Barker acceptance test. 
 As a sanity check, choosing $g(s) = \exp(s) \wedge 1$ produces the
@@
 
-\subsection{Logistic Acceptance Function}
-For our new MH test we use the logistic function:
-$g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it
+\subsection{Barker (Logistic) Acceptance Function}
+For our new MH test we use the Barker logistic function:
+$g(s)=(1+\exp(-s))^{-1}$ \cite{Barker65}. Straightforward arithmetic shows that it
 satisfies the condition in Lemma~\ref{lem:detailed_balance}.  While it
@@
 But to perform the test in equation (\ref{theexacttest}) we want $\Delta + X$ for
-a logistic random $X$ (call it $X_{\rm log}$ from now on). 
-Our approach is to decompose $X_{\rm log}$ as
+a logistic random $X$ (call it $X_{\rm log}$ from now on). In \cite{Bardenet15} it was proposed to
+use $\Delta^*$ in a Barker test anyway and tolerate the fixed error caused by this approximation.
+
+Our approach is to instead decompose $X_{\rm log}$ as
 \begin{equation}\label{eq:deconvolution}
@@
 variance $\sigma^2$, and $X_{\rm corr}$ is a zero-mean ``correction''
-term with density $C_{\sigma}(X)$.  The two variables are added
+variable with density $C_{\sigma}(X)$.  The two variables are added
 (i.e., their distributions convolve) to form $X_{\rm log}$. It will be
@@
 
+So as long as we know the first and third absolute moments $E|X|$ and $E|X|^3$, we can bound the
+error of the normal approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of variables
+$y = x \sigma(\hat{X})$, equation (\ref{eq:clt-bounds}) becomes
+\begin{equation}\label{eq:clt-bounds2}
+   \sup_y\left|{\rm Pr}(\hat{X}< y) - \Phi\left(\frac{y}{\sigma(\hat{X})}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
+\end{equation}
+which shows that the distribution of $\hat{X}$ approaches the normal distribution $\mathcal{N}(0,\sigma^2(\hat{X}))$
+whose variance is $\sigma^2(\hat{X})$, measured from the sample.
+
+In fact we can do better than this approximation by using a more
+precise limit distribution, under an additional assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$
+denote the $i^{th}$ absolute moment of $X$. Bentkus et al.'s work on Edgeworth
+expansions \cite{Bentkus97} equation 2.2 gives:
+
+\begin{lemma} Let $X_1,\ldots,X_n$ be
+a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3. Then
+\begin{equation}\label{eq:clt-bounds}
+ \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
+\end{equation}
+where
+\begin{equation}
+G_N(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x)
+\end{equation}
+\end{lemma}
+Here the limit $G_N(x,y)$ is a normal CDF plus a correction term that
+decays as $n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where $\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
+\begin{equation}
+G_N(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
+\end{equation}
+From which we see that $G_N(x,y)$ is a linear combination of $\Phi(x)$, $\phi(x)$ and $\phi^{''}(x)$.
+
 \begin{lemma} Let $P(x)$ and $Q(x)$ be two cumulative distributions satisfying $\sup_x|P(x)-Q(x)|\leq \epsilon$ with
@@
 can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
-in lots of wasted computation.
-For this particular experiment, that happened eight times.
-The mean of adaptive minibatch size is 520 which is roughly close to $\sqrt{N}$ (i.e.
-100).
-This number validates the claims of Theorem~\ref{thm:efficiency}.
-In
-contrast, our method keeps the minibatch size around 200. Figure~\ref{fig:gauss_mix_2} (in the middle) shows a histogram of the minibatch size, and the mean of our minibatch size is 247. 
+in lots of wasted computation.  For this particular experiment, that happened eight times.
 The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
@@
 
-Figure~\ref{fig:gauss_mix_3} (to the left) shows a histogram of ${\rm std}(\Delta')$.
-Figure~\ref{fig:gauss_mix_3} (to the middle) shows the confidence interval of our estimated ${\rm std}(\Delta')$ at each iteration \footnote{In order to show the confidence interval clearly, we use moving smooth to smooth the confidence interval, and plot each value by interval of 50 iterations}.
-The blue line represents the standard deviation comes from the complete data, which can be treated as the ``true'' ${\rm std}(\Delta')$, the green line represents our estimation of ${\rm std}(\Delta')$ from the minibatch, and the two red dash lines represent the confidence interval of ${\rm std}(\Delta')$ by adding / subtracting one ${\rm std(std(}\Delta'))$.
-This figure indicates that our estimation of ${\rm std}(\Delta')$ from one minibatch data is accurate.
-Figure~\ref{fig:gauss_mix_3} (to the right) shows the value of $\rho/\sigma^3$ at each iteration.
-Recall Theorem~\ref{thm:efficiency}, the number of data instance required is $\frac{4 \ell^2 C^2 \rho^2 }{\sigma^6 \varepsilon^2}$.
-This figure reveals that $\rho/\sigma^3$ is roughly a constant, which means the data required for each iteration is a constant value which has nothing to do with the $N$.
+Figure~\ref{fig:gauss_mix_3} (to the left) shows a histogram of ${\rm
+  std}(\Delta')$. Figure~\ref{fig:gauss_mix_3} (to the middle) shows
+the confidence interval of our estimated ${\rm std}(\Delta')$ at each
+iteration \footnote{In order to show the confidence interval clearly,
+  we use moving smooth to smooth the confidence interval, and plot
+  each value by interval of 50 iterations}. The blue line represents
+the standard deviation comes from the complete data, which can be
+treated as the ``true'' ${\rm std}(\Delta')$, the green line
+represents our estimation of ${\rm std}(\Delta')$ from the minibatch,
+and the two red dash lines represent the confidence interval of ${\rm
+  std}(\Delta')$ by adding / subtracting one ${\rm
+  std(std(}\Delta'))$. This figure indicates that our estimation of
+${\rm std}(\Delta')$ from one minibatch data is
+accurate. Figure~\ref{fig:gauss_mix_3} (to the right) shows the value
+of $\rho/\sigma^3$ at each iteration.
 
@@
 
-In this appendix, we describe the following topics, with an emphasis on clarity and understanding:
+In this appendix, we describe the following topics:
 
 \begin{itemize}[noitemsep]
+  \item Proofs ommitted from the main text. 
     \item More information on Section~\ref{ssec:gaussians}.
@@
 
-%% \section{Proofs}\label{app:proofs}
+\section{Proofs}\label{app:proofs}
+
+{\bf lemma 1}
+{\em  For the model of section 2.1, there
+  exists a fixed (independent of $N$) constant $c$ such that with probability $ \geq c$ over the joint
+  distribution of $(\theta, \theta', u)$, the
+  tests from \cite{icml2014c1_bardenet14} and \cite{cutting_mh_2014} consume all $N$ samples. }
+
+\begin{proof} Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and
+  $\theta -0.5 \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events
+  occur with probability $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and $p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively,
+  and guarantee that $\Lambda^*(\theta',\theta)$ is negative, which guarantees that we can find
+  a $u$ so that $\psi(u,\theta',\theta)$ equals the expected value of $\Lambda^*(\theta',\theta)$.
+  Specifically, choose $u_0$
+  to satisfy
+  \begin{equation}
+    \log u_0 = N(\Lambda^*(\theta',\theta)-\psi(1,\theta',\theta))
+  \end{equation}
+  which evaluates to 
+  \begin{equation}
+    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right) 
+  \end{equation}
+  and rewrite the test now as
+  \begin{equation}
+    \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right)
+    \hspace{0.1in} \not\approx \hspace{0.1in} 0
+  \end{equation}
+  where $\not\approx$ means ``significantly different from'' under the distribution over samples of $x_i$. 
+  The above choice of $u_0$ ensures that the terms in parenthesis above sum to 0.5. Since the $x_i$ have mean 0.5, the
+  test will never correctly succeed.
+
+  Furthermore, if we sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be sufficiently different
+  from 0.5 to allow the test to succeed.
+
+ The choices above for $\theta$ and $\theta'$ guarantee that
+  \begin{equation}
+    \log u_0 \in -[0.5,1][0.75,1.5] = -[0.375,1.5]
+        \end{equation}
+and consider the range of $u$ values
+  \begin{equation}
+    \log u \in \log u_0 + [-0.5,0.375]
+  \end{equation}
+  the size of the range in $u$ is at least $\exp([-2,-1.125]) = [0.13534,0.32465]$ and occurs
+  with probability at least $p_2=0.18932$. With $u$ in this range, we rewrite the test as:
+  \begin{equation}
+    \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
+  \end{equation}
+  so that the LHS has expected value zero.  Given our choice of intervals for the variables, the RHS is in the range $1/\sqrt{N}[-1,0.75]$.
+  The standard deviation of the LHS given the interval constraints is at least $0.5/\sqrt{b}$. And
+  so the gap between LHS and RHS is at most $2\sqrt{b/N}$ standard deviations.
+
+  The samples $\theta$, $(\theta'-\theta)$
+  and $u$ are drawn independently and so the probability of the conjunction of these events is
+  $c = p_0 p_1 p_2 = 0.0085$.
+\end{proof}
 

### 2016-07-20 rebuttal_reviewers.txt

First, we thank the reviewers for their thoughtful commentary on our paper to keep the quality of
NIPS top-notch.

General comments:

We have been working continuously on the paper in the last month, and have been able to address a
number of our reviewer comments. For instance, we have been able to demonstrate worst-case scenarios
for references [3] and [4] with Omega(N) lower bounds for fixed proposals using Gaussians. We have
also added rigorous, numerical error bounds based on t-statistics (from the empirical mean and
variance) asymptotics from [13]. We have added bounds on the equilibrium distribution given bounds
on transition kernel error (we will borrow this from [4]) with the appropriate assumptions.

To reviewers 1, 2, 3, and 4 regarding the Barker function:

You are correct, the logistic function (aka Barker function) is less efficient with \theta' as good
as \theta. However, the tradeoff is that using the Barker function lets us better control our error
rates, and our newest experiments show that we are obtaining orders of magnitude improvements.

To reviewers 1, 2, and 3 regarding the convolution (in)stability:

To make it more stable, we can use L2 regularization and move some of the variance from X_norm to
X_corr. i.e. widen X_corr by convolving with a Gaussian with variance sigma^2, while decreasing the
allowed variance of X_norm by sigma^2. This ensures that X_corr has positive pdf everywhere.  We can
achieve CDF error (X_norm + X_corr compared to X_log) of 1e-7 at sigma = 0.8 or 1e-5 at sigma = 0.9.
We are substantially reworking this part of the paper.

To reviewer 1:

TODO (Daniel: Anything else you want to address)

To reviewer 2:

Regarding the difference between ours and [BDH] sec 6.3, that is close to what we're doing, but they
don't attempt to correct for the difference between normal and logistic distributions with an
additive variable, so they are stuck with a fixed error. That allows us to run our algorithm with
arbitrarily small errors (depending on batch size of course).

TODO (Daniel: Anything else you want to address)

To reviewer 3:

Regarding your Lemma 2 comment, yes, you are correct, the CLT only applies for large n and N, and we
also need n << N. We will clarify this.

To reviewer 7:

Yes increasing the minibatch size is one of the strategies we can use to control variance. The idea
is to increase it just enough, and then to keep it fixed thereafter. We should emphasize this
distinction more, and also to emphasize that this is the least desirable tactic to use (and our
newest results show that we do not need to use it anyway).

### 2016-07-21 rebuttal_reviewers.txt
Finished most of rebuttal_reviewers.txt (for reviewers)

@@
-First, we thank the reviewers for their thoughtful commentary on our paper to keep the quality of
-NIPS top-notch.
+We thank the reviewers for their thoughtful commentary on our paper to keep the quality of NIPS
+top-notch. We start the rebuttal with a few general comments and then respond to specific reviewers.
 
@@
 on transition kernel error (we will borrow this from [4]) with the appropriate assumptions. 
+(This section addresses theoretical results)
 
-To reviewers 1, 2, 3, and 4 regarding the Barker function:
+To Reviewers 2, 3, and 4 regarding the Barker function:
 
@@
 
-To reviewers 1, 2, and 3 regarding the convolution (in)stability:
+To Reviewers 1, 2, and 3 regarding the convolution (in)stability:
 
@@
 
-To reviewer 1:
+To Reviewer 2
 
-TODO (Daniel: Anything else you want to address)
+(TODO?)
 
-To reviewer 2:
+To Reviewers 1 and 3 regarding Lemma 2:
 
-Regarding the difference between ours and [BDH] sec 6.3, that is close to what we're doing, but they
-don't attempt to correct for the difference between normal and logistic distributions with an
-additive variable, so they are stuck with a fixed error. That allows us to run our algorithm with
-arbitrarily small errors (depending on batch size of course).
+Both of you have correctly identified our implicit assumptions about Lemma 2. Reviewer_1 is right
+that we approximated it for large N, so sampling without replacement approximates sampling with
+replacement.  Reviewer_3 is correct in that the CLT only applies for large n and N, and we also need
+n << N. Note also that we generally assume n is in the several hundreds, which is often enough for
+the CLT to hold. We will clarify this in our revisions.
 
-TODO (Daniel: Anything else you want to address)
+To Reviewers 1 and 2:
 
-To reviewer 3:
+Thank you for mentioning [BDH], "MCMC Methods for Tall Data." We added this reference as part of our
+ongoing revision of the paper and will substantially address issues raised in [BDH]. We are also
+addressing Reviewer 2's other references.  In addition, we are addressing Reviewer 2's concern
+regarding the difference between ours and [BDH] Section 6.3. Yes, it's close to what we are doing,
+but they don't attempt to correct for the difference between normal and logistic distributions with
+an additive variable, so they are stuck with a fixed error.
 
-Regarding your Lemma 2 comment, yes, you are correct, the CLT only applies for large n and N, and we
-also need n << N. We will clarify this.
+To Reviewer 7:
 
-To reviewer 7:
-
-Yes increasing the minibatch size is one of the strategies we can use to control variance. The idea
-is to increase it just enough, and then to keep it fixed thereafter. We should emphasize this
-distinction more, and also to emphasize that this is the least desirable tactic to use (and our
-newest results show that we do not need to use it anyway).
+We should have clarified, that we have the option of increasing minibatch sizes, and the idea is
+that we can start by increasing it just enough for the variance condition to apply. Once we have
+identified a suitable minibatch size, we keep it fixed thereafter. In our latest revisions, we rerun
+our test cases and experiment with reducing the proposal variance or increasing the temperature. By
+doing so, we can run our algorithm using arbitrarily small minibatch sizes.

### 2016-07-21 rebuttal_reviewers.txt
Fixed rebuttals, almost done

@@
 
-General comments:
+General:
 
 We have been working continuously on the paper in the last month, and have been able to address a
-number of our reviewer comments. For instance, we have been able to demonstrate worst-case scenarios
-for references [3] and [4] with Omega(N) lower bounds for fixed proposals using Gaussians. We have
-also added rigorous, numerical error bounds based on t-statistics (from the empirical mean and
-variance) asymptotics from [13]. We have added bounds on the equilibrium distribution given bounds
-on transition kernel error (we will borrow this from [4]) with the appropriate assumptions. 
-(This section addresses theoretical results)
+number of our reviewer comments. First, we have demonstrated worst-case scenarios for references [3]
+and [4] with Omega(N) lower bounds on the number of samples consumed per iteration, using simple
+Gaussian examples.
+
+In addition, Section 4 (the theoretical results) has been vastly revamped.  We have added rigorous,
+numerical error bounds based on t-statistics (from the empirical mean and variance) asymptotics from
+[13]. We have added bounds on the equilibrium distribution given bounds on transition kernel error
+(we will borrow this from [4]) with the appropriate assumptions.
 
@@
 
-You are correct, the logistic function (aka Barker function) is less efficient with \theta' as good
-as \theta. However, the tradeoff is that using the Barker function lets us better control our error
-rates, and our newest experiments show that we are obtaining orders of magnitude improvements.
+You are correct, the logistic (Barker) function is less efficient with \theta' as good as \theta,
+and we explicitly mention this in our revision. The tradeoff is that its smoothness allows it to
+naturally tolerate substantial variance in its input argument. This in turn will lead to a much more
+efficient test on subsets of data, and our newest experiments show we have more control over error
+rates.
 
@@
 
-To make it more stable, we can use L2 regularization and move some of the variance from X_norm to
-X_corr. i.e. widen X_corr by convolving with a Gaussian with variance sigma^2, while decreasing the
-allowed variance of X_norm by sigma^2. This ensures that X_corr has positive pdf everywhere.  We can
-achieve CDF error (X_norm + X_corr compared to X_log) of 1e-7 at sigma = 0.8 or 1e-5 at sigma = 0.9.
-We are substantially reworking this part of the paper.
+To make the convolution more stable, we frame it as a least squares (with L2 regularization)
+minimization problem. We move some of the variance from X_norm to X_corr, i.e. widen X_corr by
+convolving with a Gaussian with variance sigma^2, while decreasing the allowed variance of X_norm by
+sigma^2. This ensures that X_corr has positive pdf everywhere.  We can achieve CDF error (X_norm +
+X_corr compared to X_log) of 1e-7 at sigma = 0.8 or 1e-5 at sigma = 0.9.  We have added an entire
+new section explaining this procedure.
 
-To Reviewer 2
+To Reviewer 3:
 
-(TODO?)
+(TODO respond to experiment criticism, but I really can't defend the work as it is ...)
 

### 2016-08-04 nips_2016.tex
latest edits

@@
 
-In fact we can do better than this approximation by using a more
-precise limit distribution, under an additional assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$
-denote the $i^{th}$ absolute moment of $X$. Bentkus et al.'s work on Edgeworth
-expansions \cite{Bentkus97} equation 2.2 gives:
-
-\begin{lemma} Let $X_1,\ldots,X_n$ be
-a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3. Then
-\begin{equation}\label{eq:clt-bounds}
- \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
-\end{equation}
-where
-\begin{equation}
-G_N(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x)
-\end{equation}
-\end{lemma}
-Here the limit $G_N(x,y)$ is a normal CDF plus a correction term that
-decays as $n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where $\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
-\begin{equation}
-G_N(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
-\end{equation}
-From which we see that $G_N(x,y)$ is a linear combination of $\Phi(x)$, $\phi(x)$ and $\phi^{''}(x)$.
-
 \begin{lemma} Let $P(x)$ and $Q(x)$ be two cumulative distributions satisfying $\sup_x|P(x)-Q(x)|\leq \epsilon$ with
@@
 \end{proof}
+
+In fact we can do better than this approximation (showing the error decreases as $O(n^{-1})$) by using a more
+precise limit distribution under an additional assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$
+denote the $i^{th}$ absolute moment of $X$. If Cramer's condition holds, i.e. if
+\begin{equation}
+  \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1
+\end{equation}
+
+Bentkus et al.'s work on Edgeworth
+expansions \cite{Bentkus97} equation 2.2 gives:
+
+\begin{lemma} Let $X_1,\ldots,X_n$ be
+  a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3. If $X$ satisfies Cramer's condition, 
+then
+\begin{equation}\label{eq:clt-bounds}
+ \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
+\end{equation}
+where
+\begin{equation}
+G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x)
+\end{equation}
+\end{lemma}
+This lemma shows that the average of the $X_i$ has a more precise, skewed CDF limit $G_n(x,y)$
+where the skew term has weight proportional to a certain measure of skew derived
+from the moments: $\frac{\mu_3}{b_2^{3/2}}$. Note that if the $X_i$ are symmetric,
+the weight of the correction term is zero, and the CDF of the average of the $X_i$ converges
+to $\Phi(x)$ at a rate of $O(n^{-1})$.
+
+Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that
+decays as $n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where $\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
+\begin{equation}
+  \label{eq:GNderivatives}
+G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
+\end{equation}
+From which we see that $G_n(x,y)$ is a linear combination of
+$\Phi(x)$, $\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In
+Algorithm 1, we correct for the difference in $\sigma$ between
+$\Delta^*$ and the variance needed by $X_{\rm corr}$ using $X_{\rm
+  norm}$. This same method works when we wish to estimate the error in $\Delta^*$ vs
+$G_n(x,y)$. Since all of the component functions of $G_n(x,y)$ are
+derivatives of a (unit variance) $\Phi(x)$, adding a normal variable
+with variance $\sigma'$ increases the variance of all three functions
+to $1+\sigma'$. Thus we add $X_{\rm norm}$ as per Algorithm 1
+preserving the limit in equation (\ref{eq:GNderivatives}).
+
+The deconvolution approach can be used to constuct a correction
+variable $X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard
+logistic function. An additional complexity is that $G_n(x,y)$ has
+additional parameters $y$ and $n$. Since these act as a single
+multiplier $\frac{y}{6\sqrt{n}}$ in equation
+(\label{eq:GNderivatives}), its enough to consider a function
+$g(x,y')$ parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can
+be computed and saved offline. As we have shown above, errors in the
+``limit'' function propagate directly through as errors in the
+acceptance test.  To achieve a test error of say $1e-6$ (close to single
+floating point precision), we need a $y'$ spacing of $1e-6$. It
+should not be necessary to tabulate values all the way to $y'=1$,
+since $y'$ is scaled inversely by the square root of monibatch size.
+Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.
+Since our $x$ resolution is 10,000, this leads to a table with about
+1 billion values, which can comfortably be stored in memory. However,
+if $g(x,y)$ is moderately smooth in $y$, it should be possible to
+achieve similar accuracy with a much smaller table. We leave further
+analysis and experiments with $g(x,y)$ as future work.
 

### 2016-09-05 nips_2016.tex
update paper and figures

@@
 \author{
+  John Canny \\
+  Department of Computer Science \\
+  University of California, Berkeley \\
+  \texttt{canny@berkeley.edu}
+  \And
   Haoyu Chen \\
@@
   \texttt{bjiang@berkeley.edu}
-  \And
-  John Canny \\
-  Department of Computer Science \\
-  University of California, Berkeley \\
-  \texttt{canny@berkeley.edu}
  }
@@
 
-We conduct three sets of experiments to explore the benefits of our minibatch MH test and to
+We conduct two sets of experiments to explore the benefits of our minibatch MH test and to
 benchmark it with previous work. In Section~\ref{ssec:gaussians}, we show that our test enables
 samples to converge to the posterior distribution of a heated Gaussian mixture model. In
-Section~\ref{ssec:logistic}, we analyze its efficiency on logistic regression. In
-Section~\ref{ssec:nets}, we apply the test for deep learning.
-Appendices~\ref{app:gaussian},~\ref{app:logistic}, and~\ref{app:nnet} contain more detailed
-information on these respective experiments.
+Section~\ref{ssec:logistic}, we analyze its efficiency on logistic regression. 
+%In %Section~\ref{ssec:nets}, we apply the test for deep learning. 
+Appendices~\ref{app:gaussian},~\ref{app:logistic} %and~\ref{app:nnet} 
+contain more detailed information on these respective experiments.
 
@@
 \end{equation}
-We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta = (0,1)$, we draw 10000
-data points so that the target distribution is $p(\theta)\prod_{i=1}^{10000}p(x_i\mid \theta)$, with
+We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta = (0,1)$, we draw 1,000,000
+data points so that the target distribution is $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with
 the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This
 results in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we apply a
-temperature $T=110$ to reduce ${\rm std}(\Delta')$. Taking logs, we get the target as shown in the
+temperature $T=10,000$ to reduce ${\rm std}(\Delta')$. Taking logs, we get the target as shown in the
 far left of Figure~\ref{fig:gauss_mix_1}.
 
-We run MCMC with our MH test using minibatch size $m=200$. We also run this using standard
-full-batch MCMC (``standard sampling'') with the standard test from Equation~\ref{eq:traditional}, and the method
-from~\cite{cutting_mh_2014} (``adaptive sampling''). For the latter, we also use $m=200$ and
-increment minibatches by that amount within an iteration if necessary. The tolerance for making a
-decision is $\epsilon=0.1$. To make comparisons easier, all three use the same random walk proposer
-with covariance $\Sigma={\rm diag}(0.03, 0.03)$. This is a poor proposer, but it is sufficient for
+We run MCMC with our MH test using minibatch size $m=100$. We also run this using standard
+full-batch MCMC (``standard MCMC'') with the standard test from Equation~\ref{eq:traditional}, the method
+from~\cite{cutting_mh_2014} (``Cut MH'')  and the method from ~\cite{icml2014c1_bardenet14}. For the method in \cite{cutting_mh_2014}, we also use $m=100$ and
+increment minibatches by that amount within an iteration if necessary.
+As for the method in ~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the minibatch size geometrically with a ratio of $\gamma = 1.5$.
+The tolerance for making a
+decision in \cite{cutting_mh_2014} is $\epsilon=0.005$, and the error bound control parameter in ~\cite{icml2014c1_bardenet14} is $p = 2, \delta = 0.01$.
+To make comparisons easier, all three use the same random walk proposer
+with covariance $\Sigma={\rm diag}(0.3, 0.3)$. This is a poor proposer, but it is sufficient for
 our purposes as the quality of the samples will be primarily due to the MH test. All methods are run
-10000 times to collect 10000 samples.
+5000 times to collect 5000 samples.
 
@@
     \centering
-    \includegraphics[width=1\linewidth]{cloud_v01.png}
+    % \includegraphics[width=1\linewidth]{cloud_v01.png}
+    \includegraphics[width=1\linewidth]{posterior_of_gaussian.png}
     \caption{
-    The log posterior contours (temperature 110) and three scatter plots of sampled $\theta$ values.
+    The log posterior contours and scatter plots of sampled $\theta$ values using different methods. 
     }
@@
     \centering
-    \includegraphics[width=1\linewidth]{adaptive_and_ours_information_v01.png}
+   % \includegraphics[width=1\linewidth]{adaptive_and_ours_information_v01.png}
+   \includegraphics[width=1\linewidth]{minibatch_size_gaussian.png}
     \caption{
-    Histograms of minibatch sizes and ${\rm std}(\Delta')$ values, along with ``${\rm std}(\Delta')$
-    violations'' (right).
+    Histograms of minibatch sizes %and ${\rm std}(\Delta')$ values, along with ``${\rm std}(\Delta')$
+   % violations'' (right).
     }
@@
 	\centering
-    \includegraphics[width=1\linewidth]{CI_v01.png}
-    \caption{
-    Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
-    }
-    \label{fig:gauss_mix_3}
+	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
+	\caption{Value of first and second part of equation (9) in ~\cite{icml2014c1_bardenet14} with respect to actual minibatch size}
+	\label{fig:gauss_bardenet}
+\end{figure}
+
+%\begin{figure}[t]
+    %\centering
+    %\includegraphics[width=1\linewidth]{CI_v01.png}
+   % \caption{
+   % Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
+    %}
+   % \label{fig:gauss_mix_3}
     %\vspace{-10pt}
-\end{figure}
+%\end{figure}
 
@@
 other two methods.
-However, the fourth method (i.e.
+%However, the fourth method (i.e.
 just treat the $\Delta'$ which comes from minibatch as $\Delta$ which comes from the complete data) reveals wider spread distribution than the posterior.
@@
 Meanwhile, we show the benefits of our minibatch method compared with the naive minibatch method.
-
-Figure~\ref{fig:gauss_mix_2} (to the left) shows a histogram of the final minibatch sizes used by
-the adaptive subsampling method each iteration (on a log scale for readability).  Usually, the method
-can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
-in lots of wasted computation.  For this particular experiment, that happened eight times.
-The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
-\ge 1.2$ for only consuming one minibatch data. For six minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
-our MH test five times and averaged the number of ``violations.'' As the size increases, the
-number of violations decreases, as expected.
-
+Actually, based on a measure of divergence from the ground-truth distribution, our method is better than the other three methods in terms of this measurement.
+The measurement is to calculate the poisson likelihood of the MCMC-generated parameters over a parameter range, with $\theta_1 \in (-1.5, 2.5)$ and $\theta_2 \in (-2.5, 2.5)$.
+The log likelihood value for our method, standard MCMC, cut MH method and adaptive MH method are -2144.6, -2783.5, -4007.4, -2598.4, respectively.
+
+Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used by
+the four methods in each iteration.
+It is obvious that our methods consumes significantly less data, with most of minibatch sizes smaller than 1000, with a average minibatch size of 420.
+The other three methods occasionally consumes a large proportion of the entire data set, and the average minibatch sizes are 37785, 16378, 16755 for standard MCMC, cut MH method, adaptive MH method, respectively.
+In terms of average minibatch size, our method is nearly 90 times faster than standard MCMC method, and nearly 39 times faster the cut MH method, and nearly 40 times faster than the adaptive MH method.
+
+It is useful to notice that, in ~\cite{icml2014c1_bardenet14}, they need to evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set at every iteration in order to calculate the error bound in equation (9) in ~\cite{icml2014c1_bardenet14}.
+In Figure ~\ref{fig:gauss_bardenet}, the value of the first part and second part of equation (9) in ~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual minibatch data used.
+The second part which contains  the $C_{\theta,\theta'}$ term is larger than the first part, which means that the second part is non-negligible and has to be calculated at every iteration.
+This is one of the very time-consuming parts in this method.
+
+%Usually, the method can make a decision with 200 samples.
+Occasionally, however, it must use all 10000 points, resulting in lots of wasted computation.
+For this particular experiment, that happened eight times.
+%The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
+%\ge 1.2$ for only consuming one minibatch data. For six minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
+% our MH test five times and averaged the number of ``violations.'' As the size increases, the
+% number of violations decreases, as expected.
+
+\iffalse
 Figure~\ref{fig:gauss_mix_3} (to the left) shows a histogram of ${\rm
@@
 of $\rho/\sigma^3$ at each iteration. 
+\fi
 
@@
 points, respectively (we used a random subset of the test data). The proposer is again a random walk
-with covariance matrix $0.1I$ for the $784\times 784$ identity matrix $I$. We set the posterior
-temperature at $T=3000$. We set the minibatch size $m=50$ and compare with adaptive sampling with
-tolerance 0.02.
+with covariance matrix $0.05I$ for the $784\times 784$ identity matrix $I$. We set the posterior
+temperature at $T=1000$. We set the minibatch size $m=100$ and compare with ~\cite{cutting_mh_2014} with
+tolerance 0.005 and with ~\cite{icml2014c1_bardenet14} with error bound control parameter $p = 2, \delta = 0.01$.
 
@@
     \centering
-    \includegraphics[width=1\linewidth]{exp2.png}
-    \caption{Logistic regression performance (accuracy/log likelihood) and minibatch size analysis.}
+    \begin{subfigure}{.5\textwidth}
+     \centering
+     \includegraphics[width=1\linewidth]{LogisticRegressionResult/acc_logistic}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{LogisticRegressionResult/ll_logistic}
+\end{subfigure}
+\caption{Logistic regression performance (accuracy/log likelihood) and minibatch size analysis}
 \label{fig:logistic_fig}
@@
 	\centering
-    \includegraphics[width=1\linewidth]{CI_v02.png}
-    \caption{
-    Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
-    }
-    \label{fig:exp_3}
+	\includegraphics[width=1\linewidth]{LogisticRegressionResult/minibatch_size_logistic.png}
+	\caption{Histograms of minibatch sizes of logistic regression parameter estimation}
+	\label{fig:logistic_minibatch}
+\end{figure}
+
+%\begin{figure}[t]
+  %  \centering
+    %\includegraphics[width=1\linewidth]{CI_v02.png}
+    %\caption{
+   % Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
+    %}
+    %\label{fig:exp_3}
     %\vspace{-10pt}
-\end{figure}
-
-The third plot of Figure~\ref{fig:logistic_fig} shows performance based on minibatch size. The
-parameter is evaluated on the test set after training iterations 700, 1000, and 1500. For those
-three cases, when the minibatch size increases beyond about 10, the performance of the sampled data
-will not improve with a large minibatch size. This makes sense because the error of our MH test is
-proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch increases, ${\rm
-std}(\Delta')$ decreases. Beyond a certain standard deviation (or minibatch) threshold, the error in
-our minibatch MH test is negligible.
-
-Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{fig:exp_3} to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability of $\rho/\sigma^3$.
-
+%\end{figure}
+
+The plot of Figure ~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes. Choosing an initial minibatch size of 100, 
+the standard MCMC, cut MH method, adaptive MH method achieves of final average minibatch size of 910, 585, and 2731 respectively, while our method achieves 
+that of 100, showing significant better performance than the benchmark test methods.
+
+%The third plot of Figure~\ref{fig:logistic_minibatch} shows performance based on minibatch size. The
+%parameter is evaluated on the test set after training iterations 700, 1000, and 1500. For those
+%three cases, when the minibatch size increases beyond about 10, the performance of the sampled data
+%will not improve with a large minibatch size. This makes sense because the error of our MH test is
+%proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch increases, ${\rm
+%std}(\Delta')$ decreases. Beyond a certain standard deviation (or minibatch) threshold, the error in
+%our minibatch MH test is negligible.
+
+%Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{fig:exp_3} to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability of $\rho/\sigma^3$.
+
+\iffalse
 \subsection{Neural Network Optimization}\label{ssec:nets}
@@
 \end{figure}
+
+\fi
 

### 2016-09-06 nips_2016.tex
update paper and figures

@@
 
-We run MCMC with our MH test using minibatch size $m=100$. We also run this using standard
-full-batch MCMC (``standard MCMC'') with the standard test from Equation~\ref{eq:traditional}, the method
-from~\cite{cutting_mh_2014} (``Cut MH'')  and the method from ~\cite{icml2014c1_bardenet14}. For the method in \cite{cutting_mh_2014}, we also use $m=100$ and
-increment minibatches by that amount within an iteration if necessary.
+We run MCMC with our MH test using minibatch size $m=100$. %We also run this using standard
+%full-batch MCMC (``standard MCMC'') with the standard test from Equation~\ref{eq:traditional}, 
+We also run this using the method from~\cite{cutting_mh_2014} (``Cut MH'')  and the method from 
+~\cite{icml2014c1_bardenet14}. For the method in \cite{cutting_mh_2014}, we use $m=100$ and
+increment minibatches by that amount within a test if necessary.
 As for the method in ~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the minibatch size geometrically with a ratio of $\gamma = 1.5$.
@@
 decision in \cite{cutting_mh_2014} is $\epsilon=0.005$, and the error bound control parameter in ~\cite{icml2014c1_bardenet14} is $p = 2, \delta = 0.01$.
-To make comparisons easier, all three use the same random walk proposer
-with covariance $\Sigma={\rm diag}(0.3, 0.3)$. This is a poor proposer, but it is sufficient for
+To make comparisons easier, all three use the same random walk proposer with covariance $\Sigma={\rm diag}(0.3, 0.3)$.
+This is a poor proposer, but it is sufficient for
 our purposes as the quality of the samples will be primarily due to the MH test. All methods are run
@@
     % \includegraphics[width=1\linewidth]{cloud_v01.png}
-    \includegraphics[width=1\linewidth]{posterior_of_gaussian.png}
+    \includegraphics[width=1\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
     \caption{
@@
    % \includegraphics[width=1\linewidth]{adaptive_and_ours_information_v01.png}
-   \includegraphics[width=1\linewidth]{minibatch_size_gaussian.png}
+   \includegraphics[width=1\linewidth]{GaussianMixtureResult/minibatch_size_gaussian.png}
     \caption{
@@
 Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$ samples for the four
-methods, with darker regions indicating a greater density of points. The first three methods obtain the
+methods, with darker regions indicating a greater density of points. The three methods obtain the
 same rough form of the posterior, so our MH test can indeed result in the same posterior as the
@@
 The measurement is to calculate the poisson likelihood of the MCMC-generated parameters over a parameter range, with $\theta_1 \in (-1.5, 2.5)$ and $\theta_2 \in (-2.5, 2.5)$.
-The log likelihood value for our method, standard MCMC, cut MH method and adaptive MH method are -2144.6, -2783.5, -4007.4, -2598.4, respectively.
+The log likelihood value for our method, cut MH method and adaptive MH method are -2144.6, -4007.4, -2598.4, respectively.
 
 Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used by
-the four methods in each iteration.
-It is obvious that our methods consumes significantly less data, with most of minibatch sizes smaller than 1000, with a average minibatch size of 420.
-The other three methods occasionally consumes a large proportion of the entire data set, and the average minibatch sizes are 37785, 16378, 16755 for standard MCMC, cut MH method, adaptive MH method, respectively.
-In terms of average minibatch size, our method is nearly 90 times faster than standard MCMC method, and nearly 39 times faster the cut MH method, and nearly 40 times faster than the adaptive MH method.
+the three methods in each iteration.
+It is obvious that our methods consumes significantly less data, and most minibatch sizes are smaller than 1000, with an average minibatch size of 420.
+The other two methods occasionally consumes a large proportion of the entire data set, and the average minibatch sizes are 16378, 16755 for cut MH method, adaptive MH method, respectively.
+In terms of average minibatch size, our method is nearly 39 times faster the cut MH method, and nearly 40 times faster than the adaptive MH method.
 
@@
 function of the cumulative training data points processed. Our test increments the cumulative data
-by a fixed amount per iteration, but the adaptive method may require more data per iteration.  We
+by a fixed amount per iteration, but the other two methods may require more data per iteration.  We
 see that our minibatch MH test is more efficient; it has similar or better performance while
@@
     \centering
-    \begin{subfigure}{.5\textwidth}
+    \begin{subfigure}{.4\textwidth}
      \centering
      \includegraphics[width=1\linewidth]{LogisticRegressionResult/acc_logistic}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+    \end{subfigure}
+    \begin{subfigure}{.4\textwidth}
      \centering
@@
 The plot of Figure ~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes. Choosing an initial minibatch size of 100, 
-the standard MCMC, cut MH method, adaptive MH method achieves of final average minibatch size of 910, 585, and 2731 respectively, while our method achieves 
+the cut MH method, adaptive MH method achieves final average minibatch size of 585 and 2731 respectively, while our method achieves 
 that of 100, showing significant better performance than the benchmark test methods.
@@
 simple deconvolution process allows us to use a minibatch approximation to the full data tests. We
-experimentally show the benefits of our test on Gaussian mixtures, logistic regression, and deep
-learning.  Straightforward directions for future work include running more experiments with a
+experimentally show the benefits of our test on Gaussian mixtures and logistic regression.
+Straightforward directions for future work include running more experiments with a
 particular focus on investigation of the variance precondition.  More elaborate extensions include

### 2016-09-06 nips_2016.tex
update appendix

@@
 \sigma^2I)$, where $\theta_i$ is the current sample, $\theta'$ is the proposed sample, and we choose
-the variance to be a constant $\sigma^2 = 0.01$ for all components. We initialize $\theta_0$ to be a
-vector of all ones, and set our minibatch size as $m=50$.
+the variance to be a constant $\sigma^2 = 0.05$ for all components. We initialize $\theta_0$ to be a
+vector of all ones, and set our minibatch size as $m=100$.
 
 For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we use a
-constant temperature $T = 3000$. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
-current iteration.  Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
-versus iteration count.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=1\linewidth]{./figures/appendixExp2}
-  \caption{Additional results for the logistic regression experiment.}
-  \label{fig:appendixExp2}
-\end{figure}
-
-For adaptive MH testing, our experimental settings are the same as with our MH test, except we do
-not impose a temperature. The minibatch size of adaptive MH testing is also initialized as 50, but
-it may increase by that amount each iteration. Figure~\ref{fig:appendixExp2} shows the histogram of
-the actual minibatch size at the end of each iteration in adaptive MH testing.
-
+constant temperature $T = 1000$. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
+current iteration.  %Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
+%versus iteration count.
+
+%\begin{figure}[t]
+  %\centering
+  %\includegraphics[width=1\linewidth]{./figures/appendixExp2}
+ % \caption{Additional results for the logistic regression experiment.}
+  %\label{fig:appendixExp2}
+%\end{figure}
+
+%For adaptive MH testing, our experimental settings are the same as with our MH test, except we do
+%not impose a temperature. The minibatch size of adaptive MH testing is also initialized as 50, but
+%it may increase by that amount each iteration. Figure~\ref{fig:appendixExp2} shows the histogram of
+%the actual minibatch size at the end of each iteration in adaptive MH testing.
+
+\iffalse
 \section{Neural Network Experiment Details}\label{app:nnet}
@@
 
+\fi
 % Daniel: here are example LaTeX codes for figures and tables if we want to use them.

### 2016-09-20 nips_2016.tex
Updated the paper draft with around a third of my handwritten comments. Still a lot to go.

@@
 
+\newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
 \newtheorem{lemma}{Lemma}
@@
 \author{
-  John Canny \\
-  Department of Computer Science \\
+  John Canny, Haoyu Chen, Daniel Seita, Xinlei Pan \\
   University of California, Berkeley \\
-  \texttt{canny@berkeley.edu}
-  \And
-  Haoyu Chen \\
-  Department of Computer Science \\
-  University of California, Berkeley \\
-  \texttt{haoyuchen@berkeley.edu}
-  \And
-  Daniel Seita \\
-  Department of Computer Science \\
-  University of California, Berkeley \\
-  \texttt{seita@berkeley.edu}
-  \And
-  Xinlei Pan \\
-  Department of Bioengineering \\
-  University of California, Berkeley \\
-  \texttt{xinleipan@berkeley.edu}
-  \And 
-  Biye Jiang \\
-  Department of Computer Science \\
-  University of California, Berkeley \\
-  \texttt{bjiang@berkeley.edu}
+  \texttt{\{canny,haoyuchen,seita,xinleipan\}@berkeley.edu}
  }
@@
 the cost of Metropolis-Hastings tests yield variable data consumed per
-sample, with only constant factor reductions vs. using the full
+sample, with only constant factor reductions versus using the full
 dataset for each sample.  Here we present a method that can be tuned
@@
 step size or temperature. Our approach uses the natural noise present
-in minibatch likelihood esimates to furnish the randomness in a
+in minibatch likelihood estimates to furnish the randomness in a
 Metropolis-Hastings test. Our test uses the noise-tolerant Barker
@@
 and where the number of data instances $N$ is large.  The
-Metroplis-Hastings method (M-H) generates sample candidates from a
+Metropolis-Hastings method (M-H) generates sample candidates from a
 proposal distribution $q$ which is in general different from the
@@
 
-Many state-of-the-art machine learning methods, and deep learning in
-particular, are based on minibatch updates (such as SGD) to a model.
-Minibatch updates produce many improvements to the model for each pass
-over the dataset, and have high sample efficiency. They also
-map very well onto hardware such as GPUs. In contrast, M-H requires
-calculations over the full dataset to produce a new sample.  Recent
-results from~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
-approximate (bounded error) acceptance tests using subsets
-(minibatches) of the full dataset. The tests depend on minibatch
-statistics, and on the value of an additional random variable $u$. The
-amount of data consumed for each test varies significantly from one
-minibatch to the next, and depends on the current sample, the proposed
-sample, and on the random variable $u$. By contrast,
-\cite{conf/uai/MaclaurinA14} performs exact tests but requires a lower
-bound on parameter distribution across its domain. The amount of data
-reduction depends on the accuracy of this bound, and such bounds are
-only available for relatively simple distributions.
+Many state-of-the-art machine learning methods, and deep learning in particular,
+are based on minibatch updates (such as SGD) to a model.  Minibatch updates
+produce many improvements to the model for each pass over the dataset, and have
+high sample efficiency. They also map very well onto hardware such as GPUs. In
+contrast, M-H requires calculations over the full dataset to produce a new
+sample.  Recent results from~\cite{cutting_mh_2014,icml2014c1_bardenet14}
+perform approximate (bounded error) acceptance tests using subsets (minibatches)
+of the full dataset. The tests depend on minibatch statistics, and on the value
+of an additional uniform random variable $u$. The amount of data consumed for
+each test varies significantly from one minibatch to the next, and depends on
+the current sample, the proposed sample, and on the random variable $u$. By
+contrast,~\cite{conf/uai/MaclaurinA14} performs exact tests but requires a lower
+bound on parameter distribution across its domain.  The amount of data reduction
+depends on the accuracy of this bound, and such bounds are only available for
+relatively simple distributions.
 
@@
 probability and in expected value. We use a Barker test function
-\cite{Barker65} rather than a Metropolis test, which makes the test
+\cite{Barker65} rather than a Metropolis test, which makes our test
 naturally error tolerant. The idea of using a noise-tolerant test
-using Barker's test function were suggested but not explored
+using Barker's test function was suggested but not explored
 empirically in \cite{Bardenet15} section 6.3. But the asymptotic test
@@
 
-\item Increase the minibatch size. log acceptance variance scales as $1/k$ vs the
+\item Increase the minibatch size. Log acceptance variance scales as $1/k$ vs the
   minibatch size $k$. Increased minibatch size also reduces the error rate for
-  the test.
+  the test. {\color{blue} Daniel: I'm not sure if I'd mention this here, because
+  before this list, we talk about reducing the data, and then suddenly we're
+  talking about increasing it.}
 
@@
 
-Its worth discussing at this point what are the goals (typically) of
+Its worth discussing at this point the typical goals of
 M-H sampling on very large datasets.  By the Bernstein-von Mises
-theorem, the posterior distribution of the parameter $\theta$ for a
+Theorem, the posterior distribution of the parameter $\theta$ for a
 Bayesian inference task is asymptotically normal, and has variance
@@
 sampler to find one of them and become trapped in it when run at the
-normal distribution temperature (T=1). A common solution is to anneal
+normal distribution temperature ($T=1$). A common solution is to anneal
 the sampler, running first at high temperature (scaling log
-likelihoods by 1/T) which flattens the likelihood landscape.  This is
+likelihoods by $1/T$) which flattens the likelihood landscape.  This is
 turn reduces the variance of the log acceptance probability and allows
@@
 nevertheless make steady progress to a posterior mode since their bias
-is in that direction. We demonstrate this behavior with our experiments.
+is in that direction. We demonstrate this behavior in our logistic regression
+experiments.
 
@@
   of magnitude more efficient than prior work measured as data consumed per test, and that it does not suffer
-  from long-tailed batch sizes (up to the dataset size). 
+  from long-tailed minibatch sizes (up to the dataset size). 
 \end{itemize}
@@
 
-In the Metropolis-Hastings method~\cite{gilks1996markov,brooks2011handbook}, a difficult-to-compute
-probability distribution $p(\theta)$ is sampled using a Markov chain $\theta_1,\ldots,\theta_n$. The 
-sample $\theta_{t+1}$ at time $t+1$ is generated using a candidate $\theta'$ from a (simpler) proposal
-distribution $q(\theta'\mid \theta_t)$, filtered by an acceptance test. The acceptance test is usually
-a Metropolis test. The Metropolis test has acceptance probability:
+In the Metropolis-Hastings method~\cite{gilks1996markov,brooks2011handbook}, a
+difficult-to-compute probability distribution $p(\theta)$ is sampled using a
+Markov chain $\theta_1,\ldots,\theta_n$. The sample $\theta_{t+1}$ at time $t+1$
+is generated using a candidate $\theta'$ from a (simpler) proposal distribution
+$q(\theta'\mid \theta_t)$, filtered by an acceptance test. The acceptance test
+is usually a Metropolis test. The Metropolis test has acceptance probability:
 \begin{equation}\label{eq:traditional}
@@
 \end{equation}
-where $a \wedge b$ denotes $\min(a,b)$.  With probability $\alpha(\theta_t,\theta')$, we accept $\theta'$
-and set $\theta_{t+1} = \theta'$, otherwise set
-$\theta_{t+1}=\theta_t$.  The test is often implemented with an
-auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison $u < \alpha(\theta_t,\theta')$. $\mathcal{U}(a,b)$ denotes the uniform distribution on the interval $[a,b]$. 
-For notational simplicity, from now on we will drop the subscript $t$ for the current sample $\theta_t$ and
-denote is simply as $\theta$.
+where $a \wedge b$ denotes $\min(a,b)$.  With probability
+$\alpha(\theta_t,\theta')$, we accept $\theta'$ and set $\theta_{t+1} =
+\theta'$, otherwise set $\theta_{t+1}=\theta_t$.  The test is often implemented
+with an auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison
+$u<\alpha(\theta_t,\theta')$; here, $\mathcal{U}(a,b)$ denotes the uniform
+distribution on the interval $[a,b]$.  For notational simplicity, we drop the
+subscript $t$ for the current sample $\theta_t$ and denote it as $\theta$.
 
 The acceptance test guarantees detailed balance, which means
-\begin{equation}\label{detailed_balance1}
-  p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')
-\end{equation}
-where $p(\theta'\mid\theta)$ is the probability of a transition from state $\theta$ to $\theta'$. Here
-$p(\theta'\mid\theta) = q(\theta'\mid\theta)\alpha(\theta,\theta')$ so the detailed balance equation becomes
-\begin{equation}\label{detailed_balance2}
+$p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')$,
+where $p(\theta'\mid\theta)$ is the probability of a transition from state
+$\theta$ to $\theta'$. Here $p(\theta'\mid\theta) =
+q(\theta'\mid\theta)\alpha(\theta,\theta')$ so the detailed balance equation
+becomes
+\begin{equation}\label{eq:detailed_balance2}
     p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
 \end{equation}
-The detailed balance condition, together with ergodicity, guarantee that the
-Markov chain has a unique stationary distribution $\pi(\theta) = p(\theta)$.
-
-For Bayesian inference, we would like to sample from a parameter distribution for $\theta$ based on some
-observed data $x_1,\ldots,x_N$. i.e. we want to sample from 
+This condition, together with ergodicity, guarantees that the Markov chain has a
+unique stationary distribution $\pi(\theta) = p(\theta)$.
+
+For Bayesian inference, we would like to sample from a parameter distribution
+for $\theta$ based on some observed data, so the target distribution is
 $p(\theta \mid x_1, \ldots, x_N)$. The acceptance probability is now:
-\begin{equation}
+\begin{equation}\label{eq:acceptance_probability}
     \alpha(\theta,\theta') = 
     \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
-\theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)} \wedge 1
-\end{equation}
-where $p_0(\theta)$ is a prior, and $p(x_i \mid \theta)$ are the
-probabilities of the observations. Computing samples this way requires
-the use of all $N$ training data points, but this is very expensive
-for large datasets. To address this challenge,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
-approximate Metropolis-Hasting tests using sequential hypothesis
-testing. During each iteration, they start with a small minibatch of
-data and test the hypothesis that the sample $\theta'$ based on an
-approximate version of the test $u < \alpha(\theta,\theta')$. If the
-approximate test cannot make a decision with sufficient confidence,
-then the minibatch size is increased and the test repeats. This
-process continues until a decision. The bounds depend on either a
-asymptotic central limit theorem \cite{cutting_mh_2014} or a
-concentration bound \cite{icml2014c1_bardenet14}, the latter requiring
-direct bounds on the log likelihood ratio.  The problem with both
-methods is the expense of resolving small differences.  In the worst
-case, all $N$ data points may be needed, and as we show next this
-worst case can occur about $\Omega(N)$ times during the performance of
-$N$ tests.
-
-Following ~\cite{icml2014c1_bardenet14}, we write the test $u < \alpha(\theta,\theta')$ in
-the equivalent form $\Lambda(\theta,\theta') > \psi(u,\theta,\theta')$. Our definitions differ
-from those in \cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later on. 
+    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)}
+    \wedge 1
+\end{equation}
+where $p_0(\theta)$ is a prior, and $p(x_i \mid \theta)$ are the probabilities
+of the observations. Computing samples this way requires the use of all $N$
+training data points, but this is very expensive for large datasets. To address
+this challenge,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
+Metropolis-Hasting tests using sequential hypothesis testing. During each
+iteration, they start with a small minibatch of data and test the hypothesis
+that the sample $\theta'$ should be accepted based on an approximate version of
+the test $u < \alpha(\theta,\theta')$. If the approximate test cannot make a
+decision with sufficient confidence, then the minibatch size is increased and
+the test repeats. This process continues until a decision. The bounds depend on
+either a asymptotic central limit theorem \cite{cutting_mh_2014} or a
+concentration bound \cite{icml2014c1_bardenet14}. The latter requires direct
+bounds on the log likelihood ratio, which for general distributions requires
+knowing $p(x_i \mid \theta)$ and $p(x_i \mid \theta')$ for all $N$ samples. In
+addition, both methods suffer the drawback of resolving small log likelihood
+ratio differences between the minibatch and full batch versions.  In the worst
+case, all $N$ data points may be needed, and we soon show that this worst case
+can occur about $\Omega(N)$ times during the performance of $N$ tests.
+
+Following~\cite{icml2014c1_bardenet14}, we write the test $u <
+\alpha(\theta,\theta')$ in the equivalent form $\Lambda(\theta,\theta') >
+\psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those in
+\cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later.}
 \begin{equation}
@@
 \end{equation}
-
 To reduce computational effort, an unbiased estimate of $\Lambda(\theta,\theta')$
 based on a minibatch can be used:
-
 \begin{equation}
@@
 \end{equation}
-
-Finally, it will be convenient in what follows to define the individual components that
-contribute to the sums above:
-
+Finally, it will be convenient for our analysis to define the individual
+components that contribute to the sums above:
 \begin{equation}\label{eq:individual_terms}
@@
 \end{equation}
-
-So $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over the
-entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean over the minibatch
-of $\Lambda_i(\theta,\theta')$.
-
-Since the minibatch contains randomly selected samples $x_i$, the
-values $\Lambda_i$ are independent, IID random variables.  By the central
-limit theorem, we expect $\Lambda^*(\theta,\theta')$ to be
-approximately Normal.  The acceptance test then becomes a statistical
-test of the hypothesis that $\Lambda(\theta,\theta') >
-\psi(u,\theta,\theta')$ by establishing that
+Thus, $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over
+the entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean of
+$\Lambda_i(\theta,\theta')$ over its minibatch.
+
+Since minibatches contains randomly selected samples $x_i$, the values
+$\Lambda_i$ are independent, identically distributed (iid) random variables.
+{\color{blue} Daniel: should we say sampling with or without replacement? Right
+now I think the code does it without replacement (but \emph{with} replacement
+means it's closer to the CLT, right?} By the Central Limit Theorem, we expect
+$\Lambda^*(\theta,\theta')$ to be approximately Gaussian. The acceptance test
+then becomes a statistical test of the hypothesis that
+$\Lambda(\theta,\theta')>\psi(u,\theta,\theta')$ by establishing that
 $\Lambda^*(\theta,\theta')$ is substantially larger than
-$\psi(u,\theta,\theta')$.  In ~\cite{cutting_mh_2014} an asymptotic
-central limit argument was used to derive this gap, while in
-\cite{icml2014c1_bardenet14} a concentration bound was used. In both
-cases, the resulting tests were shown to give useful reductions in
-number of samples required over using the full dataset. But no
-worst-case bounds were given on average number of samples needed.
-
-We next show that for some simple distributions, the lower
-bound of average number of data instances consumed for one iteration of
-~\cite{cutting_mh_2014} and ~\cite{icml2014c1_bardenet14} is
-$\Omega(N)$, where $N$ is the number of data points.
-
-\subsection{An Example}
-Consider a simple model where samples $x_i$ for $i=1,\ldots,N$ are
-drawn from a Normal distribution $\mathcal{N}(\theta,1)$ where
-$\theta=0.5$, and with a uniform prior on $\theta$. The log likelihood ratio is
+$\psi(u,\theta,\theta')$.  In~\cite{cutting_mh_2014} an asymptotic central limit
+argument was used to derive this gap, while in~\cite{icml2014c1_bardenet14} a
+concentration bound was used. In both cases, the resulting tests were shown to
+give useful reductions in number of samples required over using the full
+dataset, but there were no worst-case bounds given on the average number of
+samples needed per iteration.
+
+We next show that for a simple distribution, the lower bound of average
+number of data instances consumed for one iteration of~\cite{cutting_mh_2014}
+and~\cite{icml2014c1_bardenet14} is $\Omega(N)$, where $N$ is the number of data
+points.
+
+\subsection{A Worst-Case Gaussian Example}\label{ssec:gaussian_example}
+Consider a Gaussian model where $x_1,\ldots,x_N \simiid \mathcal{N}(\theta,1)$
+with a known variance $\sigma^2=1$, true mean $\theta=0.5$, and a uniform prior
+on $\theta$. The log likelihood ratio is
 \begin{equation}
@@
 \end{equation}
-which is normally distributed over selection of the Normal samples $x_i$.
-Since the $x_i$ have unit variance, their mean has variance $1/b$, and the variance $\sigma^2(\Lambda^*)$
-of $\Lambda^*(\theta,\theta')$ is $(\theta'-\theta)^2N/b$. In order to pass
-a hypothesis test that $\Lambda > \psi$, we will need to establish a
-large enough gap (several $\sigma(\Lambda^*)$) between $\Lambda^*(\theta',\theta)$
-and $\psi(u,\theta',\theta)$.
-
-We would like to sample from the posterior distribution of $\theta$
-which is a normal distribution centered on the sample mean, and with
-variance $1/N$.  The target distribution is therefore
-$\mathcal{N}(\mu,1/N)$ where $\mu$ is the sample mean, and $\mathcal{N}(\mu,\sigma^2)$ is a
-normal distribution with mean $\mu$ and variance $\sigma^2$.  An efficient
-proposal distribution has the same variance as the target distribution in one dimension
-\cite{OptimalScaling01}, so take the proposal
-$\mathcal{N}(\theta'-\theta,1/N)$, which is implemented as $q(\theta'\mid\theta) =
-\phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal density
-function with zero mean and unit variance. This proposal is symmetric
-$q(\theta'\mid\theta)=q(\theta\mid\theta')$, and we assumed uniform
-prior for $\theta$, so $\psi(u,\theta',\theta)$ reduces to $\log
-u$.
-
-\begin{lemma}\label{lem:efficienty_cutting}
-  For the model above, there
-  exists a fixed (independent of $N$) constant $c$ such that with probability $ \geq c$ over the joint
-  distribution of $(\theta, \theta', u)$, the
-  tests from \cite{icml2014c1_bardenet14} and \cite{cutting_mh_2014} consume all $N$ samples. 
+which is normally distributed over selection of the Normal samples $x_i$.  Since
+the $x_i$ have unit variance, their mean has variance $1/b$, and the variance of
+$\Lambda^*(\theta,\theta')$ is $\sigma^2(\Lambda^*) = (\theta'-\theta)^2N^2/b$.
+In order to pass a hypothesis test that $\Lambda > \psi$, there needs to be a
+large enough gap (several $\sigma(\Lambda^*)$) between
+$\Lambda^*(\theta',\theta)$ and $\psi(u,\theta',\theta)$. {\color{blue}
+Daniel: I changed the variance to be $(\theta'-\theta)^2N^2/b$; before it had
+$N$ instead of $N^2$.}
+
+Our goal is to sample from the posterior $p(\theta \mid x_1,\ldots,x_N)$, which
+is a normal distribution $\mathcal{N}(\mu, 1/N)$ centered on the sample mean
+$\mu$, and with variance $1/N$. In one dimension, an efficient proposal
+distribution has the same variance as the target
+distribution~\cite{OptimalScaling01}, so we use the proposal
+$\mathcal{N}(\theta'-\theta,1/N)$, which is implemented as
+$q(\theta'\mid\theta)=\phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal
+density function with zero mean and unit variance. This proposal is symmetric
+$q(\theta'\mid\theta)=q(\theta\mid\theta')$, and since we assumed a uniform
+prior on $\theta$, we see that $\psi(u,\theta',\theta)$ reduces to $\log u$. Our
+worst-case scenario is specified in Lemma~\ref{lem:worst_case}.
+{\color{blue} Daniel: this might be a silly question, but why do we assume the
+proposal has mean $\theta'-\theta$? I thought it would have a mean of $\theta$
+because that is the current sample and the Gaussian is centered at that point?}
+
+\begin{lemma}\label{lem:worst_case}
+    For the model in Section~\ref{ssec:gaussian_example}, there exists a fixed
+    (independent of $N$) constant $c$ such that with probability $\geq c$ over
+    the joint distribution of $(\theta, \theta', u)$, the tests
+    from~\cite{cutting_mh_2014,icml2014c1_bardenet14} consume all $N$ samples. 
 \end{lemma}
-Proof is given in the appendix.
-
-Similar results can be proved for other distributions and proposals, by identifying
-regions in product space $(\theta,\theta'-\theta,u)$ such that the hypothesis test
-needs to separate nearly-equal values.
-
-It follows that the accelerated M-H tests in \cite{icml2014c1_bardenet14} and
-\cite{cutting_mh_2014} require at least a constant fraction $\geq c$
-in the amount of data consumed per test, compared to full-dataset tests.
-i.e. their speed-up is at most $1/c$.
+
+\begin{proof}
+See Appendix~\ref{app:worst_case_proof}.
+\end{proof}
+
+Similar results can be shown for other distributions and proposals by
+identifying regions in product space $(\theta,\theta'-\theta,u)$ such that the
+hypothesis test needs to separate nearly-equal values.  It follows that the
+accelerated M-H tests in~\cite{cutting_mh_2014,icml2014c1_bardenet14} require at
+least a constant fraction $\geq c$ in the amount of data consumed per test
+compared to full-dataset tests, so their speed-up is at most $1/c$.
 
 \subsection{MCMC Posterior Inference}
-There is a separate line of MCMC work drawing principles from
-statistical physics. By viewing random variables as particles in a
-system, one can apply Hamiltonian Monte Carlo
-(HMC)~\cite{mcmc_hamiltonian_2010} methods which generate high
-acceptance \emph{and} distant proposals when run on full batches of
-data. Recently Langevin
-Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has been applied to
-Bayesian estimation on minibatches of data. This simplified dynamics
-uses local proposals and avoids MH tests by using small proposal steps
-whose acceptance approaches 1 in the limit. However, the constraint on
-proposal step size is severe, and the state space exploration reduces
-to a random walk. Full minibatch HMC for minibatches was recently
-described in ~\cite{sghmc_2014} which allows momentum-augmented
-proposals with larger step sizes. However, step sizes are still
-limited by the need to run accurately without MH tests.
-% \cite{stochastic_thermostats_2014}
-Our work opens the door to applying those methods with much
-more aggressive step sizes without loss of accuracy. We demonstrate
-this in Section~\ref{ssec:nets}.
-
-%% Daniel: for a further discussion on the 'Firefly' paper, see this blog post:
-%% https://xianblog.wordpress.com/2014/04/02/firefly-monte-carlo/
-%% Note: this guy also blogged about our earlier ICLR 2015 paper. He's the third author of the SAME
-%% paper, which explains why he knows this stuff.
+There is a separate line of MCMC work drawing principles from statistical
+physics. By viewing random variables as particles in a system, one can apply
+Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which
+generate high acceptance \emph{and} distant proposals when run on full batches
+of data. Recently Langevin Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has
+been applied to Bayesian estimation on minibatches of data. This simplified
+dynamics uses local proposals and avoids MH tests by using small proposal steps
+whose acceptance approaches 1 in the limit. However, the constraint on proposal
+step size is severe, and the state space exploration reduces to a random walk.
+Full minibatch HMC for minibatches was recently described in~\cite{sghmc_2014}
+which allows momentum-augmented proposals with larger step sizes. However, step
+sizes are still limited by the need to run accurately without MH tests.  Our
+work opens the door to applying those methods with much more aggressive step
+sizes without loss of accuracy. We demonstrate this in Section ... {\color{blue}
+Daniel: do we ever demonstrate this? I guess another issue is that this implies
+we can use larger step sizes, but earlier we \emph{also} argue that we need
+\emph{small} step sizes (to reduce variance).}
 
@@
 
-For our new M-H test, we denote the exact and approximate log likelihood ratios as $\Delta$ and $\Delta^*$.
-First, $\Delta$ is defined as:
-\begin{equation}\label{eq:delta0}
-  \Delta(\theta,\theta') = \log\frac{p(\theta')q(\theta\mid\theta')}{p(\theta)q(\theta'\mid\theta)}
-\end{equation}
-and in the Bayesian case
+{\color{blue} Daniel: cleaned up this section, deleted redundant delta equation,
+deleted a duplicate equation in Lemma~\ref{lem:detailed_balance}.}
+
+\subsection{Log-Likelihood Ratios}\label{ssec:log_likelihood_ratios}
+
+For our new M-H test, we denote the exact and approximate log likelihood ratios
+as $\Delta$ and $\Delta^*$.  First, $\Delta$ is defined as
 \begin{equation}\label{eq:delta1}
@@
     \log \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
-  \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)}
-  \end{equation}
-and we can separate out terms dependent and independent of the data $x$ as:
+    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)},
+\end{equation}
+where $p_0, p$, and $q$ match the corresponding functions within
+Equation~\ref{eq:acceptance_probability}. We separate out terms dependent and
+independent of the data $x$ as:
 \begin{equation}\label{eq:delta2}
     \Delta(\theta,\theta') = \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') =
-  \Lambda(\theta,\theta') -  \psi(1,\theta,\theta')
-\end{equation}
-A minibatch estimator of $\Delta$ can be defined as:
+    \Lambda(\theta,\theta') - \psi(1,\theta,\theta').
+\end{equation}
+A minibatch estimator of $\Delta$, denoted as $\Delta^*$, is
 \begin{equation}\label{eq:delta3}
@@
 \end{equation}
-Note that $\Delta$ and $\Delta^*$ are evaluated on the full dataset and a minibatch of size $b$
-respectively. The scaling term $\frac{N}{b}$ ensures that $\Delta^*(\theta,\theta')$ is an
-unbiased estimator of $\Delta(\theta,\theta')$.
-
-\subsection{A Full Dataset Acceptance Test}\label{ssec:deltas_full}
-
-The key to our test is a smooth acceptance function.  We consider tests other than the Metropolis
-test that satisfy the conditions needed for accurate posterior estimation. This condition is
-the detailed balance condition defined earlier. Expressed in terms of $\Delta$, the
-classical Metropolis test is $f(\Delta) = {\rm min}(\exp(\Delta),1)$. More generally we have:
+Note that $\Delta$ and $\Delta^*$ are evaluated on the full dataset and a
+minibatch of size $b$ respectively. The scaling term $N/b$ ensures that
+$\Delta^*(\theta,\theta')$ is an unbiased estimator of $\Delta(\theta,\theta')$.
+{\color{blue} Daniel: a brief question repeated from earlier, should we mention
+sampling with and/or without replacement here?}
+
+The key to our test is a smooth acceptance function.  We consider functions
+other than the classical Metropolis test that satisfy the detailed balance (see
+Equation~\ref{eq:detailed_balance2}) condition needed for accurate posterior
+estimation. A class of functions leading to detailed balance is specified as
+follows:
 
 \begin{lemma}\label{lem:detailed_balance}
-  If $g(s)$ is any function such that $g(s) = \exp(s) g(-s)$, then the acceptance function $\alpha(\theta,\theta') \triangleq  g(\Delta(\theta,\theta'))$ satisfies detailed balance. That is,
-  \begin{equation}\label{detbal}
-    p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
-    \end{equation}
-
+    If $g(s)$ is any function such that $g(s) = \exp(s) g(-s)$, then the
+    acceptance function $\alpha(\theta,\theta') \triangleq
+    g(\Delta(\theta,\theta'))$ satisfies detailed balance.
 \end{lemma}
 
-This result is used in \cite{Barker65} to define the Barker acceptance test. 
-As a sanity check, choosing $g(s) = \exp(s) \wedge 1$ produces the
-standard Metropolis acceptance test $\alpha(\theta,\theta') = g(\Delta(\theta,\theta')) = 
-\frac{p(\theta')q(\theta \mid
-  \theta')}{p(\theta)q(\theta' \mid \theta)}\wedge 1$. $g(s)$ satisfies
-the condition $g(s) = \exp(s)g(-s)$, and therefore detailed
-balance. In fact it is the optimal acceptance function in terms of
-acceptance rate, since it accepts with probability 1 for $\Delta > 0$.
-
-\subsection{Barker (Logistic) Acceptance Function}
+This result is used in~\cite{Barker65} to define the Barker acceptance test.  As
+a sanity check, choosing $g(s) = \exp(s) \wedge 1$ --- a function satisfying the
+requirement of Lemma~\ref{lem:detailed_balance} --- produces the classical
+Metropolis acceptance test $\alpha(\theta,\theta') = g(\Delta(\theta,\theta')) =
+\frac{p(\theta')q(\theta \mid \theta')}{p(\theta)q(\theta' \mid \theta)}\wedge
+1$. In fact, $g(s) =\exp(s) \wedge 1$ is the optimal acceptance function in
+terms of acceptance rate, since it accepts with probability 1 for $\Delta > 0$.
+
+\subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
 For our new MH test we use the Barker logistic function:
@@
 
-Assume we begin with the current sample $\theta$  and a
-candidate sample $\theta'$. Let $V$ be a uniform random variable $V \sim
-\mathcal{U}(0,1)$. We accept $\theta'$ if $g(\Delta(\theta,\theta')) > V$, and reject otherwise.
-Since $g(s)$ is monotonically increasing, its inverse $g^{-1}(s)$ is well-defined and
-unique. So an equivalent test is to accept $\theta'$ iff
-\begin{equation}
+Assume we begin with the current sample $\theta$  and a candidate sample
+$\theta'$. Let $V$ be a uniform random variable $V \sim \mathcal{U}(0,1)$. We
+accept $\theta'$ if $g(\Delta(\theta,\theta')) > V$, and reject otherwise.
+Since $g(s)$ is monotonically increasing, its inverse $g^{-1}(s)$ is
+well-defined and unique. So an equivalent test is to accept $\theta'$ iff
+\begin{equation}\label{eq:equivalent_test}
     \Delta(\theta,\theta') > X = g^{-1}(V)
 \end{equation}
-where $X$ is a random variable with the logistic distribution (its CDF is the logistic function). To
-see this notice that $\frac{dV}{dX} = g'$, that $g'$ is the density corresponding to a logistic
-CDF, and finally that $\frac{dV}{dX}$ is the density of $X$. The density of $X$ is symmetric,
-so we can equivalently test whether
-\begin{equation}
-  \label{theexacttest}
+where $X$ is a random variable with the logistic distribution (its CDF is the
+logistic function). To see this notice that $\frac{dV}{dX} = g'$, that $g'$ is
+the density corresponding to a logistic CDF, and finally that $\frac{dV}{dX}$ is
+the density of $X$. The density of $X$ is symmetric, so we can equivalently test
+whether
+\begin{equation}\label{eq:the_exact_test}
     \Delta(\theta,\theta') + X > 0
 \end{equation}
-for a logistic random $X$.
+for a logistic random variable $X$. {\color{blue} Daniel: I think the text
+before Equation~\ref{eq:equivalent_test} should be similar to what we had in
+our NIPS submission because I find it much clearer.}
 
@@
 $\Delta^*(\theta,\theta')$. From equation (\ref{eq:delta3}),
-$\Delta^*(\theta,\theta')$ can be represented as a constant term
-plus the mean of $b$ IID terms $\Lambda_i(\theta,\theta')$ of the form
-$N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $b$ increases, $\Delta^*(\theta,\theta')$
-therefore has a distribution which approaches a normal distribution by the central limit
-theorem. Later we will provide specific bounds between the CDF of $\Delta^*(\theta,\theta')$
-and a Normal CDF, but for now we use an asymptotic argument.
-
-In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$, we
-can write
+$\Delta^*(\theta,\theta')$ can be represented as a constant term plus the mean
+of $b$ IID terms $\Lambda_i(\theta,\theta')$ of the form
+$N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $b$ increases,
+$\Delta^*(\theta,\theta')$ therefore has a distribution which approaches a
+normal distribution by the central limit theorem. Later we will provide specific
+bounds between the CDF of $\Delta^*(\theta,\theta')$ and a Normal CDF, but for
+now we use an asymptotic argument.
+
+In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$,
+we can write
 \begin{equation}\label{eq:relationship}
@@
 \end{equation}
-where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is approximately
-normal with variance $\sigma^2(\Delta^*)$. 
-But to perform the test in equation (\ref{theexacttest}) we want $\Delta + X$ for
-a logistic random $X$ (call it $X_{\rm log}$ from now on). In \cite{Bardenet15} it was proposed to
-use $\Delta^*$ in a Barker test anyway and tolerate the fixed error caused by this approximation.
+where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is
+approximately normal with variance $\sigma^2(\Delta^*)$.  But to perform the
+test in Equation~\ref{eq:the_exact_test} we want $\Delta + X$ for a logistic
+random $X$ (call it $X_{\rm log}$ from now on). In \cite{Bardenet15} it was
+proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed error
+caused by this approximation.
 
@@
 
-Algorithm~\ref{alg:our_algorithm} describes the M-H iteration using our acceptance test.
+\section{Computing the Correction Distribution}\label{sec:correction}
+
+Our goal is to compute the distribution of the correction variable $X_{\rm corr}$ such that
+$X_{\rm norm} + X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \in \mathcal{N}(0,\sigma)$
+and $X_{\rm log}$ has a standard logistic CDF $(1+\exp(-X))^{-1}$.
+In the next section we show that the accuracy of the test depends on the absolute
+error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and the distribution of $X_{\rm log}$.
+So we minimize this in our construction.
+
+Our goal then is to minimize
+\begin{equation}
+  |\Phi_{\sigma} * C_{\sigma} - S|
+\end{equation}
+where $\Phi_{\sigma}(X) = \Phi(X/\sigma)$ and $\Phi$ is the standard
+normal CDF, $S(X)$ is the logistic function, $C_{\sigma}(X)$ is the
+{\em density} of the correction distribution, and $*$ denotes
+convolution. In continuous form, this value is
+\begin{equation}
+  \label{eq:conv1}
+  \sup_X\left|\int_{-V}^{V}\Phi_{\sigma}(X-Y) C_{\sigma}(Y) dY - S(X)\right|
+  \end{equation}
+Assuming $X_{\rm corr}$ lies in the interval $[-V,V]$. For computation of $C_{\sigma}(Y)$, we
+discretize $X$ and $Y$ into $4N+1$ and $2N+1$ values respectively, with $X \in [-2V,2V]$ and
+$Y \in [-V,V]$. Writing $X_i = i * V / N$ and $Y_j = j * V / N$, the condition in
+equation (\ref{eq:conv1}) can be written:
+\begin{equation}
+  \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|
+\end{equation}
+and defining a matrix $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, vectors $U_j = C_{\sigma}(Y_j)$ and $V_i = S(X_i)$, the
+above is equivalent to minimizing the 1-norm:
+\begin{equation}
+  \label{eq:error_norm}
+  ||MU-V||_1
+\end{equation}
+wrt $U$. We have an additional constraint that $U_j > 0$ for all $j$, since $U$ represents
+a density. We first explored optimizing this system with linear programming to find the $U$ such that:
+\begin{equation}
+  \begin{split}
+    & \min(\epsilon) \qquad \hbox{where} \\
+    & -\epsilon \leq MU - V \leq \epsilon \\
+    & U \geq 0
+  \end{split}
+\end{equation}
+which was tractable up to a few hundred dimensions. However, the discretization error is bounded by
+the curvature of the underlying functions which are here slightly less than one. i.e.
+the errors are of the order of $(V/N)^2$. Here we chose $V=20$ to provide adequate containment of
+the distributions (the CDFs are extremely close to either zero or one outside this range). So
+with 200 points, we have a discretization error of approximately 0.01. This is quite high. To yield
+higher resolution and lower error, we switched to a least squares solution. 
+Thus we seek to find $U$ that minimizes
+\begin{equation}
+  ||MU-V||_2
+\end{equation}
+Its not clear that this will yield a good 1-norm for equation (\ref{eq:error_norm}), but in practice it does.
+Because the solution of this system is not very stable, especially in high dimensions, we actually
+minimize a regularized version
+\begin{equation}
+  ||MU-V||_2 + \lambda||U||_2
+\end{equation}
+the solution is
+\begin{equation}
+  U = (M^T M + \lambda I)^{-1}V
+\end{equation}
+with this approach, there is no guarantee that $U \geq 0$. However, we have some
+flexibility in the choice of $\sigma$ in equation (\ref{eq:conv1}). As we decrease
+the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount
+and is in fact the result of convolution with a gaussian whose variance is the difference.
+Thus as $\sigma$ decreases, $C(X)$ grows and approaches the derivative of a
+logistic function at $\sigma = 0$. It retains some very weak negative values for $\sigma > 0$
+but removal of those values leads to very small error.
 
@@
 \SetKwInOut{Output}{Output}
-
 \begin{algorithm}[t]
@@
 \label{alg:our_algorithm}
-%\vspace{-10pt}
 \end{algorithm}
 
-\section{Computing the Correction Distribution}\label{sec:correction}
-Our goal is to compute the distribution of the correction variable $X_{\rm corr}$ such that
-$X_{\rm norm} + X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \in \mathcal{N}(0,\sigma)$
-and $X_{\rm log}$ has a standard logistic CDF $(1+\exp(-X))^{-1}$.
-In the next section we show that the accuracy of the test depends on the absolute
-error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and the distribution of $X_{\rm log}$.
-So we minimize this in our construction.
-
-Our goal then is to minimize
-\begin{equation}
-  |\Phi_{\sigma} * C_{\sigma} - S|
-\end{equation}
-where $\Phi_{\sigma}(X) = \Phi(X/\sigma)$ and $\Phi$ is the standard
-normal CDF, $S(X)$ is the logistic function, $C_{\sigma}(X)$ is the
-{\em density} of the correction distribution, and $*$ denotes
-convolution. In continuous form, this value is
-\begin{equation}
-  \label{eq:conv1}
-  \sup_X\left|\int_{-V}^{V}\Phi_{\sigma}(X-Y) C_{\sigma}(Y) dY - S(X)\right|
-  \end{equation}
-Assuming $X_{\rm corr}$ lies in the interval $[-V,V]$. For computation of $C_{\sigma}(Y)$, we
-discretize $X$ and $Y$ into $4N+1$ and $2N+1$ values respectively, with $X \in [-2V,2V]$ and
-$Y \in [-V,V]$. Writing $X_i = i * V / N$ and $Y_j = j * V / N$, the condition in
-equation (\ref{eq:conv1}) can be written:
-\begin{equation}
-  \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|
-\end{equation}
-and defining a matrix $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, vectors $U_j = C_{\sigma}(Y_j)$ and $V_i = S(X_i)$, the
-above is equivalent to minimizing the 1-norm:
-\begin{equation}
-  \label{eq:error_norm}
-  ||MU-V||_1
-\end{equation}
-wrt $U$. We have an additional constraint that $U_j > 0$ for all $j$, since $U$ represents
-a density. We first explored optimizing this system with linear programming to find the $U$ such that:
-\begin{equation}
-  \begin{split}
-    & \min(\epsilon) \qquad \hbox{where} \\
-    & -\epsilon \leq MU - V \leq \epsilon \\
-    & U \geq 0
-  \end{split}
-\end{equation}
-which was tractable up to a few hundred dimensions. However, the discretization error is bounded by
-the curvature of the underlying functions which are here slightly less than one. i.e.
-the errors are of the order of $(V/N)^2$. Here we chose $V=20$ to provide adequate containment of
-the distributions (the CDFs are extremely close to either zero or one outside this range). So
-with 200 points, we have a discretization error of approximately 0.01. This is quite high. To yield
-higher resolution and lower error, we switched to a least squares solution. 
-Thus we seek to find $U$ that minimizes
-\begin{equation}
-  ||MU-V||_2
-\end{equation}
-Its not clear that this will yield a good 1-norm for equation (\ref{eq:error_norm}), but in practice it does.
-Because the solution of this system is not very stable, especially in high dimensions, we actually
-minimize a regularized version
-\begin{equation}
-  ||MU-V||_2 + \lambda||U||_2
-\end{equation}
-the solution is
-\begin{equation}
-  U = (M^T M + \lambda I)^{-1}V
-\end{equation}
-with this approach, there is no guarantee that $U \geq 0$. However, we have some
-flexibility in the choice of $\sigma$ in equation (\ref{eq:conv1}). As we decrease
-the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount
-and is in fact the result of convolution with a gaussian whose variance is the difference.
-Thus as $\sigma$ decreases, $C(X)$ grows and approaches the derivative of a
-logistic function at $\sigma = 0$. It retains some very weak negative values for $\sigma > 0$
-but removal of those values leads to very small error.
-\begin{table}[h]
+\begin{table}[t]
   \centering
@@
 
+{\color{blue} Daniel: I can help improve the table and describe it. Also, I'm
+not sure right now about where Algorithm~\ref{alg:our_algorithm} should be
+placed in the paper. Still TODO for me: clean up and/or revise the optimization
+problem formulations and solutions.}
+
 \section{Analysis}\label{sec:analysis}
 
-In this section we derive error bounds for our M-H test, and for the approximate target
-distribution that it generates. We discuss generation of the correction variable in the
-next section. To cut to the chase, it is possible to generate samples $X_{\rm corr}$ with
-a CDF error of at most single-precision floating point error, and possibly better. We
-therefore treat the $X_{\rm corr}$ variable as a sample from the exact correction distribution
-and we will not analyze its errors.
+We now derive error bounds for our M-H test, and for the approximate target
+distribution that it generates. From Section~\ref{sec:correction}, we know that
+it is possible to generate the correction samples $X_{\rm corr}$ with a CDF
+error of at most single-precision floating point error, and possibly better. We
+therefore treat the $X_{\rm corr}$ variable as a sample from the exact
+correction distribution and we will not analyze its errors.
 
@@
 \begin{equation}
-  \hat{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad   \sigma(\hat{X}) = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\hat{X})^2\right)^{\frac{1}{2}}
+    \hat{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad \sigma(\hat{X}) = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\hat{X})^2\right)^{\frac{1}{2}}
 \end{equation}
 then $t=\hat{X}/\sigma(\hat{X})$ has an approximate Student's distribution which approaches a normal
-distribution in the limit. Then from \cite{explicit-clt05}:
+distribution in the limit. Then from~\cite{explicit-clt05}:
 \begin{equation}\label{eq:clt-bounds}
@@
 error of the normal approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of variables
-$y = x \sigma(\hat{X})$, equation (\ref{eq:clt-bounds}) becomes
+$y = x \sigma(\hat{X})$, Equation~\ref{eq:clt-bounds} becomes
 \begin{equation}\label{eq:clt-bounds2}
@@
 \begin{corollary}
-  We can now substitute into equation (\ref{eq:clt-bounds2}) and displace by the mean, giving:
+We can now substitute into Equation~\ref{eq:clt-bounds2} and displace by the mean, giving:
 \begin{equation}\label{eq:clt-bounds3}
@@
 
-We have shown that the distribution of $\Delta^*$ approximates a
-Normal distribution with mean $\Delta$ and variance
-$\sigma^2(\Delta^*)$, and bounded its error with estimable
-quantities. Both $E|X|$ and $E|X|^3$ can be estimated as means of
-$|\Lambda_i - \Lambda|$ and $|\Lambda_i - \Lambda|^3$ respectively on each minibatch. We expect this
-will often be accurate enough on minibatches with hundreds or
-thousands of points, but otherwise bootstrap CIs can be computed from
-those sequences. Since the bounds are monotone in $E|X|$ and $E|X|^3$,
-using upper bootstrap CI limits will provide high-confidence error bounds.
-
-So as long as we know the first and third absolute moments $E|X|$ and $E|X|^3$, we can bound the
-error of the normal approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of variables
-$y = x \sigma(\hat{X})$, equation (\ref{eq:clt-bounds}) becomes
-\begin{equation}\label{eq:clt-bounds2}
-   \sup_y\left|{\rm Pr}(\hat{X}< y) - \Phi\left(\frac{y}{\sigma(\hat{X})}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
-\end{equation}
-which shows that the distribution of $\hat{X}$ approaches the normal distribution $\mathcal{N}(0,\sigma^2(\hat{X}))$
-whose variance is $\sigma^2(\hat{X})$, measured from the sample.
-
-\begin{lemma} Let $P(x)$ and $Q(x)$ be two cumulative distributions satisfying $\sup_x|P(x)-Q(x)|\leq \epsilon$ with
-  $x$ in some real range. Let $R(y)$ be the {\em density} of another
-  random variable $y$. Let $P'$ be the convolution $P*R$ and $Q'$ be
-  the convolution $Q*R$. Then $P'(z)$ (resp. $Q'(z)$) is the CDF of sum
-  $z=x+y$ of independent random variables $x$ with CDF $P(x)$ (resp. $Q(x)$) and y
-  with density $R(y)$.  Then
+We have shown that the distribution of $\Delta^*$ approximates a Normal
+distribution with mean $\Delta$ and variance $\sigma^2(\Delta^*)$, and bounded
+its error with estimable quantities. Both $E|X|$ and $E|X|^3$ can be estimated
+as means of $|\Lambda_i - \Lambda|$ and $|\Lambda_i - \Lambda|^3$ respectively
+on each minibatch. We expect this will often be accurate enough on minibatches
+with hundreds or thousands of points, but otherwise bootstrap CIs can be
+computed from those sequences. Since the bounds are monotone in $E|X|$ and
+$E|X|^3$, using upper bootstrap CI limits will provide high-confidence error
+bounds.
+
+{\color{blue} Daniel: TODO provide a better transition for this paragraph.}
+
+\begin{lemma}\label{lem:cdf_bounds}
+Let $P(x)$ and $Q(x)$ be two cumulative distributions satisfying
+$\sup_x|P(x)-Q(x)|\leq \epsilon$ with $x$ in some real range. Let $R(y)$ be the
+{\em density} of another random variable $y$. Let $P'$ be the convolution $P*R$
+and $Q'$ be the convolution $Q*R$. Then $P'(z)$ (resp. $Q'(z)$) is the CDF of
+sum $z=x+y$ of independent random variables $x$ with CDF $P(x)$ (resp. $Q(x)$)
+and y with density $R(y)$.  Then
 \begin{equation}
@@
 \end{lemma}
+
 \begin{proof}
-  \begin{equation}
-    \begin{split}
-      P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx
-    \end{split}
+We first observe that
+\begin{equation}\label{eq:pdf_difference}
+    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
 \end{equation}
@@
 \begin{equation}
-- \epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon
+-\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon
 \end{equation}
@@
 \begin{corollary}
-  If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})| \leq \epsilon(\theta,\theta',b)$,
-  then
+If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})|
+\leq \epsilon(\theta,\theta',b)$, then
 \begin{equation}
@@
 \end{corollary}
-%% \subsection{Discussions and Practical Considerations of the New MH Test}\label{ssec:discussion}
+
 \begin{proof}
-  Apply the previous lemma twice. First take:
+We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
 \begin{equation}
@@
 \end{equation}
-  and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$ where $\sigma_n^2
-  = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of $P$ and $Q$
+and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
+where $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of
+$P$ and $Q$
 \begin{equation}
@@
 \end{equation}
-  Both steps preserve the error bound
-  $\epsilon(\theta,\theta',b)$. Finally $S(y-\Delta)$ is a logistic
-  CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm Pr}(\Delta +
-  X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We
-  conclude that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm
-    corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
+Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
+$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
+Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
+that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
+norm}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
 log} > 0)$ by at most $\epsilon$.
-
 \end{proof}
@@
 \begin{equation}
-  \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1
-\end{equation}
-
-Bentkus et al.'s work on Edgeworth
-expansions \cite{Bentkus97} equation 2.2 gives:
-
-\begin{lemma} Let $X_1,\ldots,X_n$ be
-  a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3. If $X$ satisfies Cramer's condition, 
-then
-\begin{equation}\label{eq:clt-bounds}
+  \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1.
+\end{equation}
+
+Equation 2.2 in Bentkus et al.'s work on Edgeworth expansions~\cite{Bentkus97}
+provides the following:
+
+\begin{lemma}\label{lem:clt_edgeworth}
+Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
+random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3.
+If $X$ satisfies Cramer's condition, then
+\begin{equation}\label{eq:clt-bounds_edgeworth}
     \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
@@
 \begin{equation}
-G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x)
+    G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x).
 \end{equation}
 \end{lemma}
-This lemma shows that the average of the $X_i$ has a more precise, skewed CDF limit $G_n(x,y)$
-where the skew term has weight proportional to a certain measure of skew derived
-from the moments: $\frac{\mu_3}{b_2^{3/2}}$. Note that if the $X_i$ are symmetric,
-the weight of the correction term is zero, and the CDF of the average of the $X_i$ converges
+
+Lemma~\ref{lem:clt_edgeworth} shows that the average of the $X_i$ has a more
+precise, skewed CDF limit $G_n(x,y)$ where the skew term has weight proportional
+to a certain measure of skew derived from the moments:
+$\frac{\mu_3}{b_2^{3/2}}$. Note that if the $X_i$ are symmetric, the weight of
+the correction term is zero, and the CDF of the average of the $X_i$ converges
 to $\Phi(x)$ at a rate of $O(n^{-1})$.
 
-Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that
-decays as $n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where $\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
-\begin{equation}
-  \label{eq:GNderivatives}
+Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
+$n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where
+$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
+\begin{equation}\label{eq:GNderivatives}
     G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
 \end{equation}
-From which we see that $G_n(x,y)$ is a linear combination of
-$\Phi(x)$, $\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In
-Algorithm 1, we correct for the difference in $\sigma$ between
-$\Delta^*$ and the variance needed by $X_{\rm corr}$ using $X_{\rm
-  norm}$. This same method works when we wish to estimate the error in $\Delta^*$ vs
-$G_n(x,y)$. Since all of the component functions of $G_n(x,y)$ are
-derivatives of a (unit variance) $\Phi(x)$, adding a normal variable
-with variance $\sigma'$ increases the variance of all three functions
-to $1+\sigma'$. Thus we add $X_{\rm norm}$ as per Algorithm 1
-preserving the limit in equation (\ref{eq:GNderivatives}).
-
-The deconvolution approach can be used to constuct a correction
-variable $X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard
-logistic function. An additional complexity is that $G_n(x,y)$ has
-additional parameters $y$ and $n$. Since these act as a single
-multiplier $\frac{y}{6\sqrt{n}}$ in equation
-(\label{eq:GNderivatives}), its enough to consider a function
-$g(x,y')$ parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can
-be computed and saved offline. As we have shown above, errors in the
-``limit'' function propagate directly through as errors in the
-acceptance test.  To achieve a test error of say $1e-6$ (close to single
-floating point precision), we need a $y'$ spacing of $1e-6$. It
-should not be necessary to tabulate values all the way to $y'=1$,
-since $y'$ is scaled inversely by the square root of monibatch size.
-Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.
-Since our $x$ resolution is 10,000, this leads to a table with about
-1 billion values, which can comfortably be stored in memory. However,
-if $g(x,y)$ is moderately smooth in $y$, it should be possible to
-achieve similar accuracy with a much smaller table. We leave further
-analysis and experiments with $g(x,y)$ as future work.
+From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
+$\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In Algorithm 1, we
+correct for the difference in $\sigma$ between $\Delta^*$ and the variance
+needed by $X_{\rm corr}$ using $X_{\rm norm}$. This same method works when we
+wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
+component functions of $G_n(x,y)$ are derivatives of a (unit variance)
+$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
+variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm norm}$ as
+per Algorithm 1 preserving the limit in Equation~\ref{eq:GNderivatives}.
+
+The deconvolution approach can be used to construct a correction variable
+$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
+additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
+Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
+Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
+parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
+saved offline. As we have shown above, errors in the ``limit'' function
+propagate directly through as errors in the acceptance test.  To achieve a test
+error of say $1e-6$ (close to single floating point precision), we need a $y'$
+spacing of $1e-6$. It should not be necessary to tabulate values all the way to
+$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
+Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
+resolution is 10,000, this leads to a table with about 1 billion values, which
+can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
+in $y$, it should be possible to achieve similar accuracy with a much smaller
+table. We leave further analysis and experiments with $g(x,y)$ as future work.
 
@@
 %% smaller in absolute value, thus reducing variance. The flatter posterior results in easier mixing.
-
-%% \section{Analysis}\label{sec:theory}
-
-%% In this section, we first perform an error analysis for our new
-%% acceptance test. Second, we follow the implications of this analysis for accuracy of
-%% the estimated target distribution. Third, we compare the speed and accuracy of
-%% our test with prior work. Due to space constraints, all our proofs
-%% are in Appendix~\ref{app:proofs}.
-
-%% We first introduce some notation. Let our true (full dataset) and
-%% approximate (minibatch) acceptance probabilities be $P_a = g(\Delta) =
-%% \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} > 0)$.
-%% Define $X_{\xi} = \varepsilon - X_{\rm norm}$ which means, invoking
-%% Equations~\ref{eq:relationship} and~\ref{eq:deconvolution}, that
-%% $X_\xi$ is the random variable characterizing the accuracy of our
-%% $X_{\rm norm}$.  Thus, $\Delta' = \Delta + X_{\rm norm} + X_\xi$. We
-%% assume the standard deviation of $X_{\rm norm}$ is less than
-%% $\pi/\sqrt{3}$, so that we can use Equation~\ref{eq:deconvolution}.
-%% We denote $F_X(x)$ as the CDF of $X$.
-
-%% We use some terms from standard Markov chain theory~\cite{Meyn2009}. Let the \emph{transition
-%% kernels} of MCMC on complete and minibatch data at step $i$ be $P_i$ and $P_i'$, respectively. When
-%% these kernels (e.g., $P_i$) are applied on a probability distribution $D_1$, they generate a new
-%% distribution $D_2$, which we write as $P_i \circ D_1 = D_2$. Define $\pi$ to be the stationary
-%% distribution obtained by $P_i$, and $\pi_i'$ to be the stationary distribution obtained by $P_i'$
-%% ($\pi'$ needs a subscript in this case).  Finally, we indicate the \emph{total variation distance}
-%% between two kernels using $\| \cdot \|$.
-
-%% Our first result shows that the distance $\|P_i\circ D - P_i'\circ D\| \le 4\zeta \ell$ for all
-%% distributions $D$.
-
-%% % Our first result shows that the distance between $P_i$ and $P_i'$ is proportional to our estimation
-%% % error $X_{\xi}$.
-
-%% \begin{lemma}\label{lem:theory1}
-%% If $|X_\xi| < \zeta$ and $|\nabla F_X(x)| < \ell$ for all $x$, then $\|P_i\circ D-P_i'\circ D\| \le
-%% 4\zeta \ell \;\;\forall D$.
-%% \end{lemma}
-
-%% %% \begin{proof}
-%% %% See Appendix~\ref{app:theory1}.
-%% %% \end{proof}
-
-%% Thus, the accuracy of our transition kernel $P_i'$ is related to the quality of our estimated
-%% $X_{\rm norm}$. We next show that ${\rm Var}(\Delta')$ is roughly proportional to the step size of
-%% our proposer.
-
-%% \begin{lemma}\label{lem:theory2}
-%% For one step sampling from $\theta$ to $\theta'$, if the jumping step $\|\theta - \theta'\|_2 <
-%% \epsilon$, and the gradient of the log likelihood is bounded by a constant factor, i.e. $\|\nabla
-%% (\log \Pr(x_i\mid \theta))\|_2 < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2
-%% (m-\frac{m(m-1)}{N-1})$, where $m$ is the minibatch size, and $N$ is the total data size.
-%% \end{lemma}
-
-%% %% \begin{proof}
-%% %% See Appendix~\ref{app:theory2}.
-%% %% \end{proof}
-
-%% Since ${\rm Var}(\Delta') < 4\epsilon^2 k^2 (m - \frac{m(m-1)}{N-1}) $, we can assume the maximum
-%% value of $X_{\xi}$ is proportional to the standard deviation of $\Delta'$, i.e. $\epsilon$. Thus, we
-%% can assume $\zeta=\epsilon C$, where $C$ is a constant factor. We finally introduce
-%% Theorem~\ref{thm:theory3} to characterize the convergence of our samples.
-
-%% \begin{theorem}\label{thm:theory3}
-%% Assume $\pi$ satisfies $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0 - \pi\|$, where $\eta \in [0, 1)$ for
-%% all distributions $D_0$. Also assume there exists $0 < \rho_t < 1$ such that $\|P'_i \circ D -
-%% \pi_i'\| \leq 2\rho_i \;\; \forall D$. We can get $\| P_t' \circ P_{t-1}' \circ \cdots \circ P_1' \circ D_0 - \pi
-%% \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $, where
-%% $\alpha_i = \frac{\epsilon_i C }{1-\eta}$ and $\epsilon_i$ is the step size at iteration $i$.
-%% \end{theorem}
-
-%% %% \begin{proof}
-%% %% See Appendix~\ref{app:theory3}.
-%% %% \end{proof}
-
-%% Theorem~\ref{thm:theory3} indicates that the difference between the target and sampled distributions
-%% consists of two components: $\rho_i$, which is determined by the efficiency of $P'_i$, and the
-%% $\alpha_i$s, which are determined by the error of our transition kernels. These involve competing
-%% objectives: the larger the step size, the more efficient the $P_i'$ but its error increases.  The
-%% reverse happens with a step size too small.
-
-%% We explore the efficiency of our minibatch MH test.
-Theorem~\ref{thm:efficiency} indicates that our minibatch MH test only requires a constant number of instances for each step, however, recall the Lemma~\ref{lem:efficienty_cutting}, the adaptive minibatch MH test requires $O(\sqrt{N})$ number of data instance for each step.
-Thus, We show that our proposed minibatch MH test is much higher efficient than the known minibatch algorithms.
-
-%% \begin{theorem}\label{thm:efficiency}
-%% We require $\frac{ C^2 \rho^2 }{\sigma^6 \varepsilon^2}$ minibatch size to ensure $|P_a -P_a'| < \varepsilon$, where $C$ is a positive constant which is smaller than 0.4748, $\sigma^2 $ and $\rho$ is the second and third moment of $(\nabla\log \Pr(x_i|\theta) - \nabla\log\Pr(x_i|\theta')) - E\left[ \nabla\log \Pr(x_i|\theta) - \nabla\log\Pr(x_i|\theta')\right]$, $P_a$ is the accept rate from the complete data, $P_a'$ is the accept rate from minibatch data by our algorithm, $\theta$ is the current parameter, and $\theta'$ is the proposed parameter.
-
-%% \end{theorem}
 
@@
 
-It is useful to notice that, in ~\cite{icml2014c1_bardenet14}, they need to evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set at every iteration in order to calculate the error bound in equation (9) in ~\cite{icml2014c1_bardenet14}.
+It is useful to notice that, in~\cite{icml2014c1_bardenet14}, they need to evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set at every iteration in order to calculate the error bound in equation (9) in ~\cite{icml2014c1_bardenet14}.
 In Figure ~\ref{fig:gauss_bardenet}, the value of the first part and second part of equation (9) in ~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual minibatch data used.
@@
 
-%Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{fig:exp_3} to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability of $\rho/\sigma^3$.
-
-\iffalse
-\subsection{Neural Network Optimization}\label{ssec:nets}
-
-In this experiment, we apply our minibatch MH test with a Stochastic Gradient Hamiltonian
-Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the complete parameters for a fully
-connected neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}, which
-is a binary classification task with one million instances, each of which are 28-dimensional. The
-network we use has four layers: the first is the input, the last is the softmax, and the
-intermediate ones apply the sigmoid activation unit.  We also employ dropout and batch
-normalization~\cite{icml2015_ioffe15}.  In order to control ${\rm std}(\Delta')$, we initialize the
-temperature at 1000, and adjust it at iteration $i$ according to $T_i = \max\{1,
-1000/(i+1)^{0.5}\}$.
-
-For our comparison baseline, we train the same network architecture using the Adaptive Gradient
-Descent method~\cite{adapGrad}. Since the frequencies of the features are quite different, we use
-the same strategy as in Adaptive Gradient to rescale the gradient before applying SGHMC. We
-implement our MH test and SGHMCin the open-source BIDMach project~\cite{canny2013bidmach}.
-
-Figure~\ref{fig:nnet_fig} illustrates our experiment results and reveals that SGHMC with our
-minibatch MH test achieves higher log likelihood than the baseline. By comparing SGHMC with and
-without our minibatch MH test, we see that our MH test has limited benefit. Because we decrease the
-step size quickly, the acceptance rate of SGHMC is already high, so there is little need to test.
-% Daniel's note: Haoyu said he cannot increase the step size without it violating the variance
-% condition.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=1\linewidth]{exp3}
-    \caption{Log likelihood versus number of passes for optimization of neural network.}
-    \label{fig:nnet_fig}
-    \vspace{-10pt}
-\end{figure}
-
-\fi
+%Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{fig:exp_3}
+%to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability
+%of $\rho/\sigma^3$.
+
+%% \subsection{Neural Network Optimization}\label{ssec:nets}
+%% 
+%% In this experiment, we apply our minibatch MH test with a Stochastic Gradient Hamiltonian
+%% Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the complete parameters for a fully
+%% connected neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}, which
+%% is a binary classification task with one million instances, each of which are 28-dimensional. The
+%% network we use has four layers: the first is the input, the last is the softmax, and the
+%% intermediate ones apply the sigmoid activation unit.  We also employ dropout and batch
+%% normalization~\cite{icml2015_ioffe15}.  In order to control ${\rm std}(\Delta')$, we initialize the
+%% temperature at 1000, and adjust it at iteration $i$ according to $T_i = \max\{1,
+%% 1000/(i+1)^{0.5}\}$.
+%% 
+%% For our comparison baseline, we train the same network architecture using the Adaptive Gradient
+%% Descent method~\cite{adapGrad}. Since the frequencies of the features are quite different, we use
+%% the same strategy as in Adaptive Gradient to rescale the gradient before applying SGHMC. We
+%% implement our MH test and SGHMCin the open-source BIDMach project~\cite{canny2013bidmach}.  
+%% 
+%% Figure~\ref{fig:nnet_fig} illustrates our experiment results and reveals that SGHMC with our
+%% minibatch MH test achieves higher log likelihood than the baseline. By comparing SGHMC with and
+%% without our minibatch MH test, we see that our MH test has limited benefit. Because we decrease the
+%% step size quickly, the acceptance rate of SGHMC is already high, so there is little need to test.
+%% % Daniel's note: Haoyu said he cannot increase the step size without it violating the variance
+%% % condition.
+%% 
+%% \begin{figure}[t]
+%%     \centering
+%%     \includegraphics[width=1\linewidth]{exp3}
+%%     \caption{Log likelihood versus number of passes for optimization of neural network.}
+%%     \label{fig:nnet_fig}
+%%     \vspace{-10pt}
+%% \end{figure}
 
@@
 \begin{itemize}[noitemsep]
-  \item Proofs ommitted from the main text. 
+    \item The proof of Lemma~\ref{lem:worst_case}.
     \item More information on Section~\ref{ssec:gaussians}.
     \item More information on Section~\ref{ssec:logistic}.
-    \item More information on Section~\ref{ssec:nets}.
+    %\item More information on Section~\ref{ssec:nets}. % Daniel: this is in
+    %progress.
 \end{itemize}
 
-\section{Proofs}\label{app:proofs}
-
-{\bf lemma 1}
-{\em  For the model of section 2.1, there
-  exists a fixed (independent of $N$) constant $c$ such that with probability $ \geq c$ over the joint
-  distribution of $(\theta, \theta', u)$, the
-  tests from \cite{icml2014c1_bardenet14} and \cite{cutting_mh_2014} consume all $N$ samples. }
-
-\begin{proof} Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and
-  $\theta -0.5 \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events
-  occur with probability $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and $p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively,
-  and guarantee that $\Lambda^*(\theta',\theta)$ is negative, which guarantees that we can find
-  a $u$ so that $\psi(u,\theta',\theta)$ equals the expected value of $\Lambda^*(\theta',\theta)$.
-  Specifically, choose $u_0$
-  to satisfy
-  \begin{equation}
+\section{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}
+
+Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and $\theta -0.5
+\in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events occur with
+probability ({\color{blue} Daniel: sorry for interjecting here but don't we need
+to take into account the $1/\sqrt{N}$ in the probabilities? If $N \to \infty$
+(and it's going to be large with big datasets) the probabilities should be about
+zero.}) $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and
+$p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively, and guarantee that
+$\Lambda^*(\theta',\theta)$ is negative. This then guarantees that we can find a
+$u \in [0,1]$ so that $\psi(u,\theta',\theta)$ equals
+$\mathbb{E}[\Lambda^*(\theta',\theta)]$.  Specifically, choose $u_0$ to satisfy
+\begin{equation}\label{eq:log_uo}
     \log u_0 = N(\Lambda^*(\theta',\theta)-\psi(1,\theta',\theta))
 \end{equation}
-  which evaluates to 
-  \begin{equation}
-    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right) 
-  \end{equation}
-  and rewrite the test now as
+which evaluates ({\color{blue} Daniel: sorry again for interjecting but I really
+think $\log u_0$ does not have that $N$ there. Also, for the subsequent
+equation, I think we need to be using \emph{expected} values of $\Lambda^*$}) to
+\begin{equation}
+    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
+\end{equation}
+This means we can rewrite the acceptance test as
 \begin{equation}
@@
 \end{equation}
-  where $\not\approx$ means ``significantly different from'' under the distribution over samples of $x_i$. 
-  The above choice of $u_0$ ensures that the terms in parenthesis above sum to 0.5. Since the $x_i$ have mean 0.5, the
-  test will never correctly succeed.
-
-  Furthermore, if we sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be sufficiently different
+where $\not\approx$ means ``significantly different from'' under the
+distribution over samples of $x_i$.  The above choice of $u_0$ ensures that the
+terms in parenthesis above sum to 0.5. Since the $x_i$ have mean 0.5, the test
+will never correctly succeed.  Furthermore, if we sample values of $u$ near
+enough to $u_0$, the terms in parenthesis will not be sufficiently different
 from 0.5 to allow the test to succeed.
 
- The choices above for $\theta$ and $\theta'$ guarantee that
+{\color{blue} Daniel: I would like to ask you about these probabilities because
+I do not know how we got these intervals.} The choices above for $\theta$ and
+$\theta'$ guarantee that
 \begin{equation}
@@
 \end{equation}
-  so that the LHS has expected value zero.  Given our choice of intervals for the variables, the RHS is in the range $1/\sqrt{N}[-1,0.75]$.
-  The standard deviation of the LHS given the interval constraints is at least $0.5/\sqrt{b}$. And
-  so the gap between LHS and RHS is at most $2\sqrt{b/N}$ standard deviations.
-
-  The samples $\theta$, $(\theta'-\theta)$
-  and $u$ are drawn independently and so the probability of the conjunction of these events is
-  $c = p_0 p_1 p_2 = 0.0085$.
-\end{proof}
-
-%% \subsection{Proof of Lemma~\ref{lem:efficienty_cutting}}\label{app:lem_efficiency}
-
-%% If the minibatch size is $n$, we first compute the probability that more than $n$ data instances are required for adaptive sampling method for one iteration. We denote this probability $P_{rej}$.
-%% \[P_{rej} = \Pr \left ( 1-\phi_{n-1} \left( \frac{|\bar{l} - \frac{1}{N} \log (u\frac{p(\theta)q(\theta'|\theta)}{p(\theta')q(\theta|\theta')})|}{s} \right )  \geq \varepsilon \right ) \]
-
-%% where $s = \frac{s_l}{\sqrt{n}}\sqrt{1-\frac{n-1}{N-1}}$. \\
-%% We define random variable $Z = \frac{1}{N}\log (uQ)$, where $Q=\frac{p(\theta)q(\theta'|\theta)}{p(\theta')q(\theta|\theta')}$.
-Then $Z$ has probability density function $f_Z(z) = N \exp(zN), z \in (-\infty, \frac{1}{N} \log Q]$.
-
-%% \begin{align*} 
-%% P_{rej}&=  \Pr \left( 1-\varepsilon \geq \phi_{n-1} \left( \frac{|\bar{l} - Z|}{s_l \sqrt{\frac{N-n}{N-1}}} \sqrt{n} \right) \right) \\ 
-%% &=  \Pr \left(\phi^{-1}_{n-1} \left (1-\varepsilon \right) \frac{s_l }{\sqrt{n}} \sqrt{\frac{N-n}{N-1}} \geq |\bar{l} -Z| \right) = \Pr \left( |\bar{l} - Z| \leq \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \right) \\
-%% &= \Pr \left(Z - \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \leq \bar{l} \leq Z + \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \right) \\
-%% &= \int_{-\infty} ^{\frac{\log Q}{N}} N \exp (zN)dz \int_{z - \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}}} ^{z+ \frac{\eta s_l}{\sqrt{n}}\sqrt{\frac{N-n}{N-1}}} \frac{1}{2L} d\bar{l} = \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} \int_{-\infty} ^{\frac{\log Q}{N}} N \exp(zN) dz \\
-%% &= \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} Q
-%% \end{align*}
-%%  If we define the random variable $X$ the number of instance for one minibatch, we have $\Pr (X>n) = P_{rej} = \frac{\eta s_l}{L\sqrt{n}}\sqrt{\frac{N-n}{N-1}} Q, \mbox{ for } n \in [1,N], n \in Z^+$.
-Since $X$ is nonnegative, we have $E(X) \geq \Pr (X >x)x, \forall x$.
-If we choose $x=N/2$, we have $E(X) \geq \frac{\eta s_l}{L\sqrt{N/2}}\sqrt{\frac{N-N/2}{N-1}} Q \frac{N}{2} \geq \frac{\eta s_{l}Q}{2L}\sqrt{N}$.
-
-%% \subsection{Proof of Lemma~\ref{lem:theory1}}\label{app:theory1}
-
-%% We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta
-%% \ell$ for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
-%% \begin{align*}
-%% |P_a - P_a'| &= |\Pr(\Delta + X >0) - \Pr(\Delta' + X_{\rm corr} > 0)|\\
-%% & = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
-%% &= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| \\
-%% &= |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
-%% &= |\nabla F(-\Delta) \xi + o(\xi^2)|\\
-%% &\leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2\ell \xi,
-%% \end{align*}
-%% where we use Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have smaller absolute
-%% value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| < 2 \zeta \ell$.
-
-%% Next, we show the distance of new distributions obtained by applying $P_i$ and $P_i'$ on an
-%% arbitrary distribution $D$ once is bounded if we use the same proposer, i.e.  $\|P_i\circ D -
-%% P'_i\circ D\| \leq 4\zeta \ell \;\; \forall D$. We write $P_i(\theta \mid \theta') = P_a(\theta, \theta')
-%% q(\theta'\mid \theta) + (1-P_a(\theta,\theta')) \delta_D(\theta' - \theta)$, where $\delta_D$
-%% is the Dirac delta function. The total variation distance between two generated distributions is
-%% \begin{align*}
-%% \|P_i\circ D - P'_i\circ D\| &= \int_{\theta'} \left| \int_{\theta}((P_a-P_a') q(\theta'|\theta) + (1-P_a - 1+P_a') \delta_D(\theta -\theta')) dD(\theta) \right| d\theta' \\
-%% &= \int_{\theta'} \left| \int_{\theta} (P_a- P_a')(q(\theta'|\theta) - \delta_D(\theta - \theta')) dD(\theta) \right|  d\theta'\\ 
-%% &\leq 2 \zeta \ell \int_{\theta'} \left| \int_{\theta}(q(\theta'|\theta) + \delta_D(\theta - \theta')) dD(\theta) \right| d\theta'\\
-%% &\leq 2 \zeta \ell \int_{\theta'} ( D'(\theta') + D(\theta') ) d\theta'\\
-%% &= 4 \zeta \ell,
-%% \end{align*}
-%% where $D'$ is the distribution generated if we always accept the proposed point. (In the last line,
-%% we integrate two probability density functions, resulting in a value of two.) This matches our
-%% desired bound.
-
-%% \subsection{Proof of Lemma~\ref{lem:theory2}}\label{app:theory2}
-
-%% From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
-%% \theta))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
-%% $g_i(\theta) = \log p(x_i\mid \theta)$, and $q_i = g_i(\theta') - g_i(\theta)$. Then by Taylor's Theorem:
-%% \[
-%% |q| = |g(\theta') - g(\theta)| = |\nabla g(\theta)^T(\theta' - \theta) + o(\theta' - \theta)^2| \leq 2\epsilon k.
-%% \]
-%% Thus, given $\theta$ and $\theta'$, the variance of $\Delta'$ can be written as:
-%% \begin{align*}
-%% {\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^m g_i(\theta') - g_i(\theta)\right] = {\rm Var}\left(\sum_{i=1}^m q_i\right) \\
-%% &= m {\rm Var} (q_i) + m(m-1) {\rm Cov}_{i \neq j}(q_i,q_j) \\
-%% &= m{\rm Var}(q_i) - \frac{m(m-1)}{N-1} {\rm Var}(q_i) \\
-%% &\leq 4\left(m - \frac{m(m-1)}{N-1}\right)k^2 \epsilon^2,
-%% \end{align*}
-%% as desired.
-
-%% \subsection{Proof of Theorem~\ref{thm:theory3}}\label{app:theory3}
-
-%% First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|P_i \circ D_0 - \pi\| \leq \eta \|D_0
-%% - \pi\|$ and $\|P_i\circ D-P'_i\circ D\| \leq \epsilon_i C$ for all $D$. The norm between the
-%% stationary distributions is $\|\pi - \pi_i'\|\leq \frac{\epsilon_i C}{1-\eta}$.
-
-%% Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have 
-%% \[
-%% \| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t \| \leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t
-%% \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s,
-%% \]
-%% where $\alpha = \frac{\epsilon_i C}{1-\eta}$.  Thus, we get:
-%% \begin{align*}
-%%  \| P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_0 \| &\leq \|P_t' \circ \cdots \circ P_1' \circ D_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
-%%  &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
-%% \end{align*}
-%% as desired.
-
-%% \subsection{Proof of Theorem~\ref{thm:efficiency}}\label{app:theory4}
-%% Since only the $\frac{N}{nT}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta) \right)$ term brings randomness into $\Delta'$.
-Define $X_i = \frac{N}{T}\left(\log p(x_i|\theta') - \log p (x_i|theta_t) \right)$, $\tilde{X_i} = X_i - E(X_i)$, and $Y_n = \frac{\sum_{i=1}^n \tilde{X_i}}{n}$ (i.e.
-$Y_n$ is the $\Delta' - E(\Delta')$).
-From Berry-Esseen theorem, $|F_n(x) - \Phi(x)| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$, where $\Phi()$ is the CDF of standard normal distribution, $F_n()$ is the CDF of $Y_n$, $C$ is the constant parameter which is below 0.4748, $\rho = E(|\tilde{X_i}|^3)$, and $\sigma^2 = E(\tilde{X_i}^2)$.
-Since $\Delta'$ is an unbiased estimator of $\Delta$, the CDFs between $\Delta'$ and $\Delta + X_{\rm norm}$ has the same max gap (i.e.
-$\frac{C\rho}{\sigma^3\sqrt{n}}$).
-
-%% From Equation~\ref{eq:criteria}, we show we want to use $\Delta' + X_{\rm corr}$ to approximate $\Delta + X_{\rm norm} + X_{\rm corr}$.
-Since $|F_{\Delta'} - F_{\Delta + X_{\rm norm}}| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$, for the worst case, assume $\exists \xi, ~{\rm s.t.} |f_{\Delta'}(\xi) - f_{\Delta + X_{\rm norm}} (\xi)| = \max_x |f_{\Delta'}(x) - f_{\Delta + X_{\rm norm}}(x)| \overset{\Delta}{=} \eta $, where $f()$ represents the PDF function.
-Because PDF of $X_{\rm corr}$ has the positive value and unit area, we have:
-%% \begin{align*} 
-%% & |(f_{\Delta'}*f_{X_{\rm corr}})(\xi) - (f_{\Delta+X_{\rm norm}}*f_{X_{\rm corr}})(\xi)| \\
-%% & = |\int_{-\infty}^{\infty}f_{\Delta'}(\xi-\tau)f_{X_{\rm corr}}(\tau)d\tau - \int_{-\infty}^{\infty}f_{\Delta + X_{\rm norm}} (\xi -\tau) f_{X_{\rm corr}} (\tau) d\tau| \\
-%% & = |\int_{-\infty}^{\infty} (f_{\Delta'}(\xi-\tau) - f_{\Delta + X_{\rm norm}} (\xi -\tau)) f_{X_{\rm corr}} (\tau) d\tau| \\
-%% & \leq \int_{-\infty}^{\infty} |f_{\Delta'}(\xi-\tau) - f_{\Delta + X_{\rm norm}} (\xi -\tau) |f_{X_{\rm corr}} (\tau) d\tau  \\
-%% & \leq \int_{-\infty}^{\infty} \eta f_{X_{\rm corr}} (\tau) d\tau \leq \eta
-%% \end{align*}
-
-%% where ``*'' represents the convolution operation.
-
-%% Thus, after convolution with $X_{\rm corr}$, the maximum gap between PDF of the $\Delta'$ and $\Delta + X_{\rm norm}$ decreases.
-Thus, $|F_{\Delta' + X_{\rm corr}} - F_{\Delta + X_{\rm norm} + X_{\rm corr}}| \leq |F_{\Delta'} - F_{\Delta + X_{\rm norm}}| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$.
-Thus, $|P_a - P_a'| = |\Pr (\Delta + X_{\rm norm} + X_{\rm corr} > 0) - \Pr (\Delta' + X_{\rm corr} > 0)|=|F_{\Delta + X_{\rm norm} + X_{\rm corr}}(0) - F_{\Delta' + X_{\rm corr}}(0)| \leq \frac{C\rho}{\sigma^3\sqrt{n}}$.
-We can choose $n$ so that $\frac{C\rho}{\sigma^3\sqrt{n}} < \varepsilon$ to ensure $|P_a - P_a'| < \varepsilon$.
-Thus, the requirement for $n$ is $n \geq \frac{C^2\rho^2}{\sigma^6 \varepsilon^2}$.
+so that the LHS has expected value zero.  Given our choice of intervals for the
+variables, the RHS is in the range $1/\sqrt{N}[-1,0.75]$.  The standard
+deviation of the LHS given the interval constraints is at least $0.5/\sqrt{b}$.
+And so the gap between LHS and RHS is at most $2\sqrt{b/N}$ standard deviations.
+
+The samples $\theta$, $(\theta'-\theta)$ and $u$ are drawn independently and so
+the probability of the conjunction of these events is $c = p_0 p_1 p_2 =
+0.0085$.
 
@@
 \fi
+
 % Daniel: here are example LaTeX codes for figures and tables if we want to use them.

### 2016-09-22 nips_2016.tex
More changes to Section 4, I am understanding Johns original intentions better

@@
 \newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
+\DeclareMathOperator*{\argmin}{arg\,min}
 \newtheorem{lemma}{Lemma}
@@
 
-Here we derive a new test which incorporates the variability in
-minibatch statistics as {\em a natural part of the test}. Because of
-this, the amount of data required for each test is fixed with high
-probability and in expected value. We use a Barker test function
-\cite{Barker65} rather than a Metropolis test, which makes our test
-naturally error tolerant. The idea of using a noise-tolerant test
-using Barker's test function was suggested but not explored
-empirically in \cite{Bardenet15} section 6.3. But the asymptotic test
-statistic CDF and the Barker function are different, which leads to
-fixed errors for the approach in \cite{Bardenet15}. Here we show that
-the difference between the distributions can be corrected with an
-additive random variable. This leads to a test which is fast, and
-whose error can be made arbitrarily small.
-
-Our test is applicable when the variance (over data samples) of the
-log acceptance probability is small enough (less than 1). Its not
-clear at first why this quantity should be so bounded. But we will see
-that it is ``natural'' for well-specified models running
-Metroplis-Hastings sampling with optimal proposals
-\cite{OptimalScaling01} on a full dataset. When we reduce the amount
-of data for the test, the variance goes up, and we have to reduce
-it in one of several ways. Either:
+Here we derive a new test which incorporates the variability in minibatch
+statistics as {\em a natural part of the test}. Because of this, the amount of
+data required for each test is fixed with high probability and in expected
+value. {\color{blue} Daniel: we really need to show the amount of data is fixed
+with ``high probability'' (the results suggest ``OK probability'').} We use a
+Barker test function~\cite{Barker65} rather than a Metropolis test, which makes
+our test naturally error tolerant. The idea of using a noise-tolerant test using
+Barker's test function was suggested but not explored empirically
+in~\cite{Bardenet15} section 6.3. But the asymptotic test statistic CDF and the
+Barker function are different, which leads to fixed errors for the approach
+in~\cite{Bardenet15}. Here we show that the difference between the distributions
+can be corrected with an additive random variable. This leads to a test which is
+fast, and whose error can be made arbitrarily small.
+
+Our test is applicable when the variance (over data samples) of the log
+acceptance probability is small enough (less than 1). Its not clear at first why
+this quantity should be bounded, but we will show that it is ``natural'' for
+well-specified models running Metropolis-Hastings sampling with optimal
+proposals~\cite{OptimalScaling01} on a full dataset. When we reduce the amount
+of data for the test, the variance goes up, and we have to reduce it in one of
+several ways. Either:
 
@@
 billions of samples. Simply sampling from this distribution is one
-application, but an efficient proposal distribution
-\cite{OptimalScaling01} has similar variance to the target
+application, but an efficient proposal distribution~\cite{OptimalScaling01} has similar variance to the target
 distribution and will diffuse to it extremely slowly from an
@@
 our variance constraint implies that we have to trade-off proposal
-step size $s$ and batch size $b$ ($b \propto p^2$). i.e. we can make
+step size $s$ and batch size $b$ ($b \propto p^2$), i.e. we can make
 many small steps, or one large step, with a given batch of data. One
@@
 small steps. For applications to neural networks or other models where
-the posterior is multimodel, posterior inference is arguably a search
+the posterior is multimodal, posterior inference is arguably a search
 process. Covering the search space densely with small steps is much
@@
 proposal has mean $\theta'-\theta$? I thought it would have a mean of $\theta$
-because that is the current sample and the Gaussian is centered at that point?}
+because that is the current sample and the Gaussian is centered at that point?
+That will also clarify my confusion about the $\psi$ function.}
 
@@
 compared to full-dataset tests, so their speed-up is at most $1/c$.
+{\color{blue} Daniel: doesn't our method also have this issue? In general, it
+seems like any test that requires minibatch information to show deviation from
+the full batch version will suffer from this problem. Why not just force us to
+stop if we use too much data?}
 
@@
 
-{\color{blue} Daniel: cleaned up this section, deleted redundant delta equation,
-deleted a duplicate equation in Lemma~\ref{lem:detailed_balance}.}
-
 \subsection{Log-Likelihood Ratios}\label{ssec:log_likelihood_ratios}
@@
 \subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
-For our new MH test we use the Barker logistic function:
-$g(s)=(1+\exp(-s))^{-1}$ \cite{Barker65}. Straightforward arithmetic shows that it
-satisfies the condition in Lemma~\ref{lem:detailed_balance}.  While it
-is slightly less efficient than the Metropolis test when used on the
-full dataset, we will see soon that its smoothness allows it to
-naturally tolerate substantial variance in its input argument. This in
-turn will lead to a much more efficient test on subsets of data.
+For our new MH test we use the Barker logistic~\cite{Barker65} function:
+$g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the
+condition in Lemma~\ref{lem:detailed_balance}.  While it is slightly less
+efficient than the Metropolis test when used on the full dataset, we will show
+that its smoothness allows it to naturally tolerate substantial variance in its
+input argument. This in turn will lead to a much more efficient test on subsets
+of data. {\color{blue} Daniel: aside from the experiments, where do we show it
+``naturally tolerates substantial variance'' in the input?}
 
 Assume we begin with the current sample $\theta$ and a candidate sample
-$\theta'$. Let $V$ be a uniform random variable $V \sim \mathcal{U}(0,1)$. We
+$\theta'$, and that $V \sim \mathcal{U}(0,1)$ is a uniform random variable. We
 accept $\theta'$ if $g(\Delta(\theta,\theta')) > V$, and reject otherwise.
@@
 
-We now move to acceptance testing using the minibatch estimator
-$\Delta^*(\theta,\theta')$. From equation (\ref{eq:delta3}),
+We now describe acceptance testing using the minibatch estimator
+$\Delta^*(\theta,\theta')$. From Equation~\ref{eq:delta3},
 $\Delta^*(\theta,\theta')$ can be represented as a constant term plus the mean
@@
 $\Delta^*(\theta,\theta')$ therefore has a distribution which approaches a
-normal distribution by the central limit theorem. Later we will provide specific
-bounds between the CDF of $\Delta^*(\theta,\theta')$ and a Normal CDF, but for
-now we use an asymptotic argument.
+normal distribution by the Central Limit Theorem. We now describe this using an
+asymptotic argument and defer specific bounds between the CDFs of
+$\Delta^*(\theta,\theta')$ and a Gaussian to Section~\ref{sec:analysis}.
 
@@
 test in Equation~\ref{eq:the_exact_test} we want $\Delta + X$ for a logistic
-random $X$ (call it $X_{\rm log}$ from now on). In \cite{Bardenet15} it was
-proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed error
-caused by this approximation.
+random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{Bardenet15} it
+was proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed
+error caused by this approximation.
 
@@
 \end{equation}
-where we assume $X_{\rm norm}$ has zero-mean, Normal distribution with
-variance $\sigma^2$, and $X_{\rm corr}$ is a zero-mean ``correction''
-variable with density $C_{\sigma}(X)$.  The two variables are added
-(i.e., their distributions convolve) to form $X_{\rm log}$. It will be
-convenient to express this relationship in terms of the CDFs $N_{\sigma}(X_{\rm
-  norm}) = \Phi(X_{\rm norm}/\sigma)$ and $S(X_{\rm log})$, where
-$\Phi(.)$ is the Normal CDF with zero mean and unit variance and
-$S(.)$ is the standard logistic function.
-
-Equation (\ref{eq:deconvolution}) implies that $S = N_{\sigma} * C_{\sigma}$ where $*$ denotes convolution. We will later
-derive a numerical representation of $C$ by deconvolution, and using this we can generate
-samples of $X_{\rm corr}$.
-
-The acceptance test is now:
+where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and $X_{\rm corr}$
+is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The two
+variables are added (i.e., their distributions convolve) to form $X_{\rm log}$.
+This decomposition requires an appropriate $X_{\rm corr}$ distribution. We show
+in Section~\ref{sec:correction} that it is possible to use deconvolution to
+derive a highly accurate representation of the density $C_\sigma(X)$.
+Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
+$X_{\rm corr}$ distributions that sum to $X_{\rm log}$.
+
+Using $X_{\rm corr}$ samples from $C_\sigma(X)$, the acceptance test is now
 \begin{equation}\label{eq:criteria}
@@
 \end{equation}
-i.e. given an estimate $\Delta^*$ from the current data minibatch (assuming its
-variance is small enough) we test acceptance by adding a random variable $X_{\rm corr}$ and
-then accept $\theta'$ if the result is positive.
-
-If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0, \sigma^2(\Delta^*))$,
-the above test is exact as well. And as we will see later,  if there is a maximum error $\epsilon$ between
-the CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0, \sigma^2(\Delta^*))$,
-then the acceptance test has an error of at most $\epsilon$.
+Therefore, assuming the variance of $\Delta^*$ is small enough, if we have an
+estimate of $\Delta^*$ from the current data minibatch, we test acceptance by
+adding a random variable $X_{\rm corr}$ and then accept $\theta'$ if the result
+is positive (and reject otherwise).
+
+If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0,
+\sigma^2(\Delta^*))$, the above test is exact as well, and as we show in
+Section~\ref{sec:analysis}, if there is a maximum error $\epsilon$ between the
+CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0,
+\sigma^2(\Delta^*))$, then the acceptance test has an error of at most
+$\epsilon$ relative to the full batch version. {\color{blue} Daniel: this
+subsection is improved from earlier. My main concern now is that $X_{\rm norm}$
+is described as being both approximately Gaussian $\bar{\mathcal{N}}$ and
+Gaussian $\mathcal{N}$ here. We may want to clear that up; in our NIPS
+submission we actually used a new variable $\varepsilon$ for that and assigned
+it to be the $\mathcal{N}$ distribution.}
 
@@
     \caption{
-    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
-    standard logistic distribution. We use three standard deviation values of $X_{\rm norm}$. The
-    two red curves convolve to form the logistic, etc. The $y$-axis is capped at 0.5 for
-    readability. This figure must be viewed in color.
+    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
+    convolve to form the standard logistic distribution. We use three standard
+    deviation values of $X_{\rm norm}$. The two red curves convolve to form the
+    logistic, etc. The $y$-axis is capped at 0.5 for readability. This figure
+    must be viewed in color. {\color{blue} Daniel: TODO Change figure to use
+    LaTeX for the titles.}
     }
@@
 
-%% It is important to understand why we use Equations~\ref{eq:deconvolution} and~\ref{eq:criteria}. The
-%% deconvolution process allows us to use $\Delta' + X_{\rm corr}$ as a proxy for $\Delta + X$. It
-%% works because if we know the distribution\footnote{Computing the distributions of $\varepsilon$ and
-%% $X_{\rm norm}$ reduces to finding their standard deviations.} of $\varepsilon$ (from
-%% Equation~\ref{eq:relationship}), then we can pretend that our $\Delta'$ is in fact $\Delta$. To
-%% ``insert'' $\varepsilon$ into our test, we deconvolve $X$ so that one of its components, $X_{\rm
-%% norm}$, plays the part of $\varepsilon$. In light of our revised MH test, the previous discussion
-%% raises the question: in an iteration of MCMC with current $\theta$ and proposed $\theta'$, how do
-%% we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample the
-%% \emph{values} of $X_{\rm corr}$ appropriately. This procedure consists of three major steps.
-
-%% \begin{enumerate}[noitemsep]
-%%     \item We first need to estimate the standard deviation of $X_{\rm norm}$ (i.e.
-the standard deviation of $\Delta'$).
-We use one minibatch data to complete this task.
-From the proof of Lemma~\ref{lem:gaussian}, we know the randomness comes from $\frac{N}{n}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta) \right)$.
-Thus, ${\rm var}(\Delta') = \frac{N^2}{n^2}\sum_{i=1}^n{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right ) = \frac{N^2}{n}{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$.
-We use data instances in one minibatch to estimate the ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$.
-Since there are $n$ data instances in one minibatch, the estimation of ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$ has high accuracy.
-In our experiments (see Section~\ref{sec:experiments}) we also use the complete data to obtain the population variance of $\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$, and use the nonparametric Bootstrap to estimate the ${\rm std(std}(\Delta'))$, so that we can construct the confidence interval of ${\rm std}(\Delta')$ to reveal the accuracy of our estimation.
-
-%% 	\item If the ${\rm std}(\Delta') > 1.2$, the decomposition of $X$ (i.e.
-Equation~\ref{eq:deconvolution}) is unstable.
-We consume more minibatches to decrease the ${\rm std}(\Delta')$.
-We update the value of $\Delta'$ by $\Delta' = (k\Delta'_{\rm old} + \Delta'_{\rm new}) / (k+1)$, where $\Delta'_{\rm old}$ is the existing $\Delta'$ which uses $k$ minibatches to computed (usually $k=1$), and $\Delta'_{\rm new}$ is the $\Delta'$ computed from the new minibatch.
-We also update the value of ${\rm std}(\Delta')$ by ${\rm std}(\Delta') = \sqrt{\left ( k^2 {\rm var}(\Delta'_{\rm old}) + {\rm var}(\Delta'_{\rm new}) \right ) / (k+1)^2}$.
-Thus, we can decrease the ${\rm std}(\Delta')$ by consuming more data until the ${\rm std}(\Delta') > 1.2$.
-With help of temperature (see Section~\ref{ssec:discussion}), we will show that usually the mean of minibatch consumed per iteration is below 2 in Section~\ref{sec:experiments}.
-Later our efficiency analysis in Section~\ref{sec:theory} will reveal that the number of data instance consumed by our algorithm is much fewer than that of existing algorithms.
-
-%%     \item Once we have ${\rm std}(\Delta')$, we can determine the distribution of $X_{\rm corr}$
-%%     from standard deconvolution techniques. These rely on the well-known fact that the Fourier
-%%     transform of a convolution is the product of the individual Fourier transforms. In practice, we
-%%     pre-compute distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to
-%%     facilitate the sampling process.
-%% \end{enumerate}
-
 \section{Computing the Correction Distribution}\label{sec:correction}
 
-Our goal is to compute the distribution of the correction variable $X_{\rm corr}$ such that
-$X_{\rm norm} + X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \in \mathcal{N}(0,\sigma)$
-and $X_{\rm log}$ has a standard logistic CDF $(1+\exp(-X))^{-1}$.
-In the next section we show that the accuracy of the test depends on the absolute
-error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and the distribution of $X_{\rm log}$.
-So we minimize this in our construction.
-
-Our goal then is to minimize
-\begin{equation}
-  |\Phi_{\sigma} * C_{\sigma} - S|
-\end{equation}
-where $\Phi_{\sigma}(X) = \Phi(X/\sigma)$ and $\Phi$ is the standard
-normal CDF, $S(X)$ is the logistic function, $C_{\sigma}(X)$ is the
-{\em density} of the correction distribution, and $*$ denotes
-convolution. In continuous form, this value is
-\begin{equation}
-  \label{eq:conv1}
-  \sup_X\left|\int_{-V}^{V}\Phi_{\sigma}(X-Y) C_{\sigma}(Y) dY - S(X)\right|
-  \end{equation}
-Assuming $X_{\rm corr}$ lies in the interval $[-V,V]$. For computation of $C_{\sigma}(Y)$, we
-discretize $X$ and $Y$ into $4N+1$ and $2N+1$ values respectively, with $X \in [-2V,2V]$ and
-$Y \in [-V,V]$. Writing $X_i = i * V / N$ and $Y_j = j * V / N$, the condition in
-equation (\ref{eq:conv1}) can be written:
-\begin{equation}
-  \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|
-\end{equation}
-and defining a matrix $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, vectors $U_j = C_{\sigma}(Y_j)$ and $V_i = S(X_i)$, the
-above is equivalent to minimizing the 1-norm:
-\begin{equation}
-  \label{eq:error_norm}
-  ||MU-V||_1
-\end{equation}
-wrt $U$. We have an additional constraint that $U_j > 0$ for all $j$, since $U$ represents
-a density. We first explored optimizing this system with linear programming to find the $U$ such that:
-\begin{equation}
-  \begin{split}
-    & \min(\epsilon) \qquad \hbox{where} \\
-    & -\epsilon \leq MU - V \leq \epsilon \\
-    & U \geq 0
-  \end{split}
-\end{equation}
-which was tractable up to a few hundred dimensions. However, the discretization error is bounded by
-the curvature of the underlying functions which are here slightly less than one. i.e.
-the errors are of the order of $(V/N)^2$. Here we chose $V=20$ to provide adequate containment of
-the distributions (the CDFs are extremely close to either zero or one outside this range). So
-with 200 points, we have a discretization error of approximately 0.01. This is quite high. To yield
-higher resolution and lower error, we switched to a least squares solution. 
-Thus we seek to find $U$ that minimizes
-\begin{equation}
-  ||MU-V||_2
-\end{equation}
-Its not clear that this will yield a good 1-norm for equation (\ref{eq:error_norm}), but in practice it does.
-Because the solution of this system is not very stable, especially in high dimensions, we actually
-minimize a regularized version
-\begin{equation}
-  ||MU-V||_2 + \lambda||U||_2
-\end{equation}
-the solution is
-\begin{equation}
-  U = (M^T M + \lambda I)^{-1}V
-\end{equation}
-with this approach, there is no guarantee that $U \geq 0$. However, we have some
-flexibility in the choice of $\sigma$ in equation (\ref{eq:conv1}). As we decrease
-the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount
-and is in fact the result of convolution with a gaussian whose variance is the difference.
-Thus as $\sigma$ decreases, $C(X)$ grows and approaches the derivative of a
-logistic function at $\sigma = 0$. It retains some very weak negative values for $\sigma > 0$
-but removal of those values leads to very small error.
+Our proposed test in Equation~\ref{eq:criteria} requires knowing the
+distribution of the correction variable $X_{\rm corr}$ such that $X_{\rm norm} +
+X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \sim \mathcal{N}(0,\sigma^2)$
+and $X_{\rm log}$ has a standard logistic CDF, $(1+\exp(-X))^{-1}$. In
+Section~\ref{sec:analysis}, we show that the accuracy of the test depends on the
+absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and $X_{\rm
+log}$. Consequently, we need to minimize this in our construction of $X_{\rm
+corr}$. More formally, let $\Phi_\sigma(X) = \Phi(X/\sigma)$ where $\Phi$ is the
+standard normal CDF\footnote{Hence, $\Phi_\sigma$ is the CDF of a zero-mean
+Gaussian with variance $\sigma$.}, $S(X)$ be the logistic function, and
+$C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
+distribution. Our goal, based on Equation~\ref{eq:deconvolution}, is to solve
+the following optimization problem:
+\begin{equation}\label{eq:overall_corr_problem}
+    C_\sigma^* = \argmin_{C_\sigma} |\Phi_{\sigma} * C_{\sigma} - S|
+\end{equation}
+where $*$ denotes convolution. {\color{blue} Daniel: This is very unusual use of
+convolutions (i.e., CDF * DENSITY = CDF). It's valid but I was very surprised to
+know that.} For computation of $C_\sigma$, we assume that its input $Y$ and
+another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
+{\color{blue} Daniel: I am a little confused at the different interval sizes.}
+In continuous form, the optimization problem is therefore
+\begin{equation}\label{eq:conv1}
+    C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|
+\end{equation}
+To get $C_\sigma^*$ in a practical form, we discretize $X$ and $Y$ into $4N+1$
+and $2N+1$ values respectively. Writing $X_i = i*V/N$ and $Y_j = j*V/N$, the
+condition in Equation~\ref{eq:conv1} can be re-expressed as:
+\begin{equation}\label{eq:conv2}
+    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
+\end{equation}
+Defining a matrix $M$ and vectors $u$ and $v$ such that $M_{ij} =
+\Phi_{\sigma}(X_i-Y_j)$, $u_j = C_{\sigma}(Y_j)$, and $v_i = S(X_i)$,
+Equation~\ref{eq:conv2} is equivalent to minimizing the $L_1$ norm:
+\begin{equation}\label{eq:optimization_l1}
+    u^* = \argmin_{u} \|Mu-v\|_1.
+\end{equation}
+{\color{blue} Daniel: Equation~\ref{eq:optimization_l1} is finding
+the $C_\sigma$ minimizing the $L_1$ norm, but the previous one,
+Equation~\ref{eq:conv2}, is finding the $C_\sigma$ that is minimizing the
+$L_\infty$ norm, i.e., the maximum value! If we want this to be the $L_1$ norm
+(and I think we do) then Equation~\ref{eq:conv2} needs an outer sum over the $i$
+terms. In addition, Equation~\ref{eq:conv1} also needs to be modified, perhaps
+eliminate the ``sup'' but keep $x \in [-2V,2V]$}
+
+We have the additional constraint that $u_j > 0$ for all $j$, since $u$
+represents a density. We first explored optimizing this problem with linear
+programming to find a $u$ minimizing $|Mu - v|$ subject to $u \ge 0$. This was
+tractable up to a few hundred dimensions for $u$. However, the discretization
+error is bounded by the curvature of the underlying functions which are here
+slightly less than one, i.e.  the errors are of the order of $(V/N)^2$.
+{\color{blue} Daniel: unfortunately, I do not understand what this means (i.e.,
+the curvature and how the $(V/N)^2$ is obtained).} Here we chose $V=20$ to
+provide adequate containment of the distributions (the CDFs are extremely close
+to either zero or one outside this range). So with 200 points, we have a
+discretization error of approximately 0.01, which is relatively large.  To yield
+higher resolution and lower error, we switched to a least squares solution:
+\begin{equation}\label{eq:optimization_l2}
+    u^* = \arg_u \min \; \|Mu-v\|_2^2 + \lambda \|u\|_2^2,
+\end{equation}
+where we add the standard regularization factor $\lambda$ for stability in high
+dimensions. The solution is well-known from the normal equations: $u^* = (M^TM +
+\lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_1$ norm for
+Equation~\ref{eq:optimization_l1}.
+
+With this approach, there is no guarantee that $u^* \geq 0$. However, we have
+some flexibility in the choice of $\sigma$ in Equation~\ref{eq:conv1}.
+{\color{blue} Daniel: is this the correct equation?} As we decrease the variance
+of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount and
+is in fact the result of convolution with a Gaussian whose variance is the
+difference.  Thus as $\sigma$ decreases, $C(X)$ grows and approaches the
+derivative of a logistic function at $\sigma = 0$. It retains some very weak
+negative values for $\sigma > 0$ but removal of those values leads to very small
+error.
 
@@
     }
-    Draw $X_{\rm norm} \in \mathcal{N}(0,1-\sigma^2(\Delta*))$ and
+    Draw $X_{\rm norm} \sim \mathcal{N}(0,1-\sigma^2(\Delta*))$ and
     $X_{\rm corr}$ from the correction distribution $C_1(X)$\;
@@
 }
-\caption{A description of M-H sampling with our acceptance test.}
+\caption{A description of M-H sampling with our acceptance test. {\color{blue}
+Daniel: I think we may want to be careful about the use of $C_1(X)$ because that
+implies that we have a fixed $X_{\rm corr}$ density, but that $C_1(X)$ will
+change depending on the iteration and variance of $\Delta^*$, right? Hopefully
+not by much, but it still changes sometimes?}}
 \label{alg:our_algorithm}
@@
 \begin{table}[t]
+    \caption{\color{blue} Daniel: TODO need a good title.}
+    \label{tab:xcorr}
     \centering
-  \begin{tabular}{|l|l|l|l|}
-    \hline
-     N & $\sigma$ & $\lambda$ & L1 error  \\
-        \hline
+    \begin{tabular}{l l l l}
+    \toprule
+    $N$ & $\sigma$ & $\lambda$ & $L_1$ error  \\
+    \midrule
     4000 & 0.9 & 1 & 1.0e-4  \\
     4000 & 0.8 & 0.03 & 5.0e-6 \\
-     \hline
+    \bottomrule
     \end{tabular}
@@
 
-{\color{blue} Daniel: I can help improve the table and describe it. Also, I'm
-not sure right now about where Algorithm~\ref{alg:our_algorithm} should be
-placed in the paper. Still TODO for me: clean up and/or revise the optimization
-problem formulations and solutions.}
+{\color{blue} Daniel: TODO improve/describe Table~\ref{tab:xcorr} and try to
+relocate Algorithm~\ref{alg:our_algorithm} to be somewhere nicer.}
 

### 2016-09-22 nips_2016.tex
Some cleanup to Section 5.

@@
 To get $C_\sigma^*$ in a practical form, we discretize $X$ and $Y$ into $4N+1$
-and $2N+1$ values respectively. Writing $X_i = i*V/N$ and $Y_j = j*V/N$, the
+and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}$ and $j \in \{-N,
+\ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
 condition in Equation~\ref{eq:conv1} can be re-expressed as:
@@
 \end{equation}
-Defining a matrix $M$ and vectors $u$ and $v$ such that $M_{ij} =
-\Phi_{\sigma}(X_i-Y_j)$, $u_j = C_{\sigma}(Y_j)$, and $v_i = S(X_i)$,
+To make the problem easier to understand, we can define a matrix $M$ and vectors
+$u$ and $v$ such that $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, $u_j =
+C_{\sigma}(Y_j)$, and $v_i = S(X_i)$, where the indices $i$ and $j$ are
+appropriately translated to be non-negative indices for $M, u,$ and $v$. Thus,
 Equation~\ref{eq:conv2} is equivalent to minimizing the $L_1$ norm:
@@
     }
-    Draw $X_{\rm norm} \sim \mathcal{N}(0,1-\sigma^2(\Delta*))$ and
+    Draw $X_{\rm norm} \sim \mathcal{N}(0,1-\sigma^2(\Delta^*))$ and
     $X_{\rm corr}$ from the correction distribution $C_1(X)$\;
@@
 it is possible to generate the correction samples $X_{\rm corr}$ with a CDF
-error of at most single-precision floating point error, and possibly better. We
-therefore treat the $X_{\rm corr}$ variable as a sample from the exact
-correction distribution and we will not analyze its errors.
-
-In the most similar prior works, \cite{cutting_mh_2014} uses asymptotic arguments based
-on the central limit theorem to argue that its approximate acceptance test error tends to
-zero as batch size increases. But no quantitative bounds are given. In
-\cite{icml2014c1_bardenet14}, explicit bounds are given, but they depend on a bound on
-the quantity:
-\begin{equation}
-  \max_{1\leq i\leq N}|\log p(x_i|\theta') - \log p(x_i|\theta)|
-\end{equation}
-while such bounds can be derived easily for simple models, its unclear how to derive them
-for a complex model such as a neural network.
-
-In this paper, we rely on quantitave forms of the central limit theorem which rely on measurable
-statistics from the minibatch $\{x_i\}$. Thus a sampler using our approach does not need to ``see''
-data beyond the current minibatch. This supports use of the sampler on very large datasets, and
-in online mode where the dataset is presented as a stream.
-
-The outline of this section is as follows: we first present bounds on the absolute
-and relative error (in CDF) of the distribution of $\Delta^*$ vs a Normal distribution. We then show that
-these errors bounds are preserved after the addition of other random variables, in particular
-$X_{\rm norm}$ and $X_{\rm corr}$. From this it follows that the acceptance test has the same error bound.
-
-The quantitative central-limit result below comes from \cite{explicit-clt05}:
+error of at most single-precision floating point error, and possibly better
+{\color{blue} (Daniel: I can fix this to cite Table~\ref{tab:xcorr}'s values,
+that will probably be better.)}. We therefore treat the $X_{\rm corr}$ variable
+as a sample from the exact correction distribution and we will not analyze its
+errors.
+
+In the most similar prior works,~\cite{cutting_mh_2014} uses asymptotic
+arguments based on the Central Limit Theorem to argue that its approximate
+acceptance test error tends to zero as batch size increases. But no quantitative
+bounds are given. In~\cite{icml2014c1_bardenet14}, explicit bounds are given,
+but they depend on bounding
+\begin{equation}\label{eq:bad_bound}
+    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i\mid\theta') - \log p(x_i\mid\theta)|.
+\end{equation}
+Such bounds can be derived efficiently for models such as logistic regression,
+but it is unclear how to derive them for a complex model such as a neural
+network. In general, since we obtain a new $\theta'$ value each iteration, one
+would need to iterate through all the $p(x_i\mid \theta')$ terms, defeating the
+purpose of minibatch MCMC.
+
+In this paper, we use quantitative forms of the Central Limit Theorem which rely
+on measurable statistics from a single minibatch. Thus a sampler using our
+approach does not need to ``see'' data beyond the current minibatch. This
+supports use of the sampler on very large datasets, and in online mode where the
+dataset is presented as a stream.
+
+For the analysis, in Section~\ref{ssec:delta_star_distribution}, we first
+present bounds on the absolute and relative error (in terms of the CDFs) of the
+distribution of $\Delta^*$ vs a Gaussian. We then show in
+Section~\ref{ssec:preserve_bounds} that these error bounds are preserved after
+the addition of other random variables, in particular $X_{\rm norm}$ and $X_{\rm
+corr}$. From this it follows that the acceptance test has the same error bound.
+
+\subsection{Bounding the Distribution of $\Delta^*$ Versus a Gaussian}\label{ssec:delta_star_distribution}
+
+To start, we use the following quantitative central-limit result (from
+\cite{explicit-clt05}):
 
 \begin{lemma}\label{lem:quant_clt}
-Let $X_1,\ldots,X_n$ be
-a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and
-variance $\sigma^2(\hat{X})$ where:
-\begin{equation}
-    \hat{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad \sigma(\hat{X}) = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\hat{X})^2\right)^{\frac{1}{2}}
-\end{equation}
-then $t=\hat{X}/\sigma(\hat{X})$ has an approximate Student's distribution which approaches a normal
-distribution in the limit. Then from~\cite{explicit-clt05}:
+Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
+random variables with sample mean $\hat{X}$ and variance $\sigma^2(\hat{X})$
+where:
+\begin{equation}
+    \hat{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad \sigma(\hat{X}) = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\hat{X})^2\right)^{\frac{1}{2}}.
+\end{equation}
+This means $t=\hat{X}/\sigma(\hat{X})$ has an approximate Student's distribution
+which approaches a normal distribution in the limit. Then
+from~\cite{explicit-clt05}:
 \begin{equation}\label{eq:clt-bounds}
-    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
+    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}.
 \end{equation}
@@
 
-So as long as we know the first and third absolute moments $E|X|$ and $E|X|^3$, we can bound the
-error of the normal approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of variables
-$y = x \sigma(\hat{X})$, Equation~\ref{eq:clt-bounds} becomes
+{\color{blue} Daniel: the notation here was (and still is, to a lesser extent)
+the source of considerable confusion for me. Do we want $\sigma(X)$ or
+$\sigma(\hat{X})$? I think in Lemma~\ref{lem:quant_clt} we want the former, not
+the latter. OH ALSO, this section uses $n$ as the full data size, which means
+before this section, we should use $n$, not $N$, for the full data size, right?
+The Bardenet paper uses $n$ to represent the full data size.}
+
+Lemma~\ref{lem:quant_clt} demonstrates that as long as we know the first and
+third absolute moments $E|X|$ and $E|X|^3$, we can bound the error of the normal
+approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of
+variables $y = x \sigma(\hat{X})$, Equation~\ref{eq:clt-bounds} becomes
 \begin{equation}\label{eq:clt-bounds2}
-   \sup_y\left|{\rm Pr}(\hat{X}< y) - \Phi\left(\frac{y}{\sigma(\hat{X})}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
-\end{equation}
-which shows that the distribution of $\hat{X}$ approaches the normal distribution $\mathcal{N}(0,\sigma^2(\hat{X}))$
-whose variance is $\sigma^2(\hat{X})$, measured from the sample.
-
-To apply this to our test, let $X_i = \Lambda_i(\theta,\theta') - \Lambda(\theta,\theta')$,
-then the $X_i$ are zero-mean, IID variables.
-Take $n=b$ the minibatch size, and then:
-\begin{equation}
-    \hat{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta') 
-\end{equation}
-so that $\sigma(\hat{X}) = \sigma(\Delta^*)$.
-\begin{corollary}
+   \sup_y\left|{\rm Pr}(\hat{X}< y) - \Phi\left(\frac{y}{\sigma(\hat{X})}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}},
+\end{equation}
+showing that the distribution of $\hat{X}$ approaches the normal distribution
+$\mathcal{N}(0,\sigma^2(\hat{X}))$ whose variance is $\sigma^2(\hat{X})$,
+measured from the sample.
+
+To apply this to our test, let $X_i = \Lambda_i(\theta,\theta') -
+\Lambda(\theta,\theta')$, so that the $X_i$ are zero-mean, i.i.d. variables. If
+instead of all $n$ samples, we only extract a subset of $b$ samples
+corresponding to our minibatch, we can connect $\hat{X}$ with our $\Delta^*$
+term:
+\begin{equation}\label{eq:x-hat}
+    \hat{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta'),
+\end{equation}
+so that $\sigma(\hat{X}) = \sigma(\Delta^*)$. This results in the following
+Corollary:
+
+\begin{corollary}\label{cor:our_bound_delta_prime}
 We can now substitute into Equation~\ref{eq:clt-bounds2} and displace by the mean, giving:
@@
 
-We have shown that the distribution of $\Delta^*$ approximates a Normal
-distribution with mean $\Delta$ and variance $\sigma^2(\Delta^*)$, and bounded
-its error with estimable quantities. Both $E|X|$ and $E|X|^3$ can be estimated
-as means of $|\Lambda_i - \Lambda|$ and $|\Lambda_i - \Lambda|^3$ respectively
-on each minibatch. We expect this will often be accurate enough on minibatches
-with hundreds or thousands of points, but otherwise bootstrap CIs can be
-computed from those sequences. Since the bounds are monotone in $E|X|$ and
-$E|X|^3$, using upper bootstrap CI limits will provide high-confidence error
-bounds.
-
-{\color{blue} Daniel: TODO provide a better transition for this paragraph.}
+Corollary~\ref{cor:our_bound_delta_prime} shows that the distribution of
+$\Delta^*$ approximates a Normal distribution with mean $\Delta$ and variance
+$\sigma^2(\Delta^*)$. Furthermore, it bounds the error with \emph{estimable
+quantities}: both $E|X|$ and $E|X|^3$ can be determined as means of $|\Lambda_i
+- \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each minibatch. We
+expect this will often be accurate enough on minibatches with hundreds or
+thousands of points, but otherwise bootstrap CIs can be computed from those
+sequences. Since the bounds are monotone in $E|X|$ and $E|X|^3$, using upper
+bootstrap CI limits will provide high-confidence error bounds.
+
+\subsection{Bounds are Preserved After Adding Random Variables}\label{ssec:preserve_bounds}
+
+We next relate the CDFs of distributions and show that bounds are preserved
+after adding random variables.
 
@@
 
-\begin{corollary}
+Relating Lemma~\ref{lem:cdf_bounds} to our acceptance test, we observe that:
+
+\begin{corollary}\label{cor:bounds_preserved}
 If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})|
@@
 \begin{equation}
-    P(y) = Pr(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
+    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
 \end{equation}
@@
 \begin{equation}
-    P'(y) = Pr(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
+    P'(y) = {\rm Pr}(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
 \end{equation}
@@
 \begin{equation}
-    P''(y) = Pr(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
+    P''(y) = {\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
 \end{equation}
@@
 
-In fact we can do better than this approximation (showing the error decreases as $O(n^{-1})$) by using a more
-precise limit distribution under an additional assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$
-denote the $i^{th}$ absolute moment of $X$. If Cramer's condition holds, i.e. if
-\begin{equation}
-  \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1.
-\end{equation}
-
-Equation 2.2 in Bentkus et al.'s work on Edgeworth expansions~\cite{Bentkus97}
-provides the following:
+As our discussion in Section~\ref{sec:our_algorithm} showed, as the distribution
+of $\Delta^*$ approaches a Gaussian, our new MH test becomes more accurate.
+Corollary~\ref{cor:bounds_preserved} shows that the bounds from
+Section~\ref{ssec:delta_star_distribution} are preserved after the addition of
+the random variables we use, showing that our test should remain accurate.
+
+{\color{blue} Daniel: Do we want to include this new analysis? There seem to be
+two different ways we show the bound. The first part (above) is nice and I like
+it (and it seems to show all we need for Algorithm 1, doesn't it?}
+
+In fact we can do better than this approximation (showing the error decreases as
+$O(n^{-1})$) by using a more precise limit distribution under an additional
+assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$ denote the
+$i^{th}$ absolute moment of $X$. If Cramer's condition holds:
+\begin{equation}\label{eq:cramers_condition}
+    \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1,
+\end{equation}
+then Equation 2.2 in Bentkus et al.'s work on Edgeworth
+expansions~\cite{Bentkus97} provides the following Lemma:
 
@@
 
-We conduct two sets of experiments to explore the benefits of our minibatch MH test and to
-benchmark it with previous work. In Section~\ref{ssec:gaussians}, we show that our test enables
-samples to converge to the posterior distribution of a heated Gaussian mixture model. In
-Section~\ref{ssec:logistic}, we analyze its efficiency on logistic regression. 
-%In %Section~\ref{ssec:nets}, we apply the test for deep learning. 
-Appendices~\ref{app:gaussian},~\ref{app:logistic} %and~\ref{app:nnet} 
-contain more detailed information on these respective experiments.
+We conduct two sets of experiments to explore the benefits of our minibatch MH
+test and to benchmark it with previous work. In Section~\ref{ssec:gaussians}, we
+show that our test enables samples to converge to the posterior distribution of
+a heated Gaussian mixture model. In Section~\ref{ssec:logistic}, we analyze its
+efficiency on logistic regression.  Appendices~\ref{app:gaussian}
+and~\ref{app:logistic} contain more detailed information on these respective
+experiments.
 

### 2016-09-23 nips_2016.tex

@@
 compared to full-dataset tests, so their speed-up is at most $1/c$.
-{\color{blue} Daniel: doesn't our method also have this issue? In general, it
-seems like any test that requires minibatch information to show deviation from
-the full batch version will suffer from this problem. Why not just force us to
-stop if we use too much data?}
+
+{\color{blue} Daniel: quick question: I think we should explain why our method
+\emph{doesn't} run into this issue. I think it is because we rely on an
+approximation to $\Delta + X_{\rm log} > 0$ and that test does not require any
+``difference between an approximation and an exact version'' because it is
+\emph{already} an exact version, assuming the approximation is enough.
+
+Also, in terms of this example, if we are in the special case with using a lot
+of data, why not force early stopping?}
 
@@
 for a logistic random variable $X$. {\color{blue} Daniel: I think the text
-before Equation~\ref{eq:equivalent_test} should be similar to what we had in
-our NIPS submission because I find it much clearer.}
+before Equation~\ref{eq:the_exact_test} should be similar to what we had in our
+NIPS submission because I find it much clearer.}
 
@@
 \end{equation}
-where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and $X_{\rm corr}$
-is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The two
-variables are added (i.e., their distributions convolve) to form $X_{\rm log}$.
-This decomposition requires an appropriate $X_{\rm corr}$ distribution. We show
-in Section~\ref{sec:correction} that it is possible to use deconvolution to
-derive a highly accurate representation of the density $C_\sigma(X)$.
+where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and that $X_{\rm
+corr}$ is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The
+two variables are added (i.e., their distributions convolve) to form $X_{\rm
+log}$.  This decomposition requires an appropriate $X_{\rm corr}$ distribution.
+We show in Section~\ref{sec:correction} that it is possible to use deconvolution
+to derive a highly accurate representation of $C_\sigma(X)$.
 Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
-$X_{\rm corr}$ distributions that sum to $X_{\rm log}$.
+$X_{\rm corr}$ densities that sum to $X_{\rm log}$.
 
@@
 \sigma^2(\Delta^*))$, then the acceptance test has an error of at most
-$\epsilon$ relative to the full batch version. {\color{blue} Daniel: this
-subsection is improved from earlier. My main concern now is that $X_{\rm norm}$
-is described as being both approximately Gaussian $\bar{\mathcal{N}}$ and
-Gaussian $\mathcal{N}$ here. We may want to clear that up; in our NIPS
-submission we actually used a new variable $\varepsilon$ for that and assigned
-it to be the $\mathcal{N}$ distribution.}
+$\epsilon$ relative to the full batch version.
+
+{\color{blue} Daniel: this subsection is improved from earlier. My main concern
+now is that $X_{\rm norm}$ is described as being both approximately Gaussian
+$\bar{\mathcal{N}}$ and Gaussian $\mathcal{N}$ here. We may want to clear that
+up; in our NIPS submission we actually used a new variable $\varepsilon$ for
+that and assigned it to be the $\mathcal{N}$ distribution. My current
+suggestion is to have $\Delta^*$ be the $\bar{\mathcal{N}}$ approx. normal
+distribution (but with mean $\Delta$, of course), and keep $X_{\rm norm}$ as an
+exact Gaussian.}
 
@@
     Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $\sigma^2(\Delta^*)$\;
-    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from equation (\ref{eq:clt-bounds3})\;
+    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Equation~\ref{eq:clt-bounds3}\;
     \While {$\sigma^2(\Delta^*) \ge 1$ {\bf or} $\epsilon > \delta$}{
@@
 change depending on the iteration and variance of $\Delta^*$, right? Hopefully
-not by much, but it still changes sometimes?}}
+not by much, but it still changes sometimes? ALSO bring up Xinlei's old points
+about $X_{\rm norm}$ confusion.}}
 \label{alg:our_algorithm}
@@
 but it is unclear how to derive them for a complex model such as a neural
-network. In general, since we obtain a new $\theta'$ value each iteration, one
+network. In general, since a new $\theta'$ value is obtained each iteration, one
 would need to iterate through all the $p(x_i\mid \theta')$ terms, defeating the
@@
 before this section, we should use $n$, not $N$, for the full data size, right?
-The Bardenet paper uses $n$ to represent the full data size.}
+The Bardenet paper uses $n$ to represent the full data size.
+
+PS: I'd recommend not putting sample mean and sample variance in an equation if
+the $\sigma(\hat{X})$ gets clarified, since those computations are common
+knowledge.}
 
@@
 \end{proof}
+
+{\color{blue} Daniel: Hold on, the following Lemma seems to add $X_{\rm norm}$
+to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
+didn't the previous section already justify that $\Delta^* \approx \Delta +
+X_{\rm norm}$? That is, I don't think we need to add another $X_{\rm norm}$; I
+think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
+match what we wrote earlier.}
 

### 2016-09-23 nips_2016.tex
Updated Section 6.1

@@
 \begin{equation}\label{eq:optimization_l2}
-    u^* = \arg_u \min \; \|Mu-v\|_2^2 + \lambda \|u\|_2^2,
+    u^* = \argmin_u\; \|Mu-v\|_2^2 + \lambda \|u\|_2^2,
 \end{equation}
@@
 
-To start, we use the following quantitative central-limit result (from
-\cite{explicit-clt05}):
+{\color{blue} Daniel: the following lemma is, I think, from Corollary 2
+in~\cite{explicit-clt05} but I cannot derive the exact lemma as we have it
+here.} To start, we use the following quantitative central-limit result
+(from~\cite{explicit-clt05}):
 
@@
 the $\sigma(\hat{X})$ gets clarified, since those computations are common
-knowledge.}
+knowledge.
+
+PPS: The sample mean is probably more commonly denoted as $\bar{X}$.}
 
@@
 \end{equation}
-so that $\sigma(\hat{X}) = \sigma(\Delta^*)$. This results in the following
-Corollary:
+so that $\sigma(\hat{X}) = \sigma(\Delta^*)$. {\color{blue} Daniel: This time it
+makes sense to use $\sigma(\hat{X})$ instead of $\sigma(X)$.} This results in
+the following Corollary:
 
@@
 \end{equation}
-and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
-where $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of
-$P$ and $Q$
+and convolve with the distribution of $X_n$ {\color{blue} (Daniel: to be clear I
+think this was supposed to be $X_{\rm norm}$ but see comments earlier about why
+I think it is not necessary.)} which has density $\phi(X/\sigma_n)$ where
+$\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of $P$ and
+$Q$
 \begin{equation}
@@
 then Equation 2.2 in Bentkus et al.'s work on Edgeworth
-expansions~\cite{Bentkus97} provides the following Lemma:
+expansions~\cite{Bentkus97} provides the following:
 
@@
 Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
-$n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where
+$n^{1/2}$. {\color{blue} (Daniel: how did the equation in the next sentence
+appear? I verified it and it seems correct, but it looks like it popped out of
+nowhere.)} Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where
 $\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
@@
 table. We leave further analysis and experiments with $g(x,y)$ as future work.
-
-%% \textbf{The Convolution}. Figure~\ref{fig:deconvolution} demonstrates
-%% Equation~\ref{eq:deconvolution} and provides three (color-coded) examples of densities for $X_{\rm
-%% norm}$ and $X_{\rm corr}$. The density of $X$, a logistic random variable with mean $\mu = 0$ and
-%% scale $s=1$, is known, as are the $X_{\rm norm}$ densities for different standard deviations. The
-%% deconvolution provides us with the $X_{\rm corr}$ distributions. Note, however, that as the standard
-%% deviation of $X_{\rm norm}$ increases, the $X_{\rm corr}$ distribution becomes increasingly unstable
-%% and ``bumpier'', because the logistic curve has fatter tails than Gaussians. In addition, we cap
-%% the standard deviation possibilities of $X_{\rm norm}$ at $\approx 1.2$ in our experiments. In our
-%% experiments, we discretize the possible standard deviations of $X_{\rm norm}$, and we also limit the
-%% densities to consider the range $[-10,10]$ as shown in Figure~\ref{fig:deconvolution}.
-
-%% \textbf{Variance Preconditioning}. Returning to the discussion from Section~\ref{sec:introduction},
-%% our test requires a variance check. Equation~\ref{eq:deconvolution} means that the variance of $X$,
-%% which is $\pi^2/3\approx 3.29$ is an upper bound on the variances of $X_{\rm norm}$. Therefore, the
-%% test cannot work if the variance of $X_{\rm norm}$ is too large. The standard deviation of $X$
-%% is about $\sqrt{3.29}\approx 1.81$, so this bounds ${\rm std}(X_{\rm norm})$ (as well as ${\rm
-%% std}(\Delta')$). From the discussion above, we bound the standard deviation at 1.2 so that $X_{\rm
-%% corr}$ can handle the remaining noise. Recall from Section~\ref{ssec:deltas_minibatch} that we
-%% estimate the ${\rm std}(X_{\rm norm})$ each iteration. If the variance/standard deviation test
-%% fails, then we can (a) skip the iteration, (b) increase the minibatch size, (c) change the proposal
-%% to decrease step sizes, or (d) increase temperature, as we now discuss.
-
-%% \textbf{Temperature}. One way to satisfy the variance precondition is to increase the temperature of
-%% the target distribution. For a temperature $T>1$, the augmented target distribution would become
-%% \begin{equation}\label{eq:log_temperature}
-%% \log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
-%% \end{equation}
-%% with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
-%% smaller in absolute value, thus reducing variance. The flatter posterior results in easier mixing.
+{\color{blue} Daniel: I will look at the code to understand this paragraph better.}
 
@@
 
-We start with a simple Gaussian mixture model, borrowing an experiment from~\cite{langevin_2011}.
-The parameter is 2-D, $\theta = (\theta_1,\theta_2)$, and the parameter/data generation process is
+We start with a simple Gaussian mixture model, borrowing an experiment
+from~\cite{langevin_2011}.  The parameter is 2-D, $\theta =
+(\theta_1,\theta_2)$, and the parameter/data generation process is
 \begin{equation}\label{eq:data_generation}
@@
 \end{equation}
-We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta = (0,1)$, we draw 1,000,000
-data points so that the target distribution is $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with
-the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This
-results in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we apply a
-temperature $T=10,000$ to reduce ${\rm std}(\Delta')$. Taking logs, we get the target as shown in the
-far left of Figure~\ref{fig:gauss_mix_1}.
-
-We run MCMC with our MH test using minibatch size $m=100$. %We also run this using standard
-%full-batch MCMC (``standard MCMC'') with the standard test from Equation~\ref{eq:traditional}, 
-We also run this using the method from~\cite{cutting_mh_2014} (``Cut MH'')  and the method from 
-~\cite{icml2014c1_bardenet14}. For the method in \cite{cutting_mh_2014}, we use $m=100$ and
-increment minibatches by that amount within a test if necessary.
-As for the method in ~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the minibatch size geometrically with a ratio of $\gamma = 1.5$.
-The tolerance for making a
-decision in \cite{cutting_mh_2014} is $\epsilon=0.005$, and the error bound control parameter in ~\cite{icml2014c1_bardenet14} is $p = 2, \delta = 0.01$.
-To make comparisons easier, all three use the same random walk proposer with covariance $\Sigma={\rm diag}(0.3, 0.3)$.
-This is a poor proposer, but it is sufficient for
-our purposes as the quality of the samples will be primarily due to the MH test. All methods are run
-5000 times to collect 5000 samples.
+We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta =
+(0,1)$, we draw 1,000,000 data points so that the target distribution is
+$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with the prior based on the
+$\theta$ generation process in Equation~\ref{eq:data_generation}. This results
+in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we
+apply a temperature $T=10,000$ to reduce ${\rm std}(\Delta')$. Taking logs, we
+get the target as shown in the far left of Figure~\ref{fig:gauss_mix_1}.
+{\color{blue} Daniel: the ${\rm std}(\Delta')$ stuff is outdated; let's figure
+out a way to change this. We may also have to justify such a high temperature,
+because this is far from the ``real'' (log) posterior.}
+
+We run MCMC with our MH test using minibatch size $m=100$.  We also run this
+using the methods from Korattikara 2014~\cite{cutting_mh_2014} and
+Bardenet 2014~\cite{icml2014c1_bardenet14}. For~\cite{cutting_mh_2014}, we use
+$m=100$ and increment minibatches by that amount within a test if necessary, and
+for~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the minibatch size
+geometrically with a ratio of $\gamma = 1.5$. The tolerance for making a
+decision in~\cite{cutting_mh_2014} is $\epsilon=0.005$, and the error bound
+control parameters in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta =
+0.01$. To make comparisons easier, all three use the same random walk proposer
+with covariance $\Sigma={\rm diag}(0.3, 0.3)$. This is a poor proposer, but it
+is sufficient for our purposes as the quality of the samples will be primarily
+due to the MH test. All methods are run 5000 times to collect 5000 samples.
 
@@
     \centering
-    % \includegraphics[width=1\linewidth]{cloud_v01.png}
     \includegraphics[width=1\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
@@
     \label{fig:gauss_mix_1}
-    %\vspace{-10pt}
 \end{figure}
@@
     \centering
-   % \includegraphics[width=1\linewidth]{adaptive_and_ours_information_v01.png}
     \includegraphics[width=1\linewidth]{GaussianMixtureResult/minibatch_size_gaussian.png}
     \caption{
-    Histograms of minibatch sizes %and ${\rm std}(\Delta')$ values, along with ``${\rm std}(\Delta')$
-   % violations'' (right).
+    Histograms of minibatch sizes for the three methods used in
+    Section~\ref{ssec:gaussians}. {\color{blue} Daniel: The y-axis should
+    probably just be ``Counts'' not ``Total Counts.'' Also, was this run with
+    the \texttt{plt.tight\_layout()} command? The y-axis text for the second and
+    third subplots almost overlap with the graphs to their left. Finally, I am
+    still thinking about whether the y-axis should be on a log scale. I think
+    the log scale is appropriate for the x-axis, it's a question of whether we
+    should also apply it for the y-axis.}
     }
     \label{fig:gauss_mix_2}
-    %\vspace{-10pt}
 \end{figure}
@@
 	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
-	\caption{Value of first and second part of equation (9) in ~\cite{icml2014c1_bardenet14} with respect to actual minibatch size}
+	\caption{Value of first and second part of equation (9) in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size}
 	\label{fig:gauss_bardenet}
@@
 
-%\begin{figure}[t]
-    %\centering
-    %\includegraphics[width=1\linewidth]{CI_v01.png}
-   % \caption{
-   % Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
-    %}
-   % \label{fig:gauss_mix_3}
-    %\vspace{-10pt}
-%\end{figure}
-
-Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$ samples for the four
-methods, with darker regions indicating a greater density of points. The three methods obtain the
-same rough form of the posterior, so our MH test can indeed result in the same posterior as the
-other two methods.
-%However, the fourth method (i.e.
-just treat the $\Delta'$ which comes from minibatch as $\Delta$ which comes from the complete data) reveals wider spread distribution than the posterior.
-From Figure~\ref{fig:gauss_mix_1} we illustrate our proposed method can obtain the samples which have almost the same distribution as the standard MH test from the complete data, adaptive minibatch test, and the true posterior.
-Meanwhile, we show the benefits of our minibatch method compared with the naive minibatch method.
-Actually, based on a measure of divergence from the ground-truth distribution, our method is better than the other three methods in terms of this measurement.
-The measurement is to calculate the poisson likelihood of the MCMC-generated parameters over a parameter range, with $\theta_1 \in (-1.5, 2.5)$ and $\theta_2 \in (-2.5, 2.5)$.
-The log likelihood value for our method, cut MH method and adaptive MH method are -2144.6, -4007.4, -2598.4, respectively.
-
-Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used by
-the three methods in each iteration.
-It is obvious that our methods consumes significantly less data, and most minibatch sizes are smaller than 1000, with an average minibatch size of 420.
-The other two methods occasionally consumes a large proportion of the entire data set, and the average minibatch sizes are 16378, 16755 for cut MH method, adaptive MH method, respectively.
-In terms of average minibatch size, our method is nearly 39 times faster the cut MH method, and nearly 40 times faster than the adaptive MH method.
-
-It is useful to notice that, in~\cite{icml2014c1_bardenet14}, they need to evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set at every iteration in order to calculate the error bound in equation (9) in ~\cite{icml2014c1_bardenet14}.
-In Figure ~\ref{fig:gauss_bardenet}, the value of the first part and second part of equation (9) in ~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual minibatch data used.
-The second part which contains  the $C_{\theta,\theta'}$ term is larger than the first part, which means that the second part is non-negligible and has to be calculated at every iteration.
-This is one of the very time-consuming parts in this method.
-
-%Usually, the method can make a decision with 200 samples.
-Occasionally, however, it must use all 10000 points, resulting in lots of wasted computation.
-For this particular experiment, that happened eight times.
-%The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
-%\ge 1.2$ for only consuming one minibatch data. For six minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
-% our MH test five times and averaged the number of ``violations.'' As the size increases, the
-% number of violations decreases, as expected.
-
-\iffalse
-Figure~\ref{fig:gauss_mix_3} (to the left) shows a histogram of ${\rm
-  std}(\Delta')$. Figure~\ref{fig:gauss_mix_3} (to the middle) shows
-the confidence interval of our estimated ${\rm std}(\Delta')$ at each
-iteration \footnote{In order to show the confidence interval clearly,
-  we use moving smooth to smooth the confidence interval, and plot
-  each value by interval of 50 iterations}. The blue line represents
-the standard deviation comes from the complete data, which can be
-treated as the ``true'' ${\rm std}(\Delta')$, the green line
-represents our estimation of ${\rm std}(\Delta')$ from the minibatch,
-and the two red dash lines represent the confidence interval of ${\rm
-  std}(\Delta')$ by adding / subtracting one ${\rm
-  std(std(}\Delta'))$. This figure indicates that our estimation of
-${\rm std}(\Delta')$ from one minibatch data is
-accurate. Figure~\ref{fig:gauss_mix_3} (to the right) shows the value
-of $\rho/\sigma^3$ at each iteration. 
-\fi
+Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
+samples for the three methods, with darker regions indicating a greater density
+of points. The three methods obtain the same rough form of the posterior, so our
+MH test can indeed result in similar posteriors as the other two methods.
+
+In addition, our method is superior based on the Poisson likelihood of the
+MCMC-generated parameters over the range of $\theta_1 \in (-1.5, 2.5)$ and
+$\theta_2 \in (-2.5, 2.5)$. This is a measure of divergence from the
+ground-truth distribution. The log likelihood values for our method, Korattikara
+2014, and Bardenet 2014 are -2144.6, -4007.4, and -2598.4, respectively.
+{\color{blue} Daniel: I rewrote this paragraph to make it slightly less
+offensive. Also, should we clarify what this means? Because to be honest, it
+seems hard to tell a significant difference among the methods by looking at the
+scatterplots.}
+
+Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used
+by the three methods in each iteration. Our methods consumes significantly less
+data; most minibatch sizes are smaller than 1000, and the average size is 420.
+The other two methods occasionally need to consume a large proportion of the
+entire data set of one million elements; the average minibatch sizes are 16378
+and 16755 for Korattikara 2014 and Bardenet 2014, respectively. {\color{blue}
+Daniel: the plot says our average minibatch size is 210, but the paragraph here
+says 420.}
+
+It is useful to notice that, in~\cite{icml2014c1_bardenet14}, they need to
+evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set
+at every iteration in order to calculate the error bound in equation (9)
+in~\cite{icml2014c1_bardenet14}. In Figure ~\ref{fig:gauss_bardenet}, the value
+of the first part and second part of equation (9)
+in~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual
+minibatch data used. The second part which contains  the $C_{\theta,\theta'}$
+term is larger than the first part, which means that the second part is
+non-negligible and has to be calculated at every iteration. This is one of the
+very time-consuming parts in this method.  {\color{blue} Daniel: I am not sure
+if we should include this plot. Let me think about it and if we decide to
+include it I will help rewrite the explanation.}
 

### 2016-09-23 nips_2016.tex
Modified Secdtion 6.2

@@
 	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
-	\caption{Value of first and second part of equation (9) in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size}
+	\caption{
+    Value of first and second part of equation (9)
+    in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size.
+    {\color{blue} Daniel: probably debatable on whether we should include this.
+    Let's talk it over.}
+    }
 	\label{fig:gauss_bardenet}
@@
 
-We next use logistic regression for the binary classification of 1s versus 7s in the MNIST
-dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007 and 1000 training and testing
-points, respectively (we used a random subset of the test data). The proposer is again a random walk
-with covariance matrix $0.05I$ for the $784\times 784$ identity matrix $I$. We set the posterior
-temperature at $T=1000$. We set the minibatch size $m=100$ and compare with ~\cite{cutting_mh_2014} with
-tolerance 0.005 and with ~\cite{icml2014c1_bardenet14} with error bound control parameter $p = 2, \delta = 0.01$.
-
-Figure~\ref{fig:logistic_fig} shows the prediction accuracy and log likelihood on the test set as a
-function of the cumulative training data points processed. Our test increments the cumulative data
-by a fixed amount per iteration, but the other two methods may require more data per iteration.  We
-see that our minibatch MH test is more efficient; it has similar or better performance while
-consuming fewer data points.
+We next use logistic regression for the binary classification of 1s versus 7s in
+the MNIST dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007
+training and 1000 testing elements (we used a random subset of the test data).
+For the proposer, we use a random walk with covariance matrix $0.05I$ for the
+$784\times 784$ identity matrix $I$. We set the posterior temperature at
+$T=1000$ and use the minibatch size $m=100$ and compare
+with~\cite{cutting_mh_2014} with tolerance 0.005 and
+with~\cite{icml2014c1_bardenet14} with error bound control parameters $p = 2$
+and $\delta = 0.01$. {\color{blue} Daniel: just observed this, but why are we
+using a subset of the test data? We really should use all 2000 (I think it's
+2000, right?).}
+
+Figure~\ref{fig:logistic_fig} shows the prediction accuracy and log likelihood
+on the test set as a function of the cumulative training data points processed.
+Our test increments the cumulative data by a fixed amount per iteration
+{\color{blue} (Daniel: are we sure about ``fixed amount per iteration'' because
+our Algorithm 1 implies we may have to increment the minibatch size, but it
+happens \emph{less frequently} than the other methods.)}, but the other two
+methods may require more data per iteration. We see that our minibatch MH test
+is more efficient; it has similar or better performance while consuming fewer
+data points.
 
@@
 \end{subfigure}
-\caption{Logistic regression performance (accuracy/log likelihood) and minibatch size analysis}
+\caption{
+    Logistic regression performance (accuracy/log likelihood) and minibatch size
+    analysis. {\color{blue} Daniel: let's use matplotlib's ``subplot''
+    functionality, so that the images appear the same size. We can handle this
+    later; it will be easy.}
+    }
 \label{fig:logistic_fig}
-\vspace{-10pt}
 \end{figure}
@@
 	\includegraphics[width=1\linewidth]{LogisticRegressionResult/minibatch_size_logistic.png}
-	\caption{Histograms of minibatch sizes of logistic regression parameter estimation}
+	\caption{
+    Counts of minibatch sizes needed for the three methods in the MNIST
+    classification task.
+    }
 	\label{fig:logistic_minibatch}
@@
 
-%\begin{figure}[t]
-  %  \centering
-    %\includegraphics[width=1\linewidth]{CI_v02.png}
-    %\caption{
-   % Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
-    %}
-    %\label{fig:exp_3}
-    %\vspace{-10pt}
-%\end{figure}
-
-The plot of Figure ~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes. Choosing an initial minibatch size of 100, 
-the cut MH method, adaptive MH method achieves final average minibatch size of 585 and 2731 respectively, while our method achieves 
-that of 100, showing significant better performance than the benchmark test methods.
-
-%The third plot of Figure~\ref{fig:logistic_minibatch} shows performance based on minibatch size. The
-%parameter is evaluated on the test set after training iterations 700, 1000, and 1500. For those
-%three cases, when the minibatch size increases beyond about 10, the performance of the sampled data
-%will not improve with a large minibatch size. This makes sense because the error of our MH test is
-%proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch increases, ${\rm
-%std}(\Delta')$ decreases. Beyond a certain standard deviation (or minibatch) threshold, the error in
-%our minibatch MH test is negligible.
-
-%Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{fig:exp_3}
-%to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability
-%of $\rho/\sigma^3$.
-
-%% \subsection{Neural Network Optimization}\label{ssec:nets}
-%% 
-%% In this experiment, we apply our minibatch MH test with a Stochastic Gradient Hamiltonian
-%% Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the complete parameters for a fully
-%% connected neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}, which
-%% is a binary classification task with one million instances, each of which are 28-dimensional. The
-%% network we use has four layers: the first is the input, the last is the softmax, and the
-%% intermediate ones apply the sigmoid activation unit.  We also employ dropout and batch
-%% normalization~\cite{icml2015_ioffe15}.  In order to control ${\rm std}(\Delta')$, we initialize the
-%% temperature at 1000, and adjust it at iteration $i$ according to $T_i = \max\{1,
-%% 1000/(i+1)^{0.5}\}$.
-%% 
-%% For our comparison baseline, we train the same network architecture using the Adaptive Gradient
-%% Descent method~\cite{adapGrad}. Since the frequencies of the features are quite different, we use
-%% the same strategy as in Adaptive Gradient to rescale the gradient before applying SGHMC. We
-%% implement our MH test and SGHMCin the open-source BIDMach project~\cite{canny2013bidmach}.  
-%% 
-%% Figure~\ref{fig:nnet_fig} illustrates our experiment results and reveals that SGHMC with our
-%% minibatch MH test achieves higher log likelihood than the baseline. By comparing SGHMC with and
-%% without our minibatch MH test, we see that our MH test has limited benefit. Because we decrease the
-%% step size quickly, the acceptance rate of SGHMC is already high, so there is little need to test.
-%% % Daniel's note: Haoyu said he cannot increase the step size without it violating the variance
-%% % condition.
-%% 
-%% \begin{figure}[t]
-%%     \centering
-%%     \includegraphics[width=1\linewidth]{exp3}
-%%     \caption{Log likelihood versus number of passes for optimization of neural network.}
-%%     \label{fig:nnet_fig}
-%%     \vspace{-10pt}
-%% \end{figure}
+Figure~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes for
+all three methods. With an initial minibatch size of 100, Korattikara 2014 and
+Bardenet 2014 achieve average minibatch size of 585 and 2731 respectively over
+the MNIST classification task, while our method achieves that of 100, showing
+significant better performance than the benchmark test methods. {\color{blue}
+Daniel: these numbers don't match the figure titles.}
 

### 2016-09-23 nips_2016.tex

@@
 
-{\color{blue} Daniel: Hold on, the following Lemma seems to add $X_{\rm norm}$
+{\color{blue} Daniel: Hold on, the following Corollary seems to add $X_{\rm norm}$
 to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
@@
 think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
-match what we wrote earlier.}
+match what we wrote earlier.
+
+I think the easiest way to resolve this is to do the following: we assume that
+$\Delta^*$ is already $\Delta + X_{\rm noise}$. Then all we need to do is add
+$X_{\rm corr}$. So this is the same as what we are doing, except we do not need
+to apply Lemma~\ref{lem:cdf_bounds} twice; we just need to do it once because we
+are only adding $X_{\rm corr}$, not $X_{\rm norm}$.}
 
@@
 and convolve with the distribution of $X_n$ {\color{blue} (Daniel: to be clear I
-think this was supposed to be $X_{\rm norm}$ but see comments earlier about why
-I think it is not necessary.)} which has density $\phi(X/\sigma_n)$ where
-$\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of $P$ and
-$Q$
+think this was supposed to be $X_{\rm norm}$ instead of $X_n$ but see comments
+earlier about why I think it is not necessary.)} which has density
+$\phi(X/\sigma_n)$ where $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the
+next iteration of $P$ and $Q$:
 \begin{equation}
@@
 table. We leave further analysis and experiments with $g(x,y)$ as future work.
-{\color{blue} Daniel: I will look at the code to understand this paragraph better.}
+{\color{blue} Daniel: I will look at the code to understand this paragraph
+better because right now I am confused.}
 
@@
     \item More information on Section~\ref{ssec:logistic}.
-    %\item More information on Section~\ref{ssec:nets}. % Daniel: this is in
-    %progress.
 \end{itemize}
+
+{\color{blue} Daniel: Note that everything after
+Appendix~\ref{app:worst_case_proof} has yet to be revised. I don't know what we
+will include in them.}
 

### 2016-09-28 nips_2016.tex
updates

@@
 data required for each test is fixed with high probability and in expected
-value. {\color{blue} Daniel: we really need to show the amount of data is fixed
-with ``high probability'' (the results suggest ``OK probability'').} We use a
-Barker test function~\cite{Barker65} rather than a Metropolis test, which makes
-our test naturally error tolerant. The idea of using a noise-tolerant test using
-Barker's test function was suggested but not explored empirically
-in~\cite{Bardenet15} section 6.3. But the asymptotic test statistic CDF and the
-Barker function are different, which leads to fixed errors for the approach
-in~\cite{Bardenet15}. Here we show that the difference between the distributions
-can be corrected with an additive random variable. This leads to a test which is
-fast, and whose error can be made arbitrarily small.
+value.
+%% {\color{blue} Daniel: we really need to show the amount of data is fixed
+%% with ``high probability'' (the results suggest ``OK probability'').} 
+%% Analysis show data amount is gaussian and short-tailed and therefore bounded with high p
+%% experiments back this up. 
+We use a Barker test function~\cite{Barker65} rather than a Metropolis
+test, which makes our test naturally error tolerant. The idea of using
+a noise-tolerant test using Barker's test function was suggested but
+not explored empirically in~\cite{Bardenet15} section 6.3. But the
+asymptotic test statistic CDF and the Barker function are different,
+which leads to fixed errors for the approach
+in~\cite{Bardenet15}. Here we show that the difference between the
+distributions can be corrected with an additive random variable. This
+leads to a test which is fast, and whose error can be made arbitrarily
+small.
 
@@
 proposals~\cite{OptimalScaling01} on a full dataset. When we reduce the amount
-of data for the test, the variance goes up, and we have to reduce it in one of
-several ways. Either:
+of data for the test, the variance goes up. We have to reduce variance in one
+of several ways. Either:
 
@@
 
-\item Increase the minibatch size. Log acceptance variance scales as $1/k$ vs the
-  minibatch size $k$. Increased minibatch size also reduces the error rate for
-  the test. {\color{blue} Daniel: I'm not sure if I'd mention this here, because
-  before this list, we talk about reducing the data, and then suddenly we're
-  talking about increasing it.}
+\item Increase the minibatch size when needed for certain
+  minibatches. Log acceptance variance scales as $1/k$ vs the
+  minibatch size $k$. Our test is adaptive like earlier works, but
+  unlike them, the distribution of minibatch size is gaussian, not
+  long-tailed.  Increased minibatch size also reduces the error rate
+  for the test.
+
+%% {\color{blue} Daniel: I'm not sure if I'd mention this here,
+%%    because before this list, we talk about reducing the data, and
+%%    then suddenly we're talking about increasing it.}
 
@@
 Since minibatches contains randomly selected samples $x_i$, the values
-$\Lambda_i$ are independent, identically distributed (iid) random variables.
-{\color{blue} Daniel: should we say sampling with or without replacement? Right
-now I think the code does it without replacement (but \emph{with} replacement
-means it's closer to the CLT, right?} By the Central Limit Theorem, we expect
-$\Lambda^*(\theta,\theta')$ to be approximately Gaussian. The acceptance test
-then becomes a statistical test of the hypothesis that
+$\Lambda_i$ are independent, identically distributed (iid) random
+variables\footnote{The analysis assumes sampling with replacement
+  although implementations on typical large datasets will approximate
+  this by sampling without replacement}.
+%% {\color{blue} Daniel: should we say sampling with or without replacement? Right
+%%now I think the code does it without replacement (but \emph{with} replacement
+%% means it's closer to the CLT, right?} 
+By the Central Limit Theorem, we expect $\Lambda^*(\theta,\theta')$ to
+be approximately Gaussian. The acceptance test then becomes a
+statistical test of the hypothesis that
 $\Lambda(\theta,\theta')>\psi(u,\theta,\theta')$ by establishing that
 $\Lambda^*(\theta,\theta')$ is substantially larger than
-$\psi(u,\theta,\theta')$.  In~\cite{cutting_mh_2014} an asymptotic central limit
-argument was used to derive this gap, while in~\cite{icml2014c1_bardenet14} a
-concentration bound was used. In both cases, the resulting tests were shown to
-give useful reductions in number of samples required over using the full
-dataset, but there were no worst-case bounds given on the average number of
-samples needed per iteration.
+$\psi(u,\theta,\theta')$.  In~\cite{cutting_mh_2014} an asymptotic
+central limit argument was used to derive this gap, while
+in~\cite{icml2014c1_bardenet14} a concentration bound was used. In
+both cases, the resulting tests were shown to give useful reductions
+in number of samples required over using the full dataset, but there
+were no worst-case bounds given on the average number of samples
+needed per iteration.
 
@@
 large enough gap (several $\sigma(\Lambda^*)$) between
-$\Lambda^*(\theta',\theta)$ and $\psi(u,\theta',\theta)$. {\color{blue}
-Daniel: I changed the variance to be $(\theta'-\theta)^2N^2/b$; before it had
-$N$ instead of $N^2$.}
+$\Lambda^*(\theta',\theta)$ and $\psi(u,\theta',\theta)$. 
+% {\color{blue} Daniel: I changed the variance to be $(\theta'-\theta)^2N^2/b$; before it had
+% $N$ instead of $N^2$.}
 
@@
 distribution~\cite{OptimalScaling01}, so we use the proposal
-$\mathcal{N}(\theta'-\theta,1/N)$, which is implemented as
+$\mathcal{N}(\theta,1/N)$, which is implemented as
 $q(\theta'\mid\theta)=\phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal
@@
 worst-case scenario is specified in Lemma~\ref{lem:worst_case}.
-{\color{blue} Daniel: this might be a silly question, but why do we assume the
-proposal has mean $\theta'-\theta$? I thought it would have a mean of $\theta$
-because that is the current sample and the Gaussian is centered at that point?
-That will also clarify my confusion about the $\psi$ function.}
+%{\color{blue} Daniel: this might be a silly question, but why do we assume the
+%proposal has mean $\theta'-\theta$? I thought it would have a mean of $\theta$
+%because that is the current sample and the Gaussian is centered at that point?
+%That will also clarify my confusion about the $\psi$ function.} - fixed
 
@@
 
-{\color{blue} Daniel: quick question: I think we should explain why our method
-\emph{doesn't} run into this issue. I think it is because we rely on an
-approximation to $\Delta + X_{\rm log} > 0$ and that test does not require any
-``difference between an approximation and an exact version'' because it is
-\emph{already} an exact version, assuming the approximation is enough.
-
-Also, in terms of this example, if we are in the special case with using a lot
-of data, why not force early stopping?}
+% {\color{blue} Daniel: quick question: I think we should explain why our method
+% \emph{doesn't} run into this issue. I think it is because we rely on an
+% approximation to $\Delta + X_{\rm log} > 0$ and that test does not require any
+% ``difference between an approximation and an exact version'' because it is
+% \emph{already} an exact version, assuming the approximation is enough.
+% Also, in terms of this example, if we are in the special case with using a lot
+% of data, why not force early stopping?}
+
+The problem is that these methods use tail bounds to separate $\Delta$ away
+from zero, but for certain input/random $u$ combinations, $\Delta$ can be
+arbitrarily close to zero. We will avoid this by using the {\em approximately
+normal} variation in $\Delta^*$ to {\em replace} the variation due to $u$.
 

### 2016-09-29 nips_2016.tex
addressed comments

@@
 which allows momentum-augmented proposals with larger step sizes. However, step
-sizes are still limited by the need to run accurately without MH tests.  Our
+sizes are still limited by the need to run accurately without MH tests.  By providing
+an M-H test with similar cost to standard gradient steps, our
 work opens the door to applying those methods with much more aggressive step
-sizes without loss of accuracy. We demonstrate this in Section ... {\color{blue}
-Daniel: do we ever demonstrate this? I guess another issue is that this implies
-we can use larger step sizes, but earlier we \emph{also} argue that we need
-\emph{small} step sizes (to reduce variance).}
+sizes without loss of accuracy. 
+%{\color{blue}
+%Daniel: do we ever demonstrate this? I guess another issue is that this implies
+%we can use larger step sizes, but earlier we \emph{also} argue that we need
+%\emph{small} step sizes (to reduce variance).}
 
@@
 $\Delta^*(\theta,\theta')$ is an unbiased estimator of $\Delta(\theta,\theta')$.
-{\color{blue} Daniel: a brief question repeated from earlier, should we mention
-sampling with and/or without replacement here?}
+%{\color{blue} Daniel: a brief question repeated from earlier, should we mention
+%sampling with and/or without replacement here?}
 
@@
 condition in Lemma~\ref{lem:detailed_balance}.  While it is slightly less
-efficient than the Metropolis test when used on the full dataset, we will show
+efficient than the Metropolis test when used on the full dataset, we will see
 that its smoothness allows it to naturally tolerate substantial variance in its
 input argument. This in turn will lead to a much more efficient test on subsets
-of data. {\color{blue} Daniel: aside from the experiments, where do we show it
-``naturally tolerates substantial variance'' in the input?}
+of data.
+%{\color{blue} Daniel: aside from the experiments, where do we show it
+%``naturally tolerates substantial variance'' in the input?}
 
@@
 \end{equation}
-for a logistic random variable $X$. {\color{blue} Daniel: I think the text
-before Equation~\ref{eq:the_exact_test} should be similar to what we had in our
-NIPS submission because I find it much clearer.}
+for a logistic random variable $X$.
+%{\color{blue} Daniel: I think the text
+%before Equation~\ref{eq:the_exact_test} should be similar to what we had in our
+%NIPS submission because I find it much clearer.}
 
@@
 
-{\color{blue} Daniel: this subsection is improved from earlier. My main concern
-now is that $X_{\rm norm}$ is described as being both approximately Gaussian
-$\bar{\mathcal{N}}$ and Gaussian $\mathcal{N}$ here. We may want to clear that
-up; in our NIPS submission we actually used a new variable $\varepsilon$ for
-that and assigned it to be the $\mathcal{N}$ distribution. My current
-suggestion is to have $\Delta^*$ be the $\bar{\mathcal{N}}$ approx. normal
-distribution (but with mean $\Delta$, of course), and keep $X_{\rm norm}$ as an
-exact Gaussian.}
+%% {\color{blue} Daniel: this subsection is improved from earlier. My main concern
+%% now is that $X_{\rm norm}$ is described as being both approximately Gaussian
+%% $\bar{\mathcal{N}}$ and Gaussian $\mathcal{N}$ here. We may want to clear that
+%% up; in our NIPS submission we actually used a new variable $\varepsilon$ for
+%% that and assigned it to be the $\mathcal{N}$ distribution. My current
+%% suggestion is to have $\Delta^*$ be the $\bar{\mathcal{N}}$ approx. normal
+%% distribution (but with mean $\Delta$, of course), and keep $X_{\rm norm}$ as an
+%% exact Gaussian.}
 
@@
 \end{equation}
-where $*$ denotes convolution. {\color{blue} Daniel: This is very unusual use of
-convolutions (i.e., CDF * DENSITY = CDF). It's valid but I was very surprised to
-know that.} For computation of $C_\sigma$, we assume that its input $Y$ and
+where $*$ denotes convolution.
+% {\color{blue} Daniel: This is very unusual use of
+%convolutions (i.e., CDF * DENSITY = CDF). It's valid but I was very surprised to
+% know that.}
+For computation of $C_\sigma$, we assume that its input $Y$ and
 another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
-{\color{blue} Daniel: I am a little confused at the different interval sizes.}
+% {\color{blue} Daniel: I am a little confused at the different interval sizes.}
 In continuous form, the optimization problem is therefore
@@
 \end{equation}
-To get $C_\sigma^*$ in a practical form, we discretize $X$ and $Y$ into $4N+1$
+To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
 and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}$ and $j \in \{-N,
@@
 appropriately translated to be non-negative indices for $M, u,$ and $v$. Thus,
-Equation~\ref{eq:conv2} is equivalent to minimizing the $L_1$ norm:
+Equation~\ref{eq:conv2} is equivalent to minimizing the $L_{\infty}$ norm:
 \begin{equation}\label{eq:optimization_l1}
-    u^* = \argmin_{u} \|Mu-v\|_1.
-\end{equation}
-{\color{blue} Daniel: Equation~\ref{eq:optimization_l1} is finding
-the $C_\sigma$ minimizing the $L_1$ norm, but the previous one,
-Equation~\ref{eq:conv2}, is finding the $C_\sigma$ that is minimizing the
-$L_\infty$ norm, i.e., the maximum value! If we want this to be the $L_1$ norm
-(and I think we do) then Equation~\ref{eq:conv2} needs an outer sum over the $i$
-terms. In addition, Equation~\ref{eq:conv1} also needs to be modified, perhaps
-eliminate the ``sup'' but keep $x \in [-2V,2V]$}
+    u^* = \argmin_{u} \|Mu-v\|_{\infty}.
+\end{equation}
+% {\color{blue} Daniel: Equation~\ref{eq:optimization_l1} is finding
+% the $C_\sigma$ minimizing the $L_{\infty}$ norm, but the previous one,
+% Equation~\ref{eq:conv2}, is finding the $C_\sigma$ that is minimizing the
+% $L_\infty$ norm, i.e., the maximum value! If we want this to be the $L_{\infty}$ norm
+% (and I think we do) then Equation~\ref{eq:conv2} needs an outer sum over the $i$
+% terms. In addition, Equation~\ref{eq:conv1} also needs to be modified, perhaps
+% eliminate the ``sup'' but keep $x \in [-2V,2V]$}
 
@@
 tractable up to a few hundred dimensions for $u$. However, the discretization
-error is bounded by the curvature of the underlying functions which are here
+error is bounded by the curvature of the underlying functions times the squared
+quantization step. The curvatures are here
 slightly less than one, i.e.  the errors are of the order of $(V/N)^2$.
-{\color{blue} Daniel: unfortunately, I do not understand what this means (i.e.,
-the curvature and how the $(V/N)^2$ is obtained).} Here we chose $V=20$ to
+% {\color{blue} Daniel: unfortunately, I do not understand what this means (i.e.,
+% the curvature and how the $(V/N)^2$ is obtained).} Here we chose $V=20$ to
 provide adequate containment of the distributions (the CDFs are extremely close
@@
 dimensions. The solution is well-known from the normal equations: $u^* = (M^TM +
-\lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_1$ norm for
+\lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_{\infty}$ norm for
 Equation~\ref{eq:optimization_l1}.
@@
 some flexibility in the choice of $\sigma$ in Equation~\ref{eq:conv1}.
-{\color{blue} Daniel: is this the correct equation?} As we decrease the variance
+% {\color{blue} Daniel: is this the correct equation?}
+As we decrease the variance
 of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount and
@@
 }
-\caption{A description of M-H sampling with our acceptance test. {\color{blue}
-Daniel: I think we may want to be careful about the use of $C_1(X)$ because that
-implies that we have a fixed $X_{\rm corr}$ density, but that $C_1(X)$ will
-change depending on the iteration and variance of $\Delta^*$, right? Hopefully
-not by much, but it still changes sometimes? ALSO bring up Xinlei's old points
-about $X_{\rm norm}$ confusion.}}
+\caption{A description of M-H sampling with our acceptance test.
+% {\color{blue}
+% Daniel: I think we may want to be careful about the use of $C_1(X)$ because that
+% implies that we have a fixed $X_{\rm corr}$ density, but that $C_1(X)$ will
+% change depending on the iteration and variance of $\Delta^*$, right? Hopefully
+% not by much, but it still changes sometimes? ALSO bring up Xinlei's old points
+  %about $X_{\rm norm}$ confusion.}
+}
 \label{alg:our_algorithm}
@@
 
-\begin{table}[t]
-    \caption{\color{blue} Daniel: TODO need a good title.}
+\begin{table}[h]
+    \caption{Error in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$}
     \label{tab:xcorr}
@@
     \toprule
-    $N$ & $\sigma$ & $\lambda$ & $L_1$ error  \\
+    $N$ & $\sigma$ & $\lambda$ & $L_{\infty}$ error  \\
     \midrule
@@
 
+As can be seen from table~\ref{tab:xcorr}, the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be made very small, approaching single floating precision (about $10^{-7}$).
+
 {\color{blue} Daniel: TODO improve/describe Table~\ref{tab:xcorr} and try to
@@
 We now derive error bounds for our M-H test, and for the approximate target
-distribution that it generates. From Section~\ref{sec:correction}, we know that
+distribution that it generates. From Table~\ref{tab:xcorr}, we know that
 it is possible to generate the correction samples $X_{\rm corr}$ with a CDF
-error of at most single-precision floating point error, and possibly better
-{\color{blue} (Daniel: I can fix this to cite Table~\ref{tab:xcorr}'s values,
-that will probably be better.)}. We therefore treat the $X_{\rm corr}$ variable
-as a sample from the exact correction distribution and we will not analyze its
-errors.
+error approaching single-precision floating point error.
+% {\color{blue} (Daniel: I can fix this to cite Table~\ref{tab:xcorr}'s values,
+% that will probably be better.)}.
+We therefore treat the $X_{\rm corr}$ variable as a sample from the
+exact correction distribution and we will not analyze its errors.
 

### 2016-09-29 nips_2016.tex
almost done

@@
 We show in Section~\ref{sec:correction} that it is possible to use deconvolution
-to derive a highly accurate representation of $C_\sigma(X)$.
+to derive a highly accurate representation of $C_{\sigma}(X)$.
 Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
@@
 
-Using $X_{\rm corr}$ samples from $C_\sigma(X)$, the acceptance test is now
+Using $X_{\rm corr}$ samples from $C_{\sigma}(X)$, the acceptance test is now
 \begin{equation}\label{eq:criteria}
@@
 log}$. Consequently, we need to minimize this in our construction of $X_{\rm
-corr}$. More formally, let $\Phi_\sigma(X) = \Phi(X/\sigma)$ where $\Phi$ is the
-standard normal CDF\footnote{Hence, $\Phi_\sigma$ is the CDF of a zero-mean
-Gaussian with variance $\sigma$.}, $S(X)$ be the logistic function, and
+corr}$. More formally, let $\Phi_{s_X} = \Phi(X/s_X)$ where $\Phi$ is the
+standard normal CDF\footnote{Hence, $\Phi_{s_X}$ is the CDF of a zero-mean
+Gaussian with variance $s_X$.}, $S(X)$ be the logistic function, and
 $C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
@@
     Propose a candidate $\theta'$\ from proposal distribution $q(\theta'\mid\theta_t)$\;
-    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $\sigma^2(\Delta^*)$\;
+    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*})$\;
     Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Equation~\ref{eq:clt-bounds3}\;
-    \While {$\sigma^2(\Delta^*) \ge 1$ {\bf or} $\epsilon > \delta$}{
-    	Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $\sigma^2(\Delta^*)$ and $\epsilon$ estimates\;
+    \While {$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}{
+    	Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $s^2_{\Delta^*}$ and $\epsilon$ estimates\;
     }
-    Draw $X_{\rm norm} \sim \mathcal{N}(0,1-\sigma^2(\Delta^*))$ and
+    Draw $X_{\rm nc} \sim \mathcal{N}(0,1-s^2_{\Delta^*})$ and
     $X_{\rm corr}$ from the correction distribution $C_1(X)$\;
 
-        \eIf{$\Delta^* + X_{\rm norm} + X_{\rm corr} > 0$}{
+        \eIf{$\Delta^* + X_{\rm nc} + X_{\rm corr} > 0$}{
             Accept the candidate, $\theta_{t+1} = \theta'$\;
@@
 
-{\color{blue} Daniel: TODO improve/describe Table~\ref{tab:xcorr} and try to
-relocate Algorithm~\ref{alg:our_algorithm} to be somewhere nicer.}
+The complete algorithm is given in~\ref{alg:our_algorithm}. A few points:
+\begin{itemize}
+\item The algorithm uses an adaptive step size so as to use the smallest possible
+  average minibatch size. Unlike previous work however (and see the experiments section)
+  the size distribution is short-tailed.
+\item An additional normal variable $X_{\rm nc}$ is added to $\Delta^*$ to produce a variable
+  with unit variance. This is not mathematically necessary, but allows us to use a single
+  correction distribution $C_1$ for $X_{\rm corr}$, saving on memory footprint.
+\end{itemize}
+
+The sample variance $s^2_{\Delta^*}$ is proportional to $||\theta'-\theta||^2$
+whose distribution for Normal proposals is the square of a normal variable. 
+% {\color{blue} Daniel: TODO improve/describe Table~\ref{tab:xcorr} and try to
+% relocate Algorithm~\ref{alg:our_algorithm} to be somewhere nicer.}
 
@@
 Section~\ref{ssec:preserve_bounds} that these error bounds are preserved after
-the addition of other random variables, in particular $X_{\rm norm}$ and $X_{\rm
+the addition of other random variables, in particular $X_{\rm nc}$ and $X_{\rm
 corr}$. From this it follows that the acceptance test has the same error bound.
@@
 
-{\color{blue} Daniel: the following lemma is, I think, from Corollary 2
-in~\cite{explicit-clt05} but I cannot derive the exact lemma as we have it
-here.} To start, we use the following quantitative central-limit result
-(from~\cite{explicit-clt05}):
+%{\color{blue} Daniel: the following lemma is, I think, from Corollary 2
+%in~\cite{explicit-clt05} but I cannot derive the exact lemma as we have it
+%here.} To start, we use the following quantitative central-limit result
+%(from~\cite{explicit-clt05}):
 
@@
 Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
-random variables with sample mean $\hat{X}$ and variance $\sigma^2(\hat{X})$
+random variables with sample mean $\bar{X}$ and sample variance $s^2_X$
 where:
 \begin{equation}
-    \hat{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad \sigma(\hat{X}) = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\hat{X})^2\right)^{\frac{1}{2}}.
-\end{equation}
-This means $t=\hat{X}/\sigma(\hat{X})$ has an approximate Student's distribution
+    \bar{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad s_X = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\bar{X}^2\right)^{\frac{1}{2}}.
+\end{equation}
+This implies $t=\bar{X}/s_X$ has an approximate Student's distribution
 which approaches a normal distribution in the limit. Then
@@
 
-{\color{blue} Daniel: the notation here was (and still is, to a lesser extent)
-the source of considerable confusion for me. Do we want $\sigma(X)$ or
-$\sigma(\hat{X})$? I think in Lemma~\ref{lem:quant_clt} we want the former, not
-the latter. OH ALSO, this section uses $n$ as the full data size, which means
-before this section, we should use $n$, not $N$, for the full data size, right?
-The Bardenet paper uses $n$ to represent the full data size.
-
-PS: I'd recommend not putting sample mean and sample variance in an equation if
-the $\sigma(\hat{X})$ gets clarified, since those computations are common
-knowledge.
-
-PPS: The sample mean is probably more commonly denoted as $\bar{X}$.}
+\begin{proof}
+  The following bound is given immediately after Corollary 2 in from~\cite{explicit-clt05}:
+\begin{equation}\label{eq:clt-boundsx}
+-6.4E|X|^3-2E|X| \leq    \sup_x|{\rm Pr}(t<x) - \Phi(x)|{\sqrt{n}} \leq 1.36E|X|^3
+\end{equation}
+This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
+and combining with $x>0$, we obtain the weaker but unqualified bound
+in equation (\ref{eq:clt-bounds}).
+\end{proof}
+
+%% {\color{blue} Daniel: the notation here was (and still is, to a lesser extent)
+%% the source of considerable confusion for me. Do we want $\sigma(X)$ or
+%% $\sigma(\hat{X})$? I think in Lemma~\ref{lem:quant_clt} we want the former, not
+%% the latter. OH ALSO, this section uses $n$ as the full data size, which means
+%% before this section, we should use $n$, not $N$, for the full data size, right?
+%% The Bardenet paper uses $n$ to represent the full data size.
+
+%% JFC - this bound applies to minibatches, not the full dataset so the size $n$ is
+%% deliberately not $N$.
 
@@
 approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of
-variables $y = x \sigma(\hat{X})$, Equation~\ref{eq:clt-bounds} becomes
+variables $y = x s_X)$, Equation~\ref{eq:clt-bounds} becomes
 \begin{equation}\label{eq:clt-bounds2}
-   \sup_y\left|{\rm Pr}(\hat{X}< y) - \Phi\left(\frac{y}{\sigma(\hat{X})}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}},
-\end{equation}
-showing that the distribution of $\hat{X}$ approaches the normal distribution
-$\mathcal{N}(0,\sigma^2(\hat{X}))$ whose variance is $\sigma^2(\hat{X})$,
+   \sup_y\left|{\rm Pr}(\bar{X}< y) - \Phi\left(\frac{y}{s_X}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}},
+\end{equation}
+showing that the distribution of $\bar{X}$ approaches the normal distribution
+$\mathcal{N}(0,s_X)$ whose variance is $s_X$,
 measured from the sample.
@@
 instead of all $n$ samples, we only extract a subset of $b$ samples
-corresponding to our minibatch, we can connect $\hat{X}$ with our $\Delta^*$
+corresponding to our minibatch, we can connect $\bar{X}$ with our $\Delta^*$
 term:
 \begin{equation}\label{eq:x-hat}
-    \hat{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta'),
-\end{equation}
-so that $\sigma(\hat{X}) = \sigma(\Delta^*)$. {\color{blue} Daniel: This time it
-makes sense to use $\sigma(\hat{X})$ instead of $\sigma(X)$.} This results in
-the following Corollary:
+    \bar{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta'),
+\end{equation}
+so that $s_X = s_{\Delta^*}$.
+%{\color{blue} Daniel: This time it
+%  makes sense to use $\sigma(\hat{X})$ instead of $\sigma(X)$.}
+This results in the following Corollary:
 
@@
 \begin{equation}\label{eq:clt-bounds3}
-    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} = \epsilon(\theta,\theta',b)
+    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} = \epsilon(\theta,\theta',b)
 \end{equation}
@@
 $\Delta^*$ approximates a Normal distribution with mean $\Delta$ and variance
-$\sigma^2(\Delta^*)$. Furthermore, it bounds the error with \emph{estimable
-quantities}: both $E|X|$ and $E|X|^3$ can be determined as means of $|\Lambda_i
+$s^2_{\Delta^*}$. Furthermore, it bounds the error with \emph{estimable
+quantities}: both $E|X|$ and $E|X|^3$ can be estimated as means of $|\Lambda_i
 - \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each minibatch. We
@@
 
-{\color{blue} Daniel: Hold on, the following Corollary seems to add $X_{\rm norm}$
-to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
-didn't the previous section already justify that $\Delta^* \approx \Delta +
-X_{\rm norm}$? That is, I don't think we need to add another $X_{\rm norm}$; I
-think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
-match what we wrote earlier.
-
-I think the easiest way to resolve this is to do the following: we assume that
-$\Delta^*$ is already $\Delta + X_{\rm noise}$. Then all we need to do is add
-$X_{\rm corr}$. So this is the same as what we are doing, except we do not need
-to apply Lemma~\ref{lem:cdf_bounds} twice; we just need to do it once because we
-are only adding $X_{\rm corr}$, not $X_{\rm norm}$.}
+%% {\color{blue} Daniel: Hold on, the following Corollary seems to add $X_{\rm norm}$
+%% to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
+%% didn't the previous section already justify that $\Delta^* \approx \Delta +
+%% X_{\rm norm}$? That is, I don't think we need to add another $X_{\rm norm}$; I
+%% think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
+%% match what we wrote earlier.
+
+%% I think the easiest way to resolve this is to do the following: we assume that
+%% $\Delta^*$ is already $\Delta + X_{\rm noise}$. Then all we need to do is add
+%% $X_{\rm corr}$. So this is the same as what we are doing, except we do not need
+%% to apply Lemma~\ref{lem:cdf_bounds} twice; we just need to do it once because we
+%% are only adding $X_{\rm corr}$, not $X_{\rm norm}$.}
 
@@
 \begin{corollary}\label{cor:bounds_preserved}
-If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})|
+If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{s_{\Delta^*}})|
 \leq \epsilon(\theta,\theta',b)$, then
 \begin{equation}
-    \sup_y|{\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
-\end{equation}
-where $S(x)$ is the standard logistic function, and $X_{\rm norm}$ and $X_{\rm corr}$ are generated as per Algorithm 1. 
+    \sup_y|{\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
+\end{equation}
+where $S(x)$ is the standard logistic function, and $X_{\rm nc}$ and $X_{\rm corr}$ are generated as per Algorithm 1. 
 \end{corollary}
@@
 \begin{equation}
-    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
-\end{equation}
-and convolve with the distribution of $X_n$ {\color{blue} (Daniel: to be clear I
-think this was supposed to be $X_{\rm norm}$ instead of $X_n$ but see comments
-earlier about why I think it is not necessary.)} which has density
-$\phi(X/\sigma_n)$ where $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the
+    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
+\end{equation}
+and convolve with the distribution of $X_n$ which has density
+%% {\color{blue}
+%%   (Daniel: to be clear I
+%% think this was supposed to be $X_{\rm norm}$ instead of $X_n$ but see comments
+%% earlier about why I think it is not necessary.)}
+$\phi(X/\sigma_n)$ where $\sigma_n^2 = 1 - s^2_{\Delta^*}$. This yields the
 next iteration of $P$ and $Q$:
 \begin{equation}
-    P'(y) = {\rm Pr}(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
+    P'(y) = {\rm Pr}(\Delta^*+X_{\rm nc} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
 \end{equation}
@@
 \begin{equation}
-    P''(y) = {\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
+    P''(y) = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
 \end{equation}
@@
 that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
-norm}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
+nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
 log} > 0)$ by at most $\epsilon$.
@@
 
-{\color{blue} Daniel: Do we want to include this new analysis? There seem to be
-two different ways we show the bound. The first part (above) is nice and I like
-it (and it seems to show all we need for Algorithm 1, doesn't it?}
+%% {\color{blue} Daniel: Do we want to include this new analysis? There seem to be
+%% two different ways we show the bound. The first part (above) is nice and I like
+%% it (and it seems to show all we need for Algorithm 1, doesn't it?}
 
@@
 Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
-$n^{1/2}$. {\color{blue} (Daniel: how did the equation in the next sentence
-appear? I verified it and it seems correct, but it looks like it popped out of
-nowhere.)} Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where
+$n^{1/2}$.
+%{\color{blue} (Daniel: how did the equation in the next sentence
+% appear? I verified it and it seems correct, but it looks like it popped out of
+%nowhere.)}
+% JFC: It comes from taking a second derivative. You dont need to explain it. 
+Importantly, since $\phi^{''}(x) = x^2\phi(x) - \phi(x)$ where
 $\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
@@
 correct for the difference in $\sigma$ between $\Delta^*$ and the variance
-needed by $X_{\rm corr}$ using $X_{\rm norm}$. This same method works when we
+needed by $X_{\rm corr}$ using $X_{\rm nc}$. This same method works when we
 wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
@@
 $\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
-variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm norm}$ as
+variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm nc}$ as
 per Algorithm 1 preserving the limit in Equation~\ref{eq:GNderivatives}.
@@
 table. We leave further analysis and experiments with $g(x,y)$ as future work.
-{\color{blue} Daniel: I will look at the code to understand this paragraph
-better because right now I am confused.}
+%{\color{blue} Daniel: I will look at the code to understand this paragraph
+%better because right now I am confused.}
+% JFC: the code doesnt implement this yet.
 
@@
 \end{equation}
-We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta =
-(0,1)$, we draw 1,000,000 data points so that the target distribution is
+We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
+(0,1)$. The original paper sampled 100 data points form this distribution,
+and estimated the posterior in the parameters. We are interested in performance
+on larger problems and so sampled 1,000,000 points. The posterior
 $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with the prior based on the
-$\theta$ generation process in Equation~\ref{eq:data_generation}. This results
-in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we
-apply a temperature $T=10,000$ to reduce ${\rm std}(\Delta')$. Taking logs, we
+$\theta$ generation process in Equation~\ref{eq:data_generation}. The larger
+number of samples produces a much sharper posterior with two very narrow peaks.
+Our goal is to reproduce the original posterior, so we adjust the temperature
+to $T=10,000$. Taking logs, we
 get the target as shown in the far left of Figure~\ref{fig:gauss_mix_1}.
-{\color{blue} Daniel: the ${\rm std}(\Delta')$ stuff is outdated; let's figure
-out a way to change this. We may also have to justify such a high temperature,
-because this is far from the ``real'' (log) posterior.}
-
-We run MCMC with our MH test using minibatch size $m=100$.  We also run this
-using the methods from Korattikara 2014~\cite{cutting_mh_2014} and
-Bardenet 2014~\cite{icml2014c1_bardenet14}. For~\cite{cutting_mh_2014}, we use
-$m=100$ and increment minibatches by that amount within a test if necessary, and
-for~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the minibatch size
-geometrically with a ratio of $\gamma = 1.5$. The tolerance for making a
-decision in~\cite{cutting_mh_2014} is $\epsilon=0.005$, and the error bound
-control parameters in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta =
-0.01$. To make comparisons easier, all three use the same random walk proposer
-with covariance $\Sigma={\rm diag}(0.3, 0.3)$. This is a poor proposer, but it
-is sufficient for our purposes as the quality of the samples will be primarily
-due to the MH test. All methods are run 5000 times to collect 5000 samples.
+%% {\color{blue} Daniel: the ${\rm std}(\Delta')$ stuff is outdated; let's figure
+%% out a way to change this. We may also have to justify such a high temperature,
+%% because this is far from the ``real'' (log) posterior.}
+
+We run MCMC with our MH test using minibatch size $m=100$.  We also
+run this using the methods from Korattikara
+2014~\cite{cutting_mh_2014} and Bardenet
+2014~\cite{icml2014c1_bardenet14}. For~\cite{cutting_mh_2014}, we use
+$m=100$ and increment minibatches by 100 as recommended for that test,
+and for~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the
+minibatch size geometrically with a ratio of $\gamma = 1.5$ as
+recommended. These tests use different minibatch size updates because
+of their different approaches to repeated
+testing. \cite{icml2014c1_bardenet14} uses straightforward
+discounting, allocating a fraction of the test error $\delta$ to each
+minibatch test (call these subtests). Its important to bound the
+number of subtests therefore, so that the tolerable error per subtest is
+not too small. \cite{cutting_mh_2014} uses a more aggressive repeated testing
+approach, explicitly modeling correlation between tests and allowing many
+more tests for a given $\delta$.
+
+The tolerance for making a decision in~\cite{cutting_mh_2014} is
+$\epsilon=0.005$, and the error bound control parameters
+in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.
+all three use the same random walk proposer
+with covariance $\Sigma={\rm diag}(0.3, 0.3)$. Since the proposer
+generates a random walk, all shaping of distribution of the samples
+is due to the M-H test. All methods are run 5000
+times to collect 5000 samples.
 
@@
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
-	\caption{
-    Value of first and second part of equation (9)
-    in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size.
-    {\color{blue} Daniel: probably debatable on whether we should include this.
-    Let's talk it over.}
-    }
-	\label{fig:gauss_bardenet}
-\end{figure}
+%% \begin{figure}[t]
+%% 	\centering
+%% 	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
+%% 	\caption{
+%%     Value of first and second part of equation (9)
+%%     in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size.
+%%     {\color{blue} Daniel: probably debatable on whether we should include this.
+%%     Let's talk it over.}
+%%     }
+%% 	\label{fig:gauss_bardenet}
+%% \end{figure}
 
@@
 samples for the three methods, with darker regions indicating a greater density
-of points. The three methods obtain the same rough form of the posterior, so our
-MH test can indeed result in similar posteriors as the other two methods.
-
-In addition, our method is superior based on the Poisson likelihood of the
-MCMC-generated parameters over the range of $\theta_1 \in (-1.5, 2.5)$ and
-$\theta_2 \in (-2.5, 2.5)$. This is a measure of divergence from the
-ground-truth distribution. The log likelihood values for our method, Korattikara
-2014, and Bardenet 2014 are -2144.6, -4007.4, and -2598.4, respectively.
-{\color{blue} Daniel: I rewrote this paragraph to make it slightly less
-offensive. Also, should we clarify what this means? Because to be honest, it
-seems hard to tell a significant difference among the methods by looking at the
-scatterplots.}
+of points. There are no obvious differences in these distributions. To
+better understand any differences, we measure similarity between each set
+of samples and the actual posterior.
+
+To effect the comparison, we first discretize the posterior
+coordinates into bins (non-overlapping ranges) in $\theta_1$ and
+$\theta_2$. The probability of a sample falling into one of these bins
+is the integral of the true posterior probability over the area of
+that bin. Let this probability be denoted $P_i$ where $i$ is the bin
+number. A single sample from any of the M-H methods should therefore
+be multinomial with distribution $P$, and $n$ samples (which should
+ideally be independent) should follow $\sim {\rm Multinomial}(P,n)$.
+Since the ideal distribution is simple, we can use it to measure the
+quality of the sample distributions rather than general purpose tests
+like KL-divergence or likelihood-ratio.  The latter in any case are
+problematic with zero counts in some bins as is the case here.
+
+In practice its difficult to compute multinomial probabilities for
+large $n$, and in that case the per-bin distributions are
+well-approximated by Poisson distributions with parameter
+$\lambda_i=P_i n$ in each bin. Given a set of samples $X_i$, let
+$c_j$ denote the number of samples that fall in bin $j$:
+$$c_j = |\{j : Bin(X_i) = j\}|$$
+Then the log probability for the sample is
+\begin{equation}
+  \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1))
+\end{equation}
+
+The results for each model are shown in table 2...
+
+While these results provide general guidance on the relative accuracy of the models
+its difficult interpret the quantitative scores. We can also perform significance tests to
+test the null hypothesis that the ideal distribution really generated the samples.
+We performed two tests:
+\begin{enumerate}
+  \item An exact test using simulation to generate a (approximate) distribution
+    of scores given the null hypothesis.
+  \item An approximate test using the chi-squared distribution
+    as the test statistic. 
+\end{enumerate}
+the results were...
 
 Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used
-by the three methods in each iteration. Our methods consumes significantly less
-data; most minibatch sizes are smaller than 1000, and the average size is 420.
+by the three methods in each iteration. Our method consumes significantly less
+data; most minibatch sizes are smaller than 1000, and the average size is 210.
 The other two methods occasionally need to consume a large proportion of the
-entire data set of one million elements; the average minibatch sizes are 16378
-and 16755 for Korattikara 2014 and Bardenet 2014, respectively. {\color{blue}
-Daniel: the plot says our average minibatch size is 210, but the paragraph here
-says 420.}
-
-It is useful to notice that, in~\cite{icml2014c1_bardenet14}, they need to
-evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set
-at every iteration in order to calculate the error bound in equation (9)
-in~\cite{icml2014c1_bardenet14}. In Figure ~\ref{fig:gauss_bardenet}, the value
-of the first part and second part of equation (9)
-in~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual
-minibatch data used. The second part which contains  the $C_{\theta,\theta'}$
-term is larger than the first part, which means that the second part is
-non-negligible and has to be calculated at every iteration. This is one of the
-very time-consuming parts in this method.  {\color{blue} Daniel: I am not sure
-if we should include this plot. Let me think about it and if we decide to
-include it I will help rewrite the explanation.}
+entire data set of one million elements; the average minibatch sizes are 14862
+and 18098 for Korattikara 2014 and Bardenet 2014, respectively. The average
+minibatch sizes accurately predict the running times of these methods since
+all methods have a running time proportional to the total data consumed, with
+one caveat:
+
+The method of~\cite{icml2014c1_bardenet14} requires a global bound on
+the term $C_{\theta,\theta'}$ over the entire the data set at every
+iteration in order to calculate the error bound (in equation (9) of
+\cite{icml2014c1_bardenet14}).  This bound must be either computed
+analytically using properties of the specific dataset, or it must be
+checked at each sample if a black-box test is required. We found that
+the sample code provided by the authors
+of~\cite{icml2014c1_bardenet14} in fact computed this bound explicitly
+for each sample generated. i.e. the implementation traversed the
+entire dataset for each sample, providing no performance advantage
+over the complete test. For this reason, we omitted the time to
+compute $C_{\theta,\theta'}$ from our measurements of the running time
+of the sample implementation of \cite{icml2014c1_bardenet14}, and
+consider only the time to process the minibatches.
+
+%% In Figure ~\ref{fig:gauss_bardenet}, the value
+%% of the first part and second part of equation (9)
+%% in~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual
+%% minibatch data used. The second part which contains  the $C_{\theta,\theta'}$
+%% term is larger than the first part, which means that the second part is
+%% non-negligible and has to be calculated at every iteration. This is one of the
+%% very time-consuming parts in this method.  {\color{blue} Daniel: I am not sure
+%% if we should include this plot. Let me think about it and if we decide to
+%% include it I will help rewrite the explanation.}
+
+%% JFC - it would still be interesting to see this comparison, however:
+%% (i) its important to include equation (9) and explain what the terms are.
+%% (ii) the results dont look right. I wouldnt expect the first term to be
+%% so small compared to the second and I'm fairly certain they should cross
+%% in the illustrated range. It would be better to include both curves on
+%% the same plot.
 
@@
 with~\cite{icml2014c1_bardenet14} with error bound control parameters $p = 2$
-and $\delta = 0.01$. {\color{blue} Daniel: just observed this, but why are we
-using a subset of the test data? We really should use all 2000 (I think it's
-2000, right?).}
+and $\delta = 0.01$.
+% {\color{blue} Daniel: just observed this, but why are we
+% using a subset of the test data? We really should use all 2000 (I think it's
+% 2000, right?).}
+% JFC- Would be good to fix, but not critical.
 
@@
 on the test set as a function of the cumulative training data points processed.
-Our test increments the cumulative data by a fixed amount per iteration
-{\color{blue} (Daniel: are we sure about ``fixed amount per iteration'' because
-our Algorithm 1 implies we may have to increment the minibatch size, but it
-happens \emph{less frequently} than the other methods.)}, but the other two
-methods may require more data per iteration. We see that our minibatch MH test
+We see that our minibatch MH test
 is more efficient; it has similar or better performance while consuming fewer
@@
 Figure~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes for
-all three methods. With an initial minibatch size of 100, Korattikara 2014 and
-Bardenet 2014 achieve average minibatch size of 585 and 2731 respectively over
+all three methods. With an initial minibatch size of 163, Korattikara 2014 and
+Bardenet 2014 achieve average minibatch size of 2025 and 7612 respectively over
 the MNIST classification task, while our method achieves that of 100, showing
-significant better performance than the benchmark test methods. {\color{blue}
-Daniel: these numbers don't match the figure titles.}
+significant better performance than the benchmark test methods.
 

### 2016-09-30 nips_2016.tex
Minor changes to improve John's edits. Now going to work on experiments FINALLY

@@
   minibatch size $k$. Our test is adaptive like earlier works, but
-  unlike them, the distribution of minibatch size is gaussian, not
+  unlike them, the distribution of minibatch size is Gaussian, not
   long-tailed.  Increased minibatch size also reduces the error rate
   for the test.
-
-%% {\color{blue} Daniel: I'm not sure if I'd mention this here,
-%%    because before this list, we talk about reducing the data, and
-%%    then suddenly we're talking about increasing it.}
 
@@
 the test repeats. This process continues until a decision. The bounds depend on
-either a asymptotic central limit theorem \cite{cutting_mh_2014} or a
-concentration bound \cite{icml2014c1_bardenet14}. The latter requires direct
+either a asymptotic Central Limit Theorem~\cite{cutting_mh_2014} or a
+concentration bound~\cite{icml2014c1_bardenet14}. The latter requires direct
 bounds on the log likelihood ratio, which for general distributions requires
@@
 although implementations on typical large datasets will approximate
-  this by sampling without replacement}.
-%% {\color{blue} Daniel: should we say sampling with or without replacement? Right
-%%now I think the code does it without replacement (but \emph{with} replacement
-%% means it's closer to the CLT, right?} 
+this by sampling without replacement.}.
 By the Central Limit Theorem, we expect $\Lambda^*(\theta,\theta')$ to
@@
 $\Lambda^*(\theta',\theta)$ and $\psi(u,\theta',\theta)$.
-% {\color{blue} Daniel: I changed the variance to be $(\theta'-\theta)^2N^2/b$; before it had
-% $N$ instead of $N^2$.}
 
@@
 worst-case scenario is specified in Lemma~\ref{lem:worst_case}.
-%{\color{blue} Daniel: this might be a silly question, but why do we assume the
-%proposal has mean $\theta'-\theta$? I thought it would have a mean of $\theta$
-%because that is the current sample and the Gaussian is centered at that point?
-%That will also clarify my confusion about the $\psi$ function.} - fixed
 
@@
 to derive a highly accurate representation of $C_{\sigma}(X)$.
-Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
-$X_{\rm corr}$ densities that sum to $X_{\rm log}$.
 
@@
 
+\section{Computing the Correction Distribution}\label{sec:correction}
+
 \begin{figure}[t]
@@
     Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
-    convolve to form the standard logistic distribution. We use three standard
-    deviation values of $X_{\rm norm}$. The two red curves convolve to form the
-    logistic, etc. The $y$-axis is capped at 0.5 for readability. This figure
-    must be viewed in color. {\color{blue} Daniel: TODO Change figure to use
-    LaTeX for the titles.}
+    convolve to form the standard logistic distribution. These were obtained
+    based on our original linear programming formulation, as discussed in
+    Section~\ref{sec:correction}. We use three standard deviation values of
+    $X_{\rm norm}$. The two red curves convolve to form the logistic, etc. The
+    $y$-axis is capped at 0.5 for readability. This figure must be viewed in
+    color.
     }
@@
 \end{figure}
-
-\section{Computing the Correction Distribution}\label{sec:correction}
 
@@
 where $*$ denotes convolution.
-% {\color{blue} Daniel: This is very unusual use of
-%convolutions (i.e., CDF * DENSITY = CDF). It's valid but I was very surprised to
-% know that.}
+
 For computation of $C_\sigma$, we assume that its input $Y$ and
 another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
-% {\color{blue} Daniel: I am a little confused at the different interval sizes.}
 In continuous form, the optimization problem is therefore
@@
 \end{equation}
-% {\color{blue} Daniel: Equation~\ref{eq:optimization_l1} is finding
-% the $C_\sigma$ minimizing the $L_{\infty}$ norm, but the previous one,
-% Equation~\ref{eq:conv2}, is finding the $C_\sigma$ that is minimizing the
-% $L_\infty$ norm, i.e., the maximum value! If we want this to be the $L_{\infty}$ norm
-% (and I think we do) then Equation~\ref{eq:conv2} needs an outer sum over the $i$
-% terms. In addition, Equation~\ref{eq:conv1} also needs to be modified, perhaps
-% eliminate the ``sup'' but keep $x \in [-2V,2V]$}
 
@@
 represents a density. We first explored optimizing this problem with linear
-programming to find a $u$ minimizing $|Mu - v|$ subject to $u \ge 0$. This was
-tractable up to a few hundred dimensions for $u$. However, the discretization
-error is bounded by the curvature of the underlying functions times the squared
-quantization step. The curvatures are here
-slightly less than one, i.e.  the errors are of the order of $(V/N)^2$.
-% {\color{blue} Daniel: unfortunately, I do not understand what this means (i.e.,
-% the curvature and how the $(V/N)^2$ is obtained).} Here we chose $V=20$ to
+programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
+Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
+$X_{\rm corr}$ densities that sum to $X_{\rm log}$ from this process.
+
+This was tractable up to a few hundred dimensions for $u$.  However, the
+discretization error is bounded by the curvature of the underlying functions
+times the squared quantization step. The curvatures are here slightly less than
+one, i.e.  the errors are of the order of $(V/N)^2$.  Here we chose $V=20$ to
 provide adequate containment of the distributions (the CDFs are extremely close
@@
 some flexibility in the choice of $\sigma$ in Equation~\ref{eq:conv1}.
-% {\color{blue} Daniel: is this the correct equation?}
-As we decrease the variance
-of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount and
-is in fact the result of convolution with a Gaussian whose variance is the
-difference.  Thus as $\sigma$ decreases, $C(X)$ grows and approaches the
-derivative of a logistic function at $\sigma = 0$. It retains some very weak
-negative values for $\sigma > 0$ but removal of those values leads to very small
-error.
+As we decrease the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$
+grows by the same amount and is in fact the result of convolution with a
+Gaussian whose variance is the difference.  Thus as $\sigma$ decreases,
+$C_\sigma(X)$ grows and approaches the derivative of a logistic function at
+$\sigma = 0$. It retains some very weak negative values for $\sigma > 0$ but
+removal of those values leads to very small error.
 
@@
     Propose a candidate $\theta'$\ from proposal distribution $q(\theta'\mid\theta_t)$\;
-    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*})$\;
+    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*}$\;
     Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Equation~\ref{eq:clt-bounds3}\;
@@
 }
-\caption{A description of M-H sampling with our acceptance test.
-% {\color{blue}
-% Daniel: I think we may want to be careful about the use of $C_1(X)$ because that
-% implies that we have a fixed $X_{\rm corr}$ density, but that $C_1(X)$ will
-% change depending on the iteration and variance of $\Delta^*$, right? Hopefully
-% not by much, but it still changes sometimes? ALSO bring up Xinlei's old points
-  %about $X_{\rm norm}$ confusion.}
-}
+\caption{A description of M-H sampling with our acceptance test.}
 \label{alg:our_algorithm}
@@
 
-\begin{table}[h]
+\begin{table}[t]
     \caption{Error in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$}
@@
 
-As can be seen from table~\ref{tab:xcorr}, the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be made very small, approaching single floating precision (about $10^{-7}$).
-
-The complete algorithm is given in~\ref{alg:our_algorithm}. A few points:
+As can be seen from Table~\ref{tab:xcorr}, the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be made very small, approaching single floating precision (about $10^{-7}$).
+
+The complete procedure is given in Algorithm~\ref{alg:our_algorithm}. A few points:
+
 \begin{itemize}
-\item The algorithm uses an adaptive step size so as to use the smallest possible
-  average minibatch size. Unlike previous work however (and see the experiments section)
-  the size distribution is short-tailed.
-\item An additional normal variable $X_{\rm nc}$ is added to $\Delta^*$ to produce a variable
-  with unit variance. This is not mathematically necessary, but allows us to use a single
-  correction distribution $C_1$ for $X_{\rm corr}$, saving on memory footprint.
+    \item It uses an adaptive step size so as to use the smallest possible
+    average minibatch size. Unlike previous work however (and as we show in
+    Section~\ref{sec:experiments}) the size distribution is short-tailed.
+
+    \item An additional normal variable $X_{\rm nc}$ is added to $\Delta^*$ to
+    produce a variable with unit variance. This is not mathematically necessary,
+    but allows us to use a single correction distribution $C_1$ with $\sigma=1$
+    for $X_{\rm corr}$, saving on memory footprint.
+
+    \item The sample variance $s^2_{\Delta^*}$ is proportional to
+    $\|\theta'-\theta\|_2^2$ whose distribution for Normal proposals is the
+    square of a normal variable. 
 \end{itemize}
-
-The sample variance $s^2_{\Delta^*}$ is proportional to $||\theta'-\theta||^2$
-whose distribution for Normal proposals is the square of a normal variable. 
-% {\color{blue} Daniel: TODO improve/describe Table~\ref{tab:xcorr} and try to
-% relocate Algorithm~\ref{alg:our_algorithm} to be somewhere nicer.}
 
@@
 it is possible to generate the correction samples $X_{\rm corr}$ with a CDF
-error approaching single-precision floating point error.
-% {\color{blue} (Daniel: I can fix this to cite Table~\ref{tab:xcorr}'s values,
-% that will probably be better.)}.
-We therefore treat the $X_{\rm corr}$ variable as a sample from the
-exact correction distribution and we will not analyze its errors.
+error approaching single-precision floating point error. We therefore treat the
+$X_{\rm corr}$ variable as a sample from the exact correction distribution and
+we will not analyze its errors.
 
@@
 
-%{\color{blue} Daniel: the following lemma is, I think, from Corollary 2
-%in~\cite{explicit-clt05} but I cannot derive the exact lemma as we have it
-%here.} To start, we use the following quantitative central-limit result
-%(from~\cite{explicit-clt05}):
+To start, we use the following quantitative central-limit result:
 
@@
 which approaches a normal distribution in the limit. Then
-from~\cite{explicit-clt05}:
 \begin{equation}\label{eq:clt-bounds}
@@
 and combining with $x>0$, we obtain the weaker but unqualified bound
-in equation (\ref{eq:clt-bounds}).
+in Equation~\ref{eq:clt-bounds}.
 \end{proof}
@@
 approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of
-variables $y = x s_X)$, Equation~\ref{eq:clt-bounds} becomes
+variables $y = x s_X$, Equation~\ref{eq:clt-bounds} becomes
 \begin{equation}\label{eq:clt-bounds2}
@@
 \end{equation}
-and convolve with the distribution of $X_n$ which has density
-%% {\color{blue}
-%%   (Daniel: to be clear I
-%% think this was supposed to be $X_{\rm norm}$ instead of $X_n$ but see comments
-%% earlier about why I think it is not necessary.)}
-$\phi(X/\sigma_n)$ where $\sigma_n^2 = 1 - s^2_{\Delta^*}$. This yields the
-next iteration of $P$ and $Q$:
+and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
+where $\sigma_n^2 = 1 - s^2_{\Delta^*}$. This yields the next iteration of $P$
+and $Q$:
 \begin{equation}

### 2016-09-30 nips_2016.tex
Updated Figure 3 with accompanying discussion, added table for later usage

@@
     \centering
-    \includegraphics[width=1\linewidth]{GaussianMixtureResult/minibatch_size_gaussian.png}
+    \includegraphics[width=1\linewidth]{minibatch_size_gaussian.png}
     \caption{
     Histograms of minibatch sizes for the three methods used in
-    Section~\ref{ssec:gaussians}. {\color{blue} Daniel: The y-axis should
-    probably just be ``Counts'' not ``Total Counts.'' Also, was this run with
-    the \texttt{plt.tight\_layout()} command? The y-axis text for the second and
-    third subplots almost overlap with the graphs to their left. Finally, I am
-    still thinking about whether the y-axis should be on a log scale. I think
-    the log scale is appropriate for the x-axis, it's a question of whether we
-    should also apply it for the y-axis.}
+    Section~\ref{ssec:gaussians}. Note the log-log scale, to facilitate
+    comparisons across methods; in particular, the axes have the same
+    range.
     }
@@
 \end{figure}
-
-%% \begin{figure}[t]
-%% 	\centering
-%% 	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
-%% 	\caption{
-%%     Value of first and second part of equation (9)
-%%     in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size.
-%%     {\color{blue} Daniel: probably debatable on whether we should include this.
-%%     Let's talk it over.}
-%%     }
-%% 	\label{fig:gauss_bardenet}
-%% \end{figure}
 
@@
 
-The results for each model are shown in table 2...
+The results for each model are shown in Table~\ref{tab:poissons} ...
+{\color{blue} Daniel: IN PROGRESS}
+
+\begin{table}[t]
+    \caption{{\color{blue} Daniel: TODO Poisson Probabilities}}
+    \label{tab:poissons}
+    \centering
+    \begin{tabular}{l l l l}
+    \toprule
+    Model & Ours & Korattikara 2014 & Bardenet 2014 \\
+    \midrule
+    Poission Value & TODO & TODO & TODO \\
+    \bottomrule
+    \end{tabular}
+\end{table}
 

### 2016-09-30 nips_2016.tex
Revised Figure 5, updated captions. Also gitignore to make our lives easier

@@
     \caption{
-    The log posterior contours and scatter plots of sampled $\theta$ values using different methods. 
+    The log posterior contours and scatter plots of sampled $\theta$ values
+    using different methods. 
     }
@@
     \caption{
-    Histograms of minibatch sizes for the three methods used in
-    Section~\ref{ssec:gaussians}. Note the log-log scale, to facilitate
-    comparisons across methods; in particular, the axes have the same
-    range.
+    Counts of minibatch sizes that the three methods used in the experiment of
+    Section~\ref{ssec:gaussians}. To make comparisons easier, the axes have the
+    same range (using a log-log scale).
     }
@@
 	\centering
-	\includegraphics[width=1\linewidth]{LogisticRegressionResult/minibatch_size_logistic.png}
+	\includegraphics[width=1\linewidth]{minibatch_size_logistic.png}
 	\caption{
-    Counts of minibatch sizes needed for the three methods in the MNIST
-    classification task.
+    Counts of minibatch sizes that the three methods used in the logistic
+    regression experiment from Section~\ref{ssec:logistic}. It is analogous to
+    Figure~\ref{fig:gauss_mix_2}.
     }

### 2016-09-30 nips_2016.tex
Updated Figure with logistic regression curves

@@
 
-Figure~\ref{fig:logistic_fig} shows the prediction accuracy and log likelihood
-on the test set as a function of the cumulative training data points processed.
-We see that our minibatch MH test
-is more efficient; it has similar or better performance while consuming fewer
-data points.
+Figure~\ref{fig:logistic_performance} shows the prediction accuracy and log
+likelihood on the test set as a function of the cumulative training data points
+processed.  We see that our minibatch MH test is more efficient; it has similar
+or better performance while consuming fewer data points.
 
@@
 	\centering
-    \begin{subfigure}{.4\textwidth}
-     \centering
-     \includegraphics[width=1\linewidth]{LogisticRegressionResult/acc_logistic}
-    \end{subfigure}
-    \begin{subfigure}{.4\textwidth}
-     \centering
-    \includegraphics[width=1\linewidth]{LogisticRegressionResult/ll_logistic}
-\end{subfigure}
+	\includegraphics[width=1\linewidth]{logistic_performance.png}
 	\caption{
-    Logistic regression performance (accuracy/log likelihood) and minibatch size
-    analysis. {\color{blue} Daniel: let's use matplotlib's ``subplot''
-    functionality, so that the images appear the same size. We can handle this
-    later; it will be easy.}
+    Logistic regression performance (accuracy/log likelihood) based on
+    cumulative data usage.
     }
-\label{fig:logistic_fig}
+	\label{fig:logistic_performance}
 \end{figure}
@@
 	\caption{
-    Counts of minibatch sizes that the three methods used in the logistic
+    Counts of minibatch sizes that the three methods used in the MNIST logistic
     regression experiment from Section~\ref{ssec:logistic}. It is analogous to

### 2016-09-30 nips_2016.tex
Touch ups to Section 6.1, now will read the Poisson process in detail

@@
 
-As can be seen from Table~\ref{tab:xcorr}, the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be made very small, approaching single floating precision (about $10^{-7}$).
+As can be seen from Table~\ref{tab:xcorr}, the errors between $X_{\rm
+norm}+X_{\rm corr}$ and $X_{\rm log}$ can be made very small, approaching single
+floating precision (about $10^{-7}$).
 
@@
 propagate directly through as errors in the acceptance test.  To achieve a test
-error of say $1e-6$ (close to single floating point precision), we need a $y'$
-spacing of $1e-6$. It should not be necessary to tabulate values all the way to
+error of $10^{-6}$ (close to single floating point precision), we need a $y'$
+spacing of $10^{-6}$. It should not be necessary to tabulate values all the way to
 $y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
@@
 Our goal is to reproduce the original posterior, so we adjust the temperature
-to $T=10,000$. Taking logs, we
-get the target as shown in the far left of Figure~\ref{fig:gauss_mix_1}.
-%% {\color{blue} Daniel: the ${\rm std}(\Delta')$ stuff is outdated; let's figure
-%% out a way to change this. We may also have to justify such a high temperature,
-%% because this is far from the ``real'' (log) posterior.}
+to $T=10,000$. Taking logs, we get the target as shown in the far left of
+Figure~\ref{fig:gauss_mix_1}.
 
 We run MCMC with our MH test using minibatch size $m=100$.  We also
-run this using the methods from Korattikara
-2014~\cite{cutting_mh_2014} and Bardenet
-2014~\cite{icml2014c1_bardenet14}. For~\cite{cutting_mh_2014}, we use
-$m=100$ and increment minibatches by 100 as recommended for that test,
-and for~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the
-minibatch size geometrically with a ratio of $\gamma = 1.5$ as
-recommended. These tests use different minibatch size updates because
-of their different approaches to repeated
-testing. \cite{icml2014c1_bardenet14} uses straightforward
-discounting, allocating a fraction of the test error $\delta$ to each
-minibatch test (call these subtests). Its important to bound the
-number of subtests therefore, so that the tolerable error per subtest is
-not too small. \cite{cutting_mh_2014} uses a more aggressive repeated testing
-approach, explicitly modeling correlation between tests and allowing many
-more tests for a given $\delta$.
+run this using the methods from~\cite{cutting_mh_2014} (``Korattikara 2014'')
+and~\cite{icml2014c1_bardenet14} (``Bardenet 2014'').
+For~\cite{cutting_mh_2014}, we use $m=100$ and increment minibatches by 100 as
+recommended for that test, and for~\cite{icml2014c1_bardenet14}, we use $m=100$
+and increase the minibatch size geometrically with a ratio of $\gamma = 1.5$ as
+recommended. These tests use different minibatch size updates because of their
+different approaches to repeated testing. \cite{icml2014c1_bardenet14} uses
+straightforward discounting, allocating a fraction of the test error $\delta$ to
+each minibatch test (call these subtests). Its important to bound the number of
+subtests therefore, so that the tolerable error per subtest is not too small.
+\cite{cutting_mh_2014} uses a more aggressive repeated testing approach,
+explicitly modeling correlation between tests and allowing many more tests for a
+given $\delta$.
 
@@
 $\epsilon=0.005$, and the error bound control parameters
-in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.
-all three use the same random walk proposer
-with covariance $\Sigma={\rm diag}(0.3, 0.3)$. Since the proposer
-generates a random walk, all shaping of distribution of the samples
-is due to the M-H test. All methods are run 5000
+in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.  All methods
+(including ours) use the same random walk proposer with covariance $\Sigma={\rm
+diag}(0.3, 0.3)$.  Since the proposer generates a random walk, all shaping of
+distribution of the samples is due to the M-H test. All methods are run 5000
 times to collect 5000 samples.
@@
 
-In this paper, we have derived a new MH test for minibatch MCMC methods. We demonstrated how a
-simple deconvolution process allows us to use a minibatch approximation to the full data tests. We
-experimentally show the benefits of our test on Gaussian mixtures and logistic regression.
-Straightforward directions for future work include running more experiments with a
-particular focus on investigation of the variance precondition.  More elaborate extensions include
-combining our results with Hamiltonian Monte Carlo methods, providing a recipe for how to use our
-algorithm (following the framework of~\cite{sgmcmc_2015}), or integrating parallel
+In this paper, we have derived a new MH test for minibatch MCMC methods. We
+demonstrated how a simple deconvolution process allows us to use a minibatch
+approximation to the full data tests. We experimentally show the benefits of our
+test on Gaussian mixtures and logistic regression.  Straightforward directions
+for future work include running more experiments with a particular focus on
+investigation of the variance precondition.  More elaborate extensions include
+testing on deep neural networks and combining our results with Hamiltonian Monte
+Carlo methods, providing a recipe for how to use our algorithm (following the
+framework of~\cite{sgmcmc_2015}), or integrating parallel
 MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.

### 2016-09-30 nips_2016.tex
Touch-ups and confirmed the validity of Poisson-Multinomial stuff

@@
 from~\cite{langevin_2011}.  The parameter is 2-D, $\theta =
-(\theta_1,\theta_2)$, and the parameter/data generation process is
+\langle \theta_1,\theta_2 \rangle$, and the parameter/data generation process is
+%% \begin{equation}\label{eq:data_generation}
+%%     (\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
+%%     0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
+%% \end{equation}
 \begin{equation}\label{eq:data_generation}
-    (\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
+    \begin{bmatrix} \theta_1 \\ \theta_2 \end{bmatrix} 
+    \sim \mathcal{N}\left(\begin{bmatrix}0 \\ 0\end{bmatrix}, 
+    \begin{bmatrix} \sigma_1^2 & 0 \\ 0 & \sigma_2^2\end{bmatrix} \right);\quad \quad x_i \sim
     0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
@@
 We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
-(0,1)$. The original paper sampled 100 data points form this distribution,
-and estimated the posterior in the parameters. We are interested in performance
-on larger problems and so sampled 1,000,000 points. The posterior
-$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with the prior based on the
-$\theta$ generation process in Equation~\ref{eq:data_generation}. The larger
-number of samples produces a much sharper posterior with two very narrow peaks.
-Our goal is to reproduce the original posterior, so we adjust the temperature
-to $T=10,000$. Taking logs, we get the target as shown in the far left of
+\langle 0,1 \rangle$. The original paper sampled 100 data points from this
+distribution, and estimated the posterior in the parameters. We are interested
+in performance on larger problems and so sampled 1,000,000 points to form the
+posterior of $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$ with the same
+prior from Equation~\ref{eq:data_generation}. The larger number of samples
+produces a much sharper posterior with two very narrow peaks.  Our goal is to
+reproduce the original posterior, so we adjust the temperature to $T=10,000$.
+Taking logs, we get the target as shown in the far left of
 Figure~\ref{fig:gauss_mix_1}.
 
-We run MCMC with our MH test using minibatch size $m=100$.  We also
-run this using the methods from~\cite{cutting_mh_2014} (``Korattikara 2014'')
-and~\cite{icml2014c1_bardenet14} (``Bardenet 2014'').
-For~\cite{cutting_mh_2014}, we use $m=100$ and increment minibatches by 100 as
-recommended for that test, and for~\cite{icml2014c1_bardenet14}, we use $m=100$
-and increase the minibatch size geometrically with a ratio of $\gamma = 1.5$ as
-recommended. These tests use different minibatch size updates because of their
-different approaches to repeated testing. \cite{icml2014c1_bardenet14} uses
-straightforward discounting, allocating a fraction of the test error $\delta$ to
-each minibatch test (call these subtests). Its important to bound the number of
-subtests therefore, so that the tolerable error per subtest is not too small.
-\cite{cutting_mh_2014} uses a more aggressive repeated testing approach,
-explicitly modeling correlation between tests and allowing many more tests for a
-given $\delta$.
+We run MCMC with our MH test using minibatch size $m=100$.  We also run this
+using the methods from~\cite{cutting_mh_2014} (``Korattikara 2014'')
+and~\cite{icml2014c1_bardenet14} (``Bardenet 2014''), also with $m=100$.
+For~\cite{cutting_mh_2014}, we increment minibatches by 100 as recommended, and
+for~\cite{icml2014c1_bardenet14}, we increase the minibatch size geometrically
+with a ratio of $\gamma = 1.5$ as recommended. These tests use different
+minibatch size updates because of their different approaches to repeated
+testing. \cite{icml2014c1_bardenet14} uses straightforward discounting,
+allocating a fraction of the test error $\delta$ to each minibatch test (call
+these subtests). Its important to bound the number of subtests therefore, so
+that the tolerable error per subtest is not too small.  \cite{cutting_mh_2014}
+uses a more aggressive repeated testing approach, explicitly modeling
+correlation between tests and allowing many more tests for a given $\delta$.
 
@@
 diag}(0.3, 0.3)$.  Since the proposer generates a random walk, all shaping of
-distribution of the samples is due to the M-H test. All methods are run 5000
+distribution of the samples is due to the MH test. All methods are run 5000
 times to collect 5000 samples.
@@
 samples for the three methods, with darker regions indicating a greater density
-of points. There are no obvious differences in these distributions. To
-better understand any differences, we measure similarity between each set
-of samples and the actual posterior.
-
-To effect the comparison, we first discretize the posterior
-coordinates into bins (non-overlapping ranges) in $\theta_1$ and
-$\theta_2$. The probability of a sample falling into one of these bins
-is the integral of the true posterior probability over the area of
-that bin. Let this probability be denoted $P_i$ where $i$ is the bin
-number. A single sample from any of the M-H methods should therefore
-be multinomial with distribution $P$, and $n$ samples (which should
-ideally be independent) should follow $\sim {\rm Multinomial}(P,n)$.
-Since the ideal distribution is simple, we can use it to measure the
-quality of the sample distributions rather than general purpose tests
-like KL-divergence or likelihood-ratio.  The latter in any case are
-problematic with zero counts in some bins as is the case here.
-
-In practice its difficult to compute multinomial probabilities for
-large $n$, and in that case the per-bin distributions are
-well-approximated by Poisson distributions with parameter
-$\lambda_i=P_i n$ in each bin. Given a set of samples $X_i$, let
-$c_j$ denote the number of samples that fall in bin $j$:
-$$c_j = |\{j : Bin(X_i) = j\}|$$
-Then the log probability for the sample is
-\begin{equation}
-  \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1))
+of points. There are no obvious differences in these distributions. To better
+understand any differences, we measure the similarity between each set of
+samples and the actual posterior.
+
+To effect the comparison, we first discretize the posterior coordinates into
+bins (non-overlapping ranges) with respect to the two components of the
+parameter $\theta$.  The probability of a sample falling into one of these bins
+is the integral of the true posterior probability over the area of that bin. Let
+this probability be denoted $P_i$ where $i$ is the bin number. A single sample
+from any of the MH methods should therefore be multinomial with distribution
+$P$, and a set of $n$ (ideally independent) samples should follow a ${\rm
+Multinomial}(P,n)$ distribution.  Since the ideal distribution is simple, we can
+use it to measure the quality of the sample distributions rather than use
+general purpose tests like KL-divergence or likelihood-ratio.  The latter in any
+case are problematic with zero counts in some bins as we have here.
+
+In practice, it is difficult to compute multinomial probabilities for large $n$,
+in that case the per-bin distributions are well-approximated by Poisson
+distributions with parameter $\lambda_i=P_i n$. Given a set of samples $\Theta =
+\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
+$\theta_i$ that fall in bin $j$. Then the log probability for the set is
+approximated as
+\begin{equation}\label{eq:log_prob_poissons}
+    \log p(\Theta) \approx \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
 \end{equation}
@@
 
-\begin{table}[t]
-    \caption{{\color{blue} Daniel: TODO Poisson Probabilities}}
-    \label{tab:poissons}
-    \centering
-    \begin{tabular}{l l l l}
-    \toprule
-    Model & Ours & Korattikara 2014 & Bardenet 2014 \\
-    \midrule
-    Poission Value & TODO & TODO & TODO \\
-    \bottomrule
-    \end{tabular}
-\end{table}
-
-While these results provide general guidance on the relative accuracy of the models
-its difficult interpret the quantitative scores. We can also perform significance tests to
-test the null hypothesis that the ideal distribution really generated the samples.
-We performed two tests:
+While the Poisson results provide general guidance on the relative accuracy of
+the models it is difficult interpret the quantitative scores. We therefore
+perform significance tests to test the null hypothesis that the ideal
+distribution really generated the samples.  We perform two tests:
+
 \begin{enumerate}
-  \item An exact test using simulation to generate a (approximate) distribution
+    \item An exact test using simulation to generate an approximate distribution
     of scores given the null hypothesis.
-  \item An approximate test using the chi-squared distribution
-    as the test statistic. 
+
+    \item An approximate test using the chi-squared distribution as the test
+    statistic. 
 \end{enumerate}
-the results were...
+
+The results were... {\color{blue} Daniel: IN PROGRESS}
 
@@
 data; most minibatch sizes are smaller than 1000, and the average size is 210.
-The other two methods occasionally need to consume a large proportion of the
+The other methods occasionally need to consume a large proportion of the
 entire data set of one million elements; the average minibatch sizes are 14862
@@
 all methods have a running time proportional to the total data consumed, with
-one caveat:
-
-The method of~\cite{icml2014c1_bardenet14} requires a global bound on
-the term $C_{\theta,\theta'}$ over the entire the data set at every
-iteration in order to calculate the error bound (in equation (9) of
-\cite{icml2014c1_bardenet14}).  This bound must be either computed
-analytically using properties of the specific dataset, or it must be
-checked at each sample if a black-box test is required. We found that
-the sample code provided by the authors
-of~\cite{icml2014c1_bardenet14} in fact computed this bound explicitly
-for each sample generated. i.e. the implementation traversed the
-entire dataset for each sample, providing no performance advantage
-over the complete test. For this reason, we omitted the time to
-compute $C_{\theta,\theta'}$ from our measurements of the running time
-of the sample implementation of \cite{icml2014c1_bardenet14}, and
-consider only the time to process the minibatches.
+one caveat: the method of~\cite{icml2014c1_bardenet14} requires a global bound
+on the term $C_{\theta,\theta'}$ over the entire the data set at every iteration
+in order to calculate the error bound (in Equation~\ref{eq:bad_bound} of this
+paper, i.e., Equation 9 of~\cite{icml2014c1_bardenet14}).  This bound must be
+either computed analytically using properties of the specific dataset, or it
+must be checked at each sample if a black-box test is required.  We found that
+the sample code provided by the authors of~\cite{icml2014c1_bardenet14} in fact
+computed this bound explicitly for each sample generated, i.e. the
+implementation traversed the entire dataset for each sample, providing no
+performance advantage over the complete test. For this reason, we omitted the
+time to compute $C_{\theta,\theta'}$ from our measurements of the running time
+of the sample implementation of \cite{icml2014c1_bardenet14}, and consider only
+the time to process the minibatches.
 
@@
 %% the same plot.
+
+\begin{table}[t]
+    \caption{{\color{blue} Daniel: TODO Poisson Probabilities, also we want
+    ground truth right?}}
+    \label{tab:poissons}
+    \centering
+    \begin{tabular}{l l l l l}
+    \toprule
+    Model & Ground Truth & Ours & Korattikara 2014 & Bardenet 2014 \\
+    \midrule
+    Poission Value & TODO & TODO & TODO & TODO \\
+    \bottomrule
+    \end{tabular}
+\end{table}
 

### 2016-09-30 nips_2016.tex
More touch-ups to Section 6.2.

@@
 \begin{equation}\label{eq:log_prob_poissons}
-    \log p(\Theta) \approx \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
-\end{equation}
-
+    \log p(\Theta) \approx \frac{1}{N_{\rm bins}} \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
+\end{equation}
 The results for each model are shown in Table~\ref{tab:poissons} ...
@@
 one caveat: the method of~\cite{icml2014c1_bardenet14} requires a global bound
-on the term $C_{\theta,\theta'}$ over the entire the data set at every iteration
-in order to calculate the error bound (in Equation~\ref{eq:bad_bound} of this
-paper, i.e., Equation 9 of~\cite{icml2014c1_bardenet14}).  This bound must be
-either computed analytically using properties of the specific dataset, or it
-must be checked at each sample if a black-box test is required.  We found that
-the sample code provided by the authors of~\cite{icml2014c1_bardenet14} in fact
-computed this bound explicitly for each sample generated, i.e. the
-implementation traversed the entire dataset for each sample, providing no
-performance advantage over the complete test. For this reason, we omitted the
-time to compute $C_{\theta,\theta'}$ from our measurements of the running time
-of the sample implementation of \cite{icml2014c1_bardenet14}, and consider only
-the time to process the minibatches.
+on the term $C_{\theta,\theta'}$ (see Equation~\ref{eq:bad_bound}) over the
+entire the data set at every iteration in order to calculate the error bound.
+This bound must be either computed analytically using properties of the specific
+dataset, or it must be checked at each sample if a black-box test is required.
+We found that the sample code provided by the authors
+of~\cite{icml2014c1_bardenet14} in fact computed this bound explicitly for each
+sample generated, i.e. the implementation traversed the entire dataset for each
+sample, providing no performance advantage over the complete test. For this
+reason, we omitted the time to compute $C_{\theta,\theta'}$ from our
+measurements of the running time of the sample implementation
+of~\cite{icml2014c1_bardenet14}, and consider only the time to process the
+minibatches.
 
@@
 \begin{table}[t]
-    \caption{{\color{blue} Daniel: TODO Poisson Probabilities, also we want
-    ground truth right?}}
+    \caption{\color{blue} Daniel: TODO Poisson Probabilities}
     \label{tab:poissons}
     \centering
-    \begin{tabular}{l l l l l}
+    \begin{tabular}{l l l l}
     \toprule
-    Model & Ground Truth & Ours & Korattikara 2014 & Bardenet 2014 \\
+    Model & Ours & Korattikara 2014 & Bardenet 2014 \\
     \midrule
-    Poission Value & TODO & TODO & TODO & TODO \\
+    Poisson Value (Eq.~\ref{eq:log_prob_poissons}) & TODO & TODO & TODO \\
     \bottomrule
@@
 likelihood on the test set as a function of the cumulative training data points
-processed.  We see that our minibatch MH test is more efficient; it has similar
-or better performance while consuming fewer data points.
+processed.  We see that our minibatch MH test is more efficient compared to the
+other two methods, because it has similar or better performance while consuming
+fewer data points.
 
@@
 Figure~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes for
-all three methods. With an initial minibatch size of 163, Korattikara 2014 and
-Bardenet 2014 achieve average minibatch size of 2025 and 7612 respectively over
-the MNIST classification task, while our method achieves that of 100, showing
-significant better performance than the benchmark test methods.
+all three methods. With an initial minibatch size of 100, Korattikara 2014 and
+Bardenet 2014 achieve average minibatch size of 2026 and 7612 respectively over
+the MNIST classification task, while our method achieves that of 163, showing
+significantly better performance than the benchmark test methods.
 

### 2016-09-30 nips_2016.tex
Finally fixed Appendix B (the Gaussian mixtures). That took way too long. Up next, the Logistic Regression stuff

@@
 samples for the three methods, with darker regions indicating a greater density
-of points. There are no obvious differences in these distributions. To better
-understand any differences, we measure the similarity between each set of
-samples and the actual posterior.
+of points. There are no obvious differences, so to better understand any
+differences, we measure the similarity between each set of samples and the
+actual posterior.
 
@@
 
-{\color{blue} Daniel: Note that everything after
-Appendix~\ref{app:worst_case_proof} has yet to be revised. I don't know what we
-will include in them.}
-
 \section{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}
@@
 
-In this section, we go over the math details on the Gaussian mixture model problem borrowed
-from~\cite{langevin_2011}. Our parameter is a 2-D vector $\theta =
-(\theta_1, \theta_2)$, where
+In this section, we provide more details on Section~\ref{ssec:gaussians}.  Our
+parameter is a 2-D vector
+$\theta = \langle \theta_1, \theta_2 \rangle$, where
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{N}$ indicates the normal distribution (more generally, the multivariate normal). We
-consider the above as our prior. Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and
-$\sigma_2^2=1$, so the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
+and where $\mathcal{N}$ indicates the normal distribution (more generally, the
+multivariate normal). We consider the above as our prior.
+Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and $\sigma_2^2=1$, so
+the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
 log prior probability we endow on $\theta$ is
@@
 \end{equation}
-where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$. This means the log likelihood
-of a single data instance is
-\begin{equation}
-\log p(x_i \mid \theta) = \log\left(\frac{1}{4\sqrt{\pi}}\right) +
+where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$. This means
+the log likelihood of a single data instance is
+\begin{equation}
+\log p(x_i \mid \theta) = \log \frac{1}{4\sqrt{\pi}} +
 \log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
@@
 \end{equation}
-Here is the problem statement: given some number of conditionally independent data points $x_1, x_2,
-\ldots, x_N$ generated according to~(\ref{eq:x_points}), determine the posterior
-distribution of $\theta$:
+The official problem statement is as follows: given some number of conditionally
+independent data points $x_1, x_2, \ldots, x_N$ generated according
+to~(\ref{eq:x_points}), determine the posterior distribution of $\theta$:
 \begin{equation}\label{eq:log_post}
@@
 \end{equation}
-Alternatively, if there are too many data points, we may opt to instead pick a point estimate of
-$\theta$, generally the MAP estimate. (If $N$ is extremely large, it will cause the posterior to
-peak sharply at its modes, reducing distribution estimates to point estimates.) Note that in many
-cases, we will need to take a \emph{minibatch estimate} of~(\ref{eq:log_post}). In that case, the
-literature generally uses
-\begin{equation}\label{eq:scaling_factor}
-\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
-\end{equation}
-where we only use $n \ll N$ samples, but we must scale up the likelihood contribution by $N/n$. If we
-didn't add this scaling factor, then the contribution of the likelihood terms would be weaker.
-
-One technique we use is adding a \emph{temperature} to our distribution. In general, we will want to
-add $T > 1$ so that our posterior is $p(\theta)((\prod_{i=1}^n p(x_i\mid \theta))^{N/n})^{1/T}$,
-resulting in the log posterior of 
+Alternatively, depending on the application, we may opt to instead pick a point
+estimate\footnote{We do this for the logistic regression experiment, described
+in Section~\ref{ssec:logistic} and elaborated in Appendix~\ref{app:logistic}.}
+of $\theta$, such as the MAP estimate. (If $N$ is extremely large, it will cause
+the posterior to peak sharply at its modes, reducing distribution estimates to
+point estimates.)
+
+One technique we use to make our test more applicable is to add a temperature to
+the posterior. In general, we will add $T > 1$ so that the posterior is
+$p(\theta)(\prod_{i=1}^N p(x_i\mid \theta))^{1/T}$, resulting in the log
+posterior of 
 \begin{equation}\label{eq:log_prior_temp}
-\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
-\end{equation}
-which has the extra $1/T$ to decrease the scale factor. Equation~(\ref{eq:log_prior_temp}) is what
-we use for our experiments, because warmer distributions help us satisfy our ${\rm std}(\Delta') <
-1.2$ requirement.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=1\linewidth]{contour_v1}
-  \caption{The posterior distribution, from 50 to 100k samples, with temperature set at 1.}
-  \label{fig:contour1}
-\end{figure}
+\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\sum_{i=1}^N\log p(x_i \mid \theta).
+\end{equation}
+which has the extra $1/T$ to decrease the scale factor, so the variance scales
+as $1/T^2$.
+
+% Daniel: I decided that I don't like this figure anymore.
+% \begin{figure}[t]
+%   \centering
+%   \includegraphics[width=1\linewidth]{contour_v1}
+%   \caption{The posterior distribution, from 50 to 100k samples, with temperature set at 1.}
+%   \label{fig:contour1}
+% \end{figure}
+
 \begin{figure}[t]
@@
   \includegraphics[width=1\linewidth]{contour_v2}
-  \caption{The posterior distribution, with $N=10000$ but with temperature $T$ varying.}
+  \caption{The posterior distribution, with $N=10000$ and temperature $T$ varying.}
   \label{fig:contour2}
@@
 
-To gain some intuition on what the posterior looks like, Figure~\ref{fig:contour1} shows simulated
-contour plots of the posterior based on varying numbers of data points $N$, with the temperature set
-at $T=1$. Note that because we are using all $N$ points here, the scale factor $N/n=1$. As $N$
-increases, the posterior converges to a multimodal distribution with modes at $(0,1)$ and $(1,-1)$.
-Figure~\ref{fig:contour2} is similar, except this time we fix the number of samples at $N=10000$,
-but show how changing the temperature $T$ affects the distribution. A larger $T$ implies a flatter
-posterior, one that (weakly) peaks in between the two true modes.
+To gain intuition on the shape of the posterior, Figure~\ref{fig:contour2} shows
+simulated contour plots of the posterior based on $N=10000$ data points, and
+temperature varying.  The (non-tempered) posterior is a multimodal distribution
+with modes at $(0,1)$ and $(1,-1)$. A larger $T$ implies a flatter posterior,
+which can be useful for parameter exploration so that the samples do not get
+trapped in a mode.
 

### 2016-09-30 nips_2016.tex
Revised 6.2, Deleted Appendix C. Yeah, took too long. UP NEXT: Appendix A (UH OH...)

@@
 show that our test enables samples to converge to the posterior distribution of
-a heated Gaussian mixture model. In Section~\ref{ssec:logistic}, we analyze its
-efficiency on logistic regression.  Appendices~\ref{app:gaussian}
-and~\ref{app:logistic} contain more detailed information on these respective
-experiments.
+a heated Gaussian mixture model (additional details are in
+Appendix~\ref{app:gaussian}). In Section~\ref{ssec:logistic}, we analyze its
+efficiency on logistic regression.
 
@@
 
-We next use logistic regression for the binary classification of 1s versus 7s in
-the MNIST dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007
-training and 1000 testing elements (we used a random subset of the test data).
-For the proposer, we use a random walk with covariance matrix $0.05I$ for the
-$784\times 784$ identity matrix $I$. We set the posterior temperature at
-$T=1000$ and use the minibatch size $m=100$ and compare
-with~\cite{cutting_mh_2014} with tolerance 0.005 and
-with~\cite{icml2014c1_bardenet14} with error bound control parameters $p = 2$
-and $\delta = 0.01$.
-% {\color{blue} Daniel: just observed this, but why are we
-% using a subset of the test data? We really should use all 2000 (I think it's
-% 2000, right?).}
-% JFC- Would be good to fix, but not critical.
-
-Figure~\ref{fig:logistic_performance} shows the prediction accuracy and log
-likelihood on the test set as a function of the cumulative training data points
-processed.  We see that our minibatch MH test is more efficient compared to the
-other two methods, because it has similar or better performance while consuming
-fewer data points.
-
 \begin{figure}[t]
@@
 
-Figure~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes for
-all three methods. With an initial minibatch size of 100, Korattikara 2014 and
-Bardenet 2014 achieve average minibatch size of 2026 and 7612 respectively over
-the MNIST classification task, while our method achieves that of 163, showing
-significantly better performance than the benchmark test methods.
+We next use logistic regression for the binary classification of 1s versus 7s in
+the MNIST dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007
+training and 1000 testing elements (we used a random subset of the test data).
+The pixel values are scaled to be in $(0,1)$.  We impose a flat prior on
+$\theta$. For the proposer, we again use a random walk, this time with
+covariance matrix $0.05I$ for the $784\times 784$ identity matrix $I$. The
+posterior temperature is set at $T=1000$.  We run our MH test and compare with
+Korattikara 2014 and Bardenet 2014; minibatch sizes and other parameters are set
+to be the same value as with the Gaussian Mixtures experiment from
+Section~\ref{ssec:gaussians}.
+
+Figure~\ref{fig:logistic_performance} shows the training log likelihood and the
+testing prediction accuracy, each as a function of the cumulative training data
+points processed.\footnote{Note that the curves do not span the same length over
+the x-axis, because our test consumes fewer samples throughout the MCMC
+procedure, so the corresponding curve will ``end'' before the other two.} To
+generate the curves, for each of the 5000 sampled vectors $\theta_t$,
+$t\in\{1,\ldots,5000\}$, we use $\theta_t$ as the parameter for logistic
+regression.  We see that our minibatch MH test is more efficient compared to the
+other two methods, because it achieves similar or superior performance while
+consuming fewer data points.
+
+Figure~\ref{fig:logistic_minibatch}, in a similar manner as
+Figure~\ref{fig:gauss_mix_2}, shows the histogram of minibatch sizes for all
+three methods on a log-log scale. With an initial minibatch size of 100,
+Korattikara 2014 and Bardenet 2014 achieve average minibatch size of 2026 and
+7612 respectively over the MNIST classification task, while our method achieves
+that of 163, showing significantly better performance than the benchmark test
+methods.
 
@@
     \item More information on Section~\ref{ssec:gaussians}.
-    \item More information on Section~\ref{ssec:logistic}.
 \end{itemize}
@@
 estimate\footnote{We do this for the logistic regression experiment, described
-in Section~\ref{ssec:logistic} and elaborated in Appendix~\ref{app:logistic}.}
-of $\theta$, such as the MAP estimate. (If $N$ is extremely large, it will cause
-the posterior to peak sharply at its modes, reducing distribution estimates to
-point estimates.)
+in Section~\ref{ssec:logistic}.} of $\theta$, such as the MAP estimate. (If $N$
+is extremely large, it will cause the posterior to peak sharply at its modes,
+reducing distribution estimates to point estimates.)
 
@@
 
-\section{Logistic Regression Experiment Details}\label{app:logistic}
-
-In this section, we go over some details of our logistic regression experiment. The feature vector
-for an image consists of its pixel values, normalized between 0 and 1. For simplicity, we only
-consider the binary classification case, so we only use digits one (denoted as output $y = 1$) and
-seven (denoted as $y = -1$). The probability of the $i^{\rm th}$ output $y_i \in \{-1,+1\}$ with the
-input vector $x_i$ is
-\begin{equation}
-p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)},
-\end{equation}
-where $\theta$ is the 784-length parameter vector.
-
-For our experiment, we impose a uniform prior to represent our lack of knowledge about $\theta$.  We
-use a random walk proposer, which can be modeled as $\theta' = \theta_i + \mathcal{N}(0,
-\sigma^2I)$, where $\theta_i$ is the current sample, $\theta'$ is the proposed sample, and we choose
-the variance to be a constant $\sigma^2 = 0.05$ for all components. We initialize $\theta_0$ to be a
-vector of all ones, and set our minibatch size as $m=100$.
-
-For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we use a
-constant temperature $T = 1000$. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
-current iteration.  %Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
-%versus iteration count.
-
-%\begin{figure}[t]
-  %\centering
-  %\includegraphics[width=1\linewidth]{./figures/appendixExp2}
- % \caption{Additional results for the logistic regression experiment.}
-  %\label{fig:appendixExp2}
-%\end{figure}
-
-%For adaptive MH testing, our experimental settings are the same as with our MH test, except we do
-%not impose a temperature. The minibatch size of adaptive MH testing is also initialized as 50, but
-%it may increase by that amount each iteration. Figure~\ref{fig:appendixExp2} shows the histogram of
-%the actual minibatch size at the end of each iteration in adaptive MH testing.
-
-\iffalse
-\section{Neural Network Experiment Details}\label{app:nnet}
-
-For our neural network experiment, we use the architecture discussed in Section~\ref{ssec:nets}. We
-use the SGHMC as the proposer for our minibatch MH testing. For the baseline, we use a tuned
-adaptive gradient descent optimizer, whose step size changes by $0.01 /(i+1)^{0.4}$. Both methods
-have minibatch sizes set at $m=200$. There are one million total data instances $x_i$.
-
-For SGHMC, we use the simplified update equations~\cite{sghmc_2014}:
-\begin{equation}
-\Delta \theta = v, \quad \quad\Delta v = -\eta \nabla U(\theta) - \alpha v + \mathcal{N}(0, 2(\alpha
--\hat{\beta}) \eta),
-\end{equation}
-where $v$ represents auxiliary momentum variables, and $\nabla U(\theta)$ is the gradient of the
-system. We set the hyperparameters to be $\eta_i = 0.01 /(i+1)^{0.4}$, and $\alpha=0.1$. We use the
-empirical Fisher~\cite{conf/icml/AhnBW12} information $V(\theta)$ to estimate the value of
-$\hat{\beta}$, so that $\hat{\beta}=\frac{1}{2}\eta V(\theta)$. In order to control ${\rm
-std}(\Delta')$, we initialize the temperature at 1000, and adjust it at iteration $i$ according to
-$T_i = \max\{1, 1000/(i+1)^{0.5}\}$.
-
-\fi
+% Daniel: This was some stuff from the Logistic Regression appendix, but we no
+% longer need that.
+
+%% In the logistic regression experiment of Section~\ref{ssec:logistic}, the
+%% feature vector for an image consists of its pixel values, normalized between 0
+%% and 1. For simplicity, we consider the binary classification case, so we only
+%% use digits one ($y = 1$) and seven ($y = -1$).  The probability of the $i^{\rm
+%% th}$ output $y_i \in \{-1,+1\}$ with the input vector $x_i$ is
+%% \begin{equation}\label{eq:logistic_probability}
+%% p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)},
+%% \end{equation}
+%% where $\theta$ is the 784-length parameter vector.
 

### 2016-10-01 nips_2016.tex
Did about 1/3 of my desired changes to Appendix A

@@
 \Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)}=
-  %%\frac{1}{2}\left( (x_i-\theta')^2 - (x_i - \theta)^2\right) =
   N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
@@
 
+{\color{blue} Daniel: changes I made:
+
+\begin{enumerate}
+    \item Changed all occurrences of $\Lambda^*(\theta',\theta)$ to
+    $\Lambda^*(\theta,\theta')$, i.e., switched the input ordering.
+
+    \item Originally, $\log u_0$ was defined as
+    $N(\Lambda^*(\theta,\theta')-\psi(1,\theta,\theta'))$ but I changed the
+    starting definition so it does \emph{not} have a factor of $N$ in its
+    definition (because we later get $N$ to appear in a subsequent step).
+
+    \item In addition, $\psi(1,\theta,\theta') = 0$ due to uniform priors and
+    symmetric proposals and there didn't seem to be a good enough reason to
+    include it in the definition so I deleted it (sometimes it's insightful to
+    have a nice form of zero inside, but I didn't see it here).
+
+    \item I added in an ``$E$'' because it seemed like you wanted the ``$u$''
+    that would make $\psi(u,\theta,\theta') = \log u =
+    E[\Lambda^*(\theta,\theta')]$.
+\end{enumerate}
+
+I'm hoping this is correct. I understood the general idea of what you were
+trying to do, it's a matter of getting the painful interval arithmetic details
+down.
+}
+
 Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and $\theta -0.5
@@
 $p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively, and guarantee that
-$\Lambda^*(\theta',\theta)$ is negative. This then guarantees that we can find a
-$u \in [0,1]$ so that $\psi(u,\theta',\theta)$ equals
-$\mathbb{E}[\Lambda^*(\theta',\theta)]$.  Specifically, choose $u_0$ to satisfy
-\begin{equation}\label{eq:log_uo}
-    \log u_0 = N(\Lambda^*(\theta',\theta)-\psi(1,\theta',\theta))
-\end{equation}
-which evaluates ({\color{blue} Daniel: sorry again for interjecting but I really
-think $\log u_0$ does not have that $N$ there. Also, for the subsequent
-equation, I think we need to be using \emph{expected} values of $\Lambda^*$}) to
-\begin{equation}
-    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
-\end{equation}
+$\Lambda^*(\theta,\theta')$ is negative. This then guarantees that we can find a
+$u \in [0,1]$ so that $\psi(u,\theta,\theta') = \log u$ equals
+$E[\Lambda^*(\theta,\theta')]$.
+
+Specifically, choose $u_0$ to satisfy $\log u_0 = E[\Lambda^*(\theta,\theta')]$.
+Using $E[x_i] = 0.5$, we see that
+\begin{align}
+    \log u_0 &= N(\theta'-\theta)\frac{1}{b} \cdot E\left[\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right] \\
+    &= -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
+\end{align}
 This means we can rewrite the acceptance test as

### 2016-10-01 nips_2016.tex
Halfway done with my Appendix A changes, added more math details. Next up is the harder part, interval arithmetic.

@@
     E[\Lambda^*(\theta,\theta')]$.
+
+    \item Added more detail for the $\Lambda^*(\theta,\theta') \not\approx
+    \psi(u,\theta,\theta')$ part.
 \end{enumerate}
@@
 $\Lambda^*(\theta,\theta')$ is negative. This then guarantees that we can find a
-$u \in [0,1]$ so that $\psi(u,\theta,\theta') = \log u$ equals
+$u \in [0,1]$ so that $\psi(u,\theta,\theta') = \log u < 0$ equals
 $E[\Lambda^*(\theta,\theta')]$.
@@
 \end{align}
-This means we can rewrite the acceptance test as
-\begin{equation}
-    \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right)
-    \hspace{0.1in} \not\approx \hspace{0.1in} 0
-\end{equation}
-where $\not\approx$ means ``significantly different from'' under the
-distribution over samples of $x_i$.  The above choice of $u_0$ ensures that the
-terms in parenthesis above sum to 0.5. Since the $x_i$ have mean 0.5, the test
-will never correctly succeed.  Furthermore, if we sample values of $u$ near
-enough to $u_0$, the terms in parenthesis will not be sufficiently different
-from 0.5 to allow the test to succeed.
+Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta') \not\approx
+\psi(u,\theta,\theta')$, where $\not\approx$ informally means ``significantly
+different from'' under the distribution over samples of $x_i$. This turns out to
+be
+\begin{align}
+\Lambda^*(\theta,\theta') \hspace{0.1in} \not\approx \hspace{0.1in}
+\psi(u_0,\theta,\theta') &\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2} \hspace{0.1in} \not\approx \hspace{0.1in} \log u_0 \\
+&\iff \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \hspace{0.1in} \not\approx \hspace{0.1in} 0 \\
+&\iff \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} 0.
+\end{align}
+Since the $x_i$ have mean 0.5, the resulting test with our chosen $u_0$ will
+never correctly succeed and must use all $N$ data points.  Furthermore, if we
+sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be
+sufficiently different from 0.5 to allow the test to succeed.
 

### 2016-10-01 nips_2016.tex
I think I finally understand what John meant by p_0 and p_1. About 2/3-rds done with Appendix A

@@
 large enough gap (several $\sigma(\Lambda^*)$) between
-$\Lambda^*(\theta',\theta)$ and $\psi(u,\theta',\theta)$.
+$\Lambda^*(\theta,\theta')$ and $\psi(u,\theta,\theta')$.
 
@@
 $q(\theta'\mid\theta)=q(\theta\mid\theta')$, and since we assumed a uniform
-prior on $\theta$, we see that $\psi(u,\theta',\theta)$ reduces to $\log u$. Our
+prior on $\theta$, we see that $\psi(u,\theta,\theta')$ reduces to $\log u$. Our
 worst-case scenario is specified in Lemma~\ref{lem:worst_case}.
@@
 standard normal CDF\footnote{Hence, $\Phi_{s_X}$ is the CDF of a zero-mean
-Gaussian with variance $s_X$.}, $S(X)$ be the logistic function, and
+Gaussian with standard deviation $s_X$.}, $S(X)$ be the logistic function, and
 $C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
@@
 \begin{enumerate}
+    \item (Somewhat important) I added slightly more detail on why if we assume
+    $Y$ has variance $1/N$, then $Y \in \frac{1}{\sqrt{N}}[0.5,1]$ has
+    probability $\Phi(1)-\Phi(0.5)$; note that I'm assuming the positive
+    interval only, not the negative one. However, you also made $p_0$ double the
+    value of $p_1$. I struggled to understand why, until I realized that you
+    used the phrase ``with matching sign.'' So I \emph{think} you meant that
+    $\theta'-\theta$ could be either positive or negative, but that $\theta-0.5$
+    was restricted in its sign so could be only half its value. Is my thinking
+    right?
+
     \item Changed all occurrences of $\Lambda^*(\theta',\theta)$ to
@@
 \end{enumerate}
-
-I'm hoping this is correct. I understood the general idea of what you were
-trying to do, it's a matter of getting the painful interval arithmetic details
-down.
 }
@@
 Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and $\theta -0.5
-\in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events occur with
-probability ({\color{blue} Daniel: sorry for interjecting here but don't we need
-to take into account the $1/\sqrt{N}$ in the probabilities? If $N \to \infty$
-(and it's going to be large with big datasets) the probabilities should be about
-zero.}) $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and
-$p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively, and guarantee that
-$\Lambda^*(\theta,\theta')$ is negative. This then guarantees that we can find a
-$u \in [0,1]$ so that $\psi(u,\theta,\theta') = \log u < 0$ equals
-$E[\Lambda^*(\theta,\theta')]$.
-
-Specifically, choose $u_0$ to satisfy $\log u_0 = E[\Lambda^*(\theta,\theta')]$.
-Using $E[x_i] = 0.5$, we see that
+\in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign (i.e., first establish
+$(\theta'-\theta)$, then ``force'' $\theta-0.5$ to be the same sign).  As
+discussed in Lemma~\ref{lem:worst_case}, both $q(\theta' \mid \theta)$ and
+$p(\theta \mid x_1,\ldots,x_N)$ have variance $1/N$. If we denote $\Phi$ as the
+CDF of the standard normal distribution, then the former event occurs with
+probability $p_0 = 2(\Phi(\sqrt{N}\frac{1}{\sqrt{N}}) -
+\Phi(\sqrt{N}\frac{0.5}{\sqrt{N}})) = 2(\Phi(1)-\Phi(0.5)) \approx 0.2997$. The
+latter event, because we restrict signs, occurs with probability $p_1 =
+\Phi(\sqrt{N}\frac{1}{\sqrt{N}}) - \Phi(\sqrt{N}\frac{0.5}{\sqrt{N}}) = \Phi(1)
+- \Phi(0.5) \approx 0.14988$.
+
+These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative,
+which means that we can find a $u \in [0,1]$ so that $\psi(u,\theta,\theta') =
+\log u < 0$ equals $E[\Lambda^*(\theta,\theta')]$.  Specifically, choose $u_0$
+to satisfy $\log u_0 = E[\Lambda^*(\theta,\theta')]$.  Using $E[x_i] = 0.5$, we
+see that
 \begin{align}
@@
 
-{\color{blue} Daniel: I would like to ask you about these probabilities because
-I do not know how we got these intervals.} The choices above for $\theta$ and
-$\theta'$ guarantee that
+{\color{blue} Daniel: TODO I have not yet updated the rest of this section.}
+
+The choices above for $\theta$ and $\theta'$ guarantee that
 \begin{equation}

### 2016-10-01 nips_2016.tex
update kl_divergence.ssc

@@
 Our test is applicable when the variance (over data samples) of the log
-acceptance probability is small enough (less than 1). Its not clear at first why
+acceptance probability is small enough (less than 1). It's not clear at first why
 this quantity should be bounded, but we will show that it is ``natural'' for
@@
 
-Its worth discussing at this point the typical goals of
+It's worth discussing at this point the typical goals of
 M-H sampling on very large datasets.  By the Bernstein-von Mises
@@
 the sampler, running first at high temperature (scaling log
-likelihoods by $1/T$) which flattens the likelihood landscape.  This is
+likelihoods by $1/T$) which flattens the likelihood landscape.  This in
 turn reduces the variance of the log acceptance probability and allows

### 2016-10-02 nips_2016.tex
Slight updates/clarifications to Appendix A. Still not quite done ...

@@
 on $\theta$. The log likelihood ratio is
-\begin{equation}
+\begin{equation}\label{eq:lemma_ll_ratio}
 \Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)}=
@@
 Similar results can be shown for other distributions and proposals by
-identifying regions in product space $(\theta,\theta'-\theta,u)$ such that the
+identifying regions in product space $(\theta,\theta',u)$ such that the
 hypothesis test needs to separate nearly-equal values.  It follows that the
@@
     \psi(u,\theta,\theta')$ part.
+
+    \item (TODO later) After Equation~\ref{eq:accept_test_zero}, we say that
+    ``the resulting test with our chosen $u_0$ will never correctly succeed
+    [...]'', but I think we should say it will not succeed with high
+    probability, because we could select a minibatch of samples such that the
+    mean \emph{actually is} different from 0.5 (it depends on the distribution
+    of the variable that corresponds to $Y = \frac{1}{b}\sum_{i=1}^b x_i$).
+
+    \item (TODO later) Unfortunately, I am unable to understand why ``the size
+    of the range in $u$ is at least $\exp([-2,-1.125])$'', unless the real range
+    is $\exp([-2,0])$ (as I think it should be) and you were simply cropping off
+    the extra range.
+
+    \item (TODO later) I do not understand why after
+    Equation~\ref{eq:accept_test_rewritten}, we say that the standard deviation
+    of the LHS ``given the interval constraints is at least $0.5/\sqrt{b}$.'' We
+    already know that the $X_i$ are iid $\mathcal{N}(\theta = 0.5, 1)$ variables
+    with a fixed $\theta = 0.5$ value, so we already know the standard deviation
+    is supposed to be $\sqrt{({\rm Var}(x_1) + \cdots + {\rm Var}(x_b))/b^2} =
+    1/\sqrt{b}$.
+
+    \item At the end, it said the joint distribution of
+    $(\theta,\theta'-\theta,u)$, but I changed it to $(\theta,\theta',u)$, based
+    on your change to the text of Lemma~\ref{lem:worst_case}. But here's a
+    question: how is the sample of $\theta'$ (or even $\theta'-\theta$, as it
+    was written earlier) independent of $\theta$?
 \end{enumerate}
@@
 These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative,
-which means that we can find a $u \in [0,1]$ so that $\psi(u,\theta,\theta') =
-\log u < 0$ equals $E[\Lambda^*(\theta,\theta')]$.  Specifically, choose $u_0$
-to satisfy $\log u_0 = E[\Lambda^*(\theta,\theta')]$.  Using $E[x_i] = 0.5$, we
-see that
+intuitively because the next choice of $\theta'$ must be ``further away'' from
+0.5 than $\theta$.  This fact means that we can find a $u \in (0,1)$ so that
+$\psi(u,\theta,\theta') = \log u < 0$ equals $E[\Lambda^*(\theta,\theta')]$.
+Specifically, choose $u_0$ to satisfy $\log u_0 = E[\Lambda^*(\theta,\theta')]$.
+Using $E[x_i] = 0.5$ and Equation~\ref{eq:lemma_ll_ratio}, we see that
 \begin{align}
@@
 Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta') \not\approx
-\psi(u,\theta,\theta')$, where $\not\approx$ informally means ``significantly
-different from'' under the distribution over samples of $x_i$. This turns out to
-be
+\psi(u,\theta,\theta')$ required by these methods, where $\not\approx$
+informally means ``significantly different from'' under the distribution over
+samples of $x_i$. This turns out to be
 \begin{align}
@@
 &\iff \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \hspace{0.1in} \not\approx \hspace{0.1in} 0 \\
-&\iff \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} 0.
+&\iff \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} 0. \label{eq:accept_test_zero}
 \end{align}
@@
 
-{\color{blue} Daniel: TODO I have not yet updated the rest of this section.}
-
 The choices above for $\theta$ and $\theta'$ guarantee that
-\begin{equation}
-    \log u_0 \in -[0.5,1][0.75,1.5] = -[0.375,1.5]
+\begin{equation}\label{eq:log_uo_range}
+    \log u_0 \in -[0.5,1][0.75,1.5] = [-1.5, -0.375]
 \end{equation}
 and consider the range of $u$ values
-\begin{equation}
+\begin{equation}\label{eq:log_u_range}
     \log u \in \log u_0 + [-0.5,0.375]
 \end{equation}
-the size of the range in $u$ is at least $\exp([-2,-1.125]) = [0.13534,0.32465]$ and occurs
-with probability at least $p_2=0.18932$. With $u$ in this range, we rewrite the test as:
-\begin{equation}
+the size of the range in $u$ is at least $\exp([-2,-1.125]) \approx
+[0.13534,0.32465]$ and occurs with probability at least $p_2=0.18932$. With $u$
+in this range, we rewrite the test as:
+\begin{equation}\label{eq:accept_test_rewritten}
     \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
 \end{equation}
-so that the LHS has expected value zero.  Given our choice of intervals for the
-variables, the RHS is in the range $1/\sqrt{N}[-1,0.75]$.  The standard
-deviation of the LHS given the interval constraints is at least $0.5/\sqrt{b}$.
-And so the gap between LHS and RHS is at most $2\sqrt{b/N}$ standard deviations.
-
-The samples $\theta$, $(\theta'-\theta)$ and $u$ are drawn independently and so
-the probability of the conjunction of these events is $c = p_0 p_1 p_2 =
-0.0085$.
+so that, as in Equation~\ref{eq:accept_test_zero}, the LHS has expected value
+zero.  Given our choice of intervals for the variables, the RHS is in the range
+$\frac{1}{\sqrt{N}}[-1,0.75]$. The standard deviation of the LHS given the
+interval constraints is at least $0.5/\sqrt{b}$.  And so the gap between the LHS
+and RHS in Equation~\ref{eq:accept_test_rewritten} is at most $2\sqrt{b/N}$
+standard deviations, limiting the range in which the test will be able to
+``succeed'' without requiring more samples.
+
+The samples $\theta$, $\theta'$ and $u$ are drawn independently and so the
+probability of the conjunction of these events is $c = p_0 p_1 p_2 = 0.0085$.
 

### 2016-10-03 nips_2016.tex
Updated Appendix A now send email to John

@@
 
+    \item (TODO later) My main confusion now is that I am unable to derive
+    Equation~\ref{eq:log_uo_range}, and we may want to clarify the notation used
+    when we put two intervals side by side.
+
     \item (TODO later) I do not understand why after
@@
     1/\sqrt{b}$.
+
+    \item I added Equation~\ref{eq:rhs_range}, which I think is how you
+    determined the range. Note that it doesn't seem to matter whether it is
+    $[-1,0.75]$ or $[-0.75,1]$ since the main thing that matters is that we have
+    maximum absolute value of one.
 
@@
 \begin{equation}\label{eq:log_uo_range}
-    \log u_0 \in -[0.5,1][0.75,1.5] = [-1.5, -0.375]
-\end{equation}
-and consider the range of $u$ values
+    \log u_0 \in -[0.5,1][0.75,1.5] = [-1.5, -0.375].
+\end{equation}
+Next, consider the range of $u$ values near $u_0$:
 \begin{equation}\label{eq:log_u_range}
-    \log u \in \log u_0 + [-0.5,0.375]
-\end{equation}
-the size of the range in $u$ is at least $\exp([-2,-1.125]) \approx
+    \log u \in \log u_0 + [-0.5,0.375].
+\end{equation}
+The size of the range in $u$ is at least $\exp([-2,-1.125]) \approx
 [0.13534,0.32465]$ and occurs with probability at least $p_2=0.18932$. With $u$
@@
 so that, as in Equation~\ref{eq:accept_test_zero}, the LHS has expected value
-zero.  Given our choice of intervals for the variables, the RHS is in the range
-$\frac{1}{\sqrt{N}}[-1,0.75]$. The standard deviation of the LHS given the
-interval constraints is at least $0.5/\sqrt{b}$.  And so the gap between the LHS
-and RHS in Equation~\ref{eq:accept_test_rewritten} is at most $2\sqrt{b/N}$
-standard deviations, limiting the range in which the test will be able to
-``succeed'' without requiring more samples.
+zero.  Given our choice of intervals for the variables, we can compute the range
+for the right hand side (RHS) assuming\footnote{If $\theta'-\theta<0$, then the
+range would be $\frac{1}{\sqrt{N}}[-0.75,1]$ but this does not matter for
+the purposes of our analysis.} that $\theta'-\theta > 0$:
+\begin{equation}\label{eq:rhs_range}
+    \min\{{\rm RHS}\} = \frac{-0.5}{\sqrt{N} \cdot 0.5} = -\frac{1}{\sqrt{N}}
+    \quad {\rm and} \quad 
+    \max\{{\rm RHS}\} = \frac{0.375}{\sqrt{N} \cdot 0.5} = \frac{0.75}{\sqrt{N}}
+\end{equation}
+Thus, the RHS is in $\frac{1}{\sqrt{N}}[-1,0.75]$.  The standard deviation of
+the LHS given the interval constraints is at least $0.5/\sqrt{b}$.
+Consequently, the gap between the LHS and RHS in
+Equation~\ref{eq:accept_test_rewritten} is at most $2\sqrt{b/N}$ standard
+deviations, limiting the range in which the test will be able to ``succeed''
+without requiring more samples.
 

### 2016-10-03 nips_2016.tex
Slight clarification to appendix b

@@
 \end{equation}
-where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$. This means
-the log likelihood of a single data instance is
-\begin{equation}
+where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$ and we draw
+the data with $\theta_1=0$ and $\theta_2=1$ fixed (though these values are
+arbitrary, and different choices would yield different posterior shapes).
+This means the log likelihood of a single data instance is
+\begin{equation}\label{eq:log_ll_single_instance}
     \log p(x_i \mid \theta) = \log \frac{1}{4\sqrt{\pi}} +

### 2016-10-03 nips_2016.tex
Changed two numbers because of Xinleis new figure

@@
 The other methods occasionally need to consume a large proportion of the
-entire data set of one million elements; the average minibatch sizes are 14862
-and 18098 for Korattikara 2014 and Bardenet 2014, respectively. The average
+entire data set of one million elements; the average minibatch sizes are 15562
+and 16857 for Korattikara 2014 and Bardenet 2014, respectively. The average
 minibatch sizes accurately predict the running times of these methods since

### 2016-10-03 nips_2016.tex
Few typos

@@
 allocating a fraction of the test error $\delta$ to each minibatch test (call
-these subtests). Its important to bound the number of subtests therefore, so
+these subtests). It's important to bound the number of subtests therefore, so
 that the tolerable error per subtest is not too small.  \cite{cutting_mh_2014}
@@
 (including ours) use the same random walk proposer with covariance $\Sigma={\rm
-diag}(0.3, 0.3)$.  Since the proposer generates a random walk, all shaping of
+diag}(0.15, 0.15)$.  Since the proposer generates a random walk, all shaping of
 distribution of the samples is due to the MH test. All methods are run 5000
@@
 \{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
-$\theta_i$ that fall in bin $j$. Then the log probability for the set is
-approximated as
+$\theta_i$ that fall in bin $j$ out of $N_{\rm bins}$ total. We omit bins in
+which either the source or MCMC-generated distributions did not have a single
+sample, resulting in the set of (at most $N_{\rm bins}$) bins $B$. Then the log
+probability for the set of samples is approximated as
 \begin{equation}\label{eq:log_prob_poissons}
-    \log p(\Theta) \approx \frac{1}{N_{\rm bins}} \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
-\end{equation}
-The results for each model are shown in Table~\ref{tab:poissons} ...
-{\color{blue} Daniel: IN PROGRESS}
+    %\log p(\Theta) \approx \frac{1}{N_{\rm bins}} \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
+    \log p(\Theta) \approx \sum_{j \in B} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
+\end{equation}
+The results for each model are shown in Table~\ref{tab:poissons}.
 
 While the Poisson results provide general guidance on the relative accuracy of
-the models it is difficult interpret the quantitative scores. We therefore
-perform significance tests to test the null hypothesis that the ideal
-distribution really generated the samples.  We perform two tests:
-
-\begin{enumerate}
-    \item An exact test using simulation to generate an approximate distribution
-    of scores given the null hypothesis.
-
-    \item An approximate test using the chi-squared distribution as the test
-    statistic. 
-\end{enumerate}
-
-The results were... {\color{blue} Daniel: IN PROGRESS}
+the models, it is difficult to interpret the quantitative scores. We therefore
+perform significance tests to show the difference between the MCMC-sampled
+distributions and the ground-truth.  We perform an approximate test using
+the Chi-Squared distribution as the test statistic, which we also show in
+Table~\ref{tab:poissons}.
+
+Both the Poisson values and the Chi-Squared test statistic result in similar
+conclusions: our method is slightly superior to~\cite{cutting_mh_2014}, but
+slightly worse than~\cite{icml2014c1_bardenet14}. We now show, however, that
+minibatch size results suggest that our method dominates in terms of speed and
+efficiency.
 
@@
 \begin{table}[t]
-    \caption{\color{blue} Daniel: TODO Poisson Probabilities}
+    \caption{Gaussian Mixture Model Statistics}
     \label{tab:poissons}
@@
     \toprule
-    Model & Ours & Korattikara 2014 & Bardenet 2014 \\
+    Model & Our Method & Korattikara 2014 & Bardenet 2014 \\
     \midrule
-    Poisson Value (Eq.~\ref{eq:log_prob_poissons}) & TODO & TODO & TODO \\
+    Poisson Value (Eq.~\ref{eq:log_prob_poissons}) & -1430.0 & -1578.9 & -1232.7 \\
+    Chi-Squared Test Statistic & 3313.9 & 3647.7 & 2444.1 \\
     \bottomrule
@@
 
+\begin{comment}
+
 {\color{blue} Daniel: changes I made:
@@
 }
+
+\end{comment}
 

### 2016-10-03 nips_2016.tex
Sent draft to Haoyu. Looks ready for arxiv i hope

@@
 \author{
-  John Canny, Haoyu Chen, Daniel Seita, Xinlei Pan \\
+  Haoyu Chen, Daniel Seita, Xinlei Pan, John Canny \\
   University of California, Berkeley \\
-  \texttt{\{canny,haoyuchen,seita,xinleipan\}@berkeley.edu}
+  \texttt{\{haoyuchen,seita,xinleipan,canny\}@berkeley.edu}
  }
@@
 in that case the per-bin distributions are well-approximated by Poisson
-distributions with parameter $\lambda_i=P_i n$. Given a set of samples $\Theta =
-\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
-$\theta_i$ that fall in bin $j$ out of $N_{\rm bins}$ total. We omit bins in
-which either the source or MCMC-generated distributions did not have a single
-sample, resulting in the set of (at most $N_{\rm bins}$) bins $B$. Then the log
-probability for the set of samples is approximated as
+distributions with parameter $\lambda_i=P_i n$. Given a set of samples
+$\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
+$\theta_i$ that fall in bin $j$ out of $N_{\rm bins}$ total. Then we have
 \begin{equation}\label{eq:log_prob_poissons}
-    %\log p(\Theta) \approx \frac{1}{N_{\rm bins}} \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
-    \log p(\Theta) \approx \sum_{j \in B} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
+    \log p(c_1, \ldots, c_{N_{\rm bins}} \mid P_1, \ldots, P_{N_{\rm bins}}) = \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
 \end{equation}

### 2016-10-04 nips_2016.tex
updates

@@
 \begin{proof}
-The following bound is given immediately after Corollary 2 in from~\cite{explicit-clt05}:
+The following bound is given immediately after Corollary 2 from~\cite{explicit-clt05}:
 \begin{equation}\label{eq:clt-boundsx}
@@
     \item The proof of Lemma~\ref{lem:worst_case}.
-    \item More information on Section~\ref{ssec:gaussians}.
 \end{itemize}
@@
 
-Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and $\theta -0.5
-\in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign (i.e., first establish
-$(\theta'-\theta)$, then ``force'' $\theta-0.5$ to be the same sign).  As
+Choose $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ (event 1) and $(\theta -0.5)
+\in \pm\frac{1}{\sqrt{N}}[0.5,1]$ filtered for  matching sign (event 2).  As
 discussed in Lemma~\ref{lem:worst_case}, both $q(\theta' \mid \theta)$ and
@@
 latter event, because we restrict signs, occurs with probability $p_1 =
-\Phi(\sqrt{N}\frac{1}{\sqrt{N}}) - \Phi(\sqrt{N}\frac{0.5}{\sqrt{N}}) = \Phi(1)
-- \Phi(0.5) \approx 0.14988$.
-
-These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative,
-intuitively because the next choice of $\theta'$ must be ``further away'' from
-0.5 than $\theta$.  This fact means that we can find a $u \in (0,1)$ so that
+\Phi(1) - \Phi(0.5) \approx 0.14988$.
+
+These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative
+by inspection of equation (\ref{eq:whyneg}) below.
+This implies that we can find a $u \in (0,1)$ so that
 $\psi(u,\theta,\theta') = \log u < 0$ equals $E[\Lambda^*(\theta,\theta')]$.
@@
 Using $E[x_i] = 0.5$ and Equation~\ref{eq:lemma_ll_ratio}, we see that
-\begin{align}
-    \log u_0 &= N(\theta'-\theta)\frac{1}{b} \cdot E\left[\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right] \\
-    &= -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
-\end{align}
+\begin{equation}
+  \log u_0 = N(\theta'-\theta)\frac{1}{b} \cdot E\left[\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right]
+\end{equation}
+\begin{equation}\label{eq:whyneg}
+    log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
+\end{equation}
 Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta') \not\approx
-\psi(u,\theta,\theta')$ required by these methods, where $\not\approx$
-informally means ``significantly different from'' under the distribution over
+\psi(u,\theta,\theta')$ used in ~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14} , where $\not\approx$
+means ``significantly different from'' under the distribution over
 samples of $x_i$. This turns out to be
@@
 
-\section{Gaussian Mixture Experiment Details}\label{app:gaussian}
-
-In this section, we provide more details on Section~\ref{ssec:gaussians}.  Our
-parameter is a 2-D vector
-$\theta = \langle \theta_1, \theta_2 \rangle$, where
-\begin{equation}
-    \theta_1 \sim \mathcal{N}(0, \sigma_1^2) \quad \mbox{ and } \quad \theta_2 \sim \mathcal{N}(0, \sigma_2^2)
-\end{equation}
-and where $\mathcal{N}$ indicates the normal distribution (more generally, the
-multivariate normal). We consider the above as our prior.
-Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and $\sigma_2^2=1$, so
-the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
-log prior probability we endow on $\theta$ is
-\begin{equation}
-    \log p(\theta) = \log \left(\frac{1}{2\pi\sqrt{10}}\right) - \frac{1}{2}\theta^T\Sigma^{-1}\theta.
-\end{equation}
-To generate the data, we use the following Gaussian mixture with tied means:
-\begin{equation}\label{eq:x_points}
-    x_i \sim \frac{1}{2}\mathcal{N}(\theta_1, \sigma_x^2) + \frac{1}{2}\mathcal{N}(\theta_1+\theta_2, \sigma_x^2)
-\end{equation}
-where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$ and we draw
-the data with $\theta_1=0$ and $\theta_2=1$ fixed (though these values are
-arbitrary, and different choices would yield different posterior shapes).
-This means the log likelihood of a single data instance is
-\begin{equation}\label{eq:log_ll_single_instance}
-    \log p(x_i \mid \theta) = \log \frac{1}{4\sqrt{\pi}} +
-    \log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
-    (\theta_1+\theta_2))^2\right)\right)
-\end{equation}
-The official problem statement is as follows: given some number of conditionally
-independent data points $x_1, x_2, \ldots, x_N$ generated according
-to~(\ref{eq:x_points}), determine the posterior distribution of $\theta$:
-\begin{equation}\label{eq:log_post}
-    \log p(\theta \mid x_1,\ldots,x_N) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
-\end{equation}
-Alternatively, depending on the application, we may opt to instead pick a point
-estimate\footnote{We do this for the logistic regression experiment, described
-in Section~\ref{ssec:logistic}.} of $\theta$, such as the MAP estimate. (If $N$
-is extremely large, it will cause the posterior to peak sharply at its modes,
-reducing distribution estimates to point estimates.)
-
-One technique we use to make our test more applicable is to add a temperature to
-the posterior. In general, we will add $T > 1$ so that the posterior is
-$p(\theta)(\prod_{i=1}^N p(x_i\mid \theta))^{1/T}$, resulting in the log
-posterior of 
-\begin{equation}\label{eq:log_prior_temp}
-    \log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\sum_{i=1}^N\log p(x_i \mid \theta).
-\end{equation}
-which has the extra $1/T$ to decrease the scale factor, so the variance scales
-as $1/T^2$.
+%% \section{Gaussian Mixture Experiment Details}\label{app:gaussian}
+
+%% In this section, we provide more details on Section~\ref{ssec:gaussians}.  Our
+%% parameter is a 2-D vector
+%% $\theta = \langle \theta_1, \theta_2 \rangle$, where
+%% \begin{equation}
+%%     \theta_1 \sim \mathcal{N}(0, \sigma_1^2) \quad \mbox{ and } \quad \theta_2 \sim \mathcal{N}(0, \sigma_2^2)
+%% \end{equation}
+%% and where $\mathcal{N}$ indicates the normal distribution (more generally, the
+%% multivariate normal). We consider the above as our prior.
+%% Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and $\sigma_2^2=1$, so
+%% the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
+%% log prior probability we endow on $\theta$ is
+%% \begin{equation}
+%%     \log p(\theta) = \log \left(\frac{1}{2\pi\sqrt{10}}\right) - \frac{1}{2}\theta^T\Sigma^{-1}\theta.
+%% \end{equation}
+%% To generate the data, we use the following Gaussian mixture with tied means:
+%% \begin{equation}\label{eq:x_points}
+%%     x_i \sim \frac{1}{2}\mathcal{N}(\theta_1, \sigma_x^2) + \frac{1}{2}\mathcal{N}(\theta_1+\theta_2, \sigma_x^2)
+%% \end{equation}
+%% where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$ and we draw
+%% the data with $\theta_1=0$ and $\theta_2=1$ fixed (though these values are
+%% arbitrary, and different choices would yield different posterior shapes).
+%% This means the log likelihood of a single data instance is
+%% \begin{equation}\label{eq:log_ll_single_instance}
+%%     \log p(x_i \mid \theta) = \log \frac{1}{4\sqrt{\pi}} +
+%%     \log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
+%%     (\theta_1+\theta_2))^2\right)\right)
+%% \end{equation}
+%% The problem statement is as follows: given some number of conditionally
+%% independent data points $x_1, x_2, \ldots, x_N$ generated according
+%% to~(\ref{eq:x_points}), determine the posterior distribution of $\theta$:
+%% \begin{equation}\label{eq:log_post}
+%%     \log p(\theta \mid x_1,\ldots,x_N) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
+%% \end{equation}
+%% Alternatively, depending on the application, we may opt to instead pick a point
+%% estimate\footnote{We do this for the logistic regression experiment, described
+%% in Section~\ref{ssec:logistic}.} of $\theta$, such as the MAP estimate. (If $N$
+%% is extremely large, it will cause the posterior to peak sharply at its modes,
+%% reducing distribution estimates to point estimates.)
+
+%% One technique we use to make our test more applicable is to add a temperature to
+%% the posterior. In general, we will add $T > 1$ so that the posterior is
+%% $p(\theta)(\prod_{i=1}^N p(x_i\mid \theta))^{1/T}$, resulting in the log
+%% posterior of 
+%% \begin{equation}\label{eq:log_prior_temp}
+%%     \log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\sum_{i=1}^N\log p(x_i \mid \theta).
+%% \end{equation}
+%% which has the extra $1/T$ to decrease the scale factor, so the variance scales
+%% as $1/T^2$.
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=1\linewidth]{contour_v2}
-  \caption{The posterior distribution, with $N=10000$ and temperature $T$ varying.}
-  \label{fig:contour2}
-\end{figure}
-
-To gain intuition on the shape of the posterior, Figure~\ref{fig:contour2} shows
-simulated contour plots of the posterior based on $N=10000$ data points, and
-temperature varying.  The (non-tempered) posterior is a multimodal distribution
-with modes at $(0,1)$ and $(1,-1)$. A larger $T$ implies a flatter posterior,
-which can be useful for parameter exploration so that the samples do not get
-trapped in a mode.
+%% \begin{figure}[t]
+%%   \centering
+%%   \includegraphics[width=1\linewidth]{contour_v2}
+%%   \caption{The posterior distribution, with $N=10000$ and temperature $T$ varying.}
+%%   \label{fig:contour2}
+%% \end{figure}
+
+%% To gain intuition on the shape of the posterior, Figure~\ref{fig:contour2} shows
+%% simulated contour plots of the posterior based on $N=10000$ data points, and
+%% temperature varying.  The (non-tempered) posterior is a multimodal distribution
+%% with modes at $(0,1)$ and $(1,-1)$. A larger $T$ implies a flatter posterior,
+%% which can be useful for parameter exploration so that the samples do not get
+%% trapped in a mode.
 

### 2016-10-04 nips_2016.tex
updates

@@
 
-Many state-of-the-art machine learning methods, and deep learning in particular,
-are based on minibatch updates (such as SGD) to a model.  Minibatch updates
-produce many improvements to the model for each pass over the dataset, and have
-high sample efficiency. They also map very well onto hardware such as GPUs. In
-contrast, M-H requires calculations over the full dataset to produce a new
-sample.  Recent results from~\cite{cutting_mh_2014,icml2014c1_bardenet14}
-perform approximate (bounded error) acceptance tests using subsets (minibatches)
-of the full dataset. The tests depend on minibatch statistics, and on the value
-of an additional uniform random variable $u$. The amount of data consumed for
-each test varies significantly from one minibatch to the next, and depends on
-the current sample, the proposed sample, and on the random variable $u$. By
-contrast,~\cite{conf/uai/MaclaurinA14} performs exact tests but requires a lower
-bound on parameter distribution across its domain.  The amount of data reduction
-depends on the accuracy of this bound, and such bounds are only available for
-relatively simple distributions.
+Many state-of-the-art machine learning methods, and deep learning in
+particular, are based on minibatch updates (such as SGD) to a model.
+Minibatch updates produce many improvements to the model for each pass
+over the dataset, and have high sample efficiency. They also map very
+well onto hardware such as GPUs. In contrast, M-H requires
+calculations over the full dataset to produce a new sample.  Recent
+results from~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
+approximate (bounded error) acceptance tests using subsets
+(minibatches) of the full dataset. The tests depend on minibatch
+statistics, and on the value of an additional uniform random variable
+$u$. The amount of data consumed for each test varies significantly
+from one minibatch to the next, and depends on the current sample, the
+proposed sample, and on the random variable $u$. By
+contrast,~\cite{conf/uai/MaclaurinA14}\cite{TallData15} perform exact
+tests but require a lower bound on parameter distribution across its
+domain.  The amount of data reduction depends on the accuracy of this
+bound, and such bounds are only available for relatively simple
+distributions. By contrast our approach is ``black box'' and requires
+only statistics gathered from the minibatches as they are used.
 
@@
 
-\subsection{Bounding the Distribution of $\Delta^*$ Versus a Gaussian}\label{ssec:delta_star_distribution}
+\subsection{Bounding the Error of $\Delta^*$ from Gaussian}\label{ssec:delta_star_distribution}
 
@@
 
-\subsection{Bounds are Preserved After Adding Random Variables}\label{ssec:preserve_bounds}
+\subsection{Error Bounds are Preserved After Adding Random Variables}\label{ssec:preserve_bounds}
 
@@
 
-In fact we can do better than this approximation (showing the error decreases as
+\subsection{Bounds on the Stationary Distribution}
+Bounds on the error of an M-H test imply bounds on the stationary
+distributed of the Markov chain under appropriate conditions. Such
+bounds were derived in both ~\cite{cutting_mh_2014}
+and~\cite{icml2014c1_bardenet14}. We include the result from
+~\cite{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the total
+variation distance between two distributions $P$ and $Q$.  Let
+$\mathcal{T}_0$ denote the transition kernel of the exact Markov
+chain, and $\mathcal{S}_0$ denote the exact posterior distribution.
+\begin{lemma}
+ If $\mathcal{T}_0$ satisfies the contraction
+  condition $d_v(P\mathcal{T}_0,\mathcal{S}_0) < \eta
+  d_v(P,\mathcal{S}_0)$ for some constant $\eta\in [0,1)$ and all
+    probability distributions $P$, then
+    \begin{equation}
+      d_v(S_0, S_{\epsilon}) \leq\frac{\epsilon}{1-\eta}
+    \end{equation}
+\end{lemma}
+
+\subsection{Improved Error Bounds Based on Skew Estimation}
+In fact we can do better than $O(n^{-1/2})$ error for the quantitative CLT (showing the error decreases as
 $O(n^{-1})$) by using a more precise limit distribution under an additional
@@
 Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
-$n^{1/2}$.
+$n^{-1/2}$.
 %{\color{blue} (Daniel: how did the equation in the next sentence
@@
 show that our test enables samples to converge to the posterior distribution of
-a heated Gaussian mixture model (additional details are in
-Appendix~\ref{app:gaussian}). In Section~\ref{ssec:logistic}, we analyze its
+a heated Gaussian mixture model. In Section~\ref{ssec:logistic}, we analyze its
 efficiency on logistic regression.

### 2016-10-04 nips_2016.tex
updates

@@
 $\mathcal{T}_0$ denote the transition kernel of the exact Markov
-chain, and $\mathcal{S}_0$ denote the exact posterior distribution.
+chain, $\mathcal{S}_0$ denote the exact posterior distribution,
+and  $\mathcal{S}_{\epsilon}$ denote the stationary distribution
+of the approximate transition kernel. 
 \begin{lemma}
@@
     \end{equation}
+where $\epsilon$ is the bound on the error in the acceptance test. 
 \end{lemma}

### 2016-10-04 nips_2016.tex
updates

@@
 where $p_0, p$, and $q$ match the corresponding functions within
-Equation~\ref{eq:acceptance_probability}. We separate out terms dependent and
+Equation~(\ref{eq:acceptance_probability}). We separate out terms dependent and
 independent of the data $x$ as:
@@
 other than the classical Metropolis test that satisfy the detailed balance (see
-Equation~\ref{eq:detailed_balance2}) condition needed for accurate posterior
+Equation~(\ref{eq:detailed_balance2})) condition needed for accurate posterior
 estimation. A class of functions leading to detailed balance is specified as
@@
 We now describe acceptance testing using the minibatch estimator
-$\Delta^*(\theta,\theta')$. From Equation~\ref{eq:delta3},
+$\Delta^*(\theta,\theta')$. From Equation~(\ref{eq:delta3}),
 $\Delta^*(\theta,\theta')$ can be represented as a constant term plus the mean
@@
 approximately normal with variance $\sigma^2(\Delta^*)$.  But to perform the
-test in Equation~\ref{eq:the_exact_test} we want $\Delta + X$ for a logistic
+test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
 random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{Bardenet15} it
@@
 
-Our proposed test in Equation~\ref{eq:criteria} requires knowing the
+Our proposed test in Equation~(\ref{eq:criteria}) requires knowing the
 distribution of the correction variable $X_{\rm corr}$ such that $X_{\rm norm} +
@@
 $C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
-distribution. Our goal, based on Equation~\ref{eq:deconvolution}, is to solve
+distribution. Our goal, based on Equation~(\ref{eq:deconvolution}), is to solve
 the following optimization problem:
@@
 \ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
-condition in Equation~\ref{eq:conv1} can be re-expressed as:
+condition in Equation~(\ref{eq:conv1}) can be re-expressed as:
 \begin{equation}\label{eq:conv2}
@@
 appropriately translated to be non-negative indices for $M, u,$ and $v$. Thus,
-Equation~\ref{eq:conv2} is equivalent to minimizing the $L_{\infty}$ norm:
+Equation~(\ref{eq:conv2}) is equivalent to minimizing the $L_{\infty}$ norm:
 \begin{equation}\label{eq:optimization_l1}
@@
 \lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_{\infty}$ norm for
-Equation~\ref{eq:optimization_l1}.
+Equation~(\ref{eq:optimization_l1}).
 
 With this approach, there is no guarantee that $u^* \geq 0$. However, we have
-some flexibility in the choice of $\sigma$ in Equation~\ref{eq:conv1}.
+some flexibility in the choice of $\sigma$ in Equation~(\ref{eq:conv1}).
 As we decrease the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$
@@
     Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*}$\;
-    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Equation~\ref{eq:clt-bounds3}\;
+    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Equation~(\ref{eq:clt-bounds3})\;
     \While {$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}{
@@
 and combining with $x>0$, we obtain the weaker but unqualified bound
-in Equation~\ref{eq:clt-bounds}.
+in Equation~(\ref{eq:clt-bounds}).
 \end{proof}
@@
 approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of
-variables $y = x s_X$, Equation~\ref{eq:clt-bounds} becomes
+variables $y = x s_X$, Equation~(\ref{eq:clt-bounds}) becomes
 \begin{equation}\label{eq:clt-bounds2}
@@
 \begin{corollary}\label{cor:our_bound_delta_prime}
-We can now substitute into Equation~\ref{eq:clt-bounds2} and displace by the mean, giving:
+We can now substitute into Equation~(\ref{eq:clt-bounds2}) and displace by the mean, giving:
 \begin{equation}\label{eq:clt-bounds3}
@@
 variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm nc}$ as
-per Algorithm 1 preserving the limit in Equation~\ref{eq:GNderivatives}.
+per Algorithm 1 preserving the limit in Equation~(\ref{eq:GNderivatives}).
 
@@
 Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
-Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
+Equation~(\ref{eq:GNderivatives}), its enough to consider a function $g(x,y')$
 parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
@@
 posterior of $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$ with the same
-prior from Equation~\ref{eq:data_generation}. The larger number of samples
+prior from Equation~(\ref{eq:data_generation}). The larger number of samples
 produces a much sharper posterior with two very narrow peaks.  Our goal is to
@@
 one caveat: the method of~\cite{icml2014c1_bardenet14} requires a global bound
-on the term $C_{\theta,\theta'}$ (see Equation~\ref{eq:bad_bound}) over the
+on the term $C_{\theta,\theta'}$ (see Equation~(\ref{eq:bad_bound})) over the
 entire the data set at every iteration in order to calculate the error bound.
@@
 Specifically, choose $u_0$ to satisfy $\log u_0 = E[\Lambda^*(\theta,\theta')]$.
-Using $E[x_i] = 0.5$ and Equation~\ref{eq:lemma_ll_ratio}, we see that
+Using $E[x_i] = 0.5$ and Equation~(\ref{eq:lemma_ll_ratio}), we see that
 \begin{equation}
@@
 \end{equation}
-so that, as in Equation~\ref{eq:accept_test_zero}, the LHS has expected value
+so that, as in Equation~(\ref{eq:accept_test_zero}), the LHS has expected value
 zero.  Given our choice of intervals for the variables, we can compute the range
@@
 Consequently, the gap between the LHS and RHS in
-Equation~\ref{eq:accept_test_rewritten} is at most $2\sqrt{b/N}$ standard
+Equation~(\ref{eq:accept_test_rewritten}) is at most $2\sqrt{b/N}$ standard
 deviations, limiting the range in which the test will be able to ``succeed''

### 2016-10-04 nips_2016.tex
updates

@@
 \author{
-  Haoyu Chen, Daniel Seita, Xinlei Pan, John Canny \\
-  University of California, Berkeley \\
+  Haoyu Chen\textsuperscript{*}, Daniel Seita\textsuperscript{*},
+  Xinlei Pan\textsuperscript{*}, John Canny\textsuperscript{*}$^{\dag}$ \\
+  * University of California, Berkeley \\
+$\dag$ Google Research, Mountain View, CA\\
   \texttt{\{haoyuchen,seita,xinleipan,canny\}@berkeley.edu}

### 2016-10-04 nips_2016.tex
updates

--- 2016-10-04 nips_2016.tex
@@
 \author{
-  Haoyu Chen, Daniel Seita, Xinlei Pan, John Canny \\
-  University of California, Berkeley \\
+  Haoyu Chen\textsuperscript{*}, Daniel Seita\textsuperscript{*},
+  Xinlei Pan\textsuperscript{*}, John Canny\textsuperscript{*}$^{\dag}$ \\
+  * University of California, Berkeley, CA \\
+$\dag$ Google Research, Mountain View, CA\\
   \texttt{\{haoyuchen,seita,xinleipan,canny\}@berkeley.edu}

### 2016-10-09 aistats2017.tex
first version of our paper to be submitted to AISTATS

--- 2016-10-04 nips_2016.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.: \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading nips_2016
-% to avoid loading the natbib package, add option nonatbib: \usepackage[nonatbib]{nips_2016}
-
-\PassOptionsToPackage{numbers, compress}{natbib}
-%\usepackage{nips_2016}
-% to compile a camera-ready version, add the [final] option, e.g.:
-\usepackage[final]{nips_2016}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{amssymb}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{graphicx}
+\documentclass[twoside]{article} \usepackage{aistats2017}
+
+% If your paper is accepted, change the options for the package
+% aistats2017 as follows:
+%
+%\usepackage[accepted]{aistats2017}
+%
+% This option will print headings for the title of your paper and
+% headings for the authors names, plus a copyright note at the end of
+% the first column of the first page.
+\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem,float}
+\usepackage{amssymb}
+\usepackage{natbib}  % for references
+\usepackage{amsfonts}
+\usepackage{nicefrac}  
+\usepackage{booktabs}
+\usepackage{microtype}
+\usepackage[]{algorithm2e}
 \usepackage{caption}
 \usepackage{subcaption}
-
-\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem}
-\usepackage[]{algorithm2e}
 \graphicspath{{figures/}}
-
+\usepackage{setspace}
 \lstset{
@@
 
+\renewcommand{\refname}{}
 \newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
@@
 
-\title{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
-
-% The \author macro works with any number of authors. There are two commands used to separate the
-% names and addresses of multiple authors: \And and \AND.
-% Using \And between authors leaves it to LaTeX to determine where to break the lines. Using \AND
-% forces a line break at that point. So, if LaTeX puts 3 of 4 authors names on the first line, and
-% the last on the second line, try using \AND instead of \And before the third author name.
-
-%%  David S.~Hippocampus\thanks{Use footnote for providing further
-%%    information about author (webpage, alternative
-%%    address)---\emph{not} for acknowledging funding agencies.} \\
-%%  Department of Computer Science\\
-%%  Cranberry-Lemon University\\
-%%  Pittsburgh, PA 15213 \\
-%%  \texttt{hippo@cs.cranberry-lemon.edu} \\
-
-\author{
-  Haoyu Chen, Daniel Seita, Xinlei Pan, John Canny \\
-  University of California, Berkeley \\
-  \texttt{\{haoyuchen,seita,xinleipan,canny\}@berkeley.edu}
- }
+% This portion of style definition will remove the index number before the 
+% without inducing an indent
+\makeatletter
+\renewcommand\@biblabel[1]{}
+\renewenvironment{thebibliography}[1]
+     {\section*{\refname}%
+      \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
+      \list{}%
+           {\leftmargin0pt
+            \@openbib@code
+            \usecounter{enumiv}}%
+      \sloppy
+      \clubpenalty4000
+      \@clubpenalty \clubpenalty
+      \widowpenalty4000%
+      \sfcode`\.\@m}
+     {\def\@noitemerr
+       {\@latex@warning{Empty `thebibliography' environment}}%
+      \endlist}
+\makeatother
 
 \begin{document}
-\maketitle
+
+% If your paper is accepted and the title of your paper is very long,
+% the style will print as headings an error message. Use the following
+% command to supply a shorter title of your paper so that it can be
+% used as headings.
+%
+%\runningtitle{I use this title instead because the last one was very long}
+
+% If your paper is accepted and the number of authors is large, the
+% style will print as headings an error message. Use the following
+% command to supply a shorter version of the authors names so that
+% they can be used as headings (for example, use only the surnames)
+%
+%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
+
+\twocolumn[
+
+\aistatstitle{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
+
+\aistatsauthor{ Anonymous Author 1 \And Anonymous Author 2 \And Anonymous Author 3 }
+
+\aistatsaddress{ Unknown Institution 1 \And Unknown Institution 2 \And Unknown Institution 3 } ]
 
 \begin{abstract}
+  %The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
+  %both left and right-hand margins. Use 10~point type, with a vertical
+  %spacing of 11~points. The {\bf Abstract} heading must be centered,
+  %bold, and in point size 12. Two line spaces precede the
+  %Abstract. The Abstract must be limited to one paragraph.
   Markov chain Monte Carlo (MCMC) methods have many applications in
@@
   we derive the test, analyze its performance, discuss its
-implementation, and present several experiments.
+  implementation, and present several experiments. We demonstrate 
+  several order-of-magnitude speedup over previous work, and show for
+  the first time that small expected minibatch sizes are possible. 
 \end{abstract}
@@
 \section{Introduction}\label{sec:introduction}
-
 Markov chain Monte Carlo (MCMC) sampling is a powerful method for
@@
 based on an acceptance test. The acceptance test is usually a
-Metropolis test~\cite{Metropolis1953,hastings70}. Conventional
+Metropolis test~\citep{Metropolis1953, hastings70}. Conventional
 Metropolis-Hastings requires all $N$ data instances to generate one
@@
 calculations over the full dataset to produce a new sample.  Recent
-results from~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
+results from~\citet{cutting_mh_2014} and ~\citet{icml2014c1_bardenet14} perform
 approximate (bounded error) acceptance tests using subsets
@@
 proposed sample, and on the random variable $u$. By
-contrast,~\cite{conf/uai/MaclaurinA14}\cite{TallData15} perform exact
+contrast,~\citet{conf/uai/MaclaurinA14,TallData15} perform exact
 tests but require a lower bound on parameter distribution across its
@@
 bound, and such bounds are only available for relatively simple
-distributions. By contrast our approach is ``black box'' and requires
+distributions. By contrast our approach is a ``black box'' and requires
 only statistics gathered from the minibatches as they are used.
@@
 value.
-%% {\color{blue} Daniel: we really need to show the amount of data is fixed
-%% with ``high probability'' (the results suggest ``OK probability'').} 
-%% Analysis show data amount is gaussian and short-tailed and therefore bounded with high p
-%% experiments back this up. 
-We use a Barker test function~\cite{Barker65} rather than a Metropolis
+
+We use a Barker test function~\citep{Barker65} rather than a Metropolis
 test, which makes our test naturally error tolerant. The idea of using
 a noise-tolerant test using Barker's test function was suggested but
-not explored empirically in~\cite{Bardenet15} section 6.3. But the
+not explored empirically in~\citet{TallData15} section 6.3. But the
 asymptotic test statistic CDF and the Barker function are different,
 which leads to fixed errors for the approach
-in~\cite{Bardenet15}. Here we show that the difference between the
+in~\citet{TallData15}. Here we show that the difference between the
 distributions can be corrected with an additive random variable. This
@@
 well-specified models running Metropolis-Hastings sampling with optimal
-proposals~\cite{OptimalScaling01} on a full dataset. When we reduce the amount
+proposals~\citep{OptimalScaling01} on a full dataset. When we reduce the amount
 of data for the test, the variance goes up. We have to reduce variance in one
@@
 billions of samples. Simply sampling from this distribution is one
-application, but an efficient proposal distribution~\cite{OptimalScaling01} has similar variance to the target
+application, but an efficient proposal distribution~\citep{OptimalScaling01} has similar variance to the target
 distribution and will diffuse to it extremely slowly from an
@@
 
-In the Metropolis-Hastings method~\cite{gilks1996markov,brooks2011handbook}, a
+In the Metropolis-Hastings method~\citep{gilks1996markov,brooks2011handbook}, a
 difficult-to-compute probability distribution $p(\theta)$ is sampled using a
@@
 training data points, but this is very expensive for large datasets. To address
-this challenge,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
+this challenge,~\citet{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
 Metropolis-Hasting tests using sequential hypothesis testing. During each
@@
 the test repeats. This process continues until a decision. The bounds depend on
-either a asymptotic Central Limit Theorem~\cite{cutting_mh_2014} or a
-concentration bound~\cite{icml2014c1_bardenet14}. The latter requires direct
+either a asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
+concentration bound~\citep{icml2014c1_bardenet14}. The latter requires direct
 bounds on the log likelihood ratio, which for general distributions requires
@@
 
-Following~\cite{icml2014c1_bardenet14}, we write the test $u <
+Following~\citet{icml2014c1_bardenet14}, we write the test $u <
 \alpha(\theta,\theta')$ in the equivalent form $\Lambda(\theta,\theta') >
 \psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those in
-\cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later.}
+\citet{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later.}
 \begin{equation}
-\Lambda(\theta,\theta') = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
-\hspace{0.2in} {\rm and} \hspace{0.2in}
-\psi(u,\theta,\theta') = \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right)
+\begin{split}
+\Lambda(\theta,\theta') & = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+\hspace{0.2in} \\
+{\rm and} \hspace{0.2in}
+\psi(u,\theta,\theta') & = \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right)
+\end{split}
 \end{equation}
@@
 $\Lambda^*(\theta,\theta')$ is substantially larger than
-$\psi(u,\theta,\theta')$.  In~\cite{cutting_mh_2014} an asymptotic
+$\psi(u,\theta,\theta')$.  In~\citet{cutting_mh_2014} an asymptotic
 central limit argument was used to derive this gap, while
-in~\cite{icml2014c1_bardenet14} a concentration bound was used. In
+in~\citet{icml2014c1_bardenet14} a concentration bound was used. In
 both cases, the resulting tests were shown to give useful reductions
@@
 We next show that for a simple distribution, the lower bound of average
-number of data instances consumed for one iteration of~\cite{cutting_mh_2014}
-and~\cite{icml2014c1_bardenet14} is $\Omega(N)$, where $N$ is the number of data
+number of data instances consumed for one iteration of~\citet{cutting_mh_2014}
+and~\citet{icml2014c1_bardenet14} is $\Omega(N)$, where $N$ is the number of data
 points.
@@
 \begin{equation}\label{eq:lemma_ll_ratio}
-\Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)}=
-  N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
+\begin{split}
+\Lambda^*(\theta,\theta') & =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)} \\
+& = N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right) \\
+\end{split}
 \end{equation}
@@
 distribution has the same variance as the target
-distribution~\cite{OptimalScaling01}, so we use the proposal
+distribution~\citep{OptimalScaling01}, so we use the proposal
 $\mathcal{N}(\theta,1/N)$, which is implemented as
@@
 hypothesis test needs to separate nearly-equal values.  It follows that the
-accelerated M-H tests in~\cite{cutting_mh_2014,icml2014c1_bardenet14} require at
+accelerated M-H tests in~\citet{cutting_mh_2014,icml2014c1_bardenet14} require at
 least a constant fraction $\geq c$ in the amount of data consumed per test
@@
 physics. By viewing random variables as particles in a system, one can apply
-Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which
+Hamiltonian Monte Carlo (HMC)~\citep{mcmc_hamiltonian_2010} methods which
 generate high acceptance \emph{and} distant proposals when run on full batches
-of data. Recently Langevin Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has
+of data. Recently Langevin Dynamics~\citep{langevin_2011,conf/icml/AhnBW12} has
 been applied to Bayesian estimation on minibatches of data. This simplified
@@
 step size is severe, and the state space exploration reduces to a random walk.
-Full minibatch HMC for minibatches was recently described in~\cite{sghmc_2014}
+Full minibatch HMC for minibatches was recently described in~\citet{sghmc_2014}
 which allows momentum-augmented proposals with larger step sizes. However, step
@@
 
-For our new M-H test, we denote the exact and approximate log likelihood ratios
+For our new M-H test, we denote the exact approximate log likelihood ratios
 as $\Delta$ and $\Delta^*$.  First, $\Delta$ is defined as
@@
 \begin{equation}\label{eq:delta2}
-    \Delta(\theta,\theta') = \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') =
-    \Lambda(\theta,\theta') - \psi(1,\theta,\theta').
+\begin{split}
+    \Delta(\theta,\theta') &= \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') \\
+    & = \Lambda(\theta,\theta') - \psi(1,\theta,\theta'). \\
+\end{split}
 \end{equation}
@@
 \begin{equation}\label{eq:delta3}
-    \Delta^*(\theta,\theta') =
-\frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') =
-\Lambda^*(\theta,\theta') - \psi(1,\theta,\theta')
+\begin{split}
+    \Delta^*(\theta,\theta') &=
+\frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta')\\
+&=\Lambda^*(\theta,\theta') - \psi(1,\theta,\theta')
+\end{split}
 \end{equation}
@@
 
-This result is used in~\cite{Barker65} to define the Barker acceptance test.  As
+This result is used in~\citet{Barker65} to define the Barker acceptance test.  As
 a sanity check, choosing $g(s) = \exp(s) \wedge 1$ --- a function satisfying the
@@
 \subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
-For our new MH test we use the Barker logistic~\cite{Barker65} function:
+For our new MH test we use the Barker logistic~\citep{Barker65} function:
 $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the
@@
 test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
-random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{Bardenet15} it
+random variable $X$ (call it $X_{\rm log}$ from now on). In~\citet{TallData15} it
 was proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed
@@
 
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
-    \caption{
-    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
-    convolve to form the standard logistic distribution. These were obtained
-    based on our original linear programming formulation, as discussed in
-    Section~\ref{sec:correction}. We use three standard deviation values of
-    $X_{\rm norm}$. The two red curves convolve to form the logistic, etc. The
-    $y$-axis is capped at 0.5 for readability. This figure must be viewed in
-    color.
-    }
-    \label{fig:deconvolution}
+%\begin{figure}[t]
+% \centering
+%\includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
+%    \caption{
+%   Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
+%    convolve to form the standard logistic distribution. These were obtained
+%    based on our original linear programming formulation, as discussed in
+%    Section~\ref{sec:correction}. We use three standard deviation values of
+%    $X_{\rm norm}$. The two red curves convolve to form the logistic, etc. The
+%    $y$-axis is capped at 0.5 for readability. This figure must be viewed in
+%    color.
+%    }
+%    \label{fig:deconvolution}
 %\vspace{-10pt}
-\end{figure}
+%\end{figure}
 
@@
 \begin{equation}\label{eq:conv1}
-    C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|
+    \small{C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|}
 \end{equation}
@@
 \begin{equation}\label{eq:conv2}
-    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
+    \scriptsize{C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.}
 \end{equation}
@@
 programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
-Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
-$X_{\rm corr}$ densities that sum to $X_{\rm log}$ from this process.
+%Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
+%$X_{\rm corr}$ densities that sum to $X_{\rm log}$ from this process.
 
@@
 
-\begin{table}[t]
+\begin{table}[h]
     \caption{Error in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$}
@@
     \begin{tabular}{l l l l}
-    \toprule
-    $N$ & $\sigma$ & $\lambda$ & $L_{\infty}$ error  \\
-    \midrule
+   % \toprule
+    ${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error}  \\
+    \hline \\
     4000 & 0.9 & 1 & 1.0e-4  \\
     4000 & 0.8 & 0.03 & 5.0e-6 \\
-    \bottomrule
+    %\bottomrule
     \end{tabular}
@@
 
-In the most similar prior works,~\cite{cutting_mh_2014} uses asymptotic
+In the most similar prior works,~\citet{cutting_mh_2014} uses asymptotic
 arguments based on the Central Limit Theorem to argue that its approximate
 acceptance test error tends to zero as batch size increases. But no quantitative
-bounds are given. In~\cite{icml2014c1_bardenet14}, explicit bounds are given,
+bounds are given. In~\citet{icml2014c1_bardenet14}, explicit bounds are given,
 but they depend on bounding
@@
 \begin{proof}
-The following bound is given immediately after Corollary 2 from~\cite{explicit-clt05}:
+The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
 \begin{equation}\label{eq:clt-boundsx}
--6.4E|X|^3-2E|X| \leq    \sup_x|{\rm Pr}(t<x) - \Phi(x)|{\sqrt{n}} \leq 1.36E|X|^3
+\scriptsize{-6.4E|X|^3-2E|X| \leq    \sup_x|{\rm Pr}(t<x) - \Phi(x)|{\sqrt{n}} \leq 1.36E|X|^3}
 \end{equation}
@@
 \begin{equation}\label{eq:clt-bounds3}
-    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} = \epsilon(\theta,\theta',b)
+\begin{split}
+    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| &\leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} \\
+    & = \epsilon(\theta,\theta',b) \\
+\end{split}
 \end{equation}
@@
 \begin{equation}
--\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon
+\begin{split}
+\tiny{-\epsilon} & = \tiny{\int_{-\infty}^{+\infty} -\epsilon R(x) dx } \\
+& \tiny{\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx }\\
+& \tiny{\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon}\\
+\end{split}
 \end{equation}
@@
 \begin{equation}
-    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
+\begin{split}
+    P(y) & = {\rm Pr}(\Delta^* < y) \\
+     {\rm and} \quad Q(y) & = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
+\end{split}
 \end{equation}
@@
 \begin{equation}
-    P'(y) = {\rm Pr}(\Delta^*+X_{\rm nc} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
+\begin{split}
+    P'(y) & = {\rm Pr}(\Delta^*+X_{\rm nc} < y) \\
+    \quad {\rm and}\quad Q'(y) & = \Phi\left({y-\Delta}\right)
+\end{split}
 \end{equation}
@@
 \begin{equation}
-    P''(y) = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
+\begin{split}
+    P''(y) & = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) \\
+    \quad {\rm and}\quad Q''(y) & = S\left({y-\Delta}\right)
+\end{split}
 \end{equation}
@@
 distributed of the Markov chain under appropriate conditions. Such
-bounds were derived in both ~\cite{cutting_mh_2014}
-and~\cite{icml2014c1_bardenet14}. We include the result from
-~\cite{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the total
+bounds were derived in both ~\citet{cutting_mh_2014}
+and~\citet{icml2014c1_bardenet14}. We include the result from
+~\citet{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the total
 variation distance between two distributions $P$ and $Q$.  Let
@@
 then Equation 2.2 in Bentkus et al.'s work on Edgeworth
-expansions~\cite{Bentkus97} provides the following:
+expansions~\citep{Bentkus97} provides the following:
 
@@
 We start with a simple Gaussian mixture model, borrowing an experiment
-from~\cite{langevin_2011}.  The parameter is 2-D, $\theta =
+from~\citet{langevin_2011}.  The parameter is 2-D, $\theta =
 \langle \theta_1,\theta_2 \rangle$, and the parameter/data generation process is
@@
 \begin{equation}\label{eq:data_generation}
+\begin{split}
     \begin{bmatrix} \theta_1 \\ \theta_2 \end{bmatrix} 
-    \sim \mathcal{N}\left(\begin{bmatrix}0 \\ 0\end{bmatrix}, 
-    \begin{bmatrix} \sigma_1^2 & 0 \\ 0 & \sigma_2^2\end{bmatrix} \right);\quad \quad x_i \sim
-    0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
+    & \sim \mathcal{N}\left(\begin{bmatrix}0 \\ 0\end{bmatrix}, 
+    \begin{bmatrix} \sigma_1^2 & 0 \\ 0 & \sigma_2^2\end{bmatrix} \right); \\
+    x_i & \sim 0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
+\end{split}
 \end{equation}
@@
 We run MCMC with our MH test using minibatch size $m=100$.  We also run this
-using the methods from~\cite{cutting_mh_2014} (``Korattikara 2014'')
-and~\cite{icml2014c1_bardenet14} (``Bardenet 2014''), also with $m=100$.
-For~\cite{cutting_mh_2014}, we increment minibatches by 100 as recommended, and
-for~\cite{icml2014c1_bardenet14}, we increase the minibatch size geometrically
+using the methods from~\citet{cutting_mh_2014} %(``Korattikara 2014'')
+and~\citet{icml2014c1_bardenet14}, also with $m=100$.
+For~\citet{cutting_mh_2014}, we increment minibatches by 100 as recommended, and
+for~\citet{icml2014c1_bardenet14}, we increase the minibatch size geometrically
 with a ratio of $\gamma = 1.5$ as recommended. These tests use different
 minibatch size updates because of their different approaches to repeated
-testing. \cite{icml2014c1_bardenet14} uses straightforward discounting,
+testing. \citet{icml2014c1_bardenet14} uses straightforward discounting,
 allocating a fraction of the test error $\delta$ to each minibatch test (call
 these subtests). It's important to bound the number of subtests therefore, so
-that the tolerable error per subtest is not too small.  \cite{cutting_mh_2014}
+that the tolerable error per subtest is not too small.  \citet{cutting_mh_2014}
 uses a more aggressive repeated testing approach, explicitly modeling
@@
 
-The tolerance for making a decision in~\cite{cutting_mh_2014} is
+The tolerance for making a decision in~\citet{cutting_mh_2014} is
 $\epsilon=0.005$, and the error bound control parameters
-in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.  All methods
+in~\citet{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.  All methods
 (including ours) use the same random walk proposer with covariance $\Sigma={\rm
@@
 times to collect 5000 samples.
-
-\begin{figure}[t]
+\begin{figure}
     \centering
-    \includegraphics[width=1\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
+    \includegraphics[width=\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
     \caption{
@@
     \label{fig:gauss_mix_1}
-\end{figure}
-
-\begin{figure}[t]
+%\end{figure*}
+%\begin{figure*}[t!]
     \centering
-    \includegraphics[width=1\linewidth]{minibatch_size_gaussian.png}
+    \includegraphics[width=\linewidth]{minibatch_size_gaussian.png}
     \caption{
@@
 \begin{equation}\label{eq:log_prob_poissons}
-    \log p(c_1, \ldots, c_{N_{\rm bins}} \mid P_1, \ldots, P_{N_{\rm bins}}) = \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
+\begin{split}
+    \log p(c_1, \ldots, c_{N_{\rm bins}} \mid P_1, \ldots, P_{N_{\rm bins}}) & =  \\
+    \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)). & \\
+\end{split}
 \end{equation}
@@
 Both the Poisson values and the Chi-Squared test statistic result in similar
-conclusions: our method is slightly superior to~\cite{cutting_mh_2014}, but
-slightly worse than~\cite{icml2014c1_bardenet14}. We now show, however, that
+conclusions: our method is slightly superior to~\citet{cutting_mh_2014}, but
+slightly worse than~\citet{icml2014c1_bardenet14}. We now show, however, that
 minibatch size results suggest that our method dominates in terms of speed and
@@
 all methods have a running time proportional to the total data consumed, with
-one caveat: the method of~\cite{icml2014c1_bardenet14} requires a global bound
+one caveat: the method of~\citet{icml2014c1_bardenet14} requires a global bound
 on the term $C_{\theta,\theta'}$ (see Equation~(\ref{eq:bad_bound})) over the
@@
 We found that the sample code provided by the authors
-of~\cite{icml2014c1_bardenet14} in fact computed this bound explicitly for each
+of~\citet{icml2014c1_bardenet14} in fact computed this bound explicitly for each
 sample generated, i.e. the implementation traversed the entire dataset for each
@@
 measurements of the running time of the sample implementation
-of~\cite{icml2014c1_bardenet14}, and consider only the time to process the
+of~\citet{icml2014c1_bardenet14}, and consider only the time to process the
 minibatches.
@@
 %% the same plot. 
-
-\begin{table}[t]
+\begin{table}[h]
     \caption{Gaussian Mixture Model Statistics}
@@
     \centering
+    \resizebox{.5\textwidth}{!}{% <------ Don't forget this %
     \begin{tabular}{l l l l}
-    \toprule
-    Model & Our Method & Korattikara 2014 & Bardenet 2014 \\
-    \midrule
+    %\toprule
+    {\bf Model} & {\bf Our Method} & {\bf Korattikara 2014} & {\bf Bardenet 2014} \\
+    %\midrule
+    \hline \\
     Poisson Value (Eq.~\ref{eq:log_prob_poissons}) & -1430.0 & -1578.9 & -1232.7 \\
     Chi-Squared Test Statistic & 3313.9 & 3647.7 & 2444.1 \\
-    \bottomrule
-    \end{tabular}
+   % \bottomrule
+    \end{tabular}% <------ Don't forget this %
+   }
 \end{table}
@@
 \subsection{Logistic Regression}\label{ssec:logistic}
-
-\begin{figure}[t]
+\begin{figure}
 	\centering
-	\includegraphics[width=1\linewidth]{logistic_performance.png}
+	\includegraphics[width=\linewidth]{logistic_performance.png}
 	\caption{
@@
 	\label{fig:logistic_performance}
-\end{figure}
-
-\begin{figure}[t]
+%\end{figure*}
+%\begin{figure*}[t!]
 	\centering
-	\includegraphics[width=1\linewidth]{minibatch_size_logistic.png}
+	\includegraphics[width=\linewidth]{minibatch_size_logistic.png}
 	\caption{
@@
 We next use logistic regression for the binary classification of 1s versus 7s in
-the MNIST dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007
+the MNIST dataset~\citep{lecun-mnisthandwrittendigit-2010}. The data has 12007
 training and 1000 testing elements (we used a random subset of the test data).
@@
 Carlo methods, providing a recipe for how to use our algorithm (following the
-framework of~\cite{sgmcmc_2015}), or integrating parallel
-MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
-
-\small
-\bibliography{nips_2016}
-\bibliographystyle{ieeetr}
+framework of~\citet{sgmcmc_2015}), or integrating parallel
+MCMC~\citep{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
+
+\iffalse
+\subsubsection{Figures}
+
+All artwork must be centered, neat, clean, and legible.  All lines
+should be very dark for purposes of reproduction, and art work should
+not be hand-drawn.  Figures may appear at the top of a column, at the
+top of a page spanning multiple columns, inline within a column, or
+with text wrapped around them, but the figure number and caption
+always appear immediately below the figure.  Leave 2 line spaces
+between the figure and the caption. The figure caption is initial caps
+and each figure should be numbered consecutively.
+
+Make sure that the figure caption does not get separated from the
+figure. Leave extra white space at the bottom of the page rather than
+splitting the figure and figure caption.
+\begin{figure}[h]
+\vspace{.3in}
+\centerline{\fbox{This figure intentionally left non-blank}}
+\vspace{.3in}
+\caption{Sample Figure Caption}
+\end{figure}
+
+\subsubsection{Tables}
+
+All tables must be centered, neat, clean, and legible. Do not use hand-drawn tables. Table number and title always appear above the table.
+See Table~\ref{sample-table}.
+
+Use one line space before the table title, one line space after the table title, and one line space after the table. The table title must be
+initial caps and each table numbered consecutively.
+
+\begin{table}[h]
+\caption{Sample Table Title} \label{sample-table}
+\begin{center}
+\begin{tabular}{ll}
+{\bf PART}  &{\bf DESCRIPTION} \\
+\hline \\
+Dendrite         &Input terminal \\
+Axon             &Output terminal \\
+Soma             &Cell body (contains cell nucleus) \\
+\end{tabular}
+\end{center}
+\end{table}
+
+\section{SUPPLEMENTARY MATERIAL}
+
+If you need to include additional appendices during submission, you
+can include them in the supplementary material file.
+
+\fi
+
+\newpage
+\subsubsection*{References}
+%\bibliographystyle{apalike}
+\bibliographystyle{abbrvnat}
+\setcitestyle{authoryear,open={[},close={]}}
+\bibliography{aistats2017}
 \normalsize
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \clearpage
@@
 \begin{equation}\label{eq:whyneg}
-    log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
+    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
 \end{equation}
@@
 samples of $x_i$. This turns out to be
-\begin{align}
-\Lambda^*(\theta,\theta') \hspace{0.1in} \not\approx \hspace{0.1in}
-\psi(u_0,\theta,\theta') &\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2} \hspace{0.1in} \not\approx \hspace{0.1in} \log u_0 \\
-&\iff \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \hspace{0.1in} \not\approx \hspace{0.1in} 0 \\
-&\iff \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} 0. \label{eq:accept_test_zero}
-\end{align}
+\begin{equation}
+\begin{split}
+\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') & \quad \\
+\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2} & \not\approx \log u_0\\
+\iff \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) & \not\approx  0 \\
+\iff \frac{1}{b}\sum_{i=1}^b x_i-0.5 & \not\approx 0. \label{eq:accept_test_zero}
+\end{split}
+\end{equation}
 Since the $x_i$ have mean 0.5, the resulting test with our chosen $u_0$ will
@@
 \begin{equation}\label{eq:rhs_range}
-    \min\{{\rm RHS}\} = \frac{-0.5}{\sqrt{N} \cdot 0.5} = -\frac{1}{\sqrt{N}}
-    \quad {\rm and} \quad 
-    \max\{{\rm RHS}\} = \frac{0.375}{\sqrt{N} \cdot 0.5} = \frac{0.75}{\sqrt{N}}
+\begin{split}
+    \min\{{\rm RHS}\} & = \frac{-0.5}{\sqrt{N} \cdot 0.5} = -\frac{1}{\sqrt{N}} \\
+    {\rm and} \quad \max\{{\rm RHS}\} & = \frac{0.375}{\sqrt{N} \cdot 0.5} = \frac{0.75}{\sqrt{N}}
+\end{split}
 \end{equation}

### 2016-10-09 paper_versions/nips_2016.tex
Cleaning up directory by removing nips files

--- 2016-10-04 nips_2016.tex
@@
 we derive the test, analyze its performance, discuss its
-implementation, and present several experiments.
+implementation, and present several experiments. We demonstrate 
+several order-of-magnitude speedup over previous work, and show for
+the first time that small expected minibatch sizes are possible. 
 \end{abstract}
@@
 proposed sample, and on the random variable $u$. By
-contrast,~\cite{conf/uai/MaclaurinA14}\cite{TallData15} perform exact
+contrast,~\cite{conf/uai/MaclaurinA14,TallData15} perform exact
 tests but require a lower bound on parameter distribution across its
@@
 a noise-tolerant test using Barker's test function was suggested but
-not explored empirically in~\cite{Bardenet15} section 6.3. But the
+not explored empirically in~\cite{TallData15} section 6.3. But the
 asymptotic test statistic CDF and the Barker function are different,
 which leads to fixed errors for the approach
-in~\cite{Bardenet15}. Here we show that the difference between the
+in~\cite{TallData15}. Here we show that the difference between the
 distributions can be corrected with an additive random variable. This
@@
 test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
-random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{Bardenet15} it
+random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{TallData15} it
 was proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed

### 2016-10-09 aistats2017.tex
Removed several tiny and scriptsize commands. I'm almost certain reviewers will hate that

@@
 
-%\begin{figure}[t]
-% \centering
-%\includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
-%    \caption{
-%   Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
-%    convolve to form the standard logistic distribution. These were obtained
-%    based on our original linear programming formulation, as discussed in
-%    Section~\ref{sec:correction}. We use three standard deviation values of
-%    $X_{\rm norm}$. The two red curves convolve to form the logistic, etc. The
-%    $y$-axis is capped at 0.5 for readability. This figure must be viewed in
-%    color.
-%    }
-%    \label{fig:deconvolution}
-%\vspace{-10pt}
-%\end{figure}
-
 Our proposed test in Equation~(\ref{eq:criteria}) requires knowing the
@@
 To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
-and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}$ and $j \in \{-N,
-\ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
-condition in Equation~(\ref{eq:conv1}) can be re-expressed as:
+and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}=\mathcal{I}$ and
+$j \in \{-N, \ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$,
+and the condition in Equation~(\ref{eq:conv1}) can be re-expressed as:
 \begin{equation}\label{eq:conv2}
-    \scriptsize{C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.}
+    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \mathcal{I}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
 \end{equation}
@@
 programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
-%Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
-%$X_{\rm corr}$ densities that sum to $X_{\rm log}$ from this process.
 
@@
 \begin{equation}
-    \bar{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad s_X = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\bar{X}^2\right)^{\frac{1}{2}}.
+    \bar{X} = \frac{1}{n}\sum_{i=1}^nX_i, \quad s_X = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\bar{X}^2\right)^{\frac{1}{2}}.
 \end{equation}
@@
 \begin{equation}\label{eq:clt-boundsx}
-\scriptsize{-6.4E|X|^3-2E|X| \leq    \sup_x|{\rm Pr}(t<x) - \Phi(x)|{\sqrt{n}} \leq 1.36E|X|^3}
+\begin{split}
+-6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
+&\leq 1.36E|X|^3.
+\end{split}
 \end{equation}
@@
 \end{equation}
-and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
+and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
 \begin{equation}
 \begin{split}
-\tiny{-\epsilon} & = \tiny{\int_{-\infty}^{+\infty} -\epsilon R(x) dx } \\
-& \tiny{\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx }\\
-& \tiny{\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon}\\
+-\epsilon &= \tiny{\int_{-\infty}^{+\infty} -\epsilon R(x) dx } \\
+&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
+&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon.
 \end{split}
@@
 \end{proof}
+\vspace{-10pt}
 
@@
 
-Relating Lemma~\ref{lem:cdf_bounds} to our acceptance test, we observe that:
+Relating Lemma~\ref{lem:cdf_bounds} to our acceptance test, we have the
+following Corollary:
 
@@
 \begin{center}
-{\Large Outline of Appendix}
+{\Large Supplementary Material}
 \end{center}
-
-In this appendix, we describe the following topics:
-
-\begin{itemize}[noitemsep]
-    \item The proof of Lemma~\ref{lem:worst_case}.
-\end{itemize}
 
@@
 
-%% \section{Gaussian Mixture Experiment Details}\label{app:gaussian}
-
-%% In this section, we provide more details on Section~\ref{ssec:gaussians}.  Our
-%% parameter is a 2-D vector
-%% $\theta = \langle \theta_1, \theta_2 \rangle$, where
-%% \begin{equation}
-%%     \theta_1 \sim \mathcal{N}(0, \sigma_1^2) \quad \mbox{ and } \quad \theta_2 \sim \mathcal{N}(0, \sigma_2^2)
-%% \end{equation}
-%% and where $\mathcal{N}$ indicates the normal distribution (more generally, the
-%% multivariate normal). We consider the above as our prior.
-%% Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and $\sigma_2^2=1$, so
-%% the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
-%% log prior probability we endow on $\theta$ is
-%% \begin{equation}
-%%     \log p(\theta) = \log \left(\frac{1}{2\pi\sqrt{10}}\right) - \frac{1}{2}\theta^T\Sigma^{-1}\theta.
-%% \end{equation}
-%% To generate the data, we use the following Gaussian mixture with tied means:
-%% \begin{equation}\label{eq:x_points}
-%%     x_i \sim \frac{1}{2}\mathcal{N}(\theta_1, \sigma_x^2) + \frac{1}{2}\mathcal{N}(\theta_1+\theta_2, \sigma_x^2)
-%% \end{equation}
-%% where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$ and we draw
-%% the data with $\theta_1=0$ and $\theta_2=1$ fixed (though these values are
-%% arbitrary, and different choices would yield different posterior shapes).
-%% This means the log likelihood of a single data instance is
-%% \begin{equation}\label{eq:log_ll_single_instance}
-%%     \log p(x_i \mid \theta) = \log \frac{1}{4\sqrt{\pi}} +
-%%     \log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
-%%     (\theta_1+\theta_2))^2\right)\right)
-%% \end{equation}
-%% The problem statement is as follows: given some number of conditionally
-%% independent data points $x_1, x_2, \ldots, x_N$ generated according
-%% to~(\ref{eq:x_points}), determine the posterior distribution of $\theta$:
-%% \begin{equation}\label{eq:log_post}
-%%     \log p(\theta \mid x_1,\ldots,x_N) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
-%% \end{equation}
-%% Alternatively, depending on the application, we may opt to instead pick a point
-%% estimate\footnote{We do this for the logistic regression experiment, described
-%% in Section~\ref{ssec:logistic}.} of $\theta$, such as the MAP estimate. (If $N$
-%% is extremely large, it will cause the posterior to peak sharply at its modes,
-%% reducing distribution estimates to point estimates.)
-
-%% One technique we use to make our test more applicable is to add a temperature to
-%% the posterior. In general, we will add $T > 1$ so that the posterior is
-%% $p(\theta)(\prod_{i=1}^N p(x_i\mid \theta))^{1/T}$, resulting in the log
-%% posterior of 
-%% \begin{equation}\label{eq:log_prior_temp}
-%%     \log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\sum_{i=1}^N\log p(x_i \mid \theta).
-%% \end{equation}
-%% which has the extra $1/T$ to decrease the scale factor, so the variance scales
-%% as $1/T^2$.
-
-% Daniel: I decided that I don't like this figure anymore.
-% \begin{figure}[t]
-%   \centering
-%   \includegraphics[width=1\linewidth]{contour_v1}
-%   \caption{The posterior distribution, from 50 to 100k samples, with temperature set at 1.}
-%   \label{fig:contour1}
-% \end{figure}
-
-%% \begin{figure}[t]
-%%   \centering
-%%   \includegraphics[width=1\linewidth]{contour_v2}
-%%   \caption{The posterior distribution, with $N=10000$ and temperature $T$ varying.}
-%%   \label{fig:contour2}
-%% \end{figure}
-
-%% To gain intuition on the shape of the posterior, Figure~\ref{fig:contour2} shows
-%% simulated contour plots of the posterior based on $N=10000$ data points, and
-%% temperature varying.  The (non-tempered) posterior is a multimodal distribution
-%% with modes at $(0,1)$ and $(1,-1)$. A larger $T$ implies a flatter posterior,
-%% which can be useful for parameter exploration so that the samples do not get
-%% trapped in a mode.
-
-% Daniel: This was some stuff from the Logistic Regression appendix, but we no
-% longer need that.
-
-%% In the logistic regression experiment of Section~\ref{ssec:logistic}, the
-%% feature vector for an image consists of its pixel values, normalized between 0
-%% and 1. For simplicity, we consider the binary classification case, so we only
-%% use digits one ($y = 1$) and seven ($y = -1$).  The probability of the $i^{\rm
-%% th}$ output $y_i \in \{-1,+1\}$ with the input vector $x_i$ is
-%% \begin{equation}\label{eq:logistic_probability}
-%% p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)},
-%% \end{equation}
-%% where $\theta$ is the 784-length parameter vector.
-
-% Daniel: here are example LaTeX codes for figures and tables if we want to use them.
-%\begin{figure}[h]
-%  \centering
-%  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-%  \caption{Sample figure caption.}
-%\end{figure}
-%\begin{table}[t]
-%  \caption{Sample table title}
-%  \label{sample-table}
-%  \centering
-%  \begin{tabular}{lll}
-%    \toprule
-%    \multicolumn{2}{c}{Part}                   \\
-%    \cmidrule{1-2}
-%    Name     & Description     & Size ($\mu$m) \\
-%    \midrule
-%    Dendrite & Input terminal  & $\sim$100     \\
-%    Axon     & Output terminal & $\sim$10      \\
-%    Soma     & Cell body       & up to $10^6$  \\
-%    \bottomrule
-%  \end{tabular}
-%\end{table}
-%\usepackage[pdftex]{graphicx} ...
-%\includegraphics[width=0.8\linewidth]{myfile.pdf}
-
 \end{document}

### 2016-10-10 aistats2017.tex
update the first and second section of the paper, delete some sentences, still too long

@@
 Minibatch updates produce many improvements to the model for each pass
-over the dataset, and have high sample efficiency. They also map very
-well onto hardware such as GPUs. In contrast, M-H requires
-calculations over the full dataset to produce a new sample.  Recent
+over the dataset, and have high sample efficiency. %They also map very
+% well onto hardware such as GPUs. 
+In contrast, M-H requires calculations over the full dataset to produce a new sample.  Recent
 results from~\citet{cutting_mh_2014} and ~\citet{icml2014c1_bardenet14} perform
@@
 bound, and such bounds are only available for relatively simple
-distributions. By contrast our approach is a ``black box'' and requires
-only statistics gathered from the minibatches as they are used.
-
+distributions. %By contrast our approach is a ``black box'' and requires
+%only statistics gathered from the minibatches as they are used. 
 Here we derive a new test which incorporates the variability in minibatch
@@
 test, which makes our test naturally error tolerant. The idea of using
-a noise-tolerant test using Barker's test function was suggested but
+a noise-tolerant %test using
+Barker's test function was suggested but
 not explored empirically in~\citet{TallData15} section 6.3. But the
@@
 \begin{itemize}
-
 \item Increase the temperature of the target distribution. Log likelihoods
@@
   higher temperature can be advantageous for parameter exploration.
-
 \item For continuous probability distributions, reduce the proposal
@@
   square of proposal step size. 
-
 \item Increase the minibatch size when needed for certain
@@
   for the test.
-
 \end{itemize}
@@
 Bayesian inference task is asymptotically normal, and has variance
-that scales inversely with the number of data samples $N$. This mode
-is extremely sharp for large datasets, which may contain millions or
-billions of samples. Simply sampling from this distribution is one
+that scales inversely with the number of data samples $N$. %This mode
+%is extremely sharp for large datasets, which may contain millions or
+%billions of samples. 
+Simply sampling from this distribution is one
 application, but an efficient proposal distribution~\citep{OptimalScaling01} has similar variance to the target
@@
 progress to a posterior mode with each step, taking small steps so
-that the cumulative displacement has progressively lower variance. A
-substantial part of the computational work of MCMC on massive datasets
-will be similarly in reaching a stationary distribution, which really
-means finding a deep posterior mode. Taking noisy small steps will
-nevertheless make steady progress to a posterior mode since their bias
-is in that direction. We demonstrate this behavior in our logistic regression
-experiments.
+that the cumulative displacement has progressively lower variance. %A
+%substantial part of the computational work of MCMC on massive datasets
+%will be similarly in reaching a stationary distribution, which really
+%means finding a deep posterior mode. 
+Taking noisy small steps will nevertheless make steady progress to 
+a posterior mode since their bias is in that direction. We demonstrate 
+this behavior in our logistic regression experiments.
 
 The contributions of this paper are as follows:
-
 \begin{itemize}
@@
   uses a novel additive correction variable to implement a Barker test based on minibatch mean and variance. 
-\item We analyze the test for accuracy and speed.
+%\item We analyze the test for accuracy and speed.
 \item We compared performance of our new test and prior approaches on several datasets. We demonstrate the test is several orders
@@
 unique stationary distribution $\pi(\theta) = p(\theta)$.
-
-For Bayesian inference, we would like to sample from a parameter distribution
-for $\theta$ based on some observed data, so the target distribution is
+For Bayesian inference, we would like to sample from a parameter distribution for $\theta$ based on some observed data, so the target distribution is
 $p(\theta \mid x_1, \ldots, x_N)$. The acceptance probability is now:
@@
 Metropolis-Hasting tests using sequential hypothesis testing. During each
-iteration, they start with a small minibatch of data and test the hypothesis
-that the sample $\theta'$ should be accepted based on an approximate version of
+iteration, they start with a small minibatch of data and test whether the sample $\theta'$ 
+should be accepted based on an approximate version of
 the test $u < \alpha(\theta,\theta')$. If the approximate test cannot make a
@@
 \begin{split} 
-\Lambda(\theta,\theta') & = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
-\hspace{0.2in} \\
-{\rm and} \hspace{0.2in}
-\psi(u,\theta,\theta') & = \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right)
+\Lambda(\theta,\theta') & = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right) \\  
+{\rm and} \quad \psi(u,\theta,\theta') &= \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right) \\
 \end{split}
@@
 \end{equation}
-Thus, $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over
-the entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean of
-$\Lambda_i(\theta,\theta')$ over its minibatch.
+%Thus, $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over
+%the entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean of
+%$\Lambda_i(\theta,\theta')$ over its minibatch.
 
@@
 \begin{equation}\label{eq:lemma_ll_ratio}
-\begin{split}
-\Lambda^*(\theta,\theta') & =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)} \\
-& = N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right) \\
-\end{split}
+\resizebox{\linewidth}{!}{% 
+$\Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)} = N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)$
+}%
 \end{equation}
@@
 \end{lemma}
-
+\vspace{-1em}
 \begin{proof}
@@
 \end{proof}
-
 Similar results can be shown for other distributions and proposals by

### 2016-10-10 aistats2017.tex
added noitemsep
Made figures full-width (you just have to add asterisks) and a few other minor changes to shorten paper and remove small math font.

@@
 
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item Increase the temperature of the target distribution. Log likelihoods
@@
 The contributions of this paper are as follows:
-\begin{itemize}
-\item We develop a new, more efficient (in samples per test) minibatch acceptance test with quantifiable error bounds. The test
-  uses a novel additive correction variable to implement a Barker test based on minibatch mean and variance. 
+
+\begin{itemize}[noitemsep]
+    \item We develop a new, more efficient (in samples per test) minibatch
+    acceptance test with quantifiable error bounds. The test uses a novel
+    additive correction variable to implement a Barker test based on minibatch
+    mean and variance.
+
     %\item We analyze the test for accuracy and speed.
-\item We compared performance of our new test and prior approaches on several datasets. We demonstrate the test is several orders
-  of magnitude more efficient than prior work measured as data consumed per test, and that it does not suffer
-  from long-tailed minibatch sizes (up to the dataset size). 
+
+    \item We compare performance of our new test and prior approaches on several
+    datasets. We demonstrate the test is several orders of magnitude more
+    efficient than prior work measured as data consumed per test, and that it
+    does not suffer from long-tailed minibatch sizes (up to the dataset size). 
 \end{itemize}
@@
 $u<\alpha(\theta_t,\theta')$; here, $\mathcal{U}(a,b)$ denotes the uniform
-distribution on the interval $[a,b]$.  For notational simplicity, we drop the
-subscript $t$ for the current sample $\theta_t$ and denote it as $\theta$.
+distribution on the interval $[a,b]$.  For simplicity, we drop the subscript $t$
+for the current sample $\theta_t$ and denote it as $\theta$.
 
@@
 the test repeats. This process continues until a decision. The bounds depend on
-either a asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
+either an asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
 concentration bound~\citep{icml2014c1_bardenet14}. The latter requires direct
@@
 ratio differences between the minibatch and full batch versions.  In the worst
-case, all $N$ data points may be needed, and we soon show that this worst case
-can occur about $\Omega(N)$ times during the performance of $N$ tests.
-
-Following~\citet{icml2014c1_bardenet14}, we write the test $u <
-\alpha(\theta,\theta')$ in the equivalent form $\Lambda(\theta,\theta') >
-\psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those in
-\citet{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later.}
+case, all $N$ data points may be needed, as we show in
+Section~\ref{ssec:gaussian_example}.
+
+Following~\citet{icml2014c1_bardenet14}, we write the test
+$u<\alpha(\theta,\theta')$ in the equivalent form $\Lambda(\theta,\theta') >
+\psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those
+in~\citet{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis
+later.}
 \begin{equation}
@@
 were no worst-case bounds given on the average number of samples
-needed per iteration.
-
-We next show that for a simple distribution, the lower bound of average
-number of data instances consumed for one iteration of~\citet{cutting_mh_2014}
-and~\citet{icml2014c1_bardenet14} is $\Omega(N)$, where $N$ is the number of data
-points.
+needed per iteration, which we explore in the following section.
 
@@
 \begin{equation}\label{eq:lemma_ll_ratio}
-\resizebox{\linewidth}{!}{% 
-$\Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)} = N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)$
-}%
+    \Lambda^*(\theta,\theta') = N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
 \end{equation}
@@
 
-\begin{itemize}
+\begin{itemize}[noitemsep]
     \item It uses an adaptive step size so as to use the smallest possible
@@
 times to collect 5000 samples.
-\begin{figure}
+\begin{figure*}[t]
     \centering
@@
     \label{fig:gauss_mix_1}
-%\end{figure*}
-%\begin{figure*}[t!]
+\end{figure*}
+\begin{figure*}[t]
     \centering
@@
     \label{fig:gauss_mix_2}
-\end{figure}
+\end{figure*}
 
@@
 %% the same plot. 
-\begin{table}[h]
+\begin{table}[t]
     \caption{Gaussian Mixture Model Statistics}
@@
     \centering
-    \resizebox{.5\textwidth}{!}{% <------ Don't forget this %
+    \resizebox{.45\textwidth}{!}{% <------ Don't forget this %
     \begin{tabular}{l l l l}
     %\toprule
-    {\bf Model} & {\bf Our Method} & {\bf Korattikara 2014} & {\bf Bardenet 2014} \\
+    {\bf Metric} & {\bf Ours} & {\bf Korat.'14} & {\bf Barde.'14} \\
     %\midrule
     \hline \\
-    Poisson Value (Eq.~\ref{eq:log_prob_poissons}) & -1430.0 & -1578.9 & -1232.7 \\
-    Chi-Squared Test Statistic & 3313.9 & 3647.7 & 2444.1 \\
+    Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
+    Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
     % \bottomrule
@@
 \subsection{Logistic Regression}\label{ssec:logistic}
-\begin{figure}
+
+\begin{figure*}[t]
 	\centering
@@
 	\label{fig:logistic_performance}
-%\end{figure*}
-%\begin{figure*}[t!]
+\end{figure*}
+\begin{figure*}[t]
 	\centering
@@
 	\label{fig:logistic_minibatch}
-\end{figure}
+\end{figure*}
 

### 2016-10-10 aistats2017.tex
fixed spacing by references
Shortened paper more, we are almost 3 pages over

@@
 
-\section{Introduction}\label{sec:introduction}
-Markov chain Monte Carlo (MCMC) sampling is a powerful method for
-computation on intractable distributions. We are interested primarily
-in large dataset applications, where the goal is to sample a posterior
-distribution $p(\theta \mid x_1, \ldots, x_N)$ of parameter $\theta$,
-and where the number of data instances $N$ is large.  The
-Metropolis-Hastings method (M-H) generates sample candidates from a
-proposal distribution $q$ which is in general different from the
-target distribution $p$, and decides whether to accept or reject them
-based on an acceptance test. The acceptance test is usually a
-Metropolis test~\citep{Metropolis1953, hastings70}. Conventional
-Metropolis-Hastings requires all $N$ data instances to generate one
-posterior sample.
-
-Many state-of-the-art machine learning methods, and deep learning in
-particular, are based on minibatch updates (such as SGD) to a model.
-Minibatch updates produce many improvements to the model for each pass
-over the dataset, and have high sample efficiency. %They also map very
-% well onto hardware such as GPUs. 
-In contrast, M-H requires calculations over the full dataset to produce a new sample.  Recent
-results from~\citet{cutting_mh_2014} and ~\citet{icml2014c1_bardenet14} perform
-approximate (bounded error) acceptance tests using subsets
-(minibatches) of the full dataset. The tests depend on minibatch
-statistics, and on the value of an additional uniform random variable
-$u$. The amount of data consumed for each test varies significantly
-from one minibatch to the next, and depends on the current sample, the
-proposed sample, and on the random variable $u$. By
-contrast,~\citet{conf/uai/MaclaurinA14,TallData15} perform exact
-tests but require a lower bound on parameter distribution across its
-domain.  The amount of data reduction depends on the accuracy of this
-bound, and such bounds are only available for relatively simple
-distributions. %By contrast our approach is a ``black box'' and requires
-%only statistics gathered from the minibatches as they are used. 
+\section{INTRODUCTION}\label{sec:introduction}
+
+Markov chain Monte Carlo (MCMC) sampling is a powerful method for computation on
+intractable distributions. We are interested in large dataset applications,
+where the goal is to sample a posterior distribution $p(\theta \mid x_1, \ldots,
+x_N)$ of parameter $\theta$ for large $N$.  The Metropolis-Hastings method (M-H)
+generates sample candidates from a proposal distribution $q$ which is in general
+different from the target distribution $p$, and decides whether to accept or
+reject based on an acceptance test. The acceptance test is usually a Metropolis
+test~\citep{Metropolis1953, hastings70}.
+
+Many state-of-the-art machine learning methods, and deep learning in particular,
+are based on minibatch updates (such as SGD) to a model.  Minibatch updates
+produce many improvements to the model for each pass over the dataset, and have
+high sample efficiency.  In contrast, conventional M-H requires calculations
+over the full dataset to produce a new sample.  Recent results
+from~\citet{cutting_mh_2014} and~\citet{icml2014c1_bardenet14} perform
+approximate (bounded error) acceptance tests using subsets of the full dataset.
+The tests depend on minibatch statistics, and on the value of an additional
+uniform random variable $u$. The amount of data consumed for each test varies
+significantly from one minibatch to the next, and depends on the current sample,
+the proposed sample, and on the random variable $u$. By
+contrast,~\citet{conf/uai/MaclaurinA14,TallData15} perform exact tests but
+require a lower bound on parameter distribution across its domain.  The amount
+of data reduction depends on the accuracy of this bound, and such bounds are
+only available for relatively simple distributions.
+
 Here we derive a new test which incorporates the variability in minibatch
-statistics as {\em a natural part of the test}. Because of this, the amount of
-data required for each test is fixed with high probability and in expected
-value.
-
-We use a Barker test function~\citep{Barker65} rather than a Metropolis
-test, which makes our test naturally error tolerant. The idea of using
-a noise-tolerant %test using
-Barker's test function was suggested but
-not explored empirically in~\citet{TallData15} section 6.3. But the
-asymptotic test statistic CDF and the Barker function are different,
-which leads to fixed errors for the approach
-in~\citet{TallData15}. Here we show that the difference between the
-distributions can be corrected with an additive random variable. This
-leads to a test which is fast, and whose error can be made arbitrarily
-small.
+statistics as {\em a natural part of the test} and requires less data per
+iteration than prior work. We use a Barker test function~\citep{Barker65}, which
+makes our test naturally error tolerant. The idea of using a noise-tolerant
+Barker's test function was suggested but not explored empirically
+in~\citet{TallData15} section 6.3. But the asymptotic test statistic CDF and the
+Barker function are different, which leads to fixed errors for the approach
+in~\citet{TallData15}. Here, we show that the difference between the
+distributions can be corrected with an additive random variable. This leads to a
+test which is fast, and whose error can be made arbitrarily small.
 
@@
     \item Increase the temperature of the target distribution. Log likelihoods
-  scale as $1/T$, and so the variance of the likelihood ratio will
-  vary as $1/T^2$. Our model is no longer well-specified (we are doing inference
-  at a temperature different from that assumed during data generation), but
-  higher temperature can be advantageous for parameter exploration.
-\item For continuous probability distributions, reduce the proposal
-  step size and variance (for stochastic proposals) compared to an optimal
+    scale as $1/T$, and so the variance of the likelihood ratio will vary as
+    $1/T^2$. Our model is no longer well-specified (we are doing inference at a
+    temperature different from that assumed during data generation), but higher
+    temperature can be advantageous for parameter exploration.
+
+    \item For continuous probability distributions, reduce the proposal step
+    size and variance (for stochastic proposals) compared to an optimal
     proposal. The variance of the log acceptance probability scales as the
     square of proposal step size.
-\item Increase the minibatch size when needed for certain
-  minibatches. Log acceptance variance scales as $1/k$ vs the
-  minibatch size $k$. Our test is adaptive like earlier works, but
-  unlike them, the distribution of minibatch size is Gaussian, not
-  long-tailed.  Increased minibatch size also reduces the error rate
-  for the test.
+
+    \item Increase the minibatch size when needed for certain minibatches. Log
+    acceptance variance scales as $1/k$ vs the minibatch size $k$. Our test is
+    adaptive like earlier works, but unlike them, the distribution of minibatch
+    size is Gaussian, not long-tailed.  Increased minibatch size also reduces
+    the error rate for the test.
 \end{itemize}
 
-It's worth discussing at this point the typical goals of
-M-H sampling on very large datasets.  By the Bernstein-von Mises
-Theorem, the posterior distribution of the parameter $\theta$ for a
-Bayesian inference task is asymptotically normal, and has variance
-that scales inversely with the number of data samples $N$. %This mode
-%is extremely sharp for large datasets, which may contain millions or
-%billions of samples. 
-Simply sampling from this distribution is one
-application, but an efficient proposal distribution~\citep{OptimalScaling01} has similar variance to the target
-distribution and will diffuse to it extremely slowly from an
-initialization value which is (likely to be) many standard deviations
-away. If there are any other strong modes, it is very likely for the
-sampler to find one of them and become trapped in it when run at the
-normal distribution temperature ($T=1$). A common solution is to anneal
-the sampler, running first at high temperature (scaling log
-likelihoods by $1/T$) which flattens the likelihood landscape.  This in
-turn reduces the variance of the log acceptance probability and allows
-our acceptance test to be applied.
-
-A second question concerns step size. Once we have fixed temperature,
-our variance constraint implies that we have to trade-off proposal
-step size $s$ and batch size $b$ ($b \propto p^2$), i.e. we can make
-many small steps, or one large step, with a given batch of data. One
-of the primary drivers of this work is our belief in the value of
-small steps. For applications to neural networks or other models where
-the posterior is multimodal, posterior inference is arguably a search
-process. Covering the search space densely with small steps is much
-more valuable than few sparse steps toward the nearest optimum. In
-this mode, Metropolis-Hastings can be used in similar fashion to
-Stochastic Gradient Descent. The goal in SGD is to make gradual
-progress to a posterior mode with each step, taking small steps so
-that the cumulative displacement has progressively lower variance. %A
-%substantial part of the computational work of MCMC on massive datasets
-%will be similarly in reaching a stationary distribution, which really
-%means finding a deep posterior mode. 
-Taking noisy small steps will nevertheless make steady progress to 
-a posterior mode since their bias is in that direction. We demonstrate 
-this behavior in our logistic regression experiments.
+It is worth discussing at this point the typical goals of M-H sampling on large
+datasets.  By the Bernstein-von Mises Theorem, the posterior distribution for a
+Bayesian inference task is asymptotically normal, and has variance that scales
+inversely with the $N$ data samples. Simply sampling from this distribution is
+one application, but an efficient proposal~\citep{OptimalScaling01} has similar
+variance to the target distribution and will diffuse to it extremely slowly from
+an initialization value which is (likely to be) many standard deviations away.
+If there are any other strong modes, it is very likely for the sampler to find
+one of them and become trapped in it when run at the normal distribution
+temperature ($T=1$). A common solution is to anneal the sampler, running first
+at high temperature (scaling log likelihoods by $1/T$) which flattens the
+likelihood landscape.  This in turn reduces the variance of the log acceptance
+probability and allows our acceptance test to be applied.
+
+A second question concerns step size. Once we have fixed temperature, our
+variance constraint implies that we have to trade-off proposal step size $s$ and
+batch size $b$ ($b \propto p^2$), i.e. we can make many small steps, or one
+large step, with a given batch of data. One of the primary drivers of this work
+is our belief in the value of small steps. For applications to neural networks
+or other models where the posterior is multimodal, posterior inference is
+arguably a search process. Covering the search space densely with small steps is
+much more valuable than few sparse steps toward the nearest optimum. In this
+mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
+Descent. The goal in SGD is to make gradual progress to a posterior mode with
+each step, taking small steps so that the cumulative displacement has
+progressively lower variance. We demonstrate this behavior in our logistic
+regression experiments.
 
@@
 
-\section{Preliminaries and Related Work}\label{sec:related_work}
+\section{PRELIMINARIES}\label{sec:related_work}
 
@@
 step size is severe, and the state space exploration reduces to a random walk.
-Full minibatch HMC for minibatches was recently described in~\citet{sghmc_2014}
-which allows momentum-augmented proposals with larger step sizes. However, step
-sizes are still limited by the need to run accurately without MH tests.  By providing
-an M-H test with similar cost to standard gradient steps, our
-work opens the door to applying those methods with much more aggressive step
-sizes without loss of accuracy. 
-%{\color{blue}
-%Daniel: do we ever demonstrate this? I guess another issue is that this implies
-%we can use larger step sizes, but earlier we \emph{also} argue that we need
-%\emph{small} step sizes (to reduce variance).}
-
-\section{A New Metropolis-Hastings Acceptance Test}\label{sec:our_algorithm}
+Full minibatch HMC for minibatches was described in~\citet{sghmc_2014} which
+allows momentum-augmented proposals with larger step sizes. However, step sizes
+are still limited by the need to run accurately without MH tests.  By providing
+an M-H test with similar cost to standard gradient steps, our work opens the
+door to applying those methods with much more aggressive step sizes without loss
+of accuracy.
+
+\section{A NEW MH ACCEPTANCE TEST}\label{sec:our_algorithm}
 
@@
 
-For our new M-H test, we denote the exact approximate log likelihood ratios
-as $\Delta$ and $\Delta^*$.  First, $\Delta$ is defined as
+For our new M-H test, we denote the exact and approximate log likelihood ratios
+as $\Delta$ and $\Delta^*$, respectively. First, $\Delta$ is defined as
 \begin{equation}\label{eq:delta1}
@@
 $\Delta^*(\theta,\theta')$ is an unbiased estimator of $\Delta(\theta,\theta')$.
-%{\color{blue} Daniel: a brief question repeated from earlier, should we mention
-%sampling with and/or without replacement here?}
 
@@
 of data.
-%{\color{blue} Daniel: aside from the experiments, where do we show it
-%``naturally tolerates substantial variance'' in the input?}
 
@@
 If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0,
-\sigma^2(\Delta^*))$, the above test is exact as well, and as we show in
+\sigma^2(\Delta^*))$, the above test is exact, and as we show in
 Section~\ref{sec:analysis}, if there is a maximum error $\epsilon$ between the
 CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0,
-\sigma^2(\Delta^*))$, then the acceptance test has an error of at most
-$\epsilon$ relative to the full batch version.
+\sigma^2(\Delta^*))$, then our test has an error of at most $\epsilon$ relative
+to the full batch version.
 
@@
 
-\section{Computing the Correction Distribution}\label{sec:correction}
+\section{COMPUTING THE CORRECTION DISTRIBUTION}\label{sec:correction}
 
@@
 
-For computation of $C_\sigma$, we assume that its input $Y$ and
-another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
-In continuous form, the optimization problem is therefore
-\begin{equation}\label{eq:conv1}
-    \small{C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|}
-\end{equation}
+For computation of $C_\sigma$, we assume that its input $Y$ and another variable
+$X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.  In continuous
+form, the optimization problem is to find the $C_\sigma^*$ solving
+\[
+%\begin{equation}\label{eq:conv1}
+    \argmin_{C\sigma} \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|.
+%\end{equation}
+\]
 To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
@@
 $j \in \{-N, \ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$,
-and the condition in Equation~(\ref{eq:conv1}) can be re-expressed as:
-\begin{equation}\label{eq:conv2}
+and the objective can be re-expressed as:
+\[
+%\begin{equation}\label{eq:conv2}
     C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \mathcal{I}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
-\end{equation}
-To make the problem easier to understand, we can define a matrix $M$ and vectors
-$u$ and $v$ such that $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, $u_j =
-C_{\sigma}(Y_j)$, and $v_i = S(X_i)$, where the indices $i$ and $j$ are
-appropriately translated to be non-negative indices for $M, u,$ and $v$. Thus,
-Equation~(\ref{eq:conv2}) is equivalent to minimizing the $L_{\infty}$ norm:
+%\end{equation}
+\]
+We now define a matrix $M$ and vectors $u$ and $v$ such that $M_{ij} =
+\Phi_{\sigma}(X_i-Y_j)$, $u_j = C_{\sigma}(Y_j)$, and $v_i = S(X_i)$, where the
+indices $i$ and $j$ are appropriately translated to be non-negative indices for
+$M, u,$ and $v$. Thus, the problem is now to minimize the $L_{\infty}$ norm:
 \begin{equation}\label{eq:optimization_l1}
@@
 \end{equation}
-
 We have the additional constraint that $u_j > 0$ for all $j$, since $u$
@@
 \end{equation}
-where we add the standard regularization factor $\lambda$ for stability in high
-dimensions. The solution is well-known from the normal equations: $u^* = (M^TM +
-\lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_{\infty}$ norm for
+where we add the standard regularization factor $\lambda$. The solution is
+well-known from the normal equations: $u^* = (M^TM + \lambda I)^{-1}M^Tv$, and
+in practice, it yields an acceptable $L_{\infty}$ norm for
 Equation~(\ref{eq:optimization_l1}).
-
-With this approach, there is no guarantee that $u^* \geq 0$. However, we have
-some flexibility in the choice of $\sigma$ in Equation~(\ref{eq:conv1}).
-As we decrease the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$
-grows by the same amount and is in fact the result of convolution with a
-Gaussian whose variance is the difference.  Thus as $\sigma$ decreases,
-$C_\sigma(X)$ grows and approaches the derivative of a logistic function at
-$\sigma = 0$. It retains some very weak negative values for $\sigma > 0$ but
-removal of those values leads to very small error.
 
@@
 \For{$t=\{1, \ldots, T\}$}{
-    Propose a candidate $\theta'$\ from proposal distribution $q(\theta'\mid\theta_t)$\;
+    Propose a candidate $\theta'$\ from proposal $q(\theta'\mid\theta_t)$\;
     Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*}$\;
@@
     \begin{tabular}{l l l l}
-   % \toprule
     ${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error} \\
-    \hline \\
+    \hline
     4000 & 0.9 & 1 & 1.0e-4  \\
     4000 & 0.8 & 0.03 & 5.0e-6 \\
-    %\bottomrule
     \end{tabular}
@@
 
-As can be seen from Table~\ref{tab:xcorr}, the errors between $X_{\rm
-norm}+X_{\rm corr}$ and $X_{\rm log}$ can be made very small, approaching single
-floating precision (about $10^{-7}$).
-
-The complete procedure is given in Algorithm~\ref{alg:our_algorithm}. A few points:
+With this approach, there is no guarantee that $u^* \geq 0$. However, we have
+some flexibility in the choice of $\sigma$ in Equation~(\ref{eq:overall_corr_problem}).
+As we decrease the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$
+grows by the same amount and is in fact the result of convolution with a
+Gaussian whose variance is the difference.  Thus as $\sigma$ decreases,
+$C_\sigma(X)$ grows and approaches the derivative of a logistic function at
+$\sigma = 0$. It retains some very weak negative values for $\sigma > 0$ but
+removal of those values leads to very small error. Table~\ref{tab:xcorr} shows
+that the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be
+made very small, approaching single floating precision (about $10^{-7}$).
+
+Algorithm~\ref{alg:our_algorithm} describes our procedure. A few points:
 
@@
 
-\section{Analysis}\label{sec:analysis}
+\section{ANALYSIS}\label{sec:analysis}
 
 We now derive error bounds for our M-H test, and for the approximate target
-distribution that it generates. From Table~\ref{tab:xcorr}, we know that
-it is possible to generate the correction samples $X_{\rm corr}$ with a CDF
-error approaching single-precision floating point error. We therefore treat the
-$X_{\rm corr}$ variable as a sample from the exact correction distribution and
-we will not analyze its errors.
+distribution that it generates. From Table~\ref{tab:xcorr}, we know that it is
+possible to generate the correction samples $X_{\rm corr}$ with a CDF error
+approaching single-precision floating point error. We therefore treat $X_{\rm
+corr}$ as a sample from the exact correction distribution and we will not
+analyze its errors.
 
 In the most similar prior works,~\citet{cutting_mh_2014} uses asymptotic
-arguments based on the Central Limit Theorem to argue that its approximate
-acceptance test error tends to zero as batch size increases. But no quantitative
-bounds are given. In~\citet{icml2014c1_bardenet14}, explicit bounds are given,
-but they depend on bounding
+arguments based on the CLT to argue that its approximate acceptance test error
+tends to zero as batch size increases, but no quantitative bounds are given.
+In~\citet{icml2014c1_bardenet14}, explicit bounds are given, but they depend on
+bounding:
 \begin{equation}\label{eq:bad_bound}
@@
 The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
-\begin{equation}\label{eq:clt-boundsx}
-\begin{split}
+%% \begin{equation}\label{eq:clt-boundsx}
+%% \begin{split}
+%% -6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
+%% &\leq 1.36E|X|^3.
+%% \end{split}
+%% \end{equation}
+\begin{align*}
 -6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
 &\leq 1.36E|X|^3.
-\end{split}
-\end{equation}
+\end{align*}
 This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
@@
 \end{equation}
-so that $s_X = s_{\Delta^*}$.
-%{\color{blue} Daniel: This time it
-%  makes sense to use $\sigma(\hat{X})$ instead of $\sigma(X)$.}
-This results in the following Corollary:
+so that $s_X = s_{\Delta^*}$. This results in the following:
 
@@
 \begin{split}
--\epsilon &= \tiny{\int_{-\infty}^{+\infty} -\epsilon R(x) dx } \\
+-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
 &\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
@@
 \end{equation}
-Now we convolve with the distribution of $X_{\rm corr}$ which gives:
+Now we convolve with the distribution of $X_{\rm corr}$:
 \begin{equation}
@@
 
-\section{Experiments}\label{sec:experiments}
+\section{EXPERIMENTS}\label{sec:experiments}
 
@@
 times to collect 5000 samples.
+
 \begin{figure*}[t]
     \centering
-    \includegraphics[width=\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
+    \includegraphics[width=0.9\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
     \caption{
@@
     \centering
-    \includegraphics[width=\linewidth]{minibatch_size_gaussian.png}
+    \includegraphics[width=0.9\linewidth]{minibatch_size_gaussian.png}
     \caption{
@@
 %% the same plot.
+
 \begin{table}[t]
@@
     \begin{tabular}{l l l l}
-    %\toprule
     {\bf Metric} & {\bf Ours} & {\bf Korat.'14} & {\bf Barde.'14} \\
-    %\midrule
-    \hline \\
+    \hline
     Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
     Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
-    % \bottomrule
     \end{tabular}% <------ Don't forget this %
@@
 	\centering
-	\includegraphics[width=\linewidth]{logistic_performance.png}
+	\includegraphics[width=0.9\linewidth]{logistic_performance.png}
 	\caption{
@@
 	\centering
-	\includegraphics[width=\linewidth]{minibatch_size_logistic.png}
+	\includegraphics[width=0.9\linewidth]{minibatch_size_logistic.png}
 	\caption{
@@
 
-\section{Conclusions}\label{sec:conclusion}
+\section{CONCLUSIONS}\label{sec:conclusion}
 
@@
 \subsubsection*{References}
+% Daniel: Sorry, I don't know why this is needed to get rid of the annoying
+% spaces below the "References" text.
+\vspace{-20pt} 
 %\bibliographystyle{apalike}
@@
 \bibliography{aistats2017}
-\normalsize
 

### 2016-10-10 aistats2017.tex
More shortening, now about 10.75 pages. But I want to get back to experiments.

@@
     Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*}$\;
-    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Equation~(\ref{eq:clt-bounds3})\;
+    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Corollary~\ref{cor:our_bound_delta_prime}\;
     \While {$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}{
@@
 We can now substitute into Equation~(\ref{eq:clt-bounds2}) and displace by the mean, giving:
-\begin{equation}\label{eq:clt-bounds3}
-\begin{split}
+\begin{align*}
     \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| &\leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} \\
-    & = \epsilon(\theta,\theta',b) \\
-\end{split}
-\end{equation}
+    &= \epsilon(\theta,\theta',b).
+\end{align*}
 \end{corollary}
-
 Corollary~\ref{cor:our_bound_delta_prime} shows that the distribution of
@@
 \begin{proof}
-We first observe that
-\begin{equation}\label{eq:pdf_difference}
-    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
-\end{equation}
-and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
-\begin{equation}
-\begin{split}
--\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
-&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
-&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon.
-\end{split}
-\end{equation}
+See Appendix~\ref{app:proof_cdf_bounds}.
 \end{proof}
-\vspace{-10pt}
 
@@
 posterior of $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$ with the same
-prior from Equation~(\ref{eq:data_generation}). The larger number of samples
-produces a much sharper posterior with two very narrow peaks.  Our goal is to
-reproduce the original posterior, so we adjust the temperature to $T=10,000$.
-Taking logs, we get the target as shown in the far left of
-Figure~\ref{fig:gauss_mix_1}.
-
-We run MCMC with our MH test using minibatch size $m=100$.  We also run this
-using the methods from~\citet{cutting_mh_2014} %(``Korattikara 2014'')
-and~\citet{icml2014c1_bardenet14}, also with $m=100$.
+prior from Equation~(\ref{eq:data_generation}). This produces a much sharper
+posterior with two very narrow peaks.  Our goal is to reproduce the original
+posterior, so we adjust the temperature to $T=10,000$.  Taking logs, we get the
+target as shown in the far left of Figure~\ref{fig:gauss_mix_1}.
+
+We run MCMC with our MH test and benchmark with~\citet{cutting_mh_2014}
+and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.
 For~\citet{cutting_mh_2014}, we increment minibatches by 100 as recommended, and
@@
 minibatch size updates because of their different approaches to repeated
-testing. \citet{icml2014c1_bardenet14} uses straightforward discounting,
-allocating a fraction of the test error $\delta$ to each minibatch test (call
-these subtests). It's important to bound the number of subtests therefore, so
-that the tolerable error per subtest is not too small.  \citet{cutting_mh_2014}
-uses a more aggressive repeated testing approach, explicitly modeling
-correlation between tests and allowing many more tests for a given $\delta$.
+testing. The former uses a more aggressive repeated testing approach, explicitly
+modeling correlation between tests and allowing many more tests for a given
+$\delta$.  The latter uses straightforward discounting, allocating a fraction of
+the test error $\delta$ to each minibatch test.
 
@@
 samples for the three methods, with darker regions indicating a greater density
-of points. There are no obvious differences, so to better understand any
-differences, we measure the similarity between each set of samples and the
-actual posterior.
-
-To effect the comparison, we first discretize the posterior coordinates into
-bins (non-overlapping ranges) with respect to the two components of the
-parameter $\theta$.  The probability of a sample falling into one of these bins
-is the integral of the true posterior probability over the area of that bin. Let
-this probability be denoted $P_i$ where $i$ is the bin number. A single sample
-from any of the MH methods should therefore be multinomial with distribution
-$P$, and a set of $n$ (ideally independent) samples should follow a ${\rm
-Multinomial}(P,n)$ distribution.  Since the ideal distribution is simple, we can
-use it to measure the quality of the sample distributions rather than use
-general purpose tests like KL-divergence or likelihood-ratio.  The latter in any
-case are problematic with zero counts in some bins as we have here.
+of points. There are no obvious differences, so we measure the similarity
+between each set of samples and the actual posterior.
+
+We discretize the posterior coordinates into bins (non-overlapping ranges) with
+respect to the two components of the parameter $\theta$.  The probability of a
+sample falling into one of these bins is the integral of the true posterior
+probability over the area of that bin. Let this probability be denoted $P_i$
+where $i$ is the bin number. A single sample from any of the MH methods should
+therefore be multinomial with distribution $P$, and a set of $n$ (ideally
+independent) samples should follow a ${\rm Multinomial}(P,n)$ distribution.
+Since the ideal distribution is simple, we can use it to measure the quality of
+the sample distributions rather than use general purpose tests like
+KL-divergence or likelihood-ratio.  The latter are problematic with zero counts
+in some bins as we have here.
 
@@
 While the Poisson results provide general guidance on the relative accuracy of
-the models, it is difficult to interpret the quantitative scores. We therefore
-perform significance tests to show the difference between the MCMC-sampled
-distributions and the ground-truth.  We perform an approximate test using
-the Chi-Squared distribution as the test statistic, which we also show in
-Table~\ref{tab:poissons}.
+the models, it is difficult to interpret the scores, so we perform significance
+tests to show the difference between the MCMC-sampled distributions and the
+ground-truth.  We perform an approximate test using the Chi-Squared distribution
+as the test statistic, which we also show in Table~\ref{tab:poissons}.
 
@@
 \begin{enumerate}
-    \item (Somewhat important) I added slightly more detail on why if we assume
-    $Y$ has variance $1/N$, then $Y \in \frac{1}{\sqrt{N}}[0.5,1]$ has
-    probability $\Phi(1)-\Phi(0.5)$; note that I'm assuming the positive
-    interval only, not the negative one. However, you also made $p_0$ double the
-    value of $p_1$. I struggled to understand why, until I realized that you
-    used the phrase ``with matching sign.'' So I \emph{think} you meant that
-    $\theta'-\theta$ could be either positive or negative, but that $\theta-0.5$
-    was restricted in its sign so could be only half its value. Is my thinking
-    right?
-
-    \item Changed all occurrences of $\Lambda^*(\theta',\theta)$ to
-    $\Lambda^*(\theta,\theta')$, i.e., switched the input ordering.
-
-    \item Originally, $\log u_0$ was defined as
-    $N(\Lambda^*(\theta,\theta')-\psi(1,\theta,\theta'))$ but I changed the
-    starting definition so it does \emph{not} have a factor of $N$ in its
-    definition (because we later get $N$ to appear in a subsequent step).
-
-    \item In addition, $\psi(1,\theta,\theta') = 0$ due to uniform priors and
-    symmetric proposals and there didn't seem to be a good enough reason to
-    include it in the definition so I deleted it (sometimes it's insightful to
-    have a nice form of zero inside, but I didn't see it here).
-
-    \item I added in an ``$E$'' because it seemed like you wanted the ``$u$''
-    that would make $\psi(u,\theta,\theta') = \log u =
-    E[\Lambda^*(\theta,\theta')]$.
-
-    \item Added more detail for the $\Lambda^*(\theta,\theta') \not\approx
-    \psi(u,\theta,\theta')$ part.
-
     \item (TODO later) After Equation~\ref{eq:accept_test_zero}, we say that
@@
 
+\section{Proof of Lemma~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}
+
+We first observe that
+\begin{equation}\label{eq:pdf_difference}
+    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
+\end{equation}
+and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
+\begin{align*}
+-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
+&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
+&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
+\end{align*}
+as desired.
+
 \end{document}

### 2016-10-10 aistats2017.tex
More shortening of paper. OK now I will stop for now

@@
 \Lambda(\theta,\theta') & = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right) \\  
-{\rm and} \quad \psi(u,\theta,\theta') &= \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right) \\
+{\rm and} \quad \psi(u,\theta,\theta') &= \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right).
 \end{split}
@@
 \end{equation}
-Finally, it will be convenient for our analysis to define the individual
-components that contribute to the sums above:
-\begin{equation}\label{eq:individual_terms}
-\Lambda_i(\theta,\theta') = N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
-\end{equation}
-%Thus, $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over
-%the entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean of
-%$\Lambda_i(\theta,\theta')$ over its minibatch.
+Finally, it will be convenient for our analysis to define
+$\Lambda_i(\theta,\theta') = N\log(\frac{p(x_i|\theta')}{p(x_i|\theta)})$.
+%% \begin{equation}\label{eq:individual_terms}
+%% \Lambda_i(\theta,\theta') = N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+%% \end{equation}
+Thus, $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over
+the entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean of
+$\Lambda_i(\theta,\theta')$ over its minibatch.
 
 Since minibatches contains randomly selected samples $x_i$, the values
-$\Lambda_i$ are independent, identically distributed (iid) random
+$\Lambda_i$ are i.i.d. random
 variables\footnote{The analysis assumes sampling with replacement
@@
 \subsection{A Worst-Case Gaussian Example}\label{ssec:gaussian_example}
-Consider a Gaussian model where $x_1,\ldots,x_N \simiid \mathcal{N}(\theta,1)$
+
+Consider a Gaussian model where $x_1,\ldots,x_N$ are i.i.d. $\mathcal{N}(\theta,1)$
 with a known variance $\sigma^2=1$, true mean $\theta=0.5$, and a uniform prior
@@
     \Delta(\theta,\theta') &= \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') \\
-    & = \Lambda(\theta,\theta') - \psi(1,\theta,\theta'). \\
+    & = \Lambda(\theta,\theta') - \psi(1,\theta,\theta').
 \end{split}
@@
 We show in Section~\ref{sec:correction} that it is possible to use deconvolution
-to derive a highly accurate representation of $C_{\sigma}(X)$.
-
-Using $X_{\rm corr}$ samples from $C_{\sigma}(X)$, the acceptance test is now
+to derive a highly accurate representation of $C_{\sigma}(X)$.  Using $X_{\rm
+corr}$ samples from $C_{\sigma}(X)$, the acceptance test is now
 \begin{equation}\label{eq:criteria}

### 2016-10-11 aistats2017.tex
Slight shortening to paper

@@
 Note that $\Delta$ and $\Delta^*$ are evaluated on the full dataset and a
-minibatch of size $b$ respectively. The scaling term $N/b$ ensures that
+minibatch of size $b$ respectively. The term $N/b$ means
 $\Delta^*(\theta,\theta')$ is an unbiased estimator of $\Delta(\theta,\theta')$.
@@
 \For{$t=\{1, \ldots, T\}$}{
-    Propose a candidate $\theta'$\ from proposal $q(\theta'\mid\theta_t)$\;
-    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*}$\;
-    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Corollary~\ref{cor:our_bound_delta_prime}\;
+    -Propose a candidate $\theta'$\ from proposal $q(\theta'\mid\theta_t)$\;
+    -Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*}$\;
+    -Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Corollary~\ref{cor:our_bound_delta_prime}\;
     \While {$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}{
-    	Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $s^2_{\Delta^*}$ and $\epsilon$ estimates\;
+    	-Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $s^2_{\Delta^*}$ and $\epsilon$ estimates\;
     }
-    Draw $X_{\rm nc} \sim \mathcal{N}(0,1-s^2_{\Delta^*})$ and
-    $X_{\rm corr}$ from the correction distribution $C_1(X)$\;
+    -Draw $X_{\rm nc} \sim \mathcal{N}(0,1-s^2_{\Delta^*})$ and $X_{\rm corr}$ from the correction distribution $C_1(X)$\;
 
         \eIf{$\Delta^* + X_{\rm nc} + X_{\rm corr} > 0$}{
-            Accept the candidate, $\theta_{t+1} = \theta'$\;
+            -Accept the candidate, $\theta_{t+1} = \theta'$\;
         }{
-            Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
+            -Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
         }
@@
 }
-\caption{A description of M-H sampling with our acceptance test.}
+\caption{Our acceptance test for MCMC.}
 \label{alg:our_algorithm}
@@
 
-In this paper, we use quantitative forms of the Central Limit Theorem which rely
-on measurable statistics from a single minibatch. Thus a sampler using our
-approach does not need to ``see'' data beyond the current minibatch. This
-supports use of the sampler on very large datasets, and in online mode where the
-dataset is presented as a stream.
-
-For the analysis, in Section~\ref{ssec:delta_star_distribution}, we first
-present bounds on the absolute and relative error (in terms of the CDFs) of the
-distribution of $\Delta^*$ vs a Gaussian. We then show in
-Section~\ref{ssec:preserve_bounds} that these error bounds are preserved after
-the addition of other random variables, in particular $X_{\rm nc}$ and $X_{\rm
-corr}$. From this it follows that the acceptance test has the same error bound.
+Here, we use quantitative forms of the Central Limit Theorem which rely on
+measurable statistics from a single minibatch. Thus a sampler using our approach
+does not need to see data beyond the current minibatch, and can easily be
+applied on large datasets and in online modes where the data is presented as a
+stream.
+
+In Section~\ref{ssec:delta_star_distribution}, we present bounds on the absolute
+and relative error (in terms of the CDFs) of the distribution of $\Delta^*$ vs.
+a Gaussian. We then show in Section~\ref{ssec:preserve_bounds} that these bounds
+are preserved after the addition of other random variables (e.g., $X_{\rm nc}$
+and $X_{\rm corr}$). It then follows that the acceptance test has the same error
+bound.
 
@@
 
-To start, we use the following quantitative central-limit result:
+We use the following quantitative central-limit result:
 
@@
 \begin{equation}\label{eq:clt-bounds2}
-   \sup_y\left|{\rm Pr}(\bar{X}< y) - \Phi\left(\frac{y}{s_X}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}},
+   \sup_y\left|{\rm Pr}(\bar{X}< y) - \Phi\left(\frac{y}{s_X}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
 \end{equation}
@@
 \end{align*}
-\end{corollary}
+\end{corollary} %\vspace{-10pt}
 Corollary~\ref{cor:our_bound_delta_prime} shows that the distribution of
@@
 
-Relating Lemma~\ref{lem:cdf_bounds} to our acceptance test, we have the
-following Corollary:
+From Lemma~\ref{lem:cdf_bounds}, we have the following Corollary:
 
@@
 \leq \epsilon(\theta,\theta',b)$, then
-\begin{equation}
+\[
     \sup_y|{\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
-\end{equation}
+\]
 where $S(x)$ is the standard logistic function, and $X_{\rm nc}$ and $X_{\rm corr}$ are generated as per Algorithm 1. 
@@
 and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.
-For~\citet{cutting_mh_2014}, we increment minibatches by 100 as recommended, and
+For~\citet{cutting_mh_2014}, we increment minibatches by 100 and
 for~\citet{icml2014c1_bardenet14}, we increase the minibatch size geometrically
-with a ratio of $\gamma = 1.5$ as recommended. These tests use different
-minibatch size updates because of their different approaches to repeated
-testing. The former uses a more aggressive repeated testing approach, explicitly
+with a ratio of $\gamma = 1.5$. These methods have different approaches to
+repeated testing: the former uses an aggressive testing approach, explicitly
 modeling correlation between tests and allowing many more tests for a given
-$\delta$.  The latter uses straightforward discounting, allocating a fraction of
-the test error $\delta$ to each minibatch test.
+$\delta$, while the latter uses discounting, allocating a fraction of the test
+error $\delta$ to each minibatch test.
 

### 2016-10-11 aistats2017.tex
The main part of the paper is now 10 pages.

@@
 The acceptance test guarantees detailed balance, which means
-$p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')$,
-where $p(\theta'\mid\theta)$ is the probability of a transition from state
-$\theta$ to $\theta'$. Here $p(\theta'\mid\theta) =
-q(\theta'\mid\theta)\alpha(\theta,\theta')$ so the detailed balance equation
-becomes 
-\begin{equation}\label{eq:detailed_balance2}
-    p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
-\end{equation}
+$p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')$, where
+$p(\theta'\mid\theta)$ is the probability of a transition from state $\theta$ to
+$\theta'$. Here, $p(\theta'\mid\theta)=q(\theta'\mid\theta)\alpha(\theta,\theta')$.
+% Daniel: I commented this out because people should know what it means and we
+%re-state another detailed balance equation later!
+%\begin{equation}\label{eq:detailed_balance2}
+%    p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
+%\end{equation}
 This condition, together with ergodicity, guarantees that the Markov chain has a
-unique stationary distribution $\pi(\theta) = p(\theta)$.
-For Bayesian inference, we would like to sample from a parameter distribution for $\theta$ based on some observed data, so the target distribution is
-$p(\theta \mid x_1, \ldots, x_N)$. The acceptance probability is now:
+unique stationary distribution $\pi(\theta) = p(\theta)$. For Bayesian
+inference, we would like to sample from a parameter distribution for $\theta$
+based on some observed data, so the target distribution is $p(\theta \mid x_1,
+\ldots, x_N)$. The acceptance probability is now:
 \begin{equation}\label{eq:acceptance_probability}
@@
 Following~\citet{icml2014c1_bardenet14}, we write the test
-$u<\alpha(\theta,\theta')$ in the equivalent form $\Lambda(\theta,\theta') >
+$u<\alpha(\theta,\theta')$ equivalently as $\Lambda(\theta,\theta') >
 \psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those
@@
 in number of samples required over using the full dataset, but there
-were no worst-case bounds given on the average number of samples
-needed per iteration, which we explore in the following section.
+were no worst-case bounds, which we explore in the following section.
 
@@
 
-Our goal is to sample from the posterior $p(\theta \mid x_1,\ldots,x_N)$, which
-is a normal distribution $\mathcal{N}(\mu, 1/N)$ centered on the sample mean
-$\mu$, and with variance $1/N$. In one dimension, an efficient proposal
+The posterior is a Gaussian centered on the sample mean $\mu$, and with variance
+$1/N$ (i.e., $\mathcal{N}(\mu, 1/N)$). In one dimension, an efficient proposal
 distribution has the same variance as the target
@@
 \begin{proof}
-See Appendix~\ref{app:worst_case_proof}.
+See Supplementary Material, Section~\ref{app:worst_case_proof}.
 \end{proof}
@@
 compared to full-dataset tests, so their speed-up is at most $1/c$.
-
-% {\color{blue} Daniel: quick question: I think we should explain why our method
-% \emph{doesn't} run into this issue. I think it is because we rely on an
-% approximation to $\Delta + X_{\rm log} > 0$ and that test does not require any
-% ``difference between an approximation and an exact version'' because it is
-% \emph{already} an exact version, assuming the approximation is enough.
-% Also, in terms of this example, if we are in the special case with using a lot
-% of data, why not force early stopping?}
 
@@
 The key to our test is a smooth acceptance function.  We consider functions
-other than the classical Metropolis test that satisfy the detailed balance (see
-Equation~(\ref{eq:detailed_balance2})) condition needed for accurate posterior
-estimation. A class of functions leading to detailed balance is specified as
-follows:
+other than the classical Metropolis test that satisfy the detailed balance
+condition needed for accurate posterior estimation. A class of suitable
+functions is specified as follows:
 
@@
 \[
-%\begin{equation}\label{eq:conv1}
     \argmin_{C\sigma} \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|.
-%\end{equation}
 \]
@@
 \[
-%\begin{equation}\label{eq:conv2}
     C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \mathcal{I}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
-%\end{equation}
 \]
@@
 indices $i$ and $j$ are appropriately translated to be non-negative indices for
-$M, u,$ and $v$. Thus, the problem is now to minimize the $L_{\infty}$ norm:
-\begin{equation}\label{eq:optimization_l1}
-    u^* = \argmin_{u} \|Mu-v\|_{\infty}.
-\end{equation}
-We have the additional constraint that $u_j > 0$ for all $j$, since $u$
-represents a density. We first explored optimizing this problem with linear
-programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
-
-This was tractable up to a few hundred dimensions for $u$.  However, the
-discretization error is bounded by the curvature of the underlying functions
-times the squared quantization step. The curvatures are here slightly less than
-one, i.e.  the errors are of the order of $(V/N)^2$.  Here we chose $V=20$ to
-provide adequate containment of the distributions (the CDFs are extremely close
-to either zero or one outside this range). So with 200 points, we have a
-discretization error of approximately 0.01, which is relatively large.  To yield
-higher resolution and lower error, we switched to a least squares solution:
+$M, u,$ and $v$. Thus, the problem is now to minimize $\|Mu-v\|_{\infty}$ with
+the constraint that $u > 0$ since it represents a density. We approximate this
+with a least squares solution:
 \begin{equation}\label{eq:optimization_l2}
@@
 well-known from the normal equations: $u^* = (M^TM + \lambda I)^{-1}M^Tv$, and
-in practice, it yields an acceptable $L_{\infty}$ norm for
-Equation~(\ref{eq:optimization_l1}).
+in practice, it yields an acceptable $L_{\infty}$ norm.
+
+% Daniel: This was taking up too much space, and we usually don't want to spend
+% a paragraph explaining what *didn't* work, and readers will know what it means
+% to minimize the L-infinity norm, we can put it inline.
+% \begin{equation}\label{eq:optimization_l1}
+%     u^* = \argmin_{u} \|Mu-v\|_{\infty}.
+% \end{equation}
+% We have the additional constraint that $u_j > 0$ for all $j$, since $u$
+% represents a density. We first explored optimizing this problem with linear
+% programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
+% 
+% This was tractable up to a few hundred dimensions for $u$.  However, the
+% discretization error is bounded by the curvature of the underlying functions
+% times the squared quantization step. The curvatures are here slightly less than
+% one, i.e.  the errors are of the order of $(V/N)^2$.  Here we chose $V=20$ to
+% provide adequate containment of the distributions (the CDFs are extremely close
+% to either zero or one outside this range). So with 200 points, we have a
+% discretization error of approximately 0.01, which is relatively large.  To yield
+% higher resolution and lower error, we switched to a least squares solution:
 
@@
 The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
-%% \begin{equation}\label{eq:clt-boundsx}
-%% \begin{split}
-%% -6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
-%% &\leq 1.36E|X|^3.
-%% \end{split}
-%% \end{equation}
 \begin{align*}
@@
 \begin{proof}
-See Appendix~\ref{app:proof_cdf_bounds}.
+See Supplementary Material, Section~\ref{app:proof_cdf_bounds}.
 \end{proof}
-
-%% {\color{blue} Daniel: Hold on, the following Corollary seems to add $X_{\rm norm}$
-%% to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
-%% didn't the previous section already justify that $\Delta^* \approx \Delta +
-%% X_{\rm norm}$? That is, I don't think we need to add another $X_{\rm norm}$; I
-%% think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
-%% match what we wrote earlier.
-
-%% I think the easiest way to resolve this is to do the following: we assume that
-%% $\Delta^*$ is already $\Delta + X_{\rm noise}$. Then all we need to do is add
-%% $X_{\rm corr}$. So this is the same as what we are doing, except we do not need
-%% to apply Lemma~\ref{lem:cdf_bounds} twice; we just need to do it once because we
-%% are only adding $X_{\rm corr}$, not $X_{\rm norm}$.}
 
@@
 \begin{proof}
-We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
-\begin{equation}
-\begin{split}
-    P(y) & = {\rm Pr}(\Delta^* < y) \\
-     {\rm and} \quad Q(y) & = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
-\end{split}
-\end{equation}
-and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
-where $\sigma_n^2 = 1 - s^2_{\Delta^*}$. This yields the next iteration of $P$
-and $Q$:
-\begin{equation}
-\begin{split}
-    P'(y) & = {\rm Pr}(\Delta^*+X_{\rm nc} < y) \\
-    \quad {\rm and}\quad Q'(y) & = \Phi\left({y-\Delta}\right)
-\end{split}
-\end{equation}
-Now we convolve with the distribution of $X_{\rm corr}$:
-\begin{equation}
-\begin{split}
-    P''(y) & = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) \\
-    \quad {\rm and}\quad Q''(y) & = S\left({y-\Delta}\right)
-\end{split}
-\end{equation}
-Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
-$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
-Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
-that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
-nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
-log} > 0)$ by at most $\epsilon$.
+See Supplementary Material, Section~\ref{app:bounds_preserved}.
 \end{proof}
@@
 
-%% {\color{blue} Daniel: Do we want to include this new analysis? There seem to be
-%% two different ways we show the bound. The first part (above) is nice and I like
-%% it (and it seems to show all we need for Algorithm 1, doesn't it?}
+% Daniel: this is probably less important than other parts of the paper. Hence,
+% I put it in as \begin{comment} ... \end{comment} but we can think about
+% including it back in if it is important enough.
+
+\begin{comment}
 
 \subsection{Bounds on the Stationary Distribution}
-Bounds on the error of an M-H test imply bounds on the stationary
-distributed of the Markov chain under appropriate conditions. Such
-bounds were derived in both ~\citet{cutting_mh_2014}
-and~\citet{icml2014c1_bardenet14}. We include the result from
-~\citet{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the total
-variation distance between two distributions $P$ and $Q$.  Let
-$\mathcal{T}_0$ denote the transition kernel of the exact Markov
-chain, $\mathcal{S}_0$ denote the exact posterior distribution,
-and  $\mathcal{S}_{\epsilon}$ denote the stationary distribution
-of the approximate transition kernel. 
+
+Bounds on the error of an M-H test imply bounds on the stationary distribution
+of the Markov chain under appropriate conditions. Such bounds were derived in
+both ~\citet{cutting_mh_2014} and~\citet{icml2014c1_bardenet14}. We include the
+result from~\citet{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the
+total variation distance between two distributions $P$ and $Q$.  Let
+$\mathcal{T}_0$ denote the transition kernel of the exact Markov chain,
+$\mathcal{S}_0$ denote the exact posterior distribution, and
+$\mathcal{S}_{\epsilon}$ denote the stationary distribution of the approximate
+transition kernel. 
 \begin{lemma}
- If $\mathcal{T}_0$ satisfies the contraction
-  condition $d_v(P\mathcal{T}_0,\mathcal{S}_0) < \eta
-  d_v(P,\mathcal{S}_0)$ for some constant $\eta\in [0,1)$ and all
-    probability distributions $P$, then
+If $\mathcal{T}_0$ satisfies the contraction condition
+$d_v(P\mathcal{T}_0,\mathcal{S}_0) < \eta d_v(P,\mathcal{S}_0)$ for some
+constant $\eta\in [0,1)$ and all probability distributions $P$, then
 \begin{equation}
@@
 \end{lemma}
+
+\end{comment}
 
@@
 From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
-$\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In Algorithm 1, we
+$\phi(x)$ and $\phi^{''}(x)$. In Algorithm 1, we
 correct for the difference in $\sigma$ between $\Delta^*$ and the variance
@@
 We first observe that
-\begin{equation}\label{eq:pdf_difference}
+\[
     P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
-\end{equation}
+\]
 and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
@@
 
+\section{Proof of Corollary~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}
+
+We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
+\begin{equation}
+\begin{split}
+    P(y) & = {\rm Pr}(\Delta^* < y) \\
+     {\rm and} \quad Q(y) & = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
+\end{split}
+\end{equation}
+and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
+where $\sigma_n^2 = 1 - s^2_{\Delta^*}$. This yields the next iteration of $P$
+and $Q$:
+\begin{equation}
+\begin{split}
+    P'(y) & = {\rm Pr}(\Delta^*+X_{\rm nc} < y) \\
+    \quad {\rm and}\quad Q'(y) & = \Phi\left({y-\Delta}\right)
+\end{split}
+\end{equation}
+Now we convolve with the distribution of $X_{\rm corr}$:
+\begin{equation}
+\begin{split}
+    P''(y) & = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) \\
+    \quad {\rm and}\quad Q''(y) & = S\left({y-\Delta}\right)
+\end{split}
+\end{equation}
+Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
+$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
+Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
+that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
+nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
+log} > 0)$ by at most $\epsilon$.
+
 \end{document}

### 2016-10-12 aistats2017.tex
Paper is now about 9.0 pages. Still need to eliminate a few more stuff and to update logistic regression results

@@
 approximate (bounded error) acceptance tests using subsets of the full dataset.
-The tests depend on minibatch statistics, and on the value of an additional
-uniform random variable $u$. The amount of data consumed for each test varies
-significantly from one minibatch to the next, and depends on the current sample,
-the proposed sample, and on the random variable $u$. By
-contrast,~\citet{conf/uai/MaclaurinA14,TallData15} perform exact tests but
-require a lower bound on parameter distribution across its domain.  The amount
-of data reduction depends on the accuracy of this bound, and such bounds are
-only available for relatively simple distributions.
+The amount of data consumed for each test varies significantly from one
+minibatch to the next. By contrast,~\citet{conf/uai/MaclaurinA14,TallData15}
+perform exact tests but require a lower bound on parameter distribution across
+its domain.  The amount of data reduction depends on the accuracy of this bound,
+and such bounds are only available for relatively simple distributions.
 
@@
 
-    \item For continuous probability distributions, reduce the proposal step
-    size and variance (for stochastic proposals) compared to an optimal
-    proposal. The variance of the log acceptance probability scales as the
-    square of proposal step size.
-
     \item Increase the minibatch size when needed for certain minibatches. Log
@@
     the error rate for the test.
+
+    \item For continuous probability distributions, reduce the proposal step
+    size and variance (for stochastic proposals) compared to an optimal
+    proposal. The variance of the log acceptance probability scales as the
+    square of proposal step size. 
 \end{itemize}
 
+It is worth discussing at this point the typical goals of M-H sampling on large
+datasets. By the Bernstein-von Mises Theorem, the posterior distribution for a
+Bayesian inference task has variance that scales inversely with the $N$ data
+samples. Simply sampling from it is one application, but an efficient
+proposal~\citep{OptimalScaling01} has similar variance to the target and will
+diffuse to it extremely slowly. For applications to neural networks or models
+where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
+likely get trapped in one of the modes. A common solution is to anneal the
+sampler, running first at high temperature (scaling log likelihoods by $1/T$)
+to flatten the likelihood landscape. This in turn reduces the variance of
+the log acceptance probability and allows our acceptance test to be applied. Our
+samples can cover the search space densely with small steps rather than taking a
+few sparse steps towards an optimum. In this mode, Metropolis-Hastings can be
+used in similar fashion to Stochastic Gradient Descent. The goal in SGD is to
+make gradual progress to a posterior mode with each step, taking small steps so
+that the cumulative displacement has progressively lower variance.
+
+% Daniel: I merged these two paragraphs into one, because these were too long,
+% and we also never experiment with the step size. Let me know what you think.
+\begin{comment}
 It is worth discussing at this point the typical goals of M-H sampling on large
@@
 regression experiments. 
+\end{comment}
 
@@
     \item We compare performance of our new test and prior approaches on several
-    datasets. We demonstrate the test is several orders of magnitude more
-    efficient than prior work measured as data consumed per test, and that it
-    does not suffer from long-tailed minibatch sizes (up to the dataset size). 
+    datasets. We demonstrate orders of magnitude improvements in efficiency
+    (measured as data consumed per test), and that it does not suffer from
+    long-tailed minibatch sizes.
 \end{itemize}
@@
 
-Consider a Gaussian model where $x_1,\ldots,x_N$ are i.i.d. $\mathcal{N}(\theta,1)$
-with a known variance $\sigma^2=1$, true mean $\theta=0.5$, and a uniform prior
-on $\theta$. The log likelihood ratio is
+Let $x_1,\ldots,x_N$ be i.i.d. $\mathcal{N}(\theta,1)$ with known variance
+$\sigma^2=1$ and (unknown) mean $\theta=0.5$. We use a uniform prior on
+$\theta$. The log likelihood ratio is
 \begin{equation}\label{eq:lemma_ll_ratio}
@@
 distribution has the same variance as the target
-distribution~\citep{OptimalScaling01}, so we use the proposal
-$\mathcal{N}(\theta,1/N)$, which is implemented as
-$q(\theta'\mid\theta)=\phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal
-density function with zero mean and unit variance. This proposal is symmetric
+distribution~\citep{OptimalScaling01}, so we use a proposal based on
+$\mathcal{N}(\theta,1/N)$. It is symmetric
 $q(\theta'\mid\theta)=q(\theta\mid\theta')$, and since we assumed a uniform
-prior on $\theta$, we see that $\psi(u,\theta,\theta')$ reduces to $\log u$. Our
-worst-case scenario is specified in Lemma~\ref{lem:worst_case}.
+prior, $\psi(u,\theta,\theta')=\log u$. Our worst-case scenario is specified in
+Lemma~\ref{lem:worst_case}.
 
@@
 hypothesis test needs to separate nearly-equal values.  It follows that the
-accelerated M-H tests in~\citet{cutting_mh_2014,icml2014c1_bardenet14} require at
-least a constant fraction $\geq c$ in the amount of data consumed per test
-compared to full-dataset tests, so their speed-up is at most $1/c$.
+accelerated M-H tests from prior work require at least a constant fraction $\geq
+c$ in the amount of data consumed per test compared to full-data tests, so their
+speed-up is $\le 1/c$.
 
@@
 been applied to Bayesian estimation on minibatches of data. This simplified
-dynamics uses local proposals and avoids MH tests by using small proposal steps
+dynamics uses local proposals and avoids M-H tests by using small proposal steps
 whose acceptance approaches 1 in the limit. However, the constraint on proposal
@@
 allows momentum-augmented proposals with larger step sizes. However, step sizes
-are still limited by the need to run accurately without MH tests.  By providing
+are still limited by the need to run accurately without M-H tests.  By providing
 an M-H test with similar cost to standard gradient steps, our work opens the
@@
 
-As our discussion in Section~\ref{sec:our_algorithm} showed, as the distribution
-of $\Delta^*$ approaches a Gaussian, our new MH test becomes more accurate.
+From Section~\ref{sec:our_algorithm}, as the distribution of $\Delta^*$
+approaches a Gaussian, our new MH test becomes more accurate.
 Corollary~\ref{cor:bounds_preserved} shows that the bounds from
@@
 
+% Daniel: I'm moving this paragraph over here and putting the rest of it in the
+% supplementary material since our code doesn't do this, so we should not
+% describe this in much detail.
+In fact we can do better than $O(n^{-1/2})$ error for the quantitative CLT
+(showing the error decreases as $O(n^{-1})$) by using a more precise limit
+distribution under an additional assumption. We review the details in
+the Supplementary Material, Section~\ref{app:better_error_bound}, and leave the
+implementation details to future work.
+
 % Daniel: this is probably less important than other parts of the paper. Hence,
 % I put it in as \begin{comment} ... \end{comment} but we can think about
-% including it back in if it is important enough.
+% including it back in if it is important enough. Also, the sub-section that was
+% after that, improved error bounds on skew estimation, has been moved to the
+% supplementary material. I don't see how we can discuss it in the paper without
+% taking up too much space.
 
@@
 \subsection{Bounds on the Stationary Distribution}
-
 Bounds on the error of an M-H test imply bounds on the stationary distribution
@@
 
-\subsection{Improved Error Bounds Based on Skew Estimation}
-In fact we can do better than $O(n^{-1/2})$ error for the quantitative CLT (showing the error decreases as
-$O(n^{-1})$) by using a more precise limit distribution under an additional
-assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$ denote the
-$i^{th}$ absolute moment of $X$. If Cramer's condition holds:
-\begin{equation}\label{eq:cramers_condition}
-    \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1,
-\end{equation}
-then Equation 2.2 in Bentkus et al.'s work on Edgeworth
-expansions~\citep{Bentkus97} provides the following:
-
-\begin{lemma}\label{lem:clt_edgeworth}
-Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
-random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3.
-If $X$ satisfies Cramer's condition, then
-\begin{equation}\label{eq:clt-bounds_edgeworth}
-    \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
-\end{equation}
-where
-\begin{equation}
-    G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x).
-\end{equation}
-\end{lemma}
-
-Lemma~\ref{lem:clt_edgeworth} shows that the average of the $X_i$ has a more
-precise, skewed CDF limit $G_n(x,y)$ where the skew term has weight proportional
-to a certain measure of skew derived from the moments:
-$\frac{\mu_3}{b_2^{3/2}}$. Note that if the $X_i$ are symmetric, the weight of
-the correction term is zero, and the CDF of the average of the $X_i$ converges
-to $\Phi(x)$ at a rate of $O(n^{-1})$.
-
-Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
-$n^{-1/2}$.
-%{\color{blue} (Daniel: how did the equation in the next sentence
-% appear? I verified it and it seems correct, but it looks like it popped out of
-%nowhere.)}
-% JFC: It comes from taking a second derivative. You dont need to explain it. 
-Importantly, since $\phi^{''}(x) = x^2\phi(x) - \phi(x)$ where
-$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
-\begin{equation}\label{eq:GNderivatives}
-    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
-\end{equation}
-From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
-$\phi(x)$ and $\phi^{''}(x)$. In Algorithm 1, we
-correct for the difference in $\sigma$ between $\Delta^*$ and the variance
-needed by $X_{\rm corr}$ using $X_{\rm nc}$. This same method works when we
-wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
-component functions of $G_n(x,y)$ are derivatives of a (unit variance)
-$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
-variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm nc}$ as
-per Algorithm 1 preserving the limit in Equation~(\ref{eq:GNderivatives}).
-
-The deconvolution approach can be used to construct a correction variable
-$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
-additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
-Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
-Equation~(\ref{eq:GNderivatives}), its enough to consider a function $g(x,y')$
-parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
-saved offline. As we have shown above, errors in the ``limit'' function
-propagate directly through as errors in the acceptance test.  To achieve a test
-error of $10^{-6}$ (close to single floating point precision), we need a $y'$
-spacing of $10^{-6}$. It should not be necessary to tabulate values all the way to
-$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
-Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
-resolution is 10,000, this leads to a table with about 1 billion values, which
-can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
-in $y$, it should be possible to achieve similar accuracy with a much smaller
-table. We leave further analysis and experiments with $g(x,y)$ as future work.
-%{\color{blue} Daniel: I will look at the code to understand this paragraph
-%better because right now I am confused.}
-% JFC: the code doesnt implement this yet.
-
 \section{EXPERIMENTS}\label{sec:experiments}
@@
 \subsection{Mixture of Gaussians}\label{ssec:gaussians}
+
+\begin{figure*}[t]
+    \centering
+    \includegraphics[width=0.85\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
+    \caption{
+    The log posterior contours and scatter plots of sampled $\theta$ values
+    using different methods. 
+    }
+    \label{fig:gauss_mix_1}
+\end{figure*}
+\begin{figure*}[t]
+    \centering
+    \includegraphics[width=0.85\linewidth]{minibatch_size_gaussian.png}
+    \caption{
+    Minibatch sizes used in Section~\ref{ssec:gaussians}'s experiment. The axes
+    have the same (log-log scale) range.
+    }
+    \label{fig:gauss_mix_2}
+\end{figure*}
 
@@
 \begin{split}
-    \begin{bmatrix} \theta_1 \\ \theta_2 \end{bmatrix} 
-    & \sim \mathcal{N}\left(\begin{bmatrix}0 \\ 0\end{bmatrix}, 
-    \begin{bmatrix} \sigma_1^2 & 0 \\ 0 & \sigma_2^2\end{bmatrix} \right); \\
+    \theta &\sim \mathcal{N}(0, {\rm diag}(\sigma_1^2,\sigma_2^2)) \\
     x_i & \sim 0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
@@
 
-\begin{figure*}[t]
-    \centering
-    \includegraphics[width=0.9\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
-    \caption{
-    The log posterior contours and scatter plots of sampled $\theta$ values
-    using different methods. 
-    }
-    \label{fig:gauss_mix_1}
-\end{figure*}
-\begin{figure*}[t]
-    \centering
-    \includegraphics[width=0.9\linewidth]{minibatch_size_gaussian.png}
-    \caption{
-    Counts of minibatch sizes that the three methods used in the experiment of
-    Section~\ref{ssec:gaussians}. To make comparisons easier, the axes have the
-    same range (using a log-log scale).
-    }
-    \label{fig:gauss_mix_2}
-\end{figure*}
-
 Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
@@
 respect to the two components of the parameter $\theta$.  The probability of a
-sample falling into one of these bins is the integral of the true posterior
-probability over the area of that bin. Let this probability be denoted $P_i$
-where $i$ is the bin number. A single sample from any of the MH methods should
-therefore be multinomial with distribution $P$, and a set of $n$ (ideally
-independent) samples should follow a ${\rm Multinomial}(P,n)$ distribution.
-Since the ideal distribution is simple, we can use it to measure the quality of
-the sample distributions rather than use general purpose tests like
-KL-divergence or likelihood-ratio.  The latter are problematic with zero counts
-in some bins as we have here.
+sample falling into bin $i$ (denoted as $P_i$) is the integral of the true
+posterior probability over the area of that bin.  A single sample from any of
+the MH methods should therefore be multinomial with distribution $P$, and a set
+of $n$ (ideally independent) samples should follow a ${\rm Multinomial}(P,n)$
+distribution.  Since the ideal distribution is simple, we can use it to measure
+the quality of the sample distributions rather than use general purpose tests
+like KL-divergence or likelihood-ratio.  The latter are problematic with zero
+counts in some bins as we have here.
 
@@
 	\caption{
-    Counts of minibatch sizes that the three methods used in the MNIST logistic
-    regression experiment from Section~\ref{ssec:logistic}. It is analogous to
-    Figure~\ref{fig:gauss_mix_2}.
+    Counts of minibatch sizes in the MNIST logistic regression experiment
+    (analogous to Figure~\ref{fig:gauss_mix_2}).
     }
@@
 
+\section{Improved Error Bounds Based on Skew Estimation}\label{app:better_error_bound}
+
+We demonstrate how we can show $O(n^{-1})$ error for the quantitative CLT by
+using a more precise limit distribution under an additional assumption. Let
+$\mu_i$ denote the $i^{th}$ moment, and $b_i$ denote the $i^{th}$ absolute
+moment of $X$. If Cramer's condition holds:
+\begin{equation}\label{eq:cramers_condition}
+    \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1,
+\end{equation}
+then Equation 2.2 in Bentkus et al.'s work on Edgeworth
+expansions~\citep{Bentkus97} provides the following:
+
+\begin{lemma}\label{lem:clt_edgeworth}
+Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
+random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3.
+If $X$ satisfies Cramer's condition, then
+\begin{equation}\label{eq:clt-bounds_edgeworth}
+    \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
+\end{equation}
+where
+\begin{equation}
+    G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x).
+\end{equation}
+\end{lemma}
+Lemma~\ref{lem:clt_edgeworth} shows that the average of the $X_i$ has a more
+precise, skewed CDF limit $G_n(x,y)$ where the skew term has weight proportional
+to a certain measure of skew derived from the moments:
+$\frac{\mu_3}{b_2^{3/2}}$. Note that if the $X_i$ are symmetric, the weight of
+the correction term is zero, and the CDF of the average of the $X_i$ converges
+to $\Phi(x)$ at a rate of $O(n^{-1})$.
+
+Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
+$n^{-1/2}$.
+%{\color{blue} (Daniel: how did the equation in the next sentence
+% appear? I verified it and it seems correct, but it looks like it popped out of
+%nowhere.)}
+% JFC: It comes from taking a second derivative. You dont need to explain it. 
+Importantly, since $\phi^{''}(x) = x^2\phi(x) - \phi(x)$ where
+$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
+\begin{equation}\label{eq:GNderivatives}
+    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
+\end{equation}
+From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
+$\phi(x)$ and $\phi^{''}(x)$. In Algorithm 1, we
+correct for the difference in $\sigma$ between $\Delta^*$ and the variance
+needed by $X_{\rm corr}$ using $X_{\rm nc}$. This same method works when we
+wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
+component functions of $G_n(x,y)$ are derivatives of a (unit variance)
+$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
+variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm nc}$ as
+per Algorithm 1 preserving the limit in Equation~(\ref{eq:GNderivatives}).
+
+The deconvolution approach can be used to construct a correction variable
+$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
+additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
+Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
+Equation~(\ref{eq:GNderivatives}), its enough to consider a function $g(x,y')$
+parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
+saved offline. As we have shown earlier, errors in the ``limit'' function
+propagate directly through as errors in the acceptance test.  To achieve a test
+error of $10^{-6}$ (close to single floating point precision), we need a $y'$
+spacing of $10^{-6}$. It should not be necessary to tabulate values all the way to
+$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
+Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
+resolution is 10,000, this leads to a table with about 1 billion values, which
+can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
+in $y$, it should be possible to achieve similar accuracy with a much smaller
+table. We leave further analysis and experiments with $g(x,y)$ as future work.
+%{\color{blue} Daniel: I will look at the code to understand this paragraph
+%better because right now I am confused.}
+% JFC: the code doesnt implement this yet.
+
 \end{document}

### 2016-10-12 aistats2017.tex
Removed another space-consuming equation. OK, now it really is about 9.0 pages

@@
 for a logistic random variable $X$.
-%{\color{blue} Daniel: I think the text
-%before Equation~\ref{eq:the_exact_test} should be similar to what we had in our
-%NIPS submission because I find it much clearer.}
 
@@
 to the full batch version.
-
-%% {\color{blue} Daniel: this subsection is improved from earlier. My main concern
-%% now is that $X_{\rm norm}$ is described as being both approximately Gaussian
-%% $\bar{\mathcal{N}}$ and Gaussian $\mathcal{N}$ here. We may want to clear that
-%% up; in our NIPS submission we actually used a new variable $\varepsilon$ for
-%% that and assigned it to be the $\mathcal{N}$ distribution. My current
-%% suggestion is to have $\Delta^*$ be the $\bar{\mathcal{N}}$ approx. normal
-%% distribution (but with mean $\Delta$, of course), and keep $X_{\rm norm}$ as an
-%% exact Gaussian.}
 
@@
 For computation of $C_\sigma$, we assume that its input $Y$ and another variable
-$X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.  In continuous
-form, the optimization problem is to find the $C_\sigma^*$ solving
+$X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
+% Daniel: removing this because we were explaining this in quite some detail and
+% it was taking a lot of space.
+%\[
+%    \argmin_{C\sigma} \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|.
+%\]
+We discretize the convolution by discretizing $X$ and $Y$ into $4N+1$ and $2N+1$
+values respectively. If $i \in \{-2N, \ldots, 2N\}=\mathcal{I}$ and $j \in \{-N,
+\ldots, N\}=\mathcal{J}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
+objective can be written as:
 \[
-    \argmin_{C\sigma} \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|.
-\]
-To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
-and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}=\mathcal{I}$ and
-$j \in \{-N, \ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$,
-and the objective can be re-expressed as:
-\[
-    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \mathcal{I}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
+C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \mathcal{I}}\left|\sum_{j\in\mathcal{J}} \Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
 \]
@@
 \end{equation}
-where we add the standard regularization factor $\lambda$. The solution is
-well-known from the normal equations: $u^* = (M^TM + \lambda I)^{-1}M^Tv$, and
-in practice, it yields an acceptable $L_{\infty}$ norm.
+with regularization $\lambda$. The solution is well-known from the normal
+equations ($u^* = (M^TM + \lambda I)^{-1}M^Tv$) and in practice yields an
+acceptable $L_{\infty}$ norm.
 
@@
 would need to iterate through all the $p(x_i\mid \theta')$ terms, defeating the
-purpose of minibatch MCMC.
-
-Here, we use quantitative forms of the Central Limit Theorem which rely on
-measurable statistics from a single minibatch. Thus a sampler using our approach
-does not need to see data beyond the current minibatch, and can easily be
-applied on large datasets and in online modes where the data is presented as a
-stream.
+purpose of minibatch MCMC. In contrast, we use quantitative forms of the Central
+Limit Theorem which rely on measurable statistics from a single minibatch. Thus
+a sampler using our approach does not need to see data beyond the current
+minibatch.
 

### 2016-10-12 aistats2017.tex
removing newpage before references:
Adding NIPS 2016 submission section because we need it
Added in our supplementary page regarding NIPS. I think this covers everything.

@@
 \begin{table}[t]
-    \caption{Error in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$}
+    \caption{Error ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$}
     \label{tab:xcorr}
@@
 
-\newpage
+%\newpage
 \subsubsection*{References}
@@
 
+% Daniel: don't forget this. The AISTATS website says to put it in the
+% supplementary material, so I put it at the very end on a new page (to make it
+% more noticeable). Obviously for the arXiv version, we don't want this there.
+% Also, note to John, the reviewers for AISTATS **will have access to the
+% previous NIPS reviews**, which is why I am making references to those.
+\newpage
+\section{NIPS 2016 Submission Statement}
+
+We previously submitted a much older draft of this paper to NIPS 2016. The
+current manuscript has been substantially revised since that submission. We have
+made the following changes:
+
+\begin{enumerate}
+    \item At the time of the NIPS submission, we were unaware of several
+    references which described some of what we wrote. For this submission, we
+    now cite the important work of~\citep{TallData15} which had the idea of
+    using subsampling noise for exploration (rather than the $\log u$ variable).
+    We also cite the work of~\citep{Barker65} which uses the Barker function. We
+    welcome information about other potentially missing references.
+
+    \item The previous theoretical results have been entirely scrapped and
+    replaced with different but more relevant results. The NIPS submission
+    listed a page of theoretical results which did not tie into the rest of the
+    paper's contributions and were confusing to the readers. We now have
+    theoretical results that are much clearer and also \emph{directly} show the
+    expected performance benefits of our acceptance test. Note also that
+    Lemma~\ref{lem:worst_case} is an entirely new addition to this version.
+
+    \item Several reviewers mentioned that our deconvolution approach to
+    determine the correction distribution was not unique and thus ill-defined.
+    Consequently, we have added an entirely new section of the paper
+    (Section~\ref{sec:correction}) explaining how we derived that distribution.
+
+    \item Reviewers pointed out one of our mistakes when we said our minibatch
+    sizes were fixed. We have changed this in the current manuscript to mean
+    that the per-iteration minibatch sizes for our test have a distribution with
+    \emph{shorter tails} than those of prior work, which can be observed by
+    plotting a histogram of minibatch sizes.
+
+    \item Our experiments now include new comparisons with a baseline
+    from~\citep{icml2014c1_bardenet14} (in addition to~\citep{cutting_mh_2014},
+    which we had earlier). In addition, each algorithm now runs MCMC sampling on
+    the \emph{same} distribution.  Previously, we ran our distribution at a
+    higher temperature but kept the algorithm from~\citep{cutting_mh_2014}
+    running on the distribution at temperature $T=1$. In this set of
+    experiments, we tuned hyperparameters of the other algorithms to make the
+    comparison fairer. Finally, in the Gaussian mixture model scenario
+    (Section~\ref{ssec:gaussians}), we provide more details on how ``accurate''
+    the samples are, rather than solely relying on the visualization of
+    Figure~\ref{fig:gauss_mix_1}.
+
+    \item We improved Algorithm~\ref{alg:our_algorithm} so that we show
+    explicitly when we compute the moments, and why we now only need one
+    distribution $C_\sigma$ for $\sigma=1$ due to the extra $X_{\rm nc}$
+    variable.
+
+    \item Finally, we made minor revisions addressing: differences between the
+    Barker function vs. the original MH test, and sampling with vs. without
+    replacement.
+\end{enumerate}
+
+We are confident that the current manuscript is of far better quality than the
+NIPS submission, and we appreciate the efforts of the NIPS reviewers to give us
+ideas for improvement.
+
 \end{document}

### 2016-10-12 aistats2017.tex
Whew, major update. Now to send email to John/Xinlei.

@@
 network. In general, since a new $\theta'$ value is obtained each iteration, one
-would need to iterate through all the $p(x_i\mid \theta')$ terms, defeating the
-purpose of minibatch MCMC. In contrast, we use quantitative forms of the Central
-Limit Theorem which rely on measurable statistics from a single minibatch. Thus
-a sampler using our approach does not need to see data beyond the current
+would need to iterate through all the $p(x_i\mid \theta')$ terms\footnote{The
+sample code provided by the authors of~\citet{icml2014c1_bardenet14} in fact
+computes $C_{\theta,\theta'}$ explicitly for each sample generated, i.e. it
+traverses the entire data, thus providing no performance advantage over the
+complete test.}. In contrast, we use quantitative forms of the Central Limit
+Theorem which rely on measurable statistics from a single minibatch. Thus a
+sampler using our approach does not need to see data beyond the current
 minibatch.
@@
 \begin{proof}
-The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
-\begin{align*}
--6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
-&\leq 1.36E|X|^3.
-\end{align*}
-This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
-and combining with $x>0$, we obtain the weaker but unqualified bound
-in Equation~(\ref{eq:clt-bounds}).
+See Supplementary Material, Section~\ref{app:quant_clt}.
 \end{proof}
@@
 \langle \theta_1,\theta_2 \rangle$, and the parameter/data generation process is
-%% \begin{equation}\label{eq:data_generation}
-%%     (\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
-%%     0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
-%% \end{equation}
 \begin{equation}\label{eq:data_generation}
@@
 We run MCMC with our MH test and benchmark with~\citet{cutting_mh_2014}
-and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.
-For~\citet{cutting_mh_2014}, we increment minibatches by 100 and
-for~\citet{icml2014c1_bardenet14}, we increase the minibatch size geometrically
-with a ratio of $\gamma = 1.5$. These methods have different approaches to
-repeated testing: the former uses an aggressive testing approach, explicitly
-modeling correlation between tests and allowing many more tests for a given
-$\delta$, while the latter uses discounting, allocating a fraction of the test
-error $\delta$ to each minibatch test.
-
-The tolerance for making a decision in~\citet{cutting_mh_2014} is
-$\epsilon=0.005$, and the error bound control parameters
-in~\citet{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.  All methods
-(including ours) use the same random walk proposer with covariance $\Sigma={\rm
-diag}(0.15, 0.15)$.  Since the proposer generates a random walk, all shaping of
-distribution of the samples is due to the MH test. All methods are run 5000
-times to collect 5000 samples.
+and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.  For
+the former, we increment minibatches by 100 and for the latter, we increase it
+geometrically with a ratio of $\gamma = 1.5$.  The tolerance for making a
+decision in~\citet{cutting_mh_2014} is $\epsilon=0.005$, and the error bound
+control parameters in~\citet{icml2014c1_bardenet14} are $p = 2$ and $\delta =
+0.01$.  All methods (including ours) use the same random walk proposer with
+covariance $\Sigma={\rm diag}(0.15, 0.15)$.  Since the proposer generates a
+random walk, all shaping of distribution of the samples is due to the MH test.
+All methods are run 5000 times to collect 5000 samples.
 
@@
 
-We discretize the posterior coordinates into bins (non-overlapping ranges) with
-respect to the two components of the parameter $\theta$.  The probability of a
-sample falling into bin $i$ (denoted as $P_i$) is the integral of the true
-posterior probability over the area of that bin.  A single sample from any of
-the MH methods should therefore be multinomial with distribution $P$, and a set
-of $n$ (ideally independent) samples should follow a ${\rm Multinomial}(P,n)$
-distribution.  Since the ideal distribution is simple, we can use it to measure
-the quality of the sample distributions rather than use general purpose tests
-like KL-divergence or likelihood-ratio.  The latter are problematic with zero
+We discretize the posterior coordinates into bins with respect to the two
+components of $\theta$.  The probability $P_i$ of a sample falling into bin $i$
+is the integral of the true posterior probability over the area of that bin.  A
+single sample from any of the MH methods should therefore be multinomial with
+distribution $P$, and a set of $n$ (ideally independent) samples should be ${\rm
+Multinomial}(P,n)$.  The ideal distribution is simple, so we can use it to
+measure the quality of the sample distributions rather than use general purpose
+tests like KL-divergence or likelihood-ratio, which can be problematic with zero
 counts in some bins as we have here.
 
-In practice, it is difficult to compute multinomial probabilities for large $n$,
-in that case the per-bin distributions are well-approximated by Poisson
+For large $n$, the per-bin distributions are well-approximated by Poisson
 distributions with parameter $\lambda_i=P_i n$. Given a set of samples
-$\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
-$\theta_i$ that fall in bin $j$ out of $N_{\rm bins}$ total. Then we have
+$\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual
+samples $\theta_i$ that fall in bin $j$ out of $N_{\rm bins}$ total. We have
 \begin{equation}\label{eq:log_prob_poissons}
@@
 While the Poisson results provide general guidance on the relative accuracy of
-the models, it is difficult to interpret the scores, so we perform significance
-tests to show the difference between the MCMC-sampled distributions and the
-ground-truth.  We perform an approximate test using the Chi-Squared distribution
-as the test statistic, which we also show in Table~\ref{tab:poissons}.
-
-Both the Poisson values and the Chi-Squared test statistic result in similar
-conclusions: our method is slightly superior to~\citet{cutting_mh_2014}, but
-slightly worse than~\citet{icml2014c1_bardenet14}. We now show, however, that
-minibatch size results suggest that our method dominates in terms of speed and
-efficiency.
-
-Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used
-by the three methods in each iteration. Our method consumes significantly less
-data; most minibatch sizes are smaller than 1000, and the average size is 210.
-The other methods occasionally need to consume a large proportion of the
-entire data set of one million elements; the average minibatch sizes are 15562
-and 16857 for Korattikara 2014 and Bardenet 2014, respectively. The average
-minibatch sizes accurately predict the running times of these methods since
-all methods have a running time proportional to the total data consumed, with
-one caveat: the method of~\citet{icml2014c1_bardenet14} requires a global bound
-on the term $C_{\theta,\theta'}$ (see Equation~(\ref{eq:bad_bound})) over the
-entire the data set at every iteration in order to calculate the error bound.
-This bound must be either computed analytically using properties of the specific
-dataset, or it must be checked at each sample if a black-box test is required.
-We found that the sample code provided by the authors
-of~\citet{icml2014c1_bardenet14} in fact computed this bound explicitly for each
-sample generated, i.e. the implementation traversed the entire dataset for each
-sample, providing no performance advantage over the complete test. For this
-reason, we omitted the time to compute $C_{\theta,\theta'}$ from our
-measurements of the running time of the sample implementation
-of~\citet{icml2014c1_bardenet14}, and consider only the time to process the
-minibatches.
-
-%% In Figure ~\ref{fig:gauss_bardenet}, the value
-%% of the first part and second part of equation (9)
-%% in~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual
-%% minibatch data used. The second part which contains  the $C_{\theta,\theta'}$
-%% term is larger than the first part, which means that the second part is
-%% non-negligible and has to be calculated at every iteration. This is one of the
-%% very time-consuming parts in this method.  {\color{blue} Daniel: I am not sure
-%% if we should include this plot. Let me think about it and if we decide to
-%% include it I will help rewrite the explanation.}
-
-%% JFC - it would still be interesting to see this comparison, however:
-%% (i) its important to include equation (9) and explain what the terms are.
-%% (ii) the results dont look right. I wouldnt expect the first term to be
-%% so small compared to the second and I'm fairly certain they should cross
-%% in the illustrated range. It would be better to include both curves on
-%% the same plot.
+the models, it is difficult to interpret the scores. Thus, we perform
+significance tests to show the difference between the MCMC-sampled distributions
+and the ground-truth using the Chi-Squared distribution as the test statistic,
+which we also show in Table~\ref{tab:poissons}. Both results imply that our
+method is slightly superior to~\citet{cutting_mh_2014}, but slightly worse
+than~\citet{icml2014c1_bardenet14}.
+
+Figure~\ref{fig:gauss_mix_2}, however, suggests that our method dominates in
+terms of speed and efficiency. It shows histograms of the (final) minibatch
+sizes used each iteration by the three methods. Our method consumes
+significantly less data; most minibatch sizes are smaller than 1000, and the
+average size is 210.  The other methods occasionally need to consume nearly all
+data points, and average minibatch sizes are 15562 and 16857. The average
+minibatch sizes roughly predict the running times of these methods since all
+have a running time proportional to the total data consumed, with the exception
+of~\citet{icml2014c1_bardenet14}, which requires a pass over the data to compute
+$C_{\theta,\theta'}$ (see Equation~(\ref{eq:bad_bound})). Using minibatch sizes
+as a proxy for time also avoids discrepancies due to code optimization.
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{logistic_performance.png}
+	%\includegraphics[width=0.9\linewidth]{logistic_performance.png}
+	\includegraphics[width=0.85\linewidth]{logistic_regression_ll_acc_results.png}
 	\caption{
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{minibatch_size_logistic.png}
+	%\includegraphics[width=0.9\linewidth]{minibatch_size_logistic.png}
+	\includegraphics[width=0.85\linewidth]{logistic_regression_histograms_results.png}
 	\caption{
@@
 
-We next use logistic regression for the binary classification of 1s versus 7s in
-the MNIST dataset~\citep{lecun-mnisthandwrittendigit-2010}. The data has 12007
-training and 1000 testing elements (we used a random subset of the test data).
-The pixel values are scaled to be in $(0,1)$.  We impose a flat prior on
-$\theta$. For the proposer, we again use a random walk, this time with
-covariance matrix $0.05I$ for the $784\times 784$ identity matrix $I$. The
-posterior temperature is set at $T=1000$.  We run our MH test and compare with
-Korattikara 2014 and Bardenet 2014; minibatch sizes and other parameters are set
-to be the same value as with the Gaussian Mixtures experiment from
-Section~\ref{ssec:gaussians}.
-
-Figure~\ref{fig:logistic_performance} shows the training log likelihood and the
-testing prediction accuracy, each as a function of the cumulative training data
-points processed.\footnote{Note that the curves do not span the same length over
-the x-axis, because our test consumes fewer samples throughout the MCMC
-procedure, so the corresponding curve will ``end'' before the other two.} To
-generate the curves, for each of the 5000 sampled vectors $\theta_t$,
-$t\in\{1,\ldots,5000\}$, we use $\theta_t$ as the parameter for logistic
-regression.  We see that our minibatch MH test is more efficient compared to the
-other two methods, because it achieves similar or superior performance while
-consuming fewer data points.
+We next use logistic regression for the binary classification of 1s versus 7s on
+a subset of the MNIST8M dataset, which is a larger version of
+MNIST~\citep{lecun-mnisthandwrittendigit-2010} and well-suited to our desire to
+run MCMC on large data. We randomly subsampled 450k training and 192k testing
+points. The pixels are scaled in $[0,1]$. We impose a flat prior on $\theta$ and
+again use a random walk proposer, this time with covariance matrix $0.05I$ for
+the $784\times 784$ identity matrix $I$. The posterior temperature is set at
+$T=1000$.  We run our MH test for 3000 samples and again compare with the two
+baseline algorithms with starting minibatch size 100. To
+make~\citet{cutting_mh_2014} run faster, we use a larger $\epsilon=0.05$ value.
+For~\citet{icml2014c1_bardenet14}, we tuned $\gamma \in \{1.1,1.25,1.5,1.75,2\}$
+before concluding that $\gamma=1.5$ was reasonable. In addition, we originally
+used their suggested analytic approximation to $C_{\theta,\theta'}$, but found
+that the resulting values were too high, meaning that their method needed to
+consume the entire dataset each iteration. Thus, we naively compute
+$C_{\theta,\theta'}$.
+
+Figure~\ref{fig:logistic_performance} shows the testing log likelihood and
+prediction accuracy, each as a function of the cumulative training data points
+processed.\footnote{Note that the curves do not span the same length over the
+x-axis, because our test consumes fewer samples throughout the MCMC procedure,
+so the corresponding curve will ``end'' before the other two.} To generate the
+curves, for each of the 3000 sampled vectors $\theta_t$,
+$t\in\{1,\ldots,3000\}$, we use $\theta_t$ as the parameter for logistic
+regression.  Our minibatch MH test is more efficient in terms of
+accuracy, achieving convergence using roughly half as many elements
+compared to~\citet{cutting_mh_2014} and \textbf{TODO TODO TODO} compared
+to~\citet{icml2014c1_bardenet14}, though its log likelihood results lag slightly
+behind the former algorithm during the very early stages.
 
@@
 Figure~\ref{fig:gauss_mix_2}, shows the histogram of minibatch sizes for all
-three methods on a log-log scale. With an initial minibatch size of 100,
-Korattikara 2014 and Bardenet 2014 achieve average minibatch size of 2026 and
-7612 respectively over the MNIST classification task, while our method achieves
-that of 163, showing significantly better performance than the benchmark test
-methods.
+three methods on a log-log scale. With an initial size of 100, our method
+achieves an average minibatch size of 393, far smaller than the averages of the
+other two methods, which is due to their longer-tailed distributions.
 
@@
 
+\section{Proof of Lemma~\ref{lem:quant_clt}}\label{app:quant_clt}
+
+The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
+\begin{align*}
+-6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
+&\leq 1.36E|X|^3.
+\end{align*}
+This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
+and combining with $x>0$, we obtain the weaker but unqualified bound
+in Equation~(\ref{eq:clt-bounds}).
+
 \section{Proof of Lemma~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}
@@
 then Equation 2.2 in Bentkus et al.'s work on Edgeworth
-expansions~\citep{Bentkus97} provides the following:
+expansions~\citep{Bentkus97} provides:
 
@@
 If $X$ satisfies Cramer's condition, then
-\begin{equation}\label{eq:clt-bounds_edgeworth}
+%\begin{equation}\label{eq:clt-bounds_edgeworth}
+\[
     \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
-\end{equation}
+\]
+%\end{equation}
 where
@@
 to a certain measure of skew derived from the moments:
-$\frac{\mu_3}{b_2^{3/2}}$. Note that if the $X_i$ are symmetric, the weight of
+$\mu_3/b_2^{3/2}$. Note that if the $X_i$ are symmetric, the weight of
 the correction term is zero, and the CDF of the average of the $X_i$ converges
@@
 % previous NIPS reviews**, which is why I am making references to those.
-\newpage
+\clearpage
 \section{NIPS 2016 Submission Statement}

### 2016-10-12 aistats2017.tex
Shortened paper a bit, main part is now 8.7 pages.

@@
 unique stationary distribution $\pi(\theta) = p(\theta)$. For Bayesian
-inference, we would like to sample from a parameter distribution for $\theta$
-based on some observed data, so the target distribution is $p(\theta \mid x_1,
-\ldots, x_N)$. The acceptance probability is now:
+inference, the target distribution is $p(\theta \mid x_1, \ldots, x_N)$. The
+acceptance probability is now:
 \begin{equation}\label{eq:acceptance_probability}
@@
 \end{equation}
-where $p_0(\theta)$ is a prior, and $p(x_i \mid \theta)$ are the probabilities
-of the observations. Computing samples this way requires the use of all $N$
+where $p_0(\theta)$ is the prior. Computing samples this way requires the use of all $N$
 training data points, but this is very expensive for large datasets. To address
@@
 addition, both methods suffer the drawback of resolving small log likelihood
-ratio differences between the minibatch and full batch versions.  In the worst
-case, all $N$ data points may be needed, as we show in
-Section~\ref{ssec:gaussian_example}.
+ratio differences between the minibatch and full batch versions.  We discuss
+a worst-case scenario in Section~\ref{ssec:gaussian_example}.
 
@@
 
-The problem is that these methods use tail bounds to separate $\Delta$ away
-from zero, but for certain input/random $u$ combinations, $\Delta$ can be
-arbitrarily close to zero. We will avoid this by using the {\em approximately
-normal} variation in $\Delta^*$ to {\em replace} the variation due to $u$.
+The issue is the use of tail bounds to separate $\Delta$ from zero; for certain
+input/random $u$ combinations, $\Delta$ can be arbitrarily close to zero. We
+avoid this by using the {\em approximately normal} variation in $\Delta^*$ to
+{\em replace} the variation due to $u$.
 
@@
 There is a separate line of MCMC work drawing principles from statistical
-physics. By viewing random variables as particles in a system, one can apply
+physics. One can apply
 Hamiltonian Monte Carlo (HMC)~\citep{mcmc_hamiltonian_2010} methods which
@@
 
-\section{COMPUTING THE CORRECTION DISTRIBUTION}\label{sec:correction}
-
-Our proposed test in Equation~(\ref{eq:criteria}) requires knowing the
-distribution of the correction variable $X_{\rm corr}$ such that $X_{\rm norm} +
+% Daniel: Saves space by using one line instead of two. Lame, I know.
+%\section{COMPUTING THE CORRECTION DISTRIBUTION}\label{sec:correction}
+\section{CORRECTION DISTRIBUTION}\label{sec:correction}
+
+Our test in Equation~(\ref{eq:criteria}) requires knowing the
+distribution of $X_{\rm corr}$ such that $X_{\rm norm} +
 X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \sim \mathcal{N}(0,\sigma^2)$
 and $X_{\rm log}$ has a standard logistic CDF, $(1+\exp(-X))^{-1}$. In
-Section~\ref{sec:analysis}, we show that the accuracy of the test depends on the
+Section~\ref{sec:analysis}, we show that the test accuracy depends on the
 absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and $X_{\rm
@@
 $C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
-distribution. Our goal, based on Equation~(\ref{eq:deconvolution}), is to solve
-the following optimization problem:
+distribution. Our goal is to solve:
 \begin{equation}\label{eq:overall_corr_problem}
@@
 network. In general, since a new $\theta'$ value is obtained each iteration, one
-would need to iterate through all the $p(x_i\mid \theta')$ terms\footnote{The
-sample code provided by the authors of~\citet{icml2014c1_bardenet14} in fact
-computes $C_{\theta,\theta'}$ explicitly for each sample generated, i.e. it
-traverses the entire data, thus providing no performance advantage over the
-complete test.}. In contrast, we use quantitative forms of the Central Limit
-Theorem which rely on measurable statistics from a single minibatch. Thus a
-sampler using our approach does not need to see data beyond the current
-minibatch.
+would need to use all the $p(x_i\mid \theta')$ terms\footnote{The sample code
+provided by the authors of~\citet{icml2014c1_bardenet14} in fact computes
+$C_{\theta,\theta'}$ explicitly for each sample generated, i.e. it traverses the
+entire data, thus providing no performance advantage over the complete test.}.
+In contrast, we use quantitative forms of the Central Limit Theorem which rely
+on measurable statistics from a single minibatch. Thus a sampler using our
+approach does not need to see data beyond the current minibatch.
 
@@
 \begin{lemma}\label{lem:cdf_bounds}
-Let $P(x)$ and $Q(x)$ be two cumulative distributions satisfying
+Let $P(x)$ and $Q(x)$ be two CDFs satisfying
 $\sup_x|P(x)-Q(x)|\leq \epsilon$ with $x$ in some real range. Let $R(y)$ be the
@@
 
-We start with a simple Gaussian mixture model, borrowing an experiment
-from~\citet{langevin_2011}.  The parameter is 2-D, $\theta =
-\langle \theta_1,\theta_2 \rangle$, and the parameter/data generation process is
+We borrow a Gaussian mixture experiment from~\citet{langevin_2011}.  The
+parameter is 2-D, $\theta = \langle \theta_1,\theta_2 \rangle$, and the
+generation process is
 \begin{equation}\label{eq:data_generation}
@@
 
-For large $n$, the per-bin distributions are well-approximated by Poisson
-distributions with parameter $\lambda_i=P_i n$. Given a set of samples
-$\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual
-samples $\theta_i$ that fall in bin $j$ out of $N_{\rm bins}$ total. We have
+For large $n$, the per-bin distributions are approximated by Poissons with
+parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
+$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
+out of $N_{\rm bins}$ total. We have
 \begin{equation}\label{eq:log_prob_poissons}
@@
 \end{equation}
-The results for each model are shown in Table~\ref{tab:poissons}.
-
-While the Poisson results provide general guidance on the relative accuracy of
-the models, it is difficult to interpret the scores. Thus, we perform
-significance tests to show the difference between the MCMC-sampled distributions
-and the ground-truth using the Chi-Squared distribution as the test statistic,
-which we also show in Table~\ref{tab:poissons}. Both results imply that our
-method is slightly superior to~\citet{cutting_mh_2014}, but slightly worse
+The results for each model are shown in Table~\ref{tab:poissons}. It is
+difficult, however, to interpret the scores. Thus, we perform significance tests
+to show the difference between the MCMC-sampled distributions and the
+ground-truth using the Chi-Squared distribution as the test statistic, which we
+also show in Table~\ref{tab:poissons}. Both results imply that our method is
+slightly superior to~\citet{cutting_mh_2014}, but slightly worse
 than~\citet{icml2014c1_bardenet14}.
@@
 terms of speed and efficiency. It shows histograms of the (final) minibatch
-sizes used each iteration by the three methods. Our method consumes
-significantly less data; most minibatch sizes are smaller than 1000, and the
+sizes used each iteration. Our method consumes
+significantly less data; most sizes are smaller than 1000, and the
 average size is 210.  The other methods occasionally need to consume nearly all

### 2016-10-12 aistats2017.tex
Shortened paper more, abotu 8.4 pages. We had repetitive info in sections 1/2

@@
 
-    \item Increase the minibatch size when needed for certain minibatches. Log
-    acceptance variance scales as $1/k$ vs the minibatch size $k$. Our test is
-    adaptive like earlier works, but unlike them, the distribution of minibatch
-    size is Gaussian, not long-tailed.  Increased minibatch size also reduces
-    the error rate for the test.
-
-    \item For continuous probability distributions, reduce the proposal step
-    size and variance (for stochastic proposals) compared to an optimal
-    proposal. The variance of the log acceptance probability scales as the
-    square of proposal step size. 
+    \item Increase the minibatch size when needed. Log acceptance variance
+    scales as $1/k$ vs the minibatch size $k$. Our test is adaptive like earlier
+    works, but unlike them, the distribution of minibatch size is Gaussian, not
+    long-tailed.  Increased minibatch size also reduces the error rate for the
+    test.
+
+    \item For continuous distributions, reduce the proposal step size and
+    variance compared to an optimal proposal. The variance of the log acceptance
+    probability scales as the square of proposal step size. 
 \end{itemize}
@@
 datasets. By the Bernstein-von Mises Theorem, the posterior distribution for a
-Bayesian inference task has variance that scales inversely with the $N$ data
-samples. Simply sampling from it is one application, but an efficient
+Bayesian inference task has variance that scales inversely with $N$. Simply
+sampling from it is one application, but an efficient
 proposal~\citep{OptimalScaling01} has similar variance to the target and will
@@
 likely get trapped in one of the modes. A common solution is to anneal the
-sampler, running first at high temperature (scaling log likelihoods by $1/T$)
-to flatten the likelihood landscape. This in turn reduces the variance of
-the log acceptance probability and allows our acceptance test to be applied. Our
-samples can cover the search space densely with small steps rather than taking a
-few sparse steps towards an optimum. In this mode, Metropolis-Hastings can be
-used in similar fashion to Stochastic Gradient Descent. The goal in SGD is to
-make gradual progress to a posterior mode with each step, taking small steps so
-that the cumulative displacement has progressively lower variance.
-
-% Daniel: I merged these two paragraphs into one, because these were too long,
-% and we also never experiment with the step size. Let me know what you think.
-\begin{comment}
-It is worth discussing at this point the typical goals of M-H sampling on large
-datasets.  By the Bernstein-von Mises Theorem, the posterior distribution for a
-Bayesian inference task is asymptotically normal, and has variance that scales
-inversely with the $N$ data samples. Simply sampling from this distribution is
-one application, but an efficient proposal~\citep{OptimalScaling01} has similar
-variance to the target distribution and will diffuse to it extremely slowly from
-an initialization value which is (likely to be) many standard deviations away.
-If there are any other strong modes, it is very likely for the sampler to find
-one of them and become trapped in it when run at the normal distribution
-temperature ($T=1$). A common solution is to anneal the sampler, running first
-at high temperature (scaling log likelihoods by $1/T$) which flattens the
-likelihood landscape.  This in turn reduces the variance of the log acceptance
-probability and allows our acceptance test to be applied.
-
-A second question concerns step size. Once we have fixed temperature, our
-variance constraint implies that we have to trade-off proposal step size $s$ and
-batch size $b$ ($b \propto p^2$), i.e. we can make many small steps, or one
-large step, with a given batch of data. One of the primary drivers of this work
-is our belief in the value of small steps. For applications to neural networks
-or other models where the posterior is multimodal, posterior inference is
-arguably a search process. Covering the search space densely with small steps is
-much more valuable than few sparse steps toward the nearest optimum. In this
+sampler, running first at high temperatures to flatten the likelihood landscape.
+This in turn reduces the variance of the log acceptance probability and allows
+our test to be applied. Our samples can cover the search space densely with
+small steps rather than taking a few sparse steps towards an optimum. In this
 mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
@@
 each step, taking small steps so that the cumulative displacement has
-progressively lower variance. We demonstrate this behavior in our logistic
-regression experiments. 
-\end{comment}
+progressively lower variance.
+
+% Daniel: I merged these two paragraphs into one, because these were too long,
+% and we also never experiment with the step size. Let me know what you think.
+
+%% It is worth discussing at this point the typical goals of M-H sampling on large
+%% datasets.  By the Bernstein-von Mises Theorem, the posterior distribution for a
+%% Bayesian inference task is asymptotically normal, and has variance that scales
+%% inversely with the $N$ data samples. Simply sampling from this distribution is
+%% one application, but an efficient proposal~\citep{OptimalScaling01} has similar
+%% variance to the target distribution and will diffuse to it extremely slowly from
+%% an initialization value which is (likely to be) many standard deviations away.
+%% If there are any other strong modes, it is very likely for the sampler to find
+%% one of them and become trapped in it when run at the normal distribution
+%% temperature ($T=1$). A common solution is to anneal the sampler, running first
+%% at high temperature (scaling log likelihoods by $1/T$) which flattens the
+%% likelihood landscape.  This in turn reduces the variance of the log acceptance
+%% probability and allows our acceptance test to be applied.
+%% 
+%% A second question concerns step size. Once we have fixed temperature, our
+%% variance constraint implies that we have to trade-off proposal step size $s$ and
+%% batch size $b$ ($b \propto p^2$), i.e. we can make many small steps, or one
+%% large step, with a given batch of data. One of the primary drivers of this work
+%% is our belief in the value of small steps. For applications to neural networks
+%% or other models where the posterior is multimodal, posterior inference is
+%% arguably a search process. Covering the search space densely with small steps is
+%% much more valuable than few sparse steps toward the nearest optimum. In this
+%% mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
+%% Descent. The goal in SGD is to make gradual progress to a posterior mode with
+%% each step, taking small steps so that the cumulative displacement has
+%% progressively lower variance. We demonstrate this behavior in our logistic
+%% regression experiments.
 
@@
 \end{equation}
-where $p_0(\theta)$ is the prior. Computing samples this way requires the use of all $N$
-training data points, but this is very expensive for large datasets. To address
-this challenge,~\citet{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
-Metropolis-Hasting tests using sequential hypothesis testing. During each
-iteration, they start with a small minibatch of data and test whether the sample $\theta'$ 
-should be accepted based on an approximate version of
-the test $u < \alpha(\theta,\theta')$. If the approximate test cannot make a
-decision with sufficient confidence, then the minibatch size is increased and
+where $p_0(\theta)$ is the prior. Computing samples this way requires all $N$
+data points, but this is very expensive for large datasets. To address this
+challenge,~\citet{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
+Metropolis-Hasting tests using sequential hypothesis testing. Each iteration,
+they start with a small minibatch and test whether $\theta'$ should be accepted
+based on approximating $u<\alpha(\theta,\theta')$. If the approximate test
+cannot decide with sufficient confidence, the minibatch size is increased and
 the test repeats. This process continues until a decision. The bounds depend on
@@
 knowing $p(x_i \mid \theta)$ and $p(x_i \mid \theta')$ for all $N$ samples. In
-addition, both methods suffer the drawback of resolving small log likelihood
-ratio differences between the minibatch and full batch versions.  We discuss
-a worst-case scenario in Section~\ref{ssec:gaussian_example}.
+addition, while both methods show useful reductions in the number of samples
+required, they suffer the drawback of resolving small log likelihood ratio
+differences between the minibatch and full batch versions. We discuss a
+worst-case scenario in Section~\ref{ssec:gaussian_example}.
 
@@
 Since minibatches contains randomly selected samples $x_i$, the values
-$\Lambda_i$ are i.i.d. random
-variables\footnote{The analysis assumes sampling with replacement
-although implementations on typical large datasets will approximate
-this by sampling without replacement.}.
-By the Central Limit Theorem, we expect $\Lambda^*(\theta,\theta')$ to
-be approximately Gaussian. The acceptance test then becomes a
-statistical test of the hypothesis that
+$\Lambda_i$ are i.i.d. random variables\footnote{The analysis assumes sampling
+with replacement although implementations on typical large datasets will
+approximate this by sampling without replacement.}.  By the Central Limit
+Theorem, we expect $\Lambda^*(\theta,\theta')$ to be approximately Gaussian. The
+acceptance test then becomes a statistical test of the hypothesis that
 $\Lambda(\theta,\theta')>\psi(u,\theta,\theta')$ by establishing that
 $\Lambda^*(\theta,\theta')$ is substantially larger than
-$\psi(u,\theta,\theta')$.  In~\citet{cutting_mh_2014} an asymptotic
-central limit argument was used to derive this gap, while
-in~\citet{icml2014c1_bardenet14} a concentration bound was used. In
-both cases, the resulting tests were shown to give useful reductions
-in number of samples required over using the full dataset, but there
-were no worst-case bounds, which we explore in the following section.
+$\psi(u,\theta,\theta')$.
 
@@
 hypothesis test needs to separate nearly-equal values.  It follows that the
-accelerated M-H tests from prior work require at least a constant fraction $\geq
-c$ in the amount of data consumed per test compared to full-data tests, so their
-speed-up is $\le 1/c$.
-
-The issue is the use of tail bounds to separate $\Delta$ from zero; for certain
-input/random $u$ combinations, $\Delta$ can be arbitrarily close to zero. We
-avoid this by using the {\em approximately normal} variation in $\Delta^*$ to
-{\em replace} the variation due to $u$.
+accelerated tests from prior work require at least a constant fraction $\geq c$
+in the amount of data consumed per test compared to full-data tests, so their
+speed-up is $\le 1/c$. The issue is the use of tail bounds to separate $\Delta$
+from zero; for certain input/random $u$ combinations, $\Delta$ can be
+arbitrarily close to zero. We avoid this by using the {\em approximately normal}
+variation in $\Delta^*$ to {\em replace} the variation due to $u$.
 
@@
 that its smoothness allows it to naturally tolerate substantial variance in its
-input argument. This in turn will lead to a much more efficient test on subsets
-of data.
+input. This in turn will lead to a much more efficient test on subsets of data.
 
@@
 two variables are added (i.e., their distributions convolve) to form $X_{\rm
-log}$.  This decomposition requires an appropriate $X_{\rm corr}$ distribution.
-We show in Section~\ref{sec:correction} that it is possible to use deconvolution
-to derive a highly accurate representation of $C_{\sigma}(X)$.  Using $X_{\rm
-corr}$ samples from $C_{\sigma}(X)$, the acceptance test is now
+log}$.  This decomposition requires an appropriate $C_\sigma$, which we derive
+in Section~\ref{sec:correction}. Using $X_{\rm corr}$ samples from
+$C_{\sigma}(X)$, the acceptance test is now
 \begin{equation}\label{eq:criteria}
@@
 
-Our test in Equation~(\ref{eq:criteria}) requires knowing the
-distribution of $X_{\rm corr}$ such that $X_{\rm norm} +
-X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \sim \mathcal{N}(0,\sigma^2)$
-and $X_{\rm log}$ has a standard logistic CDF, $(1+\exp(-X))^{-1}$. In
-Section~\ref{sec:analysis}, we show that the test accuracy depends on the
-absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and $X_{\rm
-log}$. Consequently, we need to minimize this in our construction of $X_{\rm
-corr}$. More formally, let $\Phi_{s_X} = \Phi(X/s_X)$ where $\Phi$ is the
-standard normal CDF\footnote{Hence, $\Phi_{s_X}$ is the CDF of a zero-mean
+Our test in Equation~(\ref{eq:criteria}) requires knowing the distribution of
+$X_{\rm corr}$. In Section~\ref{sec:analysis}, we show that the test accuracy
+depends on the absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$
+and $X_{\rm log}$. Consequently, we need to minimize this in our construction of
+$X_{\rm corr}$. More formally, let $\Phi_{s_X} = \Phi(X/s_X)$ where $\Phi$ is
+the standard normal CDF\footnote{Hence, $\Phi_{s_X}$ is the CDF of a zero-mean
 Gaussian with standard deviation $s_X$.}, $S(X)$ be the logistic function, and
@@
 that the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be
-made very small, approaching single floating precision (about $10^{-7}$).
-
+made very small, approaching single floating precision (about $10^{-7}$), and
 Algorithm~\ref{alg:our_algorithm} describes our procedure. A few points:
@@
 
-We run MCMC with our MH test and benchmark with~\citet{cutting_mh_2014}
+We run MCMC with our M-H test and benchmark with~\citet{cutting_mh_2014}
 and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.  For
-the former, we increment minibatches by 100 and for the latter, we increase it
-geometrically with a ratio of $\gamma = 1.5$.  The tolerance for making a
-decision in~\citet{cutting_mh_2014} is $\epsilon=0.005$, and the error bound
-control parameters in~\citet{icml2014c1_bardenet14} are $p = 2$ and $\delta =
-0.01$.  All methods (including ours) use the same random walk proposer with
-covariance $\Sigma={\rm diag}(0.15, 0.15)$.  Since the proposer generates a
-random walk, all shaping of distribution of the samples is due to the MH test.
-All methods are run 5000 times to collect 5000 samples.
+the former, we increment minibatches by 100 and set the tolerance
+$\epsilon=0.005$. For the latter, we increase sizes geometrically with a ratio
+of $\gamma = 1.5$ and use error parameters $p = 2$ and $\delta = 0.01$.  All
+methods collect 5000 samples using the same random walk proposer with covariance
+${\rm diag}(0.15, 0.15)$, which means all shaping of distribution of the samples
+is due to the M-H test.
 
@@
 \]
-and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
+and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
 \begin{align*}

### 2016-10-12 aistats2017.tex
Now about 8.3 pages

@@
 % describe this in much detail.
-In fact we can do better than $O(n^{-1/2})$ error for the quantitative CLT
-(showing the error decreases as $O(n^{-1})$) by using a more precise limit
-distribution under an additional assumption. We review the details in
-the Supplementary Material, Section~\ref{app:better_error_bound}, and leave the
-implementation details to future work.
+In fact we can do better ($O(n^{-1})$) by using a more precise limit
+distribution under an additional assumption. We review this in the Supplementary
+Material, Section~\ref{app:better_error_bound}, and leave the details to future
+work.
 
@@
 
-We conduct two sets of experiments to explore the benefits of our minibatch MH
-test and to benchmark it with previous work. In Section~\ref{ssec:gaussians}, we
-show that our test enables samples to converge to the posterior distribution of
-a heated Gaussian mixture model. In Section~\ref{ssec:logistic}, we analyze its
-efficiency on logistic regression.
-
-\subsection{Mixture of Gaussians}\label{ssec:gaussians}
-
+We conduct two sets of experiments. In Section~\ref{ssec:gaussians}, we analyze
+convergence to the posterior of a heated Gaussian mixture. In
+Section~\ref{ssec:logistic}, we analyze its efficiency on logistic regression.
+
+% Daniel: moving this here so figures don't overflow to next pages. We can tune
+% the positioning if needed.
 \begin{figure*}[t]
@@
 
-We borrow a Gaussian mixture experiment from~\citet{langevin_2011}.  The
-parameter is 2-D, $\theta = \langle \theta_1,\theta_2 \rangle$, and the
-generation process is
-\begin{equation}\label{eq:data_generation}
-\begin{split}
-    \theta &\sim \mathcal{N}(0, {\rm diag}(\sigma_1^2,\sigma_2^2)) \\
-    x_i & \sim 0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
-\end{split}
-\end{equation}
-We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
-\langle 0,1 \rangle$. The original paper sampled 100 data points from this
-distribution, and estimated the posterior in the parameters. We are interested
-in performance on larger problems and so sampled 1,000,000 points to form the
-posterior of $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$ with the same
-prior from Equation~(\ref{eq:data_generation}). This produces a much sharper
-posterior with two very narrow peaks.  Our goal is to reproduce the original
-posterior, so we adjust the temperature to $T=10,000$.  Taking logs, we get the
-target as shown in the far left of Figure~\ref{fig:gauss_mix_1}.
-
-We run MCMC with our M-H test and benchmark with~\citet{cutting_mh_2014}
-and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.  For
-the former, we increment minibatches by 100 and set the tolerance
-$\epsilon=0.005$. For the latter, we increase sizes geometrically with a ratio
-of $\gamma = 1.5$ and use error parameters $p = 2$ and $\delta = 0.01$.  All
-methods collect 5000 samples using the same random walk proposer with covariance
-${\rm diag}(0.15, 0.15)$, which means all shaping of distribution of the samples
-is due to the M-H test.
-
-Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
-samples for the three methods, with darker regions indicating a greater density
-of points. There are no obvious differences, so we measure the similarity
-between each set of samples and the actual posterior.
-
-We discretize the posterior coordinates into bins with respect to the two
-components of $\theta$.  The probability $P_i$ of a sample falling into bin $i$
-is the integral of the true posterior probability over the area of that bin.  A
-single sample from any of the MH methods should therefore be multinomial with
-distribution $P$, and a set of $n$ (ideally independent) samples should be ${\rm
-Multinomial}(P,n)$.  The ideal distribution is simple, so we can use it to
-measure the quality of the sample distributions rather than use general purpose
-tests like KL-divergence or likelihood-ratio, which can be problematic with zero
-counts in some bins as we have here.
-
-For large $n$, the per-bin distributions are approximated by Poissons with
-parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
-$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
-out of $N_{\rm bins}$ total. We have
-\begin{equation}\label{eq:log_prob_poissons}
-\begin{split}
-    \log p(c_1, \ldots, c_{N_{\rm bins}} \mid P_1, \ldots, P_{N_{\rm bins}}) & =  \\
-    \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)). & \\
-\end{split}
-\end{equation}
-The results for each model are shown in Table~\ref{tab:poissons}. It is
-difficult, however, to interpret the scores. Thus, we perform significance tests
-to show the difference between the MCMC-sampled distributions and the
-ground-truth using the Chi-Squared distribution as the test statistic, which we
-also show in Table~\ref{tab:poissons}. Both results imply that our method is
-slightly superior to~\citet{cutting_mh_2014}, but slightly worse
-than~\citet{icml2014c1_bardenet14}.
-
-Figure~\ref{fig:gauss_mix_2}, however, suggests that our method dominates in
-terms of speed and efficiency. It shows histograms of the (final) minibatch
-sizes used each iteration. Our method consumes
-significantly less data; most sizes are smaller than 1000, and the
-average size is 210.  The other methods occasionally need to consume nearly all
-data points, and average minibatch sizes are 15562 and 16857. The average
-minibatch sizes roughly predict the running times of these methods since all
-have a running time proportional to the total data consumed, with the exception
-of~\citet{icml2014c1_bardenet14}, which requires a pass over the data to compute
-$C_{\theta,\theta'}$ (see Equation~(\ref{eq:bad_bound})). Using minibatch sizes
-as a proxy for time also avoids discrepancies due to code optimization.
+\subsection{Mixture of Gaussians}\label{ssec:gaussians}
 
@@
 
-\subsection{Logistic Regression}\label{ssec:logistic}
-
+We borrow a Gaussian mixture experiment from~\citet{langevin_2011}.  The
+parameter is 2-D, $\theta = \langle \theta_1,\theta_2 \rangle$, and the
+generation process is
+\begin{equation}\label{eq:data_generation}
+\begin{split}
+    \theta &\sim \mathcal{N}(0, {\rm diag}(\sigma_1^2,\sigma_2^2)) \\
+    x_i & \sim 0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
+\end{split}
+\end{equation}
+We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
+\langle 0,1 \rangle$. The original paper sampled 100 data points and estimated
+the posterior. We are interested in performance on larger problems and so
+sampled 1,000,000 points to form the posterior of
+$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$ with the same prior from
+Equation~(\ref{eq:data_generation}). This produces a much sharper posterior with
+two very narrow peaks.  Our goal is to reproduce the original posterior, so we
+adjust the temperature to $T=10,000$.  Taking logs, we get the target as shown
+in the far left of Figure~\ref{fig:gauss_mix_1}.
+
+We run MCMC with our M-H test and benchmark with~\citet{cutting_mh_2014}
+and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.  For
+the former, we increment minibatches by 100 and set the tolerance
+$\epsilon=0.005$. For the latter, we increase sizes geometrically with a ratio
+of $\gamma = 1.5$ and use error parameters $p = 2$ and $\delta = 0.01$.  All
+methods collect 5000 samples using the same random walk proposer with covariance
+${\rm diag}(0.15, 0.15)$, which means all shaping of distribution of the samples
+is due to the M-H test.
+
+Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
+samples for the three methods, with darker regions indicating a greater density
+of points. There are no obvious differences, so we measure the similarity
+between each set of samples and the actual posterior.
+
+We discretize the posterior coordinates into bins with respect to the two
+components of $\theta$.  The probability $P_i$ of a sample falling into bin $i$
+is the integral of the true posterior probability over the area of that bin.  A
+single sample from any of the MH methods should therefore be multinomial with
+distribution $P$, and a set of $n$ (ideally independent) samples should be ${\rm
+Multinomial}(P,n)$.  The ideal distribution is simple, so we can use it to
+measure the quality of the sample distributions rather than use general purpose
+tests like KL-divergence or likelihood-ratio, which can be problematic with zero
+counts in some bins as we have here.
+
+For large $n$, the per-bin distributions are approximated by Poissons with
+parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
+$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
+out of $N_{\rm bins}$ total. We have
+\begin{equation}\label{eq:log_prob_poissons}
+\begin{split}
+    \log p(c_1, \ldots, c_{N_{\rm bins}} \mid P_1, \ldots, P_{N_{\rm bins}}) & =  \\
+    \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)). & \\
+\end{split}
+\end{equation}
+Table~\ref{tab:poissons} shows the results. It is difficult, however, to
+interpret the scores, so we perform significance tests to show the difference
+between the MCMC-sampled distributions and ground-truth using the Chi-Squared
+distribution as the test statistic (also in Table~\ref{tab:poissons}). Both
+imply that our method is slightly superior to~\citet{cutting_mh_2014}, but
+slightly worse than~\citet{icml2014c1_bardenet14}.
+
+Figure~\ref{fig:gauss_mix_2}, however, suggests that our method dominates in
+terms of speed and efficiency. It shows histograms of the (final) minibatch
+sizes used each iteration. Our method consumes significantly less data; most
+sizes are smaller than 1000, and the average size is 210.  The other methods
+occasionally need to consume nearly all data points, and average minibatch sizes
+are 15562 and 16857. The average minibatch sizes roughly predict the running
+times of these methods since all have a running time proportional to the total
+data consumed, with the exception of~\citet{icml2014c1_bardenet14}, which
+requires a pass over the data to compute $C_{\theta,\theta'}$ (see
+Equation~(\ref{eq:bad_bound})). Using minibatch sizes as a proxy for time also
+avoids discrepancies due to code optimization.
+
+% Daniel: moving this here so figures don't overflow to next pages. We can tune
+% the positioning if needed.
 \begin{figure*}[t]
@@
 
+\subsection{Logistic Regression}\label{ssec:logistic}
+
 We next use logistic regression for the binary classification of 1s versus 7s on
@@
 
-Figure~\ref{fig:logistic_performance} shows the testing log likelihood and
-prediction accuracy, each as a function of the cumulative training data points
+Figure~\ref{fig:logistic_performance} shows the test log likelihood and
+prediction accuracy as a function of the cumulative training points
 processed.\footnote{Note that the curves do not span the same length over the
@@
 so the corresponding curve will ``end'' before the other two.} To generate the
-curves, for each of the 3000 sampled vectors $\theta_t$,
-$t\in\{1,\ldots,3000\}$, we use $\theta_t$ as the parameter for logistic
-regression.  Our minibatch MH test is more efficient in terms of
-accuracy, achieving convergence using roughly half as many elements
-compared to~\citet{cutting_mh_2014} and \textbf{TODO TODO TODO} compared
-to~\citet{icml2014c1_bardenet14}, though its log likelihood results lag slightly
-behind the former algorithm during the very early stages.
+curves, for each of the sampled vectors $\theta_t$, $t\in\{1,\ldots,3000\}$, we
+use $\theta_t$ as the parameter for logistic regression.  Our minibatch MH test
+is more efficient in terms of accuracy, achieving convergence using roughly half
+as many elements compared to~\citet{cutting_mh_2014} and \textbf{TODO TODO TODO}
+compared to~\citet{icml2014c1_bardenet14}, though its log likelihood results lag
+slightly behind the former algorithm during the very early stages.
 
@@
 
-In this paper, we have derived a new MH test for minibatch MCMC methods. We
-demonstrated how a simple deconvolution process allows us to use a minibatch
-approximation to the full data tests. We experimentally show the benefits of our
-test on Gaussian mixtures and logistic regression.  Straightforward directions
-for future work include running more experiments with a particular focus on
-investigation of the variance precondition.  More elaborate extensions include
-testing on deep neural networks and combining our results with Hamiltonian Monte
-Carlo methods, providing a recipe for how to use our algorithm (following the
-framework of~\citet{sgmcmc_2015}), or integrating parallel
-MCMC~\citep{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
+In this paper, we derived an M-H test for minibatch MCMC which approximates full
+data tests. We provide theoretical results and experimentally show the benefits
+of our test on Gaussian mixtures and logistic regression. Directions for future
+work include running more experiments with a particular focus on controlling
+variances.  More elaborate extensions include testing on neural networks,
+combining our results with Hamiltonian Monte Carlo methods, providing a recipe
+for how to use our algorithm (following the framework of~\citet{sgmcmc_2015}),
+or integrating parallel MCMC~\citep{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14}
+concepts.
 

### 2016-10-12 aistats2017.tex
More shortening of paper, roughly 8.25 pages

@@
 
-We now derive error bounds for our M-H test, and for the approximate target
-distribution that it generates. From Table~\ref{tab:xcorr}, we know that it is
-possible to generate the correction samples $X_{\rm corr}$ with a CDF error
-approaching single-precision floating point error. We therefore treat $X_{\rm
-corr}$ as a sample from the exact correction distribution and we will not
-analyze its errors.
+We now derive error bounds for our M-H test and the approximate target
+distribution it generates. (We treat $X_{\rm corr}$ samples from the exact
+correction distribution based on the results in Table~\ref{tab:xcorr}.)
 
@@
 \end{equation}
-Such bounds can be derived efficiently for models such as logistic regression,
-but it is unclear how to derive them for a complex model such as a neural
-network. In general, since a new $\theta'$ value is obtained each iteration, one
-would need to use all the $p(x_i\mid \theta')$ terms\footnote{The sample code
+It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
+regression, but it is unclear how to do so for more complicated models, and one
+generally needs to use all $p(x_i\mid \theta')$ terms\footnote{The sample code
 provided by the authors of~\citet{icml2014c1_bardenet14} in fact computes
 $C_{\theta,\theta'}$ explicitly for each sample generated, i.e. it traverses the
-entire data, thus providing no performance advantage over the complete test.}.
-In contrast, we use quantitative forms of the Central Limit Theorem which rely
-on measurable statistics from a single minibatch. Thus a sampler using our
-approach does not need to see data beyond the current minibatch.
+entire data, thus providing no performance advantage over the complete test.} to
+get $C_{\theta,\theta'}$.  In contrast, we use quantitative forms of the Central
+Limit Theorem which rely on measurable statistics from a single minibatch. Thus
+a sampler using our approach does not need to see data beyond the current
+minibatch.
 
@@
 thousands of points, but otherwise bootstrap CIs can be computed from those
-sequences. Since the bounds are monotone in $E|X|$ and $E|X|^3$, using upper
-bootstrap CI limits will provide high-confidence error bounds.
+sequences.
+
+% Daniel: I removed this sentence to save space.
+% Since the bounds are monotone in $E|X|$ and $E|X|^3$, using upper
+% bootstrap CI limits will provide high-confidence error bounds.
 

### 2016-10-12 aistats2017.tex
About 8.05 pages. Very close ...

@@
 $C_\sigma(X)$ grows and approaches the derivative of a logistic function at
-$\sigma = 0$. It retains some very weak negative values for $\sigma > 0$ but
-removal of those values leads to very small error. Table~\ref{tab:xcorr} shows
-that the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be
-made very small, approaching single floating precision (about $10^{-7}$), and
-Algorithm~\ref{alg:our_algorithm} describes our procedure. A few points:
+$\sigma = 0$. It retains some weak negative values for $\sigma > 0$ but removal
+of those leads to small error. Table~\ref{tab:xcorr} shows that the errors
+between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ approach single floating
+precision (about $10^{-7}$), and Algorithm~\ref{alg:our_algorithm} describes our
+procedure. A few points:
 
@@
 We now derive error bounds for our M-H test and the approximate target
-distribution it generates. (We treat $X_{\rm corr}$ samples from the exact
-correction distribution based on the results in Table~\ref{tab:xcorr}.)
-
-In the most similar prior works,~\citet{cutting_mh_2014} uses asymptotic
-arguments based on the CLT to argue that its approximate acceptance test error
-tends to zero as batch size increases, but no quantitative bounds are given.
+distribution it generates. In the most similar prior
+works,~\citet{cutting_mh_2014} uses asymptotic arguments based on the CLT to
+argue that its approximate acceptance test error tends to zero as batch size
+increases, but no quantitative bounds are given.
 In~\citet{icml2014c1_bardenet14}, explicit bounds are given, but they depend on
@@
 \begin{equation}
-    \bar{X} = \frac{1}{n}\sum_{i=1}^nX_i, \quad s_X = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\bar{X}^2\right)^{\frac{1}{2}}.
+    \bar{X} = \frac{1}{n}\sum_{i=1}^nX_i, \quad s_X = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\bar{X})^2\right)^{\frac{1}{2}}.
 \end{equation}
@@
 corresponding to our minibatch, we can connect $\bar{X}$ with our $\Delta^*$
-term:
-\begin{equation}\label{eq:x-hat}
-    \bar{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta'),
-\end{equation}
-so that $s_X = s_{\Delta^*}$. This results in the following:
+term: $\bar{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta')$, so that
+$s_X = s_{\Delta^*}$. This results in the following:
 
@@
 - \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each minibatch. We
-expect this will often be accurate enough on minibatches with hundreds or
-thousands of points, but otherwise bootstrap CIs can be computed from those
-sequences.
-
-% Daniel: I removed this sentence to save space.
-% Since the bounds are monotone in $E|X|$ and $E|X|^3$, using upper
-% bootstrap CI limits will provide high-confidence error bounds.
-
-\subsection{Error Bounds are Preserved After Adding Random Variables}\label{ssec:preserve_bounds}
+expect this will often be accurate enough on minibatches with hundreds of
+points, but otherwise bootstrap CIs can be computed.
+
+% Daniel: Lame, I know, but it saves one line.
+%\subsection{Error Bounds are Preserved After Adding Random Variables}\label{ssec:preserve_bounds}
+\subsection{Adding Random Variables}\label{ssec:preserve_bounds}
 
@@
 \begin{equation}
-    \sup_x|P'(x)-Q'(x)|\leq \epsilon
+    \sup_x|P'(x)-Q'(x)|\leq \epsilon.
 \end{equation}
 \end{lemma}
-
 \begin{proof}
@@
 the random variables we use, showing that our test should remain accurate.
-
 % Daniel: I'm moving this paragraph over here and putting the rest of it in the
@@
 
-We conduct two sets of experiments. In Section~\ref{ssec:gaussians}, we analyze
-convergence to the posterior of a heated Gaussian mixture. In
-Section~\ref{ssec:logistic}, we analyze its efficiency on logistic regression.
+We conduct two sets of experiments analyzing convergence to the posterior of a
+heated Gaussian mixture, and efficiency on logistic regression.
 
@@
 
-We run MCMC with our M-H test and benchmark with~\citet{cutting_mh_2014}
-and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.  For
-the former, we increment minibatches by 100 and set the tolerance
-$\epsilon=0.005$. For the latter, we increase sizes geometrically with a ratio
-of $\gamma = 1.5$ and use error parameters $p = 2$ and $\delta = 0.01$.  All
-methods collect 5000 samples using the same random walk proposer with covariance
-${\rm diag}(0.15, 0.15)$, which means all shaping of distribution of the samples
-is due to the M-H test.
+We run MCMC with our test and benchmark with~\citet{cutting_mh_2014}
+and~\citet{icml2014c1_bardenet14}, all with initial minibatch size 100.  For the
+former, we increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For
+the latter, we increase sizes geometrically with $\gamma = 1.5$ and use
+parameters $p = 2, \delta = 0.01$.  All methods collect 5000 samples using a
+random walk proposer with covariance ${\rm diag}(0.15, 0.15)$, which means the
+M-H test is responsible for the sample distribution.
 
@@
 components of $\theta$.  The probability $P_i$ of a sample falling into bin $i$
-is the integral of the true posterior probability over the area of that bin.  A
-single sample from any of the MH methods should therefore be multinomial with
-distribution $P$, and a set of $n$ (ideally independent) samples should be ${\rm
-Multinomial}(P,n)$.  The ideal distribution is simple, so we can use it to
-measure the quality of the sample distributions rather than use general purpose
-tests like KL-divergence or likelihood-ratio, which can be problematic with zero
-counts in some bins as we have here.
+is the integral of the true posterior over the bin's area.  A single sample
+should therefore be multinomial with distribution $P$, and a set of $n$ (ideally
+independent) samples is ${\rm Multinomial}(P,n)$. This distribution is simple
+and we can use it to measure the quality of the samples rather than use general
+purpose tests like KL-divergence or likelihood-ratio, which can be problematic
+with zero counts in bins.
 
@@
 \end{equation}
-Table~\ref{tab:poissons} shows the results. It is difficult, however, to
-interpret the scores, so we perform significance tests to show the difference
-between the MCMC-sampled distributions and ground-truth using the Chi-Squared
-distribution as the test statistic (also in Table~\ref{tab:poissons}). Both
-imply that our method is slightly superior to~\citet{cutting_mh_2014}, but
-slightly worse than~\citet{icml2014c1_bardenet14}.
+Table~\ref{tab:poissons} shows the results, but it is difficult to interpret the
+scores, so we perform significance tests to show the difference between the
+MCMC-samples and ground-truth using the Chi-Squared distribution as the test
+statistic (also in Table~\ref{tab:poissons}). Both metrics imply that our method
+is slightly superior to~\citet{cutting_mh_2014}, but slightly worse
+than~\citet{icml2014c1_bardenet14}.
 
 Figure~\ref{fig:gauss_mix_2}, however, suggests that our method dominates in
-terms of speed and efficiency. It shows histograms of the (final) minibatch
-sizes used each iteration. Our method consumes significantly less data; most
-sizes are smaller than 1000, and the average size is 210.  The other methods
-occasionally need to consume nearly all data points, and average minibatch sizes
-are 15562 and 16857. The average minibatch sizes roughly predict the running
-times of these methods since all have a running time proportional to the total
-data consumed, with the exception of~\citet{icml2014c1_bardenet14}, which
-requires a pass over the data to compute $C_{\theta,\theta'}$ (see
-Equation~(\ref{eq:bad_bound})). Using minibatch sizes as a proxy for time also
-avoids discrepancies due to code optimization.
+terms of speed and efficiency. The histograms of the (final) minibatch sizes
+used each iteration show that our method consumes significantly less data; the
+distribution is short-tailed and the mean is 210. The other methods occasionally
+need to use nearly all data points, and average minibatch sizes are 15562 and
+16857. The sizes roughly predict the running times of these methods, which are
+proportional to the data consumed, with the exception
+of~\citet{icml2014c1_bardenet14}, which requires a full pass to compute
+$C_{\theta,\theta'}$ (Equation~(\ref{eq:bad_bound})). Using minibatch sizes
+as a proxy for time also avoids discrepancies due to code optimization.
 
@@
 We next use logistic regression for the binary classification of 1s versus 7s on
-a subset of the MNIST8M dataset, which is a larger version of
-MNIST~\citep{lecun-mnisthandwrittendigit-2010} and well-suited to our desire to
-run MCMC on large data. We randomly subsampled 450k training and 192k testing
-points. The pixels are scaled in $[0,1]$. We impose a flat prior on $\theta$ and
-again use a random walk proposer, this time with covariance matrix $0.05I$ for
-the $784\times 784$ identity matrix $I$. The posterior temperature is set at
-$T=1000$.  We run our MH test for 3000 samples and again compare with the two
-baseline algorithms with starting minibatch size 100. To
-make~\citet{cutting_mh_2014} run faster, we use a larger $\epsilon=0.05$ value.
+a subset of the MNIST8M dataset, a larger version of
+MNIST~\citep{lecun-mnisthandwrittendigit-2010} and well-suited to our big data
+objectives. We randomly subsampled 450k training and 192k testing points. The
+pixels are scaled in $[0,1]$. We impose a flat prior on $\theta$ and again use a
+random walk proposer, this time with covariance matrix $0.05I$ for the
+$784\times 784$ identity matrix $I$. The temperature is set at $T=1000$. We run
+the three tests for samples and again use 100 as the starting minibatch size. To
+make~\citet{cutting_mh_2014} run faster, we now use $\epsilon=0.05$.
 For~\citet{icml2014c1_bardenet14}, we tuned $\gamma \in \{1.1,1.25,1.5,1.75,2\}$
-before concluding that $\gamma=1.5$ was reasonable. In addition, we originally
-used their suggested analytic approximation to $C_{\theta,\theta'}$, but found
-that the resulting values were too high, meaning that their method needed to
-consume the entire dataset each iteration. Thus, we naively compute
-$C_{\theta,\theta'}$.
+before concluding that $\gamma=2.0$ was OK. In addition, we tried their analytic
+approximation to $C_{\theta,\theta'}$, but found that the bounds were too high,
+meaning that their method needed to consume the entire dataset each iteration.
+Thus, we naively compute $C_{\theta,\theta'}$.
 
@@
 
-Figure~\ref{fig:logistic_minibatch}, in a similar manner as
-Figure~\ref{fig:gauss_mix_2}, shows the histogram of minibatch sizes for all
-three methods on a log-log scale. With an initial size of 100, our method
-achieves an average minibatch size of 393, far smaller than the averages of the
-other two methods, which is due to their longer-tailed distributions.
+Figure~\ref{fig:logistic_minibatch} shows histograms of minibatch sizes for the
+three methods on a log-log scale. With an initial size of 100, we achieve an
+average minibatch size of 393, far smaller than the averages of the other two
+methods, which have longer-tailed distributions.
 

### 2016-10-12 aistats2017.tex
OK it is now 8 pages. Still need Xinlei's Bardenet results.

@@
 \end{equation}
-where $*$ denotes convolution.
-
-For computation of $C_\sigma$, we assume that its input $Y$ and another variable
-$X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
-% Daniel: removing this because we were explaining this in quite some detail and
-% it was taking a lot of space.
-%\[
-%    \argmin_{C\sigma} \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|.
-%\]
-We discretize the convolution by discretizing $X$ and $Y$ into $4N+1$ and $2N+1$
-values respectively. If $i \in \{-2N, \ldots, 2N\}=\mathcal{I}$ and $j \in \{-N,
-\ldots, N\}=\mathcal{J}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
-objective can be written as:
+where $*$ denotes convolution. To compute $C_\sigma$, we assume the input $Y$
+and another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$,
+respectively.  We discretize the convolution by discretizing $X$ and $Y$ into
+$4N+1$ and $2N+1$ values respectively. If $i \in \{-2N, \ldots,
+2N\}=\mathcal{I}$ and $j \in \{-N, \ldots, N\}=\mathcal{J}$, then we can write
+$X_i = i(V/N)$ and $Y_j = j(V/N)$, and the objective can be written as:
 \[
@@
 \]
-We now define a matrix $M$ and vectors $u$ and $v$ such that $M_{ij} =
+Now define matrix $M$ and vectors $u$ and $v$ such that $M_{ij} =
 \Phi_{\sigma}(X_i-Y_j)$, $u_j = C_{\sigma}(Y_j)$, and $v_i = S(X_i)$, where the
-indices $i$ and $j$ are appropriately translated to be non-negative indices for
-$M, u,$ and $v$. Thus, the problem is now to minimize $\|Mu-v\|_{\infty}$ with
-the constraint that $u > 0$ since it represents a density. We approximate this
-with a least squares solution:
+indices $i$ and $j$ are appropriately translated to be non-negative for $M, u,$
+and $v$. The problem is now to minimize $\|Mu-v\|_{\infty}$ with the density
+non-negative constraint $u > 0$. We approximate this with least squares:
 \begin{equation}\label{eq:optimization_l2}
@@
     \item It uses an adaptive step size so as to use the smallest possible
-    average minibatch size. Unlike previous work however (and as we show in
-    Section~\ref{sec:experiments}) the size distribution is short-tailed.
+    average minibatch size. Unlike previous work, the size distribution is
+    short-tailed.
 
@@
 
-    \item The sample variance $s^2_{\Delta^*}$ is proportional to
-    $\|\theta'-\theta\|_2^2$ whose distribution for Normal proposals is the
-    square of a normal variable. 
+    % Daniel: I cut some space to use this:
+    \item The sample variance is denoted as $s^2_{\Delta^*}$ and is proportional
+    to $\|\theta'-\theta\|_2^2$.
+
+    % Daniel: instead of this:
+    %\item The sample variance $s^2_{\Delta^*}$ is proportional to
+    %$\|\theta'-\theta\|_2^2$ whose distribution for Normal proposals is the
+    %square of a normal variable. 
 \end{itemize}
@@
 
-We now derive error bounds for our M-H test and the approximate target
-distribution it generates. In the most similar prior
-works,~\citet{cutting_mh_2014} uses asymptotic arguments based on the CLT to
-argue that its approximate acceptance test error tends to zero as batch size
-increases, but no quantitative bounds are given.
-In~\citet{icml2014c1_bardenet14}, explicit bounds are given, but they depend on
-bounding:
+We now derive error bounds for our M-H test and the target distribution it
+generates. In the most similar prior works,~\citet{cutting_mh_2014} use CLT
+asymptotic arguments to show that its approximate acceptance test error tends to
+zero as batch size increases, but no quantitative bounds are given.
+In~\citet{icml2014c1_bardenet14}, explicit bounds are given and depend on
 \begin{equation}\label{eq:bad_bound}
@@
 generally needs to use all $p(x_i\mid \theta')$ terms\footnote{The sample code
-provided by the authors of~\citet{icml2014c1_bardenet14} in fact computes
-$C_{\theta,\theta'}$ explicitly for each sample generated, i.e. it traverses the
-entire data, thus providing no performance advantage over the complete test.} to
-get $C_{\theta,\theta'}$.  In contrast, we use quantitative forms of the Central
-Limit Theorem which rely on measurable statistics from a single minibatch. Thus
-a sampler using our approach does not need to see data beyond the current
+provided by~\citet{icml2014c1_bardenet14} computes $C_{\theta,\theta'}$
+by traversing the entire data, thus providing no performance advantage over the
+complete test.} to get $C_{\theta,\theta'}$.  In contrast, we use quantitative
+forms of the CLT which rely on measurable statistics from a \emph{single}
 minibatch.
@@
 \end{equation}
-This implies $t=\bar{X}/s_X$ has an approximate Student's distribution
-which approaches a normal distribution in the limit. Then
+This means $t=\bar{X}/s_X$ has an approximate Student's distribution
+which approaches a Gaussian. Then
 \begin{equation}\label{eq:clt-bounds}
@@
 
-Lemma~\ref{lem:quant_clt} demonstrates that as long as we know the first and
-third absolute moments $E|X|$ and $E|X|^3$, we can bound the error of the normal
-approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of
-variables $y = x s_X$, Equation~(\ref{eq:clt-bounds}) becomes
+Lemma~\ref{lem:quant_clt} demonstrates that if we know $E|X|$ and $E|X|^3$, we
+can bound the error of the normal approximation, which decays as
+$O(n^{-\frac{1}{2}})$. Making the change of variables $y = x s_X$,
+Equation~(\ref{eq:clt-bounds}) becomes
 \begin{equation}\label{eq:clt-bounds2}
@@
 
-From Section~\ref{sec:our_algorithm}, as the distribution of $\Delta^*$
-approaches a Gaussian, our new MH test becomes more accurate.
 Corollary~\ref{cor:bounds_preserved} shows that the bounds from
-Section~\ref{ssec:delta_star_distribution} are preserved after the addition of
-the random variables we use, showing that our test should remain accurate.
-% Daniel: I'm moving this paragraph over here and putting the rest of it in the
-% supplementary material since our code doesn't do this, so we should not
-% describe this in much detail.
-In fact we can do better ($O(n^{-1})$) by using a more precise limit
+Section~\ref{ssec:delta_star_distribution} are preserved after adding 
+random variables, so our test remains accurate.  In fact we can do better
+($O(n^{-1})$ instead of $O(n^{-1/2})$) by using a more precise limit
 distribution under an additional assumption. We review this in the Supplementary
-Material, Section~\ref{app:better_error_bound}, and leave the details to future
-work.
+Material, Section~\ref{app:better_error_bound}.
+
+% Daniel: I moved the previous Section 5.4 part to the supplementary material
+% since our code doesn't do this, so we should not describe this in much detail.
 
@@
 
-We conduct two sets of experiments analyzing convergence to the posterior of a
-heated Gaussian mixture, and efficiency on logistic regression.
-
 % Daniel: moving this here so figures don't overflow to next pages. We can tune
@@
     \centering
-    \includegraphics[width=0.85\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
+    \includegraphics[width=0.8\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.85\linewidth]{minibatch_size_gaussian.png}
+    \includegraphics[width=0.8\linewidth]{minibatch_size_gaussian.png}
     \caption{
@@
 \subsection{Mixture of Gaussians}\label{ssec:gaussians}
+
+% Daniel: moving this here so figures don't overflow to next pages. We can tune
+% the positioning if needed.
+\begin{figure*}[t]
+	\centering
+	%\includegraphics[width=0.9\linewidth]{logistic_performance.png}
+	\includegraphics[width=0.8\linewidth]{logistic_regression_ll_acc_results.png}
+	\caption{
+    Logistic regression performance (accuracy/log likelihood) based on
+    cumulative data usage.
+    }
+	\label{fig:logistic_performance}
+\end{figure*}
+\begin{figure*}[t]
+	\centering
+	%\includegraphics[width=0.9\linewidth]{minibatch_size_logistic.png}
+	\includegraphics[width=0.8\linewidth]{logistic_regression_histograms_results.png}
+	\caption{
+    Counts of minibatch sizes in the MNIST logistic regression experiment
+    (analogous to Figure~\ref{fig:gauss_mix_2}).
+    }
+	\label{fig:logistic_minibatch}
+\end{figure*}
+
+We first borrow a Gaussian mixture experiment from~\citet{langevin_2011}.  The
+parameter is 2-D, $\theta = \langle \theta_1,\theta_2 \rangle$, and the
+generation process is
+\begin{equation}\label{eq:data_generation}
+\begin{split}
+    \theta &\sim \mathcal{N}(0, {\rm diag}(\sigma_1^2,\sigma_2^2)) \\
+    x_i & \sim 0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
+\end{split}
+\end{equation}
+We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
+\langle 0,1 \rangle$. The original paper sampled 100 data points and estimated
+the posterior. We are interested in performance on larger problems and so
+sampled 1,000,000 points to form the posterior of
+$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$ with the same prior from
+Equation~(\ref{eq:data_generation}). This produces a much sharper posterior with
+two very narrow peaks.  Our goal is to reproduce the original posterior, so we
+adjust the temperature to $T=10,000$.  Taking logs, we get the target as shown
+in the far left of Figure~\ref{fig:gauss_mix_1}.
+
+We run MCMC with our test and benchmark with~\citet{cutting_mh_2014}
+and~\citet{icml2014c1_bardenet14}, all with initial minibatch size 100.  For the
+former, we increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For
+the latter, we increase sizes geometrically with $\gamma = 1.5$ and use
+parameters $p = 2, \delta = 0.01$.  All methods collect 5000 samples using a
+random walk proposer with covariance ${\rm diag}(0.15, 0.15)$, which means the
+M-H test is responsible for the sample distribution.
+
+Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
+samples for the three methods, with darker regions indicating a greater density
+of points. There are no obvious differences, so we measure the similarity
+between each set of samples and the actual posterior.
+
+We discretize the posterior coordinates into bins with respect to the two
+components of $\theta$.  The probability $P_i$ of a sample falling into bin $i$
+is the integral of the true posterior over the bin's area.  A single sample
+should therefore be multinomial with distribution $P$, and a set of $n$ (ideally
+independent) samples is ${\rm Multinomial}(P,n)$. This distribution is simple
+and we can use it to measure the quality of the samples rather than use general
+purpose tests like KL-divergence or likelihood-ratio, which can be problematic
+with zero counts in bins.
+
+For large $n$, the per-bin distributions are approximated by Poissons with
+parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
+$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
+out of $N_{\rm bins}$ total. We have
+\begin{equation}\label{eq:log_prob_poissons}
+\begin{split}
+    \log p(c_1, \ldots, c_{N_{\rm bins}} \mid P_1, \ldots, P_{N_{\rm bins}}) & =  \\
+    \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)). & \\
+\end{split}
+\end{equation}
+Table~\ref{tab:poissons} shows the results, but it is difficult to interpret the
+scores, so we perform significance tests to show the difference between the
+MCMC-samples and ground-truth using the Chi-Squared distribution as the test
+statistic (also in Table~\ref{tab:poissons}). Both metrics imply that our method
+is slightly superior to~\citet{cutting_mh_2014}, but slightly worse
+than~\citet{icml2014c1_bardenet14}.
+
+Figure~\ref{fig:gauss_mix_2}, however, suggests that our method dominates in
+terms of speed and efficiency. The histograms of the (final) minibatch sizes
+used each iteration show that our method consumes significantly less data; the
+distribution is short-tailed and the mean is 210. The other methods occasionally
+need to use nearly all data points, and average minibatch sizes are 15562 and
+16857. The sizes roughly predict the running times of these methods, which are
+proportional to the data consumed, with the exception of the
+$C_{\theta,\theta'}$ time for~\citet{icml2014c1_bardenet14}.Using minibatch
+sizes as a proxy for time also avoids discrepancies due to code optimization.
+
+\subsection{Logistic Regression}\label{ssec:logistic}
 
@@
 \end{table}
-
-We borrow a Gaussian mixture experiment from~\citet{langevin_2011}.  The
-parameter is 2-D, $\theta = \langle \theta_1,\theta_2 \rangle$, and the
-generation process is
-\begin{equation}\label{eq:data_generation}
-\begin{split}
-    \theta &\sim \mathcal{N}(0, {\rm diag}(\sigma_1^2,\sigma_2^2)) \\
-    x_i & \sim 0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
-\end{split}
-\end{equation}
-We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
-\langle 0,1 \rangle$. The original paper sampled 100 data points and estimated
-the posterior. We are interested in performance on larger problems and so
-sampled 1,000,000 points to form the posterior of
-$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$ with the same prior from
-Equation~(\ref{eq:data_generation}). This produces a much sharper posterior with
-two very narrow peaks.  Our goal is to reproduce the original posterior, so we
-adjust the temperature to $T=10,000$.  Taking logs, we get the target as shown
-in the far left of Figure~\ref{fig:gauss_mix_1}.
-
-We run MCMC with our test and benchmark with~\citet{cutting_mh_2014}
-and~\citet{icml2014c1_bardenet14}, all with initial minibatch size 100.  For the
-former, we increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For
-the latter, we increase sizes geometrically with $\gamma = 1.5$ and use
-parameters $p = 2, \delta = 0.01$.  All methods collect 5000 samples using a
-random walk proposer with covariance ${\rm diag}(0.15, 0.15)$, which means the
-M-H test is responsible for the sample distribution.
-
-Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
-samples for the three methods, with darker regions indicating a greater density
-of points. There are no obvious differences, so we measure the similarity
-between each set of samples and the actual posterior.
-
-We discretize the posterior coordinates into bins with respect to the two
-components of $\theta$.  The probability $P_i$ of a sample falling into bin $i$
-is the integral of the true posterior over the bin's area.  A single sample
-should therefore be multinomial with distribution $P$, and a set of $n$ (ideally
-independent) samples is ${\rm Multinomial}(P,n)$. This distribution is simple
-and we can use it to measure the quality of the samples rather than use general
-purpose tests like KL-divergence or likelihood-ratio, which can be problematic
-with zero counts in bins.
-
-For large $n$, the per-bin distributions are approximated by Poissons with
-parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
-$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
-out of $N_{\rm bins}$ total. We have
-\begin{equation}\label{eq:log_prob_poissons}
-\begin{split}
-    \log p(c_1, \ldots, c_{N_{\rm bins}} \mid P_1, \ldots, P_{N_{\rm bins}}) & =  \\
-    \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)). & \\
-\end{split}
-\end{equation}
-Table~\ref{tab:poissons} shows the results, but it is difficult to interpret the
-scores, so we perform significance tests to show the difference between the
-MCMC-samples and ground-truth using the Chi-Squared distribution as the test
-statistic (also in Table~\ref{tab:poissons}). Both metrics imply that our method
-is slightly superior to~\citet{cutting_mh_2014}, but slightly worse
-than~\citet{icml2014c1_bardenet14}.
-
-Figure~\ref{fig:gauss_mix_2}, however, suggests that our method dominates in
-terms of speed and efficiency. The histograms of the (final) minibatch sizes
-used each iteration show that our method consumes significantly less data; the
-distribution is short-tailed and the mean is 210. The other methods occasionally
-need to use nearly all data points, and average minibatch sizes are 15562 and
-16857. The sizes roughly predict the running times of these methods, which are
-proportional to the data consumed, with the exception
-of~\citet{icml2014c1_bardenet14}, which requires a full pass to compute
-$C_{\theta,\theta'}$ (Equation~(\ref{eq:bad_bound})). Using minibatch sizes
-as a proxy for time also avoids discrepancies due to code optimization.
-
-% Daniel: moving this here so figures don't overflow to next pages. We can tune
-% the positioning if needed.
-\begin{figure*}[t]
-	\centering
-	%\includegraphics[width=0.9\linewidth]{logistic_performance.png}
-	\includegraphics[width=0.85\linewidth]{logistic_regression_ll_acc_results.png}
-	\caption{
-    Logistic regression performance (accuracy/log likelihood) based on
-    cumulative data usage.
-    }
-	\label{fig:logistic_performance}
-\end{figure*}
-\begin{figure*}[t]
-	\centering
-	%\includegraphics[width=0.9\linewidth]{minibatch_size_logistic.png}
-	\includegraphics[width=0.85\linewidth]{logistic_regression_histograms_results.png}
-	\caption{
-    Counts of minibatch sizes in the MNIST logistic regression experiment
-    (analogous to Figure~\ref{fig:gauss_mix_2}).
-    }
-	\label{fig:logistic_minibatch}
-\end{figure*}
-
-\subsection{Logistic Regression}\label{ssec:logistic}
 
@@
 prediction accuracy as a function of the cumulative training points
-processed.\footnote{Note that the curves do not span the same length over the
-x-axis, because our test consumes fewer samples throughout the MCMC procedure,
-so the corresponding curve will ``end'' before the other two.} To generate the
-curves, for each of the sampled vectors $\theta_t$, $t\in\{1,\ldots,3000\}$, we
-use $\theta_t$ as the parameter for logistic regression.  Our minibatch MH test
-is more efficient in terms of accuracy, achieving convergence using roughly half
-as many elements compared to~\citet{cutting_mh_2014} and \textbf{TODO TODO TODO}
-compared to~\citet{icml2014c1_bardenet14}, though its log likelihood results lag
-slightly behind the former algorithm during the very early stages.
+processed.\footnote{The curves do not span the same length over the x-axis since
+the methods consume different amounts of data.} To generate the curves, for each
+of the sampled vectors $\theta_t$, $t\in\{1,\ldots,3000\}$, we use $\theta_t$ as
+the parameter for logistic regression.  Our minibatch MH test is more efficient
+in terms of accuracy, achieving convergence using roughly half as many elements
+compared to~\citet{cutting_mh_2014} and \textbf{TODO TODO TODO} compared
+to~\citet{icml2014c1_bardenet14}, though its log likelihood results lag slightly
+behind the former algorithm during the very early stages.
 
@@
 
-In this paper, we derived an M-H test for minibatch MCMC which approximates full
-data tests. We provide theoretical results and experimentally show the benefits
-of our test on Gaussian mixtures and logistic regression. Directions for future
+We have derived an M-H test for minibatch MCMC which approximates full data
+tests. We present theoretical results and experimentally show the benefits of
+our test on Gaussian mixtures and logistic regression. Directions for future
 work include running more experiments with a particular focus on controlling
-variances.  More elaborate extensions include testing on neural networks,
-combining our results with Hamiltonian Monte Carlo methods, providing a recipe
-for how to use our algorithm (following the framework of~\citet{sgmcmc_2015}),
-or integrating parallel MCMC~\citep{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14}
-concepts.
+variances, testing on neural networks, combining our results with Hamiltonian
+Monte Carlo methods, providing a recipe for how to use our algorithm (following
+the framework of~\citet{sgmcmc_2015}), or integrating parallel
+MCMC~\citep{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
 

### 2016-10-13 aistats2017.tex
Adding Logistic Regression results. Leaving it over to John for review (see emails)

@@
   implementation, and present several experiments. We demonstrate 
-  several order-of-magnitude speedup over previous work, and show for
+  order-of-magnitude speedups over previous work, and show for
   the first time that small expected minibatch sizes are possible. 
@@
 	\caption{
-    Counts of minibatch sizes in the MNIST logistic regression experiment
-    (analogous to Figure~\ref{fig:gauss_mix_2}).
+    Counts of minibatch sizes in the logistic regression experiment (analogous
+    to Figure~\ref{fig:gauss_mix_2}).
     }
@@
 used each iteration show that our method consumes significantly less data; the
-distribution is short-tailed and the mean is 210. The other methods occasionally
-need to use nearly all data points, and average minibatch sizes are 15562 and
-16857. The sizes roughly predict the running times of these methods, which are
+distribution is short-tailed and the mean is 210, more than an order of
+magnitude better compared to the other two methods (averages are 15562 and
+16857). The sizes roughly predict the running times of these methods, which are
 proportional to the data consumed, with the exception of the
-$C_{\theta,\theta'}$ time for~\citet{icml2014c1_bardenet14}.Using minibatch
+$C_{\theta,\theta'}$ time for~\citet{icml2014c1_bardenet14}. Using minibatch
 sizes as a proxy for time also avoids discrepancies due to code optimization.
@@
 the parameter for logistic regression.  Our minibatch MH test is more efficient
-in terms of accuracy, achieving convergence using roughly half as many elements
-compared to~\citet{cutting_mh_2014} and \textbf{TODO TODO TODO} compared
-to~\citet{icml2014c1_bardenet14}, though its log likelihood results lag slightly
-behind the former algorithm during the very early stages.
+in terms of accuracy, achieving convergence roughly twice as fast compared
+to~\citet{cutting_mh_2014} and about 1.5 orders of magnitude faster
+than~\citet{icml2014c1_bardenet14}, though its log likelihood results lag
+slightly behind the former algorithm during the very early stages of
+exploration.
 
@@
 
-%\newpage
+\newpage
 \subsubsection*{References}
@@
 % spaces below the "References" text.
-\vspace{-20pt} 
+\vspace{-25pt} 
 %\bibliographystyle{apalike}

### 2016-10-13 aistats2017.tex
fixes to text

@@
   %Abstract. The Abstract must be limited to one paragraph.
-  Markov chain Monte Carlo (MCMC) methods have many applications in
-  machine learning. We are particularly interested in their application
-  to modeling very large datasets, where it is impractical to perform
-  Metropolis-Hastings tests on the full data. Previous work on reducing
-  the cost of Metropolis-Hastings tests yield variable data consumed per
-  sample, with only constant factor reductions versus using the full
-  dataset for each sample.  Here we present a method that can be tuned
-  to provide arbitrarily small batch sizes, by adjusting either proposal
-  step size or temperature. Our approach uses the natural noise present
-  in minibatch likelihood estimates to furnish the randomness in a
-  Metropolis-Hastings test. Our test uses the noise-tolerant Barker
-  acceptance test with a novel additive correction variable.  The
-  resulting test can be combined with minibatch proposals to yield
-  updates with the same complexity as a simple SGD update. In this paper
-  we derive the test, analyze its performance, discuss its
-  implementation, and present several experiments. We demonstrate 
-  order-of-magnitude speedups over previous work, and show for
-  the first time that small expected minibatch sizes are possible. 
+  We
+  present a novel Metropolis-Hastings method for large datasets that
+  uses small expected-size minibatches of data. Previous work on
+  reducing the cost of Metropolis-Hastings tests yield variable data
+  consumed per sample, with only constant factor reductions versus
+  using the full dataset for each sample.  Here we present a method
+  that can be tuned to provide arbitrarily small batch sizes, by
+  adjusting either proposal step size or temperature. Our test uses
+  the noise-tolerant Barker acceptance test with a novel additive
+  correction variable.  The resulting test has simlar cost to a normal
+  SGD update. Our experiments demonstrate several order-of-magnitude
+  speedup over previous work.
 \end{abstract}
@@
 
-    \item Increase the minibatch size when needed. Log acceptance variance
-    scales as $1/k$ vs the minibatch size $k$. Our test is adaptive like earlier
-    works, but unlike them, the distribution of minibatch size is Gaussian, not
-    long-tailed.  Increased minibatch size also reduces the error rate for the
-    test.
-
     \item For continuous distributions, reduce the proposal step size and
@@
     probability scales as the square of proposal step size.
+
+%  \item Increase the minibatch size when needed. Log acceptance variance
+%    scales as $1/k$ vs the minibatch size $k$. Our test is adaptive like earlier
+%    works, but unlike them, the distribution of minibatch size is Gaussian, not
+%    long-tailed.  Increased minibatch size also reduces the error rate for the
+%    test.
 \end{itemize}
@@
 This in turn reduces the variance of the log acceptance probability and allows
-our test to be applied. Our samples can cover the search space densely with
+our test to be applied. Theses samples can cover the search space densely with
 small steps rather than taking a few sparse steps towards an optimum. In this
@@
 
-This result is used in~\citet{Barker65} to define the Barker acceptance test.  As
-a sanity check, choosing $g(s) = \exp(s) \wedge 1$ --- a function satisfying the
-requirement of Lemma~\ref{lem:detailed_balance} --- produces the classical
-Metropolis acceptance test $\alpha(\theta,\theta') = g(\Delta(\theta,\theta')) =
-\frac{p(\theta')q(\theta \mid \theta')}{p(\theta)q(\theta' \mid \theta)}\wedge
-1$. In fact, $g(s) =\exp(s) \wedge 1$ is the optimal acceptance function in
-terms of acceptance rate, since it accepts with probability 1 for $\Delta > 0$.
+This result is used in~\citet{Barker65} to define the Barker acceptance test.
+%As
+%a sanity check, choosing $g(s) = \exp(s) \wedge 1$ --- a function satisfying the
+%requirement of Lemma~\ref{lem:detailed_balance} --- produces the classical
+%Metropolis acceptance test $\alpha(\theta,\theta') = g(\Delta(\theta,\theta')) =
+%\frac{p(\theta')q(\theta \mid \theta')}{p(\theta)q(\theta' \mid \theta)}\wedge
+%1$.
+%In fact, $g(s) =\exp(s) \wedge 1$ is the optimal acceptance function in
+%terms of acceptance rate, since it accepts with probability 1 for $\Delta > 0$.
 
 \subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
-For our new MH test we use the Barker logistic~\citep{Barker65} function:
-$g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the
-condition in Lemma~\ref{lem:detailed_balance}.  While it is slightly less
-efficient than the Metropolis test when used on the full dataset, we will see
-that its smoothness allows it to naturally tolerate substantial variance in its
-input. This in turn will lead to a much more efficient test on subsets of data.
+For our new MH test we use the Barker logistic~\citep{Barker65}
+function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
+that it satisfies the condition in Lemma~\ref{lem:detailed_balance}.
+It is slightly less efficient than the Metropolis test, since its
+acceptance rate for vanishing likelihood difference is 0.5. However we
+will see that its overall sample efficiency is much higher than the
+earlier methods.
 
@@
 random variable $X$ (call it $X_{\rm log}$ from now on). In~\citet{TallData15} it
-was proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed
-error caused by this approximation.
+was proposed to use $\Delta^*$ in a Barker test, and tolerate the fixed
+error between logistic and normal distributions.
 
@@
 term: $\bar{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta')$, so that
-$s_X = s_{\Delta^*}$. This results in the following:
+$s_X = s_{\Delta^*}$. We can now substitute into Equation~(\ref{eq:clt-bounds2}) and displace by the mean, giving:
 
 \begin{corollary}\label{cor:our_bound_delta_prime}
-We can now substitute into Equation~(\ref{eq:clt-bounds2}) and displace by the mean, giving:
 \begin{align*}
@@
 % taking up too much space.
-
-\begin{comment}
 
@@
 \end{lemma}
-
-\end{comment}
 
@@
 
-We first borrow a Gaussian mixture experiment from~\citet{langevin_2011}.  The
-parameter is 2-D, $\theta = \langle \theta_1,\theta_2 \rangle$, and the
+This model is adapted from~\citet{langevin_2011} by increasing the number
+of samples to 1 million.  The parameters are $\theta = \langle \theta_1,\theta_2 \rangle$, and the
 generation process is
@@
 
-We run MCMC with our test and benchmark with~\citet{cutting_mh_2014}
-and~\citet{icml2014c1_bardenet14}, all with initial minibatch size 100.  For the
-former, we increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For
-the latter, we increase sizes geometrically with $\gamma = 1.5$ and use
-parameters $p = 2, \delta = 0.01$.  All methods collect 5000 samples using a
-random walk proposer with covariance ${\rm diag}(0.15, 0.15)$, which means the
-M-H test is responsible for the sample distribution.
+We ran MCMC with with our test against the test
+from~\citet{icml2014c1_bardenet14} and a simplified version
+of~\citet{cutting_mh_2014}. Both methods use repeated testing which
+must be discounted to bound overall test
+accuracy. ~\citet{icml2014c1_bardenet14} conservatively partitions the
+test error, while~\citet{cutting_mh_2014} used a complex dynamic
+programming heuristic to estimate the overall test error. We did not
+implement this heuristic and instead set the individual test error
+equal to the overall test error. Thus our results for
+~\citet{cutting_mh_2014} should be treated as lower bounds. All
+methods were initialized with minibatch size 100.  For ~\citet{cutting_mh_2014} we
+increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For
+~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
+use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
+samples using a random walk proposer with covariance ${\rm diag}(0.15,
+0.15)$, which means the M-H test is responsible for shaping the sample
+distribution.
 
@@
 and we can use it to measure the quality of the samples rather than use general
-purpose tests like KL-divergence or likelihood-ratio, which can be problematic
-with zero counts in bins.
+purpose tests like KL-divergence or likelihood-ratio, which are problematic
+with zero counts.
 
@@
 \end{equation}
-Table~\ref{tab:poissons} shows the results, but it is difficult to interpret the
-scores, so we perform significance tests to show the difference between the
-MCMC-samples and ground-truth using the Chi-Squared distribution as the test
-statistic (also in Table~\ref{tab:poissons}). Both metrics imply that our method
-is slightly superior to~\citet{cutting_mh_2014}, but slightly worse
-than~\citet{icml2014c1_bardenet14}.
-
-Figure~\ref{fig:gauss_mix_2}, however, suggests that our method dominates in
+Table~\ref{tab:poissons} shows the likelihoods. To facilitate
+interpretation we perform significance tests using Chi-Squared
+distribution (also in Table~\ref{tab:poissons}). Scores lie
+between~\citet{cutting_mh_2014} and~\citet{icml2014c1_bardenet14}, but
+the variance of these values is high and ordering changes depending on
+the range of samples generated.
+
+Figure~\ref{fig:gauss_mix_2} shows that the new method dominates in
 terms of speed and efficiency. The histograms of the (final) minibatch sizes
@@
 magnitude better compared to the other two methods (averages are 15562 and
-16857). The sizes roughly predict the running times of these methods, which are
-proportional to the data consumed, with the exception of the
-$C_{\theta,\theta'}$ time for~\citet{icml2014c1_bardenet14}. Using minibatch
-sizes as a proxy for time also avoids discrepancies due to code optimization.
+16857). The sizes correspond to the running times of the methods, since
+they all consume linear time in data consumed.
 
@@
 
-We next use logistic regression for the binary classification of 1s versus 7s on
-a subset of the MNIST8M dataset, a larger version of
-MNIST~\citep{lecun-mnisthandwrittendigit-2010} and well-suited to our big data
-objectives. We randomly subsampled 450k training and 192k testing points. The
-pixels are scaled in $[0,1]$. We impose a flat prior on $\theta$ and again use a
-random walk proposer, this time with covariance matrix $0.05I$ for the
-$784\times 784$ identity matrix $I$. The temperature is set at $T=1000$. We run
-the three tests for samples and again use 100 as the starting minibatch size. To
-make~\citet{cutting_mh_2014} run faster, we now use $\epsilon=0.05$.
-For~\citet{icml2014c1_bardenet14}, we tuned $\gamma \in \{1.1,1.25,1.5,1.75,2\}$
-before concluding that $\gamma=2.0$ was OK. In addition, we tried their analytic
-approximation to $C_{\theta,\theta'}$, but found that the bounds were too high,
-meaning that their method needed to consume the entire dataset each iteration.
-Thus, we naively compute $C_{\theta,\theta'}$.
+We next test logistic regression for the binary classification of 1s
+versus 7s on a subset of the MNIST8M dataset, a larger version of
+MNIST~\citep{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
+450k training and 192k testing points.
+%%% The pixels are scaled in $[0,1]$  - JFC Why ????
+We impose a uniform prior on $\theta$ and again use a random walk
+proposer, this time with covariance matrix $0.05I$. The temperature is
+set at $T=1000$. We run the three tests for samples and again use 100
+as the starting minibatch size. Method~\citet{cutting_mh_2014} used
+per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
+we used $\gamma=2.0$ (discounting factor). We tried to use the
+symbolic bound for $C_{\theta,\theta'}$ for Logistic regression from
+~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead
+we used the empirical $C_{\theta,\theta'}$ from the entire dataset
+without including the test for this test, so again our results
+represent a lower bound for this method.
 
@@
 
-We demonstrate how we can show $O(n^{-1})$ error for the quantitative CLT by
+We show that the CLT error bound can be improved to $O(n^{-1})$ 
 using a more precise limit distribution under an additional assumption. Let
@@
 
-We previously submitted a much older draft of this paper to NIPS 2016. The
-current manuscript has been substantially revised since that submission. We have
+We previously submitted an early, incomplete draft of this paper to NIPS 2016. The
+current manuscript has been substantially improved since that submission. We have
 made the following changes:
@@
 \begin{enumerate}
-    \item At the time of the NIPS submission, we were unaware of several
-    references which described some of what we wrote. For this submission, we
-    now cite the important work of~\citep{TallData15} which had the idea of
-    using subsampling noise for exploration (rather than the $\log u$ variable).
-    We also cite the work of~\citep{Barker65} which uses the Barker function. We
-    welcome information about other potentially missing references.
-
-    \item The previous theoretical results have been entirely scrapped and
-    replaced with different but more relevant results. The NIPS submission
-    listed a page of theoretical results which did not tie into the rest of the
-    paper's contributions and were confusing to the readers. We now have
-    theoretical results that are much clearer and also \emph{directly} show the
-    expected performance benefits of our acceptance test. Note also that
-    Lemma~\ref{lem:worst_case} is an entirely new addition to this version.
-
-    \item Several reviewers mentioned that our deconvolution approach to
-    determine the correction distribution was not unique and thus ill-defined.
-    Consequently, we have added an entirely new section of the paper
-    (Section~\ref{sec:correction}) explaining how we derived that distribution.
-
-    \item Reviewers pointed out one of our mistakes when we said our minibatch
-    sizes were fixed. We have changed this in the current manuscript to mean
-    that the per-iteration minibatch sizes for our test have a distribution with
-    \emph{shorter tails} than those of prior work, which can be observed by
-    plotting a histogram of minibatch sizes.
+\item The (incomplete) analysis in the NIPS paper has been substantially replaced.
+  The asymptotic CLT arguments in the earlier paper have been replaced with
+  explicit quantitative bounds. There were several gaps in the earlier analysis:
+  the effects of addition of random variables on CDF errors and the construction and
+  error bounds of the correction distribution which are key parts of the present
+  paper have been added.
+
+\item Several recent relevant references have been added, and we have written
+  clearer statements about the contribution of the present work.
+  These include~\citep{TallData15} and the Barker test~\citep{Barker65}.
+
+\item Some reviewers mentioned that our deconvolution approach to
+  determine the correction distribution was not unique and thus
+  ill-defined.  We added detail description on how the correction
+  distribution is computed. While the distribution is not unique, we
+  regularize the inversion to obtain a concrete distribution with very
+  low error (close to single-point machine precision).
+
+\item The earlier draft had an inconsistency in its statement about constant
+  minibatch size. We clarify that this is true only in expected value.
 
@@
   running on the distribution at temperature $T=1$. In this set of
-    experiments, we tuned hyperparameters of the other algorithms to make the
-    comparison fairer. Finally, in the Gaussian mixture model scenario
-    (Section~\ref{ssec:gaussians}), we provide more details on how ``accurate''
-    the samples are, rather than solely relying on the visualization of
-    Figure~\ref{fig:gauss_mix_1}.
+  experiments, we tuned hyperparameters of the other algorithms to yield
+  best results. In the Gaussian mixture model scenario
+  (Section~\ref{ssec:gaussians}), we provide likelihoods and test
+  statistics to measure the accuracy of the samples.
 
@@
 
-We are confident that the current manuscript is of far better quality than the
-NIPS submission, and we appreciate the efforts of the NIPS reviewers to give us
-ideas for improvement.
+We estimate the degree of overlap between the papers in mathematics (equations) at less than
+10\%, in overall text at less than 30\%. The experiments are similar but were also redone at larger
+scale in this draft.
 

### 2016-10-13 aistats2017.tex
Minor typo changes.

@@
   the noise-tolerant Barker acceptance test with a novel additive
-  correction variable.  The resulting test has simlar cost to a normal
-  SGD update. Our experiments demonstrate several order-of-magnitude
-  speedup over previous work.
+  correction variable.  The resulting test has similar cost to a normal
+  SGD update. Our experiments demonstrate order-of-magnitude
+  speedups over previous work.
 \end{abstract}
@@
 This in turn reduces the variance of the log acceptance probability and allows
-our test to be applied. Theses samples can cover the search space densely with
+our test to be applied. These samples can cover the search space densely with
 small steps rather than taking a few sparse steps towards an optimum. In this
@@
 was proposed to use $\Delta^*$ in a Barker test, and tolerate the fixed
-error between logistic and normal distributions.
+error between the logistic and normal distributions.
 
@@
 implement this heuristic and instead set the individual test error
-equal to the overall test error. Thus our results for
-~\citet{cutting_mh_2014} should be treated as lower bounds. All
-methods were initialized with minibatch size 100.  For ~\citet{cutting_mh_2014} we
-increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For
-~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
+equal to the overall test error. Thus our results for~\citet{cutting_mh_2014} should be treated as lower bounds. All
+methods were initialized with minibatch size 100.  For~\citet{cutting_mh_2014} we
+increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
 use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
@@
 proposer, this time with covariance matrix $0.05I$. The temperature is
-set at $T=1000$. We run the three tests for samples and again use 100
+set at $T=1000$. We run the three tests for 3000 samples and again use 100
 as the starting minibatch size. Method~\citet{cutting_mh_2014} used
@@
 we used $\gamma=2.0$ (discounting factor). We tried to use the
-symbolic bound for $C_{\theta,\theta'}$ for Logistic regression from
-~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead
-we used the empirical $C_{\theta,\theta'}$ from the entire dataset
-without including the test for this test, so again our results
-represent a lower bound for this method.
+symbolic bound for $C_{\theta,\theta'}$ for Logistic regression
+from~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead we use
+the empirical $C_{\theta,\theta'}$ from the entire dataset without including the
+test for this test, so again our results represent a lower bound for this
+method.
 
@@
   determine the correction distribution was not unique and thus
-  ill-defined.  We added detail description on how the correction
-  distribution is computed. While the distribution is not unique, we
+  ill-defined.  We added detail on how the correction
+  distribution is computed (see Section~\ref{sec:correction}). While the distribution is not unique, we
   regularize the inversion to obtain a concrete distribution with very
@@
   from~\citep{icml2014c1_bardenet14} (in addition to~\citep{cutting_mh_2014},
-  which we had earlier). In addition, each algorithm now runs MCMC sampling on
+  which we had earlier). Furthermore, each algorithm now runs MCMC sampling on
   the \emph{same} distribution.  Previously, we ran our distribution at a

### 2016-10-13 aistats2017.tex
More minor typos

@@
 minibatch to the next. By contrast,~\citet{conf/uai/MaclaurinA14,TallData15}
-perform exact tests but require a lower bound on parameter distribution across
+perform exact tests but require a lower bound on the parameter distribution across
 its domain.  The amount of data reduction depends on the accuracy of this bound,
@@
     \item We compare performance of our new test and prior approaches on several
-    datasets. We demonstrate orders of magnitude improvements in efficiency
+    datasets. We show order-of-magnitude improvements in efficiency
     (measured as data consumed per test), and that it does not suffer from
@@
 difficult-to-compute probability distribution $p(\theta)$ is sampled using a
-Markov chain $\theta_1,\ldots,\theta_n$. The sample $\theta_{t+1}$ at time $t+1$
+Markov chain $\theta_1,\ldots,\theta_T$. The sample $\theta_{t+1}$ at time $t+1$
 is generated using a candidate $\theta'$ from a (simpler) proposal distribution
@@
 Since minibatches contains randomly selected samples $x_i$, the values
-$\Lambda_i$ are i.i.d. random variables\footnote{The analysis assumes sampling
+$\Lambda_i$ are i.i.d. random variables.\footnote{The analysis assumes sampling
 with replacement although implementations on typical large datasets will
-approximate this by sampling without replacement.}.  By the Central Limit
+approximate this by sampling without replacement.} By the Central Limit
 Theorem, we expect $\Lambda^*(\theta,\theta')$ to be approximately Gaussian. The
@@
 in the amount of data consumed per test compared to full-data tests, so their
-speed-up is $\le 1/c$. The issue is the use of tail bounds to separate $\Delta$
-from zero; for certain input/random $u$ combinations, $\Delta$ can be
+speed-up is $\le 1/c$. The issue is the use of tail bounds to separate $\Lambda-\psi$
+from zero; for certain input/random $u$ combinations, this difference can be
 arbitrarily close to zero. We avoid this by using the {\em approximately normal}
-variation in $\Delta^*$ to {\em replace} the variation due to $u$.
+variation in $\Lambda^*$ to {\em replace} the variation due to $u$.
 

### 2016-10-13 aistats2017.tex
Revised order of magnitude to several order of magnitude

@@
   correction variable.  The resulting test has similar cost to a normal
-  SGD update. Our experiments demonstrate order-of-magnitude
+  SGD update. Our experiments demonstrate several order-of-magnitude
   speedups over previous work.
@@
     \item We compare performance of our new test and prior approaches on several
-    datasets. We show order-of-magnitude improvements in efficiency
-    (measured as data consumed per test), and that it does not suffer from
+    datasets. We demonstrate several order-of-magnitude improvements in efficiency
+    (measured as data consumed per test), and show that it does not suffer from
     long-tailed minibatch sizes.

### 2016-10-13 aistats2017.tex
fix to empiricial c_\theta,\theta' value and clarified why we use gamma = 2

@@
     4000 & 0.9 & 1 & 1.0e-4  \\
-    4000 & 0.8 & 0.03 & 5.0e-6 \\
+    4000 & 0.8 & 0.03 & 5.0e-6
     \end{tabular}
@@
 per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
-we used $\gamma=2.0$ (discounting factor). We tried to use the
+we use their recommended discounting factor value of $\gamma=2.0$. We tried to use the
 symbolic bound for $C_{\theta,\theta'}$ for Logistic regression
 from~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead we use
-the empirical $C_{\theta,\theta'}$ from the entire dataset without including the
-test for this test, so again our results represent a lower bound for this
-method.
+the empirical $C_{\theta,\theta'}$ from the entire dataset and do not add its
+computation time when considering runtime, so again our results represent a
+lower bound for this method.
 

### 2016-10-13 aistats2017.tex
updates

@@
 implement this heuristic and instead set the individual test error
-equal to the overall test error. Thus our results for~\citet{cutting_mh_2014} should be treated as lower bounds. All
-methods were initialized with minibatch size 100.  For~\citet{cutting_mh_2014} we
-increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
-use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
-samples using a random walk proposer with covariance ${\rm diag}(0.15,
-0.15)$, which means the M-H test is responsible for shaping the sample
-distribution.
+equal to the overall test error. Thus our results
+for~\citet{cutting_mh_2014} should be treated as lower bounds. All
+methods were initialized with minibatch size 100.
+For~\citet{cutting_mh_2014} we increment sizes by 100 and set the
+tolerance $\epsilon=0.005$ to control overall test error.
+For~\citet{icml2014c1_bardenet14}, we increase sizes geometrically
+with $\gamma = 1.5$ and use parameters $p = 2, \delta = 0.01$.  All
+methods collect 5000 samples using a random walk proposer with
+covariance ${\rm diag}(0.15, 0.15)$, which means the M-H test is
+responsible for shaping the sample distribution.
 
@@
 set at $T=1000$. We run the three tests for 3000 samples and again use 100
-as the starting minibatch size. Method~\citet{cutting_mh_2014} used
+as the starting minibatch size. Our undiscounted implementation of~\citet{cutting_mh_2014} used
 per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
@@
 three methods on a log-log scale. With an initial size of 100, we achieve an
-average minibatch size of 393, far smaller than the averages of the other two
-methods, which have longer-tailed distributions.
+average minibatch size of 393, about two orders of magnitude smaller than
+~\citet{icml2014c1_bardenet14}, and an order of magnitude faster than
+the undiscounted implementation of~\citet{cutting_mh_2014}.
 

### 2017-01-09 arXiv.tex
added arxiv.tex back

--- 2016-10-09 paper_versions/nips_2016.tex
@@
 
-% if you need to pass options to natbib, use, e.g.: \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading nips_2016
-% to avoid loading the natbib package, add option nonatbib: \usepackage[nonatbib]{nips_2016}
-
-\PassOptionsToPackage{numbers, compress}{natbib}
-%\usepackage{nips_2016}
-% to compile a camera-ready version, add the [final] option, e.g.:
 \usepackage[final]{nips_2016}
 
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{amssymb}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{graphicx}
+% If your paper is accepted, change the options for the package
+% aistats2017 as follows:
+%
+%\usepackage[accepted]{aistats2017}
+%
+% This option will print headings for the title of your paper and
+% headings for the authors names, plus a copyright note at the end of
+% the first column of the first page.
+\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem,float}
+\usepackage{amssymb}
+\usepackage{natbib}  % for references
+\usepackage{amsfonts}
+\usepackage{nicefrac}  
+\usepackage{booktabs}
+\usepackage{microtype}
+\usepackage[]{algorithm2e}
 \usepackage{caption}
 \usepackage{subcaption}
-
-\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem}
-\usepackage[]{algorithm2e}
 \graphicspath{{figures/}}
-
+\usepackage{setspace}
 \lstset{
@@
 
+% This portion of style definition will remove the index number before the 
+% without inducing an indent
+\makeatletter
+\renewcommand\@biblabel[1]{}
+\renewenvironment{thebibliography}[1]
+     {\section*{\refname}%
+      \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
+      \list{}%
+           {\leftmargin0pt
+            \@openbib@code
+            \usecounter{enumiv}}%
+      \sloppy
+      \clubpenalty4000
+      \@clubpenalty \clubpenalty
+      \widowpenalty4000%
+      \sfcode`\.\@m}
+     {\def\@noitemerr
+       {\@latex@warning{Empty `thebibliography' environment}}%
+      \endlist}
+\makeatother
+
+\begin{document}
+
+% If your paper is accepted and the title of your paper is very long,
+% the style will print as headings an error message. Use the following
+% command to supply a shorter title of your paper so that it can be
+% used as headings.
+%
+%\runningtitle{I use this title instead because the last one was very long}
+
+% If your paper is accepted and the number of authors is large, the
+% style will print as headings an error message. Use the following
+% command to supply a shorter version of the authors names so that
+% they can be used as headings (for example, use only the surnames)
+%
+%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
+
 \title{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
 
-% The \author macro works with any number of authors. There are two commands used to separate the
-% names and addresses of multiple authors: \And and \AND.
-% Using \And between authors leaves it to LaTeX to determine where to break the lines. Using \AND
-% forces a line break at that point. So, if LaTeX puts 3 of 4 authors names on the first line, and
-% the last on the second line, try using \AND instead of \And before the third author name.
-
-%%  David S.~Hippocampus\thanks{Use footnote for providing further
-%%    information about author (webpage, alternative
-%%    address)---\emph{not} for acknowledging funding agencies.} \\
-%%  Department of Computer Science\\
-%%  Cranberry-Lemon University\\
-%%  Pittsburgh, PA 15213 \\
-%%  \texttt{hippo@cs.cranberry-lemon.edu} \\
-
-\author{
-  Haoyu Chen\textsuperscript{*}, Daniel Seita\textsuperscript{*},
-  Xinlei Pan\textsuperscript{*}, John Canny\textsuperscript{*}$^{\dag}$ \\
-  * University of California, Berkeley, CA \\
-$\dag$ Google Research, Mountain View, CA\\
-  \texttt{\{haoyuchen,seita,xinleipan,canny\}@berkeley.edu}
- }
-
-\begin{document}
+\author{ Anonymous Author 1 \And Anonymous Author 2 \And Anonymous Author 3 }
+
+% \address{ Unknown Institution 1 \And Unknown Institution 2 \And Unknown Institution 3 } ]
+
 \maketitle
@@
 \begin{abstract}
+  %The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
+  %both left and right-hand margins. Use 10~point type, with a vertical
+  %spacing of 11~points. The {\bf Abstract} heading must be centered,
+  %bold, and in point size 12. Two line spaces precede the
+  %Abstract. The Abstract must be limited to one paragraph.
   Markov chain Monte Carlo (MCMC) methods have many applications in
@@
 Minibatch updates produce many improvements to the model for each pass
-over the dataset, and have high sample efficiency. They also map very
-well onto hardware such as GPUs. In contrast, M-H requires
-calculations over the full dataset to produce a new sample.  Recent
-results from~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform
+over the dataset, and have high sample efficiency. %They also map very
+% well onto hardware such as GPUs. 
+In contrast, M-H requires calculations over the full dataset to produce a new sample.  Recent
+results from~\citet{cutting_mh_2014} and ~\citet{icml2014c1_bardenet14} perform
 approximate (bounded error) acceptance tests using subsets
@@
 proposed sample, and on the random variable $u$. By
-contrast,~\cite{conf/uai/MaclaurinA14,TallData15} perform exact
+contrast,~\citet{conf/uai/MaclaurinA14,TallData15} perform exact
 tests but require a lower bound on parameter distribution across its
@@
 bound, and such bounds are only available for relatively simple
-distributions. By contrast our approach is ``black box'' and requires
-only statistics gathered from the minibatches as they are used.
-
+distributions. %By contrast our approach is a ``black box'' and requires
+%only statistics gathered from the minibatches as they are used. 
 Here we derive a new test which incorporates the variability in minibatch
@@
 value.
-%% {\color{blue} Daniel: we really need to show the amount of data is fixed
-%% with ``high probability'' (the results suggest ``OK probability'').} 
-%% Analysis show data amount is gaussian and short-tailed and therefore bounded with high p
-%% experiments back this up. 
-We use a Barker test function~\cite{Barker65} rather than a Metropolis
+
+We use a Barker test function~\citep{Barker65} rather than a Metropolis
 test, which makes our test naturally error tolerant. The idea of using
-a noise-tolerant test using Barker's test function was suggested but
-not explored empirically in~\cite{TallData15} section 6.3. But the
+a noise-tolerant %test using
+Barker's test function was suggested but
+not explored empirically in~\citet{TallData15} section 6.3. But the
 asymptotic test statistic CDF and the Barker function are different,
 which leads to fixed errors for the approach
-in~\cite{TallData15}. Here we show that the difference between the
+in~\citet{TallData15}. Here we show that the difference between the
 distributions can be corrected with an additive random variable. This
@@
 
-\begin{itemize}
-
+\begin{itemize}[noitemsep]
 \item Increase the temperature of the target distribution. Log likelihoods
@@
 Bayesian inference task is asymptotically normal, and has variance
-that scales inversely with the number of data samples $N$. This mode
-is extremely sharp for large datasets, which may contain millions or
-billions of samples. Simply sampling from this distribution is one
-application, but an efficient proposal distribution~\cite{OptimalScaling01} has similar variance to the target
+that scales inversely with the number of data samples $N$. %This mode
+%is extremely sharp for large datasets, which may contain millions or
+%billions of samples. 
+Simply sampling from this distribution is one
+application, but an efficient proposal distribution~\citep{OptimalScaling01} has similar variance to the target
 distribution and will diffuse to it extremely slowly from an
@@
 progress to a posterior mode with each step, taking small steps so
-that the cumulative displacement has progressively lower variance. A
-substantial part of the computational work of MCMC on massive datasets
-will be similarly in reaching a stationary distribution, which really
-means finding a deep posterior mode. Taking noisy small steps will
-nevertheless make steady progress to a posterior mode since their bias
-is in that direction. We demonstrate this behavior in our logistic regression
-experiments.
+that the cumulative displacement has progressively lower variance. %A
+%substantial part of the computational work of MCMC on massive datasets
+%will be similarly in reaching a stationary distribution, which really
+%means finding a deep posterior mode. 
+Taking noisy small steps will nevertheless make steady progress to 
+a posterior mode since their bias is in that direction. We demonstrate 
+this behavior in our logistic regression experiments.
 
 The contributions of this paper are as follows:
-
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item We develop a new, more efficient (in samples per test) minibatch acceptance test with quantifiable error bounds. The test
   uses a novel additive correction variable to implement a Barker test based on minibatch mean and variance. 
-\item We analyze the test for accuracy and speed.
+%\item We analyze the test for accuracy and speed.
 \item We compared performance of our new test and prior approaches on several datasets. We demonstrate the test is several orders
@@
 training data points, but this is very expensive for large datasets. To address
-this challenge,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
+this challenge,~\citet{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
 Metropolis-Hasting tests using sequential hypothesis testing. During each
-iteration, they start with a small minibatch of data and test the hypothesis
-that the sample $\theta'$ should be accepted based on an approximate version of
+iteration, they start with a small minibatch of data and test whether the sample $\theta'$ 
+should be accepted based on an approximate version of
 the test $u < \alpha(\theta,\theta')$. If the approximate test cannot make a
@@
 
-Following~\cite{icml2014c1_bardenet14}, we write the test $u <
+Following~\citet{icml2014c1_bardenet14}, we write the test $u <
 \alpha(\theta,\theta')$ in the equivalent form $\Lambda(\theta,\theta') >
 \psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those in
-\cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later.}
+\citet{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later.}
 \begin{equation}
-\Lambda(\theta,\theta') = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
-\hspace{0.2in} {\rm and} \hspace{0.2in}
-\psi(u,\theta,\theta') = \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right)
+\begin{split} 
+\Lambda(\theta,\theta') & = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right) \\  
+{\rm and} \quad \psi(u,\theta,\theta') &= \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right) \\
+\end{split}
 \end{equation}
@@
 \begin{equation}\label{eq:lemma_ll_ratio}
-\Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)}=
-  N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
+\resizebox{\linewidth}{!}{% 
+$\Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)} = N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)$
+}%
 \end{equation}
@@
 test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
-random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{TallData15} it
+random variable $X$ (call it $X_{\rm log}$ from now on). In~\citet{TallData15} it
 was proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed
@@
 \section{Computing the Correction Distribution}\label{sec:correction}
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
-    \caption{
-    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
-    convolve to form the standard logistic distribution. These were obtained
-    based on our original linear programming formulation, as discussed in
-    Section~\ref{sec:correction}. We use three standard deviation values of
-    $X_{\rm norm}$. The two red curves convolve to form the logistic, etc. The
-    $y$-axis is capped at 0.5 for readability. This figure must be viewed in
-    color.
-    }
-    \label{fig:deconvolution}
-    %\vspace{-10pt}
-\end{figure}
 
@@
 \begin{equation}\label{eq:conv1}
-    C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|
+    \small{C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|}
 \end{equation}
 To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
-and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}$ and $j \in \{-N,
-\ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
-condition in Equation~(\ref{eq:conv1}) can be re-expressed as:
+and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}=\mathcal{I}$ and
+$j \in \{-N, \ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$,
+and the condition in Equation~(\ref{eq:conv1}) can be re-expressed as:
 \begin{equation}\label{eq:conv2}
-    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
+    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \mathcal{I}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
 \end{equation}
@@
 programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
-Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
-$X_{\rm corr}$ densities that sum to $X_{\rm log}$ from this process.
 
@@
 \begin{proof}
-The following bound is given immediately after Corollary 2 from~\cite{explicit-clt05}:
+The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
 \begin{equation}\label{eq:clt-boundsx}
--6.4E|X|^3-2E|X| \leq    \sup_x|{\rm Pr}(t<x) - \Phi(x)|{\sqrt{n}} \leq 1.36E|X|^3
+\begin{split}
+-6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
+&\leq 1.36E|X|^3.
+\end{split}
 \end{equation}
@@
 \end{equation}
-and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
+and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
 \begin{equation}
--\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon
+\begin{split}
+-\epsilon &= \tiny{\int_{-\infty}^{+\infty} -\epsilon R(x) dx } \\
+&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
+&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon.
+\end{split}
 \end{equation}
 \end{proof}
+\vspace{-10pt}
 
@@
 times to collect 5000 samples.
-
-\begin{figure}[t]
+\begin{figure}
     \centering
-    \includegraphics[width=1\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
+ %   \includegraphics[width=\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
     \caption{
@@
 \subsection{Logistic Regression}\label{ssec:logistic}
-
-\begin{figure}[t]
+\begin{figure}
 	\centering
-	\includegraphics[width=1\linewidth]{logistic_performance.png}
+	\includegraphics[width=\linewidth]{logistic_regression_ll_acc_results.png}
 	\caption{
@@
 	\label{fig:logistic_performance}
-\end{figure}
-
-\begin{figure}[t]
+%\end{figure*}
+%\begin{figure*}[t!]
 	\centering
-	\includegraphics[width=1\linewidth]{minibatch_size_logistic.png}
+	\includegraphics[width=\linewidth]{logistic_regression_histograms_results.png}
 	\caption{
@@
 
-We next use logistic regression for the binary classification of 1s versus 7s in
-the MNIST dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007
-training and 1000 testing elements (we used a random subset of the test data).
-The pixel values are scaled to be in $(0,1)$.  We impose a flat prior on
-$\theta$. For the proposer, we again use a random walk, this time with
-covariance matrix $0.05I$ for the $784\times 784$ identity matrix $I$. The
-posterior temperature is set at $T=1000$.  We run our MH test and compare with
-Korattikara 2014 and Bardenet 2014; minibatch sizes and other parameters are set
-to be the same value as with the Gaussian Mixtures experiment from
-Section~\ref{ssec:gaussians}.
-
-Figure~\ref{fig:logistic_performance} shows the training log likelihood and the
-testing prediction accuracy, each as a function of the cumulative training data
-points processed.\footnote{Note that the curves do not span the same length over
-the x-axis, because our test consumes fewer samples throughout the MCMC
-procedure, so the corresponding curve will ``end'' before the other two.} To
-generate the curves, for each of the 5000 sampled vectors $\theta_t$,
-$t\in\{1,\ldots,5000\}$, we use $\theta_t$ as the parameter for logistic
-regression.  We see that our minibatch MH test is more efficient compared to the
-other two methods, because it achieves similar or superior performance while
-consuming fewer data points.
-
-Figure~\ref{fig:logistic_minibatch}, in a similar manner as
-Figure~\ref{fig:gauss_mix_2}, shows the histogram of minibatch sizes for all
-three methods on a log-log scale. With an initial minibatch size of 100,
-Korattikara 2014 and Bardenet 2014 achieve average minibatch size of 2026 and
-7612 respectively over the MNIST classification task, while our method achieves
-that of 163, showing significantly better performance than the benchmark test
-methods.
+We next test logistic regression for the binary classification of 1s
+versus 7s on a subset of the MNIST8M dataset, a larger version of
+MNIST~\citep{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
+450k training and 192k testing points.
+We impose a uniform prior on $\theta$ and again use a random walk
+proposer, this time with covariance matrix $0.05I$. The temperature is
+set at $T=1000$. We run the three tests for 3000 samples and again use 100
+as the starting minibatch size. Our undiscounted implementation of~\citet{cutting_mh_2014} used
+per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
+we use their recommended discounting factor value of $\gamma=2.0$. We tried to use the
+symbolic bound for $C_{\theta,\theta'}$ for Logistic regression
+from~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead we use
+the empirical $C_{\theta,\theta'}$ from the entire dataset and do not add its
+computation time when considering runtime, so again our results represent a
+lower bound for this method.
+
+Figure~\ref{fig:logistic_performance} shows the test log likelihood and
+prediction accuracy as a function of the cumulative training points
+processed.\footnote{The curves do not span the same length over the x-axis since
+the methods consume different amounts of data.} To generate the curves, for each
+of the sampled vectors $\theta_t$, $t\in\{1,\ldots,3000\}$, we use $\theta_t$ as
+the parameter for logistic regression.  Our minibatch MH test is more efficient
+in terms of accuracy, achieving convergence roughly twice as fast compared
+to~\citet{cutting_mh_2014} and about 1.5 orders of magnitude faster
+than~\citet{icml2014c1_bardenet14}, though its log likelihood results lag
+slightly behind the former algorithm during the very early stages of
+exploration.
+
+Figure~\ref{fig:logistic_minibatch} shows histograms of minibatch sizes for the
+three methods on a log-log scale. With an initial size of 100, we achieve an
+average minibatch size of 393, more than two orders of magnitude smaller
+than~\citet{icml2014c1_bardenet14}, and almost an order of magnitude faster than the
+undiscounted implementation of~\citet{cutting_mh_2014}.
 

### 2017-02-05 icml2017/paper.tex
Whew, update to get it in ICML 2017 format.
For better organization, the most up to date version is now in icml2017 directory.

--- 2016-10-13 aistats2017.tex
@@
-\documentclass[twoside]{article} \usepackage{aistats2017}
-
-% If your paper is accepted, change the options for the package
-% aistats2017 as follows:
-%
-%\usepackage[accepted]{aistats2017}
-%
-% This option will print headings for the title of your paper and
-% headings for the authors names, plus a copyright note at the end of
-% the first column of the first page.
-\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem,float}
-\usepackage{amssymb}
-\usepackage{natbib}  % for references
-\usepackage{amsfonts}
-\usepackage{nicefrac}  
-\usepackage{booktabs}
-\usepackage{microtype}
-\usepackage[]{algorithm2e}
-\usepackage{caption}
-\usepackage{subcaption}
-\graphicspath{{figures/}}
-\usepackage{setspace}
-\lstset{
-numbers=left, 
-numberstyle=\small, 
-numbersep=8pt, 
-frame = single, 
-language=matlab, 
-framexleftmargin=20pt}
-
-\renewcommand{\refname}{}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%% ICML 2017 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass{article}
+\usepackage{times,graphicx,subfigure,natbib,algorithm,algorithmic,hyperref}
+\usepackage{amsmath,amsthm,amssymb,enumitem,verbatim}
+% Packages hyperref and algorithmic misbehave sometimes.  We can fix
+% this with the following command.
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
 \newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
+\newcommand{\mE}{\mathbb{E}}
 \DeclareMathOperator*{\argmin}{arg\,min}
@@
 \newtheorem{corollary}{Corollary}
-
-% This portion of style definition will remove the index number before the 
-% without inducing an indent
-\makeatletter
-\renewcommand\@biblabel[1]{}
-\renewenvironment{thebibliography}[1]
-     {\section*{\refname}%
-      \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
-      \list{}%
-           {\leftmargin0pt
-            \@openbib@code
-            \usecounter{enumiv}}%
-      \sloppy
-      \clubpenalty4000
-      \@clubpenalty \clubpenalty
-      \widowpenalty4000%
-      \sfcode`\.\@m}
-     {\def\@noitemerr
-       {\@latex@warning{Empty `thebibliography' environment}}%
-      \endlist}
-\makeatother
+\graphicspath{{figures/}{../figures/}}
+
+% Employ the following version of the ``usepackage'' statement for
+% submitting the draft version of the paper for review.  This will set
+% the note in the first column to ``Under review.  Do not distribute.''
+\usepackage{icml2017}
+
+% Employ this version of the ``usepackage'' statement after the paper has
+% been accepted, when creating the final version.  This will set the
+% note in the first column to ``Proceedings of the...''
+%\usepackage[accepted]{icml2017}
+
+% The \icmltitle you define below is probably too long as a header.
+% Therefore, a short form for the running title is supplied here:
+\icmltitlerunning{Submission and Formatting Instructions for ICML 2017}
 
@@
 
-% If your paper is accepted and the title of your paper is very long,
-% the style will print as headings an error message. Use the following
-% command to supply a shorter title of your paper so that it can be
-% used as headings.
-%
-%\runningtitle{I use this title instead because the last one was very long}
-
-% If your paper is accepted and the number of authors is large, the
-% style will print as headings an error message. Use the following
-% command to supply a shorter version of the authors names so that
-% they can be used as headings (for example, use only the surnames)
-%
-%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
-
 \twocolumn[
-
-\aistatstitle{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
-
-\aistatsauthor{ Anonymous Author 1 \And Anonymous Author 2 \And Anonymous Author 3 }
-
-\aistatsaddress{ Unknown Institution 1 \And Unknown Institution 2 \And Unknown Institution 3 } ]
+\icmltitle{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
+
+% It is OKAY to include author information, even for blind submissions: the
+% style file will automatically remove it for you unless you've provided the
+% [accepted] option to the icml2017 package.
+
+% list of affiliations. the first argument should be a (short) identifier you
+% will use later to specify author affiliations
+% Academic affiliations should list Department, University, City, Region, Country
+% Industry affiliations should list Company, City, Region, Country
+
+% you can specify symbols, otherwise they are numbered in order
+% ideally, you should not use this facility. affiliations will be numbered
+% in order of appearance and this is the preferred way.
+% \icmlsetsymbol{equal}{*}
+\begin{icmlauthorlist}
+\icmlauthor{Daniel Seita}{berkeley}
+\icmlauthor{Xinlei Pan}{berkeley}
+\icmlauthor{Haoyu Chen}{berkeley}
+\icmlauthor{John Canny}{berkeley,google}
+\end{icmlauthorlist}
+\icmlaffiliation{berkeley}{University of California, Berkeley, CA}
+\icmlaffiliation{google}{Google Research, Mountain View, CA}
+\icmlcorrespondingauthor{Daniel Seita}{seita@berkeley.edu}
+
+% You may provide any keywords that you 
+% find helpful for describing your paper; these are used to populate 
+% the "keywords" metadata in the PDF but will not be shown in the document
+\icmlkeywords{boring formatting information, machine learning, ICML}
+
+\vskip 0.3in
+]
+
+% this must go after the closing bracket ] following \twocolumn[ ...
+
+% This command actually creates the footnote in the first column
+% listing the affiliations and the copyright notice.
+% The command takes one argument, which is text to display at the start of the footnote.
+% The \icmlEqualContribution command is standard text for equal contribution.
+% Remove it (just {}) if you do not need this facility.
+
+\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+%\footnotetext{hi}
 
 \begin{abstract}
-  %The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
-  %both left and right-hand margins. Use 10~point type, with a vertical
-  %spacing of 11~points. The {\bf Abstract} heading must be centered,
-  %bold, and in point size 12. Two line spaces precede the
-  %Abstract. The Abstract must be limited to one paragraph.
-  We
-  present a novel Metropolis-Hastings method for large datasets that
+  We present a novel Metropolis-Hastings method for large datasets that
   uses small expected-size minibatches of data. Previous work on
@@
 
-\section{INTRODUCTION}\label{sec:introduction}
+\section{Introduction}\label{sec:introduction}
 
@@
 intractable distributions. We are interested in large dataset applications,
-where the goal is to sample a posterior distribution $p(\theta \mid x_1, \ldots,
+where the goal is to sample a posterior distribution $p(\theta | x_1, \ldots,
 x_N)$ of parameter $\theta$ for large $N$.  The Metropolis-Hastings method (M-H)
@@
 reject based on an acceptance test. The acceptance test is usually a Metropolis
-test~\citep{Metropolis1953, hastings70}.
+test~\cite{Metropolis1953, hastings70}.
 
@@
 over the full dataset to produce a new sample.  Recent results
-from~\citet{cutting_mh_2014} and~\citet{icml2014c1_bardenet14} perform
+from~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14} perform
 approximate (bounded error) acceptance tests using subsets of the full dataset.
 The amount of data consumed for each test varies significantly from one
-minibatch to the next. By contrast,~\citet{conf/uai/MaclaurinA14,TallData15}
+minibatch to the next. By contrast,~\cite{conf/uai/MaclaurinA14,TallData15}
 perform exact tests but require a lower bound on the parameter distribution across
@@
 statistics as {\em a natural part of the test} and requires less data per
-iteration than prior work. We use a Barker test function~\citep{Barker65}, which
+iteration than prior work. We use a Barker test function~\cite{Barker65}, which
 makes our test naturally error tolerant. The idea of using a noise-tolerant
 Barker's test function was suggested but not explored empirically
-in~\citet{TallData15} section 6.3. But the asymptotic test statistic CDF and the
+in~\cite{TallData15} section 6.3. But the asymptotic test statistic CDF and the
 Barker function are different, which leads to fixed errors for the approach
-in~\citet{TallData15}. Here, we show that the difference between the
+in~\cite{TallData15}. Here, we show that the difference between the
 distributions can be corrected with an additive random variable. This leads to a
@@
 well-specified models running Metropolis-Hastings sampling with optimal
-proposals~\citep{OptimalScaling01} on a full dataset. When we reduce the amount
+proposals~\cite{OptimalScaling01} on a full dataset. When we reduce the amount
 of data for the test, the variance goes up. We have to reduce variance in one
@@
 sampling from it is one application, but an efficient
-proposal~\citep{OptimalScaling01} has similar variance to the target and will
+proposal~\cite{OptimalScaling01} has similar variance to the target and will
 diffuse to it extremely slowly. For applications to neural networks or models
-where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
+where the posterior is multimodal~\cite{choromanska2014loss}, samplers will
 likely get trapped in one of the modes. A common solution is to anneal the
@@
 
-% Daniel: I merged these two paragraphs into one, because these were too long,
-% and we also never experiment with the step size. Let me know what you think.
-
-%% It is worth discussing at this point the typical goals of M-H sampling on large
-%% datasets.  By the Bernstein-von Mises Theorem, the posterior distribution for a
-%% Bayesian inference task is asymptotically normal, and has variance that scales
-%% inversely with the $N$ data samples. Simply sampling from this distribution is
-%% one application, but an efficient proposal~\citep{OptimalScaling01} has similar
-%% variance to the target distribution and will diffuse to it extremely slowly from
-%% an initialization value which is (likely to be) many standard deviations away.
-%% If there are any other strong modes, it is very likely for the sampler to find
-%% one of them and become trapped in it when run at the normal distribution
-%% temperature ($T=1$). A common solution is to anneal the sampler, running first
-%% at high temperature (scaling log likelihoods by $1/T$) which flattens the
-%% likelihood landscape.  This in turn reduces the variance of the log acceptance
-%% probability and allows our acceptance test to be applied.
-%% 
-%% A second question concerns step size. Once we have fixed temperature, our
-%% variance constraint implies that we have to trade-off proposal step size $s$ and
-%% batch size $b$ ($b \propto p^2$), i.e. we can make many small steps, or one
-%% large step, with a given batch of data. One of the primary drivers of this work
-%% is our belief in the value of small steps. For applications to neural networks
-%% or other models where the posterior is multimodal, posterior inference is
-%% arguably a search process. Covering the search space densely with small steps is
-%% much more valuable than few sparse steps toward the nearest optimum. In this
-%% mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
-%% Descent. The goal in SGD is to make gradual progress to a posterior mode with
-%% each step, taking small steps so that the cumulative displacement has
-%% progressively lower variance. We demonstrate this behavior in our logistic
-%% regression experiments.
-
 The contributions of this paper are as follows:
@@
 
-    %\item We analyze the test for accuracy and speed.
-
     \item We compare performance of our new test and prior approaches on several
@@
 
-\section{PRELIMINARIES}\label{sec:related_work}
-
-In the Metropolis-Hastings method~\citep{gilks1996markov,brooks2011handbook}, a
+\section{Preliminaries}\label{sec:related_work}
+
+In the Metropolis-Hastings method~\cite{gilks1996markov,brooks2011handbook}, a
 difficult-to-compute probability distribution $p(\theta)$ is sampled using a
@@
 is generated using a candidate $\theta'$ from a (simpler) proposal distribution
-$q(\theta'\mid \theta_t)$, filtered by an acceptance test. The acceptance test
+$q(\theta'|\theta_t)$, filtered by an acceptance test. The acceptance test
 is usually a Metropolis test. The Metropolis test has acceptance probability:
 \begin{equation}\label{eq:traditional}
-    \alpha(\theta_t,\theta') = \frac{p(\theta')q(\theta_t \mid \theta')}{p(\theta_t)q(\theta' \mid \theta_t)} \wedge 1
+    \alpha(\theta_t,\theta') = \frac{p(\theta')q(\theta_t | \theta')}{p(\theta_t)q(\theta' | \theta_t)} \wedge 1
 \end{equation}
@@
 The acceptance test guarantees detailed balance, which means
-$p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')$, where
-$p(\theta'\mid\theta)$ is the probability of a transition from state $\theta$ to
-$\theta'$. Here, $p(\theta'\mid\theta)=q(\theta'\mid\theta)\alpha(\theta,\theta')$.
-% Daniel: I commented this out because people should know what it means and we
-%re-state another detailed balance equation later!
-%\begin{equation}\label{eq:detailed_balance2}
-%    p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
-%\end{equation}
+$p(\theta)p(\theta'|\theta) = p(\theta')p(\theta|\theta')$, where
+$p(\theta'|\theta)$ is the probability of a transition from state $\theta$ to
+$\theta'$. Here, $p(\theta'|\theta)=q(\theta'|\theta)\alpha(\theta,\theta')$.
 This condition, together with ergodicity, guarantees that the Markov chain has a
 unique stationary distribution $\pi(\theta) = p(\theta)$. For Bayesian
-inference, the target distribution is $p(\theta \mid x_1, \ldots, x_N)$. The
+inference, the target distribution is $p(\theta | x_1, \ldots, x_N)$. The
 acceptance probability is now:
@@
     \alpha(\theta,\theta') = 
-    \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
-    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)}
+    \frac{p_0(\theta')\prod_{i=1}^N p(x_i | \theta')q(\theta |
+    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i | \theta)q(\theta' | \theta)}
     \wedge 1
@@
 data points, but this is very expensive for large datasets. To address this
-challenge,~\citet{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
+challenge,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
 Metropolis-Hasting tests using sequential hypothesis testing. Each iteration,
@@
 the test repeats. This process continues until a decision. The bounds depend on
-either an asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
-concentration bound~\citep{icml2014c1_bardenet14}. The latter requires direct
+either an asymptotic Central Limit Theorem~\cite{cutting_mh_2014} or a
+concentration bound~\cite{icml2014c1_bardenet14}. The latter requires direct
 bounds on the log likelihood ratio, which for general distributions requires
-knowing $p(x_i \mid \theta)$ and $p(x_i \mid \theta')$ for all $N$ samples. In
+knowing $p(x_i | \theta)$ and $p(x_i | \theta')$ for all $N$ samples. In
 addition, while both methods show useful reductions in the number of samples
@@
 
-Following~\citet{icml2014c1_bardenet14}, we write the test
+Following~\cite{icml2014c1_bardenet14}, we write the test
 $u<\alpha(\theta,\theta')$ equivalently as $\Lambda(\theta,\theta') >
 \psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those
-in~\citet{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis
+in~\cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis
 later.}
@@
 \begin{split}
-\Lambda(\theta,\theta') & = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right) \\  
-{\rm and} \quad \psi(u,\theta,\theta') &= \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right).
+\Lambda(\theta,\theta') = \sum_{i=1}^N \log\frac{p(x_i|\theta')}{p(x_i|\theta)}, \\
+\psi(u,\theta,\theta') = \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right).
 \end{split}
@@
 To reduce computational effort, an unbiased estimate of $\Lambda(\theta,\theta')$
-based on a minibatch can be used:
+based on a minibatch $\{x_1^*,\ldots,x_b^*\}$ can be used:
 \begin{equation}
-\Lambda^*(\theta,\theta') = \frac{N}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
+\Lambda^*(\theta,\theta') = \frac{N}{b}\sum_{i=1}^b 
+\log \frac{p(x_i^*|\theta')}{p(x_i^*|\theta)} 
 \end{equation}
@@
 $\Lambda_i(\theta,\theta') = N\log(\frac{p(x_i|\theta')}{p(x_i|\theta)})$.
-%% \begin{equation}\label{eq:individual_terms}
-%% \Lambda_i(\theta,\theta') = N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
-%% \end{equation}
 Thus, $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over
 the entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean of
-$\Lambda_i(\theta,\theta')$ over its minibatch.
-
-Since minibatches contains randomly selected samples $x_i$, the values
-$\Lambda_i$ are i.i.d. random variables.\footnote{The analysis assumes sampling
-with replacement although implementations on typical large datasets will
-approximate this by sampling without replacement.} By the Central Limit
-Theorem, we expect $\Lambda^*(\theta,\theta')$ to be approximately Gaussian. The
-acceptance test then becomes a statistical test of the hypothesis that
+the $\Lambda_i(\theta,\theta')$ in its minibatch.
+
+Since minibatches contains randomly selected samples, the values $\Lambda_i$ are
+i.i.d. random variables.\footnote{The analysis assumes sampling with replacement
+although implementations on typical large datasets will approximate this by
+sampling without replacement.} By the Central Limit Theorem, we expect
+$\Lambda^*(\theta,\theta')$ to be approximately Gaussian. The acceptance test
+then becomes a statistical test of the hypothesis that
 $\Lambda(\theta,\theta')>\psi(u,\theta,\theta')$ by establishing that
@@
 \begin{equation}\label{eq:lemma_ll_ratio}
-    \Lambda^*(\theta,\theta') = N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
-\end{equation}
-which is normally distributed over selection of the Normal samples $x_i$.  Since
-the $x_i$ have unit variance, their mean has variance $1/b$, and the variance of
-$\Lambda^*(\theta,\theta')$ is $\sigma^2(\Lambda^*) = (\theta'-\theta)^2N^2/b$.
-In order to pass a hypothesis test that $\Lambda > \psi$, there needs to be a
-large enough gap (several $\sigma(\Lambda^*)$) between
-$\Lambda^*(\theta,\theta')$ and $\psi(u,\theta,\theta')$.
+    \Lambda^*(\theta,\theta') = N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b
+    x_i^*-\theta-\frac{\theta'-\theta}{2}\right)
+\end{equation}
+which is normally distributed over selection of the Normal samples $x_i^*$.
+Since the $x_i^*$ have unit variance, their mean has variance $1/b$, and the
+variance of $\Lambda^*(\theta,\theta')$ is $\sigma^2(\Lambda^*) =
+(\theta'-\theta)^2N^2/b$.  In order to pass a hypothesis test that $\Lambda >
+\psi$, there needs to be a large enough gap (several $\sigma(\Lambda^*)$)
+between $\Lambda^*(\theta,\theta')$ and $\psi(u,\theta,\theta')$.
 
@@
 distribution has the same variance as the target
-distribution~\citep{OptimalScaling01}, so we use a proposal based on
+distribution~\cite{OptimalScaling01}, so we use a proposal based on
 $\mathcal{N}(\theta,1/N)$. It is symmetric
-$q(\theta'\mid\theta)=q(\theta\mid\theta')$, and since we assumed a uniform
-prior, $\psi(u,\theta,\theta')=\log u$. Our worst-case scenario is specified in
+$q(\theta'|\theta)=q(\theta|\theta')$, and since we assumed a uniform prior,
+$\psi(u,\theta,\theta')=\log u$. Our worst-case scenario is specified in
 Lemma~\ref{lem:worst_case}.
@@
 \begin{proof}
-See Supplementary Material, Section~\ref{app:worst_case_proof}.
+See Appendix, Section~\ref{app:worst_case_proof}.
 \end{proof}
@@
 physics. One can apply
-Hamiltonian Monte Carlo (HMC)~\citep{mcmc_hamiltonian_2010} methods which
+Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which
 generate high acceptance \emph{and} distant proposals when run on full batches
-of data. Recently Langevin Dynamics~\citep{langevin_2011,conf/icml/AhnBW12} has
+of data. Recently Langevin Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has
 been applied to Bayesian estimation on minibatches of data. This simplified
@@
 step size is severe, and the state space exploration reduces to a random walk.
-Full minibatch HMC for minibatches was described in~\citet{sghmc_2014} which
+Full minibatch HMC for minibatches was described in~\cite{sghmc_2014} which
 allows momentum-augmented proposals with larger step sizes. However, step sizes
@@
 
-\section{A NEW MH ACCEPTANCE TEST}\label{sec:our_algorithm}
+\section{A New MH Acceptance Test}\label{sec:our_algorithm}
 
@@
     \Delta(\theta,\theta')  =
-    \log \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
-    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)},
+    \log \frac{p_0(\theta')\prod_{i=1}^N p(x_i | \theta')q(\theta |
+    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i | \theta)q(\theta' | \theta)},
 \end{equation}
@@
 Equation~(\ref{eq:acceptance_probability}). We separate out terms dependent and
-independent of the data $x$ as:
+independent of the data as:
 \begin{equation}\label{eq:delta2}
-\begin{split}
-    \Delta(\theta,\theta') &= \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') \\
-    & = \Lambda(\theta,\theta') - \psi(1,\theta,\theta').
-\end{split}
+\Delta(\theta,\theta') =
+\underbrace{\sum_{i=1}^N\log\frac{p(x_i | \theta')}{p(x_i | \theta)}}_{\Lambda(\theta,\theta')}
+- \psi(1,\theta,\theta').
 \end{equation}
@@
 \begin{equation}\label{eq:delta3}
-\begin{split}
-    \Delta^*(\theta,\theta') &=
-\frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta')\\
-&=\Lambda^*(\theta,\theta') - \psi(1,\theta,\theta')
-\end{split}
+\Delta^*(\theta,\theta') =
+\underbrace{\frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i^* | \theta')}{p(x_i^* | \theta)}}_{\Lambda^*(\theta,\theta')}
+- \psi(1,\theta,\theta').
 \end{equation}
@@
 
-This result is used in~\citet{Barker65} to define the Barker acceptance test.
-%As
-%a sanity check, choosing $g(s) = \exp(s) \wedge 1$ --- a function satisfying the
-%requirement of Lemma~\ref{lem:detailed_balance} --- produces the classical
-%Metropolis acceptance test $\alpha(\theta,\theta') = g(\Delta(\theta,\theta')) =
-%\frac{p(\theta')q(\theta \mid \theta')}{p(\theta)q(\theta' \mid \theta)}\wedge
-%1$.
-%In fact, $g(s) =\exp(s) \wedge 1$ is the optimal acceptance function in
-%terms of acceptance rate, since it accepts with probability 1 for $\Delta > 0$.
+This result is used in~\cite{Barker65} to define the Barker acceptance test.
 
 \subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
-For our new MH test we use the Barker logistic~\citep{Barker65}
+For our new MH test we use the Barker logistic~\cite{Barker65}
 function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
@@
 of $b$ IID terms $\Lambda_i(\theta,\theta')$ of the form
-$N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $b$ increases,
+$N\log\frac{p(x_i^*|\theta')}{p(x_i^*|\theta)}$. As $b$ increases,
 $\Delta^*(\theta,\theta')$ therefore has a distribution which approaches a
@@
 test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
-random variable $X$ (call it $X_{\rm log}$ from now on). In~\citet{TallData15} it
+random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{TallData15} it
 was proposed to use $\Delta^*$ in a Barker test, and tolerate the fixed
@@
 
-% Daniel: Saves space by using one line instead of two. Lame, I know.
-%\section{COMPUTING THE CORRECTION DISTRIBUTION}\label{sec:correction}
-\section{CORRECTION DISTRIBUTION}\label{sec:correction}
+\section{Computing the Correction Distribution}\label{sec:correction}
 
@@
 
-% Daniel: This was taking up too much space, and we usually don't want to spend
-% a paragraph explaining what *didn't* work, and readers will know what it means
-% to minimize the L-infinity norm, we can put it inline.
-% \begin{equation}\label{eq:optimization_l1}
-%     u^* = \argmin_{u} \|Mu-v\|_{\infty}.
-% \end{equation}
-% We have the additional constraint that $u_j > 0$ for all $j$, since $u$
-% represents a density. We first explored optimizing this problem with linear
-% programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
-% 
-% This was tractable up to a few hundred dimensions for $u$.  However, the
-% discretization error is bounded by the curvature of the underlying functions
-% times the squared quantization step. The curvatures are here slightly less than
-% one, i.e.  the errors are of the order of $(V/N)^2$.  Here we chose $V=20$ to
-% provide adequate containment of the distributions (the CDFs are extremely close
-% to either zero or one outside this range). So with 200 points, we have a
-% discretization error of approximately 0.01, which is relatively large.  To yield
-% higher resolution and lower error, we switched to a least squares solution:
-
-\SetKwInOut{Input}{Input}
-\SetKwInOut{Output}{Output}
-\begin{algorithm}[t]
-\Input{Number of samples $T$, minibatch size $m$, error bound $\delta$, pre-computed correction $C_1(X)$ distribution, initial sample $\theta_1$.}
-\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ from $p(\theta)$\;}
-\For{$t=\{1, \ldots, T\}$}{
-    -Propose a candidate $\theta'$\ from proposal $q(\theta'\mid\theta_t)$\;
-    -Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*}$\;
-    -Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Corollary~\ref{cor:our_bound_delta_prime}\;
-    \While {$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}{
-    	-Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $s^2_{\Delta^*}$ and $\epsilon$ estimates\;
-    }
-    -Draw $X_{\rm nc} \sim \mathcal{N}(0,1-s^2_{\Delta^*})$ and $X_{\rm corr}$ from the correction distribution $C_1(X)$\;
-
-        \eIf{$\Delta^* + X_{\rm nc} + X_{\rm corr} > 0$}{
-            -Accept the candidate, $\theta_{t+1} = \theta'$\;
-        }{
-            -Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
-        }
-
-}
+% Daniel: use this algorithm style instead
+\begin{algorithm}[tb]
 \caption{Our acceptance test for MCMC.}
 \label{alg:our_algorithm}
+\begin{algorithmic}
+    \STATE {\bfseries Input:} number of samples $T$, minibatch size $m$, error
+        bound $\delta$, pre-computed correction $C_1(X)$ distribution, initial
+        sample $\theta_1$.
+    \STATE {\bfseries Output:} a chain of $T$ samples
+        $\{\theta_1,\ldots,\theta_T\}$.
+    \FOR{$t=1$ {\bfseries to} $T$}
+        \STATE -Propose a candidate $\theta'$ from proposal
+            $q(\theta' | \theta_t)$.
+        \STATE -Draw a minibatch of $m$ points $\{x_1^*,\ldots,x_m^*\}$.
+        \STATE -Compute $\Delta^*(\theta_t,\theta')$ and sample variance
+            $s^2_{\Delta^*}$.
+        \STATE -Estimate moments $\mE[|\Lambda_i-\Lambda|]$ and
+            $\mE[|\Lambda_i-\Lambda|^3]$ from the sample, and error $\epsilon$
+            from Corollary~\ref{cor:our_bound_delta_prime}.
+        \WHILE{$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}
+            \STATE -Draw $m$ more samples to augment the minibatch, update
+                $\Delta^*$, $s^2_{\Delta^*}$ and $\epsilon$ estimates.
+        \ENDWHILE
+        \STATE -Draw $X_{\rm nc} \sim \mathcal{N}(0,1-s^2_{\Delta^*})$ and
+            $X_{\rm corr} \sim C_1(X)$.
+    \IF{$\Delta^* + X_{\rm nc} + X_{\rm corr} > 0$}
+        \STATE -Accept the candidate, $\theta_{t+1} = \theta'$.
+    \ELSE
+        \STATE -Reject and re-use the old sample, $\theta_{t+1} = \theta_t$.
+    \ENDIF
+    \ENDFOR
+\end{algorithmic}
 \end{algorithm}
 
+% Daniel: also need to modify table, add more info?
 \begin{table}[t]
-    \caption{Error ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$}
+\caption{Error ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$.}
 \label{tab:xcorr}
-    \centering
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
 \begin{tabular}{l l l l}
+\hline
+\abovespace\belowspace
 ${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error} \\
 \hline
+\abovespace
 4000 & 0.9 & 1 & 1.0e-4  \\
-    4000 & 0.8 & 0.03 & 5.0e-6
+\belowspace
+4000 & 0.8 & 0.03 & 5.0e-6 \\
+\hline
 \end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
 \end{table}
@@
     % Daniel: I cut some space to use this:
-    \item The sample variance is denoted as $s^2_{\Delta^*}$ and is proportional
-    to $\|\theta'-\theta\|_2^2$.
+    \item The sample variance of $\Delta^*$ is denoted as $s^2_{\Delta^*}$ and
+    is proportional to $\|\theta'-\theta\|_2^2$.
 
@@
 
-\section{ANALYSIS}\label{sec:analysis}
+\section{Analysis}\label{sec:analysis}
 
 We now derive error bounds for our M-H test and the target distribution it
-generates. In the most similar prior works,~\citet{cutting_mh_2014} use CLT
+generates. In the most similar prior works,~\cite{cutting_mh_2014} use CLT
 asymptotic arguments to show that its approximate acceptance test error tends to
 zero as batch size increases, but no quantitative bounds are given.
-In~\citet{icml2014c1_bardenet14}, explicit bounds are given and depend on
+In~\cite{icml2014c1_bardenet14}, explicit bounds are given and depend on
 \begin{equation}\label{eq:bad_bound}
-    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i\mid\theta') - \log p(x_i\mid\theta)|.
+    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
 \end{equation}
@@
 regression, but it is unclear how to do so for more complicated models, and one
-generally needs to use all $p(x_i\mid \theta')$ terms\footnote{The sample code
-provided by~\citet{icml2014c1_bardenet14} computes $C_{\theta,\theta'}$
+generally needs to use all $p(x_i | \theta')$ terms\footnote{The sample code
+provided by~\cite{icml2014c1_bardenet14} computes $C_{\theta,\theta'}$
 by traversing the entire data, thus providing no performance advantage over the
@@
 \begin{equation}\label{eq:clt-bounds}
-    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}.
+    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4\mE[|X|^3]+2\mE[|X|]}{\sqrt{n}}.
 \end{equation}
@@
 \begin{proof}
-See Supplementary Material, Section~\ref{app:quant_clt}.
+See Appendix, Section~\ref{app:quant_clt}.
 \end{proof}
 
-%% {\color{blue} Daniel: the notation here was (and still is, to a lesser extent)
-%% the source of considerable confusion for me. Do we want $\sigma(X)$ or
-%% $\sigma(\hat{X})$? I think in Lemma~\ref{lem:quant_clt} we want the former, not
-%% the latter. OH ALSO, this section uses $n$ as the full data size, which means
-%% before this section, we should use $n$, not $N$, for the full data size, right?
-%% The Bardenet paper uses $n$ to represent the full data size.
-
-%% JFC - this bound applies to minibatches, not the full dataset so the size $n$ is
-%% deliberately not $N$.
-
-Lemma~\ref{lem:quant_clt} demonstrates that if we know $E|X|$ and $E|X|^3$, we
-can bound the error of the normal approximation, which decays as
-$O(n^{-\frac{1}{2}})$. Making the change of variables $y = x s_X$,
+Lemma~\ref{lem:quant_clt} demonstrates that if we know $\mE[|X|]$ and
+$\mE[|X|^3]$, we can bound the error of the normal approximation, which decays
+as $O(n^{-\frac{1}{2}})$. Making the change of variables $y = x s_X$,
 Equation~(\ref{eq:clt-bounds}) becomes
 \begin{equation}\label{eq:clt-bounds2}
-   \sup_y\left|{\rm Pr}(\bar{X}< y) - \Phi\left(\frac{y}{s_X}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
+   \sup_y\left|{\rm Pr}(\bar{X}< y) - \Phi\left(\frac{y}{s_X}\right)\right| \leq
+   \frac{6.4\mE[|X|^3]+2\mE[|X|]}{\sqrt{n}}
 \end{equation}
@@
 term: $\bar{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta')$, so that
-$s_X = s_{\Delta^*}$. We can now substitute into Equation~(\ref{eq:clt-bounds2}) and displace by the mean, giving:
+$s_X = s_{\Delta^*}$. We can now substitute into Equation~(\ref{eq:clt-bounds2})
+and displace by the mean, giving:
 
 \begin{corollary}\label{cor:our_bound_delta_prime}
-\begin{align*}
-    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| &\leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} \\
-    &= \epsilon(\theta,\theta',b).
-\end{align*}
+\begin{equation}
+\sup_y \left|{\rm Pr}(\Delta^*< y) -\Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| \leq \frac{6.4\mE[|X|^3]+2\mE[|X|]}{\sqrt{b}}
+\end{equation}
 \end{corollary} %\vspace{-10pt}
+where the upper bound can be expressed as $\epsilon(\theta,\theta',b)$.
 Corollary~\ref{cor:our_bound_delta_prime} shows that the distribution of
@@
 $s^2_{\Delta^*}$. Furthermore, it bounds the error with \emph{estimable
-quantities}: both $E|X|$ and $E|X|^3$ can be estimated as means of $|\Lambda_i
-- \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each minibatch. We
-expect this will often be accurate enough on minibatches with hundreds of
-points, but otherwise bootstrap CIs can be computed.
+quantities}: both $\mE[|X|]$ and $\mE[|X|^3]$ can be estimated as means of
+$|\Lambda_i - \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each
+minibatch. We expect this will often be accurate enough on minibatches with
+hundreds of points, but otherwise bootstrap CIs can be computed.
 
@@
 \begin{proof}
-See Supplementary Material, Section~\ref{app:proof_cdf_bounds}.
+See Appendix, Section~\ref{app:proof_cdf_bounds}.
 \end{proof}
@@
 \]
-where $S(x)$ is the standard logistic function, and $X_{\rm nc}$ and $X_{\rm corr}$ are generated as per Algorithm 1. 
+where $S(x)$ is the standard logistic function, and $X_{\rm nc}$ and $X_{\rm
+corr}$ are generated as per Algorithm 1.
 \end{corollary}
@@
 \begin{proof}
-See Supplementary Material, Section~\ref{app:bounds_preserved}.
+See Appendix, Section~\ref{app:bounds_preserved}.
 \end{proof}
@@
 Corollary~\ref{cor:bounds_preserved} shows that the bounds from
-Section~\ref{ssec:delta_star_distribution} are preserved after adding 
-random variables, so our test remains accurate.  In fact we can do better
-($O(n^{-1})$ instead of $O(n^{-1/2})$) by using a more precise limit
-distribution under an additional assumption. We review this in the Supplementary
-Material, Section~\ref{app:better_error_bound}.
-
-% Daniel: I moved the previous Section 5.4 part to the supplementary material
-% since our code doesn't do this, so we should not describe this in much detail.
-
-% Daniel: this is probably less important than other parts of the paper. Hence,
-% I put it in as \begin{comment} ... \end{comment} but we can think about
-% including it back in if it is important enough. Also, the sub-section that was
-% after that, improved error bounds on skew estimation, has been moved to the
-% supplementary material. I don't see how we can discuss it in the paper without
-% taking up too much space.
+Section~\ref{ssec:delta_star_distribution} are preserved after adding random
+variables, so our test remains accurate.  In fact we can do better ($O(n^{-1})$
+instead of $O(n^{-1/2})$) by using a more precise limit distribution under an
+additional assumption. We review this in Appendix~\ref{app:better_error_bound}.
 
@@
 of the Markov chain under appropriate conditions. Such bounds were derived in
-both ~\citet{cutting_mh_2014} and~\citet{icml2014c1_bardenet14}. We include the
-result from~\citet{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the
+both~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}. We include the
+result from~\cite{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the
 total variation distance between two distributions $P$ and $Q$.  Let
@@
 
-\section{EXPERIMENTS}\label{sec:experiments}
+\section{Experiments}\label{sec:experiments}
+
+\subsection{Mixture of Gaussians}\label{ssec:gaussians}
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
+    \includegraphics[width=0.9\linewidth]{posterior_of_gaussian.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{minibatch_size_gaussian.png}
+    \includegraphics[width=0.9\linewidth]{minibatch_size_gaussian.png}
     \caption{
@@
 
-\subsection{Mixture of Gaussians}\label{ssec:gaussians}
+This model is adapted from~\cite{langevin_2011} by increasing the number
+of samples to 1 million.  The parameters are $\theta = \langle \theta_1,\theta_2 \rangle$, and the
+generation process is
+\begin{equation}\label{eq:data_generation}
+\begin{split}
+    \theta &\sim \mathcal{N}(0, {\rm diag}(\sigma_1^2,\sigma_2^2)) \\
+    x_i & \sim 0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
+\end{split}
+\end{equation}
+We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
+\langle 0,1 \rangle$. The original paper sampled 100 data points and estimated
+the posterior. We are interested in performance on larger problems and so
+sampled 1,000,000 points to form the posterior of
+$p(\theta)\prod_{i=1}^{1,000,000}p(x_i | \theta)$ with the same prior from
+Equation~(\ref{eq:data_generation}). This produces a much sharper posterior with
+two very narrow peaks.  Our goal is to reproduce the original posterior, so we
+adjust the temperature to $T=10,000$.  Taking logs, we get the target as shown
+in the far left of Figure~\ref{fig:gauss_mix_1}.
+
+We ran MCMC with with our test against the test
+from~\cite{icml2014c1_bardenet14} and a simplified version
+of~\cite{cutting_mh_2014}. Both methods use repeated testing which must be
+discounted to bound overall test accuracy. ~\cite{icml2014c1_bardenet14}
+conservatively partitions the test error, while~\cite{cutting_mh_2014} used a
+complex dynamic programming heuristic to estimate the overall test error. We did
+not implement this heuristic and instead set the individual test error equal to
+the overall test error. Thus our results for~\cite{cutting_mh_2014} should be
+treated as lower bounds. All methods were initialized with minibatch size 100.
+For~\cite{cutting_mh_2014} we increment sizes by 100 and set the tolerance
+$\epsilon=0.005$ to control overall test error.
+For~\cite{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma =
+1.5$ and use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
+samples using a random walk proposer with covariance ${\rm diag}(0.15, 0.15)$,
+which means the M-H test is responsible for shaping the sample distribution.
+
+Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
+samples for the three methods, with darker regions indicating a greater density
+of points. There are no obvious differences, so we measure the similarity
+between each set of samples and the actual posterior. \textbf{Daniel: I think
+this was taking up too much space so I put it in Appendix~\ref{app:gaussians}.
+Let's state the result (Table~\ref{tab:poissons}) and prioritize other stuff.
+TODO}
+
+Figure~\ref{fig:gauss_mix_2} shows that the new method dominates in
+terms of speed and efficiency. The histograms of the (final) minibatch sizes
+used each iteration show that our method consumes significantly less data; the
+distribution is short-tailed and the mean is 210, more than an order of
+magnitude better compared to the other two methods (averages are 15562 and
+16857). The sizes correspond to the running times of the methods, since
+they all consume linear time in data consumed.
+
+\subsection{Logistic Regression}\label{ssec:logistic}
 
@@
 	\centering
-	%\includegraphics[width=0.9\linewidth]{logistic_performance.png}
-	\includegraphics[width=0.8\linewidth]{logistic_regression_ll_acc_results.png}
+	\includegraphics[width=0.9\linewidth]{logistic_regression_ll_acc_results.png}
 	\caption{
@@
 	\centering
-	%\includegraphics[width=0.9\linewidth]{minibatch_size_logistic.png}
-	\includegraphics[width=0.8\linewidth]{logistic_regression_histograms_results.png}
+	\includegraphics[width=0.9\linewidth]{logistic_regression_histograms_results.png}
 	\caption{
@@
 
-This model is adapted from~\citet{langevin_2011} by increasing the number
-of samples to 1 million.  The parameters are $\theta = \langle \theta_1,\theta_2 \rangle$, and the
-generation process is
-\begin{equation}\label{eq:data_generation}
-\begin{split}
-    \theta &\sim \mathcal{N}(0, {\rm diag}(\sigma_1^2,\sigma_2^2)) \\
-    x_i & \sim 0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
-\end{split}
-\end{equation}
-We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
-\langle 0,1 \rangle$. The original paper sampled 100 data points and estimated
-the posterior. We are interested in performance on larger problems and so
-sampled 1,000,000 points to form the posterior of
-$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$ with the same prior from
-Equation~(\ref{eq:data_generation}). This produces a much sharper posterior with
-two very narrow peaks.  Our goal is to reproduce the original posterior, so we
-adjust the temperature to $T=10,000$.  Taking logs, we get the target as shown
-in the far left of Figure~\ref{fig:gauss_mix_1}.
-
-We ran MCMC with with our test against the test
-from~\citet{icml2014c1_bardenet14} and a simplified version
-of~\citet{cutting_mh_2014}. Both methods use repeated testing which
-must be discounted to bound overall test
-accuracy. ~\citet{icml2014c1_bardenet14} conservatively partitions the
-test error, while~\citet{cutting_mh_2014} used a complex dynamic
-programming heuristic to estimate the overall test error. We did not
-implement this heuristic and instead set the individual test error
-equal to the overall test error. Thus our results for~\citet{cutting_mh_2014} should be treated as lower bounds. All
-methods were initialized with minibatch size 100.  For~\citet{cutting_mh_2014} we
-increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
-use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
-samples using a random walk proposer with covariance ${\rm diag}(0.15,
-0.15)$, which means the M-H test is responsible for shaping the sample
-distribution.
-
-Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
-samples for the three methods, with darker regions indicating a greater density
-of points. There are no obvious differences, so we measure the similarity
-between each set of samples and the actual posterior.
-
-We discretize the posterior coordinates into bins with respect to the two
-components of $\theta$.  The probability $P_i$ of a sample falling into bin $i$
-is the integral of the true posterior over the bin's area.  A single sample
-should therefore be multinomial with distribution $P$, and a set of $n$ (ideally
-independent) samples is ${\rm Multinomial}(P,n)$. This distribution is simple
-and we can use it to measure the quality of the samples rather than use general
-purpose tests like KL-divergence or likelihood-ratio, which are problematic
-with zero counts.
-
-For large $n$, the per-bin distributions are approximated by Poissons with
-parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
-$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
-out of $N_{\rm bins}$ total. We have
-\begin{equation}\label{eq:log_prob_poissons}
-\begin{split}
-    \log p(c_1, \ldots, c_{N_{\rm bins}} \mid P_1, \ldots, P_{N_{\rm bins}}) & =  \\
-    \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)). & \\
-\end{split}
-\end{equation}
-Table~\ref{tab:poissons} shows the likelihoods. To facilitate
-interpretation we perform significance tests using Chi-Squared
-distribution (also in Table~\ref{tab:poissons}). Scores lie
-between~\citet{cutting_mh_2014} and~\citet{icml2014c1_bardenet14}, but
-the variance of these values is high and ordering changes depending on
-the range of samples generated.
-
-Figure~\ref{fig:gauss_mix_2} shows that the new method dominates in
-terms of speed and efficiency. The histograms of the (final) minibatch sizes
-used each iteration show that our method consumes significantly less data; the
-distribution is short-tailed and the mean is 210, more than an order of
-magnitude better compared to the other two methods (averages are 15562 and
-16857). The sizes correspond to the running times of the methods, since
-they all consume linear time in data consumed.
-
-\subsection{Logistic Regression}\label{ssec:logistic}
-
 \begin{table}[t]
@@
 \label{tab:poissons}
-    \centering
-    \resizebox{.45\textwidth}{!}{% <------ Don't forget this %
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
 \begin{tabular}{l l l l}
-    {\bf Metric} & {\bf Ours} & {\bf Korat.'14} & {\bf Barde.'14} \\
 \hline
+\abovespace\belowspace
+Metric & Ours & Korat.'14 & Barde.'14 \\
+\hline
+\abovespace
 Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
+\belowspace
 Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
-    \end{tabular}% <------ Don't forget this %
-   }
+\hline
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
 \end{table}
 
-We next test logistic regression for the binary classification of 1s
-versus 7s on a subset of the MNIST8M dataset, a larger version of
-MNIST~\citep{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
-450k training and 192k testing points.
-%%% The pixels are scaled in $[0,1]$  - JFC Why ????
-We impose a uniform prior on $\theta$ and again use a random walk
-proposer, this time with covariance matrix $0.05I$. The temperature is
-set at $T=1000$. We run the three tests for 3000 samples and again use 100
-as the starting minibatch size. Method~\citet{cutting_mh_2014} used
-per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
-we use their recommended discounting factor value of $\gamma=2.0$. We tried to use the
-symbolic bound for $C_{\theta,\theta'}$ for Logistic regression
-from~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead we use
-the empirical $C_{\theta,\theta'}$ from the entire dataset and do not add its
-computation time when considering runtime, so again our results represent a
-lower bound for this method.
+We next test logistic regression for the binary classification of 1s versus 7s
+on a subset of the MNIST8M dataset, a larger version of
+MNIST~\cite{lecun-mnisthandwrittendigit-2010}. We randomly subsampled 450k
+training and 192k testing points.  We impose a uniform prior on $\theta$ and
+again use a random walk proposer, this time with covariance matrix $0.05I$. The
+temperature is set at $T=1000$. We run the three tests for 3000 samples and
+again use 100 as the starting minibatch size. Our undiscounted implementation
+of~\cite{cutting_mh_2014} used per-test error of $\epsilon=0.05$.
+For~\cite{icml2014c1_bardenet14}, we use their recommended discounting factor
+value of $\gamma=2.0$. We tried to use the symbolic bound for
+$C_{\theta,\theta'}$ for Logistic regression from~\cite{icml2014c1_bardenet14},
+but found it to be too high. Instead we use the empirical $C_{\theta,\theta'}$
+from the entire dataset and do not add its computation time when considering
+runtime, so again our results represent a lower bound for this method.
 
@@
 in terms of accuracy, achieving convergence roughly twice as fast compared
-to~\citet{cutting_mh_2014} and about 1.5 orders of magnitude faster
-than~\citet{icml2014c1_bardenet14}, though its log likelihood results lag
+to~\cite{cutting_mh_2014} and about 1.5 orders of magnitude faster
+than~\cite{icml2014c1_bardenet14}, though its log likelihood results lag
 slightly behind the former algorithm during the very early stages of
@@
 three methods on a log-log scale. With an initial size of 100, we achieve an
-average minibatch size of 393, far smaller than the averages of the other two
-methods, which have longer-tailed distributions.
-
-\section{CONCLUSIONS}\label{sec:conclusion}
+average minibatch size of 393, more than two orders of magnitude smaller
+than~\cite{icml2014c1_bardenet14}, and almost an order of magnitude faster than
+the undiscounted implementation of~\cite{cutting_mh_2014}.
+
+\subsection{Collaborative Filtering}\label{ssec:collab_filtering}
+
+\textbf{Daniel: TODO we wanted a third experiment, right?}
+
+\section{Conclusions}\label{sec:conclusion}
+
+\textbf{Daniel: need to revise this to take into account reviewer suggestion,
+with future work ideas. This will add some 2016 references so that'll be good
+(makes it current)}
 
@@
 Monte Carlo methods, providing a recipe for how to use our algorithm (following
-the framework of~\citet{sgmcmc_2015}), or integrating parallel
-MCMC~\citep{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
-
-\iffalse
-\subsubsection{Figures}
-
-All artwork must be centered, neat, clean, and legible.  All lines
-should be very dark for purposes of reproduction, and art work should
-not be hand-drawn.  Figures may appear at the top of a column, at the
-top of a page spanning multiple columns, inline within a column, or
-with text wrapped around them, but the figure number and caption
-always appear immediately below the figure.  Leave 2 line spaces
-between the figure and the caption. The figure caption is initial caps
-and each figure should be numbered consecutively.
-
-Make sure that the figure caption does not get separated from the
-figure. Leave extra white space at the bottom of the page rather than
-splitting the figure and figure caption.
-\begin{figure}[h]
-\vspace{.3in}
-\centerline{\fbox{This figure intentionally left non-blank}}
-\vspace{.3in}
-\caption{Sample Figure Caption}
-\end{figure}
-
-\subsubsection{Tables}
-
-All tables must be centered, neat, clean, and legible. Do not use hand-drawn tables. Table number and title always appear above the table.
-See Table~\ref{sample-table}.
-
-Use one line space before the table title, one line space after the table title, and one line space after the table. The table title must be
-initial caps and each table numbered consecutively.
-
-\begin{table}[h]
-\caption{Sample Table Title} \label{sample-table}
-\begin{center}
-\begin{tabular}{ll}
-{\bf PART}  &{\bf DESCRIPTION} \\
-\hline \\
-Dendrite         &Input terminal \\
-Axon             &Output terminal \\
-Soma             &Cell body (contains cell nucleus) \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\section{SUPPLEMENTARY MATERIAL}
-
-If you need to include additional appendices during submission, you
-can include them in the supplementary material file.
-
-\fi
-
-\newpage
-\subsubsection*{References}
-% Daniel: Sorry, I don't know why this is needed to get rid of the annoying
-% spaces below the "References" text.
-\vspace{-25pt} 
-%\bibliographystyle{apalike}
-\bibliographystyle{abbrvnat}
-\setcitestyle{authoryear,open={[},close={]}}
-\bibliography{aistats2017}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+the framework of~\cite{sgmcmc_2015}), or integrating parallel
+MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
+
+\bibliography{paper}
+\bibliographystyle{icml2017}
+
 \clearpage
@@
 
+% Daniel: I set the Appendix at single column which makes some of the math
+% easier to write. Other ICML papers from 2016 did this so it's not a big deal.
+
+\onecolumn
 \begin{center}
-{\Large Supplementary Material}
+{\Large Appendix}
 \end{center}
 
-\section{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}
-
-\begin{comment}
-
-{\color{blue} Daniel: changes I made:
-
-\begin{enumerate}
-    \item (TODO later) After Equation~\ref{eq:accept_test_zero}, we say that
-    ``the resulting test with our chosen $u_0$ will never correctly succeed
-    [...]'', but I think we should say it will not succeed with high
-    probability, because we could select a minibatch of samples such that the
-    mean \emph{actually is} different from 0.5 (it depends on the distribution
-    of the variable that corresponds to $Y = \frac{1}{b}\sum_{i=1}^b x_i$).
-
-    \item (TODO later) Unfortunately, I am unable to understand why ``the size
-    of the range in $u$ is at least $\exp([-2,-1.125])$'', unless the real range
-    is $\exp([-2,0])$ (as I think it should be) and you were simply cropping off
-    the extra range.
-
-    \item (TODO later) My main confusion now is that I am unable to derive
-    Equation~\ref{eq:log_uo_range}, and we may want to clarify the notation used
-    when we put two intervals side by side.
-
-    \item (TODO later) I do not understand why after
-    Equation~\ref{eq:accept_test_rewritten}, we say that the standard deviation
-    of the LHS ``given the interval constraints is at least $0.5/\sqrt{b}$.'' We
-    already know that the $X_i$ are iid $\mathcal{N}(\theta = 0.5, 1)$ variables
-    with a fixed $\theta = 0.5$ value, so we already know the standard deviation
-    is supposed to be $\sqrt{({\rm Var}(x_1) + \cdots + {\rm Var}(x_b))/b^2} =
-    1/\sqrt{b}$.
-
-    \item I added Equation~\ref{eq:rhs_range}, which I think is how you
-    determined the range. Note that it doesn't seem to matter whether it is
-    $[-1,0.75]$ or $[-0.75,1]$ since the main thing that matters is that we have
-    maximum absolute value of one.
-
-    \item At the end, it said the joint distribution of
-    $(\theta,\theta'-\theta,u)$, but I changed it to $(\theta,\theta',u)$, based
-    on your change to the text of Lemma~\ref{lem:worst_case}. But here's a
-    question: how is the sample of $\theta'$ (or even $\theta'-\theta$, as it
-    was written earlier) independent of $\theta$?
-\end{enumerate}
-}
-
-\end{comment}
-
-Choose $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ (event 1) and $(\theta -0.5)
-\in \pm\frac{1}{\sqrt{N}}[0.5,1]$ filtered for  matching sign (event 2).  As
-discussed in Lemma~\ref{lem:worst_case}, both $q(\theta' \mid \theta)$ and
-$p(\theta \mid x_1,\ldots,x_N)$ have variance $1/N$. If we denote $\Phi$ as the
-CDF of the standard normal distribution, then the former event occurs with
-probability $p_0 = 2(\Phi(\sqrt{N}\frac{1}{\sqrt{N}}) -
+This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
+the proofs that we omitted from the main text due to space constraints.
+Appendix~\ref{app:experiments} provides further details on our experiments to
+assist understanding and reproducibility.
+
+\section{Proofs of Lemmas and Corollaries}\label{app:proofs}
+
+\subsection{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}
+
+Choose $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ (event 1) and
+$(\theta -0.5) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ filtered for  matching sign
+(event 2).  As discussed in Lemma~\ref{lem:worst_case}, both $q(\theta' |
+\theta)$ and $p(\theta | x_1,\ldots,x_N)$ have variance $1/N$. If we denote
+$\Phi$ as the CDF of the standard normal distribution, then the former event
+occurs with probability $p_0 = 2(\Phi(\sqrt{N}\frac{1}{\sqrt{N}}) -
 \Phi(\sqrt{N}\frac{0.5}{\sqrt{N}})) = 2(\Phi(1)-\Phi(0.5)) \approx 0.2997$. The
 latter event, because we restrict signs, occurs with probability $p_1 =
-\Phi(1) - \Phi(0.5) \approx 0.14988$.
-
-These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative
-by inspection of equation (\ref{eq:whyneg}) below.
-This implies that we can find a $u \in (0,1)$ so that
-$\psi(u,\theta,\theta') = \log u < 0$ equals $E[\Lambda^*(\theta,\theta')]$.
-Specifically, choose $u_0$ to satisfy $\log u_0 = E[\Lambda^*(\theta,\theta')]$.
-Using $E[x_i] = 0.5$ and Equation~(\ref{eq:lemma_ll_ratio}), we see that
+\Phi(1)-\Phi(0.5) \approx 0.14988$.
+
+These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative by
+inspection of equation (\ref{eq:whyneg}) below.  This implies that we can find a
+$u \in (0,1)$ so that $\psi(u,\theta,\theta') = \log u < 0$ equals
+$\mE[\Lambda^*(\theta,\theta')]$.  Specifically, choose $u_0$ to satisfy $\log u_0
+= \mE[\Lambda^*(\theta,\theta')]$.  Using $\mE[x_i^*] = 0.5$ and
+Equation~(\ref{eq:lemma_ll_ratio}), we see that
 \begin{equation}
-  \log u_0 = N(\theta'-\theta)\frac{1}{b} \cdot E\left[\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right]
+  \log u_0 = N(\theta'-\theta)\frac{1}{b} \cdot \mE\left[\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2}\right]
 \end{equation}
@@
 \end{equation}
-Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta') \not\approx
-\psi(u,\theta,\theta')$ used in ~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14} , where $\not\approx$
-means ``significantly different from'' under the distribution over
-samples of $x_i$. This turns out to be
-\begin{equation}
-\begin{split}
-\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') & \quad \\
-\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2} & \not\approx \log u_0\\
-\iff \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) & \not\approx  0 \\
-\iff \frac{1}{b}\sum_{i=1}^b x_i-0.5 & \not\approx 0. \label{eq:accept_test_zero}
-\end{split}
-\end{equation}
-Since the $x_i$ have mean 0.5, the resulting test with our chosen $u_0$ will
+Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta')
+\not\approx \psi(u,\theta,\theta')$ used in ~\cite{cutting_mh_2014}
+and~\cite{icml2014c1_bardenet14} , where $\not\approx$ means ``significantly
+different from'' under the distribution over samples. This is
+\begin{align}
+\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') 
+&\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2} \not\approx \log u_0\\
+&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \not\approx  0 \\
+&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-0.5 \not\approx 0. \label{eq:accept_test_zero}
+\end{align}
+Since the $x_i^*$ have mean 0.5, the resulting test with our chosen $u_0$ will
 never correctly succeed and must use all $N$ data points.  Furthermore, if we
@@
 \begin{equation}\label{eq:accept_test_rewritten}
-    \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
+    \frac{1}{b}\sum_{i=1}^b x_i^*-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
 \end{equation}
@@
 \begin{equation}\label{eq:rhs_range}
-\begin{split}
-    \min\{{\rm RHS}\} & = \frac{-0.5}{\sqrt{N} \cdot 0.5} = -\frac{1}{\sqrt{N}} \\
-    {\rm and} \quad \max\{{\rm RHS}\} & = \frac{0.375}{\sqrt{N} \cdot 0.5} = \frac{0.75}{\sqrt{N}}
-\end{split}
+\min\{{\rm RHS}\} = \frac{-0.5}{\sqrt{N} \cdot 0.5} = -\frac{1}{\sqrt{N}}
+\quad {\rm and} \quad \max\{{\rm RHS}\} = \frac{0.375}{\sqrt{N} \cdot 0.5} = \frac{0.75}{\sqrt{N}}
 \end{equation}
@@
 
-\section{Proof of Lemma~\ref{lem:quant_clt}}\label{app:quant_clt}
-
-The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
-\begin{align*}
--6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
-&\leq 1.36E|X|^3.
-\end{align*}
+\subsection{Proof of Lemma~\ref{lem:quant_clt}}\label{app:quant_clt}
+
+The following bound is given immediately after Corollary 2 from~\cite{explicit-clt05}:
+\begin{equation}
+-6.4\mE[|X|^3]-2\mE[|X|] \leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \leq
+1.36\mE[|X|^3].
+\end{equation}
 This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
@@
 
-\section{Proof of Lemma~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}
+\subsection{Proof of Lemma~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}
 
@@
 and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
-\begin{align*}
--\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
-&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
-&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
-\end{align*}
+\begin{equation}
+-\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
+\end{equation}
+% Daniel: this is the multi-line version, but we should use the full page.
+%\begin{align*}
+%-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
+%&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
+%&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
+%\end{align*}
 as desired.
 
-\section{Proof of Corollary~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}
+\subsection{Proof of Corollary~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}
 
@@
 \begin{equation}
-\begin{split}
-    P(y) & = {\rm Pr}(\Delta^* < y) \\
-     {\rm and} \quad Q(y) & = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
-\end{split}
+    P(y) = {\rm Pr}(\Delta^* < y)
+     \quad {\rm and} \quad Q(y) = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
 \end{equation}
@@
 \begin{equation}
-\begin{split}
-    P'(y) & = {\rm Pr}(\Delta^*+X_{\rm nc} < y) \\
-    \quad {\rm and}\quad Q'(y) & = \Phi\left({y-\Delta}\right)
-\end{split}
+    P'(y) = {\rm Pr}(\Delta^*+X_{\rm nc} < y)
+    \quad {\rm and}\quad Q'(y) = \Phi\left({y-\Delta}\right)
 \end{equation}
@@
 \begin{equation}
-\begin{split}
-    P''(y) & = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) \\
-    \quad {\rm and}\quad Q''(y) & = S\left({y-\Delta}\right)
-\end{split}
+    P''(y) = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y)
+    \quad {\rm and}\quad Q''(y) = S\left({y-\Delta}\right)
 \end{equation}
 Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
-$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
-Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
+$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) =
+{\rm Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
 that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
-nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
-log} > 0)$ by at most $\epsilon$.
-
-\section{Improved Error Bounds Based on Skew Estimation}\label{app:better_error_bound}
-
-We show that the CLT error bound can be improved to $O(n^{-1})$ 
-using a more precise limit distribution under an additional assumption. Let
-$\mu_i$ denote the $i^{th}$ moment, and $b_i$ denote the $i^{th}$ absolute
-moment of $X$. If Cramer's condition holds:
+nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm log} > 0)$
+by at most $\epsilon$.
+
+\subsection{Improved Error Bounds Based on Skew Estimation}\label{app:better_error_bound}
+
+We show that the CLT error bound can be improved to $O(n^{-1})$ using a more
+precise limit distribution under an additional assumption. Let $\mu_i$ denote
+the $i^{th}$ moment, and $b_i$ denote the $i^{th}$ absolute moment of $X$. If
+Cramer's condition holds:
 \begin{equation}\label{eq:cramers_condition}
-    \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1,
+    \lim_{t \to \infty} \sup |\mE[\exp(i t X)]| < 1,
 \end{equation}
 then Equation 2.2 in Bentkus et al.'s work on Edgeworth
-expansions~\citep{Bentkus97} provides:
+expansions~\cite{Bentkus97} provides:
 
@@
 $n^{-1/2}$.
-%{\color{blue} (Daniel: how did the equation in the next sentence
-% appear? I verified it and it seems correct, but it looks like it popped out of
-%nowhere.)}
-% JFC: It comes from taking a second derivative. You dont need to explain it. 
 Importantly, since $\phi^{''}(x) = x^2\phi(x) - \phi(x)$ where
@@
 table. We leave further analysis and experiments with $g(x,y)$ as future work.
-%{\color{blue} Daniel: I will look at the code to understand this paragraph
-%better because right now I am confused.}
-% JFC: the code doesnt implement this yet.
-
-% Daniel: don't forget this. The AISTATS website says to put it in the
-% supplementary material, so I put it at the very end on a new page (to make it
-% more noticeable). Obviously for the arXiv version, we don't want this there.
-% Also, note to John, the reviewers for AISTATS **will have access to the
-% previous NIPS reviews**, which is why I am making references to those.
-\clearpage
-\section{NIPS 2016 Submission Statement}
-
-We previously submitted an early, incomplete draft of this paper to NIPS 2016. The
-current manuscript has been substantially improved since that submission. We have
-made the following changes:
-
-\begin{enumerate}
-\item The (incomplete) analysis in the NIPS paper has been substantially replaced.
-  The asymptotic CLT arguments in the earlier paper have been replaced with
-  explicit quantitative bounds. There were several gaps in the earlier analysis:
-  the effects of addition of random variables on CDF errors and the construction and
-  error bounds of the correction distribution which are key parts of the present
-  paper have been added.
-
-\item Several recent relevant references have been added, and we have written
-  clearer statements about the contribution of the present work.
-  These include~\citep{TallData15} and the Barker test~\citep{Barker65}.
-
-\item Some reviewers mentioned that our deconvolution approach to
-  determine the correction distribution was not unique and thus
-  ill-defined.  We added detail on how the correction
-  distribution is computed (see Section~\ref{sec:correction}). While the distribution is not unique, we
-  regularize the inversion to obtain a concrete distribution with very
-  low error (close to single-point machine precision).
-
-\item The earlier draft had an inconsistency in its statement about constant
-  minibatch size. We clarify that this is true only in expected value.
-
-\item Our experiments now include new comparisons with a baseline
-  from~\citep{icml2014c1_bardenet14} (in addition to~\citep{cutting_mh_2014},
-  which we had earlier). Furthermore, each algorithm now runs MCMC sampling on
-  the \emph{same} distribution.  Previously, we ran our distribution at a
-  higher temperature but kept the algorithm from~\citep{cutting_mh_2014}
-  running on the distribution at temperature $T=1$. In this set of
-  experiments, we tuned hyperparameters of the other algorithms to yield
-  best results. In the Gaussian mixture model scenario
-  (Section~\ref{ssec:gaussians}), we provide likelihoods and test
-  statistics to measure the accuracy of the samples.
-
-\item We improved Algorithm~\ref{alg:our_algorithm} so that we show
-  explicitly when we compute the moments, and why we now only need one
-  distribution $C_\sigma$ for $\sigma=1$ due to the extra $X_{\rm nc}$
-  variable.
-
-\item Finally, we made minor revisions addressing: differences between the
-  Barker function vs. the original MH test, and sampling with vs. without
-  replacement.
-\end{enumerate}
-
-We estimate the degree of overlap between the papers in mathematics (equations) at less than
-10\%, in overall text at less than 30\%. The experiments are similar but were also redone at larger
-scale in this draft.
+
+\section{Additional Experiment Details}\label{app:experiments}
+% Daniel: all three experiments should be described in more detail here.
+
+\subsection{Gaussian Mixture Model Experiment (Section~\ref{ssec:gaussians})}\label{app:gaussians}
+
+We discretize the posterior coordinates into bins with respect to the two
+components of $\theta$.  The probability $P_i$ of a sample falling into bin $i$
+is the integral of the true posterior over the bin's area.  A single sample
+should therefore be multinomial with distribution $P$, and a set of $n$ (ideally
+independent) samples is ${\rm Multinomial}(P,n)$. This distribution is simple
+and we can use it to measure the quality of the samples rather than use general
+purpose tests like KL-divergence or likelihood-ratio, which are problematic
+with zero counts.
+
+For large $n$, the per-bin distributions are approximated by Poissons with
+parameter $\lambda_i=P_i n$. Given samples $\{\theta_1,\ldots,\theta_T\}$, let
+$c_j$ denote the number of individual samples $\theta_i$ that fall in bin $j$
+out of $N_{\rm bins}$ total. We have
+\begin{equation}\label{eq:log_prob_poissons}
+\begin{split}
+    \log p(c_1, \ldots, c_{N_{\rm bins}} | P_1, \ldots, P_{N_{\rm bins}}) & =  \\
+    \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)). & \\
+\end{split}
+\end{equation}
+Table~\ref{tab:poissons} shows the likelihoods. To facilitate
+interpretation we perform significance tests using Chi-Squared
+distribution (also in Table~\ref{tab:poissons}). Scores lie
+between~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}, but
+the variance of these values is high and ordering changes depending on
+the range of samples generated.
+
+\textbf{Daniel: TODO improve this}
+
+\subsection{Logistic Regression Experiment (Section~\ref{ssec:logistic})}\label{app:logistic}
+
+\textbf{Daniel: TODO}
+
+\subsection{Collaborative Filtering Experiment (Section~\ref{ssec:collab_filtering})}\label{app:collab_filtering}
+
+\textbf{Daniel: TODO}
 
 \end{document}
+
+% Daniel: some notes to keep in mind to standardize to ICML format.
+
+%% \begin{figure}[ht]
+%% \vskip 0.2in
+%% \begin{center}
+%% \centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
+%% \caption{Historical locations and number of accepted papers for International
+%%   Machine Learning Conferences (ICML 1993 -- ICML 2008) and
+%%   International Workshops on Machine Learning (ML 1988 -- ML
+%%   1992). At the time this figure was produced, the number of
+%%   accepted papers for ICML 2008 was unknown and instead estimated.}
+%% \label{icml-historical}
+%% \end{center}
+%% \vskip -0.2in
+%% \end{figure} 
+%% 
+%% \subsection{Figures}
+%%  
+%% You may want to include figures in the paper to help readers visualize
+%% your approach and your results. Such artwork should be centered,
+%% legible, and separated from the text. Lines should be dark and at
+%% least 0.5~points thick for purposes of reproduction, and text should
+%% not appear on a gray background.
+%% 
+%% Label all distinct components of each figure. If the figure takes the
+%% form of a graph, then give a name for each axis and include a legend
+%% that briefly describes each curve. Do not include a title inside the
+%% figure; instead, the caption should serve this function.
+%% 
+%% Number figures sequentially, placing the figure number and caption
+%% {\it after\/} the graphics, with at least 0.1~inches of space before
+%% the caption and 0.1~inches after it, as in
+%% Figure~\ref{icml-historical}.  The figure caption should be set in
+%% 9~point type and centered unless it runs two or more lines, in which
+%% case it should be flush left.  You may float figures to the top or
+%% bottom of a column, and you may set wide figures across both columns
+%% (use the environment {\tt figure*} in \LaTeX), but always place
+%% two-column figures at the top or bottom of the page.
+%% 
+%% \subsection{Tables} 
+%%  
+%% You may also want to include tables that summarize material. Like 
+%% figures, these should be centered, legible, and numbered consecutively. 
+%% However, place the title {\it above\/} the table with at least 
+%% 0.1~inches of space before the title and the same after it, as in 
+%% Table~\ref{sample-table}. The table title should be set in 9~point 
+%% type and centered unless it runs two or more lines, in which case it
+%% should be flush left.
+%% 
+%% % Note use of \abovespace and \belowspace to get reasonable spacing 
+%% % above and below tabular lines. 
+%% 
+%% \begin{table}[t]
+%% \caption{Classification accuracies for naive Bayes and flexible 
+%% Bayes on various data sets.}
+%% \label{sample-table}
+%% \vskip 0.15in
+%% \begin{center}
+%% \begin{small}
+%% \begin{sc}
+%% \begin{tabular}{lcccr}
+%% \hline
+%% \abovespace\belowspace
+%% Data set & Naive & Flexible & Better? \\
+%% \hline
+%% \abovespace
+%% Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
+%% Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
+%% Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
+%% Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
+%% Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
+%% Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
+%% Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
+%% \belowspace
+%% Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
+%% \hline
+%% \end{tabular}
+%% \end{sc}
+%% \end{small}
+%% \end{center}
+%% \vskip -0.1in
+%% \end{table}
+%%
+%% % Acknowledgements should only appear in the accepted version. 
+%% \section*{Acknowledgements} 
+%% \textbf{Do not} include acknowledgements in the initial version of
+%% the paper submitted for blind review.

### 2017-02-05 icml2017/paper.tex
Fixed appendix a bit, added some of my table results.
This is for the deconvolution procedure. I think we should have more detailed tables, and then from looking at these systematically, we should find the good distribution to use and the target variance to use. I have an idea to have maybe six different sub-tables :) That's what the supplementary material is for

@@
 
-% Daniel: also need to modify table, add more info?
-\begin{table}[t]
-\caption{Error ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$.}
-\label{tab:xcorr}
-\vskip 0.15in
-\begin{center}
-\begin{small}
-\begin{sc}
-\begin{tabular}{l l l l}
-\hline
-\abovespace\belowspace
-${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error} \\
-\hline
-\abovespace
-4000 & 0.9 & 1 & 1.0e-4  \\
-\belowspace
-4000 & 0.8 & 0.03 & 5.0e-6 \\
-\hline
-\end{tabular}
-\end{sc}
-\end{small}
-\end{center}
-\vskip -0.1in
-\end{table}
-
 With this approach, there is no guarantee that $u^* \geq 0$. However, we have
@@
 $\sigma = 0$. It retains some weak negative values for $\sigma > 0$ but removal
-of those leads to small error. Table~\ref{tab:xcorr} shows that the errors
-between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ approach single floating
-precision (about $10^{-7}$), and Algorithm~\ref{alg:our_algorithm} describes our
-procedure. A few points:
+of those leads to small error. \textbf{Daniel: I moved this to
+Appendix~\ref{app:correction} so that we can have more detailed results. For the
+main text, let's state the value $\sigma$ and its error. Originally, we said:
+Table~\ref{tab:xcorr} shows that the errors between $X_{\rm norm}+X_{\rm corr}$
+and $X_{\rm log}$ approach single floating precision (about $10^{-7}$).}
+Algorithm~\ref{alg:our_algorithm} describes our procedure. A few points:
 
@@
 this was taking up too much space so I put it in Appendix~\ref{app:gaussians}.
-Let's state the result (Table~\ref{tab:poissons}) and prioritize other stuff.
-TODO}
+Let's state the result (in Table~\ref{tab:poissons}) and prioritize other stuff
+here.  TODO}
 
@@
 \end{figure*}
+
+We next test logistic regression for the binary classification of 1s versus 7s
+on a subset of the MNIST8M dataset, a larger version of
+MNIST~\cite{lecun-mnisthandwrittendigit-2010}. We randomly subsampled 450k
+training and 192k testing points.  We impose a uniform prior on $\theta$ and
+again use a random walk proposer, this time with covariance matrix $0.05I$. The
+temperature is set at $T=1000$. We run the three tests for 3000 samples and
+again use 100 as the starting minibatch size. Our undiscounted implementation
+of~\cite{cutting_mh_2014} used per-test error of $\epsilon=0.05$.
+For~\cite{icml2014c1_bardenet14}, we use their recommended discounting factor
+value of $\gamma=2.0$. We tried to use the symbolic bound for
+$C_{\theta,\theta'}$ for Logistic regression from~\cite{icml2014c1_bardenet14},
+but found it to be too high. Instead we use the empirical $C_{\theta,\theta'}$
+from the entire dataset and do not add its computation time when considering
+runtime, so again our results represent a lower bound for this method.
+
+Figure~\ref{fig:logistic_performance} shows the test log likelihood and
+prediction accuracy as a function of the cumulative training points
+processed.\footnote{The curves do not span the same length over the x-axis since
+the methods consume different amounts of data.} To generate the curves, for each
+of the sampled vectors $\theta_t$, $t\in\{1,\ldots,3000\}$, we use $\theta_t$ as
+the parameter for logistic regression.  Our minibatch MH test is more efficient
+in terms of accuracy, achieving convergence roughly twice as fast compared
+to~\cite{cutting_mh_2014} and about 1.5 orders of magnitude faster
+than~\cite{icml2014c1_bardenet14}, though its log likelihood results lag
+slightly behind the former algorithm during the very early stages of
+exploration.
+
+Figure~\ref{fig:logistic_minibatch} shows histograms of minibatch sizes for the
+three methods on a log-log scale. With an initial size of 100, we achieve an
+average minibatch size of 393, more than two orders of magnitude smaller
+than~\cite{icml2014c1_bardenet14}, and almost an order of magnitude faster than
+the undiscounted implementation of~\cite{cutting_mh_2014}.
+
+\subsection{Collaborative Filtering}\label{ssec:collab_filtering}
+
+\textbf{Daniel: TODO we wanted a third experiment, right?}
+
+\section{Conclusions}\label{sec:conclusion}
+
+\textbf{Daniel: need to revise this to take into account reviewer suggestion,
+with future work ideas. This will add some 2016 references so that'll be good
+(makes it current)}
+
+We have derived an M-H test for minibatch MCMC which approximates full data
+tests. We present theoretical results and experimentally show the benefits of
+our test on Gaussian mixtures and logistic regression. Directions for future
+work include running more experiments with a particular focus on controlling
+variances, testing on neural networks, combining our results with Hamiltonian
+Monte Carlo methods, providing a recipe for how to use our algorithm (following
+the framework of~\cite{sgmcmc_2015}), or integrating parallel
+MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
+
+\bibliography{paper}
+\bibliographystyle{icml2017}
+
+\clearpage
+\appendix
+
+% Daniel: I set the Appendix at single column which makes some of the math
+% easier to write. Other ICML papers from 2016 did this so it's not a big deal.
+
+\onecolumn
+\begin{center}
+{\Large Appendix}
+\end{center}
+
+This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
+the proofs that we omitted from the main text due to space constraints.
+Appendix~\ref{app:experiments} provides further details on the correction
+distribution derivation and on our three main experiments to assist
+understanding and reproducibility.
+
+\section{Proofs of Lemmas and Corollaries}\label{app:proofs}
+
+\subsection{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}
+
+Choose $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ (event 1) and
+$(\theta -0.5) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ filtered for  matching sign
+(event 2).  As discussed in Lemma~\ref{lem:worst_case}, both $q(\theta' |
+\theta)$ and $p(\theta | x_1,\ldots,x_N)$ have variance $1/N$. If we denote
+$\Phi$ as the CDF of the standard normal distribution, then the former event
+occurs with probability $p_0 = 2(\Phi(\sqrt{N}\frac{1}{\sqrt{N}}) -
+\Phi(\sqrt{N}\frac{0.5}{\sqrt{N}})) = 2(\Phi(1)-\Phi(0.5)) \approx 0.2997$. The
+latter event, because we restrict signs, occurs with probability $p_1 =
+\Phi(1)-\Phi(0.5) \approx 0.14988$.
+
+These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative by
+inspection of equation (\ref{eq:whyneg}) below.  This implies that we can find a
+$u \in (0,1)$ so that $\psi(u,\theta,\theta') = \log u < 0$ equals
+$\mE[\Lambda^*(\theta,\theta')]$.  Specifically, choose $u_0$ to satisfy $\log u_0
+= \mE[\Lambda^*(\theta,\theta')]$.  Using $\mE[x_i^*] = 0.5$ and
+Equation~(\ref{eq:lemma_ll_ratio}), we see that
+\begin{equation}
+  \log u_0 = N(\theta'-\theta)\frac{1}{b} \cdot \mE\left[\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2}\right]
+\end{equation}
+\begin{equation}\label{eq:whyneg}
+    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
+\end{equation}
+Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta')
+\not\approx \psi(u,\theta,\theta')$ used in ~\cite{cutting_mh_2014}
+and~\cite{icml2014c1_bardenet14} , where $\not\approx$ means ``significantly
+different from'' under the distribution over samples. This is
+\begin{align}
+\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') 
+&\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2} \not\approx \log u_0\\
+&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \not\approx  0 \\
+&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-0.5 \not\approx 0. \label{eq:accept_test_zero}
+\end{align}
+Since the $x_i^*$ have mean 0.5, the resulting test with our chosen $u_0$ will
+never correctly succeed and must use all $N$ data points.  Furthermore, if we
+sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be
+sufficiently different from 0.5 to allow the test to succeed.
+
+The choices above for $\theta$ and $\theta'$ guarantee that
+\begin{equation}\label{eq:log_uo_range}
+    \log u_0 \in -[0.5,1][0.75,1.5] = [-1.5, -0.375].
+\end{equation}
+Next, consider the range of $u$ values near $u_0$:
+\begin{equation}\label{eq:log_u_range}
+    \log u \in \log u_0 + [-0.5,0.375].
+\end{equation}
+The size of the range in $u$ is at least $\exp([-2,-1.125]) \approx
+[0.13534,0.32465]$ and occurs with probability at least $p_2=0.18932$. With $u$
+in this range, we rewrite the test as:
+\begin{equation}\label{eq:accept_test_rewritten}
+    \frac{1}{b}\sum_{i=1}^b x_i^*-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
+\end{equation}
+so that, as in Equation~(\ref{eq:accept_test_zero}), the LHS has expected value
+zero.  Given our choice of intervals for the variables, we can compute the range
+for the right hand side (RHS) assuming\footnote{If $\theta'-\theta<0$, then the
+range would be $\frac{1}{\sqrt{N}}[-0.75,1]$ but this does not matter for
+the purposes of our analysis.} that $\theta'-\theta > 0$:
+\begin{equation}\label{eq:rhs_range}
+\min\{{\rm RHS}\} = \frac{-0.5}{\sqrt{N} \cdot 0.5} = -\frac{1}{\sqrt{N}}
+\quad {\rm and} \quad \max\{{\rm RHS}\} = \frac{0.375}{\sqrt{N} \cdot 0.5} = \frac{0.75}{\sqrt{N}}
+\end{equation}
+Thus, the RHS is in $\frac{1}{\sqrt{N}}[-1,0.75]$.  The standard deviation of
+the LHS given the interval constraints is at least $0.5/\sqrt{b}$.
+Consequently, the gap between the LHS and RHS in
+Equation~(\ref{eq:accept_test_rewritten}) is at most $2\sqrt{b/N}$ standard
+deviations, limiting the range in which the test will be able to ``succeed''
+without requiring more samples.
+
+The samples $\theta$, $\theta'$ and $u$ are drawn independently and so the
+probability of the conjunction of these events is $c = p_0 p_1 p_2 = 0.0085$.
+
+\subsection{Proof of Lemma~\ref{lem:quant_clt}}\label{app:quant_clt}
+
+The following bound is given immediately after Corollary 2 from~\cite{explicit-clt05}:
+\begin{equation}
+-6.4\mE[|X|^3]-2\mE[|X|] \leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \leq
+1.36\mE[|X|^3].
+\end{equation}
+This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
+and combining with $x>0$, we obtain the weaker but unqualified bound
+in Equation~(\ref{eq:clt-bounds}).
+
+\subsection{Proof of Lemma~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}
+
+We first observe that
+\[
+    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
+\]
+and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
+\begin{equation}
+-\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
+\end{equation}
+% Daniel: this is the multi-line version, but we should use the full page.
+%\begin{align*}
+%-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
+%&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
+%&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
+%\end{align*}
+as desired.
+
+\subsection{Proof of Corollary~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}
+
+We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
+\begin{equation}
+    P(y) = {\rm Pr}(\Delta^* < y)
+     \quad {\rm and} \quad Q(y) = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
+\end{equation}
+and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
+where $\sigma_n^2 = 1 - s^2_{\Delta^*}$. This yields the next iteration of $P$
+and $Q$:
+\begin{equation}
+    P'(y) = {\rm Pr}(\Delta^*+X_{\rm nc} < y)
+    \quad {\rm and}\quad Q'(y) = \Phi\left({y-\Delta}\right)
+\end{equation}
+Now we convolve with the distribution of $X_{\rm corr}$:
+\begin{equation}
+    P''(y) = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y)
+    \quad {\rm and}\quad Q''(y) = S\left({y-\Delta}\right)
+\end{equation}
+Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
+$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) =
+{\rm Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
+that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
+nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm log} > 0)$
+by at most $\epsilon$.
+
+\subsection{Improved Error Bounds Based on Skew Estimation}\label{app:better_error_bound}
+
+We show that the CLT error bound can be improved to $O(n^{-1})$ using a more
+precise limit distribution under an additional assumption. Let $\mu_i$ denote
+the $i^{th}$ moment, and $b_i$ denote the $i^{th}$ absolute moment of $X$. If
+Cramer's condition holds:
+\begin{equation}\label{eq:cramers_condition}
+    \lim_{t \to \infty} \sup |\mE[\exp(i t X)]| < 1,
+\end{equation}
+then Equation 2.2 in Bentkus et al.'s work on Edgeworth
+expansions~\cite{Bentkus97} provides:
+
+\begin{lemma}\label{lem:clt_edgeworth}
+Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
+random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3.
+If $X$ satisfies Cramer's condition, then
+%\begin{equation}\label{eq:clt-bounds_edgeworth}
+\[
+    \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
+\]
+%\end{equation}
+where
+\begin{equation}
+    G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x).
+\end{equation}
+\end{lemma}
+Lemma~\ref{lem:clt_edgeworth} shows that the average of the $X_i$ has a more
+precise, skewed CDF limit $G_n(x,y)$ where the skew term has weight proportional
+to a certain measure of skew derived from the moments:
+$\mu_3/b_2^{3/2}$. Note that if the $X_i$ are symmetric, the weight of
+the correction term is zero, and the CDF of the average of the $X_i$ converges
+to $\Phi(x)$ at a rate of $O(n^{-1})$.
+
+Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
+$n^{-1/2}$.
+Importantly, since $\phi^{''}(x) = x^2\phi(x) - \phi(x)$ where
+$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
+\begin{equation}\label{eq:GNderivatives}
+    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
+\end{equation}
+From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
+$\phi(x)$ and $\phi^{''}(x)$. In Algorithm 1, we
+correct for the difference in $\sigma$ between $\Delta^*$ and the variance
+needed by $X_{\rm corr}$ using $X_{\rm nc}$. This same method works when we
+wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
+component functions of $G_n(x,y)$ are derivatives of a (unit variance)
+$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
+variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm nc}$ as
+per Algorithm 1 preserving the limit in Equation~(\ref{eq:GNderivatives}).
+
+The deconvolution approach can be used to construct a correction variable
+$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
+additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
+Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
+Equation~(\ref{eq:GNderivatives}), its enough to consider a function $g(x,y')$
+parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
+saved offline. As we have shown earlier, errors in the ``limit'' function
+propagate directly through as errors in the acceptance test.  To achieve a test
+error of $10^{-6}$ (close to single floating point precision), we need a $y'$
+spacing of $10^{-6}$. It should not be necessary to tabulate values all the way to
+$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
+Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
+resolution is 10,000, this leads to a table with about 1 billion values, which
+can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
+in $y$, it should be possible to achieve similar accuracy with a much smaller
+table. We leave further analysis and experiments with $g(x,y)$ as future work.
+
+\section{Additional Experiment Details}\label{app:experiments}
+% Daniel: all three experiments should be described in more detail here. And
+% also the deconvolution step to get the correction distribution.
+
+\subsection{Obtaining the Correction Distribution (Section~\ref{sec:correction})}\label{app:correction}
+
+% Daniel: moved to appendix so we can expand it. Originally we had only two
+% lines, and it looked pretty awkward. So let's add more. Note: this also
+% depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
+\begin{table}[t]
+\caption{Error ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$.}
+\label{tab:xcorr}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{l l l l}
+\hline
+\abovespace\belowspace
+${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error} \\
+\hline
+\abovespace
+4000 & 1.0 & 100    & 1.9e-3  \\
+4000 & 1.0 & 10     & 8.9e-4  \\
+4000 & 1.0 & 1      & 1.6e-3  \\
+4000 & 1.0 & 0.1    & 2.8e-3  \\
+4000 & 1.0 & 0.01   & 5.2e-3  \\
+4000 & 1.0 & 0.001  & 9.9e-3  \\
+2000 & 1.0 & 10     & 1.3e-3  \\
+8000 & 1.0 & 10     & 1.3e-3  \\
+\belowspace
+4000 & 1.0 & 10 & 6.6e-4 \\
+\hline
+\end{tabular}
+\quad
+\begin{tabular}{l l l l}
+\hline
+\abovespace\belowspace
+${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error} \\
+\hline
+\abovespace
+4000 & 0.9 & 1 & 1.0e-4  \\
+\belowspace
+4000 & 0.8 & 0.03 & 5.0e-6 \\
+\hline
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table}
+
+\textbf{Daniel: TODO, figure out what values to test, put optimal values in bold
+face here, etc., note that these are with the range $[-20,20]$ except for the
+best value in $[-10,10]$; maybe we should decrease the range?}
+
+\subsection{Gaussian Mixture Model Experiment (Section~\ref{ssec:gaussians})}\label{app:gaussians}
 
@@
 \end{table}
-
-We next test logistic regression for the binary classification of 1s versus 7s
-on a subset of the MNIST8M dataset, a larger version of
-MNIST~\cite{lecun-mnisthandwrittendigit-2010}. We randomly subsampled 450k
-training and 192k testing points.  We impose a uniform prior on $\theta$ and
-again use a random walk proposer, this time with covariance matrix $0.05I$. The
-temperature is set at $T=1000$. We run the three tests for 3000 samples and
-again use 100 as the starting minibatch size. Our undiscounted implementation
-of~\cite{cutting_mh_2014} used per-test error of $\epsilon=0.05$.
-For~\cite{icml2014c1_bardenet14}, we use their recommended discounting factor
-value of $\gamma=2.0$. We tried to use the symbolic bound for
-$C_{\theta,\theta'}$ for Logistic regression from~\cite{icml2014c1_bardenet14},
-but found it to be too high. Instead we use the empirical $C_{\theta,\theta'}$
-from the entire dataset and do not add its computation time when considering
-runtime, so again our results represent a lower bound for this method.
-
-Figure~\ref{fig:logistic_performance} shows the test log likelihood and
-prediction accuracy as a function of the cumulative training points
-processed.\footnote{The curves do not span the same length over the x-axis since
-the methods consume different amounts of data.} To generate the curves, for each
-of the sampled vectors $\theta_t$, $t\in\{1,\ldots,3000\}$, we use $\theta_t$ as
-the parameter for logistic regression.  Our minibatch MH test is more efficient
-in terms of accuracy, achieving convergence roughly twice as fast compared
-to~\cite{cutting_mh_2014} and about 1.5 orders of magnitude faster
-than~\cite{icml2014c1_bardenet14}, though its log likelihood results lag
-slightly behind the former algorithm during the very early stages of
-exploration.
-
-Figure~\ref{fig:logistic_minibatch} shows histograms of minibatch sizes for the
-three methods on a log-log scale. With an initial size of 100, we achieve an
-average minibatch size of 393, more than two orders of magnitude smaller
-than~\cite{icml2014c1_bardenet14}, and almost an order of magnitude faster than
-the undiscounted implementation of~\cite{cutting_mh_2014}.
-
-\subsection{Collaborative Filtering}\label{ssec:collab_filtering}
-
-\textbf{Daniel: TODO we wanted a third experiment, right?}
-
-\section{Conclusions}\label{sec:conclusion}
-
-\textbf{Daniel: need to revise this to take into account reviewer suggestion,
-with future work ideas. This will add some 2016 references so that'll be good
-(makes it current)}
-
-We have derived an M-H test for minibatch MCMC which approximates full data
-tests. We present theoretical results and experimentally show the benefits of
-our test on Gaussian mixtures and logistic regression. Directions for future
-work include running more experiments with a particular focus on controlling
-variances, testing on neural networks, combining our results with Hamiltonian
-Monte Carlo methods, providing a recipe for how to use our algorithm (following
-the framework of~\cite{sgmcmc_2015}), or integrating parallel
-MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
-
-\bibliography{paper}
-\bibliographystyle{icml2017}
-
-\clearpage
-\appendix
-
-% Daniel: I set the Appendix at single column which makes some of the math
-% easier to write. Other ICML papers from 2016 did this so it's not a big deal.
-
-\onecolumn
-\begin{center}
-{\Large Appendix}
-\end{center}
-
-This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
-the proofs that we omitted from the main text due to space constraints.
-Appendix~\ref{app:experiments} provides further details on our experiments to
-assist understanding and reproducibility.
-
-\section{Proofs of Lemmas and Corollaries}\label{app:proofs}
-
-\subsection{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}
-
-Choose $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ (event 1) and
-$(\theta -0.5) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ filtered for  matching sign
-(event 2).  As discussed in Lemma~\ref{lem:worst_case}, both $q(\theta' |
-\theta)$ and $p(\theta | x_1,\ldots,x_N)$ have variance $1/N$. If we denote
-$\Phi$ as the CDF of the standard normal distribution, then the former event
-occurs with probability $p_0 = 2(\Phi(\sqrt{N}\frac{1}{\sqrt{N}}) -
-\Phi(\sqrt{N}\frac{0.5}{\sqrt{N}})) = 2(\Phi(1)-\Phi(0.5)) \approx 0.2997$. The
-latter event, because we restrict signs, occurs with probability $p_1 =
-\Phi(1)-\Phi(0.5) \approx 0.14988$.
-
-These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative by
-inspection of equation (\ref{eq:whyneg}) below.  This implies that we can find a
-$u \in (0,1)$ so that $\psi(u,\theta,\theta') = \log u < 0$ equals
-$\mE[\Lambda^*(\theta,\theta')]$.  Specifically, choose $u_0$ to satisfy $\log u_0
-= \mE[\Lambda^*(\theta,\theta')]$.  Using $\mE[x_i^*] = 0.5$ and
-Equation~(\ref{eq:lemma_ll_ratio}), we see that
-\begin{equation}
-  \log u_0 = N(\theta'-\theta)\frac{1}{b} \cdot \mE\left[\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2}\right]
-\end{equation}
-\begin{equation}\label{eq:whyneg}
-    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
-\end{equation}
-Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta')
-\not\approx \psi(u,\theta,\theta')$ used in ~\cite{cutting_mh_2014}
-and~\cite{icml2014c1_bardenet14} , where $\not\approx$ means ``significantly
-different from'' under the distribution over samples. This is
-\begin{align}
-\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') 
-&\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2} \not\approx \log u_0\\
-&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \not\approx  0 \\
-&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-0.5 \not\approx 0. \label{eq:accept_test_zero}
-\end{align}
-Since the $x_i^*$ have mean 0.5, the resulting test with our chosen $u_0$ will
-never correctly succeed and must use all $N$ data points.  Furthermore, if we
-sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be
-sufficiently different from 0.5 to allow the test to succeed.
-
-The choices above for $\theta$ and $\theta'$ guarantee that
-\begin{equation}\label{eq:log_uo_range}
-    \log u_0 \in -[0.5,1][0.75,1.5] = [-1.5, -0.375].
-\end{equation}
-Next, consider the range of $u$ values near $u_0$:
-\begin{equation}\label{eq:log_u_range}
-    \log u \in \log u_0 + [-0.5,0.375].
-\end{equation}
-The size of the range in $u$ is at least $\exp([-2,-1.125]) \approx
-[0.13534,0.32465]$ and occurs with probability at least $p_2=0.18932$. With $u$
-in this range, we rewrite the test as:
-\begin{equation}\label{eq:accept_test_rewritten}
-    \frac{1}{b}\sum_{i=1}^b x_i^*-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
-\end{equation}
-so that, as in Equation~(\ref{eq:accept_test_zero}), the LHS has expected value
-zero.  Given our choice of intervals for the variables, we can compute the range
-for the right hand side (RHS) assuming\footnote{If $\theta'-\theta<0$, then the
-range would be $\frac{1}{\sqrt{N}}[-0.75,1]$ but this does not matter for
-the purposes of our analysis.} that $\theta'-\theta > 0$:
-\begin{equation}\label{eq:rhs_range}
-\min\{{\rm RHS}\} = \frac{-0.5}{\sqrt{N} \cdot 0.5} = -\frac{1}{\sqrt{N}}
-\quad {\rm and} \quad \max\{{\rm RHS}\} = \frac{0.375}{\sqrt{N} \cdot 0.5} = \frac{0.75}{\sqrt{N}}
-\end{equation}
-Thus, the RHS is in $\frac{1}{\sqrt{N}}[-1,0.75]$.  The standard deviation of
-the LHS given the interval constraints is at least $0.5/\sqrt{b}$.
-Consequently, the gap between the LHS and RHS in
-Equation~(\ref{eq:accept_test_rewritten}) is at most $2\sqrt{b/N}$ standard
-deviations, limiting the range in which the test will be able to ``succeed''
-without requiring more samples.
-
-The samples $\theta$, $\theta'$ and $u$ are drawn independently and so the
-probability of the conjunction of these events is $c = p_0 p_1 p_2 = 0.0085$.
-
-\subsection{Proof of Lemma~\ref{lem:quant_clt}}\label{app:quant_clt}
-
-The following bound is given immediately after Corollary 2 from~\cite{explicit-clt05}:
-\begin{equation}
--6.4\mE[|X|^3]-2\mE[|X|] \leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \leq
-1.36\mE[|X|^3].
-\end{equation}
-This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
-and combining with $x>0$, we obtain the weaker but unqualified bound
-in Equation~(\ref{eq:clt-bounds}).
-
-\subsection{Proof of Lemma~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}
-
-We first observe that
-\[
-    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
-\]
-and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
-\begin{equation}
--\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
-\end{equation}
-% Daniel: this is the multi-line version, but we should use the full page.
-%\begin{align*}
-%-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
-%&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
-%&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
-%\end{align*}
-as desired.
-
-\subsection{Proof of Corollary~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}
-
-We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
-\begin{equation}
-    P(y) = {\rm Pr}(\Delta^* < y)
-     \quad {\rm and} \quad Q(y) = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
-\end{equation}
-and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
-where $\sigma_n^2 = 1 - s^2_{\Delta^*}$. This yields the next iteration of $P$
-and $Q$:
-\begin{equation}
-    P'(y) = {\rm Pr}(\Delta^*+X_{\rm nc} < y)
-    \quad {\rm and}\quad Q'(y) = \Phi\left({y-\Delta}\right)
-\end{equation}
-Now we convolve with the distribution of $X_{\rm corr}$:
-\begin{equation}
-    P''(y) = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y)
-    \quad {\rm and}\quad Q''(y) = S\left({y-\Delta}\right)
-\end{equation}
-Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
-$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) =
-{\rm Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
-that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
-nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm log} > 0)$
-by at most $\epsilon$.
-
-\subsection{Improved Error Bounds Based on Skew Estimation}\label{app:better_error_bound}
-
-We show that the CLT error bound can be improved to $O(n^{-1})$ using a more
-precise limit distribution under an additional assumption. Let $\mu_i$ denote
-the $i^{th}$ moment, and $b_i$ denote the $i^{th}$ absolute moment of $X$. If
-Cramer's condition holds:
-\begin{equation}\label{eq:cramers_condition}
-    \lim_{t \to \infty} \sup |\mE[\exp(i t X)]| < 1,
-\end{equation}
-then Equation 2.2 in Bentkus et al.'s work on Edgeworth
-expansions~\cite{Bentkus97} provides:
-
-\begin{lemma}\label{lem:clt_edgeworth}
-Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
-random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3.
-If $X$ satisfies Cramer's condition, then
-%\begin{equation}\label{eq:clt-bounds_edgeworth}
-\[
-    \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
-\]
-%\end{equation}
-where
-\begin{equation}
-    G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x).
-\end{equation}
-\end{lemma}
-Lemma~\ref{lem:clt_edgeworth} shows that the average of the $X_i$ has a more
-precise, skewed CDF limit $G_n(x,y)$ where the skew term has weight proportional
-to a certain measure of skew derived from the moments:
-$\mu_3/b_2^{3/2}$. Note that if the $X_i$ are symmetric, the weight of
-the correction term is zero, and the CDF of the average of the $X_i$ converges
-to $\Phi(x)$ at a rate of $O(n^{-1})$.
-
-Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
-$n^{-1/2}$.
-Importantly, since $\phi^{''}(x) = x^2\phi(x) - \phi(x)$ where
-$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
-\begin{equation}\label{eq:GNderivatives}
-    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
-\end{equation}
-From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
-$\phi(x)$ and $\phi^{''}(x)$. In Algorithm 1, we
-correct for the difference in $\sigma$ between $\Delta^*$ and the variance
-needed by $X_{\rm corr}$ using $X_{\rm nc}$. This same method works when we
-wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
-component functions of $G_n(x,y)$ are derivatives of a (unit variance)
-$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
-variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm nc}$ as
-per Algorithm 1 preserving the limit in Equation~(\ref{eq:GNderivatives}).
-
-The deconvolution approach can be used to construct a correction variable
-$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
-additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
-Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
-Equation~(\ref{eq:GNderivatives}), its enough to consider a function $g(x,y')$
-parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
-saved offline. As we have shown earlier, errors in the ``limit'' function
-propagate directly through as errors in the acceptance test.  To achieve a test
-error of $10^{-6}$ (close to single floating point precision), we need a $y'$
-spacing of $10^{-6}$. It should not be necessary to tabulate values all the way to
-$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
-Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
-resolution is 10,000, this leads to a table with about 1 billion values, which
-can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
-in $y$, it should be possible to achieve similar accuracy with a much smaller
-table. We leave further analysis and experiments with $g(x,y)$ as future work.
-
-\section{Additional Experiment Details}\label{app:experiments}
-% Daniel: all three experiments should be described in more detail here.
-
-\subsection{Gaussian Mixture Model Experiment (Section~\ref{ssec:gaussians})}\label{app:gaussians}
 

### 2017-02-06 icml2017/paper.tex
Updated the paper to have lots of tables. I think now we can decide which sigma, lambda, etc. to set for the correction distribution
More updates to paper.
I added in a reference that a reviewer suggested. OK now focus on the BIDMach implementation...

@@
 The amount of data consumed for each test varies significantly from one
-minibatch to the next. By contrast,~\cite{conf/uai/MaclaurinA14,TallData15}
+minibatch to the next. By contrast,~\cite{conf/uai/MaclaurinA14,TallData16}
 perform exact tests but require a lower bound on the parameter distribution across
@@
 Barker's test function was suggested but not explored empirically
-in~\cite{TallData15} section 6.3. But the asymptotic test statistic CDF and the
+in~\cite{TallData16} section 6.3. But the asymptotic test statistic CDF and the
 Barker function are different, which leads to fixed errors for the approach
-in~\cite{TallData15}. Here, we show that the difference between the
+in~\cite{TallData16}. Here, we show that the difference between the
 distributions can be corrected with an additive random variable. This leads to a
@@
 test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
-random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{TallData15} it
+random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{TallData16} it
 was proposed to use $\Delta^*$ in a Barker test, and tolerate the fixed
@@
 \textbf{Daniel: need to revise this to take into account reviewer suggestion,
-with future work ideas. This will add some 2016 references so that'll be good
-(makes it current)}
+with future work ideas. One reviewer suggested
+using~\cite{DBLP:conf/icml/ChenG16}.}
 
@@
 \begin{table}[t]
-\caption{Error ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm log}$.}
+\caption{Errors ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm
+log}$, with $N=4000$ (top row) and $N=2000$ (bottom row).}
 \label{tab:xcorr}
@@
 \begin{sc}
-\begin{tabular}{l l l l}
-\hline
-\abovespace\belowspace
-${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error} \\
+%% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=4000$ & $\sigma=1.1$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
 \hline
 \abovespace
-4000 & 1.0 & 100    & 1.9e-3  \\
-4000 & 1.0 & 10     & 8.9e-4  \\
-4000 & 1.0 & 1      & 1.6e-3  \\
-4000 & 1.0 & 0.1    & 2.8e-3  \\
-4000 & 1.0 & 0.01   & 5.2e-3  \\
-4000 & 1.0 & 0.001  & 9.9e-3  \\
-2000 & 1.0 & 10     & 1.3e-3  \\
-8000 & 1.0 & 10     & 1.3e-3  \\
+100    & \textbf{4.3e-3}  \\
+10     & 6.0e-3  \\
+1      & 1.0e-2  \\
+0.1    & 1.2e-2  \\
 \belowspace
-4000 & 1.0 & 10 & 6.6e-4 \\
+0.01   & 3.5e-2  \\
 \hline
 \end{tabular}
-\quad
-\begin{tabular}{l l l l}
-\hline
-\abovespace\belowspace
-${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error} \\
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=4000$ & $\sigma=1.0$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
 \hline
 \abovespace
-4000 & 0.9 & 1 & 1.0e-4  \\
+100    & 1.9e-3  \\
+10     & \textbf{8.9e-4}  \\
+1      & 1.6e-3  \\
+0.1    & 2.8e-3  \\
 \belowspace
-4000 & 0.8 & 0.03 & 5.0e-6 \\
+0.01   & 5.2e-3  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=4000$ & $\sigma=0.9$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 1.2e-3  \\
+10     & 2.6e-4  \\
+1      & \textbf{1.0e-4}  \\
+0.1    & 2.0e-4  \\
+\belowspace
+0.01   & 3.9e-4  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=4000$ & $\sigma=0.8$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 8.3e-4  \\
+10     & 1.3e-4  \\
+1      & 2.5e-5  \\
+0.1    & \textbf{6.7e-6}  \\
+\belowspace
+0.01   & 7.4e-6  \\
+\hline
+\end{tabular}
+%% ---- NEW ROW ----
+\vskip 0.2in %% OH this works!
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=2000$ & $\sigma=1.1$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 6.8e-3  \\
+10     & \textbf{4.6e-3}  \\
+1      & 7.5e-3  \\
+0.1    & 1.3e-2  \\
+\belowspace
+0.01   & 2.4e-2  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=2000$ & $\sigma=1.0$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 4.4e-3  \\
+10     & 1.3e-3  \\
+1      & \textbf{1.1e-3}  \\
+0.1    & 2.0e-3  \\
+\belowspace
+0.01   & 3.6e-3  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=2000$ & $\sigma=0.9$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 3.3e-3  \\
+10     & 6.4e-4  \\
+1      & 1.6e-4  \\
+0.1    & \textbf{1.3e-4}  \\
+\belowspace
+0.01   & 2.7e-4  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=2000$ & $\sigma=0.8$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 2.6e-3  \\
+10     & 4.0e-4  \\
+1      & 6.7e-5  \\
+0.1    & 1.4e-5  \\
+\belowspace
+0.01   & \textbf{5.0e-6}  \\
 \hline

### 2017-02-07 icml2017/paper.tex
Updated Appendix B.1
Got the table re-sorted and a description there.

@@
 
-\textbf{Daniel: need to revise this to take into account reviewer suggestion,
-with future work ideas. One reviewer suggested
-using~\cite{DBLP:conf/icml/ChenG16}.}
+\textbf{Daniel: double check this in case we add a third experiment.}
 
@@
 our test on Gaussian mixtures and logistic regression.  Directions for future
-work include running more experiments with a particular focus on controlling
-variances, testing on neural networks, combining our results with Hamiltonian
-Monte Carlo methods, providing a recipe for how to use our algorithm (following
-the framework of~\cite{sgmcmc_2015}), or integrating parallel
-MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
+work include integrating our algorithm with~\cite{cutting_mh_2014} by applying
+both tests each iteration or utilizing the variance reduction techniques
+suggested in~\cite{DBLP:conf/icml/ChenG16}.  More elaborate extensions involve
+running more experiments on other models such as neural networks and providing a
+recipe for how to use our algorithm (following the framework
+of~\cite{sgmcmc_2015}).
 
@@
 \abovespace\belowspace
+$N=2000$ & $\sigma=0.8$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 2.6e-3  \\
+10     & 4.0e-4  \\
+1      & 6.7e-5  \\
+0.1    & 1.4e-5  \\
+\belowspace
+0.01   & \textbf{5.0e-6}  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=2000$ & $\sigma=0.9$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 3.3e-3  \\
+10     & 6.4e-4  \\
+1      & 1.6e-4  \\
+0.1    & \textbf{1.3e-4}  \\
+\belowspace
+0.01   & 2.7e-4  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=2000$ & $\sigma=1.0$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 4.4e-3  \\
+10     & 1.3e-3  \\
+1      & \textbf{1.1e-3}  \\
+0.1    & 2.0e-3  \\
+\belowspace
+0.01   & 3.6e-3  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=2000$ & $\sigma=1.1$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 6.8e-3  \\
+10     & \textbf{4.6e-3}  \\
+1      & 7.5e-3  \\
+0.1    & 1.3e-2  \\
+\belowspace
+0.01   & 2.4e-2  \\
+\hline
+\end{tabular}
+%% ---- NEW ROW ----
+\vskip 0.2in %% OH this works!
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=4000$ & $\sigma=0.8$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 8.3e-4  \\
+10     & 1.3e-4  \\
+1      & 2.5e-5  \\
+0.1    & \textbf{6.7e-6}  \\
+\belowspace
+0.01   & 7.4e-6  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=4000$ & $\sigma=0.9$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 1.2e-3  \\
+10     & 2.6e-4  \\
+1      & \textbf{1.0e-4}  \\
+0.1    & 2.0e-4  \\
+\belowspace
+0.01   & 3.9e-4  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
+$N=4000$ & $\sigma=1.0$ \\
+\hline
+\abovespace\belowspace
+$\lambda$ & $L_{\infty}$ error \\
+\hline
+\abovespace
+100    & 1.9e-3  \\
+10     & \textbf{8.9e-4}  \\
+1      & 1.6e-3  \\
+0.1    & 2.8e-3  \\
+\belowspace
+0.01   & 5.2e-3  \\
+\hline
+\end{tabular}
+\quad %% ----------------------- START OF TABLE
+\begin{tabular}{l l}
+\hline
+\abovespace\belowspace
 $N=4000$ & $\sigma=1.1$ \\
@@
 \end{tabular}
-\quad %% ----------------------- START OF TABLE
-\begin{tabular}{l l}
-\hline
-\abovespace\belowspace
-$N=4000$ & $\sigma=1.0$ \\
-\hline
-\abovespace\belowspace
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-\abovespace
-100    & 1.9e-3  \\
-10     & \textbf{8.9e-4}  \\
-1      & 1.6e-3  \\
-0.1    & 2.8e-3  \\
-\belowspace
-0.01   & 5.2e-3  \\
-\hline
-\end{tabular}
-\quad %% ----------------------- START OF TABLE
-\begin{tabular}{l l}
-\hline
-\abovespace\belowspace
-$N=4000$ & $\sigma=0.9$ \\
-\hline
-\abovespace\belowspace
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-\abovespace
-100    & 1.2e-3  \\
-10     & 2.6e-4  \\
-1      & \textbf{1.0e-4}  \\
-0.1    & 2.0e-4  \\
-\belowspace
-0.01   & 3.9e-4  \\
-\hline
-\end{tabular}
-\quad %% ----------------------- START OF TABLE
-\begin{tabular}{l l}
-\hline
-\abovespace\belowspace
-$N=4000$ & $\sigma=0.8$ \\
-\hline
-\abovespace\belowspace
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-\abovespace
-100    & 8.3e-4  \\
-10     & 1.3e-4  \\
-1      & 2.5e-5  \\
-0.1    & \textbf{6.7e-6}  \\
-\belowspace
-0.01   & 7.4e-6  \\
-\hline
-\end{tabular}
-%% ---- NEW ROW ----
-\vskip 0.2in %% OH this works!
-\begin{tabular}{l l}
-\hline
-\abovespace\belowspace
-$N=2000$ & $\sigma=1.1$ \\
-\hline
-\abovespace\belowspace
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-\abovespace
-100    & 6.8e-3  \\
-10     & \textbf{4.6e-3}  \\
-1      & 7.5e-3  \\
-0.1    & 1.3e-2  \\
-\belowspace
-0.01   & 2.4e-2  \\
-\hline
-\end{tabular}
-\quad %% ----------------------- START OF TABLE
-\begin{tabular}{l l}
-\hline
-\abovespace\belowspace
-$N=2000$ & $\sigma=1.0$ \\
-\hline
-\abovespace\belowspace
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-\abovespace
-100    & 4.4e-3  \\
-10     & 1.3e-3  \\
-1      & \textbf{1.1e-3}  \\
-0.1    & 2.0e-3  \\
-\belowspace
-0.01   & 3.6e-3  \\
-\hline
-\end{tabular}
-\quad %% ----------------------- START OF TABLE
-\begin{tabular}{l l}
-\hline
-\abovespace\belowspace
-$N=2000$ & $\sigma=0.9$ \\
-\hline
-\abovespace\belowspace
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-\abovespace
-100    & 3.3e-3  \\
-10     & 6.4e-4  \\
-1      & 1.6e-4  \\
-0.1    & \textbf{1.3e-4}  \\
-\belowspace
-0.01   & 2.7e-4  \\
-\hline
-\end{tabular}
-\quad %% ----------------------- START OF TABLE
-\begin{tabular}{l l}
-\hline
-\abovespace\belowspace
-$N=2000$ & $\sigma=0.8$ \\
-\hline
-\abovespace\belowspace
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-\abovespace
-100    & 2.6e-3  \\
-10     & 4.0e-4  \\
-1      & 6.7e-5  \\
-0.1    & 1.4e-5  \\
-\belowspace
-0.01   & \textbf{5.0e-6}  \\
-\hline
-\end{tabular}
 \end{sc}
@@
 
-\textbf{Daniel: TODO, figure out what values to test, put optimal values in bold
-face here, etc., note that these are with the range $[-20,20]$ except for the
-best value in $[-10,10]$; maybe we should decrease the range?}
+In Section~\ref{sec:correction}, we described our derivation of the correction
+distribution $C_\sigma$ for random variable $X_{\rm corr}$.
+Table~\ref{tab:xcorr} shows our $L_\infty$ error results for the convolution
+(Equation~(\ref{eq:overall_corr_problem})) based on various hyperparameter
+choices.  We test using $N=2000$ and $N=4000$ points for discretization within
+a range of $X_{\rm corr} \in [-20,20]$, covering essentially all the probability
+mass. We also vary $\sigma$ from 0.8 to 1.1.
+
+We observe a tradeoff. With smaller $\sigma$, our $C_\sigma$ is closer to the
+ideal distribution (as judged by $L_\infty$ error), but this imposes a stricter
+upper bound on the sample variance of $\Delta^*$ before our test can be applied,
+which thus results in larger minibatch sizes. Conversely, a more liberal upper
+bound means we avail ourselves of smaller minibatch sizes, but at the cost of a
+less stable derivation for $C_\sigma$.
+
+\textbf{We chose XXXX to use in our experiments, which empirically provided an
+acceptable tradeoff to us.} This is reflected in
+Algorithm~\ref{alg:our_algorithm}, which assumes that we used \textbf{TODO TODO
+TODO TODO TODO} but we reiterate that the choice is arbitrary so long as $0 <
+\sigma < \sqrt{\pi^2/3} \approx 1.814$, the standard deviation of the standard
+logistic distribution, since there must be some variance left over for $X_{\rm
+corr}$.
 

### 2017-02-20 yutian_chen/tabulate_values.m
Adding matlab file, now mu_std in [-60,60]

@@ -35,9 +35,9 @@
 %   CHECK THE OUTPUT FILE NAMES! Be careful to avoid over-writing stuff.
 
 % ratio: use 100/100000 for MNIST8M, 100/13000 for MNIST. Change file name.
-ratio = 100 / 100000; 
-K = 20;
-D = 15000;
+ratio = 100 / 13000; 
+K = 60;
+D = 25000;
 max_tests = 1. / ratio;
 assert(K > 0);
 epsilons = [0.001, 0.005, 0.01, 0.05, 0.1, 0.2];
@@ -47,7 +47,7 @@ for e = 1:length(epsilons)
     result = zeros(D, 3);
     eps = epsilons(e);
     % CHECK for mnist, mnist8m, etc., in the file name!
-    outfile_name = sprintf('mu_std_K%d_D%d_mnist8m_eps%.4f.mat', K, D, eps)
+    outfile_name = sprintf('mu_std_K%d_D%d_mnist_eps%.4f.mat', K, D, eps)
     fprintf('\tRunning tabulate_values using K = %d, D = %d\n', K, D);
     fprintf('\tratio = %f, max_tests = %d, eps = %f\n', ratio, max_tests, eps);
 

### 2017-02-22 icml2017/paper.tex
Adding figure drafts to where they would go in the paper.
I'm still using ICML format because UAI doesn't have their format out yet.

@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{logistic_regression_ll_acc_results.png}
+	\includegraphics[width=1.0\linewidth]{logistic_regression_error_regions.png}
 	\caption{
-    Logistic regression performance (accuracy/log likelihood) based on
-    cumulative data usage.
+    TODO
     }
@@
 \end{figure*}
+
 \begin{figure*}[t]
 	\centering
-	\includegraphics[width=0.9\linewidth]{logistic_regression_histograms_results.png}
+    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist.png}
 	\caption{
-    Counts of minibatch sizes in the logistic regression experiment (analogous
-    to Figure~\ref{fig:gauss_mix_2}).
+    Minibatch sizes used for one representative trial in
+    Section~\ref{ssec:logistic} using MNIST-13k (analogous to
+    Figure~\ref{fig:gauss_mix_2}).
     }
-	\label{fig:logistic_minibatch}
+	\label{fig:logistic_histograms_mnist}
 \end{figure*}
+
+\begin{figure*}[t]
+	\centering
+    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist8m.png}
+	\caption{
+    Minibatch sizes used for one representative trial in
+    Section~\ref{ssec:logistic} using MNIST8M-100k (similar to
+    Figure~\ref{fig:logistic_histograms_mnist}).
+    }
+	\label{fig:logistic_histograms_mnist8m}
+\end{figure*}
+
+\begin{table}[t]
+\caption{Log. Regression Average Minibatch Size (10 Trials)}
+\label{tab:logistic}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{l l l l}
+\hline
+\abovespace\belowspace
+Method/Data & MNIST-13k & MNIST8M-100k  \\
+\hline
+\abovespace
+Our Method & $129.8\pm 3.4$ & \textbf{TODO} \\
+AustereMH  & \textbf{TODO} & \textbf{TODO} \\
+\belowspace
+MHSubLhd   & $10752.8\pm 92.9$ & \textbf{TODO} \\
+\hline
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table}
+
+\textbf{Daniel: TODO this section is badly out of date but I will re-write it
+and it should be much better than what we had earlier. Also, define a trial
+since I say it in the table and figures. Finally, define AustereMH and MHSubLhd,
+and come up with a name for our method.}
 

### 2017-02-22 yutian_chen/tabulate_values.m
Removed a lot of outdated matlab file, have one with mu_std = 0 now.

@@ -6,10 +6,16 @@
 %   D: The number of points within [-K,K] to use for \mu_std values.
 %
 % What it saves:
-%   For now I will only be saving the `err` and `mean_j` terms. The `mean_j`
-%   value will be useful to compute \bar{\pi} estimates given the initial batch
-%   of, say, 100 elements. We can simply do 100 * mean_j. As for the error term,
-%   we'll figure out a good value to use empirically.
+%
+%   result(a,b,c) = (mu_std(a), err, mean_j)
+%
+%   where:
+%       a = mu_std index
+%       b = epsilon index (note: this is the value we pick for one
+%           accept/decision trial, it is NOT the error of the overall test
+%           (which is a _sequence_ of hypothesis tets))
+%       c = minibatch size (starting *and* the increment amount)
+%   (results is 4-D, the last index is 1, 2, or 3)
 %
 % Usage: 
 %   I just run it on my command line:
@@ -20,49 +26,60 @@
 %   watch out for the outfile_name! Set it to be something informative.
 % 
 % Notes:
-%   For logistic regression with MNIST(8M), we use a fixed minibatch size (and
-%   increment value) of 100. For regular MNIST, the training data has size 13000
-%   and for MNIST8M, we'll use 100k; otherwise anything larger means the
-%   competing methods will take absurdly long.
 %
-%   We have to fix an epsilon here (upper bound on error for each individual
-%   test). I will run with {0.001, 0.005, 0.01, 0.05, 0.1, 0.2} which roughly
-%   mirrors the values that were used in their paper. I'll run all of these
-%   together here.
-%
-%   I'll also be using their default discretization for bins.
+%   I'll be using their default discretization for bins.
 %
 %   CHECK THE OUTPUT FILE NAMES! Be careful to avoid over-writing stuff.
 
-% ratio: use 100/100000 for MNIST8M, 100/13000 for MNIST. Change file name.
-ratio = 100 / 13000; 
-K = 60;
-D = 25000;
-max_tests = 1. / ratio;
-assert(K > 0);
+% number_total_data: use 100000 for MNIST8M, 13000 for MNIST. Change file name!!
+number_total_data = 13000;
+K = 10;
 epsilons = [0.001, 0.005, 0.01, 0.05, 0.1, 0.2];
[…]

### 2017-02-22 yutian_chen/tabulate_values.m
Adding two matlabb files
one for MNIST 8m, with only mu_std = 0, and another with MNIST, but with 4000 different mu_0 values.

@@ -32,29 +32,29 @@
 %   CHECK THE OUTPUT FILE NAMES! Be careful to avoid over-writing stuff.
 
 % number_total_data: use 100000 for MNIST8M, 13000 for MNIST. Change file name!!
-number_total_data = 13000;
+number_total_data = 100000;
 K = 10;
 epsilons = [0.001, 0.005, 0.01, 0.05, 0.1, 0.2];
-sizes = [50, 100, 150, 200, 250, 300, 350, 400, 450, 500];
+sizes = [100, 200, 300, 400, 500];
 
 % If running for all mu_std values, set this to 0. Else, 1. This is mainly in
 % case I want to try finer-grained sizes and/or epsilons for the single mu_std=0
 % case. Otherwise with all the mu_std's, we can't afford to have too many
 % epsilon and size options for computational reasons.
-do_only_one = 1;
+do_only_one = 0;
 
 if do_only_one == 1
     D = 1;
     mu_std_values = [0];
     result = zeros(1, length(epsilons), length(sizes), 3);
 else 
-    D = 5000;
+    D = 4000;
     mu_std_values = linspace(-K, K, D);
     result = zeros(D, length(epsilons), length(sizes), 3);
 end
 
 % CHANGE FILE NAME IF NEEDED!!
-outfile_name = sprintf('mu_std_K%d_D%d_mnist.mat', K, D)
+outfile_name = sprintf('mu_std_K%d_D%d_mnist8m.mat', K, D)
 
 fprintf('\tRunning tabulate_values using K = %d, D = %d\n', K, D);
 for e = 1:length(epsilons)

### 2017-02-23 yutian_chen/tabulate_values.m
Updated tabulate so it saves correctly
Ie so I can load them in BIDMach

@@ -34,19 +34,20 @@
 % number_total_data: use 100000 for MNIST8M, 13000 for MNIST. Change file name!!
 number_total_data = 100000;
 K = 10;
+sizes = [50, 100, 150, 200, 250, 300, 350, 400, 450, 500];
 epsilons = [0.001, 0.005, 0.01, 0.05, 0.1, 0.2];
-sizes = [100, 200, 300, 400, 500];
 
 % If running for all mu_std values, set this to 0. Else, 1. This is mainly in
 % case I want to try finer-grained sizes and/or epsilons for the single mu_std=0
 % case. Otherwise with all the mu_std's, we can't afford to have too many
 % epsilon and size options for computational reasons.
-do_only_one = 0;
+do_only_one = 1;
 
 if do_only_one == 1
     D = 1;
     mu_std_values = [0];
-    result = zeros(1, length(epsilons), length(sizes), 3);
+    result_error = zeros(length(epsilons), length(sizes));
+    result_meanj = zeros(length(epsilons), length(sizes));
 else 
     D = 4000;
     mu_std_values = linspace(-K, K, D);
@@ -57,29 +58,34 @@ end
 outfile_name = sprintf('mu_std_K%d_D%d_mnist8m.mat', K, D)
 
 fprintf('\tRunning tabulate_values using K = %d, D = %d\n', K, D);
-for e = 1:length(epsilons)
-    eps = epsilons(e);
-
-    for j = 1:length(sizes)
-        m = sizes(j);
-
-        ratio = m / number_total_data; 
-        max_tests = ceil(1. / ratio);
-        fprintf('ratio = %f, max_tests = %d, eps = %f, m = %d\n', ...
-            ratio, max_tests, eps, m);
-
-        for i = 1:length(mu_std_values)
-            if mod(i,100) == 0
-                fprintf('i = %d, mu_std = %f\n', i, mu_std_values(i));
-            end
+for i = 1:length(mu_std_values)
+    if mod(i,100) == 0
+        fprintf('i = %d, mu_std = %f\n', i, mu_std_values(i));
+    end
+    for e = 1:length(epsilons)
+        eps = epsilons(e);
+        for j = 1:length(sizes)
+            m = sizes(j);
+            ratio = m / number_total_data; 
+            max_tests = ceil(1. / ratio);
             [err, mean_j, p_pos, p_neg, p_j] = seq_test_dynprog(mu_std_values(i), ...
                 ratio, max_tests, norminv(1 - eps));
[…]

### 2017-02-24 yutian_chen/tabulate_values.m
Updating tabulate code and added revised MATLAB files.
This will let me run the non-conservative version of Korattikara (AustereMH).

@@ -7,15 +7,8 @@
 %
 % What it saves:
 %
-%   result(a,b,c) = (mu_std(a), err, mean_j)
-%
-%   where:
-%       a = mu_std index
-%       b = epsilon index (note: this is the value we pick for one
-%           accept/decision trial, it is NOT the error of the overall test
-%           (which is a _sequence_ of hypothesis tets))
-%       c = minibatch size (starting *and* the increment amount)
-%   (results is 4-D, the last index is 1, 2, or 3)
+%   If do_only_one = 1, then it saves two matrices, result_meanj and
+%   result_error, which we can directly analyze in BIDMach.
 %
 % Usage: 
 %   I just run it on my command line:
@@ -36,12 +29,13 @@ number_total_data = 100000;
 K = 10;
 sizes = [50, 100, 150, 200, 250, 300, 350, 400, 450, 500];
 epsilons = [0.001, 0.005, 0.01, 0.05, 0.1, 0.2];
+offset_meanj = length(sizes) * length(epsilons);
 
 % If running for all mu_std values, set this to 0. Else, 1. This is mainly in
 % case I want to try finer-grained sizes and/or epsilons for the single mu_std=0
 % case. Otherwise with all the mu_std's, we can't afford to have too many
 % epsilon and size options for computational reasons.
-do_only_one = 1;
+do_only_one = 0;
 
 if do_only_one == 1
     D = 1;
@@ -49,9 +43,9 @@ if do_only_one == 1
     result_error = zeros(length(epsilons), length(sizes));
     result_meanj = zeros(length(epsilons), length(sizes));
 else 
-    D = 4000;
+    D = 4001;
     mu_std_values = linspace(-K, K, D);
-    result = zeros(D, length(epsilons), length(sizes), 3);
+    result = zeros(D, 2*offset_meanj); % for ncols, need offset doubled
 end
 
 % CHANGE FILE NAME IF NEEDED!!
@@ -74,9 +68,10 @@ for i = 1:length(mu_std_values)
                 result_error(e, j) = err;
                 result_meanj(e, j) = mean_j;
             else
-                result(i, e, j, 1) = mu_std_values(i);
-                result(i, e, j, 2) = err;
-                result(i, e, j, 3) = mean_j;
+                % Stores row by row of err/mean_j matrices into a single row.
+                index = (e-1)*length(sizes) + j; % e-1 due to 1-indexing
+                result(i, index) = err;
[…]

### 2017-02-24 uai2017/paper.tex

--- 2017-02-22 icml2017/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%% ICML 2017 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{article}
+% Sample LaTeX file for creating a paper in the Morgan Kaufmannn two
+% column, 8 1/2 by 11 inch proceedings format.
+
+\documentclass[letterpaper]{article}
+\usepackage{proceed2e}
+\usepackage[margin=1in]{geometry}
 \usepackage{times,graphicx,subfigure,natbib,algorithm,algorithmic,hyperref}
@@
 
-% Employ the following version of the ``usepackage'' statement for
-% submitting the draft version of the paper for review.  This will set
-% the note in the first column to ``Under review.  Do not distribute.''
-\usepackage{icml2017}
-
-% Employ this version of the ``usepackage'' statement after the paper has
-% been accepted, when creating the final version.  This will set the
-% note in the first column to ``Proceedings of the...''
-%\usepackage[accepted]{icml2017}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Submission and Formatting Instructions for ICML 2017}
+\title{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
+
+% Daniel Seita: upon submission, comment out all authors, un-comment the
+% following two lines:
+%\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
+%          % UAI  reviewing is double-blind.
+
+% The author names and affiliations should appear only in the accepted paper.
+%% \author{ {\bf Daniel Seita\thanks{Footnote for author to give an
+%% alternate address.}} \\
+%% Computer Science Dept. \\
+%% Cranberry University\\
+%% Pittsburgh, PA 15213 \\
+%% \And
+%% {\bf Xinlei Pan}  \\
+%% Affiliation          \\
+%% Address \\
+%% \And
+%% {\bf Haoyu Chen}   \\
+%% Affiliation \\
+%% Address    \\
+%% \And
+%% {\bf John Canny}   \\
+%% Affiliation \\
+%% Address    \\
+%% }
+\author{
+Daniel Seita$^1$, 
+Xinlei Pan$^1$,
+Haoyu Chen$^1$,
+John Canny$^{1,2}$ \\
+$^1$ University of California, Berkeley, CA \\
+$^2$ Google Research, Mountain View, CA\\ 
+\texttt{\{seita,xinleipan,haoyuchen,canny\}@berkeley.edu}
+}
 
@@
 
-\twocolumn[
-\icmltitle{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
-
-% It is OKAY to include author information, even for blind submissions: the
-% style file will automatically remove it for you unless you've provided the
-% [accepted] option to the icml2017 package.
-
-% list of affiliations. the first argument should be a (short) identifier you
-% will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% you can specify symbols, otherwise they are numbered in order
-% ideally, you should not use this facility. affiliations will be numbered
-% in order of appearance and this is the preferred way.
-% \icmlsetsymbol{equal}{*}
-\begin{icmlauthorlist}
-\icmlauthor{Daniel Seita}{berkeley}
-\icmlauthor{Xinlei Pan}{berkeley}
-\icmlauthor{Haoyu Chen}{berkeley}
-\icmlauthor{John Canny}{berkeley,google}
-\end{icmlauthorlist}
-\icmlaffiliation{berkeley}{University of California, Berkeley, CA}
-\icmlaffiliation{google}{Google Research, Mountain View, CA}
-\icmlcorrespondingauthor{Daniel Seita}{seita@berkeley.edu}
-
-% You may provide any keywords that you 
-% find helpful for describing your paper; these are used to populate 
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{boring formatting information, machine learning, ICML}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-%\footnotetext{hi}
+\maketitle
 
@@
 
-\section{Introduction}\label{sec:introduction}
+\section{INTRODUCTION}\label{sec:introduction}
 
@@
 
-\section{Preliminaries}\label{sec:related_work}
+\section{PRELIMINARIES}\label{sec:related_work}
 
@@
 
-\subsection{A Worst-Case Gaussian Example}\label{ssec:gaussian_example}
+\subsection{A WORST-CASE GAUSSIAN EXAMPLE}\label{ssec:gaussian_example}
 
@@
 
-\subsection{MCMC Posterior Inference}
+\subsection{MCMC POSTERIOR INFERENCE}
 There is a separate line of MCMC work drawing principles from statistical
@@
 
-\section{A New MH Acceptance Test}\label{sec:our_algorithm}
-
-\subsection{Log-Likelihood Ratios}\label{ssec:log_likelihood_ratios}
+\section{A NEW MH ACCEPTANCE TEST}\label{sec:our_algorithm}
+
+\subsection{LOG-LIKELIHOOD RATIOS}\label{ssec:log_likelihood_ratios}
 
@@
 
-\subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
+\subsection{BARKER (LOGISTIC) ACCEPTANCE FUNCTION}\label{ssec:barker_function}
 For our new MH test we use the Barker logistic~\cite{Barker65}
@@
 
-\subsection{A Minibatch Acceptance Test}\label{ssec:deltas_minibatch}
+\subsection{A MINIBATCH ACCEPTANCE TEST}\label{ssec:deltas_minibatch}
 
@@
 
-\section{Computing the Correction Distribution}\label{sec:correction}
+\section{THE CORRECTION DISTRIBUTION}\label{sec:correction}
 
@@
 
-\section{Analysis}\label{sec:analysis}
+\section{ANALYSIS}\label{sec:analysis}
 
@@
 
-\subsection{Bounding the Error of $\Delta^*$ from Gaussian}\label{ssec:delta_star_distribution}
+\subsection{BOUNDING THE ERROR OF $\Delta^*$ FROM A GAUSSIAN}\label{ssec:delta_star_distribution}
 
@@
 %\subsection{Error Bounds are Preserved After Adding Random Variables}\label{ssec:preserve_bounds}
-\subsection{Adding Random Variables}\label{ssec:preserve_bounds}
+\subsection{ADDING RANDOM VARIABLES}\label{ssec:preserve_bounds}
 
@@
 
-\subsection{Bounds on the Stationary Distribution}
+\subsection{BOUNDS ON THE STATIONARY DISTRIBUTION}
 Bounds on the error of an M-H test imply bounds on the stationary distribution
@@
 
-\section{Experiments}\label{sec:experiments}
-
-\subsection{Mixture of Gaussians}\label{ssec:gaussians}
+\section{EXPERIMENTS}\label{sec:experiments}
+
+\subsection{MIXTURE OF GAUSSIANS}\label{ssec:gaussians}
 
@@
 
-\subsection{Logistic Regression}\label{ssec:logistic}
+\subsection{LOGISTIC REGRESSION}\label{ssec:logistic}
 
@@
 \begin{table}[t]
-\caption{Log. Regression Average Minibatch Size (10 Trials)}
+\caption{Logistic Regression Average Minibatch Size}
 \label{tab:logistic}
@@
 \begin{center}
-\begin{small}
-\begin{sc}
 \begin{tabular}{l l l l}
-\hline
-\abovespace\belowspace
-Method/Data & MNIST-13k & MNIST8M-100k  \\
-\hline
-\abovespace
-Our Method & $129.8\pm 3.4$ & \textbf{TODO} \\
-AustereMH  & \textbf{TODO} & \textbf{TODO} \\
-\belowspace
-MHSubLhd   & $10752.8\pm 92.9$ & \textbf{TODO} \\
-\hline
+\textbf{Method/Data} & \textbf{MNIST-13k} & \textbf{MNIST-100k}  \\
+\hline \\
+Our Method & $129.8\pm 3.4$ & $2059.9\pm 135.5$ \\
+AustereMH  & $1924.3\pm 52.4$ & \textbf{TODO} \\
+MHSubLhd   & $10783.4\pm 78.9$ & \textbf{TODO} \\
 \end{tabular}
-\end{sc}
-\end{small}
 \end{center}
@@
 
-\subsection{Collaborative Filtering}\label{ssec:collab_filtering}
+\subsection{SPARSE MATRIX FACTORIZATION}\label{ssec:collab_filtering}
 
@@
 \bibliography{paper}
-\bibliographystyle{icml2017}
+\bibliographystyle{plain}
 
@@
 
-\section{Proofs of Lemmas and Corollaries}\label{app:proofs}
-
-\subsection{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}
+\section{PROOFS OF LEMMAS AND COROLLARIES}\label{app:proofs}
+
+\subsection{PROOF OF LEMMA~\ref{lem:worst_case}}\label{app:worst_case_proof}
 
@@
 These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative by
-inspection of equation (\ref{eq:whyneg}) below.  This implies that we can find a
+inspection of Equation (\ref{eq:whyneg}) below.  This implies that we can find a
 $u \in (0,1)$ so that $\psi(u,\theta,\theta') = \log u < 0$ equals
@@
 Equation~(\ref{eq:lemma_ll_ratio}), we see that
-\begin{equation}
+\begin{equation}\label{eq:whyneg}
     \log u_0 = N(\theta'-\theta)\frac{1}{b} \cdot \mE\left[\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2}\right]
-\end{equation}
-\begin{equation}\label{eq:whyneg}
-    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
+    = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
 \end{equation}
@@
 
-\subsection{Proof of Lemma~\ref{lem:quant_clt}}\label{app:quant_clt}
+\subsection{PROOF OF LEMMA~\ref{lem:quant_clt}}\label{app:quant_clt}
 
@@
 
-\subsection{Proof of Lemma~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}
+\subsection{PROOF OF LEMMA~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}
 
@@
 
-\subsection{Proof of Corollary~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}
+\subsection{PROOF OF COROLLARY~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}
 
@@
 
-\subsection{Improved Error Bounds Based on Skew Estimation}\label{app:better_error_bound}
+\subsection{IMPROVED ERROR BOUNDS BASED ON SKEW ESTIMATION}\label{app:better_error_bound}
 
@@
 
-\section{Additional Experiment Details}\label{app:experiments}
+\section{ADDITIONAL EXPERIMENT DETAILS}\label{app:experiments}
 % Daniel: all three experiments should be described in more detail here. And
@@
 
-\subsection{Obtaining the Correction Distribution (Section~\ref{sec:correction})}\label{app:correction}
-
-% Daniel: moved to appendix so we can expand it. Originally we had only two
-% lines, and it looked pretty awkward. So let's add more. Note: this also
-% depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
+\subsection{OBTAINING THE CORRECTION DISTRIBUTION (SECTION~\ref{sec:correction})}\label{app:correction}
+
+%% % Daniel: moved to appendix so we can expand it. Originally we had only two
+%% % lines, and it looked pretty awkward. So let's add more. Note: this also
+%% % depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
 \begin{table}[t]
@@
 \begin{small}
-\begin{sc}
 %% ----------------------- START OF TABLE
@@
 \hline
-\abovespace\belowspace
 $N=2000$ & $\sigma=0.8$ \\
 \hline
-\abovespace\belowspace
 $\lambda$ & $L_{\infty}$ error \\
 \hline
-\abovespace
 100    & 2.6e-3  \\
@@
 0.1    & 1.4e-5  \\
-\belowspace
 0.01   & \textbf{5.0e-6}  \\
@@
 \hline
-\abovespace\belowspace
 $N=2000$ & $\sigma=0.9$ \\
 \hline
-\abovespace\belowspace
 $\lambda$ & $L_{\infty}$ error \\
 \hline
-\abovespace
 100    & 3.3e-3  \\
@@
 0.1    & \textbf{1.3e-4}  \\
-\belowspace
 0.01   & 2.7e-4  \\
@@
 \hline
-\abovespace\belowspace
 $N=2000$ & $\sigma=1.0$ \\
 \hline
-\abovespace\belowspace
 $\lambda$ & $L_{\infty}$ error \\
 \hline
-\abovespace
 100    & 4.4e-3  \\
@@
 0.1    & 2.0e-3  \\
-\belowspace
 0.01   & 3.6e-3  \\
@@
 \hline
-\abovespace\belowspace
 $N=2000$ & $\sigma=1.1$ \\
 \hline
-\abovespace\belowspace
 $\lambda$ & $L_{\infty}$ error \\
 \hline
-\abovespace
 100    & 6.8e-3  \\
@@
 0.1    & 1.3e-2  \\
-\belowspace
 0.01   & 2.4e-2  \\
@@
 \hline
-\abovespace\belowspace
 $N=4000$ & $\sigma=0.8$ \\
 \hline
-\abovespace\belowspace
 $\lambda$ & $L_{\infty}$ error \\
 \hline
-\abovespace
 100    & 8.3e-4  \\
@@
 0.1    & \textbf{6.7e-6}  \\
-\belowspace
 0.01   & 7.4e-6  \\
@@
 \hline
-\abovespace\belowspace
 $N=4000$ & $\sigma=0.9$ \\
 \hline
-\abovespace\belowspace
 $\lambda$ & $L_{\infty}$ error \\
 \hline
-\abovespace
 100    & 1.2e-3  \\
@@
 0.1    & 2.0e-4  \\
-\belowspace
 0.01   & 3.9e-4  \\
@@
 \hline
-\abovespace\belowspace
 $N=4000$ & $\sigma=1.0$ \\
 \hline
-\abovespace\belowspace
 $\lambda$ & $L_{\infty}$ error \\
 \hline
-\abovespace
 100    & 1.9e-3  \\
@@
 0.1    & 2.8e-3  \\
-\belowspace
 0.01   & 5.2e-3  \\
@@
 \hline
-\abovespace\belowspace
 $N=4000$ & $\sigma=1.1$ \\
 \hline
-\abovespace\belowspace
 $\lambda$ & $L_{\infty}$ error \\
 \hline
-\abovespace
 100    & \textbf{4.3e-3}  \\
@@
 0.1    & 1.2e-2  \\
-\belowspace
 0.01   & 3.5e-2  \\
@@
 \end{tabular}
-\end{sc}
 \end{small}
@@
 
-\subsection{Gaussian Mixture Model Experiment (Section~\ref{ssec:gaussians})}\label{app:gaussians}
+\subsection{GAUSSIAN MIXTURE MODEL EXPERIMENT (SECTION~\ref{ssec:gaussians})}\label{app:gaussians}
 
@@
 \begin{center}
-\begin{small}
-\begin{sc}
 \begin{tabular}{l l l l}
-\hline
-\abovespace\belowspace
-Metric & Ours & Korat.'14 & Barde.'14 \\
-\hline
-\abovespace
+\textbf{Metric} & \textbf{Ours} & \textbf{AustereMH} & \textbf{MHSubLhd} \\
+\hline \\
 Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
-\belowspace
 Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
-\hline
 \end{tabular}
-\end{sc}
-\end{small}
 \end{center}
@@
 
-\subsection{Logistic Regression Experiment (Section~\ref{ssec:logistic})}\label{app:logistic}
+\subsection{LOGISTIC REGRESSION EXPERIMENT (SECTION~\ref{ssec:logistic})}\label{app:logistic}
 
@@
 
-\subsection{Collaborative Filtering Experiment (Section~\ref{ssec:collab_filtering})}\label{app:collab_filtering}
+\subsection{SPARSE MATRIX FACTORIZATION EXPERIMENT (SECTION~\ref{ssec:collab_filtering})}\label{app:collab_filtering}
 
@@
 \end{document}
-
-% Daniel: some notes to keep in mind to standardize to ICML format.
-
-%% \begin{figure}[ht]
-%% \vskip 0.2in
-%% \begin{center}
-%% \centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
-%% \caption{Historical locations and number of accepted papers for International
-%%   Machine Learning Conferences (ICML 1993 -- ICML 2008) and
-%%   International Workshops on Machine Learning (ML 1988 -- ML
-%%   1992). At the time this figure was produced, the number of
-%%   accepted papers for ICML 2008 was unknown and instead estimated.}
-%% \label{icml-historical}
-%% \end{center}
-%% \vskip -0.2in
-%% \end{figure} 
-%% 
-%% \subsection{Figures}
-%%  
-%% You may want to include figures in the paper to help readers visualize
-%% your approach and your results. Such artwork should be centered,
-%% legible, and separated from the text. Lines should be dark and at
-%% least 0.5~points thick for purposes of reproduction, and text should
-%% not appear on a gray background.
-%% 
-%% Label all distinct components of each figure. If the figure takes the
-%% form of a graph, then give a name for each axis and include a legend
-%% that briefly describes each curve. Do not include a title inside the
-%% figure; instead, the caption should serve this function.
-%% 
-%% Number figures sequentially, placing the figure number and caption
-%% {\it after\/} the graphics, with at least 0.1~inches of space before
-%% the caption and 0.1~inches after it, as in
-%% Figure~\ref{icml-historical}.  The figure caption should be set in
-%% 9~point type and centered unless it runs two or more lines, in which
-%% case it should be flush left.  You may float figures to the top or
-%% bottom of a column, and you may set wide figures across both columns
-%% (use the environment {\tt figure*} in \LaTeX), but always place
-%% two-column figures at the top or bottom of the page.
-%% 
-%% \subsection{Tables} 
-%%  
-%% You may also want to include tables that summarize material. Like 
-%% figures, these should be centered, legible, and numbered consecutively. 
-%% However, place the title {\it above\/} the table with at least 
-%% 0.1~inches of space before the title and the same after it, as in 
-%% Table~\ref{sample-table}. The table title should be set in 9~point 
-%% type and centered unless it runs two or more lines, in which case it
-%% should be flush left.
-%% 
-%% % Note use of \abovespace and \belowspace to get reasonable spacing 
-%% % above and below tabular lines. 
-%% 
-%% \begin{table}[t]
-%% \caption{Classification accuracies for naive Bayes and flexible 
-%% Bayes on various data sets.}
-%% \label{sample-table}
-%% \vskip 0.15in
-%% \begin{center}
-%% \begin{small}
-%% \begin{sc}
-%% \begin{tabular}{lcccr}
-%% \hline
-%% \abovespace\belowspace
-%% Data set & Naive & Flexible & Better? \\
-%% \hline
-%% \abovespace
-%% Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
-%% Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
-%% Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
-%% Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
-%% Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
-%% Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
-%% Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
-%% \belowspace
-%% Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
-%% \hline
-%% \end{tabular}
-%% \end{sc}
-%% \end{small}
-%% \end{center}
-%% \vskip -0.1in
-%% \end{table}
-%%
-%% % Acknowledgements should only appear in the accepted version. 
-%% \section*{Acknowledgements} 
-%% \textbf{Do not} include acknowledgements in the initial version of
-%% the paper submitted for blind review.

### 2017-03-01 uai2017/paper.tex
Whew, MASSIVE (well ok semi-massive) change for UAI preparation.
We now have a far expanded appendix, revised logistic regression stuff, etc. I still have one LAST LR thing running ... hopefully by Friday we won't have anything else to do except for SMF

@@
 \graphicspath{{figures/}{../figures/}}
+
+\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut
+\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut
 
@@
 \begin{algorithm}[tb]
-\caption{Our acceptance test for MCMC.}
+\caption{Our acceptance test, {\sc MCminibatch}.}
 \label{alg:our_algorithm}
@@
 With this approach, there is no guarantee that $u^* \geq 0$. However, we have
-some flexibility in the choice of $\sigma$ in Equation~(\ref{eq:overall_corr_problem}).
-As we decrease the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$
-grows by the same amount and is in fact the result of convolution with a
-Gaussian whose variance is the difference.  Thus as $\sigma$ decreases,
-$C_\sigma(X)$ grows and approaches the derivative of a logistic function at
-$\sigma = 0$. It retains some weak negative values for $\sigma > 0$ but removal
-of those leads to small error. \textbf{Daniel: I moved this to
-Appendix~\ref{app:correction} so that we can have more detailed results. For the
-main text, let's state the value $\sigma$ and its error. Originally, we said:
-Table~\ref{tab:xcorr} shows that the errors between $X_{\rm norm}+X_{\rm corr}$
-and $X_{\rm log}$ approach single floating precision (about $10^{-7}$).}
-Algorithm~\ref{alg:our_algorithm} describes our procedure. A few points:
+some flexibility in the choice of $\sigma$ in
+Equation~(\ref{eq:overall_corr_problem}).  As we decrease the variance of
+$X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount and is
+in fact the result of convolution with a Gaussian whose variance is the
+difference.  Thus as $\sigma$ decreases, $C_\sigma(X)$ grows and approaches the
+derivative of a logistic function at $\sigma = 0$. It retains some weak negative
+values for $\sigma > 0$ but removal of those leads to small error. We use
+$N=4000$ and $\lambda=10$ for our experiments, which empirically provided
+excellent performance. See Table~\ref{tab:xcorr} in
+Appendix~\ref{app:correction} for detailed $L_{\rm \infty}$ errors for different
+settings. Algorithm~\ref{alg:our_algorithm} describes our procedure, {\sc
+MHminibatch}. A few points:
 
@@
 \section{ANALYSIS}\label{sec:analysis}
+
+\textbf{Daniel: TODO this should be where we mention that AustereMH still
+requires computation of all likelihoods, and we should explain John's complaint
+about Gaussianity here.}
 
@@
 \section{EXPERIMENTS}\label{sec:experiments}
+
+In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark our test,
+{\sc MHminibatch}, against the test from~\cite{icml2014c1_bardenet14} ({\sc
+MHSubLhd}) and from~\cite{cutting_mh_2014}. As highlighted in
+Section~\ref{sec:analysis}, the test from~\cite{cutting_mh_2014} can be
+implemented using a conservative or a non-conservative variant; we name these
+{\sc AustereMH(c)}, and {\sc AustereMH(nc)}, respectively, and they use the
+grid-search over minibatch sizes $m$ and per-test thresholds $\epsilon$
+described in Appendix~\ref{sssec:grid_search}. Throughout our descriptions, we
+refer to a \emph{trial} as the period when an algorithm collects all its desired
+samples $\{\theta_1,\ldots,\theta_T\}$, with $T$ usually several thousand.
+\textbf{Daniel: better call temperature something other than $T$ ...}
 
@@
 
-This model is adapted from~\cite{langevin_2011} by increasing the number
-of samples to 1 million.  The parameters are $\theta = \langle \theta_1,\theta_2 \rangle$, and the
-generation process is
+This model is adapted from~\cite{langevin_2011} by increasing the number of
+samples to 1 million.  The parameters are $\theta = \langle \theta_1,\theta_2
+\rangle$, and the generation process is
 \begin{equation}\label{eq:data_generation}
@@
 
-We ran MCMC with with our test against the test
-from~\cite{icml2014c1_bardenet14} and a simplified version
-of~\cite{cutting_mh_2014}. Both methods use repeated testing which must be
-discounted to bound overall test accuracy. ~\cite{icml2014c1_bardenet14}
-conservatively partitions the test error, while~\cite{cutting_mh_2014} used a
-complex dynamic programming heuristic to estimate the overall test error. We did
-not implement this heuristic and instead set the individual test error equal to
-the overall test error. Thus our results for~\cite{cutting_mh_2014} should be
-treated as lower bounds. All methods were initialized with minibatch size 100.
-For~\cite{cutting_mh_2014} we increment sizes by 100 and set the tolerance
-$\epsilon=0.005$ to control overall test error.
-For~\cite{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma =
-1.5$ and use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
-samples using a random walk proposer with covariance ${\rm diag}(0.15, 0.15)$,
-which means the M-H test is responsible for shaping the sample distribution.
+We benchmark with {\sc AustereMH(c)} and {\sc MHSubLhd}. {\sc MHminibatch} and
+{\sc MHSubLhd} were initialized with $m=100$. For {\sc MHSubLhd}, we increase
+sizes geometrically with $\gamma = 1.5$ and use parameters $p = 2, \delta =
+0.01$.  All methods collect 5000 samples using a random walk proposer with
+covariance ${\rm diag}(0.15, 0.15)$, which means the M-H test is responsible for
+shaping the sample distribution.
 
@@
 samples for the three methods, with darker regions indicating a greater density
-of points. There are no obvious differences, so we measure the similarity
-between each set of samples and the actual posterior. \textbf{Daniel: I think
-this was taking up too much space so I put it in Appendix~\ref{app:gaussians}.
-Let's state the result (in Table~\ref{tab:poissons}) and prioritize other stuff
-here.  TODO}
-
-Figure~\ref{fig:gauss_mix_2} shows that the new method dominates in
-terms of speed and efficiency. The histograms of the (final) minibatch sizes
-used each iteration show that our method consumes significantly less data; the
+of points. There are no obvious differences, showing that {\sc MHminibatch}
+reaches an acceptable posterior. We further measure the similarity between each
+set of samples and the actual posterior. Due to space constraints, results are
+in Appendix~\ref{app:gaussians}.
+
+Figure~\ref{fig:gauss_mix_2} shows that {\sc MHminibatch} dominates in terms of
+speed and efficiency. The histograms of the (final) minibatch sizes used each
+iteration show that our method consumes significantly less data; the
 distribution is short-tailed and the mean is 210, more than an order of
 magnitude better compared to the other two methods (averages are 15562 and
-16857). The sizes correspond to the running times of the methods, since
-they all consume linear time in data consumed.
+16857). The sizes correspond to the running times of the methods, excluding the
+likelihood computation of all data points for {\sc MHSubLhd}, which would
+drastically increase running time.
 
@@
 	\caption{
-    TODO
+    Binary classification accuracy of the MCMC methods on the 1s vs 7s logistic
+    regression task for MNIST-13k (left plot) and MNIST-100k (middle plot) as a
+    function of cumulative data usage.  The right plot reports performance of
+    {\sc MHminibatch} on both datasets when the temperature starts at $T=100$
+    and drops to $T=1$ after a ``burn-in'' period of 200 samples of $\theta$.
+    For all three plots, one standard deviation is indicated by the shaded error
+    regions.
     }
@@
 	\caption{
-    Minibatch sizes used for one representative trial in
-    Section~\ref{ssec:logistic} using MNIST-13k (analogous to
-    Figure~\ref{fig:gauss_mix_2}).
+    Minibatch sizes for a representative trial of logistic regression on
+    MNIST-13k (analogous to Figure~\ref{fig:gauss_mix_2}). Both axes are on a
+    log scale and have the same ranges across the three histograms. See
+    Section~\ref{ssec:logistic} for details.
     }
@@
 	\caption{
-    Minibatch sizes used for one representative trial in
-    Section~\ref{ssec:logistic} using MNIST8M-100k (similar to
-    Figure~\ref{fig:logistic_histograms_mnist}).
+    Minibatch sizes for a representative trial of logistic regression on
+    MNIST-100k (similar to Figure~\ref{fig:logistic_histograms_mnist}). Both
+    axes are on a log scale and have the same ranges across the three
+    histograms.  See Section~\ref{ssec:logistic} for details.
     }
@@
 \begin{table}[t]
-\caption{Logistic Regression Average Minibatch Size}
+\caption{
+Average minibatch sizes ($\pm$ one standard deviation) for the methods on the
+logistic regression tasks of binary classification on MNIST-13k and MNIST-100k.
+The averages are taken over 10 independent trials (5000 samples each) for
+MNIST-13k and 5 independent trials (3000 samples each) for MNIST-100k.
+}
+\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
 \label{tab:logistic}
@@
 \hline \\
-Our Method & $129.8\pm 3.4$ & $2059.9\pm 135.5$ \\
-AustereMH  & $1924.3\pm 52.4$ & \textbf{TODO} \\
-MHSubLhd   & $10783.4\pm 78.9$ & \textbf{TODO} \\
+{\sc MHminibatch}   & $129.8\pm 3.4$    & $202.5 \pm 13.6$ \\
+{\sc AustereMH(nc)} & $973.8\pm 49.8$   & N/A \\
+{\sc AustereMH(c)}  & $1924.3\pm 52.4$  & $2795.6 \pm 364.0$ \\
+{\sc MHSubLhd}      & $10783.4\pm 78.9$ & $14977.3 \pm 582.0$ \\
 \end{tabular}
@@
 
-\textbf{Daniel: TODO this section is badly out of date but I will re-write it
-and it should be much better than what we had earlier. Also, define a trial
-since I say it in the table and figures. Finally, define AustereMH and MHSubLhd,
-and come up with a name for our method.}
-
 We next test logistic regression for the binary classification of 1s versus 7s
-on a subset of the MNIST8M dataset, a larger version of
-MNIST~\cite{lecun-mnisthandwrittendigit-2010}. We randomly subsampled 450k
-training and 192k testing points.  We impose a uniform prior on $\theta$ and
-again use a random walk proposer, this time with covariance matrix $0.05I$. The
-temperature is set at $T=1000$. We run the three tests for 3000 samples and
-again use 100 as the starting minibatch size. Our undiscounted implementation
-of~\cite{cutting_mh_2014} used per-test error of $\epsilon=0.05$.
-For~\cite{icml2014c1_bardenet14}, we use their recommended discounting factor
-value of $\gamma=2.0$. We tried to use the symbolic bound for
-$C_{\theta,\theta'}$ for Logistic regression from~\cite{icml2014c1_bardenet14},
-but found it to be too high. Instead we use the empirical $C_{\theta,\theta'}$
-from the entire dataset and do not add its computation time when considering
-runtime, so again our results represent a lower bound for this method.
-
-Figure~\ref{fig:logistic_performance} shows the test log likelihood and
-prediction accuracy as a function of the cumulative training points
-processed.\footnote{The curves do not span the same length over the x-axis since
-the methods consume different amounts of data.} To generate the curves, for each
-of the sampled vectors $\theta_t$, $t\in\{1,\ldots,3000\}$, we use $\theta_t$ as
-the parameter for logistic regression.  Our minibatch MH test is more efficient
-in terms of accuracy, achieving convergence roughly twice as fast compared
-to~\cite{cutting_mh_2014} and about 1.5 orders of magnitude faster
-than~\cite{icml2014c1_bardenet14}, though its log likelihood results lag
-slightly behind the former algorithm during the very early stages of
-exploration.
-
-Figure~\ref{fig:logistic_minibatch} shows histograms of minibatch sizes for the
-three methods on a log-log scale. With an initial size of 100, we achieve an
-average minibatch size of 393, more than two orders of magnitude smaller
-than~\cite{icml2014c1_bardenet14}, and almost an order of magnitude faster than
-the undiscounted implementation of~\cite{cutting_mh_2014}.
+on the MNIST~\cite{lecun-mnisthandwrittendigit-2010} dataset and (a subset of)
+infinite MNIST~\cite{loosli-canu-bottou-2006}. For the former, extracting all 1s
+and 7s resulted in 13,000 training samples, and for the latter, we used 87,000
+additional (augmented) 1s and 7s to get 100,000 training samples. Both datasets
+use the same test set, with 2,163 samples. Henceforth, we call them MNIST-13k
+and MNIST-100k, respectively.
+
+For all methods, we impose a uniform prior on $\theta$ and again use a random
+walk proposer, with covariance matrix $0.05I$ for MNIST-13k and $0.01I$ for
+MNIST-100k. The default temperature setting is a constant at $T=100$ for
+MNIST-13k and MNIST-100k. Performance of all methods implicitly relies on the
+step size and temperature. Setting temperature too low or step size too high
+will generally result in worse performance for all methods. For MNIST-13k, each
+method generated 5000 samples for ten independent trials; due to MNIST-100k's
+higher computational requirements, the methods generated 3000 samples for five
+independent trials.  For additional parameter settings and an investigation on
+tuning step sizes, see Appendix~\ref{app:logistic}.
+
+For {\sc MHSubLhd}, we tried to use the provided symbolic bound for
+$C_{\theta,\theta'}$ described in~\cite{icml2014c1_bardenet14}, but it was too
+high and provided no performance benefit. Instead we use the empirical
+$C_{\theta,\theta'}$ from the entire dataset.
+
+The first two subplots of Figure~\ref{fig:logistic_performance} display the
+prediction accuracy on both datasets for all methods as a function of the
+cumulative training points processed.\footnote{The curves do not span the same
+length over the x-axis since the methods consume different amounts of data.} To
+generate the curves, for each of the sampled vectors $\theta_t$,
+$t\in\{1,\ldots,N\}$ ($N=5000$ for MNIST-13k, $N=3000$ for MNIST-100k), we use
+$\theta_t$ as the parameter for logistic regression. The results indicate that
+our test is more efficient, obtaining convergence more than an order of
+magnitude faster than {\sc AustereMH(nc)} and several orders of magnitude
+compared to {\sc AustereMH(c)} and {\sc MHSubLhd}. We also observe the advantage
+of having higher temperature from the third plot in
+Figure~\ref{fig:logistic_performance}. During the exploration period, the
+accuracy rapidly increases, and then after 200 samples, we switch the
+temperature to 1 (equivalent to ignoring temperature). {\rm MHminibatch}
+requires the step size to decrease, hence the smaller changes in accuracy.
+
+Figures~\ref{fig:logistic_histograms_mnist}
+and~\ref{fig:logistic_histograms_mnist8m} show histograms of minibatch sizes for
+the three methods on a log-log scale for MNIST-13k and MNIST-100k, respectively.
+The histograms only represent one representative trial; Table~\ref{tab:logistic}
+contains the average of the average minibatch sizes ($\pm$ one standard
+deviation) across either 10 or 5 trials. The results indicate that {\sc
+MHminibatch}, with average minibatch sizes of roughly 129.8 and 202.5 for
+MNIST-13k and MNIST-100k, respectively, consumes almost an order of magnitude
+fewer data points than the next-best method, {\sc AustereMH(nc)}.  We reiterate,
+however, that both {\sc AustereMH(nc)} and {\sc MHSubLhd} require computing
+$\log p(x_i|\theta)$ and $\log p(x_i|\theta')$ for all $x_i$ each iteration.
+These results here do not count that extra data consumption. Only our method and
+{\sc AustereMH(c)} rely solely on the minibatch of data each iteration.
 
@@
 
+\begin{table}[t]
+\caption{\textbf{TODO TABLE WITH SMF RESULTS}}
+\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
+\label{tab:smf}
+\vskip 0.15in
+\begin{center}
+\begin{tabular}{l l l l}
+\textbf{RMSE} & \textbf{MovieLens 10m} & \textbf{Netflix}  \\
+\hline \\
+{\sc MHminibatch} & ??? & ??? & \\
+ADAGrad           & ??? & ??? & \\
+\end{tabular}
+\end{center}
+\vskip -0.1in
+\end{table}
+
 \textbf{Daniel: TODO we wanted a third experiment, right?}
@@
 \section{CONCLUSIONS}\label{sec:conclusion}
-
-\textbf{Daniel: double check this in case we add a third experiment.}
 
@@
 tests. We present theoretical results and experimentally show the benefits of
-our test on Gaussian mixtures and logistic regression.  Directions for future
-work include integrating our algorithm with~\cite{cutting_mh_2014} by applying
-both tests each iteration or utilizing the variance reduction techniques
-suggested in~\cite{DBLP:conf/icml/ChenG16}.  More elaborate extensions involve
-running more experiments on other models such as neural networks and providing a
-recipe for how to use our algorithm (following the framework
-of~\cite{sgmcmc_2015}).
+our test on Gaussian mixtures, logistic regression, and sparse matrix
+factorization.  Directions for future work include integrating our algorithm
+with~\cite{cutting_mh_2014} by applying both tests each iteration or utilizing
+the variance reduction techniques suggested in~\cite{DBLP:conf/icml/ChenG16}.
+More elaborate extensions involve running more experiments on other models such
+as neural networks and providing a recipe for how to use our algorithm
+(following the framework of~\cite{sgmcmc_2015}).
 
@@
 distribution derivation and on our three main experiments to assist
-understanding and reproducibility.
+understanding and reproducibility. Our code is open-source on
+GitHub.\footnote{\url{https://github.com/BIDData/BIDMach/blob/master/src/main/scala/BIDMach/updaters/MHTest.scala}}
+% Daniel: TODO remove for anonymous review later.
 
@@
 \end{equation}
-% Daniel: this is the multi-line version, but we should use the full page.
-%\begin{align*}
-%-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
-%&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
-%&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
-%\end{align*}
 as desired.
@@
 Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
-$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) =
-{\rm Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
+$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
+Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
 that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
-nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm log} > 0)$
-by at most $\epsilon$.
+nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm log}
+> 0)$ by at most $\epsilon$.
 
@@
 \hline
-$N=2000$ & $\sigma=0.8$ \\
-\hline
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-100    & 2.6e-3  \\
+\T\B  $N=2000$ & $\sigma=0.8$ \rule{0pt}{3ex} \\
+\hline
+\T\B $\lambda$ & $L_{\infty}$ error \\
+\hline
+100    & 2.6e-3 \T \\
 10     & 4.0e-4  \\
@@
 0.1    & 1.4e-5  \\
-0.01   & \textbf{5.0e-6}  \\
+0.01   & \textbf{5.0e-6} \B \\
 \hline
@@
 \hline
-$N=2000$ & $\sigma=0.9$ \\
-\hline
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-100    & 3.3e-3  \\
+\T\B $N=2000$ & $\sigma=0.9$ \\
+\hline
+\T\B $\lambda$ & $L_{\infty}$ error \\
+\hline
+100    & 3.3e-3 \T \\
 10     & 6.4e-4  \\
@@
 0.1    & \textbf{1.3e-4}  \\
-0.01   & 2.7e-4  \\
+0.01   & 2.7e-4 \B \\
 \hline
@@
 \hline
-$N=2000$ & $\sigma=1.0$ \\
-\hline
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-100    & 4.4e-3  \\
+\T\B $N=2000$ & $\sigma=1.0$ \\
+\hline
+\T\B $\lambda$ & $L_{\infty}$ error \\
+\hline
+100    & 4.4e-3 \T \\
 10     & 1.3e-3  \\
@@
 0.1    & 2.0e-3  \\
-0.01   & 3.6e-3  \\
+0.01   & 3.6e-3 \B \\
 \hline
@@
 \hline
-$N=2000$ & $\sigma=1.1$ \\
-\hline
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-100    & 6.8e-3  \\
+\T\B $N=2000$ & $\sigma=1.1$ \\
+\hline
+\T\B $\lambda$ & $L_{\infty}$ error \\
+\hline
+100    & 6.8e-3 \T \\
 10     & \textbf{4.6e-3}  \\
@@
 0.1    & 1.3e-2  \\
-0.01   & 2.4e-2  \\
+0.01   & 2.4e-2 \B \\
 \hline
@@
 \hline
-$N=4000$ & $\sigma=0.8$ \\
-\hline
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-100    & 8.3e-4  \\
+\T\B $N=4000$ & $\sigma=0.8$ \\
+\hline
+\T\B $\lambda$ & $L_{\infty}$ error \\
+\hline
+100    & 8.3e-4 \T \\
 10     & 1.3e-4  \\
@@
 0.1    & \textbf{6.7e-6}  \\
-0.01   & 7.4e-6  \\
+0.01   & 7.4e-6 \B \\
 \hline
@@
 \hline
-$N=4000$ & $\sigma=0.9$ \\
-\hline
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-100    & 1.2e-3  \\
+\T\B $N=4000$ & $\sigma=0.9$ \\
+\hline
+\T\B $\lambda$ & $L_{\infty}$ error \\
+\hline
+100    & 1.2e-3 \T \\
 10     & 2.6e-4  \\
@@
 0.1    & 2.0e-4  \\
-0.01   & 3.9e-4  \\
+0.01   & 3.9e-4 \B \\
 \hline
@@
 \hline
-$N=4000$ & $\sigma=1.0$ \\
-\hline
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-100    & 1.9e-3  \\
+\T\B $N=4000$ & $\sigma=1.0$ \\
+\hline
+\T\B $\lambda$ & $L_{\infty}$ error \\
+\hline
+100    & 1.9e-3 \T \\
 10     & \textbf{8.9e-4}  \\
@@
 0.1    & 2.8e-3  \\
-0.01   & 5.2e-3  \\
+0.01   & 5.2e-3 \B \\
 \hline
@@
 \hline
-$N=4000$ & $\sigma=1.1$ \\
-\hline
-$\lambda$ & $L_{\infty}$ error \\
-\hline
-100    & \textbf{4.3e-3}  \\
+\T\B $N=4000$ & $\sigma=1.1$ \\
+\hline
+\T\B $\lambda$ & $L_{\infty}$ error \\
+\hline
+100    & \textbf{4.3e-3} \T \\
 10     & 6.0e-3  \\
@@
 0.1    & 1.2e-2  \\
-0.01   & 3.5e-2  \\
+0.01   & 3.5e-2 \B \\
 \hline
@@
 
-We observe a tradeoff. With smaller $\sigma$, our $C_\sigma$ is closer to the
-ideal distribution (as judged by $L_\infty$ error), but this imposes a stricter
-upper bound on the sample variance of $\Delta^*$ before our test can be applied,
-which thus results in larger minibatch sizes. Conversely, a more liberal upper
-bound means we avail ourselves of smaller minibatch sizes, but at the cost of a
-less stable derivation for $C_\sigma$.
-
-\textbf{We chose XXXX to use in our experiments, which empirically provided an
-acceptable tradeoff to us.} This is reflected in
-Algorithm~\ref{alg:our_algorithm}, which assumes that we used \textbf{TODO TODO
-TODO TODO TODO} but we reiterate that the choice is arbitrary so long as $0 <
+We observe the expected tradeoff. With smaller $\sigma$, our $C_\sigma$ is
+closer to the ideal distribution (as judged by $L_\infty$ error), but this
+imposes a stricter upper bound on the sample variance of $\Delta^*$ before our
+test can be applied, which thus results in larger minibatch sizes. Conversely, a
+more liberal upper bound means we avail ourselves of smaller minibatch sizes,
+but at the cost of a less stable derivation for $C_\sigma$.
+
+We chose $N=4000, \sigma=1$, and $\lambda = 10$ to use in our experiments, which
+empirically exhibits excellent performance. This is reflected in the description
+of {\sc MHminibatch} in Algorithm~\ref{alg:our_algorithm}, which assumes that we
+used $\sigma=1$ but we reiterate that the choice is arbitrary so long as $0 <
 \sigma < \sqrt{\pi^2/3} \approx 1.814$, the standard deviation of the standard
@@
 \caption{Gaussian Mixture Model Statistics}
+\small
 \label{tab:poissons}
@@
 \begin{tabular}{l l l l}
-\textbf{Metric} & \textbf{Ours} & \textbf{AustereMH} & \textbf{MHSubLhd} \\
+\textbf{Metric} & {\sc MCminibatch} & {\sc AustereMH} & {\sc MHSubLhd} \\
 \hline \\
@@
 
+% Daniel: added this here, because it applies to both this and logistic
+% regression.
+\subsubsection{Grid Search}\label{sssec:grid_search}
+
+For the Gaussian mixture experiment, we use the conservative method
+from~\cite{cutting_mh_2014}, which avoids the need for recomputing log
+likelihoods of each data point each iteration by choosing baseline minibatch
+sizes $m$ and per-test thresholds $\epsilon$ beforehand, and then using those
+values for the entirety of the trials. We experimented with the following
+values:
+\begin{itemize}[noitemsep]
+\item $\epsilon \in \{0.001, 0.005, 0.01, 0.05, 0.1, 0.2\}$
+\item $m \in \{50, 100, 150, 200, 250, 300, 350, 400, 450, 500\}$
+\end{itemize}
+The main difference between {\sc AustereMH(c)} and {\sc
+AustereMH(nc)}\footnote{{\sc AustereMH(nc)} is used in
+Section~\ref{ssec:logistic}.}  is that the latter needs to run a grid search
+each iteration (i.e. after each time it makes an accept/reject decision for one
+sample $\theta_t$). We use the same $\epsilon$ and $m$ candidates above for {\sc
+AustereMH(nc)}.
+
+\subsubsection{Gaussian Mixture Model Metrics}\label{sssec:gauss_metrics}
+
+\textbf{Daniel: Xinlei, you should be in charge of checking this section and
+adding some stuff to Table~\ref{tab:poissons}, perhaps averages plus or minus
+one standard deviation.}
+
 We discretize the posterior coordinates into bins with respect to the two
@@
 and we can use it to measure the quality of the samples rather than use general
-purpose tests like KL-divergence or likelihood-ratio, which are problematic
-with zero counts.
+purpose tests like KL-divergence or likelihood-ratio, which are problematic with
+zero counts.
 
@@
 \begin{equation}\label{eq:log_prob_poissons}
-\begin{split}
-    \log p(c_1, \ldots, c_{N_{\rm bins}} | P_1, \ldots, P_{N_{\rm bins}}) & =  \\
-    \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)). & \\
-\end{split}
-\end{equation}
-Table~\ref{tab:poissons} shows the likelihoods. To facilitate
-interpretation we perform significance tests using Chi-Squared
-distribution (also in Table~\ref{tab:poissons}). Scores lie
-between~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}, but
-the variance of these values is high and ordering changes depending on
-the range of samples generated.
-
-\textbf{Daniel: TODO improve this}
+\log p(c_1, \ldots, c_{N_{\rm bins}} | P_1, \ldots, P_{N_{\rm bins}}) =
+\sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
+\end{equation}
+Table~\ref{tab:poissons} shows the likelihoods. To facilitate interpretation we
+perform significance tests using Chi-Squared distribution (also in
+Table~\ref{tab:poissons}). Scores lie between~\cite{cutting_mh_2014}
+and~\cite{icml2014c1_bardenet14}, but the variance of these values is high and
+ordering changes depending on the range of samples generated. \textbf{Daniel:
+one reviewer had a problem with this, and I agree. The easiest fix here is to
+simply report standard deviation values. Also, we should emphasize that we are
+not aiming to get the best score by this metric but simply something that
+exhibits some form of ``correctness.''}
 
@@
 
+\begin{figure*}[t]
+	\centering
+    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist8m.png}
+	\caption{
+    \textbf{Daniel: TODO this is a wrong figure, I'll add what I want later}
+    }
+	\label{fig:logistic_histograms_appendix}
+\end{figure*}
+
+\begin{table}[t]
+\caption{Parameters for the logistic regression experiments. Unless otherwise
+stated, values apply to all methods tested. For the methods
+from~\cite{cutting_mh_2014}, we use their notation for clarity.}
+\small
+\label{tab:logistic_settings}
+\vskip 0.15in
+\begin{center}
+\begin{tabular}{l l l}
+\textbf{Value} & \textbf{MNIST-13k} & \textbf{MNIST-100k} \\
+\hline
+\T Temperature $K$       & 100 & 100 \\
+Number of samples $T$ & 5000 & 3000 \\
+Number of trials      & 10 & 5 \\
+Step size $\sigma$ for random walk proposer with covariance $\sigma I$ & 0.05 & 0.01 \\
+{\sc MHminibatch} and {\sc MHSubLhd} minibatch size $m$ & 100 & 100 \\
+{\sc AustereMH(c)} $\Delta^*$ bound & 0.1 & 0.2 \\
+{\sc AustereMH(c)} minibatch size $m$ from grid search & 450 & 300 \\
+{\sc AustereMH(c)} per-test threshold $\epsilon$ from grid search & 0.01 & 0.01\\
+{\sc AustereMH(nc)} $\Delta^*$ bound & 0.05 & 0.1 \\
+\end{tabular}
+\end{center}
+\vskip -0.1in
+\end{table}
+
+\textbf{Daniel: TODO major work in progress.}
+
+Table~\ref{tab:logistic_settings} indicates the parameter settings for the
+logistic regression experiments.
+
+Figure~\ref{fig:logistic_histograms_appendix} indicates minibatch histograms for
+one trial of {\sc MHminibatch} on MNIST-100k for proposers $\sigma I$ with
+$\sigma \in \{0.005, 0.01, 0.05\}$.
+
+\subsection{SPARSE MATRIX FACTORIZATION EXPERIMENT (SECTION~\ref{ssec:collab_filtering})}\label{app:collab_filtering}
+
 \textbf{Daniel: TODO}
 
-\subsection{SPARSE MATRIX FACTORIZATION EXPERIMENT (SECTION~\ref{ssec:collab_filtering})}\label{app:collab_filtering}
-
-\textbf{Daniel: TODO}
-
 \end{document}

### 2017-03-01 uai2017/paper.tex
fix two typos, MC -> MH
All figures should now be up to date.
OK now let me do more changes and then I will tell Xinlei what to do. Almost there ...

@@
 \begin{algorithm}[tb]
-\caption{Our acceptance test, {\sc MCminibatch}.}
+\caption{Our acceptance test, {\sc MHminibatch}.}
 \label{alg:our_algorithm}
@@
 \begin{equation}
+\small
 \sup_y \left|{\rm Pr}(\Delta^*< y) -\Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| \leq \frac{6.4\mE[|X|^3]+2\mE[|X|]}{\sqrt{b}}
@@
 
+% Daniel: NOTE, I'm calling temperature "K" so that "T" stays reserved for
+% "number of samples of \theta we draw".
 In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark our test,
@@
 refer to a \emph{trial} as the period when an algorithm collects all its desired
-samples $\{\theta_1,\ldots,\theta_T\}$, with $T$ usually several thousand.
-\textbf{Daniel: better call temperature something other than $T$ ...}
+samples $\{\theta_1,\ldots,\theta_T\}$, generally with $T=3000$ or $T=5000$.
 
@@
 two very narrow peaks.  Our goal is to reproduce the original posterior, so we
-adjust the temperature to $T=10,000$.  Taking logs, we get the target as shown
+adjust the temperature to $K=10,000$.  Taking logs, we get the target as shown
 in the far left of Figure~\ref{fig:gauss_mix_1}.
@@
     function of cumulative data usage.  The right plot reports performance of
-    {\sc MHminibatch} on both datasets when the temperature starts at $T=100$
-    and drops to $T=1$ after a ``burn-in'' period of 200 samples of $\theta$.
+    {\sc MHminibatch} on both datasets when the temperature starts at $K=100$
+    and drops to $K=1$ after a ``burn-in'' period of 200 samples of $\theta$.
     For all three plots, one standard deviation is indicated by the shaded error
@@
 	\centering
-    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist.png}
+    \includegraphics[width=0.9\linewidth]{logistic_regression_histograms_mnist.png}
 	\caption{
@@
 	\centering
-    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist8m.png}
+    \includegraphics[width=0.9\linewidth]{logistic_regression_histograms_mnist8m.png}
 	\caption{
@@
 {\sc MHminibatch}   & $129.8\pm 3.4$    & $202.5 \pm 13.6$ \\
-{\sc AustereMH(nc)} & $973.8\pm 49.8$   & N/A \\
+{\sc AustereMH(nc)} & $973.8\pm 49.8$   & $1098.3 \pm 44.9$ \\
 {\sc AustereMH(c)}  & $1924.3\pm 52.4$  & $2795.6 \pm 364.0$ \\
@@
 walk proposer, with covariance matrix $0.05I$ for MNIST-13k and $0.01I$ for
-MNIST-100k. The default temperature setting is a constant at $T=100$ for
+MNIST-100k. The default temperature setting is a constant at $K=100$ for
 MNIST-13k and MNIST-100k. Performance of all methods implicitly relies on the
@@
 \begin{tabular}{l l l l}
-\textbf{Metric} & {\sc MCminibatch} & {\sc AustereMH} & {\sc MHSubLhd} \\
+\textbf{Metric} & {\sc MHminibatch} & {\sc AustereMH} & {\sc MHSubLhd} \\
 \hline \\
@@
 	\centering
-    \includegraphics[width=0.75\linewidth]{logistic_regression_histograms_mnist8m.png}
+    \includegraphics[width=0.9\linewidth]{logistic_regression_step_size_tests.png}
 	\caption{
-    \textbf{Daniel: TODO this is a wrong figure, I'll add what I want later}
+    \textbf{Daniel: TODO describe}
     }

### 2017-03-02 uai2017/paper.tex
Whew, did some more changes:
Added temperature description, moved prior work to section 2 and discussed gaussian assumptions and korattikara method weaknesses, minor updates everywhere. Next up, revising experiments section and the appendix

@@
 reject based on an acceptance test. The acceptance test is usually a Metropolis
-test~\cite{Metropolis1953, hastings70}.
+test~\citep{Metropolis1953, hastings70}.
 
@@
 over the full dataset to produce a new sample.  Recent results
-from~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14} perform
+from~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14} perform
 approximate (bounded error) acceptance tests using subsets of the full dataset.
 The amount of data consumed for each test varies significantly from one
-minibatch to the next. By contrast,~\cite{conf/uai/MaclaurinA14,TallData16}
+minibatch to the next. By contrast,~\citep{conf/uai/MaclaurinA14,TallData16}
 perform exact tests but require a lower bound on the parameter distribution across
@@
 statistics as {\em a natural part of the test} and requires less data per
-iteration than prior work. We use a Barker test function~\cite{Barker65}, which
+iteration than prior work. We use a Barker test function~\citep{Barker65}, which
 makes our test naturally error tolerant. The idea of using a noise-tolerant
 Barker's test function was suggested but not explored empirically
-in~\cite{TallData16} section 6.3. But the asymptotic test statistic CDF and the
+in~\citep{TallData16} section 6.3. But the asymptotic test statistic CDF and the
 Barker function are different, which leads to fixed errors for the approach
-in~\cite{TallData16}. Here, we show that the difference between the
+in~\citep{TallData16}. Here, we show that the difference between the
 distributions can be corrected with an additive random variable. This leads to a
@@
 \begin{itemize}[noitemsep]
-    \item Increase the temperature of the target distribution. Log likelihoods
-    scale as $1/T$, and so the variance of the likelihood ratio will vary as
-    $1/T^2$. Our model is no longer well-specified (we are doing inference at a
-    temperature different from that assumed during data generation), but higher
+    \item Increase the temperature $K$ of the target distribution. Log
+    likelihoods scale as $1/K$, and so the variance of the likelihood ratio will
+    vary as $1/K^2$. Our model is no longer well-specified (we are doing
+    inference at a temperature different from that assumed during data
+    generation), but as we demonstrate in Section~\ref{ssec:logistic}, higher
     temperature can be advantageous for parameter exploration.
@@
 sampling from it is one application, but an efficient
-proposal~\cite{OptimalScaling01} has similar variance to the target and will
+proposal~\citep{OptimalScaling01} has similar variance to the target and will
 diffuse to it extremely slowly. For applications to neural networks or models
-where the posterior is multimodal~\cite{choromanska2014loss}, samplers will
+where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
 likely get trapped in one of the modes. A common solution is to anneal the
@@
 small steps rather than taking a few sparse steps towards an optimum. In this
-mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
-Descent. The goal in SGD is to make gradual progress to a posterior mode with
-each step, taking small steps so that the cumulative displacement has
-progressively lower variance.
+mode, M-H tests can be used in similar fashion to Stochastic Gradient Descent.
+The goal in SGD is to make gradual progress to a posterior mode with each step,
+taking small steps so that the cumulative displacement has progressively lower
+variance.
 
@@
 where $p_0(\theta)$ is the prior. Computing samples this way requires all $N$
-data points, but this is very expensive for large datasets. To address this
-challenge,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
-Metropolis-Hasting tests using sequential hypothesis testing. Each iteration,
-they start with a small minibatch and test whether $\theta'$ should be accepted
-based on approximating $u<\alpha(\theta,\theta')$. If the approximate test
-cannot decide with sufficient confidence, the minibatch size is increased and
-the test repeats. This process continues until a decision. The bounds depend on
-either an asymptotic Central Limit Theorem~\cite{cutting_mh_2014} or a
-concentration bound~\cite{icml2014c1_bardenet14}. The latter requires direct
-bounds on the log likelihood ratio, which for general distributions requires
-knowing $p(x_i | \theta)$ and $p(x_i | \theta')$ for all $N$ samples. In
-addition, while both methods show useful reductions in the number of samples
-required, they suffer the drawback of resolving small log likelihood ratio
-differences between the minibatch and full batch versions. We discuss a
-worst-case scenario in Section~\ref{ssec:gaussian_example}.
-
-Following~\cite{icml2014c1_bardenet14}, we write the test
+data points, but this is very expensive for large datasets.
+
+To address this challenge,~\citep{cutting_mh_2014,icml2014c1_bardenet14} perform
+approximate Metropolis-Hasting tests using sequential hypothesis testing. Each
+iteration, they start with a small minibatch and test whether $\theta'$ should
+be accepted based on approximating $u<\alpha(\theta,\theta')$. If the
+approximate test cannot decide with sufficient confidence, the minibatch grows
+and the test repeats. This process continues until a decision. The bounds depend
+on either an asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
+concentration bound~\citep{icml2014c1_bardenet14}. We refer to these algorithms,
+respectively, as {\sc AustereMH} and {\sc MHSubLhd}.
+
+In~\citep{cutting_mh_2014}, CLT asymptotic arguments are used to show that their
+approximate acceptance test error (with respect to full-batch M-H) tends to zero
+as batch size increases, but no quantitative bounds are given. In addition, to
+design a test that minimizes data usage given an error bound, they rely on
+computing a standardized mean $\mu_{\rm std}$ each iteration. As the value of
+$\mu_{\rm std}$ depends on all $\log p(x_i|\theta)$ and $\log p(x_i| \theta')$
+terms, the test is technically not a minibatch test; \citep{cutting_mh_2014}
+propose a conservative variant which assumes $\mu_{\rm std}=0$ and avoids the
+need to evaluate all likelihood terms. We refer to this as {\sc AustereMH(c)}
+and the default, non-conservative variant as {\sc AustereMH(nc)}.
+
+For {\sc MHSubLhd}, the test requires direct bounds on the log likelihood ratio,
+which for general distributions requires knowing $p(x_i | \theta)$ and $p(x_i |
+\theta')$ for all $N$ samples. In~\citep{icml2014c1_bardenet14}, explicit bounds
+are given and depend on
+\begin{equation}\label{eq:bad_bound}
+    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
+\end{equation}
+It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
+regression and other simple models, but it is difficult to do so for more
+complicated models, so one generally needs to use all $p(x_i | \theta')$
+terms\footnote{The sample code provided by~\citep{icml2014c1_bardenet14}
+computes $C_{\theta,\theta'}$ by traversing the entire data, thus providing no
+performance advantage over the complete test.} to get $C_{\theta,\theta'}$, like
+the {\sc AustereMH(nc)} test.
+
+In contrast, as we show in Section~\ref{sec:analysis}, we use quantitative forms
+of the CLT which rely on measurable statistics from a \emph{single} minibatch.
+In addition, while both {\sc AustereMH(nc)} and {\sc MHSubLhd} show useful
+reductions in the number of samples required, they suffer the drawback of
+resolving small log likelihood ratio differences between the minibatch and full
+batch versions. We discuss a worst-case scenario in
+Section~\ref{ssec:gaussian_example}.
+
+\subsection{NOTATION}\label{ssec:notation}
+
+Following~\citep{icml2014c1_bardenet14}, we write the test
 $u<\alpha(\theta,\theta')$ equivalently as $\Lambda(\theta,\theta') >
 \psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those
-in~\cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis
+in~\citep{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis
 later.}
-\begin{equation}
+\begin{equation}\label{eq:lambda}
 \begin{split}
@@
 \end{equation}
+To simplify notation, we assume that temperature $K=1$ (saving $T$ to indicate
+the number of samples to draw). Temperature appears as an exponential on each
+likelihood, $p(x_i|\theta)^{1/K}$, so the effect would be to act as a $1/K$
+factor on $\Lambda(\theta,\theta')$.
+
 To reduce computational effort, an unbiased estimate of $\Lambda(\theta,\theta')$
@@
 \Lambda^*(\theta,\theta') = \frac{N}{b}\sum_{i=1}^b 
-\log \frac{p(x_i^*|\theta')}{p(x_i^*|\theta)} 
+\log \frac{p(x_i^*|\theta')}{p(x_i^*|\theta)}.
 \end{equation}
@@
 distribution has the same variance as the target
-distribution~\cite{OptimalScaling01}, so we use a proposal based on
+distribution~\citep{OptimalScaling01}, so we use a proposal based on
 $\mathcal{N}(\theta,1/N)$. It is symmetric
@@
     (independent of $N$) constant $c$ such that with probability $\geq c$ over
-    the joint distribution of $(\theta, \theta', u)$, the tests
-    from~\cite{cutting_mh_2014,icml2014c1_bardenet14} consume all $N$ samples. 
+    the joint distribution of $(\theta, \theta', u)$, {\sc AustereMH(nc)} and
+    {\sc MHSubLhd} consume all $N$ samples. 
 \end{lemma}
@@
 There is a separate line of MCMC work drawing principles from statistical
-physics. One can apply
-Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which
-generate high acceptance \emph{and} distant proposals when run on full batches
-of data. Recently Langevin Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has
-been applied to Bayesian estimation on minibatches of data. This simplified
-dynamics uses local proposals and avoids M-H tests by using small proposal steps
-whose acceptance approaches 1 in the limit. However, the constraint on proposal
-step size is severe, and the state space exploration reduces to a random walk.
-Full minibatch HMC for minibatches was described in~\cite{sghmc_2014} which
-allows momentum-augmented proposals with larger step sizes. However, step sizes
-are still limited by the need to run accurately without M-H tests.  By providing
-an M-H test with similar cost to standard gradient steps, our work opens the
-door to applying those methods with much more aggressive step sizes without loss
-of accuracy.
+physics. One can apply Hamiltonian Monte Carlo
+(HMC)~\citep{mcmc_hamiltonian_2010} methods which generate high acceptance
+\emph{and} distant proposals when run on full batches of data. Recently Langevin
+Dynamics~\citep{langevin_2011,conf/icml/AhnBW12} has been applied to Bayesian
+estimation on minibatches of data. This simplified dynamics uses local proposals
+and avoids M-H tests by using small proposal steps whose acceptance approaches 1
+in the limit. However, the constraint on proposal step size is severe, and the
+state space exploration reduces to a random walk.  Full minibatch HMC for
+minibatches was described in~\citep{sghmc_2014} which allows momentum-augmented
+proposals with larger step sizes. However, step sizes are still limited by the
+need to run accurately without M-H tests.  By providing an M-H test with similar
+cost to standard gradient steps, our work opens the door to applying those
+methods with much more aggressive step sizes without loss of accuracy.
 
@@
 
-This result is used in~\cite{Barker65} to define the Barker acceptance test.
+This result is used in~\citep{Barker65} to define the Barker acceptance test.
 
 \subsection{BARKER (LOGISTIC) ACCEPTANCE FUNCTION}\label{ssec:barker_function}
-For our new MH test we use the Barker logistic~\cite{Barker65}
+For our new MH test we use the Barker logistic~\citep{Barker65}
 function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
@@
 test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
-random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{TallData16} it
+random variable $X$ (call it $X_{\rm log}$ from now on). In~\citep{TallData16} it
 was proposed to use $\Delta^*$ in a Barker test, and tolerate the fixed
@@
 
-\textbf{Daniel: TODO this should be where we mention that AustereMH still
-requires computation of all likelihoods, and we should explain John's complaint
-about Gaussianity here.}
-
 We now derive error bounds for our M-H test and the target distribution it
-generates. In the most similar prior works,~\cite{cutting_mh_2014} use CLT
-asymptotic arguments to show that its approximate acceptance test error tends to
-zero as batch size increases, but no quantitative bounds are given.
-In~\cite{icml2014c1_bardenet14}, explicit bounds are given and depend on
-\begin{equation}\label{eq:bad_bound}
-    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
-\end{equation}
-It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
-regression, but it is unclear how to do so for more complicated models, and one
-generally needs to use all $p(x_i | \theta')$ terms\footnote{The sample code
-provided by~\cite{icml2014c1_bardenet14} computes $C_{\theta,\theta'}$
-by traversing the entire data, thus providing no performance advantage over the
-complete test.} to get $C_{\theta,\theta'}$.  In contrast, we use quantitative
-forms of the CLT which rely on measurable statistics from a \emph{single}
-minibatch.
-
-In Section~\ref{ssec:delta_star_distribution}, we present bounds on the absolute
-and relative error (in terms of the CDFs) of the distribution of $\Delta^*$ vs.
-a Gaussian. We then show in Section~\ref{ssec:preserve_bounds} that these bounds
-are preserved after the addition of other random variables (e.g., $X_{\rm nc}$
-and $X_{\rm corr}$). It then follows that the acceptance test has the same error
-bound.
+generates. In Section~\ref{ssec:delta_star_distribution}, we present bounds on
+the absolute and relative error (in terms of the CDFs) of the distribution of
+$\Delta^*$ vs.  a Gaussian. We then show in Section~\ref{ssec:preserve_bounds}
+that these bounds are preserved after the addition of other random variables
+(e.g., $X_{\rm nc}$ and $X_{\rm corr}$). It then follows that the acceptance
+test has the same error bound.
 
@@
 \begin{corollary}\label{cor:our_bound_delta_prime}
+\small
 \begin{equation}
-\small
 \sup_y \left|{\rm Pr}(\Delta^*< y) -\Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| \leq \frac{6.4\mE[|X|^3]+2\mE[|X|]}{\sqrt{b}}
 \end{equation}
+\normalsize
 \end{corollary} %\vspace{-10pt}
@@
 of the Markov chain under appropriate conditions. Such bounds were derived in
-both~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}. We include the
-result from~\cite{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the
+both~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14}. We include the
+result from~\citep{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the
 total variation distance between two distributions $P$ and $Q$.  Let
@@
 % "number of samples of \theta we draw".
-In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark our test,
-{\sc MHminibatch}, against the test from~\cite{icml2014c1_bardenet14} ({\sc
-MHSubLhd}) and from~\cite{cutting_mh_2014}. As highlighted in
-Section~\ref{sec:analysis}, the test from~\cite{cutting_mh_2014} can be
-implemented using a conservative or a non-conservative variant; we name these
-{\sc AustereMH(c)}, and {\sc AustereMH(nc)}, respectively, and they use the
-grid-search over minibatch sizes $m$ and per-test thresholds $\epsilon$
-described in Appendix~\ref{sssec:grid_search}. Throughout our descriptions, we
-refer to a \emph{trial} as the period when an algorithm collects all its desired
-samples $\{\theta_1,\ldots,\theta_T\}$, generally with $T=3000$ or $T=5000$.
+In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark {\sc
+MHminibatch} against the test from~\citep{icml2014c1_bardenet14} ({\sc
+MHSubLhd}) and from~\citep{cutting_mh_2014} ({\sc AustereMH(c)}, and {\sc
+AustereMH(nc)}). The latter two tests use a grid-search over minibatch sizes $m$
+and per-test thresholds $\epsilon$ described in
+Appendix~\ref{sssec:grid_search}. Throughout our descriptions, we refer to a
+\emph{trial} as the period when an algorithm collects all its desired samples
+$\{\theta_1,\ldots,\theta_T\}$, generally with $T=3000$ or $T=5000$.
 
@@
 
-This model is adapted from~\cite{langevin_2011} by increasing the number of
+This model is adapted from~\citep{langevin_2011} by increasing the number of
 samples to 1 million.  The parameters are $\theta = \langle \theta_1,\theta_2
@@
 sampled 1,000,000 points to form the posterior of
-$p(\theta)\prod_{i=1}^{1,000,000}p(x_i | \theta)$ with the same prior from
+$p(\theta)\prod_{i=1}^{1,000,000}p(x_i | \theta)^{1/K}$ with the same prior from
 Equation~(\ref{eq:data_generation}). This produces a much sharper posterior with
@@
 
-We benchmark with {\sc AustereMH(c)} and {\sc MHSubLhd}. {\sc MHminibatch} and
-{\sc MHSubLhd} were initialized with $m=100$. For {\sc MHSubLhd}, we increase
+We benchmark with {\sc AustereMH(c)} and {\sc MHSubLhd}. We initialized {\sc
+MHminibatch} and {\sc MHSubLhd} with $m=100$. For {\sc MHSubLhd}, we increase
 sizes geometrically with $\gamma = 1.5$ and use parameters $p = 2, \delta =
 0.01$.  All methods collect 5000 samples using a random walk proposer with
-covariance ${\rm diag}(0.15, 0.15)$, which means the M-H test is responsible for
-shaping the sample distribution.
+covariance matrix ${\rm diag}(0.15, 0.15)$, which means the M-H test is
+responsible for shaping the sample distribution.
 
@@
 set of samples and the actual posterior. Due to space constraints, results are
-in Appendix~\ref{app:gaussians}.
+in Appendix~\ref{sssec:gauss_metrics}.
 
@@
 magnitude better compared to the other two methods (averages are 15562 and
-16857). The sizes correspond to the running times of the methods, excluding the
-likelihood computation of all data points for {\sc MHSubLhd}, which would
-drastically increase running time.
+16857). \textbf{Daniel: Xinlei, you should change these numbers as you see fit,
+and also describe the number of trials you ran.} The sizes correspond to the
+running times of the methods, excluding the likelihood computation of all data
+points for {\sc MHSubLhd}, which would drastically increase running time.
 
@@
 We next test logistic regression for the binary classification of 1s versus 7s
-on the MNIST~\cite{lecun-mnisthandwrittendigit-2010} dataset and (a subset of)
-infinite MNIST~\cite{loosli-canu-bottou-2006}. For the former, extracting all 1s
+on the MNIST~\citep{lecun-mnisthandwrittendigit-2010} dataset and (a subset of)
+infinite MNIST~\citep{loosli-canu-bottou-2006}. For the former, extracting all 1s
 and 7s resulted in 13,000 training samples, and for the latter, we used 87,000
@@
 For {\sc MHSubLhd}, we tried to use the provided symbolic bound for
-$C_{\theta,\theta'}$ described in~\cite{icml2014c1_bardenet14}, but it was too
+$C_{\theta,\theta'}$ described in~\citep{icml2014c1_bardenet14}, but it was too
 high and provided no performance benefit. Instead we use the empirical
@@
 generate the curves, for each of the sampled vectors $\theta_t$,
-$t\in\{1,\ldots,N\}$ ($N=5000$ for MNIST-13k, $N=3000$ for MNIST-100k), we use
-$\theta_t$ as the parameter for logistic regression. The results indicate that
-our test is more efficient, obtaining convergence more than an order of
-magnitude faster than {\sc AustereMH(nc)} and several orders of magnitude
-compared to {\sc AustereMH(c)} and {\sc MHSubLhd}. We also observe the advantage
-of having higher temperature from the third plot in
+$t\in\{1,\ldots,T\}$, we use $\theta_t$ as the logistic regression parameter.
+The results indicate that our test is more efficient, obtaining convergence more
+than an order of magnitude faster than {\sc AustereMH(nc)} and several orders of
+magnitude compared to {\sc AustereMH(c)} and {\sc MHSubLhd}.  We also observe
+the advantage of having higher temperature from the third plot in
 Figure~\ref{fig:logistic_performance}. During the exploration period, the
@@
 Figures~\ref{fig:logistic_histograms_mnist}
-and~\ref{fig:logistic_histograms_mnist8m} show histograms of minibatch sizes for
-the three methods on a log-log scale for MNIST-13k and MNIST-100k, respectively.
-The histograms only represent one representative trial; Table~\ref{tab:logistic}
-contains the average of the average minibatch sizes ($\pm$ one standard
-deviation) across either 10 or 5 trials. The results indicate that {\sc
-MHminibatch}, with average minibatch sizes of roughly 129.8 and 202.5 for
-MNIST-13k and MNIST-100k, respectively, consumes almost an order of magnitude
-fewer data points than the next-best method, {\sc AustereMH(nc)}.  We reiterate,
-however, that both {\sc AustereMH(nc)} and {\sc MHSubLhd} require computing
-$\log p(x_i|\theta)$ and $\log p(x_i|\theta')$ for all $x_i$ each iteration.
-These results here do not count that extra data consumption. Only our method and
-{\sc AustereMH(c)} rely solely on the minibatch of data each iteration.
+and~\ref{fig:logistic_histograms_mnist8m} show log-log histograms of minibatch
+sizes for the three methods.  The histograms only represent one representative
+trial; Table~\ref{tab:logistic} contains the average of the average minibatch
+sizes ($\pm$ one standard deviation) across all trials. {\sc MHminibatch}, with
+average minibatch sizes of roughly 129.8 and 202.5 for MNIST-13k and MNIST-100k,
+respectively, consumes more than 5x fewer data points than the next-best method,
+{\sc AustereMH(nc)}.  We reiterate, however, that both {\sc AustereMH(nc)} and
+{\sc MHSubLhd} require computing $\log p(x_i|\theta)$ and $\log p(x_i|\theta')$
+for all $x_i$ each iteration. Our results here do not count that extra data
+    consumption. Only our method and {\sc AustereMH(c)} rely solely on the
+    minibatch of data each iteration.
 
@@
 
-\textbf{Daniel: TODO we wanted a third experiment, right?}
+\textbf{Daniel: TODO this is our third experiment. Results are in progress}
 
@@
 factorization.  Directions for future work include integrating our algorithm
-with~\cite{cutting_mh_2014} by applying both tests each iteration or utilizing
-the variance reduction techniques suggested in~\cite{DBLP:conf/icml/ChenG16}.
+with~\citep{cutting_mh_2014} by applying both tests each iteration or utilizing
+the variance reduction techniques suggested in~\citep{DBLP:conf/icml/ChenG16}.
 More elaborate extensions involve running more experiments on other models such
 as neural networks and providing a recipe for how to use our algorithm
-(following the framework of~\cite{sgmcmc_2015}).
+(following the framework of~\citep{sgmcmc_2015}).
 
 \bibliography{paper}
-\bibliographystyle{plain}
+%\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 
@@
 Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta')
-\not\approx \psi(u,\theta,\theta')$ used in ~\cite{cutting_mh_2014}
-and~\cite{icml2014c1_bardenet14} , where $\not\approx$ means ``significantly
+\not\approx \psi(u,\theta,\theta')$ used in ~\citep{cutting_mh_2014}
+and~\citep{icml2014c1_bardenet14} , where $\not\approx$ means ``significantly
 different from'' under the distribution over samples. This is
@@
 
-The following bound is given immediately after Corollary 2 from~\cite{explicit-clt05}:
+The following bound is given immediately after Corollary 2 from~\citep{explicit-clt05}:
 \begin{equation}
@@
 then Equation 2.2 in Bentkus et al.'s work on Edgeworth
-expansions~\cite{Bentkus97} provides:
+expansions~\citep{Bentkus97} provides:
 
@@
 \textbf{Metric} & {\sc MHminibatch} & {\sc AustereMH} & {\sc MHSubLhd} \\
-\hline \\
-Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
+\hline
+\T Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
 Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
@@
 For the Gaussian mixture experiment, we use the conservative method
-from~\cite{cutting_mh_2014}, which avoids the need for recomputing log
+from~\citep{cutting_mh_2014}, which avoids the need for recomputing log
 likelihoods of each data point each iteration by choosing baseline minibatch
@@
 perform significance tests using Chi-Squared distribution (also in
-Table~\ref{tab:poissons}). Scores lie between~\cite{cutting_mh_2014}
-and~\cite{icml2014c1_bardenet14}, but the variance of these values is high and
+Table~\ref{tab:poissons}). Scores lie between~\citep{cutting_mh_2014}
+and~\citep{icml2014c1_bardenet14}, but the variance of these values is high and
 ordering changes depending on the range of samples generated. \textbf{Daniel:
@@
 stated, values apply to all methods tested. For the methods
-from~\cite{cutting_mh_2014}, we use their notation for clarity.}
+from~\citep{cutting_mh_2014}, we use their notation for clarity.}
 \small

### 2017-03-02 uai2017/paper.tex
Whew, logistic regression done.

@@
 
-% Daniel: NOTE, I'm calling temperature "K" so that "T" stays reserved for
-% "number of samples of \theta we draw".
-In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark {\sc
-MHminibatch} against the test from~\citep{icml2014c1_bardenet14} ({\sc
-MHSubLhd}) and from~\citep{cutting_mh_2014} ({\sc AustereMH(c)}, and {\sc
-AustereMH(nc)}). The latter two tests use a grid-search over minibatch sizes $m$
-and per-test thresholds $\epsilon$ described in
-Appendix~\ref{sssec:grid_search}. Throughout our descriptions, we refer to a
-\emph{trial} as the period when an algorithm collects all its desired samples
-$\{\theta_1,\ldots,\theta_T\}$, generally with $T=3000$ or $T=5000$.
-
-\subsection{MIXTURE OF GAUSSIANS}\label{ssec:gaussians}
-
 % Daniel: moving this here so figures don't overflow to next pages. We can tune
@@
 
+% Daniel: NOTE, I'm calling temperature "K" so that "T" stays reserved for
+% "number of samples of \theta we draw".
+In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark {\sc
+MHminibatch} against the tests from~\citep{icml2014c1_bardenet14}
+and~\citep{cutting_mh_2014}, with the latter using a grid-search over minibatch
+sizes $m$ and per-test thresholds $\epsilon$ described in
+Appendix~\ref{sssec:grid_search}. Throughout our descriptions, we refer to a
+\emph{trial} as the period when an algorithm collects all its desired samples
+$\{\theta_1,\ldots,\theta_T\}$, generally with $T=3000$ or $T=5000$.
+
+\subsection{MIXTURE OF GAUSSIANS}\label{ssec:gaussians}
+
 This model is adapted from~\citep{langevin_2011} by increasing the number of
@@
 magnitude better compared to the other two methods (averages are 15562 and
-16857). \textbf{Daniel: Xinlei, you should change these numbers as you see fit,
-and also describe the number of trials you ran.} The sizes correspond to the
-running times of the methods, excluding the likelihood computation of all data
-points for {\sc MHSubLhd}, which would drastically increase running time.
+16857). Sizes correspond to the running times of the methods, excluding the
+likelihood computation of all data points for {\sc MHSubLhd}, which would
+drastically increase running time.
 
@@
     function of cumulative data usage.  The right plot reports performance of
-    {\sc MHminibatch} on both datasets when the temperature starts at $K=100$
-    and drops to $K=1$ after a ``burn-in'' period of 200 samples of $\theta$.
-    For all three plots, one standard deviation is indicated by the shaded error
-    regions.
+    {\sc MHminibatch} on both datasets when the temperature starts at 100 and
+    drops to 1 after a ``burn-in'' period of 200 samples (vertical dashed line)
+    of $\theta$.  For all three plots, one standard deviation is indicated by
+    the shaded error regions.
     }
@@
 
-\begin{figure*}[t]
-	\centering
-    \includegraphics[width=0.9\linewidth]{logistic_regression_histograms_mnist.png}
-	\caption{
-    Minibatch sizes for a representative trial of logistic regression on
-    MNIST-13k (analogous to Figure~\ref{fig:gauss_mix_2}). Both axes are on a
-    log scale and have the same ranges across the three histograms. See
-    Section~\ref{ssec:logistic} for details.
-    }
-	\label{fig:logistic_histograms_mnist}
-\end{figure*}
-
+% Moving here to get positioning better
 \begin{figure*}[t]
@@
     Minibatch sizes for a representative trial of logistic regression on
-    MNIST-100k (similar to Figure~\ref{fig:logistic_histograms_mnist}). Both
-    axes are on a log scale and have the same ranges across the three
-    histograms.  See Section~\ref{ssec:logistic} for details.
+    MNIST-100k (analogous to Figure~\ref{fig:gauss_mix_2}). Both axes are on a
+    log scale and have the same ranges across the three histograms.  See
+    Section~\ref{ssec:logistic} for details.
     }
@@
 
+We next test logistic regression for the binary classification of 1s versus 7s
+on the MNIST~\citep{lecun-mnisthandwrittendigit-2010} dataset and (a subset of)
+infinite MNIST~\citep{loosli-canu-bottou-2006}. For the former, extracting all 1s
+and 7s resulted in 13,000 training samples, and for the latter, we used 87,000
+additional (augmented) 1s and 7s to get 100,000 training samples. Both datasets
+use the same test set, with 2,163 samples. Henceforth, we call them MNIST-13k
+and MNIST-100k, respectively.
+
+For all methods, we impose a uniform prior on $\theta$ and again use a random
+walk proposer, with covariance matrix $0.05I$ for MNIST-13k and $0.01I$ for
+MNIST-100k. The default temperature setting is a constant at $K=100$ for
+MNIST-13k and MNIST-100k. Performance of all methods implicitly relies on the
+step size and temperature. Setting temperature too low or step size too high
+will result in slow convergence for all methods. For MNIST-13k, each method
+generated 5000 samples for ten independent trials; due to MNIST-100k's higher
+computational requirements, the methods generated 3000 samples for five
+independent trials.  For additional parameter settings and an investigation on
+tuning step sizes, see Appendix~\ref{app:logistic}.
+
+For {\sc MHSubLhd}, we tried to use the provided symbolic bound for
+$C_{\theta,\theta'}$ described in~\citep{icml2014c1_bardenet14}, but it was too
+high and provided no performance benefit. Instead we use the empirical
+$C_{\theta,\theta'}$ from the entire dataset.
+
+The first two subplots of Figure~\ref{fig:logistic_performance} display the
+prediction accuracy on both datasets for all methods as a function of the
+cumulative training points processed.\footnote{The curves do not span the same
+length over the x-axis since the methods consume different amounts of data.} To
+generate the curves, for each of the sampled vectors $\theta_t$,
+$t\in\{1,\ldots,T\}$, we use $\theta_t$ as the logistic regression parameter.
+The results indicate that our test is more efficient, obtaining convergence more
+than an order of magnitude faster than {\sc AustereMH(nc)} and several orders of
+magnitude compared to {\sc AustereMH(c)} and {\sc MHSubLhd}.  We also observe
+the advantage of having higher temperature from the third plot in
+Figure~\ref{fig:logistic_performance}. During the exploration period, the
+accuracy rapidly increases, and then after 200 samples, we switch the
+temperature to 1, but this requires the step size to decrease, hence the smaller
+changes in accuracy.
+
+Figure~\ref{fig:logistic_histograms_mnist8m} shows log-log histograms of
+minibatch sizes for the three methods on MNIST-100k.
+(Figure~\ref{fig:logistic_histograms_mnist} in Appendix~\ref{app:logistic}
+contains results for MNIST-13k.) The histograms only represent one
+representative trial; Table~\ref{tab:logistic} contains the average of the
+average minibatch sizes ($\pm$ one standard deviation) across all trials. {\sc
+MHminibatch}, with average minibatch sizes of roughly 129.8 and 202.5 for
+MNIST-13k and MNIST-100k, respectively, consumes more than 5x fewer data points
+than the next-best method, {\sc AustereMH(nc)}.  We reiterate, however, that
+both {\sc AustereMH(nc)} and {\sc MHSubLhd} require computing $\log
+p(x_i|\theta)$ and $\log p(x_i|\theta')$ for all $x_i$ each iteration. Our
+results here do not count that extra data consumption. Only our method and {\sc
+AustereMH(c)} rely solely on the minibatch of data each iteration.
+
+\subsection{SPARSE MATRIX FACTORIZATION}\label{ssec:collab_filtering}
+
 \begin{table}[t]
-\caption{
-Average minibatch sizes ($\pm$ one standard deviation) for the methods on the
-logistic regression tasks of binary classification on MNIST-13k and MNIST-100k.
-The averages are taken over 10 independent trials (5000 samples each) for
-MNIST-13k and 5 independent trials (3000 samples each) for MNIST-100k.
-}
+\caption{Average minibatch sizes ($\pm$ one standard deviation) on logistic
+regression on MNIST-13k and MNIST-100k.  The averages are taken over 10
+independent trials (5000 samples each) for MNIST-13k and 5 independent trials
+(3000 samples each) for MNIST-100k.}
 \small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
@@
 
-We next test logistic regression for the binary classification of 1s versus 7s
-on the MNIST~\citep{lecun-mnisthandwrittendigit-2010} dataset and (a subset of)
-infinite MNIST~\citep{loosli-canu-bottou-2006}. For the former, extracting all 1s
-and 7s resulted in 13,000 training samples, and for the latter, we used 87,000
-additional (augmented) 1s and 7s to get 100,000 training samples. Both datasets
-use the same test set, with 2,163 samples. Henceforth, we call them MNIST-13k
-and MNIST-100k, respectively.
-
-For all methods, we impose a uniform prior on $\theta$ and again use a random
-walk proposer, with covariance matrix $0.05I$ for MNIST-13k and $0.01I$ for
-MNIST-100k. The default temperature setting is a constant at $K=100$ for
-MNIST-13k and MNIST-100k. Performance of all methods implicitly relies on the
-step size and temperature. Setting temperature too low or step size too high
-will generally result in worse performance for all methods. For MNIST-13k, each
-method generated 5000 samples for ten independent trials; due to MNIST-100k's
-higher computational requirements, the methods generated 3000 samples for five
-independent trials.  For additional parameter settings and an investigation on
-tuning step sizes, see Appendix~\ref{app:logistic}.
-
-For {\sc MHSubLhd}, we tried to use the provided symbolic bound for
-$C_{\theta,\theta'}$ described in~\citep{icml2014c1_bardenet14}, but it was too
-high and provided no performance benefit. Instead we use the empirical
-$C_{\theta,\theta'}$ from the entire dataset.
-
-The first two subplots of Figure~\ref{fig:logistic_performance} display the
-prediction accuracy on both datasets for all methods as a function of the
-cumulative training points processed.\footnote{The curves do not span the same
-length over the x-axis since the methods consume different amounts of data.} To
-generate the curves, for each of the sampled vectors $\theta_t$,
-$t\in\{1,\ldots,T\}$, we use $\theta_t$ as the logistic regression parameter.
-The results indicate that our test is more efficient, obtaining convergence more
-than an order of magnitude faster than {\sc AustereMH(nc)} and several orders of
-magnitude compared to {\sc AustereMH(c)} and {\sc MHSubLhd}.  We also observe
-the advantage of having higher temperature from the third plot in
-Figure~\ref{fig:logistic_performance}. During the exploration period, the
-accuracy rapidly increases, and then after 200 samples, we switch the
-temperature to 1 (equivalent to ignoring temperature). {\rm MHminibatch}
-requires the step size to decrease, hence the smaller changes in accuracy.
-
-Figures~\ref{fig:logistic_histograms_mnist}
-and~\ref{fig:logistic_histograms_mnist8m} show log-log histograms of minibatch
-sizes for the three methods.  The histograms only represent one representative
-trial; Table~\ref{tab:logistic} contains the average of the average minibatch
-sizes ($\pm$ one standard deviation) across all trials. {\sc MHminibatch}, with
-average minibatch sizes of roughly 129.8 and 202.5 for MNIST-13k and MNIST-100k,
-respectively, consumes more than 5x fewer data points than the next-best method,
-{\sc AustereMH(nc)}.  We reiterate, however, that both {\sc AustereMH(nc)} and
-{\sc MHSubLhd} require computing $\log p(x_i|\theta)$ and $\log p(x_i|\theta')$
-for all $x_i$ each iteration. Our results here do not count that extra data
-    consumption. Only our method and {\sc AustereMH(c)} rely solely on the
-    minibatch of data each iteration.
-
-\subsection{SPARSE MATRIX FACTORIZATION}\label{ssec:collab_filtering}
-
 \begin{table}[t]
@@
 
-\textbf{Daniel: TODO this is our third experiment. Results are in progress}
+\textbf{Daniel: TODO this is our third experiment. Results are in progress...}
+
+\textbf{We benchmark with the ADAGrad~\citep{adagrad} implementation of Sparse
+Matrix Factorization described in~\citep{Canny:2013:BDA:2487575.2487677}.
+(Daniel: is this the correct citation for BIDMach? I assume this is what we will
+use)}
 
@@
 	\centering
+    \includegraphics[width=0.9\linewidth]{logistic_regression_histograms_mnist.png}
+	\caption{
+    Minibatch sizes for a representative trial of logistic regression on
+    MNIST-13k (analogous to Figure~\ref{fig:gauss_mix_2}). Both axes are on a
+    log scale and have the same ranges across the three histograms. See
+    Section~\ref{ssec:logistic} for details.
+    }
+	\label{fig:logistic_histograms_mnist}
+\end{figure*}
+
+\begin{figure*}[t]
+	\centering
     \includegraphics[width=0.9\linewidth]{logistic_regression_step_size_tests.png}
@@
 \begin{table}[t]
-\caption{Parameters for the logistic regression experiments. Unless otherwise
-stated, values apply to all methods tested. For the methods
-from~\citep{cutting_mh_2014}, we use their notation for clarity.}
+\caption{Parameters for the logistic regression experiments.}
 \small
@@
 {\sc MHminibatch} and {\sc MHSubLhd} minibatch size $m$ & 100 & 100 \\
-{\sc AustereMH(c)} $\Delta^*$ bound & 0.1 & 0.2 \\
+{\sc AustereMH(c)} chosen $\Delta^*$ bound & 0.1 & 0.2 \\
 {\sc AustereMH(c)} minibatch size $m$ from grid search & 450 & 300 \\
 {\sc AustereMH(c)} per-test threshold $\epsilon$ from grid search & 0.01 & 0.01\\
-{\sc AustereMH(nc)} $\Delta^*$ bound & 0.05 & 0.1 \\
+{\sc AustereMH(nc)} chosen $\Delta^*$ bound & 0.05 & 0.1 \\
+{\sc MHSubLhd} $\gamma$ & 2.0  & 2.0  \\
+{\sc MHSubLhd} $p$      & 2 & 2 \\
+{\sc MHSubLhd} $\delta$ & 0.01 & 0.01 \\
 \end{tabular}
@@
 
-\textbf{Daniel: TODO major work in progress.}
-
-Table~\ref{tab:logistic_settings} indicates the parameter settings for the
-logistic regression experiments.
-
-Figure~\ref{fig:logistic_histograms_appendix} indicates minibatch histograms for
-one trial of {\sc MHminibatch} on MNIST-100k for proposers $\sigma I$ with
-$\sigma \in \{0.005, 0.01, 0.05\}$.
+Figure~\ref{fig:logistic_histograms_mnist} shows the histograms for the four
+methods on one representative trial of MNIST-13k, which shows similar relative
+performance of the four methods as shown in
+Figure~\ref{fig:logistic_histograms_mnist8m} (which uses MNIST-100k). In
+particular, {\sc MHminibatch} exhibits a shorter-tailed distribution and
+consumes nearly an order of magnitude fewer data points compared to {\sc
+AustereMH(nc)}, the next-best method; see Table~\ref{tab:logistic} for details.
+
+Next, we investigate the impact of the step size $\sigma$ for the random walk
+proposers with covariance matrix $\sigma I$. Note that $I$ is $784\times 784$ as
+we did not perform any downsampling or data preprocessing other than rescaling
+the pixel values to lie in $[0,1]$.
+
+For this, we use the larger dataset MNIST-100k, and test with $\sigma \in
+\{0.005, 0.01, 0.05\}$. We keep other parameters consistent with the experiments
+in Section~\ref{ssec:logistic}, in particular, keeping the initial minibatch
+size $m=100$, which is also the amount the minibatch increments by if we need
+more data. Figure~\ref{fig:logistic_histograms_appendix} indicates minibatch
+histograms (again, using the log-log scale) for one trial of {\sc MHminibatch}
+using each of the step sizes. We observe that by tuning {\sc MHminibatch}, we
+are able to adjust the average number of data points in a minibatch across a
+wide range of values. Here, the smallest step size results in an average of just
+116.1 data points per minibatch, while increasing to $\sigma=0.05$ (the step
+size used for MNIST-13k) results in an average of 2215.6. This relative trend is
+also present for both {\sc AustereMH} variants and {\sc MHSubLhd}.
+
+Table~\ref{tab:logistic_settings} indicates the relevant parameter settings for
+the logistic regression experiments. Unless otherwise stated, values apply to
+all methods tested. For values from~\citep{cutting_mh_2014}
+or~\cite{icml2014c1_bardenet14}, we use their notation to be consistent.
 

### 2017-03-03 uai2017/paper.tex
Reorganized paper based on John's feedback.
Moved some preliminary stuff later, and added slightly more info. Also added a bit on hamiltonian dynamics. OK, still waiting for Xinlei and I to finish this.

@@
 
-%  \item Increase the minibatch size when needed. Log acceptance variance
-%    scales as $1/k$ vs the minibatch size $k$. Our test is adaptive like earlier
-%    works, but unlike them, the distribution of minibatch size is Gaussian, not
-%    long-tailed.  Increased minibatch size also reduces the error rate for the
-%    test.
+    \item Utilize Hamiltonian and Langevin Dynamics to judiciously model a system
+    so as to generate distant yet high-quality proposals.
 \end{itemize}
@@
 concentration bound~\citep{icml2014c1_bardenet14}. We refer to these algorithms,
-respectively, as {\sc AustereMH} and {\sc MHSubLhd}.
-
-In~\citep{cutting_mh_2014}, CLT asymptotic arguments are used to show that their
-approximate acceptance test error (with respect to full-batch M-H) tends to zero
-as batch size increases, but no quantitative bounds are given. In addition, to
-design a test that minimizes data usage given an error bound, they rely on
-computing a standardized mean $\mu_{\rm std}$ each iteration. As the value of
-$\mu_{\rm std}$ depends on all $\log p(x_i|\theta)$ and $\log p(x_i| \theta')$
-terms, the test is technically not a minibatch test; \citep{cutting_mh_2014}
-propose a conservative variant which assumes $\mu_{\rm std}=0$ and avoids the
-need to evaluate all likelihood terms. We refer to this as {\sc AustereMH(c)}
-and the default, non-conservative variant as {\sc AustereMH(nc)}.
-
-For {\sc MHSubLhd}, the test requires direct bounds on the log likelihood ratio,
-which for general distributions requires knowing $p(x_i | \theta)$ and $p(x_i |
-\theta')$ for all $N$ samples. In~\citep{icml2014c1_bardenet14}, explicit bounds
-are given and depend on
-\begin{equation}\label{eq:bad_bound}
-    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
-\end{equation}
-It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
-regression and other simple models, but it is difficult to do so for more
-complicated models, so one generally needs to use all $p(x_i | \theta')$
-terms\footnote{The sample code provided by~\citep{icml2014c1_bardenet14}
-computes $C_{\theta,\theta'}$ by traversing the entire data, thus providing no
-performance advantage over the complete test.} to get $C_{\theta,\theta'}$, like
-the {\sc AustereMH(nc)} test.
-
-In contrast, as we show in Section~\ref{sec:analysis}, we use quantitative forms
-of the CLT which rely on measurable statistics from a \emph{single} minibatch.
-In addition, while both {\sc AustereMH(nc)} and {\sc MHSubLhd} show useful
-reductions in the number of samples required, they suffer the drawback of
-resolving small log likelihood ratio differences between the minibatch and full
-batch versions. We discuss a worst-case scenario in
-Section~\ref{ssec:gaussian_example}.
+respectively, as {\sc AustereMH} and {\sc MHSubLhd}. While both show useful
+reductions in the number of samples required, they suffer from two drawbacks:
+having to compute all log likelihood tokens each iteration to compute certain
+statistics,\footnote{Thus, these are not technically minibatch methods.} (see
+Section~\ref{sec:analysis}), and resolving small log likelihood ratio
+differences between the minibatch and full batch versions. We discuss a
+worst-case scenario in Section~\ref{ssec:gaussian_example}.
 
@@
     (independent of $N$) constant $c$ such that with probability $\geq c$ over
-    the joint distribution of $(\theta, \theta', u)$, {\sc AustereMH(nc)} and
-    {\sc MHSubLhd} consume all $N$ samples. 
+    the joint distribution of $(\theta, \theta', u)$, {\sc AustereMH} and {\sc
+    MHSubLhd} consume all $N$ samples. 
 \end{lemma}
@@
 We now derive error bounds for our M-H test and the target distribution it
-generates. In Section~\ref{ssec:delta_star_distribution}, we present bounds on
-the absolute and relative error (in terms of the CDFs) of the distribution of
-$\Delta^*$ vs.  a Gaussian. We then show in Section~\ref{ssec:preserve_bounds}
-that these bounds are preserved after the addition of other random variables
-(e.g., $X_{\rm nc}$ and $X_{\rm corr}$). It then follows that the acceptance
-test has the same error bound.
+generates. In the most similar prior works,~\citep{cutting_mh_2014}, CLT
+asymptotic arguments are used to show that their approximate acceptance test
+error (with respect to full-batch M-H) tends to zero as batch size increases,
+but no quantitative bounds are given. In addition, to design a test that
+minimizes data usage given an error bound, they rely on computing a standardized
+mean $\mu_{\rm std}$ each iteration. As the value of $\mu_{\rm std}$ depends on
+all $\log p(x_i|\theta)$ and $\log p(x_i| \theta')$ terms, the test is
+technically not a minibatch test; \citep{cutting_mh_2014} propose a conservative
+variant which assumes $\mu_{\rm std}=0$ and avoids the need to evaluate all
+likelihood terms. We refer to this as {\sc AustereMH(c)} and the default,
+non-conservative variant as {\sc AustereMH(nc)}.
+
+For {\sc MHSubLhd}, the test requires direct bounds on the log likelihood ratio,
+which for general distributions requires knowing $p(x_i | \theta)$ and $p(x_i |
+\theta')$ for all $N$ samples. In~\citep{icml2014c1_bardenet14}, explicit bounds
+are given and depend on
+\begin{equation}\label{eq:bad_bound}
+    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
+\end{equation}
+It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
+regression and other simple models, but it is difficult to do so for more
+complicated models, so one generally needs to use all $p(x_i | \theta')$
+terms\footnote{The sample code provided by~\citep{icml2014c1_bardenet14}
+computes $C_{\theta,\theta'}$ by traversing the entire data, thus providing no
+performance advantage over the complete test.} to get $C_{\theta,\theta'}$, like
+the {\sc AustereMH(nc)} test.  In contrast, we use quantitative forms of the CLT
+which rely on measurable statistics from a \emph{single} minibatch.
+
+In Section~\ref{ssec:delta_star_distribution}, we present bounds on the absolute
+and relative error (in terms of the CDFs) of the distribution of $\Delta^*$
+versus a Gaussian. We then show in Section~\ref{ssec:preserve_bounds} that these
+bounds are preserved after the addition of other random variables (e.g., $X_{\rm
+nc}$ and $X_{\rm corr}$). It then follows that the acceptance test has the same
+error bound.
 
@@
 \bibliography{paper}
-%\bibliographystyle{plain}
 \bibliographystyle{plainnat}
@@
 \end{itemize}
+and chose the $(m,\epsilon)$ pairing which resulted in the lowest expected data
+usage given a selected upper bound on the error. Through personal communication
+with~\citet{cutting_mh_2014}, we were able to use their same code to compute
+expected data usage and errors.
+
 The main difference between {\sc AustereMH(c)} and {\sc

### 2017-03-07 uai2017/paper.tex
added another reference, probably our last one, I hope

@@
 \textbf{We benchmark with the ADAGrad~\citep{adagrad} implementation of Sparse
-Matrix Factorization described in~\citep{Canny:2013:BDA:2487575.2487677}.
+Matrix Factorization described in~\citep{Canny:2013:BDA:2487575.2487677}, which
+is a variant of Alternating Least Squares~\citep{Koren:2009:MFT:1608565.1608614}.
 (Daniel: is this the correct citation for BIDMach? I assume this is what we will

### 2017-03-15 old_experiments/test.py
Adding some brief observations when looking at SMF code.

@@ -1,0 +1,34 @@
+"""
+Test SMF.scala on the first iteration of Netflix, with MHTest updater only. NO
+INTERNAL ADAGRAD UPDATER.
+
+File name convention:
+
+accept{false,some,true} for always rejecting, normal update rule, and always
+accepting, respectively, in MHTest. We SHOULD be using the `some` (i.e. normal
+update rule) in practice.
+
+{no,yes}updates, for commenting out, or not, respectively, the two lines in the
+mupdate method which update the two bias model matrices.
+
+log_acceptfalse_noupdates.txt  
+log_acceptfalse_yesupdates.txt
+log_acceptsome_noupdates.txt   
+log_acceptsome_yesupdates.txt
+log_accepttrue_noupdates.txt   
+log_accepttrue_yesupdates.txt
+"""
+
+
+import numpy as np
+import matplotlib.pyplot as plt
+plt.style.use('seaborn-darkgrid')
+plt.figure(figsize=(11,8))
+plt.plot(np.loadtxt("log_acceptfalse_noupdates.txt"),  'k-',  lw=4, label="False/No")
+plt.plot(np.loadtxt("log_acceptfalse_yesupdates.txt"), 'k--', lw=4, label="False/Yes")
+plt.plot(np.loadtxt("log_acceptsome_noupdates.txt"),   'y-',  lw=4, label="Some/No")
+plt.plot(np.loadtxt("log_acceptsome_yesupdates.txt"),  'y--', lw=4, label="Some/Yes")
+plt.plot(np.loadtxt("log_accepttrue_noupdates.txt"),   'b-',  lw=4, label="True/No")
+plt.plot(np.loadtxt("log_accepttrue_yesupdates.txt"),  'b--', lw=4, label="True/Yes")
+plt.legend(loc=(0.0,0.2))
+plt.savefig("rmse_smf.png")

### 2017-03-16 uai2017/paper.tex
Slight modifications to logistic regression.
I had mis-run the earlier trial by not repeating samples if they were rejected. That's all taken care of now and the paper has been updated along with figures.

@@
     {\sc MHminibatch} on both datasets when the temperature starts at 100 and
-    drops to 1 after a ``burn-in'' period of 200 samples (vertical dashed line)
+    drops to 1 after a ``burn-in'' period of 400 samples (vertical dashed line)
     of $\theta$.  For all three plots, one standard deviation is indicated by
@@
 the advantage of having higher temperature from the third plot in
-Figure~\ref{fig:logistic_performance}. During the exploration period, the
-accuracy rapidly increases, and then after 200 samples, we switch the
-temperature to 1, but this requires the step size to decrease, hence the smaller
-changes in accuracy.
+Figure~\ref{fig:logistic_performance}, which plots average performance and one
+standard deviation for {\sc MHminibatch} over 10 trials. During the exploration
+period, the accuracy rapidly increases, and then after 400 samples, we switch
+the temperature to 1, but this requires the step size to decrease, hence the
+smaller changes in accuracy.
 
 Figure~\ref{fig:logistic_histograms_mnist8m} shows log-log histograms of
-minibatch sizes for the three methods on MNIST-100k.
+minibatch sizes for the methods on MNIST-100k.
 (Figure~\ref{fig:logistic_histograms_mnist} in Appendix~\ref{app:logistic}
@@
 average minibatch sizes ($\pm$ one standard deviation) across all trials. {\sc
-MHminibatch}, with average minibatch sizes of roughly 129.8 and 202.5 for
-MNIST-13k and MNIST-100k, respectively, consumes more than 5x fewer data points
-than the next-best method, {\sc AustereMH(nc)}.  We reiterate, however, that
-both {\sc AustereMH(nc)} and {\sc MHSubLhd} require computing $\log
-p(x_i|\theta)$ and $\log p(x_i|\theta')$ for all $x_i$ each iteration. Our
-results here do not count that extra data consumption. Only our method and {\sc
-AustereMH(c)} rely solely on the minibatch of data each iteration.
+MHminibatch}, with average minibatch sizes of 133.7 and 242.4 for MNIST-13k and
+MNIST-100k, respectively, consumes more than 7x and 4x fewer data points than
+the next-best method, {\sc AustereMH(nc)}.  We reiterate, however, that both
+{\sc AustereMH(nc)} and {\sc MHSubLhd} require computing $\log p(x_i|\theta)$
+and $\log p(x_i|\theta')$ for all $x_i$ each iteration. Our results here do not
+count that extra data consumption. Only our method and {\sc AustereMH(c)} rely
+solely on the minibatch of data each iteration.
 
@@
 \hline \\
-{\sc MHminibatch}   & $129.8\pm 3.4$    & $202.5 \pm 13.6$ \\
+{\sc MHminibatch}   & $133.7\pm 4.7$    & $242.4 \pm 12.4$ \\
 {\sc AustereMH(nc)} & $973.8\pm 49.8$   & $1098.3 \pm 44.9$ \\

### 2017-03-20 uai2017/paper.tex
update uai2017 paper: gaussian mixture experiment results

@@
     \centering
-    \includegraphics[width=0.9\linewidth]{posterior_of_gaussian.png}
+    %\includegraphics[width=0.9\linewidth]{posterior_of_gaussian.png}
+    \includegraphics[width=1.0\linewidth]{gaussian_mixture_scatter_results_v8.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{minibatch_size_gaussian.png}
+    %\includegraphics[width=0.9\linewidth]{minibatch_size_gaussian.png}
+    \includegraphics[width=0.9\linewidth]{gaussian_mixture_histogram_results_v8.png}
     \caption{
@@
 We benchmark with {\sc AustereMH(c)} and {\sc MHSubLhd}. We initialized {\sc
-MHminibatch} and {\sc MHSubLhd} with $m=100$. For {\sc MHSubLhd}, we increase
+MHminibatch} and {\sc MHSubLhd} with $m=50$. For {\sc AustereMH(c)}, we set
+the error bound $\epsilon$ to 0.005.  For {\sc MHSubLhd}, we increase
 sizes geometrically with $\gamma = 1.5$ and use parameters $p = 2, \delta =
-0.01$.  All methods collect 5000 samples using a random walk proposer with
+0.01$.  All methods collect 3000 samples using a random walk proposer with
 covariance matrix ${\rm diag}(0.15, 0.15)$, which means the M-H test is
@@
 iteration show that our method consumes significantly less data; the
-distribution is short-tailed and the mean is 210, more than an order of
-magnitude better compared to the other two methods (averages are 15562 and
-16857). Sizes correspond to the running times of the methods, excluding the
+distribution is short-tailed and the mean is 181, more than an order of
+magnitude better compared to the other two methods (averages are 13511 and
+73011). Sizes correspond to the running times of the methods, excluding the
 likelihood computation of all data points for {\sc MHSubLhd}, which would

### 2017-03-22 uai2017/paper.tex
Updated figures after fixing logistic regression bug with log u vs log (1)

@@
 average minibatch sizes ($\pm$ one standard deviation) across all trials. {\sc
-MHminibatch}, with average minibatch sizes of 133.7 and 242.4 for MNIST-13k and
+MHminibatch}, with average minibatch sizes of 125.4 and 216.5 for MNIST-13k and
 MNIST-100k, respectively, consumes more than 7x and 4x fewer data points than
@@
 \hline \\
-{\sc MHminibatch}   & $133.7\pm 4.7$    & $242.4 \pm 12.4$ \\
+{\sc MHminibatch}   & $125.4\pm 9.2$    & $216.5 \pm 7.9$ \\
 {\sc AustereMH(nc)} & $973.8\pm 49.8$   & $1098.3 \pm 44.9$ \\
@@
 	\caption{
-    \textbf{Daniel: TODO describe}
+    Effect of changing the proposal step size $\sigma$ for
+    {\sc MHminibatch}.
     }

### 2017-03-23 uai2017/paper.tex
update figures and results

--- 2017-03-20 uai2017/paper.tex
@@
 iteration show that our method consumes significantly less data; the
-distribution is short-tailed and the mean is 181, more than an order of
-magnitude better compared to the other two methods (averages are 13511 and
-73011). Sizes correspond to the running times of the methods, excluding the
+distribution is short-tailed and the mean is 213, more than an order of
+magnitude better compared to the other two methods (averages are 11962 and
+63408). Sizes correspond to the running times of the methods, excluding the
 likelihood computation of all data points for {\sc MHSubLhd}, which would

### 2017-03-23 uai2017/paper.tex
update paper and results

@@
 iteration show that our method consumes significantly less data; the
-distribution is short-tailed and the mean is 213, more than an order of
-magnitude better compared to the other two methods (averages are 11962 and
-63408). Sizes correspond to the running times of the methods, excluding the
+distribution is short-tailed and the mean is 172, more than an order of
+magnitude better compared to the other two methods (averages are 12562 and
+675008). We further ran 10 runs of mixture of Gaussians experiments, and
+the average and standard deviation values of the mean minibatch size 
+for {\sc MHminibatch}, {\sc AustereMH(c)}, {\sc MHSubLhd} are listed in
+table ~\ref{tab:gaussianres}. Sizes correspond to the running times of the methods, excluding the
 likelihood computation of all data points for {\sc MHSubLhd}, which would
 drastically increase running time.
+
+\begin{table}[t]
+\caption{Statistics for mean minibatch sizes on gaussian mixture model
+experiments using {\sc MHminibatch}, {\sc AustereMH(c)}, {\sc MHSubLhd}.
+The averages are taken over 10 independent trials (3000 samples each).}
+\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
+\label{tab:gaussianres}
+\vskip 0.15in
+\begin{center}
+\begin{tabular}{l l l l}
+\textbf{Method/Data} & \textbf{Average of Mean} & \textbf{Std of Mean}  \\
+\hline \\
+{\sc MHminibatch}   & 182 & 12 \\
+{\sc AustereMH(c)}  & 13540 & 1521 \\
+{\sc MHSubLhd}      & 65759 & 3222 \\
+\end{tabular}
+\end{center}
+\vskip -0.1in
+\end{table}
 

### 2017-03-24 uai2017/paper.tex
Made minor changes, added reminders, now send to John

--- 2017-03-23 uai2017/paper.tex
@@
 iteration show that our method consumes significantly less data; the
-distribution is short-tailed and the mean is 213, more than an order of
-magnitude better compared to the other two methods (averages are 11962 and
-63408). Sizes correspond to the running times of the methods, excluding the
-likelihood computation of all data points for {\sc MHSubLhd}, which would
-drastically increase running time.
+distribution is short-tailed and the mean is 172, more than an order of
+magnitude better compared to the other two methods (averages are 12562 and
+675008). \textbf{Daniel: Xinlei, these numbers don't match
+Figure~\ref{fig:gauss_mix_2} and the figures need to be updated so the scatter
+plot has the correct method names and the scaling key, while the histograms have
+the correct names and the seaborn background.} We
+further ran 10 runs of mixture of Gaussians experiments, and the average and
+standard deviation values of the mean minibatch size for {\sc MHminibatch}, {\sc
+AustereMH(c)}, {\sc MHSubLhd} are listed in table ~\ref{tab:gaussianres}. Sizes
+correspond to the running times of the methods, excluding the likelihood
+computation of all data points for {\sc MHSubLhd}, which would drastically
+increase running time.
+
+\begin{table}[t]
+\caption{Average minibatch sizes ($\pm$ one standard deviation) on the Gaussian
+mixture model. The averages are taken over 10 independent trials (3000 samples
+each). \textbf{Daniel: Xinlei, can you add in one more decimal point for all the
+numbers in this table, like I do in the Logistic Regression table?}}
+% Daniel: this `small` command was carried over from the LR table which couldn't
+% fit in a column. It's not needed here.
+%\small 
+\label{tab:gaussianres}
+\vskip 0.15in
+\begin{center}
+\begin{tabular}{l l l l}
+\textbf{Method} & \textbf{Average of MB Sizes} \\
+\hline \\
+{\sc MHminibatch}   & $182\pm 12$ \\
+{\sc AustereMH(c)}  & $13540\pm 1521$ \\
+{\sc MHSubLhd}      & $65759\pm 3222$ \\
+\end{tabular}
+\end{center}
+\vskip -0.1in
+\end{table}
 
@@
 solely on the minibatch of data each iteration.
-
-\subsection{SPARSE MATRIX FACTORIZATION}\label{ssec:collab_filtering}
 
@@
 
-\begin{table}[t]
-\caption{\textbf{TODO TABLE WITH SMF RESULTS}}
-\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
-\label{tab:smf}
-\vskip 0.15in
-\begin{center}
-\begin{tabular}{l l l l}
-\textbf{RMSE} & \textbf{MovieLens 10m} & \textbf{Netflix}  \\
-\hline \\
-{\sc MHminibatch} & ??? & ??? & \\
-ADAGrad           & ??? & ??? & \\
-\end{tabular}
-\end{center}
-\vskip -0.1in
-\end{table}
-
-\textbf{Daniel: TODO this is our third experiment. Results are in progress...}
-
-\textbf{We benchmark with the ADAGrad~\citep{adagrad} implementation of Sparse
-Matrix Factorization described in~\citep{Canny:2013:BDA:2487575.2487677}, which
-is a variant of Alternating Least Squares~\citep{Koren:2009:MFT:1608565.1608614}.
-(Daniel: is this the correct citation for BIDMach? I assume this is what we will
-use)}
-
-\section{CONCLUSIONS}\label{sec:conclusion}
+\section{CONCLUSIONS AND DISCUSSIONS}\label{sec:conclusion}
 
@@
 tests. We present theoretical results and experimentally show the benefits of
-our test on Gaussian mixtures, logistic regression, and sparse matrix
-factorization.  Directions for future work include integrating our algorithm
-with~\citep{cutting_mh_2014} by applying both tests each iteration or utilizing
-the variance reduction techniques suggested in~\citep{DBLP:conf/icml/ChenG16}.
-More elaborate extensions involve running more experiments on other models such
-as neural networks and providing a recipe for how to use our algorithm
-(following the framework of~\citep{sgmcmc_2015}).
-
+our test on Gaussian mixtures and a variety of logistic regression experiments.
+
+The most pressing direction for future work is to integrate our M-H test with
+Hamiltonian Monte Carlo and Langevin Dynamics methods to enable MCMC methods to
+scale to models such as deep neural networks. With these types of models, our
+proposal probabilities will almost certainly have to be asymmetric, but these
+typically result in an estimate plus Gaussian noise which we can absorb into our
+test.
+
+Other possibilities for future work include integrating our algorithm
+with~\citep{cutting_mh_2014} by applying both tests each iteration, utilizing
+the variance reduction techniques suggested in~\citep{DBLP:conf/icml/ChenG16},
+and providing recipe for how to use our algorithm following the framework
+of~\citep{sgmcmc_2015}.
+
+\newpage
 \bibliography{paper}
@@
 distribution derivation and on our three main experiments to assist
-understanding and reproducibility. Our code is open-source on
-GitHub.\footnote{\url{https://github.com/BIDData/BIDMach/blob/master/src/main/scala/BIDMach/updaters/MHTest.scala}}
-% Daniel: TODO remove for anonymous review later.
+understanding and reproducibility. 
+% Daniel: TODO can decide whether to include this here or not. It might actually
+% be better to do this in the logistic regression section.
+% Our code is open-source on
+% GitHub.\footnote{\url{https://github.com/BIDData/BIDMach/blob/master/src/main/scala/BIDMach/updaters/MHTest.scala}}
 
@@
 \begin{table}[t]
-\caption{Gaussian Mixture Model Statistics}
+\caption{Gaussian Mixture Model Statistics \textbf{TODO Xinlei, please fix this
+table.}}
 \small
@@
 ordering changes depending on the range of samples generated. \textbf{Daniel:
-one reviewer had a problem with this, and I agree. The easiest fix here is to
-simply report standard deviation values. Also, we should emphasize that we are
-not aiming to get the best score by this metric but simply something that
+Xinlei, one reviewer had a problem with this, and I agree. The easiest fix here
+is to simply report standard deviation values. Also, we should emphasize that we
+are not aiming to get the best score by this metric but simply something that
 exhibits some form of ``correctness.''}
@@
 
-\subsection{SPARSE MATRIX FACTORIZATION EXPERIMENT (SECTION~\ref{ssec:collab_filtering})}\label{app:collab_filtering}
-
-\textbf{Daniel: TODO}
-
 \end{document}

### 2017-03-26 uai2017/paper.tex
update new results
update paper

@@
 \hline \\
-{\sc MHminibatch}   & $182\pm 12$ \\
-{\sc AustereMH(c)}  & $13540\pm 1521$ \\
-{\sc MHSubLhd}      & $65759\pm 3222$ \\
+{\sc MHminibatch}   & $182.3\pm 11.4$ \\
+{\sc AustereMH(c)}  & $13540.5\pm 1521.4$ \\
+{\sc MHSubLhd}      & $65758.9\pm 3222.6$ \\
 \end{tabular}
@@
 \hline
-\T Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
-Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
+\T Equation~\ref{eq:log_prob_poissons} & $-1306.2 \pm 20.7$ & $-1619.6 \pm 25.3$ & $-1301.5 \pm 44.4$ \\
+Chi-Squared & $5706.8 \pm 2271.5$ & $ 10501.3 \pm 2076.5$ & $ 3711.7 \pm 1182.4$ \\
 \end{tabular}
@@
 perform significance tests using Chi-Squared distribution (also in
-Table~\ref{tab:poissons}). Scores lie between~\citep{cutting_mh_2014}
-and~\citep{icml2014c1_bardenet14}, but the variance of these values is high and
-ordering changes depending on the range of samples generated. \textbf{Daniel:
+Table~\ref{tab:poissons}). The table provides the mean likelihood value and mean 
+Chi-Squared test statistics value as well as their standard deviations.  Our likelihood values 
+lies between~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14}, 
+but the variance of their likelihood values are high and
+ordering changes depending on the range of samples generated. It should be noted 
+that we are not aiming at optimizing the likelihood values nor the Chi-Squared statistics.
+We use these values to show the extent of correctness. \textbf{Daniel:
 Xinlei, one reviewer had a problem with this, and I agree. The easiest fix here

### 2017-03-26 uai2017/paper.tex
update paper

@@
 magnitude better compared to the other two methods (averages are 12562 and
-675008). \textbf{Daniel: Xinlei, these numbers don't match
+67508). \textbf{Daniel: Xinlei, these numbers don't match
 Figure~\ref{fig:gauss_mix_2} and the figures need to be updated so the scatter

### 2017-03-29 old_paper_versions/NIPS_Reviews_Rebuttals/rebuttal_reviewers.txt

We thank the reviewers for their thoughtful commentary on our paper.
We can only agree with the comments and appreciate the reviewers
patience in reviewing the paper. It was an early draft. We noticed
most of these issues ourselves in the time since we submitted the paper
and have been working to address them. We discuss those corrections
in detail below.

General:

First we established worst-case lower bounds for references [3] and [4]
of Omega(N) bounds on the number of samples consumed per iteration,
using simple Gaussian examples.

In addition, Section 4 (the theoretical results) has been
substantially revised.  Rather than general appeal to CLT, we include
bounds based on t-statistics (using both the empirical mean and
variance) asymptotics from [A] (below) with explicit constants and an
O(n^-0.5) error bound. We revised the analysis of the test error rate
which now requires no additional assumptions. We include error bounds
on the equilibrium distribution given bounds on transition kernel
error which appeared in [3] and [4], citing those works, under the
appropriate assumptions.

To Reviewers 2, 3, and 4 regarding the Barker function:

You are correct, the logistic (Barker) function is less efficient than
the Metropolis acceptance test with \theta' close to \theta when run
on the full dataset. However, there is no comparable result when these
tests are used on minibatches since they are applied in completely
different ways. Empirically, we find that we can scale proposal step
size or temperature to achieve as small minibatch size as desired
using our corrected Barker test, but this is not true of [3] or [4]
which via the analysis above (and in experiments) still require
Omega(n) samples for some single tests. Thus the approx.  Barker test
is arbitrarily more efficient than tail tests in terms of average data
consumed per generated distinct sample. We have revised the
experiments so show that such gaps are up to 3 orders of magnitude for
logistic regression and the gaussian mixture model.

To Reviewers 1, 2, and 3 regarding the deconvolution (in)stability:

We now describe in detail the construction of the distribution of
X_corr. We explored linear programming and least squares solutions. We
find that the latter allow us to compute finer discrete approximations
to the distribution of X_corr (4000 points). It also gives smaller
absolute (i.e. L1) error. Even though least squares minimizes the L2
norm of the error, the L1 error is also bounded down to 1e-5 by L2
norm minimization. The solution accuracy benefits from fairly strong
L2 regularlization, which also makes the inversion
(i.e. deconvolution) more stable. To ensure positivity of all three
distributions, we impose a stricter bound on the variance of X_norm.
This effectively shifts variance to $X_corr$ by convolving it with
another normal distribution, and guarantees that it is positive.
We computed X_corr with variance bounds of 0.8 and 0.9 both of
which give CDF errors below 1e-4.

To Reviewer 3:

(TODO respond to experiment criticism, but I really can't defend the work as it is ...)

Both Reviewers 1 and 3 correctly identified our implicit assumptions
about Lemma 2. Reviewer_1 is right that we approximated it for large
N, so sampling without replacement approximates sampling with
replacement.  In the case of very large datasets however, it is also
typically more reasonable to treat the dataset as itself a sample
drawn from a still larger, or possibly infinite, dataset (e.g. this
holds for all the examples in our paper), and the instance
"resolution" implies that repeats have probability ~ zero. So
distinctions between sampling with replacement and without are perhaps
moot for these datasets.

To Reviewers 1 and 2:

Thank you for mentioning [BDH], "MCMC Methods for Tall Data." We added
this reference as part of our ongoing revision of the paper and will
substantially address issues raised in [BDH]. In addition, we are
addressing Reviewer 2's concern regarding the difference between ours
and [BDH] Section 6.3. The idea of using a Barker test with minbatches
was proposed in [BDH]. However, a crucial addition in our approach is
the use of an additive correction variable. Thus in our approach,
there is no lower bound on error in our approximate test, and we show
it decreases with O(n^-0.5). In [BDH] there is no correction between
normal and logistic CDFs, and the method has a minimum error set by
the max absolute difference between their CDFs. Another difference is
that the idea proposed in [BDH] was not explored experimentally. In
particular it was not shown that tests can be done with arbitrarily
small minibatches in [BDH].

To Reviewer 7:

We have improved the algorithm to use an adaptive step size rather
than fail on some tests.  Unlike [3] and [4] however, since the amount
of data required for our test scales in proportion to the proposal
step size, the minibatch size has a normal size bound.

[A] Y. Novak, "On Self-Normalized Sums And Student’s Statistic"
Theory of Probability and its Applications, 49(2), 336-344, 2005.

### 2017-03-29 old_paper_versions/ArXiv_October_2016_version/arxiv.tex

--- 2016-10-13 aistats2017.tex
@@
-\documentclass[twoside]{article} \usepackage{aistats2017}
-
-% If your paper is accepted, change the options for the package
-% aistats2017 as follows:
-%
-%\usepackage[accepted]{aistats2017}
-%
-% This option will print headings for the title of your paper and
-% headings for the authors names, plus a copyright note at the end of
-% the first column of the first page.
-\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem,float}
-\usepackage{amssymb}
-\usepackage{natbib}  % for references
-\usepackage{amsfonts}
-\usepackage{nicefrac}  
-\usepackage{booktabs}
-\usepackage{microtype}
-\usepackage[]{algorithm2e}
+\documentclass{article}
+
+% if you need to pass options to natbib, use, e.g.: \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading nips_2016
+% to avoid loading the natbib package, add option nonatbib: \usepackage[nonatbib]{nips_2016}
+
+%\usepackage{nips_2016}
+% to compile a camera-ready version, add the [final] option, e.g.:
+\usepackage[final]{nips_2016}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{amssymb}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{graphicx}
 \usepackage{caption}
 \usepackage{subcaption}
+
+\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem}
+\usepackage[]{algorithm2e}
 \graphicspath{{figures/}}
-\usepackage{setspace}
+
 \lstset{
@@
 
-\renewcommand{\refname}{}
 \newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
@@
 
-% This portion of style definition will remove the index number before the 
-% without inducing an indent
-\makeatletter
-\renewcommand\@biblabel[1]{}
-\renewenvironment{thebibliography}[1]
-     {\section*{\refname}%
-      \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
-      \list{}%
-           {\leftmargin0pt
-            \@openbib@code
-            \usecounter{enumiv}}%
-      \sloppy
-      \clubpenalty4000
-      \@clubpenalty \clubpenalty
-      \widowpenalty4000%
-      \sfcode`\.\@m}
-     {\def\@noitemerr
-       {\@latex@warning{Empty `thebibliography' environment}}%
-      \endlist}
-\makeatother
+\title{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
+
+\author{
+  Haoyu Chen\textsuperscript{*}, Daniel Seita\textsuperscript{*},
+  Xinlei Pan\textsuperscript{*}, John Canny\textsuperscript{*}$^{\dag}$ \\
+  * University of California, Berkeley, CA \\
+  $\dag$ Google Research, Mountain View, CA\\
+  \texttt{\{haoyuchen,seita,xinleipan,canny\}@berkeley.edu}
+ }
 
 \begin{document}
-
-% If your paper is accepted and the title of your paper is very long,
-% the style will print as headings an error message. Use the following
-% command to supply a shorter title of your paper so that it can be
-% used as headings.
-%
-%\runningtitle{I use this title instead because the last one was very long}
-
-% If your paper is accepted and the number of authors is large, the
-% style will print as headings an error message. Use the following
-% command to supply a shorter version of the authors names so that
-% they can be used as headings (for example, use only the surnames)
-%
-%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
-
-\twocolumn[
-
-\aistatstitle{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}
-
-\aistatsauthor{ Anonymous Author 1 \And Anonymous Author 2 \And Anonymous Author 3 }
-
-\aistatsaddress{ Unknown Institution 1 \And Unknown Institution 2 \And Unknown Institution 3 } ]
+\maketitle
 
 \begin{abstract}
-  %The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
-  %both left and right-hand margins. Use 10~point type, with a vertical
-  %spacing of 11~points. The {\bf Abstract} heading must be centered,
-  %bold, and in point size 12. Two line spaces precede the
-  %Abstract. The Abstract must be limited to one paragraph.
-  We
-  present a novel Metropolis-Hastings method for large datasets that
+  We present a novel Metropolis-Hastings method for large datasets that
   uses small expected-size minibatches of data. Previous work on
@@
 
-In the Metropolis-Hastings method~\citep{gilks1996markov,brooks2011handbook}, a
+In the Metropolis-Hastings method~\cite{gilks1996markov,brooks2011handbook}, a
 difficult-to-compute probability distribution $p(\theta)$ is sampled using a
@@
 the test repeats. This process continues until a decision. The bounds depend on
-either an asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
-concentration bound~\citep{icml2014c1_bardenet14}. The latter requires direct
+either an asymptotic Central Limit Theorem~\cite{cutting_mh_2014} or a
+concentration bound~\cite{icml2014c1_bardenet14}. The latter requires direct
 bounds on the log likelihood ratio, which for general distributions requires
@@
 
-Following~\citet{icml2014c1_bardenet14}, we write the test
+Following~\cite{icml2014c1_bardenet14}, we write the test
 $u<\alpha(\theta,\theta')$ equivalently as $\Lambda(\theta,\theta') >
 \psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those
-in~\citet{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis
+in~\cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis
 later.}
 \begin{equation}
-\begin{split} 
-\Lambda(\theta,\theta') & = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right) \\  
-{\rm and} \quad \psi(u,\theta,\theta') &= \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right).
-\end{split}
+\Lambda(\theta,\theta') = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)
+\quad {\rm and} \quad 
+\psi(u,\theta,\theta') = \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right).
 \end{equation}
@@
 distribution has the same variance as the target
-distribution~\citep{OptimalScaling01}, so we use a proposal based on
+distribution~\cite{OptimalScaling01}, so we use a proposal based on
 $\mathcal{N}(\theta,1/N)$. It is symmetric
@@
 \begin{equation}\label{eq:delta2}
-\begin{split}
-    \Delta(\theta,\theta') &= \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') \\
-    & = \Lambda(\theta,\theta') - \psi(1,\theta,\theta').
-\end{split}
+\Delta(\theta,\theta') = \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') = \Lambda(\theta,\theta') - \psi(1,\theta,\theta').
 \end{equation}
@@
 \begin{equation}\label{eq:delta3}
-\begin{split}
-    \Delta^*(\theta,\theta') &=
-\frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta')\\
-&=\Lambda^*(\theta,\theta') - \psi(1,\theta,\theta')
-\end{split}
+\Delta^*(\theta,\theta') =
+\frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} -
+\psi(1,\theta,\theta')=\Lambda^*(\theta,\theta') - \psi(1,\theta,\theta').
 \end{equation}
@@
 
-This result is used in~\citet{Barker65} to define the Barker acceptance test.
+This result is used in~\cite{Barker65} to define the Barker acceptance test.
 %As
@@
 \subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
-For our new MH test we use the Barker logistic~\citep{Barker65}
+For our new MH test we use the Barker logistic~\cite{Barker65}
 function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
@@
 
-% Daniel: This was taking up too much space, and we usually don't want to spend
-% a paragraph explaining what *didn't* work, and readers will know what it means
-% to minimize the L-infinity norm, we can put it inline.
-% \begin{equation}\label{eq:optimization_l1}
-%     u^* = \argmin_{u} \|Mu-v\|_{\infty}.
-% \end{equation}
-% We have the additional constraint that $u_j > 0$ for all $j$, since $u$
-% represents a density. We first explored optimizing this problem with linear
-% programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
-% 
-% This was tractable up to a few hundred dimensions for $u$.  However, the
-% discretization error is bounded by the curvature of the underlying functions
-% times the squared quantization step. The curvatures are here slightly less than
-% one, i.e.  the errors are of the order of $(V/N)^2$.  Here we chose $V=20$ to
-% provide adequate containment of the distributions (the CDFs are extremely close
-% to either zero or one outside this range). So with 200 points, we have a
-% discretization error of approximately 0.01, which is relatively large.  To yield
-% higher resolution and lower error, we switched to a least squares solution:
-
 \SetKwInOut{Input}{Input}
@@
     \centering
-    \begin{tabular}{l l l l}
+    \begin{tabular}{l l l l} \\
     ${\bf N}$ & ${\bf \sigma}$ & ${\bf \lambda}$ & ${\bf L_{\infty}}$ {\bf error} \\
@@
 We now derive error bounds for our M-H test and the target distribution it
-generates. In the most similar prior works,~\citet{cutting_mh_2014} use CLT
+generates. In the most similar prior works,~\cite{cutting_mh_2014} use CLT
 asymptotic arguments to show that its approximate acceptance test error tends to
 zero as batch size increases, but no quantitative bounds are given.
-In~\citet{icml2014c1_bardenet14}, explicit bounds are given and depend on
+In~\cite{icml2014c1_bardenet14}, explicit bounds are given and depend on
 \begin{equation}\label{eq:bad_bound}
@@
 generally needs to use all $p(x_i\mid \theta')$ terms\footnote{The sample code
-provided by~\citet{icml2014c1_bardenet14} computes $C_{\theta,\theta'}$
+provided by~\cite{icml2014c1_bardenet14} computes $C_{\theta,\theta'}$
 by traversing the entire data, thus providing no performance advantage over the
@@
 \begin{proof}
-See Supplementary Material, Section~\ref{app:quant_clt}.
+See Appendix, Section~\ref{app:quant_clt}.
 \end{proof}
-
-%% {\color{blue} Daniel: the notation here was (and still is, to a lesser extent)
-%% the source of considerable confusion for me. Do we want $\sigma(X)$ or
-%% $\sigma(\hat{X})$? I think in Lemma~\ref{lem:quant_clt} we want the former, not
-%% the latter. OH ALSO, this section uses $n$ as the full data size, which means
-%% before this section, we should use $n$, not $N$, for the full data size, right?
-%% The Bardenet paper uses $n$ to represent the full data size.
-
-%% JFC - this bound applies to minibatches, not the full dataset so the size $n$ is
-%% deliberately not $N$.
 
@@
 \begin{corollary}\label{cor:our_bound_delta_prime}
-\begin{align*}
-    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| &\leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} \\
-    &= \epsilon(\theta,\theta',b).
-\end{align*}
+\begin{equation}
+    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}}
+    = \epsilon(\theta,\theta',b).
+\end{equation}
 \end{corollary} %\vspace{-10pt}
@@
 
+\subsection{Mixture of Gaussians}\label{ssec:gaussians}
+
 % Daniel: moving this here so figures don't overflow to next pages. We can tune
@@
 
-\subsection{Mixture of Gaussians}\label{ssec:gaussians}
-
-% Daniel: moving this here so figures don't overflow to next pages. We can tune
-% the positioning if needed.
-\begin{figure*}[t]
-	\centering
-	%\includegraphics[width=0.9\linewidth]{logistic_performance.png}
-	\includegraphics[width=0.8\linewidth]{logistic_regression_ll_acc_results.png}
-	\caption{
-    Logistic regression performance (accuracy/log likelihood) based on
-    cumulative data usage.
-    }
-	\label{fig:logistic_performance}
-\end{figure*}
-\begin{figure*}[t]
-	\centering
-	%\includegraphics[width=0.9\linewidth]{minibatch_size_logistic.png}
-	\includegraphics[width=0.8\linewidth]{logistic_regression_histograms_results.png}
-	\caption{
-    Counts of minibatch sizes in the logistic regression experiment (analogous
-    to Figure~\ref{fig:gauss_mix_2}).
-    }
-	\label{fig:logistic_minibatch}
-\end{figure*}
-
-This model is adapted from~\citet{langevin_2011} by increasing the number
+This model is adapted from~\cite{langevin_2011} by increasing the number
 of samples to 1 million.  The parameters are $\theta = \langle \theta_1,\theta_2 \rangle$, and the
@@
 We ran MCMC with with our test against the test
-from~\citet{icml2014c1_bardenet14} and a simplified version
-of~\citet{cutting_mh_2014}. Both methods use repeated testing which
+from~\cite{icml2014c1_bardenet14} and a simplified version
+of~\cite{cutting_mh_2014}. Both methods use repeated testing which
 must be discounted to bound overall test
-accuracy. ~\citet{icml2014c1_bardenet14} conservatively partitions the
-test error, while~\citet{cutting_mh_2014} used a complex dynamic
+accuracy. ~\cite{icml2014c1_bardenet14} conservatively partitions the
+test error, while~\cite{cutting_mh_2014} used a complex dynamic
 programming heuristic to estimate the overall test error. We did not
 implement this heuristic and instead set the individual test error
-equal to the overall test error. Thus our results for~\citet{cutting_mh_2014} should be treated as lower bounds. All
-methods were initialized with minibatch size 100.  For~\citet{cutting_mh_2014} we
-increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
-use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
-samples using a random walk proposer with covariance ${\rm diag}(0.15,
-0.15)$, which means the M-H test is responsible for shaping the sample
-distribution.
+equal to the overall test error. Thus our results
+for~\cite{cutting_mh_2014} should be treated as lower bounds. All
+methods were initialized with minibatch size 100.
+For~\cite{cutting_mh_2014} we increment sizes by 100 and set the
+tolerance $\epsilon=0.005$ to control overall test error.
+For~\cite{icml2014c1_bardenet14}, we increase sizes geometrically
+with $\gamma = 1.5$ and use parameters $p = 2, \delta = 0.01$.  All
+methods collect 5000 samples using a random walk proposer with
+covariance ${\rm diag}(0.15, 0.15)$, which means the M-H test is
+responsible for shaping the sample distribution.
 
@@
 distribution (also in Table~\ref{tab:poissons}). Scores lie
-between~\citet{cutting_mh_2014} and~\citet{icml2014c1_bardenet14}, but
+between~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}, but
 the variance of these values is high and ordering changes depending on
@@
 
+% Daniel: moving this here so figures don't overflow to next pages. We can tune
+% the positioning if needed.
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=0.9\linewidth]{logistic_regression_ll_acc_results.png}
+	\caption{
+    Logistic regression performance (accuracy/log likelihood) based on
+    cumulative data usage.
+    }
+	\label{fig:logistic_performance}
+\end{figure*}
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=0.9\linewidth]{logistic_regression_histograms_results.png}
+	\caption{
+    Counts of minibatch sizes in the logistic regression experiment (analogous
+    to Figure~\ref{fig:gauss_mix_2}).
+    }
+	\label{fig:logistic_minibatch}
+\end{figure*}
+
 \begin{table}[t]
@@
     \centering
-    \resizebox{.45\textwidth}{!}{% <------ Don't forget this %
-    \begin{tabular}{l l l l}
+    \begin{tabular}{l l l l} \\
     {\bf Metric} & {\bf Ours} & {\bf Korat.'14} & {\bf Barde.'14} \\
@@
     Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
-    Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
-    \end{tabular}% <------ Don't forget this %
-   }
+    Chi-Squared & 3313.9 & 3647.7 & 2444.1
+    \end{tabular}
 \end{table}
@@
 versus 7s on a subset of the MNIST8M dataset, a larger version of
-MNIST~\citep{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
+MNIST~\cite{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
 450k training and 192k testing points.
-%%% The pixels are scaled in $[0,1]$  - JFC Why ????
 We impose a uniform prior on $\theta$ and again use a random walk
@@
 set at $T=1000$. We run the three tests for 3000 samples and again use 100
-as the starting minibatch size. Method~\citet{cutting_mh_2014} used
-per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
+as the starting minibatch size. Our undiscounted implementation of~\cite{cutting_mh_2014} used
+per-test error of $\epsilon=0.05$.  For~\cite{icml2014c1_bardenet14},
 we use their recommended discounting factor value of $\gamma=2.0$. We tried to use the
 symbolic bound for $C_{\theta,\theta'}$ for Logistic regression
-from~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead we use
+from~\cite{icml2014c1_bardenet14}, but found it to be too high. Instead we use
 the empirical $C_{\theta,\theta'}$ from the entire dataset and do not add its
@@
 three methods on a log-log scale. With an initial size of 100, we achieve an
-average minibatch size of 393, far smaller than the averages of the other two
-methods, which have longer-tailed distributions.
+average minibatch size of 393, more than two orders of magnitude smaller
+than~\cite{icml2014c1_bardenet14}, and almost an order of magnitude faster than the
+undiscounted implementation of~\cite{cutting_mh_2014}.
 
@@
 Monte Carlo methods, providing a recipe for how to use our algorithm (following
-the framework of~\citet{sgmcmc_2015}), or integrating parallel
-MCMC~\citep{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
-
-\iffalse
-\subsubsection{Figures}
-
-All artwork must be centered, neat, clean, and legible.  All lines
-should be very dark for purposes of reproduction, and art work should
-not be hand-drawn.  Figures may appear at the top of a column, at the
-top of a page spanning multiple columns, inline within a column, or
-with text wrapped around them, but the figure number and caption
-always appear immediately below the figure.  Leave 2 line spaces
-between the figure and the caption. The figure caption is initial caps
-and each figure should be numbered consecutively.
-
-Make sure that the figure caption does not get separated from the
-figure. Leave extra white space at the bottom of the page rather than
-splitting the figure and figure caption.
-\begin{figure}[h]
-\vspace{.3in}
-\centerline{\fbox{This figure intentionally left non-blank}}
-\vspace{.3in}
-\caption{Sample Figure Caption}
-\end{figure}
-
-\subsubsection{Tables}
-
-All tables must be centered, neat, clean, and legible. Do not use hand-drawn tables. Table number and title always appear above the table.
-See Table~\ref{sample-table}.
-
-Use one line space before the table title, one line space after the table title, and one line space after the table. The table title must be
-initial caps and each table numbered consecutively.
-
-\begin{table}[h]
-\caption{Sample Table Title} \label{sample-table}
-\begin{center}
-\begin{tabular}{ll}
-{\bf PART}  &{\bf DESCRIPTION} \\
-\hline \\
-Dendrite         &Input terminal \\
-Axon             &Output terminal \\
-Soma             &Cell body (contains cell nucleus) \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\section{SUPPLEMENTARY MATERIAL}
-
-If you need to include additional appendices during submission, you
-can include them in the supplementary material file.
-
-\fi
+the framework of~\cite{sgmcmc_2015}), or integrating parallel
+MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
 
 \newpage
-\subsubsection*{References}
-% Daniel: Sorry, I don't know why this is needed to get rid of the annoying
-% spaces below the "References" text.
-\vspace{-25pt} 
-%\bibliographystyle{apalike}
-\bibliographystyle{abbrvnat}
-\setcitestyle{authoryear,open={[},close={]}}
-\bibliography{aistats2017}
+\bibliographystyle{plain}
+\bibliography{arxiv}
 
@@
 \begin{center}
-{\Large Supplementary Material}
+{\Large Appendix}
 \end{center}
@@
 \section{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}
-
-\begin{comment}
-
-{\color{blue} Daniel: changes I made:
-
-\begin{enumerate}
-    \item (TODO later) After Equation~\ref{eq:accept_test_zero}, we say that
-    ``the resulting test with our chosen $u_0$ will never correctly succeed
-    [...]'', but I think we should say it will not succeed with high
-    probability, because we could select a minibatch of samples such that the
-    mean \emph{actually is} different from 0.5 (it depends on the distribution
-    of the variable that corresponds to $Y = \frac{1}{b}\sum_{i=1}^b x_i$).
-
-    \item (TODO later) Unfortunately, I am unable to understand why ``the size
-    of the range in $u$ is at least $\exp([-2,-1.125])$'', unless the real range
-    is $\exp([-2,0])$ (as I think it should be) and you were simply cropping off
-    the extra range.
-
-    \item (TODO later) My main confusion now is that I am unable to derive
-    Equation~\ref{eq:log_uo_range}, and we may want to clarify the notation used
-    when we put two intervals side by side.
-
-    \item (TODO later) I do not understand why after
-    Equation~\ref{eq:accept_test_rewritten}, we say that the standard deviation
-    of the LHS ``given the interval constraints is at least $0.5/\sqrt{b}$.'' We
-    already know that the $X_i$ are iid $\mathcal{N}(\theta = 0.5, 1)$ variables
-    with a fixed $\theta = 0.5$ value, so we already know the standard deviation
-    is supposed to be $\sqrt{({\rm Var}(x_1) + \cdots + {\rm Var}(x_b))/b^2} =
-    1/\sqrt{b}$.
-
-    \item I added Equation~\ref{eq:rhs_range}, which I think is how you
-    determined the range. Note that it doesn't seem to matter whether it is
-    $[-1,0.75]$ or $[-0.75,1]$ since the main thing that matters is that we have
-    maximum absolute value of one.
-
-    \item At the end, it said the joint distribution of
-    $(\theta,\theta'-\theta,u)$, but I changed it to $(\theta,\theta',u)$, based
-    on your change to the text of Lemma~\ref{lem:worst_case}. But here's a
-    question: how is the sample of $\theta'$ (or even $\theta'-\theta$, as it
-    was written earlier) independent of $\theta$?
-\end{enumerate}
-}
-
-\end{comment}
 
@@
 samples of $x_i$. This turns out to be
-\begin{equation}
-\begin{split}
-\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') & \quad \\
-\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2} & \not\approx \log u_0\\
-\iff \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) & \not\approx  0 \\
-\iff \frac{1}{b}\sum_{i=1}^b x_i-0.5 & \not\approx 0. \label{eq:accept_test_zero}
-\end{split}
-\end{equation}
+\begin{align}
+\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') 
+&\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2} \not\approx \log u_0\\
+&\iff \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \not\approx  0 \\
+&\iff \frac{1}{b}\sum_{i=1}^b x_i-0.5 \not\approx 0. \label{eq:accept_test_zero}
+\end{align}
 Since the $x_i$ have mean 0.5, the resulting test with our chosen $u_0$ will
@@
 
-The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
-\begin{align*}
--6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
-&\leq 1.36E|X|^3.
-\end{align*}
+The following bound is given immediately after Corollary 2 from~\cite{explicit-clt05}:
+\begin{equation}
+-6.4E|X|^3-2E|X| \leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \leq 1.36E|X|^3.
+\end{equation}
 This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
@@
 \begin{equation}
-\begin{split}
-    P''(y) & = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) \\
-    \quad {\rm and}\quad Q''(y) & = S\left({y-\Delta}\right)
-\end{split}
+    P''(y) = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y)
+    \quad {\rm and}\quad Q''(y) = S\left({y-\Delta}\right)
 \end{equation}
@@
 then Equation 2.2 in Bentkus et al.'s work on Edgeworth
-expansions~\citep{Bentkus97} provides:
+expansions~\cite{Bentkus97} provides:
 
@@
 table. We leave further analysis and experiments with $g(x,y)$ as future work.
-%{\color{blue} Daniel: I will look at the code to understand this paragraph
-%better because right now I am confused.}
-% JFC: the code doesnt implement this yet.
-
-% Daniel: don't forget this. The AISTATS website says to put it in the
-% supplementary material, so I put it at the very end on a new page (to make it
-% more noticeable). Obviously for the arXiv version, we don't want this there.
-% Also, note to John, the reviewers for AISTATS **will have access to the
-% previous NIPS reviews**, which is why I am making references to those.
-\clearpage
-\section{NIPS 2016 Submission Statement}
-
-We previously submitted an early, incomplete draft of this paper to NIPS 2016. The
-current manuscript has been substantially improved since that submission. We have
-made the following changes:
-
-\begin{enumerate}
-\item The (incomplete) analysis in the NIPS paper has been substantially replaced.
-  The asymptotic CLT arguments in the earlier paper have been replaced with
-  explicit quantitative bounds. There were several gaps in the earlier analysis:
-  the effects of addition of random variables on CDF errors and the construction and
-  error bounds of the correction distribution which are key parts of the present
-  paper have been added.
-
-\item Several recent relevant references have been added, and we have written
-  clearer statements about the contribution of the present work.
-  These include~\citep{TallData15} and the Barker test~\citep{Barker65}.
-
-\item Some reviewers mentioned that our deconvolution approach to
-  determine the correction distribution was not unique and thus
-  ill-defined.  We added detail on how the correction
-  distribution is computed (see Section~\ref{sec:correction}). While the distribution is not unique, we
-  regularize the inversion to obtain a concrete distribution with very
-  low error (close to single-point machine precision).
-
-\item The earlier draft had an inconsistency in its statement about constant
-  minibatch size. We clarify that this is true only in expected value.
-
-\item Our experiments now include new comparisons with a baseline
-  from~\citep{icml2014c1_bardenet14} (in addition to~\citep{cutting_mh_2014},
-  which we had earlier). Furthermore, each algorithm now runs MCMC sampling on
-  the \emph{same} distribution.  Previously, we ran our distribution at a
-  higher temperature but kept the algorithm from~\citep{cutting_mh_2014}
-  running on the distribution at temperature $T=1$. In this set of
-  experiments, we tuned hyperparameters of the other algorithms to yield
-  best results. In the Gaussian mixture model scenario
-  (Section~\ref{ssec:gaussians}), we provide likelihoods and test
-  statistics to measure the accuracy of the samples.
-
-\item We improved Algorithm~\ref{alg:our_algorithm} so that we show
-  explicitly when we compute the moments, and why we now only need one
-  distribution $C_\sigma$ for $\sigma=1$ due to the extra $X_{\rm nc}$
-  variable.
-
-\item Finally, we made minor revisions addressing: differences between the
-  Barker function vs. the original MH test, and sampling with vs. without
-  replacement.
-\end{enumerate}
-
-We estimate the degree of overlap between the papers in mathematics (equations) at less than
-10\%, in overall text at less than 30\%. The experiments are similar but were also redone at larger
-scale in this draft.
 

### 2017-03-29 uai2017/paper.tex

@@
 adjust the temperature to $K=10,000$.  Taking logs, we get the target as shown
-in the far left of Figure~\ref{fig:gauss_mix_1}.
+in the far right of Figure~\ref{fig:gauss_mix_1}.
 
@@
 magnitude better compared to the other two methods (averages are 12562 and
-67508). \textbf{Daniel: Xinlei, these numbers don't match
-Figure~\ref{fig:gauss_mix_2} and the figures need to be updated so the scatter
-plot has the correct method names and the scaling key, while the histograms have
-the correct names and the seaborn background.} We
-further ran 10 runs of mixture of Gaussians experiments, and the average and
-standard deviation values of the mean minibatch size for {\sc MHminibatch}, {\sc
-AustereMH(c)}, {\sc MHSubLhd} are listed in table ~\ref{tab:gaussianres}. Sizes
-correspond to the running times of the methods, excluding the likelihood
-computation of all data points for {\sc MHSubLhd}, which would drastically
-increase running time.
+67508). We further ran 10 runs of mixture of Gaussians experiments and report
+minibatch sizes in Table~\ref{tab:gaussianres}. Sizes correspond to the running
+times of the methods, excluding the likelihood computation of all data points
+for {\sc AustereMH(nc)} and {\sc MHSubLhd}, which would drastically increase
+    running time.
 
@@
 mixture model. The averages are taken over 10 independent trials (3000 samples
-each). \textbf{Daniel: Xinlei, can you add in one more decimal point for all the
-numbers in this table, like I do in the Logistic Regression table?}}
+each).}
 % Daniel: this `small` command was carried over from the LR table which couldn't
@@
 \begin{table}[t]
-\caption{Gaussian Mixture Model Statistics \textbf{TODO Xinlei, please fix this
-table.}}
+\caption{Gaussian Mixture Model Statistics}
 \small
@@
 \begin{tabular}{l l l l}
-\textbf{Metric} & {\sc MHminibatch} & {\sc AustereMH} & {\sc MHSubLhd} \\
+\textbf{Metric/Method} & {\sc MHminibatch} & {\sc AustereMH(c)} & {\sc MHSubLhd} \\
 \hline
@@
 
-\textbf{Daniel: Xinlei, you should be in charge of checking this section and
-adding some stuff to Table~\ref{tab:poissons}, perhaps averages plus or minus
-one standard deviation.}
-
 We discretize the posterior coordinates into bins with respect to the two
@@
 Table~\ref{tab:poissons}). The table provides the mean likelihood value and mean
-Chi-Squared test statistics value as well as their standard deviations.  Our likelihood values 
-lies between~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14}, 
-but the variance of their likelihood values are high and
-ordering changes depending on the range of samples generated. It should be noted 
-that we are not aiming at optimizing the likelihood values nor the Chi-Squared statistics.
-We use these values to show the extent of correctness. \textbf{Daniel:
-Xinlei, one reviewer had a problem with this, and I agree. The easiest fix here
-is to simply report standard deviation values. Also, we should emphasize that we
-are not aiming to get the best score by this metric but simply something that
-exhibits some form of ``correctness.''}
+Chi-Squared test statistics value as well as their standard deviations.  Our
+likelihood values lies between~\citep{cutting_mh_2014}
+and~\citep{icml2014c1_bardenet14}, but we note that we are not aiming to
+optimize the likelihood values or the Chi-Squared statistics.  We use these
+values to show the extent of correctness.
 

### 2017-03-29 uai2017/paper.tex
tweaks to intro and conclusions

@@
 
-Our test is applicable when the variance (over data samples) of the log
-acceptance probability is small enough (less than 1). It's not clear at first why
-this quantity should be bounded, but we will show that it is ``natural'' for
-well-specified models running Metropolis-Hastings sampling with optimal
-proposals~\citep{OptimalScaling01} on a full dataset. When we reduce the amount
-of data for the test, the variance goes up. We have to reduce variance in one
-of several ways. Either:
-
-\begin{itemize}[noitemsep]
+We note that this approach is fundamentally different from prior
+work. It makes no assumptions and requires no global bounds on the
+posterior parameter distribution in order to work. It is exact in the
+limit as batch size increases by the Central Limit Theorem. This is
+not true of ~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14}
+which use tail bounds and provide only approximate tests even with
+arbitratily large batches of data. Our test is also exact under the
+assumptions of \citep{cutting_mh_2014} that per-sample parameter
+posteriors are normally distributed about the population mean.  Rather
+than tail bounds, our approach uses moment estimates from the data to
+determine how far the minibatch posteriors deviate from a normal
+distribution.  These bounds carry through to the overall accuracy of
+the test.
+
+Our test is applicable when the variance (over data samples) of the
+log probability ratio between proposal and current state is small
+enough (less than 1). It's not clear at first why this quantity should
+be bounded, but it is natural for well-specified models running
+Metropolis-Hastings sampling with optimal
+proposals~\citep{OptimalScaling01} on a full dataset. If the posterior
+parameter distribution is a unit-variance normal distribution, then
+the posterior for $N$ samples will have variance $1/N$. There is
+simply not enough information in $M << N$ samples to locate and
+efficiently sample from this posterior. This is not a property of any
+particular proposal or test, but of the information carried by the
+data. The variance condition succinctly captures the condition that
+the minibatch carries enough information to generate a sample.  While
+we cannot expect to generate independent samples from the posterior
+using only a small subset of the data, there are three situations
+where we can exploit small minibatches:
+
+\begin{enumerate}[noitemsep]
     \item Increase the temperature $K$ of the target distribution. Log
     likelihoods scale as $1/K$, and so the variance of the likelihood ratio will
-    vary as $1/K^2$. Our model is no longer well-specified (we are doing
-    inference at a temperature different from that assumed during data
-    generation), but as we demonstrate in Section~\ref{ssec:logistic}, higher
+    vary as $1/K^2$. As we demonstrate in Section~\ref{ssec:logistic}, higher
     temperature can be advantageous for parameter exploration.
 
-    \item For continuous distributions, reduce the proposal step size and
-    variance compared to an optimal proposal. The variance of the log acceptance
-    probability scales as the square of proposal step size.
-
-    \item Utilize Hamiltonian and Langevin Dynamics to judiciously model a system
-    so as to generate distant yet high-quality proposals.
-\end{itemize}
-
+    \item For continuous distributions, reduce the proposal step size
+      (i.e. generate correlated samples). The variance of the log
+      acceptance probability scales as the square of proposal step
+      size.
+
+  \item Utilize Hamiltonian Dynamics for proposals and tests. Here the
+    dynamics itself provide shaping to the posterior distribution, and
+    the M-H is only needed to correct for quantization error. In terms
+    of the information carried by the samples, this approach is not
+    limited by the data in a particular minibatch since momentum is
+    carried over time and ``remembered'' across multiple
+    minibatches.
+\end{enumerate}
+
+We note that 2 above is characteristic of Gibbs samplers applied to
+large datasets~\citep{dupuy2016}. Such samplers represent a
+model posterior via counts over an entire dataset of $N$ samples. When
+a minibatch of $M$ samples is used to update the model, the counts for
+these samples only are updated. This creates ``steps'' of $O(M/N)$
+in the model parameters, and correlated samples from the model
+posterior. Correlated samples are still very useful in high-dimensional
+ML problems with multi-model posteriors since they correspond to a finer-scale
+random walk through the posterior landscape. 
+\begin{comment}
 It is worth discussing at this point the typical goals of M-H sampling on large
@@
 variance.
-
+\end{comment}
 The contributions of this paper are as follows:
@@
 
-    \item We compare performance of our new test and prior approaches on several
-    datasets. We demonstrate several order-of-magnitude improvements in efficiency
-    (measured as data consumed per test), and show that it does not suffer from
-    long-tailed minibatch sizes.
+    \item We compare our new test and prior approaches on several
+      datasets. We demonstrate several order-of-magnitude improvements in sample efficiency,
+      and that batch size distribution is short-tailed. 
 \end{itemize}
@@
 
-The most pressing direction for future work is to integrate our M-H test with
-Hamiltonian Monte Carlo and Langevin Dynamics methods to enable MCMC methods to
-scale to models such as deep neural networks. With these types of models, our
-proposal probabilities will almost certainly have to be asymmetric, but these
-typically result in an estimate plus Gaussian noise which we can absorb into our
-test.
+A priority is to extend our work to methods such as Hamiltonian Monte Carlo and Langevin Dynamics
+which use efficient but asymmetric proposals. While there are various approaches
+to symmetrizing these proposals, they have high cost in the context of minibatch
+MCMC. Instead we plan to extend our method to log proposal ratios which have
+similar structure (whole-dataset mean plus additive noise) to the log probability
+ratio. These can be similarly absorbed in the Barker test.
 

### 2017-03-30 uai2017/paper.tex
Final touch-ups. Will send to arxiv, UAI

@@
 Our test is applicable when the variance (over data samples) of the
-log probability ratio between proposal and current state is small
+log probability ratio between the proposal and the current state is small
 enough (less than 1). It's not clear at first why this quantity should
@@
 the posterior for $N$ samples will have variance $1/N$. There is
-simply not enough information in $M << N$ samples to locate and
+simply not enough information in $M \ll N$ samples to locate and
 efficiently sample from this posterior. This is not a property of any
@@
 
-We note that 2 above is characteristic of Gibbs samplers applied to
+We note that case two above is characteristic of Gibbs samplers applied to
 large datasets~\citep{dupuy2016}. Such samplers represent a
@@
 posterior. Correlated samples are still very useful in high-dimensional
-ML problems with multi-model posteriors since they correspond to a finer-scale
-random walk through the posterior landscape. 
-\begin{comment}
-It is worth discussing at this point the typical goals of M-H sampling on large
-datasets. By the Bernstein-von Mises Theorem, the posterior distribution for a
-Bayesian inference task has variance that scales inversely with $N$. Simply
-sampling from it is one application, but an efficient
-proposal~\citep{OptimalScaling01} has similar variance to the target and will
-diffuse to it extremely slowly. For applications to neural networks or models
-where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
-likely get trapped in one of the modes. A common solution is to anneal the
-sampler, running first at high temperatures to flatten the likelihood landscape.
-This in turn reduces the variance of the log acceptance probability and allows
-our test to be applied. These samples can cover the search space densely with
-small steps rather than taking a few sparse steps towards an optimum. In this
-mode, M-H tests can be used in similar fashion to Stochastic Gradient Descent.
-The goal in SGD is to make gradual progress to a posterior mode with each step,
-taking small steps so that the cumulative displacement has progressively lower
-variance.
-\end{comment}
-The contributions of this paper are as follows:
+ML problems with multi-modal posteriors since they correspond to a finer-scale
+random walk through the posterior landscape.  The contributions of this paper
+are as follows:
 
@@
       datasets. We demonstrate several order-of-magnitude improvements in sample efficiency,
-      and that batch size distribution is short-tailed. 
+      and that the batch size distribution is short-tailed. 
 \end{itemize}
@@
 all methods tested. For values from~\citep{cutting_mh_2014}
-or~\cite{icml2014c1_bardenet14}, we use their notation to be consistent.
+or~\cite{icml2014c1_bardenet14}, we use their notation ($\Delta^*, m, \epsilon,
+\gamma, p$, and $\delta$) to be consistent.
 

### 2017-03-30 uai2017/paper.tex
Made author names anonymous, added student-t distribution.

@@
 % following two lines:
-%\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
-%          % UAI  reviewing is double-blind.
+\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
+          % UAI  reviewing is double-blind.
 
 % The author names and affiliations should appear only in the accepted paper.
-%% \author{ {\bf Daniel Seita\thanks{Footnote for author to give an
-%% alternate address.}} \\
-%% Computer Science Dept. \\
-%% Cranberry University\\
-%% Pittsburgh, PA 15213 \\
-%% \And
-%% {\bf Xinlei Pan}  \\
-%% Affiliation          \\
-%% Address \\
-%% \And
-%% {\bf Haoyu Chen}   \\
-%% Affiliation \\
-%% Address    \\
-%% \And
-%% {\bf John Canny}   \\
-%% Affiliation \\
-%% Address    \\
+%% \author{
+%% Daniel Seita$^1$, 
+%% Xinlei Pan$^1$,
+%% Haoyu Chen$^1$,
+%% John Canny$^{1,2}$ \\
+%% $^1$ University of California, Berkeley, CA \\
+%% $^2$ Google Research, Mountain View, CA\\ 
+%% \texttt{\{seita,xinleipan,haoyuchen,canny\}@berkeley.edu}
 %% }
-\author{
-Daniel Seita$^1$, 
-Xinlei Pan$^1$,
-Haoyu Chen$^1$,
-John Canny$^{1,2}$ \\
-$^1$ University of California, Berkeley, CA \\
-$^2$ Google Research, Mountain View, CA\\ 
-\texttt{\{seita,xinleipan,haoyuchen,canny\}@berkeley.edu}
-}
 
@@
 \end{equation}
-This means $t=\bar{X}/s_X$ has an approximate Student's distribution
+This means $t=\bar{X}/s_X$ has an approximate Student's $t$ distribution
 which approaches a Gaussian. Then

### 2017-03-30 uai2017/paper.tex
fixed wording of lemma 3
fixed a typo

@@
 \end{equation}
-This means $t=\bar{X}/s_X$ has an approximate Student's $t$ distribution
-which approaches a Gaussian. Then
+Then the t-statistic $t=\bar{X}/s_X$ has a distribution which is approximately normal, with
+error bounded by:
 \begin{equation}\label{eq:clt-bounds}
@@
 showing that the distribution of $\bar{X}$ approaches the normal distribution
-$\mathcal{N}(0,s_X)$ whose variance is $s_X$,
-measured from the sample.
+$\mathcal{N}(0,s_X)$ whose standard deviation is $s_X$, as measured
+from the sample.
 

### 2017-03-30 uai2017/paper.tex
fix organization

@@
 We note that this approach is fundamentally different from prior
-work. It makes no assumptions and requires no global bounds on the
-posterior parameter distribution in order to work. It is exact in the
+work. It makes no assumptions about the form of, and requires no global bounds on the
+posterior parameter distribution. It is exact in the
 limit as batch size increases by the Central Limit Theorem. This is
@@
 arbitratily large batches of data. Our test is also exact under the
-assumptions of \citep{cutting_mh_2014} that per-sample parameter
-posteriors are normally distributed about the population mean.  Rather
+assumptions of \citep{cutting_mh_2014} that the log probability ratios
+of batches are normally distributed about their mean.  Rather
 than tail bounds, our approach uses moment estimates from the data to
@@
     dynamics itself provide shaping to the posterior distribution, and
-    the M-H is only needed to correct for quantization error. In terms
+    the M-H test is only needed to correct quantization error. In terms
     of the information carried by the samples, this approach is not
@@
 
-To address this challenge,~\citep{cutting_mh_2014,icml2014c1_bardenet14} perform
-approximate Metropolis-Hasting tests using sequential hypothesis testing. Each
-iteration, they start with a small minibatch and test whether $\theta'$ should
-be accepted based on approximating $u<\alpha(\theta,\theta')$. If the
-approximate test cannot decide with sufficient confidence, the minibatch grows
-and the test repeats. This process continues until a decision. The bounds depend
-on either an asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
-concentration bound~\citep{icml2014c1_bardenet14}. We refer to these algorithms,
-respectively, as {\sc AustereMH} and {\sc MHSubLhd}. While both show useful
-reductions in the number of samples required, they suffer from two drawbacks:
-having to compute all log likelihood tokens each iteration to compute certain
-statistics,\footnote{Thus, these are not technically minibatch methods.} (see
-Section~\ref{sec:analysis}), and resolving small log likelihood ratio
-differences between the minibatch and full batch versions. We discuss a
+To address this
+challenge,~\citep{cutting_mh_2014,icml2014c1_bardenet14} perform
+approximate Metropolis-Hasting tests using sequential hypothesis
+testing. At each iteration, a subset of data is sampled and used to
+test whether to accept $\theta'$ using an approximation to
+$\alpha(\theta,\theta')$. If the approximate test does not yield a
+decision, the minibatch size is increased and the test repeated. This
+process continues until a decision. These methods either invoke the
+asymptotic CLT and assume that finite  batch errors are normally
+distributed~\citep{cutting_mh_2014} or use a concentration
+bound~\citep{icml2014c1_bardenet14}. We refer to these algorithms,
+respectively, as {\sc AustereMH} and {\sc MHSubLhd}. While both show
+useful reductions in the number of samples required, they suffer from
+two drawbacks: (i) They are approximate, and always yield a decision
+with a finite error, (ii) They both require exact, dataset-wide bounds
+that depend on $\theta$.
+\footnote{We obtained the authors code for both and found that they scanned the entire
+dataset at each step to obtain these estimates.} (see
+Section~\ref{sec:analysis}). We discuss a
 worst-case scenario in Section~\ref{ssec:gaussian_example}.
@@
 \begin{algorithm}[tb]
-\caption{Our acceptance test, {\sc MHminibatch}.}
+\caption{{\sc MHminibatch} acceptance test.}
 \label{alg:our_algorithm}
@@
 We now derive error bounds for our M-H test and the target distribution it
-generates. In the most similar prior works,~\citep{cutting_mh_2014}, CLT
-asymptotic arguments are used to show that their approximate acceptance test
-error (with respect to full-batch M-H) tends to zero as batch size increases,
-but no quantitative bounds are given. In addition, to design a test that
-minimizes data usage given an error bound, they rely on computing a standardized
-mean $\mu_{\rm std}$ each iteration. As the value of $\mu_{\rm std}$ depends on
-all $\log p(x_i|\theta)$ and $\log p(x_i| \theta')$ terms, the test is
-technically not a minibatch test; \citep{cutting_mh_2014} propose a conservative
-variant which assumes $\mu_{\rm std}=0$ and avoids the need to evaluate all
-likelihood terms. We refer to this as {\sc AustereMH(c)} and the default,
-non-conservative variant as {\sc AustereMH(nc)}.
-
-For {\sc MHSubLhd}, the test requires direct bounds on the log likelihood ratio,
-which for general distributions requires knowing $p(x_i | \theta)$ and $p(x_i |
-\theta')$ for all $N$ samples. In~\citep{icml2014c1_bardenet14}, explicit bounds
-are given and depend on
-\begin{equation}\label{eq:bad_bound}
-    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
-\end{equation}
-It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
-regression and other simple models, but it is difficult to do so for more
-complicated models, so one generally needs to use all $p(x_i | \theta')$
-terms\footnote{The sample code provided by~\citep{icml2014c1_bardenet14}
-computes $C_{\theta,\theta'}$ by traversing the entire data, thus providing no
-performance advantage over the complete test.} to get $C_{\theta,\theta'}$, like
-the {\sc AustereMH(nc)} test.  In contrast, we use quantitative forms of the CLT
-which rely on measurable statistics from a \emph{single} minibatch.
-
+generates. 
 In Section~\ref{ssec:delta_star_distribution}, we present bounds on the absolute
@@
 \section{EXPERIMENTS}\label{sec:experiments}
+In one of the most closely-related prior
+works,~\citep{cutting_mh_2014}, asymptotic CLT arguments are used to
+argue that a modified standard M-H test can be used on subsets of data
+with modest error. This assumes knowledge of dataset-wide mean
+$\mu_{\rm std}$ each iteration (it depends on $\theta$). Determining
+$\mu_{\rm std}$ exactly requires a scan over the entire dataset (so
+the method in practice provides no data reduction), or some
+model-specific bounds must be derived.  \citep{cutting_mh_2014} also
+propose a conservative variant which assumes $\mu_{\rm std}=0$ and
+avoids the scan.  We refer to the conservative version as {\sc
+  AustereMH(c)} and the non-conservative variant as {\sc
+  AustereMH(nc)}. We analyze both in the sequel.
+
+In another related work~\citep{icml2014c1_bardenet14} use concentration bounds
+and a similar modification to the standard M-H test ({\sc MHSubLhd} method). 
+For {\sc MHSubLhd}, the test requires direct bounds on the log likelihood ratio,
+which for general distributions requires knowing $p(x_i | \theta)$ and $p(x_i |
+\theta')$ for all $N$ samples. In~\citep{icml2014c1_bardenet14}, explicit bounds
+are given and depend on
+\begin{equation}\label{eq:bad_bound}
+    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
+\end{equation}
+It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
+regression and other simple models, but it is difficult to do so for more
+complicated models, so one generally needs to use all $p(x_i | \theta')$
+terms\footnote{The sample code provided by~\citep{icml2014c1_bardenet14}
+computes $C_{\theta,\theta'}$ by traversing the entire data, thus providing no
+performance advantage over the complete test.} to get $C_{\theta,\theta'}$, like
+the {\sc AustereMH(nc)} test.  In contrast, we use quantitative forms of the CLT
+which rely on measurable statistics from a \emph{single} minibatch.
 

### 2017-03-30 uai2017/paper.tex
fix organization

@@
 \section{EXPERIMENTS}\label{sec:experiments}
-In one of the most closely-related prior
-works,~\citep{cutting_mh_2014}, asymptotic CLT arguments are used to
-argue that a modified standard M-H test can be used on subsets of data
-with modest error. This assumes knowledge of dataset-wide mean
-$\mu_{\rm std}$ each iteration (it depends on $\theta$). Determining
-$\mu_{\rm std}$ exactly requires a scan over the entire dataset (so
-the method in practice provides no data reduction), or some
-model-specific bounds must be derived.  \citep{cutting_mh_2014} also
-propose a conservative variant which assumes $\mu_{\rm std}=0$ and
-avoids the scan.  We refer to the conservative version as {\sc
-  AustereMH(c)} and the non-conservative variant as {\sc
-  AustereMH(nc)}. We analyze both in the sequel.
-
-In another related work~\citep{icml2014c1_bardenet14} use concentration bounds
-and a similar modification to the standard M-H test ({\sc MHSubLhd} method). 
-For {\sc MHSubLhd}, the test requires direct bounds on the log likelihood ratio,
-which for general distributions requires knowing $p(x_i | \theta)$ and $p(x_i |
-\theta')$ for all $N$ samples. In~\citep{icml2014c1_bardenet14}, explicit bounds
-are given and depend on
-\begin{equation}\label{eq:bad_bound}
-    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
-\end{equation}
-It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
-regression and other simple models, but it is difficult to do so for more
-complicated models, so one generally needs to use all $p(x_i | \theta')$
-terms\footnote{The sample code provided by~\citep{icml2014c1_bardenet14}
-computes $C_{\theta,\theta'}$ by traversing the entire data, thus providing no
-performance advantage over the complete test.} to get $C_{\theta,\theta'}$, like
-the {\sc AustereMH(nc)} test.  In contrast, we use quantitative forms of the CLT
+Here we compare with the most similar prior works
+~\citep{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}.
+In~\citep{cutting_mh_2014}, an asymptotic CLT is used to argue that a
+modified standard M-H test can be used on subsets of the data. This
+assumes knowledge of dataset-wide mean $\mu_{\rm std}$ each iteration
+(it depends on $\theta$). Determining $\mu_{\rm std}$ exactly requires
+a scan over the entire dataset, or some model-specific bounds.
+\citep{cutting_mh_2014} also propose a conservative variant which
+assumes $\mu_{\rm std}=0$ and avoids the scan.  We refer to the
+conservative version as {\sc AustereMH(c)} and the non-conservative
+variant as {\sc AustereMH(nc)}. We analyze both in this section.
+
+In~\citep{icml2014c1_bardenet14}
+concentration bounds are used with a similar modification to the standard M-H
+test ({\sc MHSubLhd} method).  For {\sc MHSubLhd}, the required global
+bound is denoted $C_{\theta, \theta'}$ which once again depends on
+$\theta$ and so must be recomputed at each step, or estimated in a
+model-specific way.   We obtained
+sample code for both methods from the authors, and found that both
+{\sc AustereMN(nc)} and {\sc MHSubLhd} scanned the entire dataset at
+each iteration to derive these bounds. We do not include the cost of
+doing this in our experiments, since otherwise there would be no
+improvement over testing the full dataset. However, it should be kept
+in mind that such bounds must be provided to these methods.
+Our test by contrast uses a quantitative form of the CLT
 which rely on measurable statistics from a \emph{single} minibatch.
+It therefore requires no dataset-wide scans, and can be used, e.g. on
+streams of data.
 
@@
 In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark {\sc
-MHminibatch} against the tests from~\citep{icml2014c1_bardenet14}
-and~\citep{cutting_mh_2014}, with the latter using a grid-search over minibatch
+MHminibatch} against {\sc MHSubLhd}, {\sc AustereMH(c)} and {\sc AustereMH(nc)}.
+Hyperparameters for the latter were optimized using a grid-search over minibatch
 sizes $m$ and per-test thresholds $\epsilon$ described in

### 2017-03-30 uai2017/paper.tex
fix organization
update with 10 runs of gaussian mixture model experiment results

@@
 tests. We present theoretical results and experimentally show the benefits of
-our test on Gaussian mixtures and a variety of logistic regression experiments.
+our test on Gaussian mixtures and a logistic regression experiment.
 
@@
 \hline
-\T Equation~\ref{eq:log_prob_poissons} & $-1306.2 \pm 20.7$ & $-1619.6 \pm 25.3$ & $-1301.5 \pm 44.4$ \\
-Chi-Squared & $5706.8 \pm 2271.5$ & $ 10501.3 \pm 2076.5$ & $ 3711.7 \pm 1182.4$ \\
+\T Equation~\ref{eq:log_prob_poissons} & $-1307.0 \pm 229.5$ & $-1386.9 \pm 322.4$ & $-1295.1 \pm 278.0$ \\
+Chi-Squared & $4502.3 \pm 1821.8$ & $ 5216.9 \pm 3315.8$ & $ 3462.3 \pm 1519.5$ \\
 \end{tabular}

### 2017-03-30 uai2017/paper.tex
Very minor changes, in appendix mostly

@@
 with a finite error, (ii) They both require exact, dataset-wide bounds
-that depend on $\theta$.
-\footnote{We obtained the authors code for both and found that they scanned the entire
-dataset at each step to obtain these estimates.} (see
-Section~\ref{sec:analysis}). We discuss a
+that depend on $\theta$ (see Section~\ref{sec:analysis}).\footnote{We obtained
+the authors code for both and found that they scanned the entire dataset at each
+step to obtain these estimates.} We discuss a
 worst-case scenario in Section~\ref{ssec:gaussian_example}.
@@
 values for the entirety of the trials. We experimented with the following
-values:
+values, which are similar to the values reported in~\citep{cutting_mh_2014}:
 \begin{itemize}[noitemsep]
@@
 Figure~\ref{fig:logistic_histograms_mnist} shows the histograms for the four
-methods on one representative trial of MNIST-13k, which shows similar relative
-performance of the four methods as shown in
+methods on one representative trial of MNIST-13k, indicating similar relative
+performance of the four methods as in
 Figure~\ref{fig:logistic_histograms_mnist8m} (which uses MNIST-100k). In
@@
 all methods tested. For values from~\citep{cutting_mh_2014}
-or~\cite{icml2014c1_bardenet14}, we use their notation ($\Delta^*, m, \epsilon,
+or~\citep{icml2014c1_bardenet14}, we use their notation ($\Delta^*, m, \epsilon,
 \gamma, p$, and $\delta$) to be consistent.

### 2017-03-30 uai2017/paper.tex
Fixed table caption to say 10 trials. NOW let's re-submit to UAI.
Adding UAI arxiv version and UAI paper submission (without names)

@@
 % following two lines:
-\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
-          % UAI  reviewing is double-blind.
+%\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
+%          % UAI  reviewing is double-blind.
 
 % The author names and affiliations should appear only in the accepted paper.
-%% \author{
-%% Daniel Seita$^1$, 
-%% Xinlei Pan$^1$,
-%% Haoyu Chen$^1$,
-%% John Canny$^{1,2}$ \\
-%% $^1$ University of California, Berkeley, CA \\
-%% $^2$ Google Research, Mountain View, CA\\ 
-%% \texttt{\{seita,xinleipan,haoyuchen,canny\}@berkeley.edu}
-%% }
+\author{
+Daniel Seita$^1$, 
+Xinlei Pan$^1$,
+Haoyu Chen$^1$,
+John Canny$^{1,2}$ \\
+$^1$ University of California, Berkeley, CA \\
+$^2$ Google Research, Mountain View, CA\\ 
+\texttt{\{seita,xinleipan,haoyuchen,canny\}@berkeley.edu}
+}
 
@@
 \begin{table}[t]
-\caption{Gaussian Mixture Model Statistics}
+\caption{Gaussian Mixture Model statistics ($\pm$ one standard deviation over 10
+trials).}
 \small

### 2017-07-02 uai2017/paper.tex
Updated UAI paper with reviewer comments, +copyright form

@@
 limit as batch size increases by the Central Limit Theorem. This is
-not true of ~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14}
+not true of~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14}
 which use tail bounds and provide only approximate tests even with
-arbitratily large batches of data. Our test is also exact under the
-assumptions of \citep{cutting_mh_2014} that the log probability ratios
+arbitrarily large batches of data. Our test is also exact under the
+assumptions of~\citet{cutting_mh_2014} that the log probability ratios
 of batches are normally distributed about their mean.  Rather
@@
 will see that its overall sample efficiency is much higher than the
-earlier methods.
+earlier methods. See Appendix~\ref{app:why_barker} for details.
 
@@
 
-This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
+This appendix is divided into three major sections. Appendix~\ref{app:proofs} provides
 the proofs that we omitted from the main text due to space constraints.
-Appendix~\ref{app:experiments} provides further details on the correction
-distribution derivation and on our three main experiments to assist
+Appendix~\ref{app:why_barker} elaborates on our choice of the Barker logistic
+function. Finally, Appendix~\ref{app:experiments} presents further details on the correction
+distribution numerical derivation and on our three main experiments to assist
 understanding and reproducibility.
+
 % Daniel: TODO can decide whether to include this here or not. It might actually
@@
 
+\section{WHY THE BARKER LOGISTIC FUNCTION?}\label{app:why_barker}
+% Daniel: adding this one due to reviewer concerns
+
+Regarding our choice of the Logistic function, a test function $f(x)$ for
+Metropolis-Hastings must satisfy Lemma~\ref{lem:detailed_balance}. In addition,
+it must be monotone, bounded by $[0,1]$ and be such that $\lim_{x \to
+-\infty}f(x) = 0$ and $\lim_{x \to \infty}f(x) = 1$. While many functions
+satisfy this, including the classical test $f(x) = \min\{\exp(x), 1\}$, the
+Logistic function is the \emph{unique} function in this class which is
+anti-symmetric about 0.5, so it represents the (unique) CDF of a symmetric
+random variable. Our method requires approximating this with the sum of a
+Gaussian random variable (which is symmetric) and a correction. The Logistic CDF
+$L$ and Gaussian CDF $\Phi$ are extremely close even without correction; more
+precisely, the CDF error from the closest Gaussian CDF numerically satisfies
+$\sup_x |L(x) - \Phi(x/1.7)| < 0.01$, i.e. the sup error between Logistic and
+Gaussian CDFs is less than 1\%. With our correction we can make this error
+orders of magnitude smaller.
+
+While not a proof of optimality, it is unlikely that a non-symmetric test
+function $f(x)$ --- representing a skewed variable --- would do better. It would
+require a highly-skewed correction variable, and likely require a much narrower
+normal distribution (and hence more samples).
+
 \section{ADDITIONAL EXPERIMENT DETAILS}\label{app:experiments}

### 2017-07-09 uai2017/paper.tex
Final version for UAI 2017 and arXiv

@@
 will see that its overall sample efficiency is much higher than the
-earlier methods. See Appendix~\ref{app:why_barker} for details.
+earlier methods. See Appendix~\ref{app:why_barker} for additional discussion.
 
@@
 $L$ and Gaussian CDF $\Phi$ are extremely close even without correction; more
-precisely, the CDF error from the closest Gaussian CDF numerically satisfies
-$\sup_x |L(x) - \Phi(x/1.7)| < 0.01$, i.e. the sup error between Logistic and
-Gaussian CDFs is less than 1\%. With our correction we can make this error
+precisely, the CDF error from the closest Gaussian CDF --- which we numerically
+determined to have standard deviation approximately 1.7 --- satisfies $\sup_x
+|L(x) - \Phi(x/1.7)| < 0.01$. Said another way, the error between the Logistic
+and Gaussian CDFs is less than 1\%. With our correction we can make this error
 orders of magnitude smaller.
